### 2022-08-17 code/L0/L0_1_Ecuador_plant_species_lists.Rmd
set up file and folder structure

---
title: "Ecuador plant species list"
authors: "Hazel Anderson"
collaborators: "None"
data input: "None"
data output: "Ecuador_plant_species_list.csv"
project: "Frugivoria"
date: "2022-08-17"
output: html_document
---
# Set Working directory & file path
```{r}
setwd()
data_path <- 'G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0'
```

# Load required packages
```{r}
library(BIEN)
library(kewr)
library(gbifdb)
```

# Pull data from various sources

## BIEN
```{r}
Ecuador <- BIEN_occurrence_country("Ecuador")
```

## Kew
```{r}

```

## GBIF
```{r}

```

### 2022-08-23 code/L0/L0_1_Ecuador_plant_species_lists.Rmd
add code for POWO and GBIF files

@@
 ```{r}
-setwd()
+#setwd() *WHAT SHOULD THIS BE???
 data_path <- 'G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0'
@@
 library(BIEN)
-library(kewr)
-library(gbifdb)
 ```
 
-# Pull data from various sources
+# Pull Ecuador plant species lists from various sources
 
@@
 ```{r}
-Ecuador <- BIEN_occurrence_country("Ecuador")
+Ecuador_BIEN <- BIEN_occurrence_country("Ecuador")
+write.csv(Ecuador_BIEN, "Ecuador_BIEN_species_list.csv")
 ```
 
-## Kew
+## Kew Plants of the World Online
 ```{r}
-
+library(readxl)
+Ecuador_POWO <- read_excel("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/PlantsOfTheWorldOnline_Ecuador_species.xlsx")
 ```
@@
 ```{r}
-
+Ecuador_GBIF <- read.csv("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/GBIF_Ecuador_PlantSpeciesList.csv")
 ```

### 2022-08-31 code/L0/L0_1_Ecuador_plant_species_lists.Rmd
clean code
Combine data into one csv and remove duplicate species Need to fix working directory issue; not saving files in correct Google Drive folder

@@
 collaborators: "None"
-data input: "None"
-data output: "Ecuador_plant_species_list.csv"
+data input: "PlantsOfTheWorldOnline_Ecuador_species.xlsx and GBIF_Ecuador_PlantSpeciesList.csv"
+data output: "Ecuador_unique_species_list.csv"
 project: "Frugivoria"
@@
 ---
-# Set Working directory & file path
+# Set Working directory
 ```{r}
-#setwd() *WHAT SHOULD THIS BE???
-data_path <- 'G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0'
+setwd('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
 ```
@@
 library(BIEN)
+library(tidyr)
 ```
@@
 Ecuador_POWO <- read_excel("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/PlantsOfTheWorldOnline_Ecuador_species.xlsx")
+#won't read in csv version for some reason
 ```
@@
 ```
+
+# Combine data
+```{r}
+species_GBIF <- as.list(Ecuador_GBIF$species)
+species_BIEN <- as.list(Ecuador_BIEN$scrubbed_species_binomial)
+species_POWO <- as.list(Ecuador_POWO$genus)
+genus_POWO <- as.list(Ecuador_POWO$species)
+
+#need to combine species_POWO and genus_POWO
+species_POWO <- as.list(paste(genus_POWO, species_POWO, sep=" "))
+
+#combine all 3 species list together
+all_species <- c(species_BIEN, species_GBIF, species_POWO)
+
+#remove duplicates
+unique_species <- unique(all_species)
+
+```
+
+# Output data to csv
+```{r}
+write.csv(unique_species, "Ecuador_unique_species_list.csv")
+```

### 2022-08-31 code/L0/L0_2_Ecuador_plant_species_IUCN.Rmd
inital commit Add libraries and outline of code

---
title: "Ecudor plant species IUCN statuses"
author: "Hazel Anderson"
collaborators: "None"
data input: "Ecuador_unique_species_list.csv"
data output: "TBD"
project: "Frugivoria"
date: "2022-08-31"
output: html_document
---
# Set Working directory
```{r}
setwd('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(jsonlite)
library(rredlist)
```

```{r}
## Token for accessing IUCN data through their API
## User must obtain a token (https://apiv3.iucnredlist.org/api/v3/token) before using the 'rredlist package'.
token <- 'f632af67ee5eef8446a8cb0ca394e1f1cb369ca5eab626edc4ea46e5698deb25'
api_url <- 'https://apiv3.iucnredlist.org/api/v3/'

#Ecuador
ecuador_species <- rl_sp_country("EC", key = token, parse = TRUE)
list <- ecuador_species$result[1]

# Get codes for the different species groups
grp_codes <- GET(url = paste0(api_url, 'comp-group/list?token=', token))
code <-fromJSON(content(grp_codes, as = 'text'), simplifyDataFrame = TRUE, flatten = TRUE)$result
```

### 2022-09-12 code/L0/L0_1_Ecuador_plant_species_lists.Rmd
change set working directory to set file paths revamp how packages are installed/loaded

@@
 ---
-# Set Working directory
+# Set file paths
 ```{r}
-setwd('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
 ```
@@
 ```{r}
-library(BIEN)
-library(tidyr)
+for (package in c("BIEN", "tidyr")) {
+  if (!require(package, character.only=T, quietly=T)) {
+    install.packages(package)
+    library(package, character.only=T)
+  }
+}
 ```

### 2022-09-12 code/L0/L0_2_Ecuador_plant_species_IUCN.Rmd
change set working directory to set file paths revamp how packages are installed/loaded

@@
 ---
-# Set Working directory
+# Set file paths
 ```{r}
-setwd('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
 ```
 
-# Load required packages
+# Install and load required packages
 ```{r}
-library(jsonlite)
-library(rredlist)
+for (package in c("httr", "jsonlite", "rredlist")) {
+  if (!require(package, character.only=T, quietly=T)) {
+    install.packages(package)
+    library(package, character.only=T)
+  }
+}
 ```
@@
 code <-fromJSON(content(grp_codes, as = 'text'), simplifyDataFrame = TRUE, flatten = TRUE)$result
+
+# Get species list of all plants across all countries. 
+all_plants <-rl_comp_groups(group = c('mangrove_plants', 'seagrasses', 'cycads', 'magnolias', 'conifers', 'cacti'), key = token)
+
 ```

### 2022-09-12 code/L0/L0_1_Ecuador_plant_species_lists.Rmd
tidy up write csv to use output_path

@@
 Ecuador_BIEN <- BIEN_occurrence_country("Ecuador")
-write.csv(Ecuador_BIEN, "Ecuador_BIEN_species_list.csv")
+Ecuador_BIEN_species_list <- file.path(output_path,"Ecuador_BIEN_species_list.csv") 
+write.csv(Ecuador_BIEN, file = Ecuador_BIEN_species_list)
 ```
@@
 ```{r}
-write.csv(unique_species, "Ecuador_unique_species_list.csv")
+Ecuador_unique_species_list <- file.path(output_path,"Ecuador_unique_species_list.csv") 
+write.csv(unique_species, file = Ecuador_unique_species_list)
 ```

### 2022-09-20 code/L0/L0_2_Ecuador_plant_species_IUCN.Rmd
new, working code for species lists for plant groups Issues with merging ecuador plants; need to combine with all data not just taxon id

--- 2022-08-31 code/L0/L0_2_Ecuador_plant_species_IUCN.Rmd
@@
 ---
-# Set Working directory
+# Set file paths
 ```{r}
-setwd('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
 ```
 
-# Load required packages
+# Install and load required packages
 ```{r}
-library(jsonlite)
-library(rredlist)
+for (package in c("httr", "jsonlite", "rredlist", "dplyr")) {
+  if (!require(package, character.only=T, quietly=T)) {
+    install.packages(package)
+    library(package, character.only=T)
+  }
+}
 ```
@@
 ecuador_species <- rl_sp_country("EC", key = token, parse = TRUE)
-list <- ecuador_species$result[1]
+ecuador_species <- ecuador_species$result[1]
 
@@
 ```
+
+```{r}
+# Get species list of all plants across all countries
+mangrove_plants <- rl_comp_groups(group = 'mangrove_plants', key = token)
+p1 <- mangrove_plants$result[1]
+seagrasses<- rl_comp_groups(group = 'seagrasses', key = token)
+p2 <- seagrasses$result[1]
+cycads <- rl_comp_groups(group = 'cycads', key = token)
+p3 <- cycads$result[1]
+magnolias <- rl_comp_groups(group = 'magnolias', key = token)
+p4 <- magnolias$result[1]
+confiers <- rl_comp_groups(group = 'conifers', key = token)
+p5 <- confiers$result[1]
+cacti <- rl_comp_groups(group = 'cacti', key = token)
+p6 <- cacti$result[1]
+
+all_plants <- rbind(p1, p2, p3, p4, p5, p6)
+```
+
+```{r}
+ecuador_plants <- merge(all_plants, ecuador_species, by="taxonid")
+ecuador_plants_unique <- distinct(ecuador_plants, scientific_name, .keep_all = TRUE)
+```

### 2022-09-21 code/L0/L0_2_Ecuador_plants_IUCN.Rmd
Rename L0_2_Ecuador_plant_species_IUCN.Rmd to L0_2_Ecuador_plants_IUCN.Rmd

---
title: "Ecudor plant species IUCN statuses"
author: "Hazel Anderson"
collaborators: "None"
data input: "Ecuador_unique_species_list.csv"
data output: "TBD"
project: "Frugivoria"
date: "2022-08-31"
output: html_document
---
# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Install and load required packages
```{r}
for (package in c("httr", "jsonlite", "rredlist", "dplyr")) {
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package, character.only=T)
  }
}
```
# Set up token and api url
```{r}
## Token for accessing IUCN data through their API
## User must obtain a token (https://apiv3.iucnredlist.org/api/v3/token) before using the 'rredlist package'.
token <- 'f632af67ee5eef8446a8cb0ca394e1f1cb369ca5eab626edc4ea46e5698deb25'
api_url <- 'https://apiv3.iucnredlist.org/api/v3/'

```

# Ecuador species list
```{r}
#Ecuador
ecuador_species <- rl_sp_country("EC", key = token, parse = TRUE)
ecuador_species <- ecuador_species$result[1]
```

# Plant species list
```{r}
# Get codes for the different species groups
grp_codes <- GET(url = paste0(api_url, 'comp-group/list?token=', token))
code <-fromJSON(content(grp_codes, as = 'text'), simplifyDataFrame = TRUE, flatten = TRUE)$result
```

```{r}
# Get species list of all plants across all countries
mangrove_plants <- rl_comp_groups(group = 'mangrove_plants', key = token)
p1 <- as.data.frame(mangrove_plants$result)
seagrasses<- rl_comp_groups(group = 'seagrasses', key = token)
p2 <- as.data.frame(seagrasses$result)
cycads <- rl_comp_groups(group = 'cycads', key = token)
p3 <- as.data.frame(cycads$result)
magnolias <- rl_comp_groups(group = 'magnolias', key = token)
p4 <- as.data.frame(magnolias$result)
confiers <- rl_comp_groups(group = 'conifers', key = token)
p5 <- as.data.frame(confiers$result)
cacti <- rl_comp_groups(group = 'cacti', key = token)
p6 <- as.data.frame(cacti$result)

all_plants <- rbind(p1, p2, p3, p4, p5, p6)
```

# Generate Ecuador plant list
```{r}
ecuador_plants <- merge(all_plants, ecuador_species, by="taxonid")
ecuador_plants_unique <- distinct(ecuador_plants, scientific_name, .keep_all = TRUE)
```

# Summary of data
```{r}
#need better summary info
summary(ecuador_plants)
```

# Write data to csv
```{r}
ecuador_plants_ICUN <- file.path(output_path,"ecuador_plants_IUCN.csv")
write.csv(ecuador_plants, ecuador_plants_ICUN)
```

```{r}
all_habitat_plants <- ecuador_plants %>%
  rowwise %>%
  do(rl_habitats(.$taxonid))
```

### 2022-09-21 archive/BIENDataVisualization.Rmd
add archive files -- old code from Hazel undergrad

---
title: "BIEN data visualization"
author: "Hazel Anderson"
date: "3/1/2021"
output: html_document
---

```{r}
library(BIEN)
library(sp)
library(sf)
library(raster)
library(ggplot2)
library(dplyr)
```

```{r}
#load files
traits <- read.csv("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/L0/Ecuador_species_withTraits.csv")
allplant_rich <- raster("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/analysis/MODIS_data_2019/plant_ric.tif")
fruiting_rich <- raster("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/analysis/MODIS_data_2019/fruitingplant_ric.tif")
forestfruiting_rich <- raster("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/analysis/MODIS_data_2019/fruitingplant_rich_great_50.tif")
```

```{r}
allplant_rich@crs
fruiting_rich@crs
forestfruiting_rich@crs

plot(allplant_rich)
plot(fruiting_rich)
plot(forestfruiting_rich)
```
```{r}
allplant_rich@extent
fruiting_rich@extent
forestfruiting_rich@extent
```
```{r}
allplant_rich.pts <- rasterToPoints(allplant_rich, spatial=TRUE)
fruiting_rich.pts <- rasterToPoints(fruiting_rich, spatial=TRUE)
forestfruiting_rich.pts <- rasterToPoints(forestfruiting_rich, spatial=TRUE)
```

```{r}
# extract values from maps
all_rich.data <- extract(allplant_rich, allplant_rich.pts)
fruiting_rich.data <- extract(fruiting_rich, fruiting_rich.pts)
forestfruiting_rich.data <- extract(forestfruiting_rich, forestfruiting_rich.pts)
```

```{r}
elevation <-raster("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/analysis/MODIS_data_2019/srtm_1km.tif")
elevation@crs
#THE LINE BELOW DOESN"T WORK
ecuador_elevation <-crop(elevation, c(xmin=-91.3884, xmax=-70.00002, ymin=--10.0014, ymax=10))
#suggestion--intersect elevation with richness points; check to make sure they are the same resolution
```
```{r}
#Plotting fruiting species richness vs elevation
#Combine .data and .pts into a dataframe
fruiting_rich.pts$richness <- fruiting_rich.data

#NEED to get richness data and elevation data to be the same length
plot(fruiting_rich.pts, ecuador_elevation)
```

```{r}

subset_traits <- table(traits$scrubbed_species_binomial, traits$trait_name)
subset_traits

#WANT TO PLOT Trait_name by number of species with record for that trait
#ALSO want to map one trait in Space; NEED to subset to all records with values for selected trait
```

### 2022-09-26 code/L0/L0_Ecuador_plants_BIEN.Rmd
Rename L0_1_Ecuador_plants_BIEN.Rmd to L0_Ecuador_plants_BIEN.Rmd

---
title: "Ecuador plant species list"
authors: "Hazel Anderson"
collaborators: "None"
data input: "PlantsOfTheWorldOnline_Ecuador_species.xlsx and GBIF_Ecuador_PlantSpeciesList.csv"
data output: "Ecuador_unique_species_list.csv"
project: "Frugivoria"
date: "2022-08-17"
output: html_document
---
# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
for (package in c("BIEN", "tidyr", "mde")) {
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package, character.only=T)
  }
}
```

# Get data for Ecuador from BIEN database

```{r}
# occurrence data
Ecuador_BIEN_occ <- BIEN_occurrence_country("Ecuador")
# species list
Ecuador_BIEN_species_list <- BIEN_list_country(country = "Ecuador")
# trait data
Ecuador_BIEN_traits <- BIEN_trait_country("Ecuador")
```

## Data summary
```{r}
summary(Ecuador_BIEN_occ)
na_summary(Ecuador_BIEN_occ, exclude_cols = c("date_collected"))
```

```{r}
summary(Ecuador_BIEN_species_list)
na_summary(Ecuador_BIEN_species_list)
```

```{r}
summary(Ecuador_BIEN_traits)
na_summary(Ecuador_BIEN_traits)
```

# Write data to csv
```{r}
Ecuador_BIEN_occurrence_path <- file.path(output_path,"Ecuador_BIEN_occurrence.csv") 
write.csv(Ecuador_BIEN_occ, file = Ecuador_BIEN_occurrence_path)

Ecuador_BIEN_species_list_path <- file.path(output_path, "Ecuador_BIEN_species_list.csv")
write.csv(Ecuador_BIEN_species_list, file = Ecuador_BIEN_species_list_path)

Ecuador_BIEN_traits_path <- file.path(output_path, "Ecuador_BIEN_traits.csv")
write.csv(Ecuador_BIEN_traits, file = Ecuador_BIEN_traits_path)
```

### 2022-09-26 code/L0/L0_Ecuador_plants_IUCN.Rmd
Rename L0_2_Ecuador_plants_IUCN.Rmd to L0_Ecuador_plants_IUCN.Rmd

--- 2022-09-21 code/L0/L0_2_Ecuador_plants_IUCN.Rmd
@@
 ```{r}
-for (package in c("httr", "jsonlite", "rredlist", "dplyr")) {
+for (package in c("httr", "jsonlite", "rredlist", "dplyr", "mde")) {
   if (!require(package, character.only=T, quietly=T)) {
@@
 
-# Generate Ecuador plant list
+# Generate Ecuador plant species list
 ```{r}
 ecuador_plants <- merge(all_plants, ecuador_species, by="taxonid")
-ecuador_plants_unique <- distinct(ecuador_plants, scientific_name, .keep_all = TRUE)
+```
+
+# Get occurences for Ecuador using species list
+```{r}
+get_occ <- function(ID) {
+  rl_occ_country<- rl_occ_country(id=ID,key=token)$result
+  if (class(rl_occ_country) == 'data.frame') {
+    data.frame(taxonid=ID, rl_occ_country)
+  } else {
+    data.frame(taxonid=ID)
+  }
+}
+ecuador_IUCN_plant_occ <- ecuador_plants %>%
+  rowwise %>%
+  do(get_occ(.$taxonid))
+ecuador_IUCN_plant_occ <- as.data.frame(ecuador_IUCN_plant_occ)
+```
+
+# Get habitat for species list
+```{r}
+get_habitat <- function(ID) {
+  habitat_by_sp<- rl_habitats(id=ID,key=token)$result
+  if (class(habitat_by_sp) == 'data.frame') {
+    data.frame(taxonid=ID, habitat_by_sp)
+  } else {
+    data.frame(taxonid=ID)
+  }
+}
+ecuador_IUCN_plant_habitat <- ecuador_plants %>%
+  rowwise %>%
+  do(get_habitat(.$taxonid))
+ecuador_IUCN_plant_habitat <- as.data.frame(ecuador_IUCN_plant_habitat)
 ```
@@
 ```{r}
-#need better summary info
 summary(ecuador_plants)
+na_summary(ecuador_plants, exclude_cols = c("subpopulation"))
+```
+
+```{r}
+summary(ecuador_IUCN_plant_occ)
+na_summary(ecuador_IUCN_plant_occ)
+```
+
+```{r}
+summary(ecuador_IUCN_plant_habitat)
+na_summary(ecuador_IUCN_plant_habitat)
 ```
@@
 ```{r}
-ecuador_plants_ICUN <- file.path(output_path,"ecuador_plants_IUCN.csv")
-write.csv(ecuador_plants, ecuador_plants_ICUN)
+ecuador_ICUN_species_list_path <- file.path(output_path,"Ecuador_IUCN_plant_species.csv")
+write.csv(ecuador_plants, file = ecuador_ICUN_species_list_path)
+
+ecuador_IUCN_plant_occ_path <- file.path(output_path, "Ecuador_IUCN_plant_occ.csv")
+write.csv(ecuador_IUCN_plant_occ, file = ecuador_IUCN_plant_occ_path)
+
+ecuador_IUCN_plant_habitat_path <- file.path(output_path, "Ecuador_IUCN_plant_habitat.csv")
+write.csv(ecuador_IUCN_plant_habitat, file = ecuador_IUCN_plant_habitat_path)
 ```
-
-```{r}
-all_habitat_plants <- ecuador_plants %>%
-  rowwise %>%
-  do(rl_habitats(.$taxonid))
-```

### 2022-10-10 code/L0/L0_2_Ecuador_plants_IUCN.Rmd
update top header information

@@
 ---
-title: "Ecudor plant species IUCN statuses"
-author: "Hazel Anderson"
+title: "Ecudor plant species IUCN data"
+author: "Hazel J. Anderson"
 collaborators: "None"
-data input: "Ecuador_unique_species_list.csv"
-data output: "TBD"
+data input: "None"
+data output: "Ecuador_IUCN_plant_species.csv, Ecuador_IUCN_plant_occ.csv, Ecuador_IUCN_plant_habitat.csv"
 project: "Frugivoria"

### 2022-10-10 code/L0/L0_1_Ecuador_plants_BIEN.Rmd
update top header information

--- 2022-09-26 code/L0/L0_Ecuador_plants_BIEN.Rmd
@@
 ---
-title: "Ecuador plant species list"
-authors: "Hazel Anderson"
+title: "Ecuador plant species BIEN data"
+authors: "Hazel J. Anderson"
 collaborators: "None"
-data input: "PlantsOfTheWorldOnline_Ecuador_species.xlsx and GBIF_Ecuador_PlantSpeciesList.csv"
-data output: "Ecuador_unique_species_list.csv"
+data input: "none"
+data output: "Ecuador_BIEN_species_list.csv, Ecuador_BIEN_occurrence.csv, Ecuador_BIEN_traits.csv"
 project: "Frugivoria"

### 2022-10-10 code/L0/L0_Ecuador_elevation.Rmd
Initalizing

---
title: "Ecuador Elevation"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-10"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Install and load required packages
```{r}
library(elevatr)
library(sf)
library(raster)
```
# Take location coordinates and convert to sf object
```{r}
#x= long, y = lat
ecuador <- data.frame(x = -79.3666965, y = -1.3397668) %>%
  st_as_sf(coords = c("x", "y"), crs = 4326)
```

```{r}
elevation <- get_elev_raster(ecuador, z = 9)
```

### 2022-10-10 code/L0/L0_Ecuador_plants_GBIF.Rmd
Initalizing

---
title: "Ecuador plant species GBIF"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-10"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Install and load required packages
```{r}
library(rgbif)
```

# Set GBIF User information
```{r}
#only run once
#install.packages("usethis")
#usethis::edit_r_environ()
```
# Get Taxon Key for Plants
```{r}
name_backbone("Plantae")
```

```{r}
occ_download(
pred("taxonKey", 6), 
pred("hasGeospatialIssue", FALSE),
pred("hasCoordinate", TRUE),
pred("country","EC"),
pred("occurrenceStatus","PRESENT"), 
pred_not(pred_in("basisOfRecord",c("FOSSIL_SPECIMEN","LIVING_SPECIMEN"))),
format = "SIMPLE_CSV"
)
```

### 2022-10-10 code/L0/L0_Ecuador_plants_GBIF.Rmd
added summary information and write data to csv

@@
 library(rgbif)
+library(mde)
 ```
@@
 ```
+# Retrive GBIF download, save file, and load into r
+```{r}
+d <- occ_download_get("0065583-220831081235567", path = output_path) %>%
+  occ_download_import()
+```
+
+# Summary 
+```{r}
+summary(d)
+na_summary(d, exclude_cols = c("gbifID"))
+```

### 2022-10-10 code/L1/L1_Ecuador_plant_species_BIEN.Rmd
Inital commit; create L1 folder

---
title: "Ecuador plant species BIEN Cleaning"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "Ecuador_BIEN_species_list.csv, Ecuador_BIEN_occurrence.csv, Ecuador_BIEN_traits.csv"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-10"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(taxize)
library(taxList)
```

# Read in BIEN data
```{r}
Ecuador_BIEN_occurrence <- read.csv(file.path(data_path,"Ecuador_BIEN_occurrence.csv"))

Ecuador_BIEN_species_list <- read.csv(file.path(data_path, "Ecuador_BIEN_species_list.csv"))

Ecuador_BIEN_traits_path <- read.csv(file.path(data_path, "Ecuador_BIEN_traits.csv"))
```

```{r}

```

### 2022-10-10 code/L0/L0_Ecuador_plants_POWO.Rmd
Inital commit Download data Timeout error on line 31

---
title: "Ecuador plant species POWO"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "Ecuador_POWO_species_list.csv"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-10"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(kewr)
library(dplyr)
```

# Search Plants of the World Online [https://powo.science.kew.org/] for species in Ecuador
## Best to break up request into smaller chunks so R handle it without timing out
```{r}
Ecuador_POWO_page1 <- search_powo(list(distribution="Ecuador"), limit = 5000)
Ecuador_POWO_page2 <- request_next(Ecuador_POWO_page1)
Ecuador_POWO_page3 <- request_next(Ecuador_POWO_page2)
Ecuador_POWO_page4 <- request_next(Ecuador_POWO_page3)
Ecuador_POWO_page5 <- request_next(Ecuador_POWO_page4)
```

# Combine pages
```{r}
Ecuador_POWO <- bind_rows(
  tidy(Ecuador_POWO_page1),
  tidy(Ecuador_POWO_page2),
  tidy(Ecuador_POWO_page3),
  tidy(Ecuador_POWO_page4),
  tidy(Ecuador_POWO_page5)
)
```

# Summary
```{r}
summary(Ecuador_POWO)
na_summary(Ecuador_POWO)
```

# Write data to csv
```{r}
Ecuador_POWO_path <- file.path(output_path, "Ecuador_POWO.csv")
write.csv(Ecuador_POWO, file = Ecuador_POWO_path)
```

### 2022-10-12 code/L0/L0_Ecuador_plants_POWO.Rmd
fix na summary save as csv

@@
 summary(Ecuador_POWO)
-na_summary(Ecuador_POWO)
+na_summary(Ecuador_POWO, exclude_cols = c("accepted", "url", "images"))
 ```
@@
 ```{r}
+#convert Ecuador_POWO from tbl_df to dataframe
+Ecuador_POWO <- as.data.frame(Ecuador_POWO)
+# remove image column from dataframe
+Ecuador_POWO <- Ecuador_POWO %>% select(-images)
+
 Ecuador_POWO_path <- file.path(output_path, "Ecuador_POWO.csv")

### 2022-10-12 code/L0/L0_Ecuador_elevation.Rmd
change package used to geodata Still getting an error and can't download data (connection error)

@@
 ```{r}
-library(elevatr)
-library(sf)
-library(raster)
-```
-# Take location coordinates and convert to sf object
-```{r}
-#x= long, y = lat
-ecuador <- data.frame(x = -79.3666965, y = -1.3397668) %>%
-  st_as_sf(coords = c("x", "y"), crs = 4326)
+library(geodata)
 ```
@@
 ```{r}
-elevation <- get_elev_raster(ecuador, z = 9)
+ecuador_outline <- gadm(country = "ECU", level=1, path = data_path, version="latest", resolution=1)
+
 ```
+
+# Download elevation data for Ecuador
+```{r}
+ecuador <- elevation_30s(country = "ECU", path = data_path)
+ecuador
+```

### 2022-10-12 code/L1/L1_Ecuador_plants_GBIF.Rmd
Intial commit Set file path, load packages, read in data, standardize data, validate data

---
title: "Ecuador plant species GBIF cleaning"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "TBD"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-12"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
install.packages("remotes")
library("remotes")
install_github("LimaRAF/plantR")
library("plantR")
```

# Read in GBIF data
```{r}
gbif_data <- read.csv(file.path(data_path,"Ecuador_GBIF_occurences.csv"))
# number of rows (records) in imported GBIF data
nrow(gbif_data)
```

# Standardize data
```{r}
# Field names
occs <- formatDwc(gbif_data = gbif_data, drop = TRUE)
dim(occs)
```

```{r}
# Collection codes, people names, collector number and dates
occs <- formatOcc(occs)
# Locality information
occs <- formatLoc(occs)
# Geographical coordinates
occs <- formatCoord(occs)
# Species and family names
occs <- formatTax(occs)
```

# Data validation
```{r}
# Locality information
occs <- validateLoc(occs)
# Geographical coordinates
occs <- validateCoord(occs)
# Species taxonomy and identification
occs <- validateTax(occs)
# Duplicate specimens
occs <- validateDup(occs)
```

# Data Summary and Export
```{r}
summ <- summaryData(occs)
flags <- summaryFlags(occs)
```

```{r}
Ecuador_GBIF_path <- file.path(output_path,"Ecuador_GBIF_occurences_plantR.csv") 
write.csv(d, file = Ecuador_GBIF_path)
```

### 2022-10-12 code/L0/L0_Ecuador_plants_WFO.Rmd
Initial commit download WFO backbone

---
title: "Ecuador plant species WorldFlora"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "TBD"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-11"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(WorldFlora)
```

```{r}
if(! file.exists(file.path(data_path,'WFO.zip'))){
 download.file("http://104.198.143.165/files/WFO_Backbone/_WFOCompleteBackbone/WFO_Backbone.zip", dest=file.path(data_path,"WFO.zip"), mode="wb") 
}
```

```{r}
unzip (file.path(data_path,"WFO.zip"), exdir = data_path)
```

### 2022-10-17 code/L1/L1_Ecuador_plants_GBIF.Rmd
Add checks in data validation Add duplicate specimen code

@@
 ```{r}
-install.packages("remotes")
-library("remotes")
-install_github("LimaRAF/plantR")
+#remotes::install_github("LimaRAF/plantR")
 library("plantR")
@@
 # Field names
-occs <- formatDwc(gbif_data = gbif_data, drop = TRUE)
+occs <- formatDwc(gbif_data = gbif_data)
 dim(occs)
@@
 occs <- validateLoc(occs)
+unique(occs$loc.check)
+```
+
+```{r}
 # Geographical coordinates
 occs <- validateCoord(occs)
+unique(occs$geo.check)
+table(occs$cult.check)
+```
+
+```{r}
 # Species taxonomy and identification
 occs <- validateTax(occs)
+table(occs$tax.check)/dim(occs)[1]
+```
+
+```{r}
 # Duplicate specimens
+dups <- prepDup(occs, 
+                comb.fields =   list(c("family","col.last.name","col.number","col.loc"),
+     c("family","col.last.name","col.number","col.year")))
+dups <- getDup(dups)
+head(dups[order(dups$dup.ID),],6)
+table(dups$dup.prop)
+occs <- cbind.data.frame(occs, 
+                         dups[,c("dup.ID", "dup.numb", "dup.prop")],
+                         stringsAsFactors = FALSE)
+occs <- mergeDup(occs)
+table(occs$tax.check, occs$tax.check1)
+table(occs$geo.check, occs$geo.check1)
+table(occs$loc.check, occs$loc.check1)
 occs <- validateDup(occs)
@@
 flags <- summaryFlags(occs)
+spp.check <- checkList(occs)
 ```
@@
 ```{r}
+saveData(occs, by = "family")
 Ecuador_GBIF_path <- file.path(output_path,"Ecuador_GBIF_occurences_plantR.csv") 

### 2022-10-19 code/L0/L0_Ecuador_plants_WFO.Rmd
Download WFO_Backbone to data folder

@@
 collaborators: "None"
-data input: "TBD"
-data output: "TBD"
+data input: "None"
+data output: "WFO_Backbone"
 project: "Frugivoria"
@@
 ```{r}
-if(! file.exists(file.path(data_path,'WFO.zip'))){
- download.file("http://104.198.143.165/files/WFO_Backbone/_WFOCompleteBackbone/WFO_Backbone.zip", dest=file.path(data_path,"WFO.zip"), mode="wb") 
-}
+dir.create(file.path(data_path, "WFO_Backbone"))
+WFO_path <- file.path(data_path, "WFO_Backbone")
+setwd(WFO_path)
+WFO.download()
+
 ```
-
-```{r}
-unzip (file.path(data_path,"WFO.zip"), exdir = data_path)
-```

### 2022-10-19 code/L0/L0_Ecuador_elevation.Rmd
add save to file for ecuador_geodata_elevation downloading data from elevatr still needs work

@@
 
-# Install and load required packages
+# Load required packages
 ```{r}
 library(geodata)
+library(elevatr)
 ```
 
+# Download elevation data for Ecuador using geodata
 ```{r}
+# get outline of Ecuador
 ecuador_outline <- gadm(country = "ECU", level=1, path = data_path, version="latest", resolution=1)
+ecuador_outline
 
+# get elevation data for Ecuador
+ecuador_geodata <- elevation_30s(country = "ECU", path = data_path)
+ecuador_geodata
+
+# look at elevation data
+plot(ecuador_geodata)
+
+#save elevation data to file
+ecuador_geodata_elevation_path <- file.path(output_path,"ecuador_geodata_elevation.RData") 
+save(ecuador_geodata, file = ecuador_geodata_elevation_path)
 ```
 
-# Download elevation data for Ecuador
+# Download elevation data for Ecuador using 'elevatr'
 ```{r}
-ecuador <- elevation_30s(country = "ECU", path = data_path)
-ecuador
+#NEED TO CONVERT ecuador_outline to spatial dataframe
+ecuador_outline.df <- as.data.frame(ecuador_outline)
+ecuador_elevatr <- get_elev_raster(ecuador_outline.df, z= 5)
 ```

### 2022-10-19 code/L0/L0_Ecuador_plants_GBIF.Rmd
Write aquired data to csv
add more summary information, including plots

@@
 data input: "None"
-data output: "TBD"
+data output: "Ecuador_GBIF_occurences.csv"
 project: "Frugivoria"
@@
 library(mde)
+library(DataExplorer)
+library(summarytools)
 ```
@@
 summary(d)
+```
+
+```{r}
+# using mde package
 na_summary(d, exclude_cols = c("gbifID"))
 ```
+
+```{r}
+# using DataExplorer package
+plot_str(d)
+introduce(d)
+plot_intro(d)
+plot_missing(d)
+plot_bar(d)
+```
+
+```{r}
+# using summarytools package
+dfSummary(d, 
+          plain.ascii  = FALSE, 
+          style        = "grid", 
+          graph.magnif = 0.75, 
+          valid.col    = FALSE,
+          tmp.img.dir  = "/tmp")
+```
+
+# Write data to csv
+```{r}
+Ecuador_GBIF_path <- file.path(output_path,"Ecuador_GBIF_occurences.csv") 
+write.csv(d, file = Ecuador_GBIF_path)
+```

### 2022-10-21 code/L0/L0_Ecuador_elevation.Rmd
get elevation data using elevatr package

@@
 data input: "None"
-data output: "TBD"
+data output: "ecuador_geodata_elevation.RData and ecuador_elevatr_elevation.RData"
 project: "Frugivoria"
-date: "2022-10-10"
+date: "2022-10-21"
 output: html_document
@@
 # get outline of Ecuador
-ecuador_outline <- gadm(country = "ECU", level=1, path = data_path, version="latest", resolution=1)
+ecuador_outline <- gadm(country = "ECU", level=0, path = data_path, version="latest", resolution=1)
 ecuador_outline
@@
 plot(ecuador_geodata)
+plot(ecuador_outline, add = T)
 
@@
 ```{r}
-#NEED TO CONVERT ecuador_outline to spatial dataframe
-ecuador_outline.df <- as.data.frame(ecuador_outline)
+# convert ecuador_outline to spatial dataframe
+ecuador_outline.df <- sf::st_as_sf(ecuador_outline)
+
+# get elevation data for Ecuador
 ecuador_elevatr <- get_elev_raster(ecuador_outline.df, z= 5)
+
+# plot elevation
+plot(ecuador_elevatr)
+plot(ecuador_outline, add = T)
+
+#save elevation data to file
+ecuador_elevatr_elevation_path <- file.path(output_path,"ecuador_elevatr_elevation.RData") 
+save(ecuador_elevatr, file = ecuador_elevatr_elevation_path)
 ```

### 2022-10-26 code/L0/L0_Ecuador_plants_BIEN.Rmd
update summary methods

@@
 ```{r}
-data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
-output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
 ```
@@
 ```{r}
-for (package in c("BIEN", "tidyr", "mde")) {
-  if (!require(package, character.only=T, quietly=T)) {
-    install.packages(package)
-    library(package, character.only=T)
-  }
-}
+library(BIEN)
+library(tidyr)
+library(mde)
+library(DataExplorer)
+library(summarytools)
 ```
@@
 
+### using DataExplorer package
+```{r}
+plot_str(Ecuador_BIEN_occ)
+introduce(Ecuador_BIEN_occ)
+plot_intro(Ecuador_BIEN_occ)
+plot_missing(Ecuador_BIEN_occ)
+plot_bar(Ecuador_BIEN_occ)
+```
+
+```{r}
+plot_str(Ecuador_BIEN_species_list)
+introduce(Ecuador_BIEN_species_list)
+plot_intro(Ecuador_BIEN_species_list)
+plot_missing(Ecuador_BIEN_species_list)
+plot_bar(Ecuador_BIEN_species_list)
+```
+
+```{r}
+plot_str(Ecuador_BIEN_traits)
+introduce(Ecuador_BIEN_traits)
+plot_intro(Ecuador_BIEN_traits)
+plot_missing(Ecuador_BIEN_traits)
+plot_bar(Ecuador_BIEN_traits)
+```
+
+### using summarytools package
+```{r}
+dfSummary(Ecuador_BIEN_occ, 
+          plain.ascii  = FALSE, 
+          style        = "grid", 
+          graph.magnif = 0.75, 
+          valid.col    = FALSE,
+          tmp.img.dir  = "/tmp")
+```
+
+```{r}
+dfSummary(Ecuador_BIEN_species_list, 
+          plain.ascii  = FALSE, 
+          style        = "grid", 
+          graph.magnif = 0.75, 
+          valid.col    = FALSE,
+          tmp.img.dir  = "/tmp")
+```
+
+```{r}
+dfSummary(Ecuador_BIEN_traits, 
+          plain.ascii  = FALSE, 
+          style        = "grid", 
+          graph.magnif = 0.75, 
+          valid.col    = FALSE,
+          tmp.img.dir  = "/tmp")
+```
+
 # Write data to csv

### 2022-10-26 code/L0/L0_Ecuador_plants_BIEN.Rmd
output summarytools results to html in output file path

@@
 ```{r}
-dfSummary(Ecuador_BIEN_occ, 
-          plain.ascii  = FALSE, 
-          style        = "grid", 
-          graph.magnif = 0.75, 
-          valid.col    = FALSE,
-          tmp.img.dir  = "/tmp")
+Ecuador_BIEN_occ_summary <- dfSummary(Ecuador_BIEN_occ, 
+            na.col       = FALSE,
+            style        = "multiline",
+            plain.ascii  = TRUE,
+            graph.magnif = .8)
+print(Ecuador_BIEN_occ_summary, method = "render")
+view(Ecuador_BIEN_occ_summary, file = file.path(output_path,"Ecuador_BIEN_occ_summary.html"))
 ```
@@
 ```{r}
-dfSummary(Ecuador_BIEN_species_list, 
-          plain.ascii  = FALSE, 
-          style        = "grid", 
-          graph.magnif = 0.75, 
-          valid.col    = FALSE,
-          tmp.img.dir  = "/tmp")
+Ecuador_BIEN_species_list_summary <- dfSummary(Ecuador_BIEN_species_list, 
+            na.col       = FALSE,
+            style        = "multiline",
+            plain.ascii  = TRUE,
+            graph.magnif = .8)
+print(Ecuador_BIEN_species_list_summary, method = "render")
+view(Ecuador_BIEN_species_list_summary, file = file.path(output_path,"Ecuador_BIEN_species_list_summary.html"))
 ```
@@
 ```{r}
-dfSummary(Ecuador_BIEN_traits, 
-          plain.ascii  = FALSE, 
-          style        = "grid", 
-          graph.magnif = 0.75, 
-          valid.col    = FALSE,
-          tmp.img.dir  = "/tmp")
+Ecuador_BIEN_traits_summary <- dfSummary(Ecuador_BIEN_traits, 
+            na.col       = FALSE,
+            style        = "multiline",
+            plain.ascii  = TRUE,
+            graph.magnif = .8)
+print(Ecuador_BIEN_traits_summary, method = "render")
+view(Ecuador_BIEN_traits_summary, file = file.path(output_path,"Ecuador_BIEN_traits_summary.html"))
 ```

### 2022-10-27 code/L0/L0_Ecuador_plants_IUCN.Rmd
update title, correct spelling error

@@
 ---
-title: "Ecudor plant species IUCN statuses"
-author: "Hazel Anderson"
+title: "Ecuador plant species IUCN data"
+author: "Hazel J. Anderson"
 collaborators: "None"
-data input: "Ecuador_unique_species_list.csv"
-data output: "TBD"
+data input: "None"
+data output: "Ecuador_IUCN_plant_species.csv, Ecuador_IUCN_plant_occ.csv, Ecuador_IUCN_plant_habitat.csv"
 project: "Frugivoria"

### 2022-10-27 code/L0/L0_Ecuador_plants_TRY.Rmd
inital commit with header information

---
title: "Ecuador plant species TRY data"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-27"
output: html_document
---

Public data path "G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/TRY_Ecuador_PublicData_10192020.txt"

### 2022-10-31 code/L0/L0_Ecuador_plants_TRY.Rmd
add file paths, load required packages, read in data, explore TRY data using rtry package

@@
 
-Public data path "G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/TRY_Ecuador_PublicData_10192020.txt"
+# Set file paths
+```{r}
+data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+```
+
+# Load required packages
+```{r}
+library(rtry)
+```
+
+# Read in TRY data
+Data was obtained from https://www.try-db.org/TryWeb/Prop0.php 
+```{r}
+data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/TRY_Ecuador_PublicData_10192020.txt")
+```
+# Explore TRY data
+```{r}
+# explore by trait id and trait name
+TRYdata_explore_trait <- rtry_explore(data, TraitID, TraitName)
+TRYdata_explore_trait
+
+# explore by species id, species name, trait id, and trait name
+TRYdata_explore_species <- rtry_explore(data, AccSpeciesID, AccSpeciesName, TraitID, TraitName)
+TRYdata_explore_species
+
+# explore by data id, data name, trait id, and trait name sorted by trait id
+TRYdata_explore_anc <- rtry_explore(data, DataID, DataName, TraitID, TraitName, sortBy = TraitID)
+TRYdata_explore_anc
+```

### 2022-10-31 code/L0/L0_Ecuador_plants_TRY.Rmd
add summary information using mde, DataExplorer, and summarytools packages summarytools summary is giving error

@@
 library(rtry)
+library(mde)
+library(DataExplorer)
+library(summarytools)
 ```
@@
 ```
+
+# Summary of data
+```{r}
+summary(data)
+na_summary(data, , exclude_cols = c("RelUncertaintyPercent", "V28"))
+```
+
+### using DataExplorer package
+```{r}
+plot_str(data)
+introduce(data)
+plot_intro(data)
+plot_missing(data)
+plot_bar(data)
+```
+
+### using summarytools package
+```{r}
+ecuador_TRY_plant_traits_summary <- dfSummary(data, 
+            na.col       = FALSE,
+            style        = "multiline",
+            plain.ascii  = TRUE,
+            graph.magnif = .8)
+print(ecuador_TRY_plant_traits_summary, method = "render")
+view(ecuador_TRY_plant_traits_summary, file = file.path(output_path,"ecuador_TRY_plant_traits_summary.html"))
+# gives the following error: Error in withCallingHandlers(expr, warning = function(w) if (inherits(w, : 
+#   invalid multibyte string at '<b0>45''
+```
+
+# Write data to csv
+```{r}
+ecuador_TRY_plant_traits_path <- file.path(output_path,"Ecuador_TRY_plant_traits.csv")
+write.csv(data, file = ecuador_TRY_plant_traits_path)
+
+```

### 2022-10-31 code/L0/L0_Ecuador_plants_GBIF.Rmd
add summary with summarytools package

@@
           tmp.img.dir  = "/tmp")
+d_summary <- dfSummary(d, 
+            na.col       = FALSE,
+            style        = "multiline",
+            plain.ascii  = TRUE,
+            graph.magnif = .8)
+print(d_summary, method = "render")
+view(d_summary, file = file.path(output_path,"Ecuador_GBIF_occurences_summary.html"))
 ```

### 2022-10-31 code/L1/L1_Ecuador_plants_GBIF.Rmd
formatting changes and removing unnecessary duplicate code

@@
 collaborators: "None"
-data input: "TBD"
+data input: "Ecuador_GBIF_occurences.csv"
 data output: "TBD"
@@
 occs <- formatOcc(occs)
+```
+
+```{r}
 # Locality information
 occs <- formatLoc(occs)
+```
+
+```{r}
 # Geographical coordinates
 occs <- formatCoord(occs)
+```
+
+```{r}
 # Species and family names
@@
 # Duplicate specimens
-dups <- prepDup(occs, 
-                comb.fields =   list(c("family","col.last.name","col.number","col.loc"),
-     c("family","col.last.name","col.number","col.year")))
-dups <- getDup(dups)
-head(dups[order(dups$dup.ID),],6)
-table(dups$dup.prop)
-occs <- cbind.data.frame(occs, 
-                         dups[,c("dup.ID", "dup.numb", "dup.prop")],
-                         stringsAsFactors = FALSE)
-occs <- mergeDup(occs)
-table(occs$tax.check, occs$tax.check1)
-table(occs$geo.check, occs$geo.check1)
-table(occs$loc.check, occs$loc.check1)
 occs <- validateDup(occs)
@@
 ```{r}
-saveData(occs, by = "family")
+saveData(occs, by = "geo", path = file.path(output_path, "Ecuador_GBIF_occurences_plantR_bygeo.csv"))
+saveData(occs, by = "tax", path = file.path(output_path, "Ecuador_GBIF_occurences_plantR_bytax.csv"))
 Ecuador_GBIF_path <- file.path(output_path,"Ecuador_GBIF_occurences_plantR.csv") 

### 2022-10-31 code/L0/L0_Ecuador_plants_IUCN.Rmd
add summary information using DataExplorer and summarytools packages

@@
 ```{r}
-for (package in c("httr", "jsonlite", "rredlist", "dplyr", "mde")) {
-  if (!require(package, character.only=T, quietly=T)) {
-    install.packages(package)
-    library(package, character.only=T)
-  }
-}
+library(httr)
+library(jsonlite)
+library(rredlist)
+library(dplyr)
+library(mde)
+library(DataExplorer)
+library(summarytools)
 ```
+
 # Set up token and api url

### 2022-10-31 code/L0/L0_Ecuador_plants_IUCN.Rmd
add summary information using DataExplorer and summarytools packages

@@
 
+### using DataExplorer package
+```{r}
+plot_str(ecuador_plants)
+introduce(ecuador_plants)
+plot_intro(ecuador_plants)
+plot_missing(ecuador_plants)
+plot_bar(ecuador_plants)
+```
+
+```{r}
+plot_str(ecuador_IUCN_plant_occ)
+introduce(ecuador_IUCN_plant_occ)
+plot_intro(ecuador_IUCN_plant_occ)
+plot_missing(ecuador_IUCN_plant_occ)
+plot_bar(ecuador_IUCN_plant_occ)
+```
+
+```{r}
+plot_str(ecuador_IUCN_plant_habitat)
+introduce(ecuador_IUCN_plant_habitat)
+plot_intro(ecuador_IUCN_plant_habitat)
+plot_missing(ecuador_IUCN_plant_habitat)
+plot_bar(ecuador_IUCN_plant_habitat)
+```
+
+### using summarytools package
+```{r}
+ecuador_plants_summary <- dfSummary(ecuador_plants, 
+            na.col       = FALSE,
+            style        = "multiline",
+            plain.ascii  = TRUE,
+            graph.magnif = .8)
+print(ecuador_plants_summary, method = "render")
+view(ecuador_plants_summary, file = file.path(output_path,"ecuador_plants_summary.html"))
+```
+
+```{r}
+ecuador_IUCN_plant_occ_summary <- dfSummary(ecuador_IUCN_plant_occ, 
+            na.col       = FALSE,
+            style        = "multiline",
+            plain.ascii  = TRUE,
+            graph.magnif = .8)
+print(ecuador_IUCN_plant_occ_summary, method = "render")
+view(ecuador_IUCN_plant_occ_summary, file = file.path(output_path,"ecuador_IUCN_plant_occ_summary.html"))
+```
+
+```{r}
+ecuador_IUCN_plant_habitat_summary <- dfSummary(ecuador_IUCN_plant_habitat, 
+            na.col       = FALSE,
+            style        = "multiline",
+            plain.ascii  = TRUE,
+            graph.magnif = .8)
+print(ecuador_IUCN_plant_habitat_summary, method = "render")
+view(ecuador_IUCN_plant_habitat_summary, file = file.path(output_path,"ecuador_IUCN_plant_habitat_summary.html"))
+```
+
 # Write data to csv

### 2022-11-09 code/L1/L1_Ecuador_plants_BIEN.Rmd
change packages, add pre-filtering code for BIEN_occurrences

---
title: "Ecuador plant species BIEN Cleaning"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "Ecuador_BIEN_species_list.csv, Ecuador_BIEN_occurrence.csv, Ecuador_BIEN_traits.csv"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-10"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(bdc)
```

# Read in BIEN data
```{r}
Ecuador_BIEN_occurrence <- read.csv(file.path(data_path,"Ecuador_BIEN_occurrence.csv"))
Ecuador_BIEN_species_list <- read.csv(file.path(data_path, "Ecuador_BIEN_species_list.csv"))
Ecuador_BIEN_traits_path <- read.csv(file.path(data_path, "Ecuador_BIEN_traits.csv"))
```

# Pre-filter
```{r}
# Flag records missing species names
check_pf <-
  bdc_scientificName_empty(
  data = Ecuador_BIEN_occ,
  sci_name = "scrubbed_species_binomial")
```

```{r}
# Flag records missing partial or complete information on geographic coordinates
check_pf <- bdc_coordinates_empty(
  data = check_pf,
  lat = "latitude",
  lon = "longitude")
```

```{r}
# Flag records with out-of-range coordinates: latitude > 90 or -90; longitude >180 or -180
check_pf <- bdc_coordinates_outOfRange(
  data = check_pf,
  lat = "latitude",
  lon = "longitude")

```
```{r}
# Deriving country names for records missing country names
check_pf <- bdc_country_from_coordinates(
  data = check_pf,
  lat = "latitude",
  lon = "longitude",
  country = "country")
```

```{r}
# Country names are standardized against a list of country names in several languages retrieved from Wikipedia
check_pf <- bdc_country_standardized(
  data = check_pf,
  country = "country"
)
```

```{r}
# Correcting latitude and longitude transposed
check_pf <-
  bdc_coordinates_transposed(
    data = check_pf,
    id = "database_id",
    sci_names = "scrubbed_species_binomial",
    lat = "decimalLatitude",
    lon = "decimalLongitude",
    country = "country_suggested",
    countryCode = "countryCode",
    border_buffer = 0.2, # in decimal degrees (~22 km at the equator)
    save_outputs = FALSE
  )
```

```{r}
# Records outside one or multiple reference countries
check_pf <-
  bdc_coordinates_country_inconsistent(
    data = check_pf,
    country_name = "Ecuador",
    country = "country_suggested",
    lon = "longitude",
    lat = "latitude",
    dist = 0.1 # in decimal degrees (~11 km at the equator)
  )
```

```{r}
# Report
check_pf <- bdc_summary_col(data = check_pf)
report <-
  bdc_create_report(data = check_pf,
                    database_id = "database_id",
                    workflow_step = "prefilter",
                    save_report = FALSE)
report
```

```{r}
# Figures
figures <-
bdc_create_figures(data = check_pf,
                   database_id = "database_id",
                   workflow_step = "prefilter",
                   save_figures = FALSE)

# Check figures using
figures$.coordinates_country_inconsistent

```

```{r}
# Filter out flags
output <-
  check_pf %>%
  dplyr::filter(.summary == TRUE) %>%
  bdc_filter_out_flags(data = ., col_to_remove = "all")
```

### 2022-11-14 code/L1/L1_Ecuador_plants_TRY.Rmd
inital commit set file paths, load packages, read in try data

---
title: "Ecuador plant trait TRY Cleaning"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "Ecuador_TRY_plant_traits.csv"
data output: "TBD"
project: "Frugivoria"
date: "2022-11-14"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(rtry)
```

# Read in IUCN occurrence data
```{r}
try_data <- read.csv(file.path(data_path,"Ecuador_TRY_plant_traits.csv"))
```

# Explore data
```{r}
TRYdata_explore_trait <- rtry_explore(try_data, TraitID, TraitName)
View(TRYdata_explore_trait)
```

### 2022-11-16 code/L1/L1_Ecuador_plants_TRY.Rmd
Add code to transform dataframe to wide table

@@
 
-# Explore data
+# Transform to wide table
 ```{r}
-TRYdata_explore_trait <- rtry_explore(try_data, TraitID, TraitName)
-View(TRYdata_explore_trait)
+#-------------------------------------------------
+# Exclude
+# 1. All entries with "" in TraitID
+# 2. Potential categorical traits that don't have a StdValue
+# 3. Traits that have not yet been standardized in TRY
+# Then select the relevant columns for transformation
+# Note: The complete.cases() is used to ensure the cases are complete,
+#       i.e. have no missing values
+#-------------------------------------------------
+num_traits <- rtry_select_row(try_data, complete.cases(TraitID) & complete.cases(StdValue))
+num_traits <- rtry_select_col(num_traits, ObservationID, AccSpeciesID, AccSpeciesName, TraitID, TraitName, StdValue, UnitName)
+# Extract the unique value of latitude (DataID 59) and the corresponding ObservationID
+workdata_lat <- rtry_select_anc(try_data, 59)
+
+# Extract the unique value of longitude (DataID 60) and the corresponding ObservationID
+workdata_lon <- rtry_select_anc(try_data, 60)
+
+# To merge the extracted ancillary data with the numerical traits
+# Merge the relevant data frames based on the ObservationID using rtry_join_left (left join)
+num_traits_georef <- rtry_join_left(num_traits, workdata_lat, baseOn = ObservationID)
+num_traits_georef <- rtry_join_left(num_traits_georef, workdata_lon, baseOn = ObservationID)
+
+# Perform wide table transformation on TraitID, TraitName and UnitName
+# With cell values to be the mean values calculated for StdValue
+num_traits_georef_wider <- rtry_trans_wider(num_traits_georef, names_from = c(TraitID, TraitName, UnitName), values_from = c(StdValue), values_fn = list(StdValue = mean))
 ```

### 2022-11-28 code/L0/L0_Ecuador_plants_BIEN.Rmd
updated data pulls from BIEN to include all taxonomy, observation type and colleciton info

@@
 # occurrence data
-Ecuador_BIEN_occ <- BIEN_occurrence_country("Ecuador")
+Ecuador_BIEN_occ <- BIEN_occurrence_country("Ecuador", all.taxonomy = TRUE, observation.type = TRUE, collection.info = TRUE)
 # species list
@@
 # trait data
-Ecuador_BIEN_traits <- BIEN_trait_country("Ecuador")
+Ecuador_BIEN_traits <- BIEN_trait_country("Ecuador", all.taxonomy = TRUE)
 ```
@@
 summary(Ecuador_BIEN_occ)
-na_summary(Ecuador_BIEN_occ, exclude_cols = c("date_collected"))
+na_summary(Ecuador_BIEN_occ, exclude_cols = c("date_collected", "date_identified"))
 ```
@@
 plot_intro(Ecuador_BIEN_occ)
-plot_missing(Ecuador_BIEN_occ)
-plot_bar(Ecuador_BIEN_occ)
+#plot_missing(Ecuador_BIEN_occ)
+#plot_bar(Ecuador_BIEN_occ)
 ```
@@
 plot_intro(Ecuador_BIEN_species_list)
-plot_missing(Ecuador_BIEN_species_list)
-plot_bar(Ecuador_BIEN_species_list)
 ```

### 2022-11-30 code/L2/L2_Ecuador_plant_occ.Rmd
initial commit set up data merge with bdc package

---
title: "Ecuador plant occurrence data merging"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "TBD"
data output: "TBD"
project: "Frugivoria"
date: "2022-11-30"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
```

# Load required packages
```{r}
library(bdc)
```

# Read in the configuration table
```{r}
metadata <-
  readr::read_csv(system.file("ADD FILE PATH HERE",
                              package = "bdc"),
                  show_col_types = FALSE)
```

# Read in BIEN, GBIF, and IUCN occurrence data
```{r}
# Path to the folder containing the example datasets. For instance:
path <- data_path

# Change in the Configuration table the path to the folder in your computer containing the example datasets
metadata$fileName <-
  gsub(pattern = "https://raw.githubusercontent.com/brunobrr/bdc/master/inst/extdata/input_files/",
       replacement = path,
       x = metadata$fileName)
```

```{r}
database <-
bdc_standardize_datasets(metadata = metadata,
                         format = "csv",
                         overwrite = TRUE,
                         save_database = TRUE)
```

### 2022-11-30 code/L1/L1_Ecuador_plants_BIEN.Rmd
change variable name

@@
   bdc_scientificName_empty(
-  data = Ecuador_BIEN_occ,
+  data = Ecuador_BIEN_occurrence,
   sci_name = "scrubbed_species_binomial")

### 2022-11-30 code/L1/L1_Ecuador_plants_TRY.Rmd
remove duplicates and preprocess TRY data export preprocessed data to csv

@@
 
-# Read in IUCN occurrence data
+# Import TRY data
 ```{r}
 try_data <- read.csv(file.path(data_path,"Ecuador_TRY_plant_traits.csv"))
+```
+
+```{r}
+rtry_explore(try_data, DataID, DataName, TraitID, TraitName, sortBy = TraitID)
+```
+
+```{r}
+# Remove duplicates
+workdata <- rtry_remove_dup(try_data)
 ```
@@
 #-------------------------------------------------
-num_traits <- rtry_select_row(try_data, complete.cases(TraitID) & complete.cases(StdValue))
+num_traits <- rtry_select_row(workdata, complete.cases(TraitID) & complete.cases(StdValue))
 num_traits <- rtry_select_col(num_traits, ObservationID, AccSpeciesID, AccSpeciesName, TraitID, TraitName, StdValue, UnitName)
@@
 ```
+
+# Export preprocessed TRY data
+```{r}
+# Export the data into a CSV file
+ecuador_TRY_plant_traits_path <- file.path(output_path,"Ecuador_TRY_plant_traits.csv")
+write.csv(num_traits_georef_wider, file = ecuador_TRY_plant_traits_path)
+```

### 2022-12-02 code/L1/L1_Ecuador_plants_BIEN_traits.Rmd
use consistent data cleaning methods seperate out BIEN occ cleaning from BIEN trait cleaning

---
title: "Ecuador plant species BIEN trait cleaning"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "Ecuador_BIEN_traits.csv"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-10"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(bdc)
library(bdDwC)
```

# Read in BIEN trait data
```{r}
database <- read.csv(file.path(data_path, "Ecuador_BIEN_traits.csv"))
```

# Configure BIEN trait data to Darwin Core Standards
```{r}
# darwinizing a dataset
config <- read.csv(file.path(output_path,"configuration table.csv"))

# rename dataset fields
database <- renameUserData(database, config)
```

# Pre-filter
```{r}
# Flag records missing species names
check_pf <-
  bdc_scientificName_empty(
  data = database,
  sci_name = "scientificName")
```

```{r}
# Flag records missing partial or complete information on geographic coordinates
check_pf <- bdc_coordinates_empty(
  data = check_pf,
  lat = "decimalLatitude",
  lon = "decimalLongitude")
```

```{r}
# Flag records with out-of-range coordinates: latitude > 90 or -90; longitude >180 or -180
check_pf <- bdc_coordinates_outOfRange(
  data = check_pf,
  lat = "decimalLatitude",
  lon = "decimalLongitude")

```

```{r}
# Check record sources of your dataset using:
check_pf %>%
  dplyr::group_by(basisOfRecord) %>%
  dplyr::summarise(n = dplyr::n())

check_pf <- bdc_basisOfRecords_notStandard(
  data = check_pf,
  basisOfRecord = "basisOfRecord",
  names_to_keep = "all")
```

```{r}
# Deriving country names for records missing country names
check_pf <- bdc_country_from_coordinates(
  data = check_pf,
  lat = "decimalLatitude",
  lon = "decimalLongitude",
  country = "country")
```

```{r}
# Standardizing country names and getting country code information
check_pf <- bdc_country_standardized(
  data = check_pf,
  country = "country"
)
```

```{r}
# Correcting latitude and longitude transposed
check_pf <-
  bdc_coordinates_transposed(
    data = check_pf,
    id = "collectionID",
    sci_names = "scientificName",
    lat = "decimalLatitude",
    lon = "decimalLongitude",
    country = "country_suggested",
    countryCode = "countryCode",
    border_buffer = 0.2, # in decimal degrees (~22 km at the equator)
    save_outputs = FALSE
  )
```

```{r}
# Records outside one or multiple reference countries
check_pf <-
  bdc_coordinates_country_inconsistent(
    data = check_pf,
    country_name = "Ecuador",
    country = "country_suggested",
    lon = "decimalLongitude",
    lat = "decimalLatitude",
    dist = 0.1 # in decimal degrees (~11 km at the equator)
  )
```

```{r}
# Report
check_pf <- bdc_summary_col(data = check_pf)
report <-
  bdc_create_report(data = check_pf,
                    database_id = "database_id",
                    workflow_step = "prefilter",
                    save_report = FALSE)
report
```

```{r}
# Figures
figures <-
bdc_create_figures(data = check_pf,
                   database_id = "database_id",
                   workflow_step = "prefilter",
                   save_figures = FALSE)

# Check figures using
figures$.coordinates_country_inconsistent

```

```{r}
# Filter out flags
database <-
  check_pf %>%
  dplyr::filter(.summary == TRUE) %>%
  bdc_filter_out_flags(data = ., col_to_remove = "all")
```

# Flagging common time issues
```{r}
check_time <-
  bdc_eventDate_empty(data = database, eventDate = "verbatimEventDate")
```

```{r}
check_time <-
  bdc_year_from_eventDate(data = check_time, eventDate = "verbatimEventDate")
```

```{r}
check_time <-
  bdc_year_outOfRange(data = check_time,
                      eventDate = "year",
                      year_threshold = 1800)
```

```{r}
check_time <- bdc_summary_col(data = check_time)
```

```{r}
report <-
  bdc_create_report(data = check_time,
                    database_id = "database_id",
                    workflow_step = "time",
                    save_report = FALSE)

report
```

```{r}
figures <-
bdc_create_figures(data = check_time,
                   database_id = "database_id",
                   workflow_step = "time",
                   save_figures = FALSE)

# Check figures using
figures$time_year_BAR
```

```{r}
database <-
  check_time %>%
  dplyr::filter(.summary == TRUE) %>%
  bdc_filter_out_flags(data = ., col_to_remove = "all")
```

# Flagging common spatial issues
```{r}
check_space <-
  bdc_coordinates_precision(
    data = database,
    lon = "decimalLongitude",
    lat = "decimalLatitude",
    ndec = c(0, 1) # number of decimals to be tested
  )
```

```{r}
check_space <-
  CoordinateCleaner::clean_coordinates(
    x =  check_space,
    lon = "decimalLongitude",
    lat = "decimalLatitude",
    species = "scientificName",
    countries = ,
    tests = c(
      "capitals",     # records within 2km around country and province centroids
      "centroids",    # records within 1km of capitals centroids
      "duplicates",   # duplicated records
      "equal",        # records with equal coordinates
      "gbif",         # records within 1 degree (~111km) of GBIF headsquare
      "institutions", # records within 100m of zoo and herbaria
      "outliers",     # outliers
      "zeros",        # records with coordinates 0,0
      "urban"         # records within urban areas
    ),
    capitals_rad = 2000,
    centroids_rad = 1000,
    centroids_detail = "both", # test both country and province centroids
    inst_rad = 100, # remove zoo and herbaria within 100m
    outliers_method = "quantile",
    outliers_mtp = 5,
    outliers_td = 1000,
    outliers_size = 10,
    range_rad = 0,
    zeros_rad = 0.5,
    capitals_ref = NULL,
    centroids_ref = NULL,
    country_ref = NULL,
    country_refcol = "countryCode",
    inst_ref = NULL,
    range_ref = NULL,
    # seas_ref = continent_border,
    # seas_scale = 110,
    urban_ref = NULL,
    value = "spatialvalid" # result of tests are appended in separate columns
  )
```

```{r}
check_space <- bdc_summary_col(data = check_space)

check_space %>%
  dplyr::filter(.summary == FALSE) %>% # map only records flagged as FALSE
  bdc_quickmap(
    data = .,
    lon = "decimalLongitude",
    lat = "decimalLatitude",
    col_to_map = ".summary",
    size = 0.9
  )
```

```{r}
report <-
  bdc_create_report(data = check_space,
                    database_id = "database_id",
                    workflow_step = "space",
                    save_report = FALSE)

report
```

```{r}
figures <-
bdc_create_figures(data = check_space,
                   database_id = "database_id",
                   workflow_step = "space",
                   save_figures = TRUE)

# Check figures using
figures$.rou
```

```{r}
database <-
  check_space %>%
  dplyr::filter(.summary == TRUE) %>%
  bdc_filter_out_flags(data = ., col_to_remove = "all")
```

# Clean and parse species names
```{r}
parse_names <-
  bdc_clean_names(sci_names = database$scientificName, save_outputs = FALSE)
```

```{r}
parse_names <-
  parse_names %>%
  dplyr::select(.uncer_terms, names_clean)

database <- dplyr::bind_cols(database, parse_names)
```

```{r}
query_names <- bdc_query_names_taxadb(
  sci_name            = database$names_clean,
  replace_synonyms    = TRUE, # replace synonyms by accepted names?
  suggest_names       = TRUE, # try to found a candidate name for misspelled names?
  suggestion_distance = 0.9, # distance between the searched and suggested names
  db                  = "gbif", # taxonomic database
  parallel            = FALSE, # should parallel processing be used?
  ncores              = 2, # number of cores to be used in the parallelization process
  export_accepted     = FALSE # save names linked to multiple accepted names
)
```

```{r}
database <-
  database %>%
  dplyr::rename(verbatim_scientificName = scientificName) %>%
  dplyr::select(-names_clean) %>%
  dplyr::bind_cols(., query_names)
```

```{r}
report <-
  bdc_create_report(data = database,
                    database_id = "database_id",
                    workflow_step = "taxonomy",
                    save_report = FALSE)

report
```

```{r}
unresolved_names <-
  bdc_filter_out_names(data = database,
                       col_name = "notes",
                       taxonomic_status = "accepted",
                       opposite = TRUE)
```

```{r}
output <-
   bdc_filter_out_names(
     data = database,
     taxonomic_notes = "accepted",
     opposite = FALSE
   )
```

# Save output data to csv
```{r}
# Export the data into a CSV file
ecuador_BIEN_plant_traits_cleaned_path <- file.path(output_path,"Ecuador_BIEN_plant_traits_cleaned.csv")
write.csv(output, file = ecuador_BIEN_plant_traits_cleaned_path)
ecuador_BIEN_plant_traits_unresolved_path <- file.path(output_path,"Ecuador_BIEN_plant_traits_unresolved.csv")
write.csv(unresolved_names, file = ecuador_BIEN_plant_traits_unresolved_path)
```

### 2022-12-02 code/L1/L1_Ecuador_plants_TRY.Rmd
simple rename

@@
 
-# Export preprocessed TRY data
+# Export TRY data
 ```{r}

### 2022-12-02 code/L1/L1_Ecuador_plants_BIEN_occ.Rmd
Rename L1_Ecuador_plants_BIEN.Rmd to L1_Ecuador_plants_BIEN_occ.Rmd

--- 2022-12-02 code/L1/L1_Ecuador_plants_BIEN_traits.Rmd
@@
 ---
-title: "Ecuador plant species BIEN trait cleaning"
+title: "Ecuador plant species BIEN occurrence cleaning"
 author: "Hazel J. Anderson"
 collaborators: "None"
-data input: "Ecuador_BIEN_traits.csv"
-data output: "TBD"
+data input: "Ecuador_BIEN_occurrence.csv"
+data output: "Ecuador_BIEN_plant_occ_cleaned.csv", "Ecuador_BIEN_plant_occ_unresolved.csv"
 project: "Frugivoria"
@@
 
-# Read in BIEN trait data
+# Read in BIEN occ data
 ```{r}
-database <- read.csv(file.path(data_path, "Ecuador_BIEN_traits.csv"))
+database <- read.csv(file.path(data_path,"Ecuador_BIEN_occurrence.csv"))
 ```
 
-# Configure BIEN trait data to Darwin Core Standards
+# Configure BIEN data to Darwin Core Standards
 ```{r}
@@
 # Export the data into a CSV file
-ecuador_BIEN_plant_traits_cleaned_path <- file.path(output_path,"Ecuador_BIEN_plant_traits_cleaned.csv")
-write.csv(output, file = ecuador_BIEN_plant_traits_cleaned_path)
-ecuador_BIEN_plant_traits_unresolved_path <- file.path(output_path,"Ecuador_BIEN_plant_traits_unresolved.csv")
-write.csv(unresolved_names, file = ecuador_BIEN_plant_traits_unresolved_path)
+ecuador_BIEN_plant_occ_cleaned_path <- file.path(output_path,"Ecuador_BIEN_plant_occ_cleaned.csv")
+write.csv(output, file = ecuador_BIEN_plant_occ_cleaned_path)
+ecuador_BIEN_plant_occ_unresolved_path <- file.path(output_path,"Ecuador_BIEN_plant_occ_unresolved.csv")
+write.csv(unresolved_names, file = ecuador_BIEN_plant_occ_unresolved_path)
 ```

### 2022-12-02 code/L1/L1_Ecuador_plants_TRY.Rmd
add cleaning data and move widetransform to the end

@@
 library(rtry)
+library(bdc)
+library(bdDwC)
 ```
@@
 try_data <- read.csv(file.path(data_path,"Ecuador_TRY_plant_traits.csv"))
+```
+
+# Configure TRY data to Darwin Core Standards
+```{r}
+# darwinizing a dataset
+config <- read.csv(file.path(output_path,"Ecuador_plant_TRY_trait_ConfigurationTable"))
+
+# rename dataset fields
+database <- renameUserData(database, config)
+```
+
+# Pre-filter
+```{r}
+# Flag records missing species names
+check_pf <-
+  bdc_scientificName_empty(
+  data = database,
+  sci_name = "scientificName")
+```
+
+```{r}
+# Flag records missing partial or complete information on geographic coordinates
+check_pf <- bdc_coordinates_empty(
+  data = check_pf,
+  lat = "decimalLatitude",
+  lon = "decimalLongitude")
+```
+
+```{r}
+# Flag records with out-of-range coordinates: latitude > 90 or -90; longitude >180 or -180
+check_pf <- bdc_coordinates_outOfRange(
+  data = check_pf,
+  lat = "decimalLatitude",
+  lon = "decimalLongitude")
+
+```
+
+```{r}
+# Check record sources of your dataset using:
+check_pf %>%
+  dplyr::group_by(basisOfRecord) %>%
+  dplyr::summarise(n = dplyr::n())
+
+check_pf <- bdc_basisOfRecords_notStandard(
+  data = check_pf,
+  basisOfRecord = "basisOfRecord",
+  names_to_keep = "all")
+```
+
+```{r}
+# Deriving country names for records missing country names
+check_pf <- bdc_country_from_coordinates(
+  data = check_pf,
+  lat = "decimalLatitude",
+  lon = "decimalLongitude",
+  country = "country")
+```
+
+```{r}
+# Standardizing country names and getting country code information
+check_pf <- bdc_country_standardized(
+  data = check_pf,
+  country = "country"
+)
+```
+
+```{r}
+# Correcting latitude and longitude transposed
+check_pf <-
+  bdc_coordinates_transposed(
+    data = check_pf,
+    id = "collectionID",
+    sci_names = "scientificName",
+    lat = "decimalLatitude",
+    lon = "decimalLongitude",
+    country = "country_suggested",
+    countryCode = "countryCode",
+    border_buffer = 0.2, # in decimal degrees (~22 km at the equator)
+    save_outputs = FALSE
+  )
+```
+
+```{r}
+# Records outside one or multiple reference countries
+check_pf <-
+  bdc_coordinates_country_inconsistent(
+    data = check_pf,
+    country_name = "Ecuador",
+    country = "country_suggested",
+    lon = "decimalLongitude",
+    lat = "decimalLatitude",
+    dist = 0.1 # in decimal degrees (~11 km at the equator)
+  )
+```
+
+```{r}
+# Report
+check_pf <- bdc_summary_col(data = check_pf)
+report <-
+  bdc_create_report(data = check_pf,
+                    database_id = "database_id",
+                    workflow_step = "prefilter",
+                    save_report = FALSE)
+report
+```
+
+```{r}
+# Figures
+figures <-
+bdc_create_figures(data = check_pf,
+                   database_id = "database_id",
+                   workflow_step = "prefilter",
+                   save_figures = FALSE)
+
+# Check figures using
+figures$.coordinates_country_inconsistent
+
+```
+
+```{r}
+# Filter out flags
+database <-
+  check_pf %>%
+  dplyr::filter(.summary == TRUE) %>%
+  bdc_filter_out_flags(data = ., col_to_remove = "all")
+```
+
+# Flagging common time issues
+```{r}
+check_time <-
+  bdc_eventDate_empty(data = database, eventDate = "verbatimEventDate")
+```
+
+```{r}
+check_time <-
+  bdc_year_from_eventDate(data = check_time, eventDate = "verbatimEventDate")
+```
+
+```{r}
+check_time <-
+  bdc_year_outOfRange(data = check_time,
+                      eventDate = "year",
+                      year_threshold = 1800)
+```
+
+```{r}
+check_time <- bdc_summary_col(data = check_time)
+```
+
+```{r}
+report <-
+  bdc_create_report(data = check_time,
+                    database_id = "database_id",
+                    workflow_step = "time",
+                    save_report = FALSE)
+
+report
+```
+
+```{r}
+figures <-
+bdc_create_figures(data = check_time,
+                   database_id = "database_id",
+                   workflow_step = "time",
+                   save_figures = FALSE)
+
+# Check figures using
+figures$time_year_BAR
+```
+
+```{r}
+database <-
+  check_time %>%
+  dplyr::filter(.summary == TRUE) %>%
+  bdc_filter_out_flags(data = ., col_to_remove = "all")
+```
+
+# Flagging common spatial issues
+```{r}
+check_space <-
+  bdc_coordinates_precision(
+    data = database,
+    lon = "decimalLongitude",
+    lat = "decimalLatitude",
+    ndec = c(0, 1) # number of decimals to be tested
+  )
+```
+
+```{r}
+check_space <-
+  CoordinateCleaner::clean_coordinates(
+    x =  check_space,
+    lon = "decimalLongitude",
+    lat = "decimalLatitude",
+    species = "scientificName",
+    countries = ,
+    tests = c(
+      "capitals",     # records within 2km around country and province centroids
+      "centroids",    # records within 1km of capitals centroids
+      "duplicates",   # duplicated records
+      "equal",        # records with equal coordinates
+      "gbif",         # records within 1 degree (~111km) of GBIF headsquare
+      "institutions", # records within 100m of zoo and herbaria
+      "outliers",     # outliers
+      "zeros",        # records with coordinates 0,0
+      "urban"         # records within urban areas
+    ),
+    capitals_rad = 2000,
+    centroids_rad = 1000,
+    centroids_detail = "both", # test both country and province centroids
+    inst_rad = 100, # remove zoo and herbaria within 100m
+    outliers_method = "quantile",
+    outliers_mtp = 5,
+    outliers_td = 1000,
+    outliers_size = 10,
+    range_rad = 0,
+    zeros_rad = 0.5,
+    capitals_ref = NULL,
+    centroids_ref = NULL,
+    country_ref = NULL,
+    country_refcol = "countryCode",
+    inst_ref = NULL,
+    range_ref = NULL,
+    # seas_ref = continent_border,
+    # seas_scale = 110,
+    urban_ref = NULL,
+    value = "spatialvalid" # result of tests are appended in separate columns
+  )
+```
+
+```{r}
+check_space <- bdc_summary_col(data = check_space)
+
+check_space %>%
+  dplyr::filter(.summary == FALSE) %>% # map only records flagged as FALSE
+  bdc_quickmap(
+    data = .,
+    lon = "decimalLongitude",
+    lat = "decimalLatitude",
+    col_to_map = ".summary",
+    size = 0.9
+  )
+```
+
+```{r}
+report <-
+  bdc_create_report(data = check_space,
+                    database_id = "database_id",
+                    workflow_step = "space",
+                    save_report = FALSE)
+
+report
+```
+
+```{r}
+figures <-
+bdc_create_figures(data = check_space,
+                   database_id = "database_id",
+                   workflow_step = "space",
+                   save_figures = TRUE)
+
+# Check figures using
+figures$.rou
+```
+
+```{r}
+database <-
+  check_space %>%
+  dplyr::filter(.summary == TRUE) %>%
+  bdc_filter_out_flags(data = ., col_to_remove = "all")
+```
+
+# Clean and parse species names
+```{r}
+parse_names <-
+  bdc_clean_names(sci_names = database$scientificName, save_outputs = FALSE)
+```
+
+```{r}
+parse_names <-
+  parse_names %>%
+  dplyr::select(.uncer_terms, names_clean)
+
+database <- dplyr::bind_cols(database, parse_names)
+```
+
+```{r}
+query_names <- bdc_query_names_taxadb(
+  sci_name            = database$names_clean,
+  replace_synonyms    = TRUE, # replace synonyms by accepted names?
+  suggest_names       = TRUE, # try to found a candidate name for misspelled names?
+  suggestion_distance = 0.9, # distance between the searched and suggested names
+  db                  = "gbif", # taxonomic database
+  parallel            = FALSE, # should parallel processing be used?
+  ncores              = 2, # number of cores to be used in the parallelization process
+  export_accepted     = FALSE # save names linked to multiple accepted names
+)
+```
+
+```{r}
+database <-
+  database %>%
+  dplyr::rename(verbatim_scientificName = scientificName) %>%
+  dplyr::select(-names_clean) %>%
+  dplyr::bind_cols(., query_names)
+```
+
+```{r}
+report <-
+  bdc_create_report(data = database,
+                    database_id = "database_id",
+                    workflow_step = "taxonomy",
+                    save_report = FALSE)
+
+report
+```
+
+```{r}
+unresolved_names <-
+  bdc_filter_out_names(data = database,
+                       col_name = "notes",
+                       taxonomic_status = "accepted",
+                       opposite = TRUE)
+```
+
+```{r}
+output <-
+   bdc_filter_out_names(
+     data = database,
+     taxonomic_notes = "accepted",
+     opposite = FALSE
+   )
 ```
@@
 # With cell values to be the mean values calculated for StdValue
-num_traits_georef_wider <- rtry_trans_wider(num_traits_georef, names_from = c(TraitID, TraitName, UnitName), values_from = c(StdValue), values_fn = list(StdValue = mean))
+database <- rtry_trans_wider(num_traits_georef, names_from = c(TraitID, TraitName, UnitName), values_from = c(StdValue), values_fn = list(StdValue = mean))
 ```
@@
 # Export the data into a CSV file
-ecuador_TRY_plant_traits_path <- file.path(output_path,"Ecuador_TRY_plant_traits.csv")
-write.csv(num_traits_georef_wider, file = ecuador_TRY_plant_traits_path)
+ecuador_TRY_plant_traits_path <- file.path(output_path,"Ecuador_TRY_plant_traits_cleaned.csv")
+write.csv(#NEED#, file = ecuador_TRY_plant_traits_path)
 ```

### 2023-01-12 code/L0/L0_Ecuador_plants_BIEN.Rmd
Add bdvis package for summary

@@
 library(summarytools)
+library(bdvis)
 ```
@@
 plot_intro(Ecuador_BIEN_occ)
-#plot_missing(Ecuador_BIEN_occ)
-#plot_bar(Ecuador_BIEN_occ)
 ```
@@
 ```
+```{r}
+#using bdvis package
+Ecuador_BIEN_occ <- format_bdvis(Ecuador_BIEN_occ, Latitude = "latitude", Longitude = "longitude", Date_collected = "date_collected", Scientific_name = "verbatim_scientific_name")
+bdsummary(Ecuador_BIEN_occ)
+```
+```{r}
+bdcalendarheat(Ecuador_BIEN_occ)
+bdcomplete(Ecuador_BIEN_occ)
+distrigraph(Ecuador_BIEN_occ, ptype = "species")
+distrigraph(Ecuador_BIEN_occ, ptype = "efforts")
+distrigraph(Ecuador_BIEN_occ, ptype = "cell")
+distrigraph(Ecuador_BIEN_occ, ptype = "effortspecies")
+mapgrid(Ecuador_BIEN_occ, ptype = "species", region = "Ecuador")
+```
 

### 2023-01-12 code/L0/L0_Ecuador_plants_GBIF.Rmd
Add bdvis package for summary

--- 2022-10-19 code/L0/L0_Ecuador_plants_GBIF.Rmd
@@
 # using summarytools package
-dfSummary(d, 
-          plain.ascii  = FALSE, 
-          style        = "grid", 
-          graph.magnif = 0.75, 
-          valid.col    = FALSE,
-          tmp.img.dir  = "/tmp")
+d_summary <- dfSummary(d, 
+            na.col       = FALSE,
+            style        = "multiline",
+            plain.ascii  = TRUE,
+            graph.magnif = .8)
+print(d_summary, method = "render")
+view(d_summary, file = file.path(output_path,"Ecuador_BIEN_traits_summary.html"))
+```
+
+```{r}
+#using bdvis package
+d <- format_bdvis(d, , Latitude = "Latitude", Longitude = "Longitude", Date_collected = "Date_collected", Scientific_name = "scientificName")
+bdsummary(d)
+d$Date_collected <- as.Date(d$Date_collected)
+bdcalendarheat(d)
+chronohorogram(d)
+getcellid(d, gridscale = 10)
+distrigraph(d, ptype = "species")
+distrigraph(d, ptype = "efforts")
+distrigraph(d, ptype = "cell")
+distrigraph(d, ptype = "effortspecies")
+tempolar(d, timescale = "m")
+bdcomplete(d)
+mapgrid(d, ptype = "species", region = "Ecuador")
 ```

### 2023-01-12 code/L0/L0_Ecuador_plants_TRY.Rmd
Add output information to header

@@
 data input: "None"
-data output: "TBD"
+data output: "Ecuador_TRY_plant_traits.csv"
 project: "Frugivoria"
@@
 print(ecuador_TRY_plant_traits_summary, method = "render")
-view(ecuador_TRY_plant_traits_summary, file = file.path(output_path,"ecuador_TRY_plant_traits_summary.html"))
+view(ecuador_TRY_plant_traits_summary, file = file.path(output_path,"Ecuador_TRY_summary.html"))
 # gives the following error: Error in withCallingHandlers(expr, warning = function(w) if (inherits(w, : 

### 2023-01-12 code/L0/L0_Ecuador_plants_GIFT.Rmd
Inital commit with file paths, load packages and header information

---
title: "Ecuador plant species GIFT data"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "TBD"
project: "Frugivoria"
date: "2023-01-12"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
remotes::install_github("https://github.com/BioGeoMacro/GIFT",
dependencies = TRUE)
library(GIFT)
library(mde)
library(DataExplorer)
library(summarytools)
```

# Get data for Ecuador from GIFT database

### 2023-01-22 code/L0/L0_Ecuador_plants_TRY.Rmd
add bdvis package & corresponding code convert TRY data to wide table format

--- 2022-10-31 code/L0/L0_Ecuador_plants_TRY.Rmd
@@
 library(rtry)
+library(mde)
+library(DataExplorer)
+library(summarytools)
+library(bdvis)
 ```
@@
 ```{r}
-data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/TRY_Ecuador_PublicData_10192020.txt")
+data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/TRY_Ecuador_PublicData_10192020.txt", encoding = "Latin-1")
+data <- as.data.frame(data)
 ```
+
 # Explore TRY data
@@
 ```
+
+# Convert to wide table
+From https://cran.r-project.org/web/packages/rtry/vignettes/rtry-workflow-general.html
+```{r}
+num_traits <- rtry_select_row(data, complete.cases(TraitID) & complete.cases(StdValue), getAncillary = TRUE, rmDuplicates = TRUE)
+num_traits <- rtry_select_col(num_traits, ObservationID, AccSpeciesID, AccSpeciesName, TraitID, TraitName, StdValue, UnitName)
+
+# Extract the unique value of latitude (DataID 59) and the corresponding ObservationID
+workdata_lat <- rtry_select_anc(data, 59)
+
+# Extract the unique value of longitude (DataID 60) and the corresponding ObservationID
+workdata_lon <- rtry_select_anc(data, 60)
+
+# To merge the extracted ancillary data with the numerical traits
+# Merge the relevant data frames based on the ObservationID using rtry_join_left (left join)
+num_traits_georef <- rtry_join_left(num_traits, workdata_lat, baseOn = ObservationID)
+num_traits_georef <- rtry_join_left(num_traits_georef, workdata_lon, baseOn = ObservationID)
+
+TRY_data <- rtry_trans_wider(num_traits_georef, names_from = c(TraitID, TraitName, UnitName), values_from = c(StdValue), values_fn = list(StdValue = mean))
+```
+
+# Summary of data
+```{r}
+summary(TRY_data)
+```
+
+### using DataExplorer package
+```{r}
+plot_str(TRY_data)
+introduce(TRY_data)
+plot_intro(TRY_data)
+plot_missing(TRY_data)
+plot_bar(TRY_data)
+```
+
+### using summarytools package
+```{r}
+ecuador_TRY_plant_traits_summary <- dfSummary(TRY_data, 
+            na.col       = FALSE,
+            style        = "multiline",
+            plain.ascii  = TRUE,
+            graph.magnif = .8)
+print(ecuador_TRY_plant_traits_summary, method = "render")
+view(ecuador_TRY_plant_traits_summary, file = file.path(output_path,"Ecuador_TRY_summary.html"))
+# gives the following error: Error in withCallingHandlers(expr, warning = function(w) if (inherits(w, : 
+#   invalid multibyte string at '<b0>45''
+```
+
+```{r}
+#using bdvis package
+d <- format_bdvis(d, , Latitude = "Latitude", Longitude = "Longitude", Date_collected = "Date_collected", Scientific_name = "AccSpeciesName")
+bdsummary(d)
+d$Date_collected <- as.Date(d$Date_collected)
+bdcalendarheat(d)
+getcellid(d, gridscale = 10)
+distrigraph(d, ptype = "species")
+distrigraph(d, ptype = "efforts")
+distrigraph(d, ptype = "cell")
+distrigraph(d, ptype = "effortspecies")
+bdcomplete(d)
+mapgrid(d, ptype = "species", region = "Ecuador")
+```
+
+# Write data to csv
+```{r}
+ecuador_TRY_plant_traits_path <- file.path(output_path,"Ecuador_TRY_plant_traits.csv")
+write.csv(data, file = ecuador_TRY_plant_traits_path)
+
+```

### 2023-01-22 code/L0/L0_Ecuador_plants_GIFT.Rmd
retrive data for Ecuador get checklist get richness and trait coverage = 0

@@
 ```{r}
-remotes::install_github("https://github.com/BioGeoMacro/GIFT",
-dependencies = TRUE)
+#run first time only
+#remotes::install_github("https://github.com/BioGeoMacro/GIFT",dependencies = TRUE)
 library(GIFT)
@@
 # Get data for Ecuador from GIFT database
+```{r}
+# retrive entity_ID for Ecuador
+regions <- GIFT_regions()
+str(regions)
+regions[regions$geo_entity %in% "Ecuador",]
+
+```
+
+```{r}
+Ecuador_shape <- GIFT_shape(entity_ID = 10348)
+```
+
+```{r}
+GIFT_data <- GIFT_checklist(floristic_group = 'all', geo_type = 'All', shp = Ecuador_shape, overlap = 'extent_intersect', taxonomic_group = TRUE)
+```
+
+```{r}
+richness <- GIFT_richness(what = "species_richness")
+trait_coverage <- GIFT_richness(what = "trait_coverage")
+```
+
+```{r}
+richness[richness$geo_entity %in% 10348,]
+trait_coverage[trait_coverage$geo_entity %in% 10348,]
+```
+
+```{r}
+
+```

### 2023-01-22 code/L0/L0_Ecuador_plants_GBIF.Rmd
update download number

@@
 ```{r}
-d <- occ_download_get("0065583-220831081235567", path = output_path) %>%
+d <- occ_download_get("0253091-220831081235567", path = output_path) %>%
   occ_download_import()
@@
 #using bdvis package
-d <- format_bdvis(d, , Latitude = "Latitude", Longitude = "Longitude", Date_collected = "Date_collected", Scientific_name = "scientificName")
+d <- format_bdvis(d, , Latitude = "decimalLatitude", Longitude = "decimalLongitude", Date_collected = "eventDate", Scientific_name = "scientificName", Family = "family", Genus = "genus")
+```
+
+```{r}
 bdsummary(d)
@@
 bdcalendarheat(d)
-chronohorogram(d)
 getcellid(d, gridscale = 10)
@@
 distrigraph(d, ptype = "effortspecies")
-tempolar(d, timescale = "m")
 bdcomplete(d)

### 2023-01-23 code/L0/L0_Ecuador_plants_BIEN.Rmd
knit to html

--- 2022-11-28 code/L0/L0_Ecuador_plants_BIEN.Rmd
@@
 # Get data for Ecuador from BIEN database
-
 ```{r}
@@
 Ecuador_BIEN_traits <- BIEN_trait_country("Ecuador", all.taxonomy = TRUE)
+```
+
+```{r}
+#remove duplicate data_collected column from Ecuador_BIEN_occ
+Ecuador_BIEN_occ <- subset (Ecuador_BIEN_occ, select = -date_collected)
 ```
@@
 plot_intro(Ecuador_BIEN_traits)
-plot_missing(Ecuador_BIEN_traits)
 plot_bar(Ecuador_BIEN_traits)
@@
 
-### using summarytools package
 ```{r}
+# using summarytools package
 Ecuador_BIEN_occ_summary <- dfSummary(Ecuador_BIEN_occ, 
@@
 ```
+```{r}
+#using bdvis package
+Ecuador_BIEN_occ <- format_bdvis(Ecuador_BIEN_occ, Latitude = "latitude", Longitude = "longitude", Date_collected = "date_collected", Scientific_name = "verbatim_scientific_name")
+bdsummary(Ecuador_BIEN_occ)
+```
+```{r}
+## ERROR - Error in !dbPreExists : invalid argument type
+## ERROR - Error in postgresqlNewConnection(drv, ...) :
+# RPosgreSQL error: could not connect postgres@localhost:5432 on dbname "test": could not connect to  server: Connection refused (0x0000274D/10061)
+#bdcalendarheat(Ecuador_BIEN_occ)
+#bdcomplete(Ecuador_BIEN_occ)
+#distrigraph(Ecuador_BIEN_occ, ptype = "species")
+#distrigraph(Ecuador_BIEN_occ, ptype = "efforts")
+#distrigraph(Ecuador_BIEN_occ, ptype = "cell")
+#distrigraph(Ecuador_BIEN_occ, ptype = "effortspecies")
+#mapgrid(Ecuador_BIEN_occ, ptype = "species", region = "Ecuador")
+```
 

### 2023-01-23 code/L0/L0_Ecuador_plants_GBIF.Rmd
knit to html

@@
 # using summarytools package
-d_summary <- dfSummary(d, 
-            na.col       = FALSE,
-            style        = "multiline",
-            plain.ascii  = TRUE,
-            graph.magnif = .8)
-print(d_summary, method = "render")
-view(d_summary, file = file.path(output_path,"Ecuador_BIEN_traits_summary.html"))
+# ERROR -- Error in hist.default(data, breaks = breaks_x, plot = FALSE) :
+# some 'x' not counted; maybe 'breaks' do not span range of 'x'
+#d_summary <- dfSummary(d, 
+#            na.col       = FALSE,
+#            style        = "multiline",
+#            plain.ascii  = TRUE,
+#            graph.magnif = .8)
+#print(d_summary, method = "render")
+#view(d_summary, file = file.path(output_path,"Ecuador_BIEN_traits_summary.html"))
 ```
@@
 #using bdvis package
-d <- format_bdvis(d, , Latitude = "decimalLatitude", Longitude = "decimalLongitude", Date_collected = "eventDate", Scientific_name = "scientificName", Family = "family", Genus = "genus")
+d <- format_bdvis(d, Latitude = "decimalLatitude", Longitude = "decimalLongitude", Date_collected = "eventDate", Scientific_name = "scientificName", Family = "family", Genus = "genus")
+bdsummary(d)
 ```
@@
 ```{r}
-bdsummary(d)
-d$Date_collected <- as.Date(d$Date_collected)
-bdcalendarheat(d)
-getcellid(d, gridscale = 10)
-distrigraph(d, ptype = "species")
-distrigraph(d, ptype = "efforts")
-distrigraph(d, ptype = "cell")
-distrigraph(d, ptype = "effortspecies")
-bdcomplete(d)
-mapgrid(d, ptype = "species", region = "Ecuador")
+## ERROR - Error in !dbPreExists : invalid argument type
+## ERROR - Error in postgresqlNewConnection(drv, ...) :
+# RPosgreSQL error: could not connect postgres@localhost:5432 on dbname "test": could not connect to server: Connection refused (0x0000274D/10061)
+#bdcalendarheat(d)
+#distrigraph(d, ptype = "species")
+#distrigraph(d, ptype = "efforts")
+#distrigraph(d, ptype = "cell")
+#distrigraph(d, ptype = "effortspecies")
+#bdcomplete(d)
+#mapgrid(d, ptype = "species", region = "Ecuador")
 ```

### 2023-01-23 code/L2/L2_Ecuador_plant_occ.Rmd
add metadata file add summary code (same as L0)

@@
 library(bdc)
+library(DataExplorer)
+library(summarytools)
+library(bdvis)
 ```
@@
 metadata <-
-  readr::read_csv(system.file("ADD FILE PATH HERE",
-                              package = "bdc"),
+  readr::read_csv(file.path(data_path, "occ_metadata.csv"),
                   show_col_types = FALSE)
@@
 # Read in BIEN, GBIF, and IUCN occurrence data
-```{r}
-# Path to the folder containing the example datasets. For instance:
-path <- data_path
-
-# Change in the Configuration table the path to the folder in your computer containing the example datasets
-metadata$fileName <-
-  gsub(pattern = "https://raw.githubusercontent.com/brunobrr/bdc/master/inst/extdata/input_files/",
-       replacement = path,
-       x = metadata$fileName)
-```
 
@@
                          overwrite = TRUE,
-                         save_database = TRUE)
+                         save_database = FALSE)
 ```
+
+# Summary of data
+```{r}
+summary(database)
+```
+
+### using DataExplorer package
+```{r}
+plot_str(database)
+introduce(database)
+plot_intro(database)
+plot_missing(database)
+```
+
+### using summarytools package
+```{r}
+database_summary <- dfSummary(database, 
+            na.col       = FALSE,
+            style        = "multiline",
+            plain.ascii  = TRUE,
+            graph.magnif = .8)
+print(database_summary, method = "render")
+view(database_summary, file = file.path(output_path,"database_summary.html"))
+```
+
+```{r}
+#using bdvis package
+d <- format_bdvis(d, , Latitude = "Latitude", Longitude = "Longitude", Date_collected = "Date_collected", Scientific_name = "AccSpeciesName")
+bdsummary(d)
+d$Date_collected <- as.Date(d$Date_collected)
+bdcalendarheat(d)
+getcellid(d, gridscale = 10)
+distrigraph(d, ptype = "species")
+distrigraph(d, ptype = "efforts")
+distrigraph(d, ptype = "cell")
+distrigraph(d, ptype = "effortspecies")
+bdcomplete(d)
+mapgrid(d, ptype = "species", region = "Ecuador")
+```

### 2023-01-23 code/L0/L0_Ecuador_plants_NPP.Rmd
read in data downloaded from Neotropical plant portal add summary information and export to csv

---
title: "Ecuador plant species Neotropical Plant Portal data"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "Ecuador_NPP_plant_occ.csv"
project: "Frugivoria"
date: "2023-1-23"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(DataExplorer)
library(summarytools)
library(bdvis)
```

# Read in Neotropical Plant Portal data
Data was obtained from https://serv.biokic.asu.edu/neotrop/plantae/collections/map/index.php?clid=1&cltype=vouchers&taxonfilter=&db=all&type=1&reset=1
```{r}
npp_data <- read.csv(file.path(data_path,"NeotropicalPlantPortal_SymbOutput_2022-11-28_084151_DwC-A/occurrences.csv"))
```

# Summary of data
```{r}
summary(npp_data)
```

### using DataExplorer package
```{r}
plot_str(npp_data)
introduce(npp_data)
plot_intro(npp_data)
plot_missing(npp_data)
```

### using summarytools package
```{r}
ecuador_NPP_plant_occ_summary <- dfSummary(npp_data, 
            na.col       = FALSE,
            style        = "multiline",
            plain.ascii  = TRUE,
            graph.magnif = .8)
print(ecuador_NPP_plant_occ_summary, method = "render")
view(ecuador_NPP_plant_occ_summary, file = file.path(output_path,"Ecuador_NPP_summary.html"))
```

```{r}
#using bdvis package
d <- format_bdvis(npp_data, , Latitude = "decimalLatitude", Longitude = "decimalLongitude", Date_collected = "eventDate", Scientific_name = "scientificName")
bdsummary(d)

bdcalendarheat(d)
getcellid(d, gridscale = 10)
distrigraph(d, ptype = "species")
distrigraph(d, ptype = "efforts")
distrigraph(d, ptype = "cell")
distrigraph(d, ptype = "effortspecies")
bdcomplete(d)
mapgrid(d, ptype = "species", region = "Ecuador")
```

# Write data to csv
```{r}
ecuador_npp_path <- file.path(output_path,"Ecuador_NPP_plant_occ.csv")
write.csv(npp_data, file = ecuador_npp_path)
```

### 2023-01-23 code/L0/L0_Ecuador_plants_BIEN.Rmd
update html

@@
 library(tidyr)
-library(mde)
+library(dplyr)
 library(DataExplorer)
@@
 library(bdvis)
+library(visdat)
 ```
@@
 # occurrence data
-Ecuador_BIEN_occ <- BIEN_occurrence_country("Ecuador", all.taxonomy = TRUE, observation.type = TRUE, collection.info = TRUE)
+Ecuador_BIEN_occ <- BIEN_occurrence_country("Ecuador", all.taxonomy = TRUE, observation.type = TRUE, collection.info = TRUE, political.boundaries = TRUE, native.status = TRUE)
 # species list
@@
 # trait data
-Ecuador_BIEN_traits <- BIEN_trait_country("Ecuador", all.taxonomy = TRUE)
+Ecuador_BIEN_traits <- BIEN_trait_country("Ecuador", all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
 ```
@@
 summary(Ecuador_BIEN_occ)
-na_summary(Ecuador_BIEN_occ, exclude_cols = c("date_collected", "date_identified"))
 ```
@@
 summary(Ecuador_BIEN_species_list)
-na_summary(Ecuador_BIEN_species_list)
 ```
@@
 summary(Ecuador_BIEN_traits)
-na_summary(Ecuador_BIEN_traits)
+```
+
+```{r}
+glimpse(Ecuador_BIEN_occ)
+```
+
+```{r}
+glimpse(Ecuador_BIEN_species_list)
+```
+
+```{r}
+glimpse(Ecuador_BIEN_traits)
 ```
@@
 plot_intro(Ecuador_BIEN_occ)
+plot_missing(Ecuador_BIEN_occ)
+plot_bar(Ecuador_BIEN_occ)
 ```
@@
 plot_intro(Ecuador_BIEN_traits)
+plot_missing(Ecuador_BIEN_traits)
 plot_bar(Ecuador_BIEN_traits)
@@
 
+```{r}
+#using visdat package
+vis_dat(Ecuador_BIEN_occ, warn_large_data = FALSE)
+vis_miss(Ecuador_BIEN_occ, warn_large_data = FALSE)
+Ecuador_BIEN_occ %>%
+  select_if(is.numeric)%>%
+  vis_cor(warn_large_data = FALSE)
+```
+
+```{r}
+#using visdat package
+vis_dat(Ecuador_BIEN_traits, warn_large_data = FALSE)
+vis_miss(Ecuador_BIEN_traits, warn_large_data = FALSE)
+Ecuador_BIEN_traits %>%
+  select_if(is.numeric)%>%
+  vis_cor(warn_large_data = FALSE)
+```
+
 # Write data to csv

### 2023-02-27 code/L0/L0_Ecuador_plants_GBIF.Rmd
Add visdat package visualizations

@@
 ---
-title: "Ecuador plant species GBIF"
+title: "Northern Andes plant species GBIF"
 author: "Hazel J. Anderson"
@@
 data input: "None"
-data output: "Ecuador_GBIF_occurences.csv"
+data output: "NorthAndes_GBIF_occurences.csv"
 project: "Frugivoria"
-date: "2022-10-10"
+date: "2023-02-13"
 output: html_document
@@
 library(rgbif)
-library(mde)
 library(DataExplorer)
@@
 pred("hasCoordinate", TRUE),
-pred("country","EC"),
-pred("occurrenceStatus","PRESENT"), 
+pred_in("country",c("EC","CO","VE")), 
 pred_not(pred_in("basisOfRecord",c("FOSSIL_SPECIMEN","LIVING_SPECIMEN"))),
@@
 ```{r}
-d <- occ_download_get("0253091-220831081235567", path = output_path) %>%
+d <- occ_download_get("0282867-220831081235567", path = output_path) %>%
   occ_download_import()
@@
 ```{r}
-summary(d)
-```
-
-```{r}
-# using mde package
-na_summary(d, exclude_cols = c("gbifID"))
+glimpse(d)
 ```
@@
 #using bdvis package
-d <- format_bdvis(d, Latitude = "decimalLatitude", Longitude = "decimalLongitude", Date_collected = "eventDate", Scientific_name = "scientificName", Family = "family", Genus = "genus")
-bdsummary(d)
+d_vis <- format_bdvis(d, Latitude = "decimalLatitude", Longitude = "decimalLongitude", Date_collected = "eventDate", Scientific_name = "scientificName", Family = "family", Genus = "genus")
+bdsummary(d_vis)
 ```
@@
 # RPosgreSQL error: could not connect postgres@localhost:5432 on dbname "test": could not connect to server: Connection refused (0x0000274D/10061)
-#bdcalendarheat(d)
-#distrigraph(d, ptype = "species")
-#distrigraph(d, ptype = "efforts")
-#distrigraph(d, ptype = "cell")
-#distrigraph(d, ptype = "effortspecies")
-#bdcomplete(d)
-#mapgrid(d, ptype = "species", region = "Ecuador")
+#bdcalendarheat(d_vis)
+#distrigraph(d_vis, ptype = "species")
+#distrigraph(d_vis, ptype = "efforts")
+#distrigraph(d_vis, ptype = "cell")
+#distrigraph(d_vis, ptype = "effortspecies")
+#bdcomplete(d_vis)
+```
+
+```{r}
+#using visdat package
+vis_dat(d, warn_large_data = FALSE)
+vis_miss(d, warn_large_data = FALSE)
+d %>%
+  select_if(is.numeric)%>%
+  vis_cor(warn_large_data = FALSE)
 ```
@@
 ```{r}
-Ecuador_GBIF_path <- file.path(output_path,"Ecuador_GBIF_occurences.csv") 
-write.csv(d, file = Ecuador_GBIF_path)
+NorthAndes_GBIF_path <- file.path(output_path,"NorthAndes_GBIF_occurences.csv") 
+write.csv(d, file = NorthAndes_GBIF_path)
 ```

### 2023-02-27 code/L0/L0_Ecuador_plants_IUCN.Rmd
Update for Northern Andes

--- 2022-10-31 code/L0/L0_Ecuador_plants_IUCN.Rmd
@@
 ---
-title: "Ecuador plant species IUCN data"
+title: "Northern Andes plant species IUCN data"
 author: "Hazel J. Anderson"
@@
 data input: "None"
-data output: "Ecuador_IUCN_plant_species.csv, Ecuador_IUCN_plant_occ.csv, Ecuador_IUCN_plant_habitat.csv"
+data output: "NorthAndes_IUCN_plant_species.csv, NorthAndes_IUCN_plant_occ.csv, NorthAndes_IUCN_plant_habitat.csv"
 project: "Frugivoria"
-date: "2022-08-31"
+date: "2023-02-27"
 output: html_document
@@
 
-# Ecuador species list
+# North Andes species list
 ```{r}
@@
 ecuador_species <- ecuador_species$result[1]
+#Colombia
+colombia_species <- rl_sp_country("CO", key = token, parse = TRUE)
+colombia_species <- colombia_species$result[1]
+#Venezuela
+venezuela_species <- rl_sp_country("VE", key = token, parse = TRUE)
+venezuela_species <- venezuela_species$result[1]
+
+#combine all 3
+northAndes_species <- rbind(ecuador_species, colombia_species, venezuela_species)
 ```
@@
 
-# Generate Ecuador plant species list
+# Generate Northern Andes plant species list
 ```{r}
-ecuador_plants <- merge(all_plants, ecuador_species, by="taxonid")
+northAndes_plants <- merge(all_plants, northAndes_species, by="taxonid")
 ```
 
-# Get occurences for Ecuador using species list
+# Get occurences for Northern Andes using species list
 ```{r}
@@
 }
-ecuador_IUCN_plant_occ <- ecuador_plants %>%
+northAndes_IUCN_plant_occ <- northAndes_plants %>%
   rowwise %>%
   do(get_occ(.$taxonid))
-ecuador_IUCN_plant_occ <- as.data.frame(ecuador_IUCN_plant_occ)
+northAndes_IUCN_plant_occ <- as.data.frame(northAndes_IUCN_plant_occ)
 ```
@@
 }
-ecuador_IUCN_plant_habitat <- ecuador_plants %>%
+northAndes_IUCN_plant_habitat <- northAndes_plants %>%
   rowwise %>%
   do(get_habitat(.$taxonid))
-ecuador_IUCN_plant_habitat <- as.data.frame(ecuador_IUCN_plant_habitat)
+northAndes_IUCN_plant_habitat <- as.data.frame(northAndes_IUCN_plant_habitat)
 ```
@@
 ```{r}
-summary(ecuador_plants)
-na_summary(ecuador_plants, exclude_cols = c("subpopulation"))
+summary(northAndes_plants)
+na_summary(northAndes_plants, exclude_cols = c("subpopulation"))
 ```
@@
 ```{r}
-summary(ecuador_IUCN_plant_occ)
-na_summary(ecuador_IUCN_plant_occ)
+summary(northAndes_IUCN_plant_occ)
+na_summary(northAndes_IUCN_plant_occ)
 ```
@@
 ```{r}
-summary(ecuador_IUCN_plant_habitat)
-na_summary(ecuador_IUCN_plant_habitat)
+summary(northAndes_IUCN_plant_habitat)
+na_summary(northAndes_IUCN_plant_habitat)
+```
+
+### using DataExplorer package
+```{r}
+plot_str(northAndes_plants)
+introduce(northAndes_plants)
+plot_intro(northAndes_plants)
+plot_missing(northAndes_plants)
+plot_bar(northAndes_plants)
+```
+
+```{r}
+plot_str(northAndes_IUCN_plant_occ)
+introduce(northAndes_IUCN_plant_occ)
+plot_intro(northAndes_IUCN_plant_occ)
+plot_missing(northAndes_IUCN_plant_occ)
+plot_bar(northAndes_IUCN_plant_occ)
+```
+
+```{r}
+plot_str(northAndes_IUCN_plant_habitat)
+introduce(northAndes_IUCN_plant_habitat)
+plot_intro(northAndes_IUCN_plant_habitat)
+plot_missing(northAndes_IUCN_plant_habitat)
+plot_bar(northAndes_IUCN_plant_habitat)
+```
+
+### using summarytools package
+```{r}
+northAndes_plants_summary <- dfSummary(northAndes_plants, 
+            na.col       = FALSE,
+            style        = "multiline",
+            plain.ascii  = TRUE,
+            graph.magnif = .8)
+print(northAndes_plants_summary, method = "render")
+view(northAndes_plants_summary, file = file.path(output_path,"northAndes_plants_summary.html"))
+```
+
+```{r}
+northAndes_IUCN_plant_occ_summary <- dfSummary(northAndes_IUCN_plant_occ, 
+            na.col       = FALSE,
+            style        = "multiline",
+            plain.ascii  = TRUE,
+            graph.magnif = .8)
+print(northAndes_IUCN_plant_occ_summary, method = "render")
+view(northAndes_IUCN_plant_occ_summary, file = file.path(output_path,"northAndes_IUCN_plant_occ_summary.html"))
+```
+
+```{r}
+northAndes_IUCN_plant_habitat_summary <- dfSummary(northAndes_IUCN_plant_habitat, 
+            na.col       = FALSE,
+            style        = "multiline",
+            plain.ascii  = TRUE,
+            graph.magnif = .8)
+print(northAndes_IUCN_plant_habitat_summary, method = "render")
+view(northAndes_IUCN_plant_habitat_summary, file = file.path(output_path,"northAndes_IUCN_plant_habitat_summary.html"))
 ```
@@
 ```{r}
-ecuador_ICUN_species_list_path <- file.path(output_path,"Ecuador_IUCN_plant_species.csv")
-write.csv(ecuador_plants, file = ecuador_ICUN_species_list_path)
+northAndes_ICUN_species_list_path <- file.path(output_path,"northAndes_IUCN_plant_species.csv")
+write.csv(northAndes_plants, file = northAndes_ICUN_species_list_path)
 
-ecuador_IUCN_plant_occ_path <- file.path(output_path, "Ecuador_IUCN_plant_occ.csv")
-write.csv(ecuador_IUCN_plant_occ, file = ecuador_IUCN_plant_occ_path)
+northAndes_IUCN_plant_occ_path <- file.path(output_path, "northAndes_IUCN_plant_occ.csv")
+write.csv(northAndes_IUCN_plant_occ, file = northAndes_IUCN_plant_occ_path)
 
-ecuador_IUCN_plant_habitat_path <- file.path(output_path, "Ecuador_IUCN_plant_habitat.csv")
-write.csv(ecuador_IUCN_plant_habitat, file = ecuador_IUCN_plant_habitat_path)
+northAndes_IUCN_plant_habitat_path <- file.path(output_path, "northAndes_IUCN_plant_habitat.csv")
+write.csv(northAndes_IUCN_plant_habitat, file = northAndes_IUCN_plant_habitat_path)
 ```

### 2023-02-27 code/L2/L2_Ecuador_plant_occ.Rmd
Add code acknowledgments

@@
 # Read in the configuration table
+Using bdc package; modified from https://brunobrr.github.io/bdc/articles/integrate_datasets.html
 ```{r}
@@
 # Read in BIEN, GBIF, and IUCN occurrence data
-
+Using bdc package; modified from https://brunobrr.github.io/bdc/articles/integrate_datasets.html
 ```{r}

### 2023-02-27 code/L0/L0_Ecuador_plants_POWO.Rmd
add summary code

@@
 library(dplyr)
+library(DataExplorer)
+library(summarytools)
 ```
@@
 ```{r}
-Ecuador_POWO_page1 <- search_powo(list(distribution="Ecuador"), limit = 5000)
+Ecuador_POWO_page1 <- search_powo(list(distribution="Ecuador"), limit = 4000)
 Ecuador_POWO_page2 <- request_next(Ecuador_POWO_page1)
@@
 Ecuador_POWO_page5 <- request_next(Ecuador_POWO_page4)
+Ecuador_POWO_page6 <- request_next(Ecuador_POWO_page5)
 ```
@@
   tidy(Ecuador_POWO_page4),
-  tidy(Ecuador_POWO_page5)
+  tidy(Ecuador_POWO_page5),
+  tidy(Ecuador_POWO_page6)
 )
@@
 na_summary(Ecuador_POWO, exclude_cols = c("accepted", "url", "images"))
+```
+
+```{r}
+# using DataExplorer package
+plot_str(Ecuador_POWO)
+introduce(Ecuador_POWO)
+plot_intro(Ecuador_POWO)
+plot_missing(Ecuador_POWO)
+plot_bar(Ecuador_POWO)
+```
+
+```{r}
+# using summarytools package
+Ecuador_POWO_summary <- dfSummary(Ecuador_POWO, 
+            na.col       = FALSE,
+            style        = "multiline",
+            plain.ascii  = TRUE,
+            graph.magnif = .8)
+print(Ecuador_POWO_summary, method = "render")
+view(Ecuador_POWO_summary, file = file.path(output_path,"Ecuador_POWO_summary.html"))
 ```

### 2023-03-02 code/L0/L0_Ecuador_plants_IUCN.Rmd
switch to reading in IUCN data downloaded from website that includes all plants

@@
 data input: "None"
-data output: "NorthAndes_IUCN_plant_species.csv, NorthAndes_IUCN_plant_occ.csv, NorthAndes_IUCN_plant_habitat.csv"
+data output: "NorthAndes_IUCN_plant_species.csv, NorthAndes_IUCN_plant_occ.csv"
 project: "Frugivoria"
@@
 ```{r}
-library(httr)
-library(jsonlite)
-library(rredlist)
 library(dplyr)
@@
 
-# Set up token and api url
+# Read in IUCN data
+Data was obtained from https://www.iucnredlist.org/search?dl=true&permalink=bd0e6db7-6fb0-4ad3-ad6c-63da82cc0999
 ```{r}
-## Token for accessing IUCN data through their API
-## User must obtain a token (https://apiv3.iucnredlist.org/api/v3/token) before using the 'rredlist package'.
-token <- 'f632af67ee5eef8446a8cb0ca394e1f1cb369ca5eab626edc4ea46e5698deb25'
-api_url <- 'https://apiv3.iucnredlist.org/api/v3/'
-
+assessments <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/assessments.csv"))
+common_names <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/common_names.csv"))
+conservation_needed <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/conservation_needed.csv"))
+countries <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/countries.csv"))
+credits <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/credits.csv"))
+dois <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/dois.csv"))
+fao <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/fao.csv"))
+habitats <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/habitats.csv"))
+plant_specific <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/plant_specific.csv"))
+references <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/references.csv"))
+research_needed <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/research_needed.csv"))
+simple_summary <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/simple_summary.csv"))
+synonyms <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/synonyms.csv"))
+taxonomy <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/taxonomy.csv"))
+threats <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/threats.csv"))
+usetrade <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/usetrade.csv"))
+points_data <- read.csv(file.path(data_path,"IUCN/Northern Andes 3 - Range data Points/points_data.csv"))
 ```
 
-# North Andes species list
+the different tables joined using the "internalTaxonId" when viewing taxon information, or the "assessmentID" field to see assessment information, especially in cases where downloads include multiple assesments for a taxon (e.g.
+regional and global assessments).
 ```{r}
-#Ecuador
-ecuador_species <- rl_sp_country("EC", key = token, parse = TRUE)
-ecuador_species <- ecuador_species$result[1]
-#Colombia
-colombia_species <- rl_sp_country("CO", key = token, parse = TRUE)
-colombia_species <- colombia_species$result[1]
-#Venezuela
-venezuela_species <- rl_sp_country("VE", key = token, parse = TRUE)
-venezuela_species <- venezuela_species$result[1]
 
-#combine all 3
-northAndes_species <- rbind(ecuador_species, colombia_species, venezuela_species)
-```
-
-# Plant species list
-```{r}
-# Get codes for the different species groups
-grp_codes <- GET(url = paste0(api_url, 'comp-group/list?token=', token))
-code <-fromJSON(content(grp_codes, as = 'text'), simplifyDataFrame = TRUE, flatten = TRUE)$result
-```
-
-```{r}
-# Get species list of all plants across all countries
-mangrove_plants <- rl_comp_groups(group = 'mangrove_plants', key = token)
-p1 <- as.data.frame(mangrove_plants$result)
-seagrasses<- rl_comp_groups(group = 'seagrasses', key = token)
-p2 <- as.data.frame(seagrasses$result)
-cycads <- rl_comp_groups(group = 'cycads', key = token)
-p3 <- as.data.frame(cycads$result)
-magnolias <- rl_comp_groups(group = 'magnolias', key = token)
-p4 <- as.data.frame(magnolias$result)
-confiers <- rl_comp_groups(group = 'conifers', key = token)
-p5 <- as.data.frame(confiers$result)
-cacti <- rl_comp_groups(group = 'cacti', key = token)
-p6 <- as.data.frame(cacti$result)
-
-all_plants <- rbind(p1, p2, p3, p4, p5, p6)
-```
-
-# Generate Northern Andes plant species list
-```{r}
-northAndes_plants <- merge(all_plants, northAndes_species, by="taxonid")
-```
-
-# Get occurences for Northern Andes using species list
-```{r}
-get_occ <- function(ID) {
-  rl_occ_country<- rl_occ_country(id=ID,key=token)$result
-  if (class(rl_occ_country) == 'data.frame') {
-    data.frame(taxonid=ID, rl_occ_country)
-  } else {
-    data.frame(taxonid=ID)
-  }
-}
-northAndes_IUCN_plant_occ <- northAndes_plants %>%
-  rowwise %>%
-  do(get_occ(.$taxonid))
-northAndes_IUCN_plant_occ <- as.data.frame(northAndes_IUCN_plant_occ)
-```
-
-# Get habitat for species list
-```{r}
-get_habitat <- function(ID) {
-  habitat_by_sp<- rl_habitats(id=ID,key=token)$result
-  if (class(habitat_by_sp) == 'data.frame') {
-    data.frame(taxonid=ID, habitat_by_sp)
-  } else {
-    data.frame(taxonid=ID)
-  }
-}
-northAndes_IUCN_plant_habitat <- northAndes_plants %>%
-  rowwise %>%
-  do(get_habitat(.$taxonid))
-northAndes_IUCN_plant_habitat <- as.data.frame(northAndes_IUCN_plant_habitat)
 ```

### 2023-03-23 code/L0/L0_Ecuador_plants_IUCN.Rmd
updated with new IUCN link

@@
 # Read in IUCN data
-Data was obtained from https://www.iucnredlist.org/search?dl=true&permalink=bd0e6db7-6fb0-4ad3-ad6c-63da82cc0999
+Data was obtained from https://www.iucnredlist.org/search?dl=true&permalink=a44f4942-2b56-47b9-8122-9896ac4af647
 ```{r}

### 2023-03-23 code/L1/L1_NorthAndes_plants_IUCN.Rmd
rename from Ecuador to Northern Andes

---
title: "Ecuador plant species IUCN Cleaning"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "Ecuador_IUCN_plant_occ.csv, Ecuador_IUCN_plant_species.csv"
data output: "TBD"
project: "Frugivoria"
date: "2022-11-14"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(WorldFlora)
library(taxize)
```

# Read in IUCN occurrence data
```{r}
iucn_occ <- read.csv(file.path(data_path,"Ecuador_IUCN_plant_occ.csv"))
iucn_names <- read.csv(file.path(data_path, "Ecuador_IUCN_plant_species.csv"))
iucn_data <- merge(iucn_names, iucn_occ, by = "taxonid")
iucn_data$X.x <- NULL
iucn_data$X.y <- NULL
```

# Read in World Flora Backbone
```{r}
WFO.backbone <- read.delim(file.path(data_path,"WFO_Backbone/classification.txt"))
```

# Prepare IUCN data for analysis with WFO.match
```{r}
iucn_data.prepare <- WFO.prepare(spec.data = iucn_data, spec.full = "scientific_name", verbose = TRUE)
```

# Standarize plant names according to the World Flora Online taxonimic backbone
```{r}
iucn_data.standard <- WFO.match(spec.data = iucn_data.prepare, WFO.data = WFO.backbone, spec.name = "scientific_name", counter = TRUE)
```

# Resolve IUCN data taxonomic names with taxize
```{r}
iucn_data.resolved <- gnr_resolve(iucn_data$scientific_name)
```

# Merge resolved and standardized data
```{r}

```

### 2023-03-23 archive/L0_Ecuador_elevation.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Ecuador Elevation},
  pdfauthor={Hazel J. Anderson},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Ecuador Elevation}
\author{Hazel J. Anderson}
\date{2022-10-21}

\begin{document}
\maketitle

\hypertarget{set-file-paths}{%
\section{Set file paths}\label{set-file-paths}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{data\_path}\OtherTok{\textless{}{-}}\FunctionTok{file.path}\NormalTok{(}\StringTok{\textquotesingle{}G:/Shared drives/SpaCE\_Lab\_FRUGIVORIA/data/plants/L0\textquotesingle{}}\NormalTok{)}
\NormalTok{output\_path}\OtherTok{\textless{}{-}} \FunctionTok{file.path}\NormalTok{(}\StringTok{\textquotesingle{}G:/Shared drives/SpaCE\_Lab\_FRUGIVORIA/data/plants/L0\textquotesingle{}}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{load-required-packages}{%
\section{Load required packages}\label{load-required-packages}}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(geodata)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Loading required package: terra
\end{verbatim}

\begin{verbatim}
## terra 1.6.47
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(elevatr)}
\end{Highlighting}
\end{Shaded}

\hypertarget{download-elevation-data-for-ecuador-using-geodata}{%
\section{Download elevation data for Ecuador using
geodata}\label{download-elevation-data-for-ecuador-using-geodata}}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# get outline of Ecuador}
\NormalTok{ecuador\_outline }\OtherTok{\textless{}{-}} \FunctionTok{gadm}\NormalTok{(}\AttributeTok{country =} \StringTok{"ECU"}\NormalTok{, }\AttributeTok{level=}\DecValTok{0}\NormalTok{, }\AttributeTok{path =}\NormalTok{ data\_path, }\AttributeTok{version=}\StringTok{"latest"}\NormalTok{, }\AttributeTok{resolution=}\DecValTok{1}\NormalTok{)}
\NormalTok{ecuador\_outline}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  class       : SpatVector 
##  geometry    : polygons 
##  dimensions  : 1, 2  (geometries, attributes)
##  extent      : -92.00854, -75.18715, -5.015803, 1.681835  (xmin, xmax, ymin, ymax)
##  coord. ref. : lon/lat WGS 84 (EPSG:4326) 
##  names       : GID_0 COUNTRY
##  type        : <chr>   <chr>
##  values      :   ECU Ecuador
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# get elevation data for Ecuador}
\NormalTok{ecuador\_geodata }\OtherTok{\textless{}{-}} \FunctionTok{elevation\_30s}\NormalTok{(}\AttributeTok{country =} \StringTok{"ECU"}\NormalTok{, }\AttributeTok{path =}\NormalTok{ data\_path)}
\NormalTok{ecuador\_geodata}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## class       : SpatRaster 
## dimensions  : 816, 792, 1  (nrow, ncol, nlyr)
## resolution  : 0.008333333, 0.008333333  (x, y)
## extent      : -81.6, -75, -5.2, 1.6  (xmin, xmax, ymin, ymax)
## coord. ref. : lon/lat WGS 84 (EPSG:4326) 
## source      : ECU_elv_msk.tif 
## name        : ECU_elv_msk 
## min value   :          -6 
## max value   :        6169
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# look at elevation data}
\FunctionTok{plot}\NormalTok{(ecuador\_geodata)}
\FunctionTok{plot}\NormalTok{(ecuador\_outline, }\AttributeTok{add =}\NormalTok{ T)}
\end{Highlighting}
\end{Shaded}

\includegraphics{L0_Ecuador_elevation_files/figure-latex/unnamed-chunk-3-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#save elevation data to file}
\NormalTok{ecuador\_geodata\_elevation\_path }\OtherTok{\textless{}{-}} \FunctionTok{file.path}\NormalTok{(output\_path,}\StringTok{"ecuador\_geodata\_elevation.RData"}\NormalTok{) }
\FunctionTok{save}\NormalTok{(ecuador\_geodata, }\AttributeTok{file =}\NormalTok{ ecuador\_geodata\_elevation\_path)}
\end{Highlighting}
\end{Shaded}

\hypertarget{download-elevation-data-for-ecuador-using-elevatr}{%
\section{Download elevation data for Ecuador using
`elevatr'}\label{download-elevation-data-for-ecuador-using-elevatr}}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# convert ecuador\_outline to spatial dataframe}
\NormalTok{ecuador\_outline.df }\OtherTok{\textless{}{-}}\NormalTok{ sf}\SpecialCharTok{::}\FunctionTok{st\_as\_sf}\NormalTok{(ecuador\_outline)}

\CommentTok{\# get elevation data for Ecuador}
\NormalTok{ecuador\_elevatr }\OtherTok{\textless{}{-}} \FunctionTok{get\_elev\_raster}\NormalTok{(ecuador\_outline.df, }\AttributeTok{z=} \DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Mosaicing & Projecting
\end{verbatim}

\begin{verbatim}
## Note: Elevation units are in meters.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# plot elevation}
\FunctionTok{plot}\NormalTok{(ecuador\_elevatr)}
\FunctionTok{plot}\NormalTok{(ecuador\_outline, }\AttributeTok{add =}\NormalTok{ T)}
\end{Highlighting}
\end{Shaded}

\includegraphics{L0_Ecuador_elevation_files/figure-latex/unnamed-chunk-4-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#save elevation data to file}
\NormalTok{ecuador\_elevatr\_elevation\_path }\OtherTok{\textless{}{-}} \FunctionTok{file.path}\NormalTok{(output\_path,}\StringTok{"ecuador\_elevatr\_elevation.RData"}\NormalTok{) }
\FunctionTok{save}\NormalTok{(ecuador\_elevatr, }\AttributeTok{file =}\NormalTok{ ecuador\_elevatr\_elevation\_path)}
\end{Highlighting}
\end{Shaded}

\end{document}

### 2023-03-23 archive/L0_Ecuador_plants_POWO.Rmd

---
title: "Northern Andes plant species POWO"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "NorthAndes_POWO_species_list.csv"
project: "Frugivoria"
date: "2023-02-27"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
devtools::install_github("barnabywalker/kewr")
library(kewr)
library(dplyr)
library(DataExplorer)
library(summarytools)
library(visdat)
```

# Search Plants of the World Online [https://powo.science.kew.org/] for species in Northern Andes
Best to break up request into smaller chunks so R handle it without timing out
## Ecuador
```{r}
Ecuador_POWO_page1 <- search_powo(list(distribution="Ecuador"), limit = 4000)
Ecuador_POWO_page2 <- request_next(Ecuador_POWO_page1)
Ecuador_POWO_page3 <- request_next(Ecuador_POWO_page2)
Ecuador_POWO_page4 <- request_next(Ecuador_POWO_page3)
Ecuador_POWO_page5 <- request_next(Ecuador_POWO_page4)
Ecuador_POWO_page6 <- request_next(Ecuador_POWO_page5)
```

```{r}
# Combine pages
Ecuador_POWO <- bind_rows(
  tidy(Ecuador_POWO_page1),
  tidy(Ecuador_POWO_page2),
  tidy(Ecuador_POWO_page3),
  tidy(Ecuador_POWO_page4),
  tidy(Ecuador_POWO_page5),
  tidy(Ecuador_POWO_page6)
)
```

## Colombia
```{r}
Colombia_POWO_page1 <- search_powo(list(distribution="Colombia"), limit = 4000)
Colombia_POWO_page2 <- request_next(Colombia_POWO_page1)
Colombia_POWO_page3 <- request_next(Colombia_POWO_page2)
Colombia_POWO_page4 <- request_next(Colombia_POWO_page3)
Colombia_POWO_page5 <- request_next(Colombia_POWO_page4)
Colombia_POWO_page6 <- request_next(Colombia_POWO_page5)
Colombia_POWO_page7 <- request_next(Colombia_POWO_page6)
Colombia_POWO_page8 <- request_next(Colombia_POWO_page7)
```

```{r}
# Combine pages
Colombia_POWO <- bind_rows(
  tidy(Colombia_POWO_page1),
  tidy(Colombia_POWO_page2),
  tidy(Colombia_POWO_page3),
  tidy(Colombia_POWO_page4),
  tidy(Colombia_POWO_page5),
  tidy(Colombia_POWO_page6)
)
```

## Venezuela
```{r}
Venezuela_POWO_page1 <- search_powo(list(distribution="Venezuela"), limit = 4000)
Venezuela_POWO_page2 <- request_next(Venezuela_POWO_page1)
Venezuela_POWO_page3 <- request_next(Venezuela_POWO_page2)
Venezuela_POWO_page4 <- request_next(Venezuela_POWO_page3)
Venezuela_POWO_page5 <- request_next(Venezuela_POWO_page4)
Venezuela_POWO_page6 <- request_next(Venezuela_POWO_page5)
```

```{r}
# Combine pages
Venezuela_POWO <- bind_rows(
  tidy(Venezuela_POWO_page1),
  tidy(Venezuela_POWO_page2),
  tidy(Venezuela_POWO_page3),
  tidy(Venezuela_POWO_page4),
  tidy(Venezuela_POWO_page5),
  tidy(Venezuela_POWO_page6)
)
```

## Combine Ecuador, Colombia, and Venezuela
```{r}
NorthAndes_POWO <- rbind(Ecuador_POWO,Colombia_POWO, Venezuela_POWO)
```

# Summary
```{r}
summary(NorthAndes_POWO)
```

```{r}
# using DataExplorer package
plot_str(NorthAndes_POWO)
introduce(NorthAndes_POWO)
plot_intro(NorthAndes_POWO)
plot_missing(NorthAndes_POWO)
plot_bar(NorthAndes_POWO)
```

```{r}
# using summarytools package
NorthAndes_POWO_summary <- dfSummary(NorthAndes_POWO, 
            na.col       = FALSE,
            style        = "multiline",
            plain.ascii  = TRUE,
            graph.magnif = .8)
print(NorthAndes_POWO_summary, method = "render")
view(ENorthAndes_POWO_summary, file = file.path(output_path,"Ecuador_POWO_summary.html"))
```

```{r}
#using visdat package
vis_dat(NorthAndes_POWO, warn_large_data = FALSE)
vis_miss(NorthAndes_POWO, warn_large_data = FALSE)
NorthAndes_POWO %>%
  select_if(is.numeric)%>%
  vis_cor(warn_large_data = FALSE)
```

# Write data to csv
```{r}
#convert NorthAndes_POWOO from tbl_df to dataframe
NorthAndes_POWO <- as.data.frame(NorthAndes_POWO)
# remove image column from dataframe
NorthAndes_POWO <- NorthAndes_POWO %>% select(-images)

NorthAndes_POWO_path <- file.path(output_path, "NorthAndes_POWO.csv")
write.csv(NorthAndes_POWO, file = NorthAndes_POWO)
```

### 2023-03-23 code/L0/L0_NorthAndes_plants_BIEN.Rmd

---
title: "Northern Andes plant species BIEN data"
authors: "Hazel J. Anderson"
collaborators: "None"
data input: "none"
data output: "NorthAndes_BIEN_species_list.csv, NorthAndes_BIEN_occurrence.csv, NorthAndes_BIEN_traits.csv"
project: "Frugivoria"
date: "2023-02-13"
output: html_document
---
# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(BIEN)
library(tidyr)
library(dplyr)
library(DataExplorer)
library(summarytools)
library(bdvis)
library(visdat)
```

# Get data for Ecuador from BIEN database
```{r}
# occurrence data
Ecuador_BIEN_occ <- BIEN_occurrence_country("Ecuador", all.taxonomy = TRUE, observation.type = TRUE, collection.info = TRUE, political.boundaries = TRUE, native.status = TRUE)
# species list
Ecuador_BIEN_species_list <- BIEN_list_country(country = "Ecuador")
# trait data
Ecuador_BIEN_traits <- BIEN_trait_country("Ecuador", all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
```

```{r}
#remove duplicate data_collected column from Ecuador_BIEN_occ
Ecuador_BIEN_occ <- subset (Ecuador_BIEN_occ, select = -date_collected)
```

# Get data for Venezuela from BIEN database
```{r}
# occurrence data
Venezuela_BIEN_occ <- BIEN_occurrence_country("Venezuela", all.taxonomy = TRUE, observation.type = TRUE, collection.info = TRUE, political.boundaries = TRUE, native.status = TRUE)
# species list
Venezuela_BIEN_species_list <- BIEN_list_country(country = "Venezuela")
# trait data
Venezuela_BIEN_traits <- BIEN_trait_country("Venezuela", all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
```

```{r}
#remove duplicate data_collected column from Ecuador_BIEN_occ
Venezuela_BIEN_occ <- subset (Venezuela_BIEN_occ, select = -date_collected)
```

# Get data for Colombia from BIEN database
```{r}
# occurrence data
Colombia_BIEN_occ <- BIEN_occurrence_country("Colombia", all.taxonomy = TRUE, observation.type = TRUE, collection.info = TRUE, political.boundaries = TRUE, native.status = TRUE)
# species list
Colombia_BIEN_species_list <- BIEN_list_country(country = "Colombia")
# trait data
Colombia_BIEN_traits <- BIEN_trait_country("Colombia", all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
```

```{r}
#remove duplicate data_collected column from Ecuador_BIEN_occ
Colombia_BIEN_occ <- subset (Colombia_BIEN_occ, select = -date_collected)
```

# Combine BIEN data for Ecuador, Colombia, and Venezuela
```{r}
NorthAndes_BIEN_occ <- rbind(Ecuador_BIEN_occ, Venezuela_BIEN_occ, Colombia_BIEN_occ)
NorthAndes_BIEN_species_list <- rbind(Ecuador_BIEN_species_list, Venezuela_BIEN_species_list, Colombia_BIEN_species_list)
NorthAndes_BIEN_traits <- rbind(Ecuador_BIEN_traits, Venezuela_BIEN_traits, Colombia_BIEN_traits)
```

# Data summary
```{r}
glimpse(NorthAndes_BIEN_occ)
```

```{r}
glimpse(NorthAndes_BIEN_species_list)
```

```{r}
glimpse(NorthAndes_BIEN_traits)
```

### using DataExplorer package
```{r}
plot_str(NorthAndes_BIEN_occ)
introduce(NorthAndes_BIEN_occ)
plot_intro(NorthAndes_BIEN_occ)
plot_missing(NorthAndes_BIEN_occ)
plot_bar(NorthAndes_BIEN_occ)
```

```{r}
plot_str(NorthAndes_BIEN_species_list)
introduce(NorthAndes_BIEN_species_list)
plot_intro(NorthAndes_BIEN_species_list)
```

```{r}
plot_str(NorthAndes_BIEN_traits)
introduce(NorthAndes_BIEN_traits)
plot_intro(NorthAndes_BIEN_traits)
plot_missing(NorthAndes_BIEN_traits)
plot_bar(NorthAndes_BIEN_traits)
```

```{r}
# using summarytools package
NorthAndes_BIEN_occ_summary <- dfSummary(NorthAndes_BIEN_occ, 
            na.col       = FALSE,
            style        = "multiline",
            plain.ascii  = TRUE,
            graph.magnif = .8)
print(NorthAndes_BIEN_occ_summary, method = "render")
view(NorthAndes_BIEN_occ_summary, file = file.path(output_path,"NorthAndes_BIEN_occ_summary.html"))
```

```{r}
NorthAndes_BIEN_species_list_summary <- dfSummary(NorthAndes_BIEN_species_list, 
            na.col       = FALSE,
            style        = "multiline",
            plain.ascii  = TRUE,
            graph.magnif = .8)
print(NorthAndes_BIEN_species_list_summary, method = "render")
view(NorthAndes_BIEN_species_list_summary, file = file.path(output_path,"NorthAndes_BIEN_species_list_summary.html"))
```

```{r}
NorthAndes_BIEN_traits_summary <- dfSummary(NorthAndes_BIEN_traits, 
            na.col       = FALSE,
            style        = "multiline",
            plain.ascii  = TRUE,
            graph.magnif = .8)
print(NorthAndes_BIEN_traits_summary, method = "render")
view(NorthAndes_BIEN_traits_summary, file = file.path(output_path,"NorthAndes_BIEN_traits_summary.html"))
```
```{r}
#using bdvis package
NorthAndes_BIEN_occ_vis <- format_bdvis(NorthAndes_BIEN_occ, Latitude = "latitude", Longitude = "longitude", Date_collected = "date_collected", Scientific_name = "verbatim_scientific_name")
bdsummary(NorthAndes_BIEN_occ_vis)
```
```{r}
## ERROR - Error in !dbPreExists : invalid argument type
## ERROR - Error in postgresqlNewConnection(drv, ...) :
# RPosgreSQL error: could not connect postgres@localhost:5432 on dbname "test": could not connect to  server: Connection refused (0x0000274D/10061)

#bdcomplete(NorthAndes_BIEN_occ_vis)
#distrigraph(NorthAndes_BIEN_occ_vis, ptype = "species")
#distrigraph(NorthAndes_BIEN_occ_vis, ptype = "efforts")
#distrigraph(NorthAndes_BIEN_occ_vis, ptype = "cell")
#distrigraph(NorthAndes_BIEN_occ_vis, ptype = "effortspecies")
```

```{r}
#using visdat package
vis_dat(NorthAndes_BIEN_occ, warn_large_data = FALSE)
vis_miss(NorthAndes_BIEN_occ, warn_large_data = FALSE)
NorthAndes_BIEN_occ %>%
  select_if(is.numeric)%>%
  vis_cor(warn_large_data = FALSE)
```

```{r}
#using visdat package
vis_dat(NorthAndes_BIEN_traits, warn_large_data = FALSE)
vis_miss(NorthAndes_BIEN_traits, warn_large_data = FALSE)
NorthAndes_BIEN_traits %>%
  select_if(is.numeric)%>%
  vis_cor(warn_large_data = FALSE)
```

# Write data to csv
```{r}
NorthAndes_BIEN_occurrence_path <- file.path(output_path,"NorthAndes_BIEN_occurrence.csv") 
write.csv(NorthAndes_BIEN_occ, file = NorthAndes_BIEN_occurrence_path)

NorthAndes_BIEN_species_list_path <- file.path(output_path, "NorthAndes_BIEN_species_list.csv")
write.csv(NorthAndes_BIEN_species_list, file = NorthAndes_BIEN_species_list_path)

NorthAndes_BIEN_traits_path <- file.path(output_path, "NorthAndes_BIEN_traits.csv")
write.csv(NorthAndes_BIEN_traits, file = NorthAndes_BIEN_traits_path)
```

### 2023-03-23 code/L1/L1_NorthAndes_plants_BIEN_occ.Rmd

--- 2022-12-02 code/L1/L1_Ecuador_plants_BIEN_occ.Rmd
@@
 # darwinizing a dataset
-config <- read.csv(file.path(output_path,"configuration table.csv"))
+config <- read.csv(file.path(output_path,"Ecuador_plant_BIEN_occ_ConfigurationTable"))
 

### 2023-03-23 archive/L0_Ecuador_plants_GIFT.Rmd

--- 2023-01-22 code/L0/L0_Ecuador_plants_GIFT.Rmd
@@
 ```{r}
-#run first time only
-#remotes::install_github("https://github.com/BioGeoMacro/GIFT",dependencies = TRUE)
+remotes::install_github("https://github.com/BioGeoMacro/GIFT",dependencies = TRUE)
 library(GIFT)

### 2023-03-23 code/L0/L0_NorthAndes_plants_TRY.Rmd

--- 2023-01-22 code/L0/L0_Ecuador_plants_TRY.Rmd
@@
 library(rtry)
-library(mde)
 library(DataExplorer)
@@
 plot_missing(TRY_data)
-plot_bar(TRY_data)
 ```

### 2023-03-23 code/L0/L0_NorthAndes_frugivores_GBIF.Rmd
inital commit

---
title: "Northern Andes frugivore species GBIF"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "NorthAndes_GBIF_frugivore_occ.csv"
project: "Frugivoria"
date: "2023-03-23"
output: html_document
---
# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Install and load required packages
```{r}
library(rgbif)
library(DataExplorer)
library(bdvis)
library(visdat)
library(dplyr)
```

# Get Taxon Key for Mammals & Birds
```{r}
name_backbone("Mammalia")
name_backbone("Aves")
```

```{r}
occ_download(
pred("taxonKey", c("359", "212")), 
pred("hasGeospatialIssue", FALSE),
pred("hasCoordinate", TRUE),
pred_in("country",c("EC","CO","VE")), 
pred_not(pred_in("basisOfRecord",c("FOSSIL_SPECIMEN","LIVING_SPECIMEN"))),
format = "SIMPLE_CSV"
)
```

# Retrive GBIF download, save file, and load into r
```{r}
d <- occ_download_get("0282867-220831081235567", path = output_path) %>%
  occ_download_import()
```

# Summary 
```{r}
glimpse(d)
```

```{r}
# using DataExplorer package
plot_str(d)
introduce(d)
plot_intro(d)
plot_missing(d)
plot_bar(d)
```

```{r}
#using bdvis package
d_vis <- format_bdvis(d, Latitude = "decimalLatitude", Longitude = "decimalLongitude", Date_collected = "eventDate", Scientific_name = "scientificName", Family = "family", Genus = "genus")
bdsummary(d_vis)
```

```{r}
#using visdat package
vis_dat(d, warn_large_data = FALSE)
vis_miss(d, warn_large_data = FALSE)
d %>%
  select_if(is.numeric)%>%
  vis_cor(warn_large_data = FALSE)
```

# Write data to csv
```{r}
NorthAndes_GBIF_path <- file.path(output_path,"NorthAndes_GBIF_frugivore_occ.csv") 
write.csv(d, file = NorthAndes_GBIF_path)
```

### 2023-04-06 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
Initial commit

---
title: "Tropical Andes montane and lowland forest plant species GBIF"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "TropAndes_GBIF_plant_occ.csv"
project: "Frugivoria"
date: "2023-04-06"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(dplyr)
library(rgbif) 
```

# Read in IUCN data
Data was obtained from https://www.iucnredlist.org/search?permalink=9621ab77-2341-4bf8-9557-b7f19f7f2a35
```{r}
plants_habitats <- read.csv(file.path(data_path,"IUCN/Tropical Andes Plants - Search Results/habitats.csv"))
```

# Create species list from IUCN data
```{r}
plant_species <- unique(plants_habitats$scientificName)
```

# Match species names to GBIF
```{r}
gbif_taxon_keys <- plant_species %>% 
name_backbone_checklist() %>% # match to backbone 
filter(!matchType == "NONE") %>% # get matched names
pull(usageKey) 
```

# Download data
```{r}
occ_download(
pred_in("taxonKey", gbif_taxon_keys), # important to use pred_in
pred("hasCoordinate", TRUE),
pred("hasGeospatialIssue", FALSE),
pred_in("country",c("EC","CO","VE")), 
format = "SIMPLE_CSV"
)
```

# Retrive GBIF download, save file, and load into r
```{r}
d <- occ_download_get("0149390-230224095556074", path = output_path) %>%
  occ_download_import()
```

# Summary 
```{r}
glimpse(d)
```

# Write data to csv
```{r}
TropicalAndes_GBIF_path <- file.path(output_path,"TropAndes_GBIF_plant_occ.csv") 
write.csv(d, file = TropicalAndes_GBIF_path)
```

### 2023-04-06 code/L0/L0_TropicalAndes_frugivores_GBIF.Rmd
Initial commit

--- 2023-04-06 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
@@
 ---
-title: "Tropical Andes montane and lowland forest plant species GBIF"
+title: "Tropical Andes montane and lowland forest bird and mammal species GBIF"
 author: "Hazel J. Anderson"
@@
 data input: "None"
-data output: "TropAndes_GBIF_plant_occ.csv"
+data output: "TropAndes_GBIF_frugivore_occ.csv"
 project: "Frugivoria"
@@
 # Read in IUCN data
-Data was obtained from https://www.iucnredlist.org/search?permalink=9621ab77-2341-4bf8-9557-b7f19f7f2a35
+Data was obtained from https://www.iucnredlist.org/search?permalink=f4407d96-c039-413a-8d46-707f262271ee
 ```{r}
-plants_habitats <- read.csv(file.path(data_path,"IUCN/Tropical Andes Plants - Search Results/habitats.csv"))
+frugivore_habitats <- read.csv(file.path(data_path,"IUCN/Tropical Andes Birds and Mammals - Search Results/habitats.csv"))
 ```
@@
 ```{r}
-plant_species <- unique(plants_habitats$scientificName)
+frugivore_species <- unique(frugivore_habitats$scientificName)
 ```
@@
 ```{r}
-gbif_taxon_keys <- plant_species %>% 
+gbif_taxon_keys <- frugivore_species %>% 
 name_backbone_checklist() %>% # match to backbone 
@@
 ```{r}
-TropicalAndes_GBIF_path <- file.path(output_path,"TropAndes_GBIF_plant_occ.csv") 
+TropicalAndes_GBIF_path <- file.path(output_path,"TropAndes_GBIF_frugivore_occ.csv") 
 write.csv(d, file = TropicalAndes_GBIF_path)

### 2023-04-10 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
update for Tropical Andes

@@
 plant_species <- unique(plants_habitats$scientificName)
+write.csv2(plant_species, file = file.path(output_path, "TropicalPlantSpecies.csv"))
 ```
@@
 occ_download(
-pred_in("taxonKey", gbif_taxon_keys), # important to use pred_in
+pred_in("taxonKey", gbif_taxon_keys),
 pred("hasCoordinate", TRUE),
 pred("hasGeospatialIssue", FALSE),
-pred_in("country",c("EC","CO","VE")), 
+pred_in("country",c("EC","CO","VE", "PE", "BO")),
 format = "SIMPLE_CSV"
@@
 ```{r}
-d <- occ_download_get("0149390-230224095556074", path = output_path) %>%
+d <- occ_download_get("0154091-230224095556074", path = output_path) %>%
   occ_download_import()

### 2023-04-10 code/L0/L0_TropicalAndes_frugivores_GBIF.Rmd
update for Tropical Andes and species list

@@
 frugivore_species <- unique(frugivore_habitats$scientificName)
+write.csv2(frugivore_species, file = file.path(output_path, "TropicalFrugivoreSpecies.csv"))
 ```
@@
 occ_download(
-pred_in("taxonKey", gbif_taxon_keys), # important to use pred_in
+pred_in("taxonKey", gbif_taxon_keys),
 pred("hasCoordinate", TRUE),
 pred("hasGeospatialIssue", FALSE),
-pred_in("country",c("EC","CO","VE")), 
+pred_in("country",c("EC","CO","VE", "PE", "BO")), 
 format = "SIMPLE_CSV"
@@
 ```{r}
-d <- occ_download_get("0149390-230224095556074", path = output_path) %>%
+d <- occ_download_get("0154158-230224095556074", path = output_path) %>%
   occ_download_import()

### 2023-04-10 code/L0/L0_TropicalAndes_plants_BIEN_occ.Rmd
update for Tropical Andes

---
title: "Tropical Andes plant species BIEN data"
authors: "Hazel J. Anderson"
collaborators: "None"
data input: "none"
data output: "TropicalAndes_BIEN_species_list.csv, TropicalAndes_BIEN_occurrence.csv"
project: "Frugivoria"
date: "2023-02-13"
output: html_document
---
# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(BIEN)
library(tidyr)
library(dplyr)
```

# Read in IUCN data
Data was obtained from https://www.iucnredlist.org/search?permalink=9621ab77-2341-4bf8-9557-b7f19f7f2a35
```{r}
plants_habitats <- read.csv(file.path(data_path,"IUCN/Tropical Andes Plants - Search Results/habitats.csv"))
```

# Create species list from IUCN data
```{r}
plant_species <- unique(plants_habitats$scientificName)
```

# List of countries of interest
```{r}
country_vector <- c("Ecuador", "Venezuela", "Peru", "Colombia", "Bolivia")
```

# Get data for Tropical Andes from BIEN database
```{r}
# occurrence data
#TropicalAndes_BIEN_occ <- BIEN_occurrence_box(min.lat = -94, max.lat = -54, min.long = -24, max.long = 14,species = plant_species, all.taxonomy = TRUE, political.boundaries = TRUE, observation.type = TRUE)
#line above returned 0 records
TropicalAndes_BIEN_occ <- BIEN_occurrence_country(country = country_vector, all.taxonomy = TRUE, observation.type = TRUE, political.boundaries = TRUE, collection.info = TRUE)

# species list
TropicalAndes_BIEN_species_list <- BIEN_list_country(country = country_vector)
```

# Data summary
```{r}
glimpse(TropicalAndes_BIEN_occ)
```

```{r}
glimpse(TropicalAndes_BIEN_species_list)
```

# Write data to csv
```{r}
TropicalAndes_BIEN_occurrence_path <- file.path(output_path,"TropicalAndes_BIEN_occurrence.csv") 
write.csv(TropicalAndes_BIEN_occ, file = TropicalAndes_BIEN_occurrence_path)

TropicalAndes_BIEN_species_list_path <- file.path(output_path, "TropicalAndes_BIEN_species_list.csv")
write.csv(TropicalAndes_BIEN_species_list, file = TropicalAndes_BIEN_species_list_path)
```

### 2023-04-12 MappingGBIF_Richness.RMD
Raw GBIF Richness mapping

Mapping GBIF Richness
Adapted from https://luisdva.github.io/rstats/richness/

# Load required packages
```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(smoothr)
library(purrr)
library(raster)
library(scico)
library(ggspatial)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Read in Data
```{r}
NorthAndes_GBIF_plant_occ <- read.csv(file.path(data_path,"TropAndes_GBIF_plant_occ.csv"))
NorthAndes_GBIF_frugivore_occ <- read.csv(file.path(data_path,"TropAndes_GBIF_frugivore_occ.csv"))
```

# Polygons of countries
```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
#polygon of Tropical Andes
TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
```

```{r}
plants.sf <- st_as_sf(NorthAndes_GBIF_plant_occ, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
```

```{r}
frugivores.sf <- st_as_sf(NorthAndes_GBIF_frugivore_occ, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
```

```{r}
plants_sf_species <- plants.sf %>%
  group_by(species) %>%
  summarise()

frugivores_sf_species <- frugivores.sf %>%
  group_by(species) %>%
  summarise()
```

# Plot Base Map
```{r}
basePlot <-
ggplot() +
geom_sf(data = worldMap, fill = "white") +
geom_sf(data = TApoly) +
labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
annotation_scale(location = "bl", width_hint = 0.5) +
annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
theme(panel.background = element_rect(fill = "lightblue"))

basePlot
```

# Plot points
```{r}
plantsPointsPlot <-
  ggplot() +
  geom_sf(data = worldMap, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = plants_sf_species, pch = 21) +
  labs(title = "Tropical Andes Plant Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  )
plantsPointsPlot
```

```{r}
frugivoresPointsPlot <-
  ggplot() +
  geom_sf(data = worldMap, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = frugivores_sf_species, pch = 21) +
  labs(title = "Tropical Andes Bird and Mammal Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  )
frugivoresPointsPlot
```

# Make 0.5 Grid
```{r}
# make grid
TAGrid0.5 <- TApoly %>%
  st_make_grid(cellsize = c(0.5)) %>%
  st_intersection(TApoly) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())
```

# Plot Empty Grid
```{r}
# empty grid
gridPlot <-
  ggplot() +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly) +
  geom_sf(data = TAGrid0.5) +
  labs(title = "Tropical Andes Grid [0.5°]", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  )
gridPlot
```

# Cell Richness
```{r}
plant_richness_grid0.5 <- TAGrid0.5 %>%
  st_join(plants_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

```{r}
frugivore_richness_grid0.5 <- TAGrid0.5 %>%
  st_join(frugivores_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

# Richness plots
```{r}
plantgridRichTA0.5 <-
  ggplot(plant_richness_grid0.5) +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1, end = 0.9, midpoint = 1500) +
  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
  labs(title = "Tropical Andes Plant Richness [0.5°]", x = "Latitude", y = "Longitude", fill = "Number of species")
plantgridRichTA0.5
```

```{r}
frugivoregridRichTA0.5 <-
  ggplot(frugivore_richness_grid0.5) +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1, end = 0.9, midpoint = 1500) +
  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  ) +   
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
  labs(title = "Tropical Andes Bird and Mammal Richness [0.5°]", x = "Latitude", y = "Longitude", fill = "Number of species")
frugivoregridRichTA0.5
```

# Extract cell values
```{r}
plant_cellRichness0.5 <- plantgridRichTA0.5[["data"]]
frugivore_cellRichness0.5 <- frugivoregridRichTA0.5[["data"]]
```

```{r}
plot(plant_cellRichness0.5$num_species, frugivore_cellRichness0.5$num_species,
     main = "Plant Richness vs Bird & Mammal Richness by cell  [0.5°]",
     xlab = "Plant Richness by cell", ylab = "Bird & Mammal Richness by cell")
abline(lm(frugivore_cellRichness0.5$num_species ~ plant_cellRichness0.5$num_species))
summary(lm(frugivore_cellRichness0.5$num_species ~ plant_cellRichness0.5$num_species))
```

```{r}
hist(plant_cellRichness0.5$num_species,
     main = "Histogram of Plant Richness by cell  [0.5°]",
     xlab = "Plant Richness by cell")
hist(frugivore_cellRichness0.5$num_species,
     main = "Histogram of Bird & Mammal Richness by cell  [0.5°]",
     xlab = "Bird & Mammal Richness by cell")
```

```{r}
summary(plant_cellRichness0.5$num_species)
summary(frugivore_cellRichness0.5$num_species)
```

# Make 0.9 Grid
```{r}
# make grid
TAGrid0.9 <- TApoly %>%
  st_make_grid(cellsize = c(0.9)) %>%
  st_intersection(TApoly) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())
```

# Plot Empty Grid
```{r}
# empty grid
gridPlot <-
  ggplot() +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly) +
  geom_sf(data = TAGrid0.9) +
  labs(title = "Tropical Andes Grid [0.9°]", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  )
gridPlot
```

# Cell Richness
```{r}
plant_richness_grid0.9 <- TAGrid0.9 %>%
  st_join(plants_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

```{r}
frugivore_richness_grid0.9 <- TAGrid0.9 %>%
  st_join(frugivores_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

# Richness plots
```{r}
plantgridRichTA0.9 <-
  ggplot(plant_richness_grid0.9) +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1, end = 0.9, midpoint = 1500) +
  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
  labs(title = "Tropical Andes Plant Richness [0.9°]", x = "Latitude", y = "Longitude", fill = "Number of species")
plantgridRichTA0.9
```

```{r}
frugivoregridRichTA0.9 <-
  ggplot(frugivore_richness_grid0.9) +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1, end = 0.9, midpoint = 1500) +
  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  ) +   
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
  labs(title = "Tropical Andes Bird and Mammal Richness [0.9°]", x = "Latitude", y = "Longitude", fill = "Number of species")
frugivoregridRichTA0.9
```

# Extract cell values
```{r}
plant_cellRichness0.9 <- plantgridRichTA0.9[["data"]]
frugivore_cellRichness0.9 <- frugivoregridRichTA0.9[["data"]]
```

```{r}
plot(plant_cellRichness0.9$num_species, frugivore_cellRichness0.9$num_species,
     main = "Plant Richness vs Bird & Mammal Richness by cell  [0.9°]",
     xlab = "Plant Richness by cell", ylab = "Bird & Mammal Richness by cell")
abline(lm(frugivore_cellRichness0.9$num_species ~ plant_cellRichness0.9$num_species))
summary(lm(frugivore_cellRichness0.9$num_species ~ plant_cellRichness0.9$num_species))
```

```{r}
hist(plant_cellRichness0.9$num_species,
     main = "Histogram of Plant Richness by cell  [0.9°]",
     xlab = "Plant Richness by cell")
hist(frugivore_cellRichness0.9$num_species,
     main = "Histogram of Bird & Mammal Richness by cell  [0.9°]",
     xlab = "Bird & Mammal Richness by cell")
```

```{r}
summary(plant_cellRichness0.9$num_species)
summary(frugivore_cellRichness0.9$num_species)
```

# Make 1.2 Grid
```{r}
# make grid
TAGrid1.4 <- TApoly %>%
  st_make_grid(cellsize = c(1.4)) %>%
  st_intersection(TApoly) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())
```

# Plot Empty Grid
```{r}
# empty grid
gridPlot <-
  ggplot() +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly) +
  geom_sf(data = TAGrid1.4) +
  labs(title = "Tropical Andes Grid  [1.4°]", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  )
gridPlot
```

# Cell Richness
```{r}
plant_richness_grid1.4 <- TAGrid1.4 %>%
  st_join(plants_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

```{r}
frugivore_richness_grid1.4 <- TAGrid1.4 %>%
  st_join(frugivores_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

# Richness plots
```{r}
plantgridRichTA1.4 <-
  ggplot(plant_richness_grid1.4) +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1, end = 0.9) +
  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
  labs(title = "Tropical Andes Plant Richness  [1.4°]", x = "Latitude", y = "Longitude", fill = "Number of species")
plantgridRichTA1.4
```

```{r}
frugivoregridRichTA1.4 <-
  ggplot(frugivore_richness_grid1.4) +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1, end = 0.9, midpoint = 1500) +
  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  ) +   
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
  labs(title = "Tropical Andes Bird and Mammal Richness [1.4°]", x = "Latitude", y = "Longitude", fill = "Number of species")
frugivoregridRichTA1.4
```

# Extract cell values
```{r}
plant_cellRichness1.4 <- plantgridRichTA1.4[["data"]]
frugivore_cellRichness1.4 <- frugivoregridRichTA1.4[["data"]]
```

```{r}
plot(plant_cellRichness1.4$num_species, frugivore_cellRichness1.4$num_species,
     main = "Plant Richness vs Bird & Mammal Richness by cell  [1.4°]",
     xlab = "Plant Richness by cell", ylab = "Bird & Mammal Richness by cell")
abline(lm(frugivore_cellRichness1.4$num_species ~ plant_cellRichness1.4$num_species))
summary(lm(frugivore_cellRichness1.4$num_species ~ plant_cellRichness1.4$num_species))
```

```{r}
hist(plant_cellRichness1.4$num_species,
     main = "Histogram of Plant Richness by cell  [0.5°]",
     xlab = "Plant Richness by cell")
hist(frugivore_cellRichness1.4$num_species,
     main = "Histogram of Bird & Mammal Richness by cell  [1.4°]",
     xlab = "Bird & Mammal Richness by cell")
```

```{r}
summary(plant_cellRichness1.4$num_species)
summary(frugivore_cellRichness1.4$num_species)
```

### 2023-07-11 archive/MappingNorthernAndes.Rmd
update import file

Map of Northern Andes

```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(ggspatial)
```

```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
# country subset. 
ECpoly <- worldMap %>% filter(sovereignt == "Ecuador")
VEpoly <- worldMap %>% filter(sovereignt == "Venezuela")
COpoly <- worldMap %>% filter(sovereignt == "Colombia")
# set color for Northern Andes
Northern_Andes <- "lightgreen"
```

```{r}
MapPlot <-
ggplot() +
geom_sf(data = worldMap, fill = "White") +
geom_sf(data = ECpoly, fill = Northern_Andes) +
geom_sf(data = VEpoly, fill = Northern_Andes) +
geom_sf(data = COpoly, fill = Northern_Andes) +
labs(title = "Northern Andes", x = "Latitude", y = "Longitude") +
coord_sf(xlim = c(-97, -55), ylim = c(-12, 15), expand = FALSE) +
annotation_scale(location = "bl", width_hint = 0.5) +
annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
theme(panel.background = element_rect(fill = "lightblue"))

MapPlot
```

```{r}
ggsave(filename = "Northern Andes Map.png", dpi = 300, device="png")
```

### 2023-07-11 code/L1/CleanGBIFData.qmd
inital commit of cleaning plant gbif data

---
title: "Clean GBIF Data"
format: html
editor: visual
---

```{r}
library(countrycode)
library(CoordinateCleaner)
library(dplyr)
library(ggplot2)
library(rgbif)
library(sp)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Read in Data
```{r}
TropicalAndes_GBIF_plant_occ <- read.csv(file.path(data_path,"TropAndes_GBIF_plant_occ.csv"))
TropicalAndes_GBIF_frugivore_occ <- read.csv(file.path(data_path,"TropAndes_GBIF_frugivore_occ.csv"))
```

```{r}
names(TropicalAndes_GBIF_plant_occ)
```

```{r}
dim(TropicalAndes_GBIF_plant_occ)
dim(TropicalAndes_GBIF_frugivore_occ)
```

```{r}
# remove records without coordinates
TropicalAndes_GBIF_plant_occ_cleaned <- TropicalAndes_GBIF_plant_occ%>%
  filter(!is.na(decimalLongitude))%>%
  filter(!is.na(decimalLatitude))
TropicalAndes_GBIF_frugivore_occ_cleaned <- TropicalAndes_GBIF_frugivore_occ%>%
  filter(!is.na(decimalLongitude))%>%
  filter(!is.na(decimalLatitude))
```

```{r}
dim(TropicalAndes_GBIF_plant_occ_cleaned)
dim(TropicalAndes_GBIF_frugivore_occ_cleaned)
```

```{r}
#convert country code from ISO2c to ISO3c
TropicalAndes_GBIF_plant_occ_cleaned$countryCode <-  countrycode(TropicalAndes_GBIF_plant_occ_cleaned$countryCode, origin =  'iso2c', destination = 'iso3c')
TropicalAndes_GBIF_frugivore_occ_cleaned$countryCode <-  countrycode(TropicalAndes_GBIF_frugivore_occ_cleaned$countryCode, origin =  'iso2c', destination = 'iso3c')
```

```{r}
#flag problems
TropicalAndes_GBIF_plant_occ_cleaned <- data.frame(TropicalAndes_GBIF_plant_occ_cleaned)
plant_flags <- clean_coordinates(x = TropicalAndes_GBIF_plant_occ_cleaned, 
                           lon = "decimalLongitude", 
                           lat = "decimalLatitude",
                           countries = "countryCode",
                           species = "species",
                          tests = c("capitals", "centroids", "equal","gbif", "institutions",
                                    "duplicates","zeros", "countries"))
TropicalAndes_GBIF_frugivore_occ_cleaned<- data.frame(TropicalAndes_GBIF_frugivore_occ_cleaned)
frugivore_flags <- clean_coordinates(x = TropicalAndes_GBIF_frugivore_occ_cleaned, 
                           lon = "decimalLongitude", 
                           lat = "decimalLatitude",
                           countries = "countryCode",
                           species = "species",
                          tests = c("capitals", "centroids", "equal","gbif", "institutions",
                                    "duplicates","zeros", "countries"))
```

```{r}
summary(plant_flags)
plot(plant_flags, lon = "decimalLongitude", lat = "decimalLatitude")
summary(frugivore_flags)
plot(frugivore_flags, lon = "decimalLongitude", lat = "decimalLatitude")
```

```{r}
#Exclude problematic records
TropicalAndes_GBIF_plant_occ_cleaned <- TropicalAndes_GBIF_plant_occ_cleaned[plant_flags$.summary,]
TropicalAndes_GBIF_frugivore_occ_cleaned <- TropicalAndes_GBIF_frugivore_occ_cleaned[frugivore_flags$.summary,]
```

```{r}
dim(TropicalAndes_GBIF_plant_occ_cleaned)
dim(TropicalAndes_GBIF_frugivore_occ_cleaned)
```

# Sample Bias
```{r}
library(sampbias)
library(dplyr)
```

```{r}
#remove columns
TropicalAndes_GBIF_plant_occ_cleaned <- subset(TropicalAndes_GBIF_plant_occ_cleaned, select = -c(X, decimallongitude, decimallatitude))

TropicalAndes_GBIF_frugivore_occ_cleaned <- subset(TropicalAndes_GBIF_frugivore_occ_cleaned, select = -c(X, decimallongitude, decimallatitude))
```

```{r}
bias.out.plants <- calculate_bias(x = TropicalAndes_GBIF_plant_occ_cleaned, res = 1)
```

```{r}
bias.out.frugivores <- calculate_bias(x = TropicalAndes_GBIF_frugivore_occ_cleaned, res = 1)
```

### 2023-07-11 code/L2/GBIF_PlantData_GapAnalysis.Rmd
inital commit

GBIF Plant Gap Analyis

# Load required packages
```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(smoothr)
library(purrr)
library(raster)
library(scico)
library(ggspatial)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Read in Data
```{r}
TropicalAndes_GBIF_plant_occ <- read.csv(file.path(data_path,"TropAndes_GBIF_plant_occ.csv"))
```

```{r}
summary(TropicalAndes_GBIF_plant_occ)
```

```{r}
TropicalAndes_GBIF_plant_occ %>% count(family, sort = TRUE, name = "Occurrences records")
```

```{r}
occ_by_country <- TropicalAndes_GBIF_plant_occ %>% count(countryCode, sort = TRUE, name = "occurrencesRecords") 
occ_by_country
occ_by_country$occurrencesRecords <- as.numeric(occ_by_country$occurrencesRecords)
occ_by_country
tbl <- with(occ_by_country, table(occ_by_country$countryCode, occ_by_country$occurrencesRecords))
barplot(tbl, beside = TRUE, legend = TRUE)
```

```{r}
TropicalAndes_GBIF_plant_occ %>% count(species, sort = TRUE, name = "Occurrences records")
```

```{r}
spc <- tapply(TropicalAndes_GBIF_plant_occ$species, TropicalAndes_GBIF_plant_occ$countryCode, function(x)length(unique(x)) )
as.data.frame(spc)
```

# Sample Bias
```{r}
library(sampbias)
```

```{r}
bias.out.plants <- calculate_bias(x = TropicalAndes_GBIF_plant_occ, res = 1)
```

```{r}
bias.out.frugivores <- calculate_bias(x = TropicalAndes_GBIF_frugivore_occ, res = 1)
```

```{r}
summary(bias.out.plants)
```

```{r}
summary(bias.out.frugivores)
```

```{r}
plot(bias.out.plants)
proj.plants <- project_bias(bias.out.plants)
map_bias(proj.plants, type = "log_sampling_rate")
map_bias(proj.plants, type = "sampling_rate")
map_bias(proj.plants, type = "diff_to_max")
```

```{r}
plot(bias.out.frugivores)
proj.frugivore <- project_bias(bias.out.frugivores)
map_bias(proj.frugivore, type = "log_sampling_rate")
map_bias(proj.frugivore, type = "sampling_rate")
map_bias(proj.frugivore, type = "diff_to_max")
```

# Number of Occurences per grid cell
Based on https://azizka.github.io/Biodiversity_Data_from_Field_to_Yield/04_species_richness_maps_tutorial.html

```{r}
library(readr)
library(ggplot2)
library(magrittr)
library(raster)
library(rgdal)
devtools::install_github("azizka/speciesgeocodeR")
library(speciesgeocodeR)
options(scipen=10000)
```

```{r}
# Visualize
world.inp <- map_data("world")

# The number of occurrences per grid cell
gocc <- RichnessGrid(TropicalAndes_GBIF_plant_occ, type = "abu")

plot(gocc)

# Or more sofisticated plotting Prepare for plotting
plo_gocc <- data.frame(rasterToPoints(gocc))

# Visualize
ggplot() + 
  geom_map(data = world.inp, map = world.inp,
           aes(x = long, y = lat, map_id = region), fill = "grey80") +
  xlim(min(TropicalAndes_GBIF_plant_occ$decimalLongitude, na.rm = T),
    max(TropicalAndes_GBIF_plant_occ$decimalLongitude, na.rm = T)) +
  ylim(min(TropicalAndes_GBIF_plant_occ$decimalLatitude, na.rm = T), 
    max(TropicalAndes_GBIF_plant_occ$decimalLatitude, na.rm = T)) + 
  geom_raster(data = plo_gocc, aes(x = x, y = y, fill = layer)) +
  scale_fill_viridis_c(name="Count") +
  coord_fixed() +
  theme_bw() +
  ggtitle("Number of GBIF Plant occurrence records") +
  ylab("Latitude") +
  xlab("Longitude")

# The number of species per grid cell
gri <- RichnessGrid(TropicalAndes_GBIF_plant_occ)

plot(gri)
```

```{r}
# Visualize
world.inp <- map_data("world")

# The number of occurrences per grid cell
gocc <- RichnessGrid(TropicalAndes_GBIF_frugivore_occ, type = "abu")

plot(gocc)

# Or more sofisticated plotting Prepare for plotting
plo_gocc <- data.frame(rasterToPoints(gocc))

# The number of species per grid cell
gri <- RichnessGrid(TropicalAndes_GBIF_frugivore_occ)

plot(gri)
```

```{r}
ggplot() +
  geom_map(data = world.inp, map = world.inp,
           aes(x = long, y = lat, map_id = region), fill = "grey80") +
  xlim(min(TropicalAndes_GBIF_frugivore_occ$decimalLongitude, na.rm = T), 
    max(TropicalAndes_GBIF_frugivore_occ$decimalLongitude, na.rm = T)) +
  ylim(min(TropicalAndes_GBIF_frugivore_occ$decimalLatitude, na.rm = T), 
    max(TropicalAndes_GBIF_frugivore_occ$decimalLatitude, na.rm = T)) +
  geom_raster(data = plo_gocc, aes(x = x, y = y, fill = layer)) +
  scale_fill_viridis_c(name="Count") +
  coord_fixed() +
  theme_bw() +
  ggtitle("Number of GBIF Bird & Mammal occurrence records") +
  ylab("Latitude") +
  xlab("Longitude")
```

### 2023-07-11 code/L2/EvaluatingIUCNAssessedTropicalAndesPlantSpecies.Rmd
inital commit

Evaluating Assessed Tropical Andes plant species from the IUCN and POWO

Modified from https://barnabywalker.github.io/kewr/articles/mapping-assessed-species.html

# Load required packages
```{r}
library(kewr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# 1. Getting assessment information

```{r}
# Load in data
data <- read.csv(file.path(data_path,"IUCN/Tropical Andes Plants - Search Results/simple_summary.csv"))
```

```{r}
# Subset to just species names
TropicalAndes_plants <- data[ , c('internalTaxonId', 'scientificName', 'authority', 'redlistCategory')]
```

# 2. Match names to an IPNI ID
For large data sets use http://namematch.science.kew.org/csv
```{r}
TropicalAndes_plants <-
  TropicalAndes_plants %>%
  unite("full_name", scientificName, authority, sep=" ", remove=FALSE)
```

```{r}
full_matches <- match_knms(TropicalAndes_plants$full_name)
full_matches
```

```{r}
full_matches <- tidy(full_matches)

unmatched <- filter(full_matches, !matched)

to_match <- 
  TropicalAndes_plants %>%
  filter(full_name %in% unmatched$submitted)

part_matches <- match_knms(to_match$scientificName)
part_matches
```

```{r}
part_matches <- tidy(part_matches)

full_matches <- 
  full_matches %>%
  filter(matched) %>%
  left_join(
    TropicalAndes_plants %>% dplyr::select(internalTaxonId, full_name, redlistCategory),
    by=c("submitted"="full_name")
  )

part_matches <- 
  part_matches %>%
  filter(matched) %>%
  left_join(
    TropicalAndes_plants %>% dplyr::select(internalTaxonId, scientificName, redlistCategory),
    by=c("submitted"="scientificName")
  )

matched_names <- bind_rows(full_matches, part_matches)

head(matched_names)
```

# 3. Resolve assessments to accepted names
```{r}
# wrap up the lookup_wcvp function to make sure it comes back as a list
f <- function(internalTaxonId) {
  list(lookup_wcvp(internalTaxonId))
}

resolved_names <- 
  matched_names %>%
  dplyr::nest_by(internalTaxonId, ipni_id) %>%
  dplyr::mutate(matched_record=f(ipni_id))

resolved_names <-
  resolved_names %>%
  dplyr::mutate(status=matched_record$status)
```

```{r}
resolved_names <-
  resolved_names %>%
  filter(status %in% c("accepted", "homotypic synonym"))
```

```{r}
resolved_names %>%
  ungroup() %>%
  add_count(internalTaxonid) %>%
  summarise(multiple_matches=sum(n > 1))
```

```{r}
resolved_names <-
  resolved_names %>%
  mutate(accepted_id=ifelse(status != "accepted",
                              wcvp_record$accepted$id,
                              wcvp_record$id),
         accepted_name=ifelse(status != "accepted",
                              wcvp_record$accepted$name,
                              wcvp_record$name),
         accepted_author=ifelse(status != "accepted",
                              wcvp_record$accepted$author,
                              wcvp_record$authors),
         accepted_rank=ifelse(status != "accepted",
                              wcvp_record$accepted$rank,
                              wcvp_record$rank))

resolved_names <- 
  resolved_names %>%
  select(-wcvp_record) %>%
  unnest(cols=c(data)) %>%
  rename(match_id=ipni_id)

accepted_species <-
  resolved_names %>%
  filter(accepted_rank == "Species") %>%
  ungroup()

nrow(accepted_species)
```

# 4. Get a list of all species in the country
```{r}
Ecuador_checklist <- search_powo(list(distribution= "Ecuador"),
                         filters=c("accepted", "species"),
                        limit=5000)
Ecuador_checklist
Venezuela_checklist <- search_powo(list(distribution= "Venezuela"),
                         filters=c("accepted", "species"),
                        limit=5000)
Venezuela_checklist

Colombia_checklist <- search_powo(list(distribution= "Colombia"),
                         filters=c("accepted", "species"),
                        limit=5000)
Colombia_checklist

Peru_checklist <- search_powo(list(distribution= "Peru"),
                         filters=c("accepted", "species"),
                        limit=5000)
Peru_checklist

Bolivia_checklist <- search_powo(list(distribution= "Bolivia"),
                         filters=c("accepted", "species"),
                        limit=5000)
Bolivia_checklist
```
```{r}
Bolivia_checklist <- tidy(Bolivia_checklist)
Colombia_checklist <- tidy(Colombia_checklist)
Ecuador_checklist <- tidy(Ecuador_checklist)
Peru_checklist <- tidy(Peru_checklist)
Venezuela_checklist <- tidy(Venezuela_checklist)
```

```{r}
checklist <- rbind(Bolivia_checklist, Colombia_checklist, Ecuador_checklist, Peru_checklist, Venezuela_checklist)
```

```{r}
checklist <-
  checklist %>%
  mutate(ipni_id=str_extract(fqId, "[0-9\\-]+")) %>%
  select(ipni_id, family, name, author) %>%
  distinct(ipni_id, .keep_all=TRUE) %>%
  left_join(
    accepted_species %>% select(accepted_id, category),
    by=c("ipni_id"="accepted_id")
  )
```

# 5. Calculating the proportion of assessed species
```{r}
checklist %>%
  summarise(p_assessed=mean(! is.na(category)))
```

```{r}
iucn_colours <- c("NE"="#ffffff", 
                  "DD"="#d1d1d6", 
                  "LC"="#60c659", 
                  "NT"="#cce226", 
                  "VU"="#f9e814", 
                  "EN"="#fc7f3f", 
                  "CR"="d81e05", 
                  "EW"="#542344", 
                  "EX"="#000000")
checklist %>%
  replace_na(list(category="NE")) %>%
  mutate(category=factor(category, levels=names(iucn_colours),
                         ordered=TRUE)) %>%
  ggplot(mapping=aes(y=category, fill=category,
                     colour=category == "NE")) +
  geom_bar() +
  scale_fill_manual(values=iucn_colours, drop=FALSE) +
  scale_colour_manual(values=c(`TRUE`="black", `FALSE`=NA)) +
  scale_y_discrete(drop=FALSE) +
  guides(fill=FALSE, colour=FALSE) +
  labs(x="Number of species", y="IUCN Red List category")
```

### 2023-07-11 archive/L2_NorthAndes_plant_occ_option2.Rmd
move old code to archive

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(bdDwC)
```

# Load data
```{r}
gbif_data <- read.csv(file.path(data_path,"Ecuador_GBIF_occurences.csv"))
iucn_data <- read.csv(file.path(data_path, "Ecuador_IUCN_plant_occ.csv"))
bien_data <- read.csv(file.path(data_path, "Ecuador_BIEN_occurrence.csv"))
```

```{r}
# get newest version of Darwin Cloud Data
downloadCloudData()
```

```{r}
# darwinize GBIF dataframe
gbif_result <- darwinizeNames(dataUser = gbif_data,
                            dataDWC   = bdDwC:::dataDarwinCloud$data)
new_gbif_data <- renameUserData(gbif_data, gbif_result)
```

```{r}
# darwinize IUCN dataframe
iucn_result <- darwinizeNames(dataUser = iucn_data,
                            dataDWC   = bdDwC:::dataDarwinCloud$data)
new_iucn_data <- renameUserData(iucn_data, iucn_result)
```

```{r}
# darwinize BIEN dataframe
bien_result <- darwinizeNames(dataUser = bien_data,
                            dataDWC   = bdDwC:::dataDarwinCloud$data)
new_bien_data <- renameUserData(bien_data, bien_result)
```

# Write data to csv
```{r}
# renamed GBIF data
ecuador_GBIF_path <- file.path(output_path,"Ecuador_GBIF_plant_occ.csv")
write.csv(new_gbif_data, file = ecuador_GBIF_path)
# GBIF configuration table
ecuador_GBIF_table_path <- file.path(output_path,"Ecuador_plant_GBIF_ConfigurationTable.csv")
write.csv(gbif_result, file = ecuador_GBIF_table_path)
```

```{r}
# renamed IUCN data
ecuador_IUCN_path <- file.path(output_path,"Ecuador_IUCN_plant_occ.csv")
write.csv(new_iucn_data, file = ecuador_IUCN_path)
# IUCN configuration table
ecuador_IUCN_table_path <- file.path(output_path,"Ecuador_plant_IUCN_ConfigurationTable.csv")
write.csv(iucn_result, file = ecuador_IUCN_table_path)
```

```{r}
# renamed BIEN data
ecuador_BIEN_path <- file.path(output_path,"Ecuador_BIEN_plant_occ.csv")
write.csv(new_bien_data, file = ecuador_BIEN_path)

```

### 2023-07-13 archive/L0_NorthAndes_plants_GBIF.Rmd
move old code to archive

--- 2023-02-27 code/L0/L0_Ecuador_plants_GBIF.Rmd
@@
 library(DataExplorer)
-library(summarytools)
 library(bdvis)
 library(visdat)
+library(dplyr)
 ```
@@
 ```
+
 # Get Taxon Key for Plants
@@
 ```
+
 # Retrive GBIF download, save file, and load into r
@@
 ```{r}
-# using summarytools package
-# ERROR -- Error in hist.default(data, breaks = breaks_x, plot = FALSE) :
-# some 'x' not counted; maybe 'breaks' do not span range of 'x'
-#d_summary <- dfSummary(d, 
-#            na.col       = FALSE,
-#            style        = "multiline",
-#            plain.ascii  = TRUE,
-#            graph.magnif = .8)
-#print(d_summary, method = "render")
-#view(d_summary, file = file.path(output_path,"Ecuador_BIEN_traits_summary.html"))
-```
-
-```{r}
 #using bdvis package
@@
 bdsummary(d_vis)
-```
-
-```{r}
-## ERROR - Error in !dbPreExists : invalid argument type
-## ERROR - Error in postgresqlNewConnection(drv, ...) :
-# RPosgreSQL error: could not connect postgres@localhost:5432 on dbname "test": could not connect to server: Connection refused (0x0000274D/10061)
-#bdcalendarheat(d_vis)
-#distrigraph(d_vis, ptype = "species")
-#distrigraph(d_vis, ptype = "efforts")
-#distrigraph(d_vis, ptype = "cell")
-#distrigraph(d_vis, ptype = "effortspecies")
-#bdcomplete(d_vis)
 ```
@@
 ```{r}
-NorthAndes_GBIF_path <- file.path(output_path,"NorthAndes_GBIF_occurences.csv") 
+NorthAndes_GBIF_path <- file.path(output_path,"NorthAndes_GBIF_plant_occ.csv") 
 write.csv(d, file = NorthAndes_GBIF_path)

### 2023-07-13 archive/L0_NorthAndes_plants_IUCN.Rmd
move old code to archive

--- 2023-03-23 code/L0/L0_Ecuador_plants_IUCN.Rmd
@@
 ```{r}
-assessments <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/assessments.csv"))
-common_names <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/common_names.csv"))
-conservation_needed <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/conservation_needed.csv"))
-countries <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/countries.csv"))
-credits <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/credits.csv"))
-dois <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/dois.csv"))
-fao <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/fao.csv"))
-habitats <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/habitats.csv"))
-plant_specific <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/plant_specific.csv"))
-references <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/references.csv"))
-research_needed <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/research_needed.csv"))
-simple_summary <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/simple_summary.csv"))
-synonyms <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/synonyms.csv"))
-taxonomy <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/taxonomy.csv"))
-threats <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/threats.csv"))
-usetrade <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/usetrade.csv"))
+assessments <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/assessments.csv"))
+common_names <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/common_names.csv"))
+conservation_needed <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/conservation_needed.csv"))
+countries <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/countries.csv"))
+credits <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/credits.csv"))
+dois <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/dois.csv"))
+fao <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/fao.csv"))
+habitats <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/habitats.csv"))
+plant_specific <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/plant_specific.csv"))
+references <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/references.csv"))
+research_needed <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/research_needed.csv"))
+simple_summary <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/simple_summary.csv"))
+synonyms <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/synonyms.csv"))
+taxonomy <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/taxonomy.csv"))
+threats <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/threats.csv"))
+usetrade <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/usetrade.csv"))
 points_data <- read.csv(file.path(data_path,"IUCN/Northern Andes 3 - Range data Points/points_data.csv"))

### 2023-07-13 archive/L0_NorthAndes_plants_TRY.Rmd
move old code to archive

--- 2023-03-23 code/L0/L0_NorthAndes_plants_TRY.Rmd
@@
 ```{r}
-data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/TRY_Ecuador_PublicData_10192020.txt", encoding = "Latin-1")
+data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/23650_24112022083128/23650.txt", encoding = "Latin-1")
 data <- as.data.frame(data)

### 2023-07-13 archive/L1_NorthAndes_plants_TRY.Rmd
move old code to archive

--- 2022-12-02 code/L1/L1_Ecuador_plants_TRY.Rmd
@@
 # darwinizing a dataset
-config <- read.csv(file.path(output_path,"Ecuador_plant_TRY_trait_ConfigurationTable"))
+config <- read.csv(file.path(output_path,"Ecuador_plant_TRY_trait_ConfigurationTable.csv"))
 
@@
 # Pre-filter
+Modified from https://brunobrr.github.io/bdc/articles/prefilter.html
 ```{r}
@@
 # Flagging common time issues
+Modified from https://brunobrr.github.io/bdc/articles/time.html
 ```{r}
@@
 # Flagging common spatial issues
+Modified from https://brunobrr.github.io/bdc/articles/space.html
 ```{r}
@@
 # Clean and parse species names
+Modified from https://brunobrr.github.io/bdc/articles/taxonomy.html
 ```{r}
@@
 # Transform to wide table
+Modified from https://cran.r-project.org/web/packages/rtry/vignettes/rtry-workflow-general.html
 ```{r}

### 2023-07-18 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
inital commit set file paths, load packages

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(raster)
```

# Load in Ecosystem Functional Groups
Data downloaded from www.zendo.org/record/5090419
```{r}
iucn_habitats <- raster(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004.tif"))
```

### 2023-07-18 code/L0/L0_TropicalAndes_plants_BIEN_traits.Rmd
inital commit

---
title: ""
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: ""
data output: ""
date: "2023-07-18"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(BIEN)
library(tidyr)
library(dplyr)
```

# List of countries of interest
```{r}
country_vector <- c("Ecuador", "Venezuela", "Peru", "Colombia", "Bolivia")
```

```{r}
TropicalAndes_BIEN_traits <- BIEN_trait_country(country_vector, all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
```

# Data summary
```{r}
glimpse(TropicalAndes_BIEN_traits)
```

```{r}

```

# Write data to csv
```{r}
TropicalAndes_BIEN_traits_path <- file.path(output_path,"TropicalAndes_BIEN_traits.csv") 
write.csv(TropicalAndes_BIEN_traits, file = TropicalAndes_BIEN_traits_path)
```

### 2023-07-18 code/L0/L0_TropicalAndes_plants_BIEN_occ.Rmd
update file path

@@
 ```{r}
-TropicalAndes_BIEN_occurrence_path <- file.path(output_path,"TropicalAndes_BIEN_occurrence.csv") 
+TropicalAndes_BIEN_occurrence_path <- file.path(output_path,"TropicalAndes_BIEN_occ.csv") 
 write.csv(TropicalAndes_BIEN_occ, file = TropicalAndes_BIEN_occurrence_path)

### 2023-07-18 code/L0/L0_TropicalAndes_plants_BIEN_occ.Rmd
remove unneccessary code and update file paths

@@
 
-# Get data for Tropical Andes from BIEN database
+# Get occurrence data for Tropical Andes from BIEN database
 ```{r}
-# occurrence data
-#TropicalAndes_BIEN_occ <- BIEN_occurrence_box(min.lat = -94, max.lat = -54, min.long = -24, max.long = 14,species = plant_species, all.taxonomy = TRUE, political.boundaries = TRUE, observation.type = TRUE)
-#line above returned 0 records
 TropicalAndes_BIEN_occ <- BIEN_occurrence_country(country = country_vector, all.taxonomy = TRUE, observation.type = TRUE, political.boundaries = TRUE, collection.info = TRUE)
@@
 ```{r}
-TropicalAndes_BIEN_occurrence_path <- file.path(output_path,"TropicalAndes_BIEN_occ.csv") 
+TropicalAndes_BIEN_occurrence_path <- file.path(output_path,"TropicalAndes_BIEN_plant_occ.csv") 
 write.csv(TropicalAndes_BIEN_occ, file = TropicalAndes_BIEN_occurrence_path)
 
-TropicalAndes_BIEN_species_list_path <- file.path(output_path, "TropicalAndes_BIEN_species_list.csv")
+TropicalAndes_BIEN_species_list_path <- file.path(output_path, "TropicalAndes_BIEN_plant_species_list.csv")
 write.csv(TropicalAndes_BIEN_species_list, file = TropicalAndes_BIEN_species_list_path)

### 2023-07-18 code/L0/L0_TropicalAndes_plants_BIEN_traits.Rmd
add summary, fill in overview, add title, save output to csv

@@
 ---
-title: ""
+title: "Tropical Andes plant trait data BIEN"
 author: "Hazel J. Anderson"
@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: ""
-data input: ""
-data output: ""
+overview: "This script retrieves plant trait data from the BIEN database for plant species in the countries that make up the Tropical Andes."
+data input: "none"
+data output: "TropicalAndes_BIEN_traits.csv"
 date: "2023-07-18"
@@
 ```{r}
-
+print("The number of records is")
+nrow(TropicalAndes_BIEN_traits)
+print("The number of species is")
+length(unique(TropicalAndes_BIEN_traits$scrubbed_species_binomial))
+print("The number of families is")
+length(unique(TropicalAndes_BIEN_traits$family_matched))
 ```
@@
 ```{r}
-TropicalAndes_BIEN_traits_path <- file.path(output_path,"TropicalAndes_BIEN_traits.csv") 
+TropicalAndes_BIEN_traits_path <- file.path(output_path,"TropicalAndes_BIEN_plant_traits.csv") 
 write.csv(TropicalAndes_BIEN_traits, file = TropicalAndes_BIEN_traits_path)

### 2023-07-18 code/L0/L0_TropicalAndes_frugivores_frugivoria_traits.Rmd
inital committ

---
title: "Tropical Andes frugivore trait data Frugivoria"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: ""
data output: ""
date: "2023-07-18"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}

```

```{r}

```

# Summary 
```{r}
glimpse(d)
```

```{r}
print("The number of occurrences is")
nrow(d)
print("The number of species is" )
length(unique(d$species))
print("The number of families is" )
length(unique(d$family))
```

# Write data to csv
```{r}
TropicalAndes_Frugivoria_path <- file.path(output_path,"TropicalAndes_Frugivoria_frugivore_traits.csv") 
write.csv(d2, file = TropicalAndes_Frugivoria_path)
```

### 2023-07-18 code/L0/L0_TropicalAndes_plants_TRY_traits.Rmd
inital commit

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}

```

# Load TRY data
Data was obtained from https://www.try-db.org/TryWeb/Prop0.php 
```{r}
try_data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/23650_24112022083128/23650.txt", encoding = "Latin-1")
try_data <- as.data.frame(try_data)
```

```{r}

```

### 2023-07-18 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
update top info parse out IUCN habitat subset and running without habitat subset add additional summary information

--- 2023-04-06 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
@@
 ---
-title: "Tropical Andes montane and lowland forest plant species GBIF"
+title: "Tropical Andes plant occurrence data GBIF"
 author: "Hazel J. Anderson"
-collaborators: "None"
-data input: "None"
+project: "Plant-Frugivore Diversity"
+collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
+overview: "This script retrives plant occurrence data from GBIF using a subset of species with IUCN habitat designations ‘Forest-Subtropical/Tropical Moist Montane’ and/or ‘Forest-Subtropical/Tropical Moist Lowland’.
+Data is also retrived without the subset for species in the countries that make up the Tropical Andes to be spatially subset later, as not many plant species are assessed on the IUCN RedList.
+"
+data input: "IUCN/Tropical Andes Plants - Search Results/habitats.csv"
 data output: "TropAndes_GBIF_plant_occ.csv"
-project: "Frugivoria"
-date: "2023-04-06"
+date: "2023-07-18"
 output: html_document
@@
 
-# Read in IUCN data
+# Plant occurrences using IUCN habitat subset
+## Read in IUCN data
 Data was obtained from https://www.iucnredlist.org/search?permalink=9621ab77-2341-4bf8-9557-b7f19f7f2a35
@@
 
-# Create species list from IUCN data
+## Create species list from IUCN data
 ```{r}
 plant_species <- unique(plants_habitats$scientificName)
+print("The number of species on the IUCN is")
+length(plant_species)
 ```
 
-# Match species names to GBIF
+## Match species names to GBIF
 ```{r}
@@
 
-# Download data
+## Download data
 ```{r}
 occ_download(
-pred_in("taxonKey", gbif_taxon_keys), # important to use pred_in
+pred_in("taxonKey", gbif_taxon_keys),
 pred("hasCoordinate", TRUE),
 pred("hasGeospatialIssue", FALSE),
-pred_in("country",c("EC","CO","VE")), 
+pred_in("country",c("EC","CO","VE", "PE", "BO")),
 format = "SIMPLE_CSV"
@@
 
-# Retrive GBIF download, save file, and load into r
+## Retrive GBIF download, save file, and load into r
 ```{r}
-d <- occ_download_get("0149390-230224095556074", path = output_path) %>%
+d <- occ_download_get("0097296-230530130749713", path = output_path) %>%
   occ_download_import()
@@
 
-# Summary 
+## Summary 
 ```{r}
@@
 
-# Write data to csv
+```{r}
+print("The number of occurrences is")
+nrow(d)
+print("The number of species is" )
+length(unique(d$species))
+print("The number of families is" )
+length(unique(d$family))
+```
+
+# Plant occurrences without habitat subset
+
+## Download data
+```{r}
+occ_download(
+pred("hasCoordinate", TRUE),
+pred("hasGeospatialIssue", FALSE),
+pred_in("country",c("EC","CO","VE", "PE", "BO")),
+format = "SIMPLE_CSV"
+)
+```
+
+## Retrive GBIF download, save file, and load into r
+```{r}
+d2 <- occ_download_get("0097297-230530130749713", path = output_path, overwrite = TRUE) %>%
+  occ_download_import()
+```
+
+## Summary 
+```{r}
+glimpse(d2)
+```
+
+```{r}
+print("The number of occurrences is")
+nrow(d2)
+print("The number of species is" )
+length(unique(d2$species))
+print("The number of families is" )
+length(unique(d2$family))
+```
+
+## Write data to csv
 ```{r}
 TropicalAndes_GBIF_path <- file.path(output_path,"TropAndes_GBIF_plant_occ.csv") 
-write.csv(d, file = TropicalAndes_GBIF_path)
+write.csv(d2, file = TropicalAndes_GBIF_path)
 ```

### 2023-07-18 code/L0/L0_TropicalAndes_frugivores_frugivoria_traits.Rmd
inital commit add code from edi repo

@@
 
+# Retrive Frugivoria data
+Using r code provided at https://portal.edirepository.org/nis/codeGeneration?packageId=edi.1220.6&statisticalFileType=r
 ```{r}
+# Package ID: edi.1220.6 Cataloging System:https://pasta.edirepository.org.
+# Data set title: Frugivoria: A trait database for birds and mammals exhibiting frugivory across contiguous Neotropical moist forests .
+# Data set creator:  Beth Gerstner - Michigan State University 
+# Data set creator:  Phoebe Zarnetske - Michigan State University 
+# Data set creator:  Patrick Bills - Michigan State University 
+# Contact:  Beth Gerstner -  Michigan State University  - gerstn11@msu.edu
+# Stylesheet v2.11 for metadata conversion into program: John H. Porter, Univ. Virginia, jporter@virginia.edu
 
+inUrl1  <- "https://pasta.lternet.edu/package/data/eml/edi/1220/6/f2503359fb87b67d22e4a9a33f014107" 
+infile1 <- tempfile()
+try(download.file(inUrl1,infile1,method="curl"))
+if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")
+
+ dt1 <-read.csv(infile1,header=F 
+          ,skip=1
+            ,sep=","  
+                ,quot='"' 
+        , col.names=c(
+                    "IUCN_species_name",     
+                    "taxonid",     
+                    "common_name",     
+                    "family",     
+                    "genus",     
+                    "species",     
+                    "subspecies",     
+                    "rank",     
+                    "elton_species_name",     
+                    "taxonomic_disparity",     
+                    "species_notes",     
+                    "diet_inv_e",     
+                    "diet_vend_e",     
+                    "diet_vect_e",     
+                    "diet_vfish_e",     
+                    "diet_vunk_e",     
+                    "diet_scav_e",     
+                    "diet_fruit_e",     
+                    "diet_nect_e",     
+                    "diet_seed_e",     
+                    "diet_plant_e",     
+                    "diet_cat",     
+                    "diet_source_e",     
+                    "diet_certainty_e",     
+                    "diet_breadth",     
+                    "diet_level",     
+                    "for_strat_value_e",     
+                    "for_strat_certainty_e",     
+                    "activity_nocturnal_e",     
+                    "activity_crepuscular_e",     
+                    "activity_diurnal_e",     
+                    "activity_source_e",     
+                    "activity_certainty_e",     
+                    "activity_level",     
+                    "body_mass_e",     
+                    "body_mass_notes",     
+                    "body_mass_level_e",     
+                    "body_mass_source_e",     
+                    "min_body_size_mm",     
+                    "max_body_size_mm",     
+                    "body_size_mm",     
+                    "body_size_notes",     
+                    "body_size_level",     
+                    "body_size_source",     
+                    "sexual_dimorphism",     
+                    "sexual_dimorphism_notes",     
+                    "sexual_dimorphism_level",     
+                    "sexual_dimorphism_source",     
+                    "min_longevity",     
+                    "max_longevity",     
+                    "longevity",     
+                    "longevity_notes",     
+                    "longevity_level",     
+                    "longevity_source",     
+                    "min_home_range",     
+                    "max_home_range",     
+                    "home_range_size",     
+                    "home_range_notes",     
+                    "home_range_level",     
+                    "home_range_source",     
+                    "habitat_specialization",     
+                    "habitat_specialization_notes",     
+                    "habitat_specialization_source",     
+                    "min_generation_time",     
+                    "max_generation_time",     
+                    "generation_time",     
+                    "generation_time_notes",     
+                    "generation_time_level",     
+                    "generation_time_source",     
+                    "IUCN_category",     
+                    "habitat",     
+                    "habitat_suitability_montane",     
+                    "habitat_suitability_lowland",     
+                    "habitat_season_montane",     
+                    "habitat_season_lowland",     
+                    "habitat_major_importance_montane",     
+                    "habitat_major_importance_lowland",     
+                    "habitat_breadth",     
+                    "habitat_breadth_source",     
+                    "habitat_level",     
+                    "mean_CHELSA_bio1_1981.2010_V.2.1",     
+                    "mean_CHELSA_bio12_1981.2010_V.2.1",     
+                    "mean_human_fp_range_2010",     
+                    "mean_human_fp_range_2020",     
+                    "percent_change_hf_2010_2020",     
+                    "observed_range_sqkm",     
+                    "inferred_range_sqkm",     
+                    "date_data_obtained",     
+                    "filled_by",     
+                    "MSW05_order_p",     
+                    "MSW05_family_p",     
+                    "MSW05_genus_p",     
+                    "MSW05_species_p",     
+                    "PanTHERIA_level",     
+                    "activity_cycle_p",     
+                    "adult_body_mass_g_p",     
+                    "adult_forearm_len_mm_p",     
+                    "adult_head_body_len_mm_p",     
+                    "age_at_eye_opening_d_p",     
+                    "age_at_first_birth_d_p",     
+                    "basal_met_rate_mLO2hr_p",     
+                    "basal_met_rate_mass_g_p",     
+                    "diet_breadth_p",     
+                    "dispersal_age_d_p",     
+                    "gestation_len_d_p",     
+                    "habitat_breadth_p",     
+                    "home_range_km2_p",     
+                    "home_range_indiv_km2_p",     
+                    "Inter_birth_interval_d_p",     
+                    "litter_size_p",     
+                    "litters_per_year_p",     
+                    "max_longevity_m_p",     
+                    "neonate_body_mass_g_p",     
+                    "neonate_head_body_len_mm_p",     
+                    "population_density_n.km2_p",     
+                    "population_grp_size_p",     
+                    "sexual_maturity_age_d_p",     
+                    "social_grp_size_p",     
+                    "teat_number_p",     
+                    "terrestriality_p",     
+                    "trophic_level_p",     
+                    "weaning_age_d_p",     
+                    "weaning_body_mass_g_p",     
+                    "weaning_head_body_len_mm_p",     
+                    "references_p",     
+                    "adult_body_mass_g_EXT_p",     
+                    "litters_per_year_EXT_p",     
+                    "neonate_body_mass_g_EXT_p",     
+                    "weaning_body_mass_g_EXT_p",     
+                    "GR_area_km2_p",     
+                    "GR_max_lat_dd_p",     
+                    "GR_min_lat_dd_p",     
+                    "GR_mid_range_lat_dd_p",     
+                    "GR_max_long_dd_p",     
+                    "GR_min_long_dd_p",     
+                    "GR_mid_range_long_dd_p",     
+                    "hu_pop_den_min_n.km2_p_OD",     
+                    "hu_pop_den_mean_n.km2_p_OD",     
+                    "hu_pop_den_5p_n.km2_p_OD",     
+                    "hu_pop_den_change_p_OD",     
+                    "precip_mean_mm_p_OD",     
+                    "temp_mean_01degC_p_OD",     
+                    "AET_mean_mm_p_OD",     
+                    "PET_mean_mm_p_OD"    ), check.names=TRUE)
+
+unlink(infile1)
+
+# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
+
+if (class(dt1$IUCN_species_name)!="factor") dt1$IUCN_species_name<- as.factor(dt1$IUCN_species_name)
+if (class(dt1$taxonid)=="factor") dt1$taxonid <-as.numeric(levels(dt1$taxonid))[as.integer(dt1$taxonid) ]               
+if (class(dt1$taxonid)=="character") dt1$taxonid <-as.numeric(dt1$taxonid)
+if (class(dt1$common_name)!="factor") dt1$common_name<- as.factor(dt1$common_name)
+if (class(dt1$family)!="factor") dt1$family<- as.factor(dt1$family)
+if (class(dt1$genus)!="factor") dt1$genus<- as.factor(dt1$genus)
+if (class(dt1$species)!="factor") dt1$species<- as.factor(dt1$species)
+if (class(dt1$subspecies)!="factor") dt1$subspecies<- as.factor(dt1$subspecies)
+if (class(dt1$rank)!="factor") dt1$rank<- as.factor(dt1$rank)
+if (class(dt1$elton_species_name)!="factor") dt1$elton_species_name<- as.factor(dt1$elton_species_name)
+if (class(dt1$taxonomic_disparity)!="factor") dt1$taxonomic_disparity<- as.factor(dt1$taxonomic_disparity)
+if (class(dt1$species_notes)!="factor") dt1$species_notes<- as.factor(dt1$species_notes)
+if (class(dt1$diet_inv_e)=="factor") dt1$diet_inv_e <-as.numeric(levels(dt1$diet_inv_e))[as.integer(dt1$diet_inv_e) ]               
+if (class(dt1$diet_inv_e)=="character") dt1$diet_inv_e <-as.numeric(dt1$diet_inv_e)
+if (class(dt1$diet_vend_e)=="factor") dt1$diet_vend_e <-as.numeric(levels(dt1$diet_vend_e))[as.integer(dt1$diet_vend_e) ]               
+if (class(dt1$diet_vend_e)=="character") dt1$diet_vend_e <-as.numeric(dt1$diet_vend_e)
+if (class(dt1$diet_vect_e)=="factor") dt1$diet_vect_e <-as.numeric(levels(dt1$diet_vect_e))[as.integer(dt1$diet_vect_e) ]               
+if (class(dt1$diet_vect_e)=="character") dt1$diet_vect_e <-as.numeric(dt1$diet_vect_e)
+if (class(dt1$diet_vfish_e)=="factor") dt1$diet_vfish_e <-as.numeric(levels(dt1$diet_vfish_e))[as.integer(dt1$diet_vfish_e) ]               
+if (class(dt1$diet_vfish_e)=="character") dt1$diet_vfish_e <-as.numeric(dt1$diet_vfish_e)
+if (class(dt1$diet_vunk_e)=="factor") dt1$diet_vunk_e <-as.numeric(levels(dt1$diet_vunk_e))[as.integer(dt1$diet_vunk_e) ]               
+if (class(dt1$diet_vunk_e)=="character") dt1$diet_vunk_e <-as.numeric(dt1$diet_vunk_e)
+if (class(dt1$diet_scav_e)=="factor") dt1$diet_scav_e <-as.numeric(levels(dt1$diet_scav_e))[as.integer(dt1$diet_scav_e) ]               
+if (class(dt1$diet_scav_e)=="character") dt1$diet_scav_e <-as.numeric(dt1$diet_scav_e)
+if (class(dt1$diet_fruit_e)=="factor") dt1$diet_fruit_e <-as.numeric(levels(dt1$diet_fruit_e))[as.integer(dt1$diet_fruit_e) ]               
+if (class(dt1$diet_fruit_e)=="character") dt1$diet_fruit_e <-as.numeric(dt1$diet_fruit_e)
+if (class(dt1$diet_nect_e)=="factor") dt1$diet_nect_e <-as.numeric(levels(dt1$diet_nect_e))[as.integer(dt1$diet_nect_e) ]               
+if (class(dt1$diet_nect_e)=="character") dt1$diet_nect_e <-as.numeric(dt1$diet_nect_e)
+if (class(dt1$diet_seed_e)=="factor") dt1$diet_seed_e <-as.numeric(levels(dt1$diet_seed_e))[as.integer(dt1$diet_seed_e) ]               
+if (class(dt1$diet_seed_e)=="character") dt1$diet_seed_e <-as.numeric(dt1$diet_seed_e)
+if (class(dt1$diet_plant_e)=="factor") dt1$diet_plant_e <-as.numeric(levels(dt1$diet_plant_e))[as.integer(dt1$diet_plant_e) ]               
+if (class(dt1$diet_plant_e)=="character") dt1$diet_plant_e <-as.numeric(dt1$diet_plant_e)
+if (class(dt1$diet_cat)!="factor") dt1$diet_cat<- as.factor(dt1$diet_cat)
+if (class(dt1$diet_source_e)!="factor") dt1$diet_source_e<- as.factor(dt1$diet_source_e)
+if (class(dt1$diet_certainty_e)!="factor") dt1$diet_certainty_e<- as.factor(dt1$diet_certainty_e)
+if (class(dt1$diet_breadth)=="factor") dt1$diet_breadth <-as.numeric(levels(dt1$diet_breadth))[as.integer(dt1$diet_breadth) ]               
+if (class(dt1$diet_breadth)=="character") dt1$diet_breadth <-as.numeric(dt1$diet_breadth)
+if (class(dt1$diet_level)!="factor") dt1$diet_level<- as.factor(dt1$diet_level)
+if (class(dt1$for_strat_value_e)!="factor") dt1$for_strat_value_e<- as.factor(dt1$for_strat_value_e)
+if (class(dt1$for_strat_certainty_e)!="factor") dt1$for_strat_certainty_e<- as.factor(dt1$for_strat_certainty_e)
+if (class(dt1$activity_nocturnal_e)!="factor") dt1$activity_nocturnal_e<- as.factor(dt1$activity_nocturnal_e)
+if (class(dt1$activity_crepuscular_e)!="factor") dt1$activity_crepuscular_e<- as.factor(dt1$activity_crepuscular_e)
+if (class(dt1$activity_diurnal_e)!="factor") dt1$activity_diurnal_e<- as.factor(dt1$activity_diurnal_e)
+if (class(dt1$activity_source_e)!="factor") dt1$activity_source_e<- as.factor(dt1$activity_source_e)
+if (class(dt1$activity_certainty_e)!="factor") dt1$activity_certainty_e<- as.factor(dt1$activity_certainty_e)
+if (class(dt1$activity_level)!="factor") dt1$activity_level<- as.factor(dt1$activity_level)
+if (class(dt1$body_mass_e)=="factor") dt1$body_mass_e <-as.numeric(levels(dt1$body_mass_e))[as.integer(dt1$body_mass_e) ]               
+if (class(dt1$body_mass_e)=="character") dt1$body_mass_e <-as.numeric(dt1$body_mass_e)
+if (class(dt1$body_mass_notes)!="factor") dt1$body_mass_notes<- as.factor(dt1$body_mass_notes)
+if (class(dt1$body_mass_level_e)!="factor") dt1$body_mass_level_e<- as.factor(dt1$body_mass_level_e)
+if (class(dt1$body_mass_source_e)!="factor") dt1$body_mass_source_e<- as.factor(dt1$body_mass_source_e)
+if (class(dt1$min_body_size_mm)=="factor") dt1$min_body_size_mm <-as.numeric(levels(dt1$min_body_size_mm))[as.integer(dt1$min_body_size_mm) ]               
+if (class(dt1$min_body_size_mm)=="character") dt1$min_body_size_mm <-as.numeric(dt1$min_body_size_mm)
+if (class(dt1$max_body_size_mm)=="factor") dt1$max_body_size_mm <-as.numeric(levels(dt1$max_body_size_mm))[as.integer(dt1$max_body_size_mm) ]               
+if (class(dt1$max_body_size_mm)=="character") dt1$max_body_size_mm <-as.numeric(dt1$max_body_size_mm)
+if (class(dt1$body_size_mm)=="factor") dt1$body_size_mm <-as.numeric(levels(dt1$body_size_mm))[as.integer(dt1$body_size_mm) ]               
+if (class(dt1$body_size_mm)=="character") dt1$body_size_mm <-as.numeric(dt1$body_size_mm)
+if (class(dt1$body_size_notes)!="factor") dt1$body_size_notes<- as.factor(dt1$body_size_notes)
+if (class(dt1$body_size_level)!="factor") dt1$body_size_level<- as.factor(dt1$body_size_level)
+if (class(dt1$body_size_source)!="factor") dt1$body_size_source<- as.factor(dt1$body_size_source)
+if (class(dt1$sexual_dimorphism)!="factor") dt1$sexual_dimorphism<- as.factor(dt1$sexual_dimorphism)
+if (class(dt1$sexual_dimorphism_notes)!="factor") dt1$sexual_dimorphism_notes<- as.factor(dt1$sexual_dimorphism_notes)
+if (class(dt1$sexual_dimorphism_level)!="factor") dt1$sexual_dimorphism_level<- as.factor(dt1$sexual_dimorphism_level)
+if (class(dt1$sexual_dimorphism_source)!="factor") dt1$sexual_dimorphism_source<- as.factor(dt1$sexual_dimorphism_source)
+if (class(dt1$min_longevity)=="factor") dt1$min_longevity <-as.numeric(levels(dt1$min_longevity))[as.integer(dt1$min_longevity) ]               
+if (class(dt1$min_longevity)=="character") dt1$min_longevity <-as.numeric(dt1$min_longevity)
+if (class(dt1$max_longevity)=="factor") dt1$max_longevity <-as.numeric(levels(dt1$max_longevity))[as.integer(dt1$max_longevity) ]               
+if (class(dt1$max_longevity)=="character") dt1$max_longevity <-as.numeric(dt1$max_longevity)
+if (class(dt1$longevity)=="factor") dt1$longevity <-as.numeric(levels(dt1$longevity))[as.integer(dt1$longevity) ]               
+if (class(dt1$longevity)=="character") dt1$longevity <-as.numeric(dt1$longevity)
+if (class(dt1$longevity_notes)!="factor") dt1$longevity_notes<- as.factor(dt1$longevity_notes)
+if (class(dt1$longevity_level)!="factor") dt1$longevity_level<- as.factor(dt1$longevity_level)
+if (class(dt1$longevity_source)!="factor") dt1$longevity_source<- as.factor(dt1$longevity_source)
+if (class(dt1$min_home_range)=="factor") dt1$min_home_range <-as.numeric(levels(dt1$min_home_range))[as.integer(dt1$min_home_range) ]               
+if (class(dt1$min_home_range)=="character") dt1$min_home_range <-as.numeric(dt1$min_home_range)
+if (class(dt1$max_home_range)=="factor") dt1$max_home_range <-as.numeric(levels(dt1$max_home_range))[as.integer(dt1$max_home_range) ]               
+if (class(dt1$max_home_range)=="character") dt1$max_home_range <-as.numeric(dt1$max_home_range)
+if (class(dt1$home_range_size)=="factor") dt1$home_range_size <-as.numeric(levels(dt1$home_range_size))[as.integer(dt1$home_range_size) ]               
+if (class(dt1$home_range_size)=="character") dt1$home_range_size <-as.numeric(dt1$home_range_size)
+if (class(dt1$home_range_notes)!="factor") dt1$home_range_notes<- as.factor(dt1$home_range_notes)
+if (class(dt1$home_range_level)!="factor") dt1$home_range_level<- as.factor(dt1$home_range_level)
+if (class(dt1$home_range_source)!="factor") dt1$home_range_source<- as.factor(dt1$home_range_source)
+if (class(dt1$habitat_specialization)!="factor") dt1$habitat_specialization<- as.factor(dt1$habitat_specialization)
+if (class(dt1$habitat_specialization_notes)!="factor") dt1$habitat_specialization_notes<- as.factor(dt1$habitat_specialization_notes)
+if (class(dt1$habitat_specialization_source)!="factor") dt1$habitat_specialization_source<- as.factor(dt1$habitat_specialization_source)
+if (class(dt1$min_generation_time)=="factor") dt1$min_generation_time <-as.numeric(levels(dt1$min_generation_time))[as.integer(dt1$min_generation_time) ]               
+if (class(dt1$min_generation_time)=="character") dt1$min_generation_time <-as.numeric(dt1$min_generation_time)
+if (class(dt1$max_generation_time)=="factor") dt1$max_generation_time <-as.numeric(levels(dt1$max_generation_time))[as.integer(dt1$max_generation_time) ]               
+if (class(dt1$max_generation_time)=="character") dt1$max_generation_time <-as.numeric(dt1$max_generation_time)
+if (class(dt1$generation_time)=="factor") dt1$generation_time <-as.numeric(levels(dt1$generation_time))[as.integer(dt1$generation_time) ]               
+if (class(dt1$generation_time)=="character") dt1$generation_time <-as.numeric(dt1$generation_time)
+if (class(dt1$generation_time_notes)!="factor") dt1$generation_time_notes<- as.factor(dt1$generation_time_notes)
+if (class(dt1$generation_time_level)!="factor") dt1$generation_time_level<- as.factor(dt1$generation_time_level)
+if (class(dt1$generation_time_source)!="factor") dt1$generation_time_source<- as.factor(dt1$generation_time_source)
+if (class(dt1$IUCN_category)!="factor") dt1$IUCN_category<- as.factor(dt1$IUCN_category)
+if (class(dt1$habitat)!="factor") dt1$habitat<- as.factor(dt1$habitat)
+if (class(dt1$habitat_suitability_montane)!="factor") dt1$habitat_suitability_montane<- as.factor(dt1$habitat_suitability_montane)
+if (class(dt1$habitat_suitability_lowland)!="factor") dt1$habitat_suitability_lowland<- as.factor(dt1$habitat_suitability_lowland)
+if (class(dt1$habitat_season_montane)!="factor") dt1$habitat_season_montane<- as.factor(dt1$habitat_season_montane)
+if (class(dt1$habitat_season_lowland)!="factor") dt1$habitat_season_lowland<- as.factor(dt1$habitat_season_lowland)
+if (class(dt1$habitat_major_importance_montane)!="factor") dt1$habitat_major_importance_montane<- as.factor(dt1$habitat_major_importance_montane)
+if (class(dt1$habitat_major_importance_lowland)!="factor") dt1$habitat_major_importance_lowland<- as.factor(dt1$habitat_major_importance_lowland)
+if (class(dt1$habitat_breadth)=="factor") dt1$habitat_breadth <-as.numeric(levels(dt1$habitat_breadth))[as.integer(dt1$habitat_breadth) ]               
+if (class(dt1$habitat_breadth)=="character") dt1$habitat_breadth <-as.numeric(dt1$habitat_breadth)
+if (class(dt1$habitat_breadth_source)!="factor") dt1$habitat_breadth_source<- as.factor(dt1$habitat_breadth_source)
+if (class(dt1$habitat_level)!="factor") dt1$habitat_level<- as.factor(dt1$habitat_level)
+if (class(dt1$mean_CHELSA_bio1_1981.2010_V.2.1)=="factor") dt1$mean_CHELSA_bio1_1981.2010_V.2.1 <-as.numeric(levels(dt1$mean_CHELSA_bio1_1981.2010_V.2.1))[as.integer(dt1$mean_CHELSA_bio1_1981.2010_V.2.1) ]
+if (class(dt1$mean_CHELSA_bio1_1981.2010_V.2.1)=="character") dt1$mean_CHELSA_bio1_1981.2010_V.2.1 <-as.numeric(dt1$mean_CHELSA_bio1_1981.2010_V.2.1)
+if (class(dt1$mean_CHELSA_bio12_1981.2010_V.2.1)=="factor") dt1$mean_CHELSA_bio12_1981.2010_V.2.1 <-as.numeric(levels(dt1$mean_CHELSA_bio12_1981.2010_V.2.1))[as.integer(dt1$mean_CHELSA_bio12_1981.2010_V.2.1) ]
+if (class(dt1$mean_CHELSA_bio12_1981.2010_V.2.1)=="character") dt1$mean_CHELSA_bio12_1981.2010_V.2.1 <-as.numeric(dt1$mean_CHELSA_bio12_1981.2010_V.2.1)
+if (class(dt1$mean_human_fp_range_2010)=="factor") dt1$mean_human_fp_range_2010 <-as.numeric(levels(dt1$mean_human_fp_range_2010))[as.integer(dt1$mean_human_fp_range_2010) ]               
+if (class(dt1$mean_human_fp_range_2010)=="character") dt1$mean_human_fp_range_2010 <-as.numeric(dt1$mean_human_fp_range_2010)
+if (class(dt1$mean_human_fp_range_2020)=="factor") dt1$mean_human_fp_range_2020 <-as.numeric(levels(dt1$mean_human_fp_range_2020))[as.integer(dt1$mean_human_fp_range_2020) ]               
+if (class(dt1$mean_human_fp_range_2020)=="character") dt1$mean_human_fp_range_2020 <-as.numeric(dt1$mean_human_fp_range_2020)
+if (class(dt1$percent_change_hf_2010_2020)=="factor") dt1$percent_change_hf_2010_2020 <-as.numeric(levels(dt1$percent_change_hf_2010_2020))[as.integer(dt1$percent_change_hf_2010_2020) ]               
+if (class(dt1$percent_change_hf_2010_2020)=="character") dt1$percent_change_hf_2010_2020 <-as.numeric(dt1$percent_change_hf_2010_2020)
+if (class(dt1$observed_range_sqkm)=="factor") dt1$observed_range_sqkm <-as.numeric(levels(dt1$observed_range_sqkm))[as.integer(dt1$observed_range_sqkm) ]               
+if (class(dt1$observed_range_sqkm)=="character") dt1$observed_range_sqkm <-as.numeric(dt1$observed_range_sqkm)
+if (class(dt1$inferred_range_sqkm)=="factor") dt1$inferred_range_sqkm <-as.numeric(levels(dt1$inferred_range_sqkm))[as.integer(dt1$inferred_range_sqkm) ]               
+if (class(dt1$inferred_range_sqkm)=="character") dt1$inferred_range_sqkm <-as.numeric(dt1$inferred_range_sqkm)
+if (class(dt1$date_data_obtained)!="factor") dt1$date_data_obtained<- as.factor(dt1$date_data_obtained)
+if (class(dt1$filled_by)!="factor") dt1$filled_by<- as.factor(dt1$filled_by)
+if (class(dt1$MSW05_order_p)!="factor") dt1$MSW05_order_p<- as.factor(dt1$MSW05_order_p)
+if (class(dt1$MSW05_family_p)!="factor") dt1$MSW05_family_p<- as.factor(dt1$MSW05_family_p)
+if (class(dt1$MSW05_genus_p)!="factor") dt1$MSW05_genus_p<- as.factor(dt1$MSW05_genus_p)
+if (class(dt1$MSW05_species_p)!="factor") dt1$MSW05_species_p<- as.factor(dt1$MSW05_species_p)
+if (class(dt1$PanTHERIA_level)!="factor") dt1$PanTHERIA_level<- as.factor(dt1$PanTHERIA_level)
+if (class(dt1$activity_cycle_p)!="factor") dt1$activity_cycle_p<- as.factor(dt1$activity_cycle_p)
+if (class(dt1$adult_body_mass_g_p)=="factor") dt1$adult_body_mass_g_p <-as.numeric(levels(dt1$adult_body_mass_g_p))[as.integer(dt1$adult_body_mass_g_p) ]               
+if (class(dt1$adult_body_mass_g_p)=="character") dt1$adult_body_mass_g_p <-as.numeric(dt1$adult_body_mass_g_p)
+if (class(dt1$adult_forearm_len_mm_p)=="factor") dt1$adult_forearm_len_mm_p <-as.numeric(levels(dt1$adult_forearm_len_mm_p))[as.integer(dt1$adult_forearm_len_mm_p) ]               
+if (class(dt1$adult_forearm_len_mm_p)=="character") dt1$adult_forearm_len_mm_p <-as.numeric(dt1$adult_forearm_len_mm_p)
+if (class(dt1$adult_head_body_len_mm_p)=="factor") dt1$adult_head_body_len_mm_p <-as.numeric(levels(dt1$adult_head_body_len_mm_p))[as.integer(dt1$adult_head_body_len_mm_p) ]               
+if (class(dt1$adult_head_body_len_mm_p)=="character") dt1$adult_head_body_len_mm_p <-as.numeric(dt1$adult_head_body_len_mm_p)
+if (class(dt1$age_at_eye_opening_d_p)=="factor") dt1$age_at_eye_opening_d_p <-as.numeric(levels(dt1$age_at_eye_opening_d_p))[as.integer(dt1$age_at_eye_opening_d_p) ]               
+if (class(dt1$age_at_eye_opening_d_p)=="character") dt1$age_at_eye_opening_d_p <-as.numeric(dt1$age_at_eye_opening_d_p)
+if (class(dt1$age_at_first_birth_d_p)=="factor") dt1$age_at_first_birth_d_p <-as.numeric(levels(dt1$age_at_first_birth_d_p))[as.integer(dt1$age_at_first_birth_d_p) ]               
+if (class(dt1$age_at_first_birth_d_p)=="character") dt1$age_at_first_birth_d_p <-as.numeric(dt1$age_at_first_birth_d_p)
+if (class(dt1$basal_met_rate_mLO2hr_p)=="factor") dt1$basal_met_rate_mLO2hr_p <-as.numeric(levels(dt1$basal_met_rate_mLO2hr_p))[as.integer(dt1$basal_met_rate_mLO2hr_p) ]               
+if (class(dt1$basal_met_rate_mLO2hr_p)=="character") dt1$basal_met_rate_mLO2hr_p <-as.numeric(dt1$basal_met_rate_mLO2hr_p)
+if (class(dt1$basal_met_rate_mass_g_p)=="factor") dt1$basal_met_rate_mass_g_p <-as.numeric(levels(dt1$basal_met_rate_mass_g_p))[as.integer(dt1$basal_met_rate_mass_g_p) ]               
+if (class(dt1$basal_met_rate_mass_g_p)=="character") dt1$basal_met_rate_mass_g_p <-as.numeric(dt1$basal_met_rate_mass_g_p)
+if (class(dt1$diet_breadth_p)=="factor") dt1$diet_breadth_p <-as.numeric(levels(dt1$diet_breadth_p))[as.integer(dt1$diet_breadth_p) ]               
+if (class(dt1$diet_breadth_p)=="character") dt1$diet_breadth_p <-as.numeric(dt1$diet_breadth_p)
+if (class(dt1$dispersal_age_d_p)=="factor") dt1$dispersal_age_d_p <-as.numeric(levels(dt1$dispersal_age_d_p))[as.integer(dt1$dispersal_age_d_p) ]               
+if (class(dt1$dispersal_age_d_p)=="character") dt1$dispersal_age_d_p <-as.numeric(dt1$dispersal_age_d_p)
+if (class(dt1$gestation_len_d_p)=="factor") dt1$gestation_len_d_p <-as.numeric(levels(dt1$gestation_len_d_p))[as.integer(dt1$gestation_len_d_p) ]               
+if (class(dt1$gestation_len_d_p)=="character") dt1$gestation_len_d_p <-as.numeric(dt1$gestation_len_d_p)
+if (class(dt1$habitat_breadth_p)=="factor") dt1$habitat_breadth_p <-as.numeric(levels(dt1$habitat_breadth_p))[as.integer(dt1$habitat_breadth_p) ]               
+if (class(dt1$habitat_breadth_p)=="character") dt1$habitat_breadth_p <-as.numeric(dt1$habitat_breadth_p)
+if (class(dt1$home_range_km2_p)=="factor") dt1$home_range_km2_p <-as.numeric(levels(dt1$home_range_km2_p))[as.integer(dt1$home_range_km2_p) ]               
+if (class(dt1$home_range_km2_p)=="character") dt1$home_range_km2_p <-as.numeric(dt1$home_range_km2_p)
+if (class(dt1$home_range_indiv_km2_p)=="factor") dt1$home_range_indiv_km2_p <-as.numeric(levels(dt1$home_range_indiv_km2_p))[as.integer(dt1$home_range_indiv_km2_p) ]               
+if (class(dt1$home_range_indiv_km2_p)=="character") dt1$home_range_indiv_km2_p <-as.numeric(dt1$home_range_indiv_km2_p)
+if (class(dt1$Inter_birth_interval_d_p)=="factor") dt1$Inter_birth_interval_d_p <-as.numeric(levels(dt1$Inter_birth_interval_d_p))[as.integer(dt1$Inter_birth_interval_d_p) ]               
+if (class(dt1$Inter_birth_interval_d_p)=="character") dt1$Inter_birth_interval_d_p <-as.numeric(dt1$Inter_birth_interval_d_p)
+if (class(dt1$litter_size_p)=="factor") dt1$litter_size_p <-as.numeric(levels(dt1$litter_size_p))[as.integer(dt1$litter_size_p) ]               
+if (class(dt1$litter_size_p)=="character") dt1$litter_size_p <-as.numeric(dt1$litter_size_p)
+if (class(dt1$litters_per_year_p)=="factor") dt1$litters_per_year_p <-as.numeric(levels(dt1$litters_per_year_p))[as.integer(dt1$litters_per_year_p) ]               
+if (class(dt1$litters_per_year_p)=="character") dt1$litters_per_year_p <-as.numeric(dt1$litters_per_year_p)
+if (class(dt1$max_longevity_m_p)=="factor") dt1$max_longevity_m_p <-as.numeric(levels(dt1$max_longevity_m_p))[as.integer(dt1$max_longevity_m_p) ]               
+if (class(dt1$max_longevity_m_p)=="character") dt1$max_longevity_m_p <-as.numeric(dt1$max_longevity_m_p)
+if (class(dt1$neonate_body_mass_g_p)=="factor") dt1$neonate_body_mass_g_p <-as.numeric(levels(dt1$neonate_body_mass_g_p))[as.integer(dt1$neonate_body_mass_g_p) ]               
+if (class(dt1$neonate_body_mass_g_p)=="character") dt1$neonate_body_mass_g_p <-as.numeric(dt1$neonate_body_mass_g_p)
+if (class(dt1$neonate_head_body_len_mm_p)=="factor") dt1$neonate_head_body_len_mm_p <-as.numeric(levels(dt1$neonate_head_body_len_mm_p))[as.integer(dt1$neonate_head_body_len_mm_p) ]               
+if (class(dt1$neonate_head_body_len_mm_p)=="character") dt1$neonate_head_body_len_mm_p <-as.numeric(dt1$neonate_head_body_len_mm_p)
+if (class(dt1$population_density_n.km2_p)=="factor") dt1$population_density_n.km2_p <-as.numeric(levels(dt1$population_density_n.km2_p))[as.integer(dt1$population_density_n.km2_p) ]               
+if (class(dt1$population_density_n.km2_p)=="character") dt1$population_density_n.km2_p <-as.numeric(dt1$population_density_n.km2_p)
+if (class(dt1$population_grp_size_p)=="factor") dt1$population_grp_size_p <-as.numeric(levels(dt1$population_grp_size_p))[as.integer(dt1$population_grp_size_p) ]               
+if (class(dt1$population_grp_size_p)=="character") dt1$population_grp_size_p <-as.numeric(dt1$population_grp_size_p)
+if (class(dt1$sexual_maturity_age_d_p)=="factor") dt1$sexual_maturity_age_d_p <-as.numeric(levels(dt1$sexual_maturity_age_d_p))[as.integer(dt1$sexual_maturity_age_d_p) ]               
+if (class(dt1$sexual_maturity_age_d_p)=="character") dt1$sexual_maturity_age_d_p <-as.numeric(dt1$sexual_maturity_age_d_p)
+if (class(dt1$social_grp_size_p)=="factor") dt1$social_grp_size_p <-as.numeric(levels(dt1$social_grp_size_p))[as.integer(dt1$social_grp_size_p) ]               
+if (class(dt1$social_grp_size_p)=="character") dt1$social_grp_size_p <-as.numeric(dt1$social_grp_size_p)
+if (class(dt1$teat_number_p)=="factor") dt1$teat_number_p <-as.numeric(levels(dt1$teat_number_p))[as.integer(dt1$teat_number_p) ]               
+if (class(dt1$teat_number_p)=="character") dt1$teat_number_p <-as.numeric(dt1$teat_number_p)
+if (class(dt1$terrestriality_p)!="factor") dt1$terrestriality_p<- as.factor(dt1$terrestriality_p)
+if (class(dt1$trophic_level_p)!="factor") dt1$trophic_level_p<- as.factor(dt1$trophic_level_p)
+if (class(dt1$weaning_age_d_p)=="factor") dt1$weaning_age_d_p <-as.numeric(levels(dt1$weaning_age_d_p))[as.integer(dt1$weaning_age_d_p) ]               
+if (class(dt1$weaning_age_d_p)=="character") dt1$weaning_age_d_p <-as.numeric(dt1$weaning_age_d_p)
+if (class(dt1$weaning_body_mass_g_p)=="factor") dt1$weaning_body_mass_g_p <-as.numeric(levels(dt1$weaning_body_mass_g_p))[as.integer(dt1$weaning_body_mass_g_p) ]               
+if (class(dt1$weaning_body_mass_g_p)=="character") dt1$weaning_body_mass_g_p <-as.numeric(dt1$weaning_body_mass_g_p)
+if (class(dt1$weaning_head_body_len_mm_p)=="factor") dt1$weaning_head_body_len_mm_p <-as.numeric(levels(dt1$weaning_head_body_len_mm_p))[as.integer(dt1$weaning_head_body_len_mm_p) ]               
+if (class(dt1$weaning_head_body_len_mm_p)=="character") dt1$weaning_head_body_len_mm_p <-as.numeric(dt1$weaning_head_body_len_mm_p)
+if (class(dt1$references_p)!="factor") dt1$references_p<- as.factor(dt1$references_p)
+if (class(dt1$adult_body_mass_g_EXT_p)=="factor") dt1$adult_body_mass_g_EXT_p <-as.numeric(levels(dt1$adult_body_mass_g_EXT_p))[as.integer(dt1$adult_body_mass_g_EXT_p) ]               
+if (class(dt1$adult_body_mass_g_EXT_p)=="character") dt1$adult_body_mass_g_EXT_p <-as.numeric(dt1$adult_body_mass_g_EXT_p)
+if (class(dt1$litters_per_year_EXT_p)=="factor") dt1$litters_per_year_EXT_p <-as.numeric(levels(dt1$litters_per_year_EXT_p))[as.integer(dt1$litters_per_year_EXT_p) ]               
+if (class(dt1$litters_per_year_EXT_p)=="character") dt1$litters_per_year_EXT_p <-as.numeric(dt1$litters_per_year_EXT_p)
+if (class(dt1$neonate_body_mass_g_EXT_p)=="factor") dt1$neonate_body_mass_g_EXT_p <-as.numeric(levels(dt1$neonate_body_mass_g_EXT_p))[as.integer(dt1$neonate_body_mass_g_EXT_p) ]               
+if (class(dt1$neonate_body_mass_g_EXT_p)=="character") dt1$neonate_body_mass_g_EXT_p <-as.numeric(dt1$neonate_body_mass_g_EXT_p)
+if (class(dt1$weaning_body_mass_g_EXT_p)=="factor") dt1$weaning_body_mass_g_EXT_p <-as.numeric(levels(dt1$weaning_body_mass_g_EXT_p))[as.integer(dt1$weaning_body_mass_g_EXT_p) ]               
+if (class(dt1$weaning_body_mass_g_EXT_p)=="character") dt1$weaning_body_mass_g_EXT_p <-as.numeric(dt1$weaning_body_mass_g_EXT_p)
+if (class(dt1$GR_area_km2_p)=="factor") dt1$GR_area_km2_p <-as.numeric(levels(dt1$GR_area_km2_p))[as.integer(dt1$GR_area_km2_p) ]               
+if (class(dt1$GR_area_km2_p)=="character") dt1$GR_area_km2_p <-as.numeric(dt1$GR_area_km2_p)
+if (class(dt1$GR_max_lat_dd_p)=="factor") dt1$GR_max_lat_dd_p <-as.numeric(levels(dt1$GR_max_lat_dd_p))[as.integer(dt1$GR_max_lat_dd_p) ]               
+if (class(dt1$GR_max_lat_dd_p)=="character") dt1$GR_max_lat_dd_p <-as.numeric(dt1$GR_max_lat_dd_p)
+if (class(dt1$GR_min_lat_dd_p)=="factor") dt1$GR_min_lat_dd_p <-as.numeric(levels(dt1$GR_min_lat_dd_p))[as.integer(dt1$GR_min_lat_dd_p) ]               
+if (class(dt1$GR_min_lat_dd_p)=="character") dt1$GR_min_lat_dd_p <-as.numeric(dt1$GR_min_lat_dd_p)
+if (class(dt1$GR_mid_range_lat_dd_p)=="factor") dt1$GR_mid_range_lat_dd_p <-as.numeric(levels(dt1$GR_mid_range_lat_dd_p))[as.integer(dt1$GR_mid_range_lat_dd_p) ]               
+if (class(dt1$GR_mid_range_lat_dd_p)=="character") dt1$GR_mid_range_lat_dd_p <-as.numeric(dt1$GR_mid_range_lat_dd_p)
+if (class(dt1$GR_max_long_dd_p)=="factor") dt1$GR_max_long_dd_p <-as.numeric(levels(dt1$GR_max_long_dd_p))[as.integer(dt1$GR_max_long_dd_p) ]               
+if (class(dt1$GR_max_long_dd_p)=="character") dt1$GR_max_long_dd_p <-as.numeric(dt1$GR_max_long_dd_p)
+if (class(dt1$GR_min_long_dd_p)=="factor") dt1$GR_min_long_dd_p <-as.numeric(levels(dt1$GR_min_long_dd_p))[as.integer(dt1$GR_min_long_dd_p) ]               
+if (class(dt1$GR_min_long_dd_p)=="character") dt1$GR_min_long_dd_p <-as.numeric(dt1$GR_min_long_dd_p)
+if (class(dt1$GR_mid_range_long_dd_p)=="factor") dt1$GR_mid_range_long_dd_p <-as.numeric(levels(dt1$GR_mid_range_long_dd_p))[as.integer(dt1$GR_mid_range_long_dd_p) ]               
+if (class(dt1$GR_mid_range_long_dd_p)=="character") dt1$GR_mid_range_long_dd_p <-as.numeric(dt1$GR_mid_range_long_dd_p)
+if (class(dt1$hu_pop_den_min_n.km2_p_OD)=="factor") dt1$hu_pop_den_min_n.km2_p_OD <-as.numeric(levels(dt1$hu_pop_den_min_n.km2_p_OD))[as.integer(dt1$hu_pop_den_min_n.km2_p_OD) ]               
+if (class(dt1$hu_pop_den_min_n.km2_p_OD)=="character") dt1$hu_pop_den_min_n.km2_p_OD <-as.numeric(dt1$hu_pop_den_min_n.km2_p_OD)
+if (class(dt1$hu_pop_den_mean_n.km2_p_OD)=="factor") dt1$hu_pop_den_mean_n.km2_p_OD <-as.numeric(levels(dt1$hu_pop_den_mean_n.km2_p_OD))[as.integer(dt1$hu_pop_den_mean_n.km2_p_OD) ]               
+if (class(dt1$hu_pop_den_mean_n.km2_p_OD)=="character") dt1$hu_pop_den_mean_n.km2_p_OD <-as.numeric(dt1$hu_pop_den_mean_n.km2_p_OD)
+if (class(dt1$hu_pop_den_5p_n.km2_p_OD)=="factor") dt1$hu_pop_den_5p_n.km2_p_OD <-as.numeric(levels(dt1$hu_pop_den_5p_n.km2_p_OD))[as.integer(dt1$hu_pop_den_5p_n.km2_p_OD) ]               
+if (class(dt1$hu_pop_den_5p_n.km2_p_OD)=="character") dt1$hu_pop_den_5p_n.km2_p_OD <-as.numeric(dt1$hu_pop_den_5p_n.km2_p_OD)
+if (class(dt1$hu_pop_den_change_p_OD)=="factor") dt1$hu_pop_den_change_p_OD <-as.numeric(levels(dt1$hu_pop_den_change_p_OD))[as.integer(dt1$hu_pop_den_change_p_OD) ]               
+if (class(dt1$hu_pop_den_change_p_OD)=="character") dt1$hu_pop_den_change_p_OD <-as.numeric(dt1$hu_pop_den_change_p_OD)
+if (class(dt1$precip_mean_mm_p_OD)=="factor") dt1$precip_mean_mm_p_OD <-as.numeric(levels(dt1$precip_mean_mm_p_OD))[as.integer(dt1$precip_mean_mm_p_OD) ]               
+if (class(dt1$precip_mean_mm_p_OD)=="character") dt1$precip_mean_mm_p_OD <-as.numeric(dt1$precip_mean_mm_p_OD)
+if (class(dt1$temp_mean_01degC_p_OD)=="factor") dt1$temp_mean_01degC_p_OD <-as.numeric(levels(dt1$temp_mean_01degC_p_OD))[as.integer(dt1$temp_mean_01degC_p_OD) ]               
+if (class(dt1$temp_mean_01degC_p_OD)=="character") dt1$temp_mean_01degC_p_OD <-as.numeric(dt1$temp_mean_01degC_p_OD)
+if (class(dt1$AET_mean_mm_p_OD)=="factor") dt1$AET_mean_mm_p_OD <-as.numeric(levels(dt1$AET_mean_mm_p_OD))[as.integer(dt1$AET_mean_mm_p_OD) ]               
+if (class(dt1$AET_mean_mm_p_OD)=="character") dt1$AET_mean_mm_p_OD <-as.numeric(dt1$AET_mean_mm_p_OD)
+if (class(dt1$PET_mean_mm_p_OD)=="factor") dt1$PET_mean_mm_p_OD <-as.numeric(levels(dt1$PET_mean_mm_p_OD))[as.integer(dt1$PET_mean_mm_p_OD) ]               
+if (class(dt1$PET_mean_mm_p_OD)=="character") dt1$PET_mean_mm_p_OD <-as.numeric(dt1$PET_mean_mm_p_OD)
+
+# Convert Missing Values to NA for non-dates
+
+dt1$common_name <- as.factor(ifelse((trimws(as.character(dt1$common_name))==trimws("NA")),NA,as.character(dt1$common_name)))
+dt1$subspecies <- as.factor(ifelse((trimws(as.character(dt1$subspecies))==trimws("NA")),NA,as.character(dt1$subspecies)))
+dt1$rank <- as.factor(ifelse((trimws(as.character(dt1$rank))==trimws("NA")),NA,as.character(dt1$rank)))
+dt1$species_notes <- as.factor(ifelse((trimws(as.character(dt1$species_notes))==trimws("NA")),NA,as.character(dt1$species_notes)))
+dt1$diet_certainty_e <- as.factor(ifelse((trimws(as.character(dt1$diet_certainty_e))==trimws("NA")),NA,as.character(dt1$diet_certainty_e)))
+dt1$diet_level <- as.factor(ifelse((trimws(as.character(dt1$diet_level))==trimws("NA")),NA,as.character(dt1$diet_level)))
+dt1$activity_level <- as.factor(ifelse((trimws(as.character(dt1$activity_level))==trimws("NA")),NA,as.character(dt1$activity_level)))
+dt1$body_mass_notes <- as.factor(ifelse((trimws(as.character(dt1$body_mass_notes))==trimws("NA")),NA,as.character(dt1$body_mass_notes)))
+dt1$min_body_size_mm <- ifelse((trimws(as.character(dt1$min_body_size_mm))==trimws("NA")),NA,dt1$min_body_size_mm)               
+suppressWarnings(dt1$min_body_size_mm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$min_body_size_mm))==as.character(as.numeric("NA"))),NA,dt1$min_body_size_mm))
+dt1$max_body_size_mm <- ifelse((trimws(as.character(dt1$max_body_size_mm))==trimws("NA")),NA,dt1$max_body_size_mm)               
+suppressWarnings(dt1$max_body_size_mm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$max_body_size_mm))==as.character(as.numeric("NA"))),NA,dt1$max_body_size_mm))
+dt1$body_size_mm <- ifelse((trimws(as.character(dt1$body_size_mm))==trimws("NA")),NA,dt1$body_size_mm)               
+suppressWarnings(dt1$body_size_mm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$body_size_mm))==as.character(as.numeric("NA"))),NA,dt1$body_size_mm))
+dt1$body_size_notes <- as.factor(ifelse((trimws(as.character(dt1$body_size_notes))==trimws("NA")),NA,as.character(dt1$body_size_notes)))
+dt1$body_size_level <- as.factor(ifelse((trimws(as.character(dt1$body_size_level))==trimws("NA")),NA,as.character(dt1$body_size_level)))
+dt1$body_size_source <- as.factor(ifelse((trimws(as.character(dt1$body_size_source))==trimws("NA")),NA,as.character(dt1$body_size_source)))
+dt1$sexual_dimorphism <- as.factor(ifelse((trimws(as.character(dt1$sexual_dimorphism))==trimws("NA")),NA,as.character(dt1$sexual_dimorphism)))
+dt1$sexual_dimorphism <- as.factor(ifelse((trimws(as.character(dt1$sexual_dimorphism))==trimws("NA ")),NA,as.character(dt1$sexual_dimorphism)))
+dt1$sexual_dimorphism_notes <- as.factor(ifelse((trimws(as.character(dt1$sexual_dimorphism_notes))==trimws("NA")),NA,as.character(dt1$sexual_dimorphism_notes)))
+dt1$sexual_dimorphism_level <- as.factor(ifelse((trimws(as.character(dt1$sexual_dimorphism_level))==trimws("NA")),NA,as.character(dt1$sexual_dimorphism_level)))
+dt1$sexual_dimorphism_source <- as.factor(ifelse((trimws(as.character(dt1$sexual_dimorphism_source))==trimws("NA")),NA,as.character(dt1$sexual_dimorphism_source)))
+dt1$min_longevity <- ifelse((trimws(as.character(dt1$min_longevity))==trimws("NA")),NA,dt1$min_longevity)               
+suppressWarnings(dt1$min_longevity <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$min_longevity))==as.character(as.numeric("NA"))),NA,dt1$min_longevity))
+dt1$max_longevity <- ifelse((trimws(as.character(dt1$max_longevity))==trimws("NA")),NA,dt1$max_longevity)               
+suppressWarnings(dt1$max_longevity <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$max_longevity))==as.character(as.numeric("NA"))),NA,dt1$max_longevity))
+dt1$longevity <- ifelse((trimws(as.character(dt1$longevity))==trimws("NA")),NA,dt1$longevity)               
+suppressWarnings(dt1$longevity <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$longevity))==as.character(as.numeric("NA"))),NA,dt1$longevity))
+dt1$longevity_notes <- as.factor(ifelse((trimws(as.character(dt1$longevity_notes))==trimws("NA")),NA,as.character(dt1$longevity_notes)))
+dt1$longevity_level <- as.factor(ifelse((trimws(as.character(dt1$longevity_level))==trimws("NA")),NA,as.character(dt1$longevity_level)))
+dt1$longevity_source <- as.factor(ifelse((trimws(as.character(dt1$longevity_source))==trimws("NA")),NA,as.character(dt1$longevity_source)))
+dt1$min_home_range <- ifelse((trimws(as.character(dt1$min_home_range))==trimws("NA")),NA,dt1$min_home_range)               
+suppressWarnings(dt1$min_home_range <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$min_home_range))==as.character(as.numeric("NA"))),NA,dt1$min_home_range))
+dt1$max_home_range <- ifelse((trimws(as.character(dt1$max_home_range))==trimws("NA")),NA,dt1$max_home_range)               
+suppressWarnings(dt1$max_home_range <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$max_home_range))==as.character(as.numeric("NA"))),NA,dt1$max_home_range))
+dt1$home_range_size <- ifelse((trimws(as.character(dt1$home_range_size))==trimws("NA")),NA,dt1$home_range_size)               
+suppressWarnings(dt1$home_range_size <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$home_range_size))==as.character(as.numeric("NA"))),NA,dt1$home_range_size))
+dt1$home_range_notes <- as.factor(ifelse((trimws(as.character(dt1$home_range_notes))==trimws("NA")),NA,as.character(dt1$home_range_notes)))
+dt1$home_range_level <- as.factor(ifelse((trimws(as.character(dt1$home_range_level))==trimws("NA")),NA,as.character(dt1$home_range_level)))
+dt1$home_range_source <- as.factor(ifelse((trimws(as.character(dt1$home_range_source))==trimws("NA")),NA,as.character(dt1$home_range_source)))
+dt1$habitat_specialization <- as.factor(ifelse((trimws(as.character(dt1$habitat_specialization))==trimws("NA")),NA,as.character(dt1$habitat_specialization)))
+dt1$habitat_specialization_notes <- as.factor(ifelse((trimws(as.character(dt1$habitat_specialization_notes))==trimws("NA")),NA,as.character(dt1$habitat_specialization_notes)))
+dt1$habitat_specialization_source <- as.factor(ifelse((trimws(as.character(dt1$habitat_specialization_source))==trimws("NA")),NA,as.character(dt1$habitat_specialization_source)))
+dt1$min_generation_time <- ifelse((trimws(as.character(dt1$min_generation_time))==trimws("NA")),NA,dt1$min_generation_time)               
+suppressWarnings(dt1$min_generation_time <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$min_generation_time))==as.character(as.numeric("NA"))),NA,dt1$min_generation_time))
+dt1$max_generation_time <- ifelse((trimws(as.character(dt1$max_generation_time))==trimws("NA")),NA,dt1$max_generation_time)               
+suppressWarnings(dt1$max_generation_time <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$max_generation_time))==as.character(as.numeric("NA"))),NA,dt1$max_generation_time))
+dt1$generation_time <- ifelse((trimws(as.character(dt1$generation_time))==trimws("NA")),NA,dt1$generation_time)               
+suppressWarnings(dt1$generation_time <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$generation_time))==as.character(as.numeric("NA"))),NA,dt1$generation_time))
+dt1$generation_time_notes <- as.factor(ifelse((trimws(as.character(dt1$generation_time_notes))==trimws("NA")),NA,as.character(dt1$generation_time_notes)))
+dt1$generation_time_level <- as.factor(ifelse((trimws(as.character(dt1$generation_time_level))==trimws("NA")),NA,as.character(dt1$generation_time_level)))
+dt1$generation_time_source <- as.factor(ifelse((trimws(as.character(dt1$generation_time_source))==trimws("NA")),NA,as.character(dt1$generation_time_source)))
+dt1$habitat_suitability_montane <- as.factor(ifelse((trimws(as.character(dt1$habitat_suitability_montane))==trimws("NA")),NA,as.character(dt1$habitat_suitability_montane)))
+dt1$habitat_suitability_lowland <- as.factor(ifelse((trimws(as.character(dt1$habitat_suitability_lowland))==trimws("NA")),NA,as.character(dt1$habitat_suitability_lowland)))
+dt1$habitat_season_montane <- as.factor(ifelse((trimws(as.character(dt1$habitat_season_montane))==trimws("NA")),NA,as.character(dt1$habitat_season_montane)))
+dt1$habitat_season_lowland <- as.factor(ifelse((trimws(as.character(dt1$habitat_season_lowland))==trimws("NA")),NA,as.character(dt1$habitat_season_lowland)))
+dt1$habitat_major_importance_montane <- as.factor(ifelse((trimws(as.character(dt1$habitat_major_importance_montane))==trimws("NA")),NA,as.character(dt1$habitat_major_importance_montane)))
+dt1$habitat_major_importance_lowland <- as.factor(ifelse((trimws(as.character(dt1$habitat_major_importance_lowland))==trimws("NA")),NA,as.character(dt1$habitat_major_importance_lowland)))
+dt1$habitat_breadth <- ifelse((trimws(as.character(dt1$habitat_breadth))==trimws("NA")),NA,dt1$habitat_breadth)               
+suppressWarnings(dt1$habitat_breadth <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$habitat_breadth))==as.character(as.numeric("NA"))),NA,dt1$habitat_breadth))
+dt1$habitat_level <- as.factor(ifelse((trimws(as.character(dt1$habitat_level))==trimws("NA")),NA,as.character(dt1$habitat_level)))
+dt1$mean_CHELSA_bio1_1981.2010_V.2.1 <- ifelse((trimws(as.character(dt1$mean_CHELSA_bio1_1981.2010_V.2.1))==trimws("NA")),NA,dt1$mean_CHELSA_bio1_1981.2010_V.2.1)               
+suppressWarnings(dt1$mean_CHELSA_bio1_1981.2010_V.2.1 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$mean_CHELSA_bio1_1981.2010_V.2.1))==as.character(as.numeric("NA"))),NA,dt1$mean_CHELSA_bio1_1981.2010_V.2.1))
+dt1$mean_CHELSA_bio12_1981.2010_V.2.1 <- ifelse((trimws(as.character(dt1$mean_CHELSA_bio12_1981.2010_V.2.1))==trimws("NA")),NA,dt1$mean_CHELSA_bio12_1981.2010_V.2.1)               
+suppressWarnings(dt1$mean_CHELSA_bio12_1981.2010_V.2.1 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$mean_CHELSA_bio12_1981.2010_V.2.1))==as.character(as.numeric("NA"))),NA,dt1$mean_CHELSA_bio12_1981.2010_V.2.1))
+dt1$mean_human_fp_range_2010 <- ifelse((trimws(as.character(dt1$mean_human_fp_range_2010))==trimws("NA")),NA,dt1$mean_human_fp_range_2010)               
+suppressWarnings(dt1$mean_human_fp_range_2010 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$mean_human_fp_range_2010))==as.character(as.numeric("NA"))),NA,dt1$mean_human_fp_range_2010))
+dt1$mean_human_fp_range_2020 <- ifelse((trimws(as.character(dt1$mean_human_fp_range_2020))==trimws("NA")),NA,dt1$mean_human_fp_range_2020)               
+suppressWarnings(dt1$mean_human_fp_range_2020 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$mean_human_fp_range_2020))==as.character(as.numeric("NA"))),NA,dt1$mean_human_fp_range_2020))
+dt1$percent_change_hf_2010_2020 <- ifelse((trimws(as.character(dt1$percent_change_hf_2010_2020))==trimws("NA")),NA,dt1$percent_change_hf_2010_2020)               
+suppressWarnings(dt1$percent_change_hf_2010_2020 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$percent_change_hf_2010_2020))==as.character(as.numeric("NA"))),NA,dt1$percent_change_hf_2010_2020))
+dt1$observed_range_sqkm <- ifelse((trimws(as.character(dt1$observed_range_sqkm))==trimws("NA")),NA,dt1$observed_range_sqkm)               
+suppressWarnings(dt1$observed_range_sqkm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$observed_range_sqkm))==as.character(as.numeric("NA"))),NA,dt1$observed_range_sqkm))
+dt1$inferred_range_sqkm <- ifelse((trimws(as.character(dt1$inferred_range_sqkm))==trimws("NA")),NA,dt1$inferred_range_sqkm)               
+suppressWarnings(dt1$inferred_range_sqkm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$inferred_range_sqkm))==as.character(as.numeric("NA"))),NA,dt1$inferred_range_sqkm))
+dt1$date_data_obtained <- as.factor(ifelse((trimws(as.character(dt1$date_data_obtained))==trimws("NA")),NA,as.character(dt1$date_data_obtained)))
+dt1$PanTHERIA_level <- as.factor(ifelse((trimws(as.character(dt1$PanTHERIA_level))==trimws("NA")),NA,as.character(dt1$PanTHERIA_level)))
+dt1$activity_cycle_p <- as.factor(ifelse((trimws(as.character(dt1$activity_cycle_p))==trimws("NA")),NA,as.character(dt1$activity_cycle_p)))
+dt1$adult_body_mass_g_p <- ifelse((trimws(as.character(dt1$adult_body_mass_g_p))==trimws("NA")),NA,dt1$adult_body_mass_g_p)               
+suppressWarnings(dt1$adult_body_mass_g_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$adult_body_mass_g_p))==as.character(as.numeric("NA"))),NA,dt1$adult_body_mass_g_p))
+dt1$adult_forearm_len_mm_p <- ifelse((trimws(as.character(dt1$adult_forearm_len_mm_p))==trimws("NA")),NA,dt1$adult_forearm_len_mm_p)               
+suppressWarnings(dt1$adult_forearm_len_mm_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$adult_forearm_len_mm_p))==as.character(as.numeric("NA"))),NA,dt1$adult_forearm_len_mm_p))
+dt1$adult_head_body_len_mm_p <- ifelse((trimws(as.character(dt1$adult_head_body_len_mm_p))==trimws("NA")),NA,dt1$adult_head_body_len_mm_p)               
+suppressWarnings(dt1$adult_head_body_len_mm_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$adult_head_body_len_mm_p))==as.character(as.numeric("NA"))),NA,dt1$adult_head_body_len_mm_p))
+dt1$age_at_eye_opening_d_p <- ifelse((trimws(as.character(dt1$age_at_eye_opening_d_p))==trimws("NA")),NA,dt1$age_at_eye_opening_d_p)               
+suppressWarnings(dt1$age_at_eye_opening_d_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$age_at_eye_opening_d_p))==as.character(as.numeric("NA"))),NA,dt1$age_at_eye_opening_d_p))
+dt1$age_at_first_birth_d_p <- ifelse((trimws(as.character(dt1$age_at_first_birth_d_p))==trimws("NA")),NA,dt1$age_at_first_birth_d_p)               
+suppressWarnings(dt1$age_at_first_birth_d_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$age_at_first_birth_d_p))==as.character(as.numeric("NA"))),NA,dt1$age_at_first_birth_d_p))
+dt1$basal_met_rate_mLO2hr_p <- ifelse((trimws(as.character(dt1$basal_met_rate_mLO2hr_p))==trimws("NA")),NA,dt1$basal_met_rate_mLO2hr_p)               
+suppressWarnings(dt1$basal_met_rate_mLO2hr_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$basal_met_rate_mLO2hr_p))==as.character(as.numeric("NA"))),NA,dt1$basal_met_rate_mLO2hr_p))
+dt1$basal_met_rate_mass_g_p <- ifelse((trimws(as.character(dt1$basal_met_rate_mass_g_p))==trimws("NA")),NA,dt1$basal_met_rate_mass_g_p)               
+suppressWarnings(dt1$basal_met_rate_mass_g_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$basal_met_rate_mass_g_p))==as.character(as.numeric("NA"))),NA,dt1$basal_met_rate_mass_g_p))
+dt1$diet_breadth_p <- ifelse((trimws(as.character(dt1$diet_breadth_p))==trimws("NA")),NA,dt1$diet_breadth_p)               
+suppressWarnings(dt1$diet_breadth_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$diet_breadth_p))==as.character(as.numeric("NA"))),NA,dt1$diet_breadth_p))
+dt1$dispersal_age_d_p <- ifelse((trimws(as.character(dt1$dispersal_age_d_p))==trimws("NA")),NA,dt1$dispersal_age_d_p)               
+suppressWarnings(dt1$dispersal_age_d_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$dispersal_age_d_p))==as.character(as.numeric("NA"))),NA,dt1$dispersal_age_d_p))
+dt1$gestation_len_d_p <- ifelse((trimws(as.character(dt1$gestation_len_d_p))==trimws("NA")),NA,dt1$gestation_len_d_p)               
+suppressWarnings(dt1$gestation_len_d_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$gestation_len_d_p))==as.character(as.numeric("NA"))),NA,dt1$gestation_len_d_p))
+dt1$habitat_breadth_p <- ifelse((trimws(as.character(dt1$habitat_breadth_p))==trimws("NA")),NA,dt1$habitat_breadth_p)               
+suppressWarnings(dt1$habitat_breadth_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$habitat_breadth_p))==as.character(as.numeric("NA"))),NA,dt1$habitat_breadth_p))
+dt1$home_range_km2_p <- ifelse((trimws(as.character(dt1$home_range_km2_p))==trimws("NA")),NA,dt1$home_range_km2_p)               
+suppressWarnings(dt1$home_range_km2_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$home_range_km2_p))==as.character(as.numeric("NA"))),NA,dt1$home_range_km2_p))
+dt1$home_range_indiv_km2_p <- ifelse((trimws(as.character(dt1$home_range_indiv_km2_p))==trimws("NA")),NA,dt1$home_range_indiv_km2_p)               
+suppressWarnings(dt1$home_range_indiv_km2_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$home_range_indiv_km2_p))==as.character(as.numeric("NA"))),NA,dt1$home_range_indiv_km2_p))
+dt1$Inter_birth_interval_d_p <- ifelse((trimws(as.character(dt1$Inter_birth_interval_d_p))==trimws("NA")),NA,dt1$Inter_birth_interval_d_p)               
+suppressWarnings(dt1$Inter_birth_interval_d_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$Inter_birth_interval_d_p))==as.character(as.numeric("NA"))),NA,dt1$Inter_birth_interval_d_p))
+dt1$litter_size_p <- ifelse((trimws(as.character(dt1$litter_size_p))==trimws("NA")),NA,dt1$litter_size_p)               
+suppressWarnings(dt1$litter_size_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$litter_size_p))==as.character(as.numeric("NA"))),NA,dt1$litter_size_p))
+dt1$litters_per_year_p <- ifelse((trimws(as.character(dt1$litters_per_year_p))==trimws("NA")),NA,dt1$litters_per_year_p)               
+suppressWarnings(dt1$litters_per_year_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$litters_per_year_p))==as.character(as.numeric("NA"))),NA,dt1$litters_per_year_p))
+dt1$max_longevity_m_p <- ifelse((trimws(as.character(dt1$max_longevity_m_p))==trimws("NA")),NA,dt1$max_longevity_m_p)               
+suppressWarnings(dt1$max_longevity_m_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$max_longevity_m_p))==as.character(as.numeric("NA"))),NA,dt1$max_longevity_m_p))
+dt1$neonate_body_mass_g_p <- ifelse((trimws(as.character(dt1$neonate_body_mass_g_p))==trimws("NA")),NA,dt1$neonate_body_mass_g_p)               
+suppressWarnings(dt1$neonate_body_mass_g_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$neonate_body_mass_g_p))==as.character(as.numeric("NA"))),NA,dt1$neonate_body_mass_g_p))
+dt1$neonate_head_body_len_mm_p <- ifelse((trimws(as.character(dt1$neonate_head_body_len_mm_p))==trimws("NA")),NA,dt1$neonate_head_body_len_mm_p)               
+suppressWarnings(dt1$neonate_head_body_len_mm_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$neonate_head_body_len_mm_p))==as.character(as.numeric("NA"))),NA,dt1$neonate_head_body_len_mm_p))
+dt1$population_density_n.km2_p <- ifelse((trimws(as.character(dt1$population_density_n.km2_p))==trimws("NA")),NA,dt1$population_density_n.km2_p)               
+suppressWarnings(dt1$population_density_n.km2_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$population_density_n.km2_p))==as.character(as.numeric("NA"))),NA,dt1$population_density_n.km2_p))
+dt1$population_grp_size_p <- ifelse((trimws(as.character(dt1$population_grp_size_p))==trimws("NA")),NA,dt1$population_grp_size_p)               
+suppressWarnings(dt1$population_grp_size_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$population_grp_size_p))==as.character(as.numeric("NA"))),NA,dt1$population_grp_size_p))
+dt1$sexual_maturity_age_d_p <- ifelse((trimws(as.character(dt1$sexual_maturity_age_d_p))==trimws("NA")),NA,dt1$sexual_maturity_age_d_p)               
+suppressWarnings(dt1$sexual_maturity_age_d_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$sexual_maturity_age_d_p))==as.character(as.numeric("NA"))),NA,dt1$sexual_maturity_age_d_p))
+dt1$social_grp_size_p <- ifelse((trimws(as.character(dt1$social_grp_size_p))==trimws("NA")),NA,dt1$social_grp_size_p)               
+suppressWarnings(dt1$social_grp_size_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$social_grp_size_p))==as.character(as.numeric("NA"))),NA,dt1$social_grp_size_p))
+dt1$teat_number_p <- ifelse((trimws(as.character(dt1$teat_number_p))==trimws("NA")),NA,dt1$teat_number_p)               
+suppressWarnings(dt1$teat_number_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$teat_number_p))==as.character(as.numeric("NA"))),NA,dt1$teat_number_p))
+dt1$terrestriality_p <- as.factor(ifelse((trimws(as.character(dt1$terrestriality_p))==trimws("NA")),NA,as.character(dt1$terrestriality_p)))
+dt1$trophic_level_p <- as.factor(ifelse((trimws(as.character(dt1$trophic_level_p))==trimws("NA")),NA,as.character(dt1$trophic_level_p)))
+dt1$weaning_age_d_p <- ifelse((trimws(as.character(dt1$weaning_age_d_p))==trimws("NA")),NA,dt1$weaning_age_d_p)               
+suppressWarnings(dt1$weaning_age_d_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$weaning_age_d_p))==as.character(as.numeric("NA"))),NA,dt1$weaning_age_d_p))
+dt1$weaning_body_mass_g_p <- ifelse((trimws(as.character(dt1$weaning_body_mass_g_p))==trimws("NA")),NA,dt1$weaning_body_mass_g_p)               
+suppressWarnings(dt1$weaning_body_mass_g_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$weaning_body_mass_g_p))==as.character(as.numeric("NA"))),NA,dt1$weaning_body_mass_g_p))
+dt1$weaning_head_body_len_mm_p <- ifelse((trimws(as.character(dt1$weaning_head_body_len_mm_p))==trimws("NA")),NA,dt1$weaning_head_body_len_mm_p)               
+suppressWarnings(dt1$weaning_head_body_len_mm_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$weaning_head_body_len_mm_p))==as.character(as.numeric("NA"))),NA,dt1$weaning_head_body_len_mm_p))
+dt1$references_p <- as.factor(ifelse((trimws(as.character(dt1$references_p))==trimws("NA")),NA,as.character(dt1$references_p)))
+dt1$adult_body_mass_g_EXT_p <- ifelse((trimws(as.character(dt1$adult_body_mass_g_EXT_p))==trimws("NA")),NA,dt1$adult_body_mass_g_EXT_p)               
+suppressWarnings(dt1$adult_body_mass_g_EXT_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$adult_body_mass_g_EXT_p))==as.character(as.numeric("NA"))),NA,dt1$adult_body_mass_g_EXT_p))
+dt1$litters_per_year_EXT_p <- ifelse((trimws(as.character(dt1$litters_per_year_EXT_p))==trimws("NA")),NA,dt1$litters_per_year_EXT_p)               
+suppressWarnings(dt1$litters_per_year_EXT_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$litters_per_year_EXT_p))==as.character(as.numeric("NA"))),NA,dt1$litters_per_year_EXT_p))
+dt1$neonate_body_mass_g_EXT_p <- ifelse((trimws(as.character(dt1$neonate_body_mass_g_EXT_p))==trimws("NA")),NA,dt1$neonate_body_mass_g_EXT_p)               
+suppressWarnings(dt1$neonate_body_mass_g_EXT_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$neonate_body_mass_g_EXT_p))==as.character(as.numeric("NA"))),NA,dt1$neonate_body_mass_g_EXT_p))
+dt1$weaning_body_mass_g_EXT_p <- ifelse((trimws(as.character(dt1$weaning_body_mass_g_EXT_p))==trimws("NA")),NA,dt1$weaning_body_mass_g_EXT_p)               
+suppressWarnings(dt1$weaning_body_mass_g_EXT_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$weaning_body_mass_g_EXT_p))==as.character(as.numeric("NA"))),NA,dt1$weaning_body_mass_g_EXT_p))
+dt1$GR_area_km2_p <- ifelse((trimws(as.character(dt1$GR_area_km2_p))==trimws("NA")),NA,dt1$GR_area_km2_p)               
+suppressWarnings(dt1$GR_area_km2_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$GR_area_km2_p))==as.character(as.numeric("NA"))),NA,dt1$GR_area_km2_p))
+dt1$GR_max_lat_dd_p <- ifelse((trimws(as.character(dt1$GR_max_lat_dd_p))==trimws("NA")),NA,dt1$GR_max_lat_dd_p)               
+suppressWarnings(dt1$GR_max_lat_dd_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$GR_max_lat_dd_p))==as.character(as.numeric("NA"))),NA,dt1$GR_max_lat_dd_p))
+dt1$GR_min_lat_dd_p <- ifelse((trimws(as.character(dt1$GR_min_lat_dd_p))==trimws("NA")),NA,dt1$GR_min_lat_dd_p)               
+suppressWarnings(dt1$GR_min_lat_dd_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$GR_min_lat_dd_p))==as.character(as.numeric("NA"))),NA,dt1$GR_min_lat_dd_p))
+dt1$GR_mid_range_lat_dd_p <- ifelse((trimws(as.character(dt1$GR_mid_range_lat_dd_p))==trimws("NA")),NA,dt1$GR_mid_range_lat_dd_p)               
+suppressWarnings(dt1$GR_mid_range_lat_dd_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$GR_mid_range_lat_dd_p))==as.character(as.numeric("NA"))),NA,dt1$GR_mid_range_lat_dd_p))
+dt1$GR_max_long_dd_p <- ifelse((trimws(as.character(dt1$GR_max_long_dd_p))==trimws("NA")),NA,dt1$GR_max_long_dd_p)               
+suppressWarnings(dt1$GR_max_long_dd_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$GR_max_long_dd_p))==as.character(as.numeric("NA"))),NA,dt1$GR_max_long_dd_p))
+dt1$GR_min_long_dd_p <- ifelse((trimws(as.character(dt1$GR_min_long_dd_p))==trimws("NA")),NA,dt1$GR_min_long_dd_p)               
+suppressWarnings(dt1$GR_min_long_dd_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$GR_min_long_dd_p))==as.character(as.numeric("NA"))),NA,dt1$GR_min_long_dd_p))
+dt1$GR_mid_range_long_dd_p <- ifelse((trimws(as.character(dt1$GR_mid_range_long_dd_p))==trimws("NA")),NA,dt1$GR_mid_range_long_dd_p)               
+suppressWarnings(dt1$GR_mid_range_long_dd_p <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$GR_mid_range_long_dd_p))==as.character(as.numeric("NA"))),NA,dt1$GR_mid_range_long_dd_p))
+dt1$hu_pop_den_min_n.km2_p_OD <- ifelse((trimws(as.character(dt1$hu_pop_den_min_n.km2_p_OD))==trimws("NA")),NA,dt1$hu_pop_den_min_n.km2_p_OD)               
+suppressWarnings(dt1$hu_pop_den_min_n.km2_p_OD <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$hu_pop_den_min_n.km2_p_OD))==as.character(as.numeric("NA"))),NA,dt1$hu_pop_den_min_n.km2_p_OD))
+dt1$hu_pop_den_mean_n.km2_p_OD <- ifelse((trimws(as.character(dt1$hu_pop_den_mean_n.km2_p_OD))==trimws("NA")),NA,dt1$hu_pop_den_mean_n.km2_p_OD)               
+suppressWarnings(dt1$hu_pop_den_mean_n.km2_p_OD <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$hu_pop_den_mean_n.km2_p_OD))==as.character(as.numeric("NA"))),NA,dt1$hu_pop_den_mean_n.km2_p_OD))
+dt1$hu_pop_den_5p_n.km2_p_OD <- ifelse((trimws(as.character(dt1$hu_pop_den_5p_n.km2_p_OD))==trimws("NA")),NA,dt1$hu_pop_den_5p_n.km2_p_OD)               
+suppressWarnings(dt1$hu_pop_den_5p_n.km2_p_OD <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$hu_pop_den_5p_n.km2_p_OD))==as.character(as.numeric("NA"))),NA,dt1$hu_pop_den_5p_n.km2_p_OD))
+dt1$hu_pop_den_change_p_OD <- ifelse((trimws(as.character(dt1$hu_pop_den_change_p_OD))==trimws("NA")),NA,dt1$hu_pop_den_change_p_OD)               
+suppressWarnings(dt1$hu_pop_den_change_p_OD <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$hu_pop_den_change_p_OD))==as.character(as.numeric("NA"))),NA,dt1$hu_pop_den_change_p_OD))
+dt1$precip_mean_mm_p_OD <- ifelse((trimws(as.character(dt1$precip_mean_mm_p_OD))==trimws("NA")),NA,dt1$precip_mean_mm_p_OD)               
+suppressWarnings(dt1$precip_mean_mm_p_OD <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$precip_mean_mm_p_OD))==as.character(as.numeric("NA"))),NA,dt1$precip_mean_mm_p_OD))
+dt1$temp_mean_01degC_p_OD <- ifelse((trimws(as.character(dt1$temp_mean_01degC_p_OD))==trimws("NA")),NA,dt1$temp_mean_01degC_p_OD)               
+suppressWarnings(dt1$temp_mean_01degC_p_OD <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$temp_mean_01degC_p_OD))==as.character(as.numeric("NA"))),NA,dt1$temp_mean_01degC_p_OD))
+dt1$AET_mean_mm_p_OD <- ifelse((trimws(as.character(dt1$AET_mean_mm_p_OD))==trimws("NA")),NA,dt1$AET_mean_mm_p_OD)               
+suppressWarnings(dt1$AET_mean_mm_p_OD <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$AET_mean_mm_p_OD))==as.character(as.numeric("NA"))),NA,dt1$AET_mean_mm_p_OD))
+dt1$PET_mean_mm_p_OD <- ifelse((trimws(as.character(dt1$PET_mean_mm_p_OD))==trimws("NA")),NA,dt1$PET_mean_mm_p_OD)               
+suppressWarnings(dt1$PET_mean_mm_p_OD <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$PET_mean_mm_p_OD))==as.character(as.numeric("NA"))),NA,dt1$PET_mean_mm_p_OD))
+
+# Here is the structure of the input data frame:
+str(dt1)                            
+attach(dt1)                            
+# The analyses below are basic descriptions of the variables. After testing, they should be replaced.
+
+summary(IUCN_species_name)
+summary(taxonid)
+summary(common_name)
+summary(family)
+summary(genus)
+summary(species)
+summary(subspecies)
+summary(rank)
+summary(elton_species_name)
+summary(taxonomic_disparity)
+summary(species_notes)
+summary(diet_inv_e)
+summary(diet_vend_e)
+summary(diet_vect_e)
+summary(diet_vfish_e)
+summary(diet_vunk_e)
+summary(diet_scav_e)
+summary(diet_fruit_e)
+summary(diet_nect_e)
+summary(diet_seed_e)
+summary(diet_plant_e)
+summary(diet_cat)
+summary(diet_source_e)
+summary(diet_certainty_e)
+summary(diet_breadth)
+summary(diet_level)
+summary(for_strat_value_e)
+summary(for_strat_certainty_e)
+summary(activity_nocturnal_e)
+summary(activity_crepuscular_e)
+summary(activity_diurnal_e)
+summary(activity_source_e)
+summary(activity_certainty_e)
+summary(activity_level)
+summary(body_mass_e)
+summary(body_mass_notes)
+summary(body_mass_level_e)
+summary(body_mass_source_e)
+summary(min_body_size_mm)
+summary(max_body_size_mm)
+summary(body_size_mm)
+summary(body_size_notes)
+summary(body_size_level)
+summary(body_size_source)
+summary(sexual_dimorphism)
+summary(sexual_dimorphism_notes)
+summary(sexual_dimorphism_level)
+summary(sexual_dimorphism_source)
+summary(min_longevity)
+summary(max_longevity)
+summary(longevity)
+summary(longevity_notes)
+summary(longevity_level)
+summary(longevity_source)
+summary(min_home_range)
+summary(max_home_range)
+summary(home_range_size)
+summary(home_range_notes)
+summary(home_range_level)
+summary(home_range_source)
+summary(habitat_specialization)
+summary(habitat_specialization_notes)
+summary(habitat_specialization_source)
+summary(min_generation_time)
+summary(max_generation_time)
+summary(generation_time)
+summary(generation_time_notes)
+summary(generation_time_level)
+summary(generation_time_source)
+summary(IUCN_category)
+summary(habitat)
+summary(habitat_suitability_montane)
+summary(habitat_suitability_lowland)
+summary(habitat_season_montane)
+summary(habitat_season_lowland)
+summary(habitat_major_importance_montane)
+summary(habitat_major_importance_lowland)
+summary(habitat_breadth)
+summary(habitat_breadth_source)
+summary(habitat_level)
+summary(mean_CHELSA_bio1_1981.2010_V.2.1)
+summary(mean_CHELSA_bio12_1981.2010_V.2.1)
+summary(mean_human_fp_range_2010)
+summary(mean_human_fp_range_2020)
+summary(percent_change_hf_2010_2020)
+summary(observed_range_sqkm)
+summary(inferred_range_sqkm)
+summary(date_data_obtained)
+summary(filled_by)
+summary(MSW05_order_p)
+summary(MSW05_family_p)
+summary(MSW05_genus_p)
+summary(MSW05_species_p)
+summary(PanTHERIA_level)
+summary(activity_cycle_p)
+summary(adult_body_mass_g_p)
+summary(adult_forearm_len_mm_p)
+summary(adult_head_body_len_mm_p)
+summary(age_at_eye_opening_d_p)
+summary(age_at_first_birth_d_p)
+summary(basal_met_rate_mLO2hr_p)
+summary(basal_met_rate_mass_g_p)
+summary(diet_breadth_p)
+summary(dispersal_age_d_p)
+summary(gestation_len_d_p)
+summary(habitat_breadth_p)
+summary(home_range_km2_p)
+summary(home_range_indiv_km2_p)
+summary(Inter_birth_interval_d_p)
+summary(litter_size_p)
+summary(litters_per_year_p)
+summary(max_longevity_m_p)
+summary(neonate_body_mass_g_p)
+summary(neonate_head_body_len_mm_p)
+summary(population_density_n.km2_p)
+summary(population_grp_size_p)
+summary(sexual_maturity_age_d_p)
+summary(social_grp_size_p)
+summary(teat_number_p)
+summary(terrestriality_p)
+summary(trophic_level_p)
+summary(weaning_age_d_p)
+summary(weaning_body_mass_g_p)
+summary(weaning_head_body_len_mm_p)
+summary(references_p)
+summary(adult_body_mass_g_EXT_p)
+summary(litters_per_year_EXT_p)
+summary(neonate_body_mass_g_EXT_p)
+summary(weaning_body_mass_g_EXT_p)
+summary(GR_area_km2_p)
+summary(GR_max_lat_dd_p)
+summary(GR_min_lat_dd_p)
+summary(GR_mid_range_lat_dd_p)
+summary(GR_max_long_dd_p)
+summary(GR_min_long_dd_p)
+summary(GR_mid_range_long_dd_p)
+summary(hu_pop_den_min_n.km2_p_OD)
+summary(hu_pop_den_mean_n.km2_p_OD)
+summary(hu_pop_den_5p_n.km2_p_OD)
+summary(hu_pop_den_change_p_OD)
+summary(precip_mean_mm_p_OD)
+summary(temp_mean_01degC_p_OD)
+summary(AET_mean_mm_p_OD)
+summary(PET_mean_mm_p_OD) 
+                # Get more details on character variables
+
+summary(as.factor(dt1$IUCN_species_name)) 
+summary(as.factor(dt1$common_name)) 
+summary(as.factor(dt1$family)) 
+summary(as.factor(dt1$genus)) 
+summary(as.factor(dt1$species)) 
+summary(as.factor(dt1$subspecies)) 
+summary(as.factor(dt1$rank)) 
+summary(as.factor(dt1$elton_species_name)) 
+summary(as.factor(dt1$taxonomic_disparity)) 
+summary(as.factor(dt1$species_notes)) 
+summary(as.factor(dt1$diet_cat)) 
+summary(as.factor(dt1$diet_source_e)) 
+summary(as.factor(dt1$diet_certainty_e)) 
+summary(as.factor(dt1$diet_level)) 
+summary(as.factor(dt1$for_strat_value_e)) 
+summary(as.factor(dt1$for_strat_certainty_e)) 
+summary(as.factor(dt1$activity_nocturnal_e)) 
+summary(as.factor(dt1$activity_crepuscular_e)) 
+summary(as.factor(dt1$activity_diurnal_e)) 
+summary(as.factor(dt1$activity_source_e)) 
+summary(as.factor(dt1$activity_certainty_e)) 
+summary(as.factor(dt1$activity_level)) 
+summary(as.factor(dt1$body_mass_notes)) 
+summary(as.factor(dt1$body_mass_level_e)) 
+summary(as.factor(dt1$body_mass_source_e)) 
+summary(as.factor(dt1$body_size_notes)) 
+summary(as.factor(dt1$body_size_level)) 
+summary(as.factor(dt1$body_size_source)) 
+summary(as.factor(dt1$sexual_dimorphism)) 
+summary(as.factor(dt1$sexual_dimorphism_notes)) 
+summary(as.factor(dt1$sexual_dimorphism_level)) 
+summary(as.factor(dt1$sexual_dimorphism_source)) 
+summary(as.factor(dt1$longevity_notes)) 
+summary(as.factor(dt1$longevity_level)) 
+summary(as.factor(dt1$longevity_source)) 
+summary(as.factor(dt1$home_range_notes)) 
+summary(as.factor(dt1$home_range_level)) 
+summary(as.factor(dt1$home_range_source)) 
+summary(as.factor(dt1$habitat_specialization)) 
+summary(as.factor(dt1$habitat_specialization_notes)) 
+summary(as.factor(dt1$habitat_specialization_source)) 
+summary(as.factor(dt1$generation_time_notes)) 
+summary(as.factor(dt1$generation_time_level)) 
+summary(as.factor(dt1$generation_time_source)) 
+summary(as.factor(dt1$IUCN_category)) 
+summary(as.factor(dt1$habitat)) 
+summary(as.factor(dt1$habitat_suitability_montane)) 
+summary(as.factor(dt1$habitat_suitability_lowland)) 
+summary(as.factor(dt1$habitat_season_montane)) 
+summary(as.factor(dt1$habitat_season_lowland)) 
+summary(as.factor(dt1$habitat_major_importance_montane)) 
+summary(as.factor(dt1$habitat_major_importance_lowland)) 
+summary(as.factor(dt1$habitat_breadth_source)) 
+summary(as.factor(dt1$habitat_level)) 
+summary(as.factor(dt1$date_data_obtained)) 
+summary(as.factor(dt1$filled_by)) 
+summary(as.factor(dt1$MSW05_order_p)) 
+summary(as.factor(dt1$MSW05_family_p)) 
+summary(as.factor(dt1$MSW05_genus_p)) 
+summary(as.factor(dt1$MSW05_species_p)) 
+summary(as.factor(dt1$PanTHERIA_level)) 
+summary(as.factor(dt1$activity_cycle_p)) 
+summary(as.factor(dt1$terrestriality_p)) 
+summary(as.factor(dt1$trophic_level_p)) 
+summary(as.factor(dt1$references_p))
+detach(dt1)
+
+inUrl2  <- "https://pasta.lternet.edu/package/data/eml/edi/1220/6/b020ef4c1a1310c3db45504557f99bae" 
+infile2 <- tempfile()
+try(download.file(inUrl2,infile2,method="curl"))
+if (is.na(file.size(infile2))) download.file(inUrl2,infile2,method="auto")
+
+ dt2 <-read.csv(infile2,header=F 
+          ,skip=1
+            ,sep=","  
+                ,quot='"' 
+        , col.names=c(
+                    "IUCN_species_name",     
+                    "taxonid",     
+                    "genus",     
+                    "species",     
+                    "common_name",     
+                    "elton_species_name",     
+                    "taxonomic_disparity",     
+                    "pass_non_pass_e",     
+                    "order_e",     
+                    "family_e",     
+                    "family_english_e",     
+                    "diet_inv_e",     
+                    "diet_vend_e",     
+                    "diet_vect_e",     
+                    "diet_vfish_e",     
+                    "diet_vunk_e",     
+                    "diet_scav_e",     
+                    "diet_fruit_e",     
+                    "diet_nect_e",     
+                    "diet_seed_e",     
+                    "diet_plant_e",     
+                    "diet_cat_e",     
+                    "diet_source_e",     
+                    "diet_certainty_e",     
+                    "diet_breadth",     
+                    "diet_level",     
+                    "for_strat_water",     
+                    "for_strat_ground_e",     
+                    "for_strat_understory_e",     
+                    "for_strat_midhigh_e",     
+                    "for_strat_canopy_e",     
+                    "for_strat_aerial_e",     
+                    "for_strat_spec_level",     
+                    "for_strat_source_e",     
+                    "activity_nocturnal_e",     
+                    "body_mass_e",     
+                    "body_mass_notes",     
+                    "body_mass_level_e",     
+                    "body_mass_source_e",     
+                    "min_body_size_mm",     
+                    "max_body_size_mm",     
+                    "body_size_mm",     
+                    "body_size_notes",     
+                    "body_size_level",     
+                    "body_size_source",     
+                    "sexual_dimorphism",     
+                    "sexual_dimorphism_notes",     
+                    "sexual_dimorphism_level",     
+                    "sexual_dimorphism_source",     
+                    "min_longevity",     
+                    "max_longevity",     
+                    "longevity_notes",     
+                    "longevity",     
+                    "longevity_source",     
+                    "longevity_level",     
+                    "min_home_range",     
+                    "max_home_range",     
+                    "home_range_size",     
+                    "home_range_notes",     
+                    "home_range_level",     
+                    "home_range_source",     
+                    "habitat_specialization",     
+                    "habitat_specialization_notes",     
+                    "habitat_specialization_source",     
+                    "generation_time",     
+                    "generation_time_level",     
+                    "generation_time_notes",     
+                    "generation_time_source",     
+                    "IUCN_category",     
+                    "habitat",     
+                    "habitat_suitability_montane",     
+                    "habitat_suitability_lowland",     
+                    "habitat_season_montane",     
+                    "habitat_season_lowland",     
+                    "habitat_major_importance_montane",     
+                    "habitat_major_importance_lowland",     
+                    "habitat_breadth",     
+                    "habitat_breadth_source",     
+                    "habitat_level",     
+                    "mean_CHELSA_bio1_1981.2010_V.2.1",     
+                    "mean_CHELSA_bio12_1981.2010_V.2.1",     
+                    "mean_human_fp_range_2010",     
+                    "mean_human_fp_range_2020",     
+                    "percent_change_hf_2010_2020",     
+                    "observed_range_sqkm",     
+                    "inferred_range_sqkm",     
+                    "date_data_obtained",     
+                    "filled_by"    ), check.names=TRUE)
+
+unlink(infile2)
+
+# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
+
+if (class(dt2$IUCN_species_name)!="factor") dt2$IUCN_species_name<- as.factor(dt2$IUCN_species_name)
+if (class(dt2$taxonid)=="factor") dt2$taxonid <-as.numeric(levels(dt2$taxonid))[as.integer(dt2$taxonid) ]               
+if (class(dt2$taxonid)=="character") dt2$taxonid <-as.numeric(dt2$taxonid)
+if (class(dt2$genus)!="factor") dt2$genus<- as.factor(dt2$genus)
+if (class(dt2$species)!="factor") dt2$species<- as.factor(dt2$species)
+if (class(dt2$common_name)!="factor") dt2$common_name<- as.factor(dt2$common_name)
+if (class(dt2$elton_species_name)!="factor") dt2$elton_species_name<- as.factor(dt2$elton_species_name)
+if (class(dt2$taxonomic_disparity)!="factor") dt2$taxonomic_disparity<- as.factor(dt2$taxonomic_disparity)
+if (class(dt2$pass_non_pass_e)!="factor") dt2$pass_non_pass_e<- as.factor(dt2$pass_non_pass_e)
+if (class(dt2$order_e)!="factor") dt2$order_e<- as.factor(dt2$order_e)
+if (class(dt2$family_e)!="factor") dt2$family_e<- as.factor(dt2$family_e)
+if (class(dt2$family_english_e)!="factor") dt2$family_english_e<- as.factor(dt2$family_english_e)
+if (class(dt2$diet_inv_e)=="factor") dt2$diet_inv_e <-as.numeric(levels(dt2$diet_inv_e))[as.integer(dt2$diet_inv_e) ]               
+if (class(dt2$diet_inv_e)=="character") dt2$diet_inv_e <-as.numeric(dt2$diet_inv_e)
+if (class(dt2$diet_vend_e)=="factor") dt2$diet_vend_e <-as.numeric(levels(dt2$diet_vend_e))[as.integer(dt2$diet_vend_e) ]               
+if (class(dt2$diet_vend_e)=="character") dt2$diet_vend_e <-as.numeric(dt2$diet_vend_e)
+if (class(dt2$diet_vect_e)=="factor") dt2$diet_vect_e <-as.numeric(levels(dt2$diet_vect_e))[as.integer(dt2$diet_vect_e) ]               
+if (class(dt2$diet_vect_e)=="character") dt2$diet_vect_e <-as.numeric(dt2$diet_vect_e)
+if (class(dt2$diet_vfish_e)=="factor") dt2$diet_vfish_e <-as.numeric(levels(dt2$diet_vfish_e))[as.integer(dt2$diet_vfish_e) ]               
+if (class(dt2$diet_vfish_e)=="character") dt2$diet_vfish_e <-as.numeric(dt2$diet_vfish_e)
+if (class(dt2$diet_vunk_e)=="factor") dt2$diet_vunk_e <-as.numeric(levels(dt2$diet_vunk_e))[as.integer(dt2$diet_vunk_e) ]               
+if (class(dt2$diet_vunk_e)=="character") dt2$diet_vunk_e <-as.numeric(dt2$diet_vunk_e)
+if (class(dt2$diet_scav_e)=="factor") dt2$diet_scav_e <-as.numeric(levels(dt2$diet_scav_e))[as.integer(dt2$diet_scav_e) ]               
+if (class(dt2$diet_scav_e)=="character") dt2$diet_scav_e <-as.numeric(dt2$diet_scav_e)
+if (class(dt2$diet_fruit_e)=="factor") dt2$diet_fruit_e <-as.numeric(levels(dt2$diet_fruit_e))[as.integer(dt2$diet_fruit_e) ]               
+if (class(dt2$diet_fruit_e)=="character") dt2$diet_fruit_e <-as.numeric(dt2$diet_fruit_e)
+if (class(dt2$diet_nect_e)=="factor") dt2$diet_nect_e <-as.numeric(levels(dt2$diet_nect_e))[as.integer(dt2$diet_nect_e) ]               
+if (class(dt2$diet_nect_e)=="character") dt2$diet_nect_e <-as.numeric(dt2$diet_nect_e)
+if (class(dt2$diet_seed_e)=="factor") dt2$diet_seed_e <-as.numeric(levels(dt2$diet_seed_e))[as.integer(dt2$diet_seed_e) ]               
+if (class(dt2$diet_seed_e)=="character") dt2$diet_seed_e <-as.numeric(dt2$diet_seed_e)
+if (class(dt2$diet_plant_e)=="factor") dt2$diet_plant_e <-as.numeric(levels(dt2$diet_plant_e))[as.integer(dt2$diet_plant_e) ]               
+if (class(dt2$diet_plant_e)=="character") dt2$diet_plant_e <-as.numeric(dt2$diet_plant_e)
+if (class(dt2$diet_cat_e)!="factor") dt2$diet_cat_e<- as.factor(dt2$diet_cat_e)
+if (class(dt2$diet_source_e)!="factor") dt2$diet_source_e<- as.factor(dt2$diet_source_e)
+if (class(dt2$diet_certainty_e)!="factor") dt2$diet_certainty_e<- as.factor(dt2$diet_certainty_e)
+if (class(dt2$diet_breadth)=="factor") dt2$diet_breadth <-as.numeric(levels(dt2$diet_breadth))[as.integer(dt2$diet_breadth) ]               
+if (class(dt2$diet_breadth)=="character") dt2$diet_breadth <-as.numeric(dt2$diet_breadth)
+if (class(dt2$diet_level)!="factor") dt2$diet_level<- as.factor(dt2$diet_level)
+if (class(dt2$for_strat_water)=="factor") dt2$for_strat_water <-as.numeric(levels(dt2$for_strat_water))[as.integer(dt2$for_strat_water) ]               
+if (class(dt2$for_strat_water)=="character") dt2$for_strat_water <-as.numeric(dt2$for_strat_water)
+if (class(dt2$for_strat_ground_e)=="factor") dt2$for_strat_ground_e <-as.numeric(levels(dt2$for_strat_ground_e))[as.integer(dt2$for_strat_ground_e) ]               
+if (class(dt2$for_strat_ground_e)=="character") dt2$for_strat_ground_e <-as.numeric(dt2$for_strat_ground_e)
+if (class(dt2$for_strat_understory_e)=="factor") dt2$for_strat_understory_e <-as.numeric(levels(dt2$for_strat_understory_e))[as.integer(dt2$for_strat_understory_e) ]               
+if (class(dt2$for_strat_understory_e)=="character") dt2$for_strat_understory_e <-as.numeric(dt2$for_strat_understory_e)
+if (class(dt2$for_strat_midhigh_e)=="factor") dt2$for_strat_midhigh_e <-as.numeric(levels(dt2$for_strat_midhigh_e))[as.integer(dt2$for_strat_midhigh_e) ]               
+if (class(dt2$for_strat_midhigh_e)=="character") dt2$for_strat_midhigh_e <-as.numeric(dt2$for_strat_midhigh_e)
+if (class(dt2$for_strat_canopy_e)=="factor") dt2$for_strat_canopy_e <-as.numeric(levels(dt2$for_strat_canopy_e))[as.integer(dt2$for_strat_canopy_e) ]               
+if (class(dt2$for_strat_canopy_e)=="character") dt2$for_strat_canopy_e <-as.numeric(dt2$for_strat_canopy_e)
+if (class(dt2$for_strat_aerial_e)=="factor") dt2$for_strat_aerial_e <-as.numeric(levels(dt2$for_strat_aerial_e))[as.integer(dt2$for_strat_aerial_e) ]               
+if (class(dt2$for_strat_aerial_e)=="character") dt2$for_strat_aerial_e <-as.numeric(dt2$for_strat_aerial_e)
+if (class(dt2$for_strat_spec_level)!="factor") dt2$for_strat_spec_level<- as.factor(dt2$for_strat_spec_level)
+if (class(dt2$for_strat_source_e)!="factor") dt2$for_strat_source_e<- as.factor(dt2$for_strat_source_e)
+if (class(dt2$activity_nocturnal_e)!="factor") dt2$activity_nocturnal_e<- as.factor(dt2$activity_nocturnal_e)
+if (class(dt2$body_mass_e)=="factor") dt2$body_mass_e <-as.numeric(levels(dt2$body_mass_e))[as.integer(dt2$body_mass_e) ]               
+if (class(dt2$body_mass_e)=="character") dt2$body_mass_e <-as.numeric(dt2$body_mass_e)
+if (class(dt2$body_mass_notes)!="factor") dt2$body_mass_notes<- as.factor(dt2$body_mass_notes)
+if (class(dt2$body_mass_level_e)!="factor") dt2$body_mass_level_e<- as.factor(dt2$body_mass_level_e)
+if (class(dt2$body_mass_source_e)!="factor") dt2$body_mass_source_e<- as.factor(dt2$body_mass_source_e)
+if (class(dt2$min_body_size_mm)=="factor") dt2$min_body_size_mm <-as.numeric(levels(dt2$min_body_size_mm))[as.integer(dt2$min_body_size_mm) ]               
+if (class(dt2$min_body_size_mm)=="character") dt2$min_body_size_mm <-as.numeric(dt2$min_body_size_mm)
+if (class(dt2$max_body_size_mm)=="factor") dt2$max_body_size_mm <-as.numeric(levels(dt2$max_body_size_mm))[as.integer(dt2$max_body_size_mm) ]               
+if (class(dt2$max_body_size_mm)=="character") dt2$max_body_size_mm <-as.numeric(dt2$max_body_size_mm)
+if (class(dt2$body_size_mm)=="factor") dt2$body_size_mm <-as.numeric(levels(dt2$body_size_mm))[as.integer(dt2$body_size_mm) ]               
+if (class(dt2$body_size_mm)=="character") dt2$body_size_mm <-as.numeric(dt2$body_size_mm)
+if (class(dt2$body_size_notes)!="factor") dt2$body_size_notes<- as.factor(dt2$body_size_notes)
+if (class(dt2$body_size_level)!="factor") dt2$body_size_level<- as.factor(dt2$body_size_level)
+if (class(dt2$body_size_source)!="factor") dt2$body_size_source<- as.factor(dt2$body_size_source)
+if (class(dt2$sexual_dimorphism)!="factor") dt2$sexual_dimorphism<- as.factor(dt2$sexual_dimorphism)
+if (class(dt2$sexual_dimorphism_notes)!="factor") dt2$sexual_dimorphism_notes<- as.factor(dt2$sexual_dimorphism_notes)
+if (class(dt2$sexual_dimorphism_level)!="factor") dt2$sexual_dimorphism_level<- as.factor(dt2$sexual_dimorphism_level)
+if (class(dt2$sexual_dimorphism_source)!="factor") dt2$sexual_dimorphism_source<- as.factor(dt2$sexual_dimorphism_source)
+if (class(dt2$min_longevity)=="factor") dt2$min_longevity <-as.numeric(levels(dt2$min_longevity))[as.integer(dt2$min_longevity) ]               
+if (class(dt2$min_longevity)=="character") dt2$min_longevity <-as.numeric(dt2$min_longevity)
+if (class(dt2$max_longevity)=="factor") dt2$max_longevity <-as.numeric(levels(dt2$max_longevity))[as.integer(dt2$max_longevity) ]               
+if (class(dt2$max_longevity)=="character") dt2$max_longevity <-as.numeric(dt2$max_longevity)
+if (class(dt2$longevity_notes)!="factor") dt2$longevity_notes<- as.factor(dt2$longevity_notes)
+if (class(dt2$longevity)=="factor") dt2$longevity <-as.numeric(levels(dt2$longevity))[as.integer(dt2$longevity) ]               
+if (class(dt2$longevity)=="character") dt2$longevity <-as.numeric(dt2$longevity)
+if (class(dt2$longevity_source)!="factor") dt2$longevity_source<- as.factor(dt2$longevity_source)
+if (class(dt2$longevity_level)!="factor") dt2$longevity_level<- as.factor(dt2$longevity_level)
+if (class(dt2$min_home_range)=="factor") dt2$min_home_range <-as.numeric(levels(dt2$min_home_range))[as.integer(dt2$min_home_range) ]               
+if (class(dt2$min_home_range)=="character") dt2$min_home_range <-as.numeric(dt2$min_home_range)
+if (class(dt2$max_home_range)=="factor") dt2$max_home_range <-as.numeric(levels(dt2$max_home_range))[as.integer(dt2$max_home_range) ]               
+if (class(dt2$max_home_range)=="character") dt2$max_home_range <-as.numeric(dt2$max_home_range)
+if (class(dt2$home_range_size)=="factor") dt2$home_range_size <-as.numeric(levels(dt2$home_range_size))[as.integer(dt2$home_range_size) ]               
+if (class(dt2$home_range_size)=="character") dt2$home_range_size <-as.numeric(dt2$home_range_size)
+if (class(dt2$home_range_notes)!="factor") dt2$home_range_notes<- as.factor(dt2$home_range_notes)
+if (class(dt2$home_range_level)!="factor") dt2$home_range_level<- as.factor(dt2$home_range_level)
+if (class(dt2$home_range_source)!="factor") dt2$home_range_source<- as.factor(dt2$home_range_source)
+if (class(dt2$habitat_specialization)!="factor") dt2$habitat_specialization<- as.factor(dt2$habitat_specialization)
+if (class(dt2$habitat_specialization_notes)!="factor") dt2$habitat_specialization_notes<- as.factor(dt2$habitat_specialization_notes)
+if (class(dt2$habitat_specialization_source)!="factor") dt2$habitat_specialization_source<- as.factor(dt2$habitat_specialization_source)
+if (class(dt2$generation_time)=="factor") dt2$generation_time <-as.numeric(levels(dt2$generation_time))[as.integer(dt2$generation_time) ]               
+if (class(dt2$generation_time)=="character") dt2$generation_time <-as.numeric(dt2$generation_time)
+if (class(dt2$generation_time_level)!="factor") dt2$generation_time_level<- as.factor(dt2$generation_time_level)
+if (class(dt2$generation_time_notes)!="factor") dt2$generation_time_notes<- as.factor(dt2$generation_time_notes)
+if (class(dt2$generation_time_source)!="factor") dt2$generation_time_source<- as.factor(dt2$generation_time_source)
+if (class(dt2$IUCN_category)!="factor") dt2$IUCN_category<- as.factor(dt2$IUCN_category)
+if (class(dt2$habitat)!="factor") dt2$habitat<- as.factor(dt2$habitat)
+if (class(dt2$habitat_suitability_montane)!="factor") dt2$habitat_suitability_montane<- as.factor(dt2$habitat_suitability_montane)
+if (class(dt2$habitat_suitability_lowland)!="factor") dt2$habitat_suitability_lowland<- as.factor(dt2$habitat_suitability_lowland)
+if (class(dt2$habitat_season_montane)!="factor") dt2$habitat_season_montane<- as.factor(dt2$habitat_season_montane)
+if (class(dt2$habitat_season_lowland)!="factor") dt2$habitat_season_lowland<- as.factor(dt2$habitat_season_lowland)
+if (class(dt2$habitat_major_importance_montane)!="factor") dt2$habitat_major_importance_montane<- as.factor(dt2$habitat_major_importance_montane)
+if (class(dt2$habitat_major_importance_lowland)!="factor") dt2$habitat_major_importance_lowland<- as.factor(dt2$habitat_major_importance_lowland)
+if (class(dt2$habitat_breadth)=="factor") dt2$habitat_breadth <-as.numeric(levels(dt2$habitat_breadth))[as.integer(dt2$habitat_breadth) ]               
+if (class(dt2$habitat_breadth)=="character") dt2$habitat_breadth <-as.numeric(dt2$habitat_breadth)
+if (class(dt2$habitat_breadth_source)!="factor") dt2$habitat_breadth_source<- as.factor(dt2$habitat_breadth_source)
+if (class(dt2$habitat_level)!="factor") dt2$habitat_level<- as.factor(dt2$habitat_level)
+if (class(dt2$mean_CHELSA_bio1_1981.2010_V.2.1)=="factor") dt2$mean_CHELSA_bio1_1981.2010_V.2.1 <-as.numeric(levels(dt2$mean_CHELSA_bio1_1981.2010_V.2.1))[as.integer(dt2$mean_CHELSA_bio1_1981.2010_V.2.1) ]
+if (class(dt2$mean_CHELSA_bio1_1981.2010_V.2.1)=="character") dt2$mean_CHELSA_bio1_1981.2010_V.2.1 <-as.numeric(dt2$mean_CHELSA_bio1_1981.2010_V.2.1)
+if (class(dt2$mean_CHELSA_bio12_1981.2010_V.2.1)=="factor") dt2$mean_CHELSA_bio12_1981.2010_V.2.1 <-as.numeric(levels(dt2$mean_CHELSA_bio12_1981.2010_V.2.1))[as.integer(dt2$mean_CHELSA_bio12_1981.2010_V.2.1) ]
+if (class(dt2$mean_CHELSA_bio12_1981.2010_V.2.1)=="character") dt2$mean_CHELSA_bio12_1981.2010_V.2.1 <-as.numeric(dt2$mean_CHELSA_bio12_1981.2010_V.2.1)
+if (class(dt2$mean_human_fp_range_2010)=="factor") dt2$mean_human_fp_range_2010 <-as.numeric(levels(dt2$mean_human_fp_range_2010))[as.integer(dt2$mean_human_fp_range_2010) ]               
+if (class(dt2$mean_human_fp_range_2010)=="character") dt2$mean_human_fp_range_2010 <-as.numeric(dt2$mean_human_fp_range_2010)
+if (class(dt2$mean_human_fp_range_2020)=="factor") dt2$mean_human_fp_range_2020 <-as.numeric(levels(dt2$mean_human_fp_range_2020))[as.integer(dt2$mean_human_fp_range_2020) ]               
+if (class(dt2$mean_human_fp_range_2020)=="character") dt2$mean_human_fp_range_2020 <-as.numeric(dt2$mean_human_fp_range_2020)
+if (class(dt2$percent_change_hf_2010_2020)=="factor") dt2$percent_change_hf_2010_2020 <-as.numeric(levels(dt2$percent_change_hf_2010_2020))[as.integer(dt2$percent_change_hf_2010_2020) ]               
+if (class(dt2$percent_change_hf_2010_2020)=="character") dt2$percent_change_hf_2010_2020 <-as.numeric(dt2$percent_change_hf_2010_2020)
+if (class(dt2$observed_range_sqkm)=="factor") dt2$observed_range_sqkm <-as.numeric(levels(dt2$observed_range_sqkm))[as.integer(dt2$observed_range_sqkm) ]               
+if (class(dt2$observed_range_sqkm)=="character") dt2$observed_range_sqkm <-as.numeric(dt2$observed_range_sqkm)
+if (class(dt2$inferred_range_sqkm)=="factor") dt2$inferred_range_sqkm <-as.numeric(levels(dt2$inferred_range_sqkm))[as.integer(dt2$inferred_range_sqkm) ]               
+if (class(dt2$inferred_range_sqkm)=="character") dt2$inferred_range_sqkm <-as.numeric(dt2$inferred_range_sqkm)
+if (class(dt2$date_data_obtained)!="factor") dt2$date_data_obtained<- as.factor(dt2$date_data_obtained)
+if (class(dt2$filled_by)!="factor") dt2$filled_by<- as.factor(dt2$filled_by)
+
+# Convert Missing Values to NA for non-dates
+
+dt2$IUCN_species_name <- as.factor(ifelse((trimws(as.character(dt2$IUCN_species_name))==trimws("NA")),NA,as.character(dt2$IUCN_species_name)))
+dt2$taxonid <- ifelse((trimws(as.character(dt2$taxonid))==trimws("NA")),NA,dt2$taxonid)               
+suppressWarnings(dt2$taxonid <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$taxonid))==as.character(as.numeric("NA"))),NA,dt2$taxonid))
+dt2$genus <- as.factor(ifelse((trimws(as.character(dt2$genus))==trimws("NA")),NA,as.character(dt2$genus)))
+dt2$common_name <- as.factor(ifelse((trimws(as.character(dt2$common_name))==trimws("NA")),NA,as.character(dt2$common_name)))
+dt2$elton_species_name <- as.factor(ifelse((trimws(as.character(dt2$elton_species_name))==trimws("NA")),NA,as.character(dt2$elton_species_name)))
+dt2$taxonomic_disparity <- as.factor(ifelse((trimws(as.character(dt2$taxonomic_disparity))==trimws("NA")),NA,as.character(dt2$taxonomic_disparity)))
+dt2$order_e <- as.factor(ifelse((trimws(as.character(dt2$order_e))==trimws("NA")),NA,as.character(dt2$order_e)))
+dt2$family_e <- as.factor(ifelse((trimws(as.character(dt2$family_e))==trimws("NA")),NA,as.character(dt2$family_e)))
+dt2$family_english_e <- as.factor(ifelse((trimws(as.character(dt2$family_english_e))==trimws("NA")),NA,as.character(dt2$family_english_e)))
+dt2$diet_inv_e <- ifelse((trimws(as.character(dt2$diet_inv_e))==trimws("NA")),NA,dt2$diet_inv_e)               
+suppressWarnings(dt2$diet_inv_e <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$diet_inv_e))==as.character(as.numeric("NA"))),NA,dt2$diet_inv_e))
+dt2$diet_vend_e <- ifelse((trimws(as.character(dt2$diet_vend_e))==trimws("NA")),NA,dt2$diet_vend_e)               
+suppressWarnings(dt2$diet_vend_e <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$diet_vend_e))==as.character(as.numeric("NA"))),NA,dt2$diet_vend_e))
+dt2$diet_vect_e <- ifelse((trimws(as.character(dt2$diet_vect_e))==trimws("NA")),NA,dt2$diet_vect_e)               
+suppressWarnings(dt2$diet_vect_e <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$diet_vect_e))==as.character(as.numeric("NA"))),NA,dt2$diet_vect_e))
+dt2$diet_vfish_e <- ifelse((trimws(as.character(dt2$diet_vfish_e))==trimws("NA")),NA,dt2$diet_vfish_e)               
+suppressWarnings(dt2$diet_vfish_e <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$diet_vfish_e))==as.character(as.numeric("NA"))),NA,dt2$diet_vfish_e))
+dt2$diet_vunk_e <- ifelse((trimws(as.character(dt2$diet_vunk_e))==trimws("NA")),NA,dt2$diet_vunk_e)               
+suppressWarnings(dt2$diet_vunk_e <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$diet_vunk_e))==as.character(as.numeric("NA"))),NA,dt2$diet_vunk_e))
+dt2$diet_scav_e <- ifelse((trimws(as.character(dt2$diet_scav_e))==trimws("NA")),NA,dt2$diet_scav_e)               
+suppressWarnings(dt2$diet_scav_e <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$diet_scav_e))==as.character(as.numeric("NA"))),NA,dt2$diet_scav_e))
+dt2$diet_fruit_e <- ifelse((trimws(as.character(dt2$diet_fruit_e))==trimws("NA")),NA,dt2$diet_fruit_e)               
+suppressWarnings(dt2$diet_fruit_e <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$diet_fruit_e))==as.character(as.numeric("NA"))),NA,dt2$diet_fruit_e))
+dt2$diet_nect_e <- ifelse((trimws(as.character(dt2$diet_nect_e))==trimws("NA")),NA,dt2$diet_nect_e)               
+suppressWarnings(dt2$diet_nect_e <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$diet_nect_e))==as.character(as.numeric("NA"))),NA,dt2$diet_nect_e))
+dt2$diet_seed_e <- ifelse((trimws(as.character(dt2$diet_seed_e))==trimws("NA")),NA,dt2$diet_seed_e)               
+suppressWarnings(dt2$diet_seed_e <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$diet_seed_e))==as.character(as.numeric("NA"))),NA,dt2$diet_seed_e))
+dt2$diet_plant_e <- ifelse((trimws(as.character(dt2$diet_plant_e))==trimws("NA")),NA,dt2$diet_plant_e)               
+suppressWarnings(dt2$diet_plant_e <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$diet_plant_e))==as.character(as.numeric("NA"))),NA,dt2$diet_plant_e))
+dt2$diet_cat_e <- as.factor(ifelse((trimws(as.character(dt2$diet_cat_e))==trimws("NA")),NA,as.character(dt2$diet_cat_e)))
+dt2$diet_level <- as.factor(ifelse((trimws(as.character(dt2$diet_level))==trimws("NA")),NA,as.character(dt2$diet_level)))
+dt2$for_strat_spec_level <- as.factor(ifelse((trimws(as.character(dt2$for_strat_spec_level))==trimws("NA")),NA,as.character(dt2$for_strat_spec_level)))
+dt2$for_strat_source_e <- as.factor(ifelse((trimws(as.character(dt2$for_strat_source_e))==trimws("NA")),NA,as.character(dt2$for_strat_source_e)))
+dt2$activity_nocturnal_e <- as.factor(ifelse((trimws(as.character(dt2$activity_nocturnal_e))==trimws("NA")),NA,as.character(dt2$activity_nocturnal_e)))
+dt2$body_mass_e <- ifelse((trimws(as.character(dt2$body_mass_e))==trimws("NA")),NA,dt2$body_mass_e)               
+suppressWarnings(dt2$body_mass_e <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$body_mass_e))==as.character(as.numeric("NA"))),NA,dt2$body_mass_e))
+dt2$body_mass_notes <- as.factor(ifelse((trimws(as.character(dt2$body_mass_notes))==trimws("NA")),NA,as.character(dt2$body_mass_notes)))
+dt2$body_mass_level_e <- as.factor(ifelse((trimws(as.character(dt2$body_mass_level_e))==trimws("NA")),NA,as.character(dt2$body_mass_level_e)))
+dt2$body_mass_source_e <- as.factor(ifelse((trimws(as.character(dt2$body_mass_source_e))==trimws("NA")),NA,as.character(dt2$body_mass_source_e)))
+dt2$min_body_size_mm <- ifelse((trimws(as.character(dt2$min_body_size_mm))==trimws("NA")),NA,dt2$min_body_size_mm)               
+suppressWarnings(dt2$min_body_size_mm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$min_body_size_mm))==as.character(as.numeric("NA"))),NA,dt2$min_body_size_mm))
+dt2$max_body_size_mm <- ifelse((trimws(as.character(dt2$max_body_size_mm))==trimws("NA")),NA,dt2$max_body_size_mm)               
+suppressWarnings(dt2$max_body_size_mm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$max_body_size_mm))==as.character(as.numeric("NA"))),NA,dt2$max_body_size_mm))
+dt2$body_size_mm <- ifelse((trimws(as.character(dt2$body_size_mm))==trimws("NA")),NA,dt2$body_size_mm)               
+suppressWarnings(dt2$body_size_mm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$body_size_mm))==as.character(as.numeric("NA"))),NA,dt2$body_size_mm))
+dt2$body_size_notes <- as.factor(ifelse((trimws(as.character(dt2$body_size_notes))==trimws("NA")),NA,as.character(dt2$body_size_notes)))
+dt2$body_size_level <- as.factor(ifelse((trimws(as.character(dt2$body_size_level))==trimws("NA")),NA,as.character(dt2$body_size_level)))
+dt2$body_size_source <- as.factor(ifelse((trimws(as.character(dt2$body_size_source))==trimws("NA")),NA,as.character(dt2$body_size_source)))
+dt2$sexual_dimorphism <- as.factor(ifelse((trimws(as.character(dt2$sexual_dimorphism))==trimws("NA")),NA,as.character(dt2$sexual_dimorphism)))
+dt2$sexual_dimorphism_notes <- as.factor(ifelse((trimws(as.character(dt2$sexual_dimorphism_notes))==trimws("NA")),NA,as.character(dt2$sexual_dimorphism_notes)))
+dt2$sexual_dimorphism_level <- as.factor(ifelse((trimws(as.character(dt2$sexual_dimorphism_level))==trimws("NA")),NA,as.character(dt2$sexual_dimorphism_level)))
+dt2$sexual_dimorphism_source <- as.factor(ifelse((trimws(as.character(dt2$sexual_dimorphism_source))==trimws("NA")),NA,as.character(dt2$sexual_dimorphism_source)))
+dt2$min_longevity <- ifelse((trimws(as.character(dt2$min_longevity))==trimws("NA")),NA,dt2$min_longevity)               
+suppressWarnings(dt2$min_longevity <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$min_longevity))==as.character(as.numeric("NA"))),NA,dt2$min_longevity))
+dt2$max_longevity <- ifelse((trimws(as.character(dt2$max_longevity))==trimws("NA")),NA,dt2$max_longevity)               
+suppressWarnings(dt2$max_longevity <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$max_longevity))==as.character(as.numeric("NA"))),NA,dt2$max_longevity))
+dt2$longevity_notes <- as.factor(ifelse((trimws(as.character(dt2$longevity_notes))==trimws("NA")),NA,as.character(dt2$longevity_notes)))
+dt2$longevity <- ifelse((trimws(as.character(dt2$longevity))==trimws("NA")),NA,dt2$longevity)               
+suppressWarnings(dt2$longevity <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$longevity))==as.character(as.numeric("NA"))),NA,dt2$longevity))
+dt2$longevity_source <- as.factor(ifelse((trimws(as.character(dt2$longevity_source))==trimws("NA")),NA,as.character(dt2$longevity_source)))
+dt2$longevity_level <- as.factor(ifelse((trimws(as.character(dt2$longevity_level))==trimws("NA")),NA,as.character(dt2$longevity_level)))
+dt2$min_home_range <- ifelse((trimws(as.character(dt2$min_home_range))==trimws("NA")),NA,dt2$min_home_range)               
+suppressWarnings(dt2$min_home_range <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$min_home_range))==as.character(as.numeric("NA"))),NA,dt2$min_home_range))
+dt2$max_home_range <- ifelse((trimws(as.character(dt2$max_home_range))==trimws("NA")),NA,dt2$max_home_range)               
+suppressWarnings(dt2$max_home_range <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$max_home_range))==as.character(as.numeric("NA"))),NA,dt2$max_home_range))
+dt2$home_range_size <- ifelse((trimws(as.character(dt2$home_range_size))==trimws("NA")),NA,dt2$home_range_size)               
+suppressWarnings(dt2$home_range_size <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$home_range_size))==as.character(as.numeric("NA"))),NA,dt2$home_range_size))
+dt2$home_range_notes <- as.factor(ifelse((trimws(as.character(dt2$home_range_notes))==trimws("NA")),NA,as.character(dt2$home_range_notes)))
+dt2$home_range_level <- as.factor(ifelse((trimws(as.character(dt2$home_range_level))==trimws("NA")),NA,as.character(dt2$home_range_level)))
+dt2$home_range_source <- as.factor(ifelse((trimws(as.character(dt2$home_range_source))==trimws("NA")),NA,as.character(dt2$home_range_source)))
+dt2$habitat_specialization <- as.factor(ifelse((trimws(as.character(dt2$habitat_specialization))==trimws("NA")),NA,as.character(dt2$habitat_specialization)))
+dt2$habitat_specialization_notes <- as.factor(ifelse((trimws(as.character(dt2$habitat_specialization_notes))==trimws("NA")),NA,as.character(dt2$habitat_specialization_notes)))
+dt2$habitat_specialization_source <- as.factor(ifelse((trimws(as.character(dt2$habitat_specialization_source))==trimws("NA")),NA,as.character(dt2$habitat_specialization_source)))
+dt2$generation_time <- ifelse((trimws(as.character(dt2$generation_time))==trimws("NA")),NA,dt2$generation_time)               
+suppressWarnings(dt2$generation_time <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$generation_time))==as.character(as.numeric("NA"))),NA,dt2$generation_time))
+dt2$generation_time_notes <- as.factor(ifelse((trimws(as.character(dt2$generation_time_notes))==trimws("NA")),NA,as.character(dt2$generation_time_notes)))
+dt2$generation_time_source <- as.factor(ifelse((trimws(as.character(dt2$generation_time_source))==trimws("NA")),NA,as.character(dt2$generation_time_source)))
+dt2$habitat_suitability_montane <- as.factor(ifelse((trimws(as.character(dt2$habitat_suitability_montane))==trimws("NA")),NA,as.character(dt2$habitat_suitability_montane)))
+dt2$habitat_suitability_lowland <- as.factor(ifelse((trimws(as.character(dt2$habitat_suitability_lowland))==trimws("NA")),NA,as.character(dt2$habitat_suitability_lowland)))
+dt2$habitat_season_montane <- as.factor(ifelse((trimws(as.character(dt2$habitat_season_montane))==trimws("NA")),NA,as.character(dt2$habitat_season_montane)))
+dt2$habitat_season_lowland <- as.factor(ifelse((trimws(as.character(dt2$habitat_season_lowland))==trimws("NA")),NA,as.character(dt2$habitat_season_lowland)))
+dt2$habitat_major_importance_montane <- as.factor(ifelse((trimws(as.character(dt2$habitat_major_importance_montane))==trimws("NA")),NA,as.character(dt2$habitat_major_importance_montane)))
+dt2$habitat_major_importance_lowland <- as.factor(ifelse((trimws(as.character(dt2$habitat_major_importance_lowland))==trimws("NA")),NA,as.character(dt2$habitat_major_importance_lowland)))
+dt2$habitat_breadth <- ifelse((trimws(as.character(dt2$habitat_breadth))==trimws("NA")),NA,dt2$habitat_breadth)               
+suppressWarnings(dt2$habitat_breadth <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$habitat_breadth))==as.character(as.numeric("NA"))),NA,dt2$habitat_breadth))
+dt2$habitat_level <- as.factor(ifelse((trimws(as.character(dt2$habitat_level))==trimws("NA")),NA,as.character(dt2$habitat_level)))
+dt2$mean_CHELSA_bio1_1981.2010_V.2.1 <- ifelse((trimws(as.character(dt2$mean_CHELSA_bio1_1981.2010_V.2.1))==trimws("NA")),NA,dt2$mean_CHELSA_bio1_1981.2010_V.2.1)               
+suppressWarnings(dt2$mean_CHELSA_bio1_1981.2010_V.2.1 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$mean_CHELSA_bio1_1981.2010_V.2.1))==as.character(as.numeric("NA"))),NA,dt2$mean_CHELSA_bio1_1981.2010_V.2.1))
+dt2$mean_CHELSA_bio12_1981.2010_V.2.1 <- ifelse((trimws(as.character(dt2$mean_CHELSA_bio12_1981.2010_V.2.1))==trimws("NA")),NA,dt2$mean_CHELSA_bio12_1981.2010_V.2.1)               
+suppressWarnings(dt2$mean_CHELSA_bio12_1981.2010_V.2.1 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$mean_CHELSA_bio12_1981.2010_V.2.1))==as.character(as.numeric("NA"))),NA,dt2$mean_CHELSA_bio12_1981.2010_V.2.1))
+dt2$mean_human_fp_range_2010 <- ifelse((trimws(as.character(dt2$mean_human_fp_range_2010))==trimws("NA")),NA,dt2$mean_human_fp_range_2010)               
+suppressWarnings(dt2$mean_human_fp_range_2010 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$mean_human_fp_range_2010))==as.character(as.numeric("NA"))),NA,dt2$mean_human_fp_range_2010))
+dt2$mean_human_fp_range_2020 <- ifelse((trimws(as.character(dt2$mean_human_fp_range_2020))==trimws("NA")),NA,dt2$mean_human_fp_range_2020)               
+suppressWarnings(dt2$mean_human_fp_range_2020 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$mean_human_fp_range_2020))==as.character(as.numeric("NA"))),NA,dt2$mean_human_fp_range_2020))
+dt2$percent_change_hf_2010_2020 <- ifelse((trimws(as.character(dt2$percent_change_hf_2010_2020))==trimws("NA")),NA,dt2$percent_change_hf_2010_2020)               
+suppressWarnings(dt2$percent_change_hf_2010_2020 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$percent_change_hf_2010_2020))==as.character(as.numeric("NA"))),NA,dt2$percent_change_hf_2010_2020))
+dt2$observed_range_sqkm <- ifelse((trimws(as.character(dt2$observed_range_sqkm))==trimws("NA")),NA,dt2$observed_range_sqkm)               
+suppressWarnings(dt2$observed_range_sqkm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$observed_range_sqkm))==as.character(as.numeric("NA"))),NA,dt2$observed_range_sqkm))
+dt2$inferred_range_sqkm <- ifelse((trimws(as.character(dt2$inferred_range_sqkm))==trimws("NA")),NA,dt2$inferred_range_sqkm)               
+suppressWarnings(dt2$inferred_range_sqkm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$inferred_range_sqkm))==as.character(as.numeric("NA"))),NA,dt2$inferred_range_sqkm))
+dt2$date_data_obtained <- as.factor(ifelse((trimws(as.character(dt2$date_data_obtained))==trimws("NA")),NA,as.character(dt2$date_data_obtained)))
+
+# Here is the structure of the input data frame:
+str(dt2)                            
+attach(dt2)                            
+# The analyses below are basic descriptions of the variables. After testing, they should be replaced.
+
+summary(IUCN_species_name)
+summary(taxonid)
+summary(genus)
+summary(species)
+summary(common_name)
+summary(elton_species_name)
+summary(taxonomic_disparity)
+summary(pass_non_pass_e)
+summary(order_e)
+summary(family_e)
+summary(family_english_e)
+summary(diet_inv_e)
+summary(diet_vend_e)
+summary(diet_vect_e)
+summary(diet_vfish_e)
+summary(diet_vunk_e)
+summary(diet_scav_e)
+summary(diet_fruit_e)
+summary(diet_nect_e)
+summary(diet_seed_e)
+summary(diet_plant_e)
+summary(diet_cat_e)
+summary(diet_source_e)
+summary(diet_certainty_e)
+summary(diet_breadth)
+summary(diet_level)
+summary(for_strat_water)
+summary(for_strat_ground_e)
+summary(for_strat_understory_e)
+summary(for_strat_midhigh_e)
+summary(for_strat_canopy_e)
+summary(for_strat_aerial_e)
+summary(for_strat_spec_level)
+summary(for_strat_source_e)
+summary(activity_nocturnal_e)
+summary(body_mass_e)
+summary(body_mass_notes)
+summary(body_mass_level_e)
+summary(body_mass_source_e)
+summary(min_body_size_mm)
+summary(max_body_size_mm)
+summary(body_size_mm)
+summary(body_size_notes)
+summary(body_size_level)
+summary(body_size_source)
+summary(sexual_dimorphism)
+summary(sexual_dimorphism_notes)
+summary(sexual_dimorphism_level)
+summary(sexual_dimorphism_source)
+summary(min_longevity)
+summary(max_longevity)
+summary(longevity_notes)
+summary(longevity)
+summary(longevity_source)
+summary(longevity_level)
+summary(min_home_range)
+summary(max_home_range)
+summary(home_range_size)
+summary(home_range_notes)
+summary(home_range_level)
+summary(home_range_source)
+summary(habitat_specialization)
+summary(habitat_specialization_notes)
+summary(habitat_specialization_source)
+summary(generation_time)
+summary(generation_time_level)
+summary(generation_time_notes)
+summary(generation_time_source)
+summary(IUCN_category)
+summary(habitat)
+summary(habitat_suitability_montane)
+summary(habitat_suitability_lowland)
+summary(habitat_season_montane)
+summary(habitat_season_lowland)
+summary(habitat_major_importance_montane)
+summary(habitat_major_importance_lowland)
+summary(habitat_breadth)
+summary(habitat_breadth_source)
+summary(habitat_level)
+summary(mean_CHELSA_bio1_1981.2010_V.2.1)
+summary(mean_CHELSA_bio12_1981.2010_V.2.1)
+summary(mean_human_fp_range_2010)
+summary(mean_human_fp_range_2020)
+summary(percent_change_hf_2010_2020)
+summary(observed_range_sqkm)
+summary(inferred_range_sqkm)
+summary(date_data_obtained)
+summary(filled_by) 
+                # Get more details on character variables
+
+summary(as.factor(dt2$IUCN_species_name)) 
+summary(as.factor(dt2$genus)) 
+summary(as.factor(dt2$species)) 
+summary(as.factor(dt2$common_name)) 
+summary(as.factor(dt2$elton_species_name)) 
+summary(as.factor(dt2$taxonomic_disparity)) 
+summary(as.factor(dt2$pass_non_pass_e)) 
+summary(as.factor(dt2$order_e)) 
+summary(as.factor(dt2$family_e)) 
+summary(as.factor(dt2$family_english_e)) 
+summary(as.factor(dt2$diet_cat_e)) 
+summary(as.factor(dt2$diet_source_e)) 
+summary(as.factor(dt2$diet_certainty_e)) 
+summary(as.factor(dt2$diet_level)) 
+summary(as.factor(dt2$for_strat_spec_level)) 
+summary(as.factor(dt2$for_strat_source_e)) 
+summary(as.factor(dt2$activity_nocturnal_e)) 
+summary(as.factor(dt2$body_mass_notes)) 
+summary(as.factor(dt2$body_mass_level_e)) 
+summary(as.factor(dt2$body_mass_source_e)) 
+summary(as.factor(dt2$body_size_notes)) 
+summary(as.factor(dt2$body_size_level)) 
+summary(as.factor(dt2$body_size_source)) 
+summary(as.factor(dt2$sexual_dimorphism)) 
+summary(as.factor(dt2$sexual_dimorphism_notes)) 
+summary(as.factor(dt2$sexual_dimorphism_level)) 
+summary(as.factor(dt2$sexual_dimorphism_source)) 
+summary(as.factor(dt2$longevity_notes)) 
+summary(as.factor(dt2$longevity_source)) 
+summary(as.factor(dt2$longevity_level)) 
+summary(as.factor(dt2$home_range_notes)) 
+summary(as.factor(dt2$home_range_level)) 
+summary(as.factor(dt2$home_range_source)) 
+summary(as.factor(dt2$habitat_specialization)) 
+summary(as.factor(dt2$habitat_specialization_notes)) 
+summary(as.factor(dt2$habitat_specialization_source)) 
+summary(as.factor(dt2$generation_time_level)) 
+summary(as.factor(dt2$generation_time_notes)) 
+summary(as.factor(dt2$generation_time_source)) 
+summary(as.factor(dt2$IUCN_category)) 
+summary(as.factor(dt2$habitat)) 
+summary(as.factor(dt2$habitat_suitability_montane)) 
+summary(as.factor(dt2$habitat_suitability_lowland)) 
+summary(as.factor(dt2$habitat_season_montane)) 
+summary(as.factor(dt2$habitat_season_lowland)) 
+summary(as.factor(dt2$habitat_major_importance_montane)) 
+summary(as.factor(dt2$habitat_major_importance_lowland)) 
+summary(as.factor(dt2$habitat_breadth_source)) 
+summary(as.factor(dt2$habitat_level)) 
+summary(as.factor(dt2$date_data_obtained)) 
+summary(as.factor(dt2$filled_by))
+detach(dt2)
+
+inUrl3  <- "https://pasta.lternet.edu/package/data/eml/edi/1220/6/3c655f2ab1d525d1b1f05ee78153e875" 
+infile3 <- tempfile()
+try(download.file(inUrl3,infile3,method="curl"))
+if (is.na(file.size(infile3))) download.file(inUrl3,infile3,method="auto")
+
+ dt3 <-read.csv(infile3,header=F 
+          ,skip=1
+            ,sep=","  
+                ,quot='"' 
+        , col.names=c(
+                    "IUCN_species_name",     
+                    "common_name",     
+                    "family",     
+                    "genus",     
+                    "species",     
+                    "subspecies",     
+                    "elton_species_name",     
+                    "diet_cat",     
+                    "diet_source_e",     
+                    "diet_breadth",     
+                    "diet_level",     
+                    "activity_nocturnal_e",     
+                    "activity_source_e",     
+                    "activity_level",     
+                    "for_strat_value_e",     
+                    "for_strat_certainty_e",     
+                    "body_mass_e",     
+                    "body_mass_level_e",     
+                    "body_mass_source_e",     
+                    "body_size_mm",     
+                    "body_size_notes",     
+                    "body_size_level",     
+                    "body_size_source",     
+                    "sexual_dimorphism",     
+                    "sexual_dimorphism_notes",     
+                    "sexual_dimorphism_level",     
+                    "sexual_dimorphism_source",     
+                    "longevity",     
+                    "longevity_notes",     
+                    "longevity_level",     
+                    "longevity_source",     
+                    "home_range_size",     
+                    "home_range_notes",     
+                    "home_range_level",     
+                    "home_range_source",     
+                    "habitat_specialization",     
+                    "habitat_specialization_source",     
+                    "generation_time",     
+                    "generation_time_notes",     
+                    "generation_time_level",     
+                    "generation_time_source",     
+                    "IUCN_category",     
+                    "habitat",     
+                    "habitat_breadth",     
+                    "habitat_breadth_source",     
+                    "habitat_level",     
+                    "mean_CHELSA_bio1_1981.2010_V.2.1",     
+                    "mean_CHELSA_bio12_1981.2010_V.2.1",     
+                    "mean_human_fp_range_2010",     
+                    "mean_human_fp_range_2020",     
+                    "percent_change_hf_2010_2020",     
+                    "inferred_range_sqkm",     
+                    "date_data_obtained",     
+                    "filled_by"    ), check.names=TRUE)
+
+unlink(infile3)
+
+# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
+
+if (class(dt3$IUCN_species_name)!="factor") dt3$IUCN_species_name<- as.factor(dt3$IUCN_species_name)
+if (class(dt3$common_name)!="factor") dt3$common_name<- as.factor(dt3$common_name)
+if (class(dt3$family)!="factor") dt3$family<- as.factor(dt3$family)
+if (class(dt3$genus)!="factor") dt3$genus<- as.factor(dt3$genus)
+if (class(dt3$species)!="factor") dt3$species<- as.factor(dt3$species)
+if (class(dt3$subspecies)!="factor") dt3$subspecies<- as.factor(dt3$subspecies)
+if (class(dt3$elton_species_name)!="factor") dt3$elton_species_name<- as.factor(dt3$elton_species_name)
+if (class(dt3$diet_cat)!="factor") dt3$diet_cat<- as.factor(dt3$diet_cat)
+if (class(dt3$diet_source_e)!="factor") dt3$diet_source_e<- as.factor(dt3$diet_source_e)
+if (class(dt3$diet_breadth)=="factor") dt3$diet_breadth <-as.numeric(levels(dt3$diet_breadth))[as.integer(dt3$diet_breadth) ]               
+if (class(dt3$diet_breadth)=="character") dt3$diet_breadth <-as.numeric(dt3$diet_breadth)
+if (class(dt3$diet_level)!="factor") dt3$diet_level<- as.factor(dt3$diet_level)
+if (class(dt3$activity_nocturnal_e)!="factor") dt3$activity_nocturnal_e<- as.factor(dt3$activity_nocturnal_e)
+if (class(dt3$activity_source_e)!="factor") dt3$activity_source_e<- as.factor(dt3$activity_source_e)
+if (class(dt3$activity_level)!="factor") dt3$activity_level<- as.factor(dt3$activity_level)
+if (class(dt3$for_strat_value_e)!="factor") dt3$for_strat_value_e<- as.factor(dt3$for_strat_value_e)
+if (class(dt3$for_strat_certainty_e)!="factor") dt3$for_strat_certainty_e<- as.factor(dt3$for_strat_certainty_e)
+if (class(dt3$body_mass_e)=="factor") dt3$body_mass_e <-as.numeric(levels(dt3$body_mass_e))[as.integer(dt3$body_mass_e) ]               
+if (class(dt3$body_mass_e)=="character") dt3$body_mass_e <-as.numeric(dt3$body_mass_e)
+if (class(dt3$body_mass_level_e)!="factor") dt3$body_mass_level_e<- as.factor(dt3$body_mass_level_e)
+if (class(dt3$body_mass_source_e)!="factor") dt3$body_mass_source_e<- as.factor(dt3$body_mass_source_e)
+if (class(dt3$body_size_mm)=="factor") dt3$body_size_mm <-as.numeric(levels(dt3$body_size_mm))[as.integer(dt3$body_size_mm) ]               
+if (class(dt3$body_size_mm)=="character") dt3$body_size_mm <-as.numeric(dt3$body_size_mm)
+if (class(dt3$body_size_notes)!="factor") dt3$body_size_notes<- as.factor(dt3$body_size_notes)
+if (class(dt3$body_size_level)!="factor") dt3$body_size_level<- as.factor(dt3$body_size_level)
+if (class(dt3$body_size_source)!="factor") dt3$body_size_source<- as.factor(dt3$body_size_source)
+if (class(dt3$sexual_dimorphism)!="factor") dt3$sexual_dimorphism<- as.factor(dt3$sexual_dimorphism)
+if (class(dt3$sexual_dimorphism_notes)!="factor") dt3$sexual_dimorphism_notes<- as.factor(dt3$sexual_dimorphism_notes)
+if (class(dt3$sexual_dimorphism_level)!="factor") dt3$sexual_dimorphism_level<- as.factor(dt3$sexual_dimorphism_level)
+if (class(dt3$sexual_dimorphism_source)!="factor") dt3$sexual_dimorphism_source<- as.factor(dt3$sexual_dimorphism_source)
+if (class(dt3$longevity)=="factor") dt3$longevity <-as.numeric(levels(dt3$longevity))[as.integer(dt3$longevity) ]               
+if (class(dt3$longevity)=="character") dt3$longevity <-as.numeric(dt3$longevity)
+if (class(dt3$longevity_notes)!="factor") dt3$longevity_notes<- as.factor(dt3$longevity_notes)
+if (class(dt3$longevity_level)!="factor") dt3$longevity_level<- as.factor(dt3$longevity_level)
+if (class(dt3$longevity_source)!="factor") dt3$longevity_source<- as.factor(dt3$longevity_source)
+if (class(dt3$home_range_size)=="factor") dt3$home_range_size <-as.numeric(levels(dt3$home_range_size))[as.integer(dt3$home_range_size) ]               
+if (class(dt3$home_range_size)=="character") dt3$home_range_size <-as.numeric(dt3$home_range_size)
+if (class(dt3$home_range_notes)!="factor") dt3$home_range_notes<- as.factor(dt3$home_range_notes)
+if (class(dt3$home_range_level)!="factor") dt3$home_range_level<- as.factor(dt3$home_range_level)
+if (class(dt3$home_range_source)!="factor") dt3$home_range_source<- as.factor(dt3$home_range_source)
+if (class(dt3$habitat_specialization)!="factor") dt3$habitat_specialization<- as.factor(dt3$habitat_specialization)
+if (class(dt3$habitat_specialization_source)!="factor") dt3$habitat_specialization_source<- as.factor(dt3$habitat_specialization_source)
+if (class(dt3$generation_time)=="factor") dt3$generation_time <-as.numeric(levels(dt3$generation_time))[as.integer(dt3$generation_time) ]               
+if (class(dt3$generation_time)=="character") dt3$generation_time <-as.numeric(dt3$generation_time)
+if (class(dt3$generation_time_notes)!="factor") dt3$generation_time_notes<- as.factor(dt3$generation_time_notes)
+if (class(dt3$generation_time_level)!="factor") dt3$generation_time_level<- as.factor(dt3$generation_time_level)
+if (class(dt3$generation_time_source)!="factor") dt3$generation_time_source<- as.factor(dt3$generation_time_source)
+if (class(dt3$IUCN_category)!="factor") dt3$IUCN_category<- as.factor(dt3$IUCN_category)
+if (class(dt3$habitat)!="factor") dt3$habitat<- as.factor(dt3$habitat)
+if (class(dt3$habitat_breadth)=="factor") dt3$habitat_breadth <-as.numeric(levels(dt3$habitat_breadth))[as.integer(dt3$habitat_breadth) ]               
+if (class(dt3$habitat_breadth)=="character") dt3$habitat_breadth <-as.numeric(dt3$habitat_breadth)
+if (class(dt3$habitat_breadth_source)!="factor") dt3$habitat_breadth_source<- as.factor(dt3$habitat_breadth_source)
+if (class(dt3$habitat_level)!="factor") dt3$habitat_level<- as.factor(dt3$habitat_level)
+if (class(dt3$mean_CHELSA_bio1_1981.2010_V.2.1)=="factor") dt3$mean_CHELSA_bio1_1981.2010_V.2.1 <-as.numeric(levels(dt3$mean_CHELSA_bio1_1981.2010_V.2.1))[as.integer(dt3$mean_CHELSA_bio1_1981.2010_V.2.1) ]
+if (class(dt3$mean_CHELSA_bio1_1981.2010_V.2.1)=="character") dt3$mean_CHELSA_bio1_1981.2010_V.2.1 <-as.numeric(dt3$mean_CHELSA_bio1_1981.2010_V.2.1)
+if (class(dt3$mean_CHELSA_bio12_1981.2010_V.2.1)=="factor") dt3$mean_CHELSA_bio12_1981.2010_V.2.1 <-as.numeric(levels(dt3$mean_CHELSA_bio12_1981.2010_V.2.1))[as.integer(dt3$mean_CHELSA_bio12_1981.2010_V.2.1) ]
+if (class(dt3$mean_CHELSA_bio12_1981.2010_V.2.1)=="character") dt3$mean_CHELSA_bio12_1981.2010_V.2.1 <-as.numeric(dt3$mean_CHELSA_bio12_1981.2010_V.2.1)
+if (class(dt3$mean_human_fp_range_2010)=="factor") dt3$mean_human_fp_range_2010 <-as.numeric(levels(dt3$mean_human_fp_range_2010))[as.integer(dt3$mean_human_fp_range_2010) ]               
+if (class(dt3$mean_human_fp_range_2010)=="character") dt3$mean_human_fp_range_2010 <-as.numeric(dt3$mean_human_fp_range_2010)
+if (class(dt3$mean_human_fp_range_2020)=="factor") dt3$mean_human_fp_range_2020 <-as.numeric(levels(dt3$mean_human_fp_range_2020))[as.integer(dt3$mean_human_fp_range_2020) ]               
+if (class(dt3$mean_human_fp_range_2020)=="character") dt3$mean_human_fp_range_2020 <-as.numeric(dt3$mean_human_fp_range_2020)
+if (class(dt3$percent_change_hf_2010_2020)=="factor") dt3$percent_change_hf_2010_2020 <-as.numeric(levels(dt3$percent_change_hf_2010_2020))[as.integer(dt3$percent_change_hf_2010_2020) ]               
+if (class(dt3$percent_change_hf_2010_2020)=="character") dt3$percent_change_hf_2010_2020 <-as.numeric(dt3$percent_change_hf_2010_2020)
+if (class(dt3$inferred_range_sqkm)=="factor") dt3$inferred_range_sqkm <-as.numeric(levels(dt3$inferred_range_sqkm))[as.integer(dt3$inferred_range_sqkm) ]               
+if (class(dt3$inferred_range_sqkm)=="character") dt3$inferred_range_sqkm <-as.numeric(dt3$inferred_range_sqkm)
+if (class(dt3$date_data_obtained)!="factor") dt3$date_data_obtained<- as.factor(dt3$date_data_obtained)
+if (class(dt3$filled_by)!="factor") dt3$filled_by<- as.factor(dt3$filled_by)
+
+# Convert Missing Values to NA for non-dates
+
+dt3$common_name <- as.factor(ifelse((trimws(as.character(dt3$common_name))==trimws("NA")),NA,as.character(dt3$common_name)))
+dt3$subspecies <- as.factor(ifelse((trimws(as.character(dt3$subspecies))==trimws("NA")),NA,as.character(dt3$subspecies)))
+dt3$diet_level <- as.factor(ifelse((trimws(as.character(dt3$diet_level))==trimws("NA")),NA,as.character(dt3$diet_level)))
+dt3$activity_level <- as.factor(ifelse((trimws(as.character(dt3$activity_level))==trimws("NA")),NA,as.character(dt3$activity_level)))
+dt3$body_size_mm <- ifelse((trimws(as.character(dt3$body_size_mm))==trimws("NA")),NA,dt3$body_size_mm)               
+suppressWarnings(dt3$body_size_mm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt3$body_size_mm))==as.character(as.numeric("NA"))),NA,dt3$body_size_mm))
+dt3$body_size_notes <- as.factor(ifelse((trimws(as.character(dt3$body_size_notes))==trimws("NA")),NA,as.character(dt3$body_size_notes)))
+dt3$body_size_level <- as.factor(ifelse((trimws(as.character(dt3$body_size_level))==trimws("NA")),NA,as.character(dt3$body_size_level)))
+dt3$body_size_source <- as.factor(ifelse((trimws(as.character(dt3$body_size_source))==trimws("NA")),NA,as.character(dt3$body_size_source)))
+dt3$sexual_dimorphism <- as.factor(ifelse((trimws(as.character(dt3$sexual_dimorphism))==trimws("NA")),NA,as.character(dt3$sexual_dimorphism)))
+dt3$sexual_dimorphism <- as.factor(ifelse((trimws(as.character(dt3$sexual_dimorphism))==trimws("NA ")),NA,as.character(dt3$sexual_dimorphism)))
+dt3$sexual_dimorphism_notes <- as.factor(ifelse((trimws(as.character(dt3$sexual_dimorphism_notes))==trimws("NA")),NA,as.character(dt3$sexual_dimorphism_notes)))
+dt3$sexual_dimorphism_level <- as.factor(ifelse((trimws(as.character(dt3$sexual_dimorphism_level))==trimws("NA")),NA,as.character(dt3$sexual_dimorphism_level)))
+dt3$sexual_dimorphism_source <- as.factor(ifelse((trimws(as.character(dt3$sexual_dimorphism_source))==trimws("NA")),NA,as.character(dt3$sexual_dimorphism_source)))
+dt3$longevity <- ifelse((trimws(as.character(dt3$longevity))==trimws("NA")),NA,dt3$longevity)               
+suppressWarnings(dt3$longevity <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt3$longevity))==as.character(as.numeric("NA"))),NA,dt3$longevity))
+dt3$longevity_notes <- as.factor(ifelse((trimws(as.character(dt3$longevity_notes))==trimws("NA")),NA,as.character(dt3$longevity_notes)))
+dt3$longevity_level <- as.factor(ifelse((trimws(as.character(dt3$longevity_level))==trimws("NA")),NA,as.character(dt3$longevity_level)))
+dt3$longevity_source <- as.factor(ifelse((trimws(as.character(dt3$longevity_source))==trimws("NA")),NA,as.character(dt3$longevity_source)))
+dt3$home_range_size <- ifelse((trimws(as.character(dt3$home_range_size))==trimws("NA")),NA,dt3$home_range_size)               
+suppressWarnings(dt3$home_range_size <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt3$home_range_size))==as.character(as.numeric("NA"))),NA,dt3$home_range_size))
+dt3$home_range_notes <- as.factor(ifelse((trimws(as.character(dt3$home_range_notes))==trimws("NA")),NA,as.character(dt3$home_range_notes)))
+dt3$home_range_level <- as.factor(ifelse((trimws(as.character(dt3$home_range_level))==trimws("NA")),NA,as.character(dt3$home_range_level)))
+dt3$home_range_source <- as.factor(ifelse((trimws(as.character(dt3$home_range_source))==trimws("NA")),NA,as.character(dt3$home_range_source)))
+dt3$habitat_specialization <- as.factor(ifelse((trimws(as.character(dt3$habitat_specialization))==trimws("NA")),NA,as.character(dt3$habitat_specialization)))
+dt3$habitat_specialization_source <- as.factor(ifelse((trimws(as.character(dt3$habitat_specialization_source))==trimws("NA")),NA,as.character(dt3$habitat_specialization_source)))
+dt3$generation_time <- ifelse((trimws(as.character(dt3$generation_time))==trimws("NA")),NA,dt3$generation_time)               
+suppressWarnings(dt3$generation_time <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt3$generation_time))==as.character(as.numeric("NA"))),NA,dt3$generation_time))
+dt3$generation_time_notes <- as.factor(ifelse((trimws(as.character(dt3$generation_time_notes))==trimws("NA")),NA,as.character(dt3$generation_time_notes)))
+dt3$generation_time_level <- as.factor(ifelse((trimws(as.character(dt3$generation_time_level))==trimws("NA")),NA,as.character(dt3$generation_time_level)))
+dt3$generation_time_source <- as.factor(ifelse((trimws(as.character(dt3$generation_time_source))==trimws("NA")),NA,as.character(dt3$generation_time_source)))
+dt3$habitat_breadth <- ifelse((trimws(as.character(dt3$habitat_breadth))==trimws("NA")),NA,dt3$habitat_breadth)               
+suppressWarnings(dt3$habitat_breadth <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt3$habitat_breadth))==as.character(as.numeric("NA"))),NA,dt3$habitat_breadth))
+dt3$habitat_level <- as.factor(ifelse((trimws(as.character(dt3$habitat_level))==trimws("NA")),NA,as.character(dt3$habitat_level)))
+dt3$mean_CHELSA_bio1_1981.2010_V.2.1 <- ifelse((trimws(as.character(dt3$mean_CHELSA_bio1_1981.2010_V.2.1))==trimws("NA")),NA,dt3$mean_CHELSA_bio1_1981.2010_V.2.1)               
+suppressWarnings(dt3$mean_CHELSA_bio1_1981.2010_V.2.1 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt3$mean_CHELSA_bio1_1981.2010_V.2.1))==as.character(as.numeric("NA"))),NA,dt3$mean_CHELSA_bio1_1981.2010_V.2.1))
+dt3$mean_CHELSA_bio12_1981.2010_V.2.1 <- ifelse((trimws(as.character(dt3$mean_CHELSA_bio12_1981.2010_V.2.1))==trimws("NA")),NA,dt3$mean_CHELSA_bio12_1981.2010_V.2.1)               
+suppressWarnings(dt3$mean_CHELSA_bio12_1981.2010_V.2.1 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt3$mean_CHELSA_bio12_1981.2010_V.2.1))==as.character(as.numeric("NA"))),NA,dt3$mean_CHELSA_bio12_1981.2010_V.2.1))
+dt3$mean_human_fp_range_2010 <- ifelse((trimws(as.character(dt3$mean_human_fp_range_2010))==trimws("NA")),NA,dt3$mean_human_fp_range_2010)               
+suppressWarnings(dt3$mean_human_fp_range_2010 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt3$mean_human_fp_range_2010))==as.character(as.numeric("NA"))),NA,dt3$mean_human_fp_range_2010))
+dt3$mean_human_fp_range_2020 <- ifelse((trimws(as.character(dt3$mean_human_fp_range_2020))==trimws("NA")),NA,dt3$mean_human_fp_range_2020)               
+suppressWarnings(dt3$mean_human_fp_range_2020 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt3$mean_human_fp_range_2020))==as.character(as.numeric("NA"))),NA,dt3$mean_human_fp_range_2020))
+dt3$percent_change_hf_2010_2020 <- ifelse((trimws(as.character(dt3$percent_change_hf_2010_2020))==trimws("NA")),NA,dt3$percent_change_hf_2010_2020)               
+suppressWarnings(dt3$percent_change_hf_2010_2020 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt3$percent_change_hf_2010_2020))==as.character(as.numeric("NA"))),NA,dt3$percent_change_hf_2010_2020))
+dt3$inferred_range_sqkm <- ifelse((trimws(as.character(dt3$inferred_range_sqkm))==trimws("NA")),NA,dt3$inferred_range_sqkm)               
+suppressWarnings(dt3$inferred_range_sqkm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt3$inferred_range_sqkm))==as.character(as.numeric("NA"))),NA,dt3$inferred_range_sqkm))
+dt3$date_data_obtained <- as.factor(ifelse((trimws(as.character(dt3$date_data_obtained))==trimws("NA")),NA,as.character(dt3$date_data_obtained)))
+
+# Here is the structure of the input data frame:
+str(dt3)                            
+attach(dt3)                            
+# The analyses below are basic descriptions of the variables. After testing, they should be replaced.
+
+summary(IUCN_species_name)
+summary(common_name)
+summary(family)
+summary(genus)
+summary(species)
+summary(subspecies)
+summary(elton_species_name)
+summary(diet_cat)
+summary(diet_source_e)
+summary(diet_breadth)
+summary(diet_level)
+summary(activity_nocturnal_e)
+summary(activity_source_e)
+summary(activity_level)
+summary(for_strat_value_e)
+summary(for_strat_certainty_e)
+summary(body_mass_e)
+summary(body_mass_level_e)
+summary(body_mass_source_e)
+summary(body_size_mm)
+summary(body_size_notes)
+summary(body_size_level)
+summary(body_size_source)
+summary(sexual_dimorphism)
+summary(sexual_dimorphism_notes)
+summary(sexual_dimorphism_level)
+summary(sexual_dimorphism_source)
+summary(longevity)
+summary(longevity_notes)
+summary(longevity_level)
+summary(longevity_source)
+summary(home_range_size)
+summary(home_range_notes)
+summary(home_range_level)
+summary(home_range_source)
+summary(habitat_specialization)
+summary(habitat_specialization_source)
+summary(generation_time)
+summary(generation_time_notes)
+summary(generation_time_level)
+summary(generation_time_source)
+summary(IUCN_category)
+summary(habitat)
+summary(habitat_breadth)
+summary(habitat_breadth_source)
+summary(habitat_level)
+summary(mean_CHELSA_bio1_1981.2010_V.2.1)
+summary(mean_CHELSA_bio12_1981.2010_V.2.1)
+summary(mean_human_fp_range_2010)
+summary(mean_human_fp_range_2020)
+summary(percent_change_hf_2010_2020)
+summary(inferred_range_sqkm)
+summary(date_data_obtained)
+summary(filled_by) 
+                # Get more details on character variables
+
+summary(as.factor(dt3$IUCN_species_name)) 
+summary(as.factor(dt3$common_name)) 
+summary(as.factor(dt3$family)) 
+summary(as.factor(dt3$genus)) 
+summary(as.factor(dt3$species)) 
+summary(as.factor(dt3$subspecies)) 
+summary(as.factor(dt3$elton_species_name)) 
+summary(as.factor(dt3$diet_cat)) 
+summary(as.factor(dt3$diet_source_e)) 
+summary(as.factor(dt3$diet_level)) 
+summary(as.factor(dt3$activity_nocturnal_e)) 
+summary(as.factor(dt3$activity_source_e)) 
+summary(as.factor(dt3$activity_level)) 
+summary(as.factor(dt3$for_strat_value_e)) 
+summary(as.factor(dt3$for_strat_certainty_e)) 
+summary(as.factor(dt3$body_mass_level_e)) 
+summary(as.factor(dt3$body_mass_source_e)) 
+summary(as.factor(dt3$body_size_notes)) 
+summary(as.factor(dt3$body_size_level)) 
+summary(as.factor(dt3$body_size_source)) 
+summary(as.factor(dt3$sexual_dimorphism)) 
+summary(as.factor(dt3$sexual_dimorphism_notes)) 
+summary(as.factor(dt3$sexual_dimorphism_level)) 
+summary(as.factor(dt3$sexual_dimorphism_source)) 
+summary(as.factor(dt3$longevity_notes)) 
+summary(as.factor(dt3$longevity_level)) 
+summary(as.factor(dt3$longevity_source)) 
+summary(as.factor(dt3$home_range_notes)) 
+summary(as.factor(dt3$home_range_level)) 
+summary(as.factor(dt3$home_range_source)) 
+summary(as.factor(dt3$habitat_specialization)) 
+summary(as.factor(dt3$habitat_specialization_source)) 
+summary(as.factor(dt3$generation_time_notes)) 
+summary(as.factor(dt3$generation_time_level)) 
+summary(as.factor(dt3$generation_time_source)) 
+summary(as.factor(dt3$IUCN_category)) 
+summary(as.factor(dt3$habitat)) 
+summary(as.factor(dt3$habitat_breadth_source)) 
+summary(as.factor(dt3$habitat_level)) 
+summary(as.factor(dt3$date_data_obtained)) 
+summary(as.factor(dt3$filled_by))
+detach(dt3)
+
+inUrl4  <- "https://pasta.lternet.edu/package/data/eml/edi/1220/6/5a86fde71322a1ff64d94ace0ed1982c" 
+infile4 <- tempfile()
+try(download.file(inUrl4,infile4,method="curl"))
+if (is.na(file.size(infile4))) download.file(inUrl4,infile4,method="auto")
+
+ dt4 <-read.csv(infile4,header=F 
+          ,skip=1
+            ,sep=","  
+                ,quot='"' 
+        , col.names=c(
+                    "IUCN_species_name",     
+                    "common_name",     
+                    "family_e",     
+                    "genus",     
+                    "species",     
+                    "elton_species_name",     
+                    "diet_cat_e",     
+                    "diet_source_e",     
+                    "diet_breadth",     
+                    "diet_level",     
+                    "activity_nocturnal_e",     
+                    "for_strat_ground_e",     
+                    "for_strat_understory_e",     
+                    "for_strat_midhigh_e",     
+                    "for_strat_canopy_e",     
+                    "for_strat_aerial_e",     
+                    "for_strat_spec_level",     
+                    "for_strat_source_e",     
+                    "body_mass_e",     
+                    "body_mass_level_e",     
+                    "body_mass_source_e",     
+                    "body_size_mm",     
+                    "body_size_notes",     
+                    "body_size_level",     
+                    "body_size_source",     
+                    "sexual_dimorphism",     
+                    "sexual_dimorphism_notes",     
+                    "sexual_dimorphism_level",     
+                    "sexual_dimorphism_source",     
+                    "longevity",     
+                    "longevity_notes",     
+                    "longevity_level",     
+                    "longevity_source",     
+                    "home_range_size",     
+                    "home_range_notes",     
+                    "home_range_level",     
+                    "home_range_source",     
+                    "habitat_specialization",     
+                    "habitat_specialization_source",     
+                    "generation_time",     
+                    "generation_time_notes",     
+                    "generation_time_level",     
+                    "generation_time_source",     
+                    "IUCN_category",     
+                    "habitat",     
+                    "habitat_breadth",     
+                    "habitat_breadth_source",     
+                    "habitat_level",     
+                    "mean_CHELSA_bio1_1981.2010_V.2.1",     
+                    "mean_CHELSA_bio12_1981.2010_V.2.1",     
+                    "mean_human_fp_range_2010",     
+                    "mean_human_fp_range_2020",     
+                    "percent_change_hf_2010_2020",     
+                    "inferred_range_sqkm",     
+                    "date_data_obtained",     
+                    "filled_by"    ), check.names=TRUE)
+
+unlink(infile4)
+
+# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
+
+if (class(dt4$IUCN_species_name)!="factor") dt4$IUCN_species_name<- as.factor(dt4$IUCN_species_name)
+if (class(dt4$common_name)!="factor") dt4$common_name<- as.factor(dt4$common_name)
+if (class(dt4$family_e)!="factor") dt4$family_e<- as.factor(dt4$family_e)
+if (class(dt4$genus)!="factor") dt4$genus<- as.factor(dt4$genus)
+if (class(dt4$species)!="factor") dt4$species<- as.factor(dt4$species)
+if (class(dt4$elton_species_name)!="factor") dt4$elton_species_name<- as.factor(dt4$elton_species_name)
+if (class(dt4$diet_cat_e)!="factor") dt4$diet_cat_e<- as.factor(dt4$diet_cat_e)
+if (class(dt4$diet_source_e)!="factor") dt4$diet_source_e<- as.factor(dt4$diet_source_e)
+if (class(dt4$diet_breadth)=="factor") dt4$diet_breadth <-as.numeric(levels(dt4$diet_breadth))[as.integer(dt4$diet_breadth) ]               
+if (class(dt4$diet_breadth)=="character") dt4$diet_breadth <-as.numeric(dt4$diet_breadth)
+if (class(dt4$diet_level)!="factor") dt4$diet_level<- as.factor(dt4$diet_level)
+if (class(dt4$activity_nocturnal_e)!="factor") dt4$activity_nocturnal_e<- as.factor(dt4$activity_nocturnal_e)
+if (class(dt4$for_strat_ground_e)=="factor") dt4$for_strat_ground_e <-as.numeric(levels(dt4$for_strat_ground_e))[as.integer(dt4$for_strat_ground_e) ]               
+if (class(dt4$for_strat_ground_e)=="character") dt4$for_strat_ground_e <-as.numeric(dt4$for_strat_ground_e)
+if (class(dt4$for_strat_understory_e)=="factor") dt4$for_strat_understory_e <-as.numeric(levels(dt4$for_strat_understory_e))[as.integer(dt4$for_strat_understory_e) ]               
+if (class(dt4$for_strat_understory_e)=="character") dt4$for_strat_understory_e <-as.numeric(dt4$for_strat_understory_e)
+if (class(dt4$for_strat_midhigh_e)=="factor") dt4$for_strat_midhigh_e <-as.numeric(levels(dt4$for_strat_midhigh_e))[as.integer(dt4$for_strat_midhigh_e) ]               
+if (class(dt4$for_strat_midhigh_e)=="character") dt4$for_strat_midhigh_e <-as.numeric(dt4$for_strat_midhigh_e)
+if (class(dt4$for_strat_canopy_e)=="factor") dt4$for_strat_canopy_e <-as.numeric(levels(dt4$for_strat_canopy_e))[as.integer(dt4$for_strat_canopy_e) ]               
+if (class(dt4$for_strat_canopy_e)=="character") dt4$for_strat_canopy_e <-as.numeric(dt4$for_strat_canopy_e)
+if (class(dt4$for_strat_aerial_e)=="factor") dt4$for_strat_aerial_e <-as.numeric(levels(dt4$for_strat_aerial_e))[as.integer(dt4$for_strat_aerial_e) ]               
+if (class(dt4$for_strat_aerial_e)=="character") dt4$for_strat_aerial_e <-as.numeric(dt4$for_strat_aerial_e)
+if (class(dt4$for_strat_spec_level)!="factor") dt4$for_strat_spec_level<- as.factor(dt4$for_strat_spec_level)
+if (class(dt4$for_strat_source_e)!="factor") dt4$for_strat_source_e<- as.factor(dt4$for_strat_source_e)
+if (class(dt4$body_mass_e)=="factor") dt4$body_mass_e <-as.numeric(levels(dt4$body_mass_e))[as.integer(dt4$body_mass_e) ]               
+if (class(dt4$body_mass_e)=="character") dt4$body_mass_e <-as.numeric(dt4$body_mass_e)
+if (class(dt4$body_mass_level_e)!="factor") dt4$body_mass_level_e<- as.factor(dt4$body_mass_level_e)
+if (class(dt4$body_mass_source_e)!="factor") dt4$body_mass_source_e<- as.factor(dt4$body_mass_source_e)
+if (class(dt4$body_size_mm)=="factor") dt4$body_size_mm <-as.numeric(levels(dt4$body_size_mm))[as.integer(dt4$body_size_mm) ]               
+if (class(dt4$body_size_mm)=="character") dt4$body_size_mm <-as.numeric(dt4$body_size_mm)
+if (class(dt4$body_size_notes)!="factor") dt4$body_size_notes<- as.factor(dt4$body_size_notes)
+if (class(dt4$body_size_level)!="factor") dt4$body_size_level<- as.factor(dt4$body_size_level)
+if (class(dt4$body_size_source)!="factor") dt4$body_size_source<- as.factor(dt4$body_size_source)
+if (class(dt4$sexual_dimorphism)!="factor") dt4$sexual_dimorphism<- as.factor(dt4$sexual_dimorphism)
+if (class(dt4$sexual_dimorphism_notes)!="factor") dt4$sexual_dimorphism_notes<- as.factor(dt4$sexual_dimorphism_notes)
+if (class(dt4$sexual_dimorphism_level)!="factor") dt4$sexual_dimorphism_level<- as.factor(dt4$sexual_dimorphism_level)
+if (class(dt4$sexual_dimorphism_source)!="factor") dt4$sexual_dimorphism_source<- as.factor(dt4$sexual_dimorphism_source)
+if (class(dt4$longevity)=="factor") dt4$longevity <-as.numeric(levels(dt4$longevity))[as.integer(dt4$longevity) ]               
+if (class(dt4$longevity)=="character") dt4$longevity <-as.numeric(dt4$longevity)
+if (class(dt4$longevity_notes)!="factor") dt4$longevity_notes<- as.factor(dt4$longevity_notes)
+if (class(dt4$longevity_level)!="factor") dt4$longevity_level<- as.factor(dt4$longevity_level)
+if (class(dt4$longevity_source)!="factor") dt4$longevity_source<- as.factor(dt4$longevity_source)
+if (class(dt4$home_range_size)=="factor") dt4$home_range_size <-as.numeric(levels(dt4$home_range_size))[as.integer(dt4$home_range_size) ]               
+if (class(dt4$home_range_size)=="character") dt4$home_range_size <-as.numeric(dt4$home_range_size)
+if (class(dt4$home_range_notes)!="factor") dt4$home_range_notes<- as.factor(dt4$home_range_notes)
+if (class(dt4$home_range_level)!="factor") dt4$home_range_level<- as.factor(dt4$home_range_level)
+if (class(dt4$home_range_source)!="factor") dt4$home_range_source<- as.factor(dt4$home_range_source)
+if (class(dt4$habitat_specialization)!="factor") dt4$habitat_specialization<- as.factor(dt4$habitat_specialization)
+if (class(dt4$habitat_specialization_source)!="factor") dt4$habitat_specialization_source<- as.factor(dt4$habitat_specialization_source)
+if (class(dt4$generation_time)=="factor") dt4$generation_time <-as.numeric(levels(dt4$generation_time))[as.integer(dt4$generation_time) ]               
+if (class(dt4$generation_time)=="character") dt4$generation_time <-as.numeric(dt4$generation_time)
+if (class(dt4$generation_time_notes)!="factor") dt4$generation_time_notes<- as.factor(dt4$generation_time_notes)
+if (class(dt4$generation_time_level)!="factor") dt4$generation_time_level<- as.factor(dt4$generation_time_level)
+if (class(dt4$generation_time_source)!="factor") dt4$generation_time_source<- as.factor(dt4$generation_time_source)
+if (class(dt4$IUCN_category)!="factor") dt4$IUCN_category<- as.factor(dt4$IUCN_category)
+if (class(dt4$habitat)!="factor") dt4$habitat<- as.factor(dt4$habitat)
+if (class(dt4$habitat_breadth)=="factor") dt4$habitat_breadth <-as.numeric(levels(dt4$habitat_breadth))[as.integer(dt4$habitat_breadth) ]               
+if (class(dt4$habitat_breadth)=="character") dt4$habitat_breadth <-as.numeric(dt4$habitat_breadth)
+if (class(dt4$habitat_breadth_source)!="factor") dt4$habitat_breadth_source<- as.factor(dt4$habitat_breadth_source)
+if (class(dt4$habitat_level)!="factor") dt4$habitat_level<- as.factor(dt4$habitat_level)
+if (class(dt4$mean_CHELSA_bio1_1981.2010_V.2.1)=="factor") dt4$mean_CHELSA_bio1_1981.2010_V.2.1 <-as.numeric(levels(dt4$mean_CHELSA_bio1_1981.2010_V.2.1))[as.integer(dt4$mean_CHELSA_bio1_1981.2010_V.2.1) ]
+if (class(dt4$mean_CHELSA_bio1_1981.2010_V.2.1)=="character") dt4$mean_CHELSA_bio1_1981.2010_V.2.1 <-as.numeric(dt4$mean_CHELSA_bio1_1981.2010_V.2.1)
+if (class(dt4$mean_CHELSA_bio12_1981.2010_V.2.1)=="factor") dt4$mean_CHELSA_bio12_1981.2010_V.2.1 <-as.numeric(levels(dt4$mean_CHELSA_bio12_1981.2010_V.2.1))[as.integer(dt4$mean_CHELSA_bio12_1981.2010_V.2.1) ]
+if (class(dt4$mean_CHELSA_bio12_1981.2010_V.2.1)=="character") dt4$mean_CHELSA_bio12_1981.2010_V.2.1 <-as.numeric(dt4$mean_CHELSA_bio12_1981.2010_V.2.1)
+if (class(dt4$mean_human_fp_range_2010)=="factor") dt4$mean_human_fp_range_2010 <-as.numeric(levels(dt4$mean_human_fp_range_2010))[as.integer(dt4$mean_human_fp_range_2010) ]               
+if (class(dt4$mean_human_fp_range_2010)=="character") dt4$mean_human_fp_range_2010 <-as.numeric(dt4$mean_human_fp_range_2010)
+if (class(dt4$mean_human_fp_range_2020)=="factor") dt4$mean_human_fp_range_2020 <-as.numeric(levels(dt4$mean_human_fp_range_2020))[as.integer(dt4$mean_human_fp_range_2020) ]               
+if (class(dt4$mean_human_fp_range_2020)=="character") dt4$mean_human_fp_range_2020 <-as.numeric(dt4$mean_human_fp_range_2020)
+if (class(dt4$percent_change_hf_2010_2020)=="factor") dt4$percent_change_hf_2010_2020 <-as.numeric(levels(dt4$percent_change_hf_2010_2020))[as.integer(dt4$percent_change_hf_2010_2020) ]               
+if (class(dt4$percent_change_hf_2010_2020)=="character") dt4$percent_change_hf_2010_2020 <-as.numeric(dt4$percent_change_hf_2010_2020)
+if (class(dt4$inferred_range_sqkm)=="factor") dt4$inferred_range_sqkm <-as.numeric(levels(dt4$inferred_range_sqkm))[as.integer(dt4$inferred_range_sqkm) ]               
+if (class(dt4$inferred_range_sqkm)=="character") dt4$inferred_range_sqkm <-as.numeric(dt4$inferred_range_sqkm)
+if (class(dt4$date_data_obtained)!="factor") dt4$date_data_obtained<- as.factor(dt4$date_data_obtained)
+if (class(dt4$filled_by)!="factor") dt4$filled_by<- as.factor(dt4$filled_by)
+
+# Convert Missing Values to NA for non-dates
+
+dt4$IUCN_species_name <- as.factor(ifelse((trimws(as.character(dt4$IUCN_species_name))==trimws("NA")),NA,as.character(dt4$IUCN_species_name)))
+dt4$common_name <- as.factor(ifelse((trimws(as.character(dt4$common_name))==trimws("NA")),NA,as.character(dt4$common_name)))
+dt4$family_e <- as.factor(ifelse((trimws(as.character(dt4$family_e))==trimws("NA")),NA,as.character(dt4$family_e)))
+dt4$genus <- as.factor(ifelse((trimws(as.character(dt4$genus))==trimws("NA")),NA,as.character(dt4$genus)))
+dt4$elton_species_name <- as.factor(ifelse((trimws(as.character(dt4$elton_species_name))==trimws("NA")),NA,as.character(dt4$elton_species_name)))
+dt4$diet_cat_e <- as.factor(ifelse((trimws(as.character(dt4$diet_cat_e))==trimws("NA")),NA,as.character(dt4$diet_cat_e)))
+dt4$diet_level <- as.factor(ifelse((trimws(as.character(dt4$diet_level))==trimws("NA")),NA,as.character(dt4$diet_level)))
+dt4$activity_nocturnal_e <- as.factor(ifelse((trimws(as.character(dt4$activity_nocturnal_e))==trimws("NA")),NA,as.character(dt4$activity_nocturnal_e)))
+dt4$for_strat_spec_level <- as.factor(ifelse((trimws(as.character(dt4$for_strat_spec_level))==trimws("NA")),NA,as.character(dt4$for_strat_spec_level)))
+dt4$for_strat_source_e <- as.factor(ifelse((trimws(as.character(dt4$for_strat_source_e))==trimws("NA")),NA,as.character(dt4$for_strat_source_e)))
+dt4$body_mass_e <- ifelse((trimws(as.character(dt4$body_mass_e))==trimws("NA")),NA,dt4$body_mass_e)               
+suppressWarnings(dt4$body_mass_e <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt4$body_mass_e))==as.character(as.numeric("NA"))),NA,dt4$body_mass_e))
+dt4$body_mass_level_e <- as.factor(ifelse((trimws(as.character(dt4$body_mass_level_e))==trimws("NA")),NA,as.character(dt4$body_mass_level_e)))
+dt4$body_mass_source_e <- as.factor(ifelse((trimws(as.character(dt4$body_mass_source_e))==trimws("NA")),NA,as.character(dt4$body_mass_source_e)))
+dt4$body_size_mm <- ifelse((trimws(as.character(dt4$body_size_mm))==trimws("NA")),NA,dt4$body_size_mm)               
+suppressWarnings(dt4$body_size_mm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt4$body_size_mm))==as.character(as.numeric("NA"))),NA,dt4$body_size_mm))
+dt4$body_size_notes <- as.factor(ifelse((trimws(as.character(dt4$body_size_notes))==trimws("NA")),NA,as.character(dt4$body_size_notes)))
+dt4$body_size_level <- as.factor(ifelse((trimws(as.character(dt4$body_size_level))==trimws("NA")),NA,as.character(dt4$body_size_level)))
+dt4$body_size_source <- as.factor(ifelse((trimws(as.character(dt4$body_size_source))==trimws("NA")),NA,as.character(dt4$body_size_source)))
+dt4$sexual_dimorphism <- as.factor(ifelse((trimws(as.character(dt4$sexual_dimorphism))==trimws("NA")),NA,as.character(dt4$sexual_dimorphism)))
+dt4$sexual_dimorphism_notes <- as.factor(ifelse((trimws(as.character(dt4$sexual_dimorphism_notes))==trimws("NA")),NA,as.character(dt4$sexual_dimorphism_notes)))
+dt4$sexual_dimorphism_level <- as.factor(ifelse((trimws(as.character(dt4$sexual_dimorphism_level))==trimws("NA")),NA,as.character(dt4$sexual_dimorphism_level)))
+dt4$sexual_dimorphism_source <- as.factor(ifelse((trimws(as.character(dt4$sexual_dimorphism_source))==trimws("NA")),NA,as.character(dt4$sexual_dimorphism_source)))
+dt4$longevity <- ifelse((trimws(as.character(dt4$longevity))==trimws("NA")),NA,dt4$longevity)               
+suppressWarnings(dt4$longevity <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt4$longevity))==as.character(as.numeric("NA"))),NA,dt4$longevity))
+dt4$longevity_notes <- as.factor(ifelse((trimws(as.character(dt4$longevity_notes))==trimws("NA")),NA,as.character(dt4$longevity_notes)))
+dt4$longevity_level <- as.factor(ifelse((trimws(as.character(dt4$longevity_level))==trimws("NA")),NA,as.character(dt4$longevity_level)))
+dt4$longevity_source <- as.factor(ifelse((trimws(as.character(dt4$longevity_source))==trimws("NA")),NA,as.character(dt4$longevity_source)))
+dt4$home_range_size <- ifelse((trimws(as.character(dt4$home_range_size))==trimws("NA")),NA,dt4$home_range_size)               
+suppressWarnings(dt4$home_range_size <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt4$home_range_size))==as.character(as.numeric("NA"))),NA,dt4$home_range_size))
+dt4$home_range_notes <- as.factor(ifelse((trimws(as.character(dt4$home_range_notes))==trimws("NA")),NA,as.character(dt4$home_range_notes)))
+dt4$home_range_level <- as.factor(ifelse((trimws(as.character(dt4$home_range_level))==trimws("NA")),NA,as.character(dt4$home_range_level)))
+dt4$home_range_source <- as.factor(ifelse((trimws(as.character(dt4$home_range_source))==trimws("NA")),NA,as.character(dt4$home_range_source)))
+dt4$habitat_specialization <- as.factor(ifelse((trimws(as.character(dt4$habitat_specialization))==trimws("NA")),NA,as.character(dt4$habitat_specialization)))
+dt4$habitat_specialization_source <- as.factor(ifelse((trimws(as.character(dt4$habitat_specialization_source))==trimws("NA")),NA,as.character(dt4$habitat_specialization_source)))
+dt4$generation_time <- ifelse((trimws(as.character(dt4$generation_time))==trimws("NA")),NA,dt4$generation_time)               
+suppressWarnings(dt4$generation_time <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt4$generation_time))==as.character(as.numeric("NA"))),NA,dt4$generation_time))
+dt4$generation_time_notes <- as.factor(ifelse((trimws(as.character(dt4$generation_time_notes))==trimws("NA")),NA,as.character(dt4$generation_time_notes)))
+dt4$generation_time_source <- as.factor(ifelse((trimws(as.character(dt4$generation_time_source))==trimws("NA")),NA,as.character(dt4$generation_time_source)))
+dt4$habitat_breadth <- ifelse((trimws(as.character(dt4$habitat_breadth))==trimws("NA")),NA,dt4$habitat_breadth)               
+suppressWarnings(dt4$habitat_breadth <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt4$habitat_breadth))==as.character(as.numeric("NA"))),NA,dt4$habitat_breadth))
+dt4$habitat_level <- as.factor(ifelse((trimws(as.character(dt4$habitat_level))==trimws("NA")),NA,as.character(dt4$habitat_level)))
+dt4$mean_CHELSA_bio1_1981.2010_V.2.1 <- ifelse((trimws(as.character(dt4$mean_CHELSA_bio1_1981.2010_V.2.1))==trimws("NA")),NA,dt4$mean_CHELSA_bio1_1981.2010_V.2.1)               
+suppressWarnings(dt4$mean_CHELSA_bio1_1981.2010_V.2.1 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt4$mean_CHELSA_bio1_1981.2010_V.2.1))==as.character(as.numeric("NA"))),NA,dt4$mean_CHELSA_bio1_1981.2010_V.2.1))
+dt4$mean_CHELSA_bio12_1981.2010_V.2.1 <- ifelse((trimws(as.character(dt4$mean_CHELSA_bio12_1981.2010_V.2.1))==trimws("NA")),NA,dt4$mean_CHELSA_bio12_1981.2010_V.2.1)               
+suppressWarnings(dt4$mean_CHELSA_bio12_1981.2010_V.2.1 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt4$mean_CHELSA_bio12_1981.2010_V.2.1))==as.character(as.numeric("NA"))),NA,dt4$mean_CHELSA_bio12_1981.2010_V.2.1))
+dt4$mean_human_fp_range_2010 <- ifelse((trimws(as.character(dt4$mean_human_fp_range_2010))==trimws("NA")),NA,dt4$mean_human_fp_range_2010)               
+suppressWarnings(dt4$mean_human_fp_range_2010 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt4$mean_human_fp_range_2010))==as.character(as.numeric("NA"))),NA,dt4$mean_human_fp_range_2010))
+dt4$mean_human_fp_range_2020 <- ifelse((trimws(as.character(dt4$mean_human_fp_range_2020))==trimws("NA")),NA,dt4$mean_human_fp_range_2020)               
+suppressWarnings(dt4$mean_human_fp_range_2020 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt4$mean_human_fp_range_2020))==as.character(as.numeric("NA"))),NA,dt4$mean_human_fp_range_2020))
+dt4$percent_change_hf_2010_2020 <- ifelse((trimws(as.character(dt4$percent_change_hf_2010_2020))==trimws("NA")),NA,dt4$percent_change_hf_2010_2020)               
+suppressWarnings(dt4$percent_change_hf_2010_2020 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt4$percent_change_hf_2010_2020))==as.character(as.numeric("NA"))),NA,dt4$percent_change_hf_2010_2020))
+dt4$inferred_range_sqkm <- ifelse((trimws(as.character(dt4$inferred_range_sqkm))==trimws("NA")),NA,dt4$inferred_range_sqkm)               
+suppressWarnings(dt4$inferred_range_sqkm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt4$inferred_range_sqkm))==as.character(as.numeric("NA"))),NA,dt4$inferred_range_sqkm))
+dt4$date_data_obtained <- as.factor(ifelse((trimws(as.character(dt4$date_data_obtained))==trimws("NA")),NA,as.character(dt4$date_data_obtained)))
+
+# Here is the structure of the input data frame:
+str(dt4)                            
+attach(dt4)                            
+# The analyses below are basic descriptions of the variables. After testing, they should be replaced.
+
+summary(IUCN_species_name)
+summary(common_name)
+summary(family_e)
+summary(genus)
+summary(species)
+summary(elton_species_name)
+summary(diet_cat_e)
+summary(diet_source_e)
+summary(diet_breadth)
+summary(diet_level)
+summary(activity_nocturnal_e)
+summary(for_strat_ground_e)
+summary(for_strat_understory_e)
+summary(for_strat_midhigh_e)
+summary(for_strat_canopy_e)
+summary(for_strat_aerial_e)
+summary(for_strat_spec_level)
+summary(for_strat_source_e)
+summary(body_mass_e)
+summary(body_mass_level_e)
+summary(body_mass_source_e)
+summary(body_size_mm)
+summary(body_size_notes)
+summary(body_size_level)
+summary(body_size_source)
+summary(sexual_dimorphism)
+summary(sexual_dimorphism_notes)
+summary(sexual_dimorphism_level)
+summary(sexual_dimorphism_source)
+summary(longevity)
+summary(longevity_notes)
+summary(longevity_level)
+summary(longevity_source)
+summary(home_range_size)
+summary(home_range_notes)
+summary(home_range_level)
+summary(home_range_source)
+summary(habitat_specialization)
+summary(habitat_specialization_source)
+summary(generation_time)
+summary(generation_time_notes)
+summary(generation_time_level)
+summary(generation_time_source)
+summary(IUCN_category)
+summary(habitat)
+summary(habitat_breadth)
+summary(habitat_breadth_source)
+summary(habitat_level)
+summary(mean_CHELSA_bio1_1981.2010_V.2.1)
+summary(mean_CHELSA_bio12_1981.2010_V.2.1)
+summary(mean_human_fp_range_2010)
+summary(mean_human_fp_range_2020)
+summary(percent_change_hf_2010_2020)
+summary(inferred_range_sqkm)
+summary(date_data_obtained)
+summary(filled_by) 
+                # Get more details on character variables
+
+summary(as.factor(dt4$IUCN_species_name)) 
+summary(as.factor(dt4$common_name)) 
+summary(as.factor(dt4$family_e)) 
+summary(as.factor(dt4$genus)) 
+summary(as.factor(dt4$species)) 
+summary(as.factor(dt4$elton_species_name)) 
+summary(as.factor(dt4$diet_cat_e)) 
+summary(as.factor(dt4$diet_source_e)) 
+summary(as.factor(dt4$diet_level)) 
+summary(as.factor(dt4$activity_nocturnal_e)) 
+summary(as.factor(dt4$for_strat_spec_level)) 
+summary(as.factor(dt4$for_strat_source_e)) 
+summary(as.factor(dt4$body_mass_level_e)) 
+summary(as.factor(dt4$body_mass_source_e)) 
+summary(as.factor(dt4$body_size_notes)) 
+summary(as.factor(dt4$body_size_level)) 
+summary(as.factor(dt4$body_size_source)) 
+summary(as.factor(dt4$sexual_dimorphism)) 
+summary(as.factor(dt4$sexual_dimorphism_notes)) 
+summary(as.factor(dt4$sexual_dimorphism_level)) 
+summary(as.factor(dt4$sexual_dimorphism_source)) 
+summary(as.factor(dt4$longevity_notes)) 
+summary(as.factor(dt4$longevity_level)) 
+summary(as.factor(dt4$longevity_source)) 
+summary(as.factor(dt4$home_range_notes)) 
+summary(as.factor(dt4$home_range_level)) 
+summary(as.factor(dt4$home_range_source)) 
+summary(as.factor(dt4$habitat_specialization)) 
+summary(as.factor(dt4$habitat_specialization_source)) 
+summary(as.factor(dt4$generation_time_notes)) 
+summary(as.factor(dt4$generation_time_level)) 
+summary(as.factor(dt4$generation_time_source)) 
+summary(as.factor(dt4$IUCN_category)) 
+summary(as.factor(dt4$habitat)) 
+summary(as.factor(dt4$habitat_breadth_source)) 
+summary(as.factor(dt4$habitat_level)) 
+summary(as.factor(dt4$date_data_obtained)) 
+summary(as.factor(dt4$filled_by))
+detach(dt4)
+
+inUrl5  <- "https://pasta.lternet.edu/package/data/eml/edi/1220/6/1c814c7d28285498076f6ebe2fb67655" 
+infile5 <- tempfile()
+try(download.file(inUrl5,infile5,method="curl"))
+if (is.na(file.size(infile5))) download.file(inUrl5,infile5,method="auto")
+
+ dt5 <-read.csv(infile5,header=F 
+          ,skip=1
+            ,sep=","  
+                ,quot='"' 
+        , col.names=c(
+                    "IUCN_species_name",     
+                    "elton_species_name"    ), check.names=TRUE)
+
+unlink(infile5)
+
+# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
+
+if (class(dt5$IUCN_species_name)!="factor") dt5$IUCN_species_name<- as.factor(dt5$IUCN_species_name)
+if (class(dt5$elton_species_name)!="factor") dt5$elton_species_name<- as.factor(dt5$elton_species_name)
+
+# Convert Missing Values to NA for non-dates
+
+# Here is the structure of the input data frame:
+str(dt5)                            
+attach(dt5)                            
+# The analyses below are basic descriptions of the variables. After testing, they should be replaced.
+
+summary(IUCN_species_name)
+summary(elton_species_name) 
+                # Get more details on character variables
+
+summary(as.factor(dt5$IUCN_species_name)) 
+summary(as.factor(dt5$elton_species_name))
+detach(dt5)
+
+inUrl6  <- "https://pasta.lternet.edu/package/data/eml/edi/1220/6/140e29ffdeac84310c3e07e9a93b9fcf" 
+infile6 <- tempfile()
+try(download.file(inUrl6,infile6,method="curl"))
+if (is.na(file.size(infile6))) download.file(inUrl6,infile6,method="auto")
+
+ dt6 <-read.csv(infile6,header=F 
+          ,skip=1
+            ,sep=","  
+                ,quot='"' 
+        , col.names=c(
+                    "IUCN_species_name",     
+                    "elton_species_name"    ), check.names=TRUE)
+
+unlink(infile6)
+
+# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
+
+if (class(dt6$IUCN_species_name)!="factor") dt6$IUCN_species_name<- as.factor(dt6$IUCN_species_name)
+if (class(dt6$elton_species_name)!="factor") dt6$elton_species_name<- as.factor(dt6$elton_species_name)
+
+# Convert Missing Values to NA for non-dates
+
+# Here is the structure of the input data frame:
+str(dt6)                            
+attach(dt6)                            
+# The analyses below are basic descriptions of the variables. After testing, they should be replaced.
+
+summary(IUCN_species_name)
+summary(elton_species_name) 
+                # Get more details on character variables
+
+summary(as.factor(dt6$IUCN_species_name)) 
+summary(as.factor(dt6$elton_species_name))
+detach(dt6)
+
+inUrl7  <- "https://pasta.lternet.edu/package/data/eml/edi/1220/6/a31672f05368bbcb88f166a8f959c857" 
+infile7 <- tempfile()
+try(download.file(inUrl7,infile7,method="curl"))
+if (is.na(file.size(infile7))) download.file(inUrl7,infile7,method="auto")
+
+ dt7 <-read.csv(infile7,header=F 
+          ,skip=1
+            ,sep=","  
+                ,quot='"' 
+        , col.names=c(
+                    "Trait",     
+                    "Full.citation",     
+                    "In.text.citation"    ), check.names=TRUE)
+
+unlink(infile7)
+
+# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
+
+if (class(dt7$Trait)!="factor") dt7$Trait<- as.factor(dt7$Trait)
+if (class(dt7$Full.citation)!="factor") dt7$Full.citation<- as.factor(dt7$Full.citation)
+if (class(dt7$In.text.citation)!="factor") dt7$In.text.citation<- as.factor(dt7$In.text.citation)
+
+# Convert Missing Values to NA for non-dates
+
+# Here is the structure of the input data frame:
+str(dt7)                            
+attach(dt7)                            
+# The analyses below are basic descriptions of the variables. After testing, they should be replaced.
+
+summary(Trait)
+summary(Full.citation)
+summary(In.text.citation) 
+                # Get more details on character variables
+
+summary(as.factor(dt7$Trait)) 
+summary(as.factor(dt7$Full.citation)) 
+summary(as.factor(dt7$In.text.citation))
+detach(dt7)
+
+inUrl8  <- "https://pasta.lternet.edu/package/data/eml/edi/1220/6/655382c5f780337ecd85fae8345bfb10" 
+infile8 <- tempfile()
+try(download.file(inUrl8,infile8,method="curl"))
+if (is.na(file.size(infile8))) download.file(inUrl8,infile8,method="auto")
+
+ dt8 <-read.csv(infile8,header=F 
+          ,skip=1
+            ,sep=","  
+                ,quot='"' 
+        , col.names=c(
+                    "Trait",     
+                    "Full.citation",     
+                    "In.text.citation"    ), check.names=TRUE)
+
+unlink(infile8)
+
+# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
+
+if (class(dt8$Trait)!="factor") dt8$Trait<- as.factor(dt8$Trait)
+if (class(dt8$Full.citation)!="factor") dt8$Full.citation<- as.factor(dt8$Full.citation)
+if (class(dt8$In.text.citation)!="factor") dt8$In.text.citation<- as.factor(dt8$In.text.citation)
+
+# Convert Missing Values to NA for non-dates
+
+# Here is the structure of the input data frame:
+str(dt8)                            
+attach(dt8)                            
+# The analyses below are basic descriptions of the variables. After testing, they should be replaced.
+
+summary(Trait)
+summary(Full.citation)
+summary(In.text.citation) 
+                # Get more details on character variables
+
+summary(as.factor(dt8$Trait)) 
+summary(as.factor(dt8$Full.citation)) 
+summary(as.factor(dt8$In.text.citation))
+detach(dt8)
+
+inUrl9  <- "https://pasta.lternet.edu/package/data/eml/edi/1220/6/8f154e65cc499cdbf7b22f6f4485ab18" 
+infile9 <- tempfile()
+try(download.file(inUrl9,infile9,method="curl"))
+if (is.na(file.size(infile9))) download.file(inUrl9,infile9,method="auto")
+
+ dt9 <-read.csv(infile9,header=F 
+          ,skip=1
+            ,sep=","  
+                ,quot='"' 
+        , col.names=c(
+                    "Grouping.feature",     
+                    "Trait",     
+                    "Column.name",     
+                    "Description",     
+                    "Unit",     
+                    "Citation"    ), check.names=TRUE)
+
+unlink(infile9)
+
+# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
+
+if (class(dt9$Grouping.feature)!="factor") dt9$Grouping.feature<- as.factor(dt9$Grouping.feature)
+if (class(dt9$Trait)!="factor") dt9$Trait<- as.factor(dt9$Trait)
+if (class(dt9$Column.name)!="factor") dt9$Column.name<- as.factor(dt9$Column.name)
+if (class(dt9$Description)!="factor") dt9$Description<- as.factor(dt9$Description)
+if (class(dt9$Unit)!="factor") dt9$Unit<- as.factor(dt9$Unit)
+if (class(dt9$Citation)!="factor") dt9$Citation<- as.factor(dt9$Citation)
+
+# Convert Missing Values to NA for non-dates
+
+# Here is the structure of the input data frame:
+str(dt9)                            
+attach(dt9)                            
+# The analyses below are basic descriptions of the variables. After testing, they should be replaced.
+
+summary(Grouping.feature)
+summary(Trait)
+summary(Column.name)
+summary(Description)
+summary(Unit)
+summary(Citation) 
+                # Get more details on character variables
+
+summary(as.factor(dt9$Grouping.feature)) 
+summary(as.factor(dt9$Trait)) 
+summary(as.factor(dt9$Column.name)) 
+summary(as.factor(dt9$Description)) 
+summary(as.factor(dt9$Unit)) 
+summary(as.factor(dt9$Citation))
+detach(dt9)
+
+inUrl10  <- "https://pasta.lternet.edu/package/data/eml/edi/1220/6/f73cd1b16a4ed908dac4cf532690a12b" 
+infile10 <- tempfile()
+try(download.file(inUrl10,infile10,method="curl"))
+if (is.na(file.size(infile10))) download.file(inUrl10,infile10,method="auto")
+
+ dt10 <-read.csv(infile10,header=F 
+          ,skip=1
+            ,sep=","  
+                ,quot='"' 
+        , col.names=c(
+                    "IUCN_species_name",     
+                    "elton_species_name",     
+                    "PanTHERIA_species_name",     
+                    "habitat",     
+                    "new_species",     
+                    "reclassified",     
+                    "spelling_differences"    ), check.names=TRUE)
+
+unlink(infile10)
+
+# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
+
+if (class(dt10$IUCN_species_name)!="factor") dt10$IUCN_species_name<- as.factor(dt10$IUCN_species_name)
+if (class(dt10$elton_species_name)!="factor") dt10$elton_species_name<- as.factor(dt10$elton_species_name)
+if (class(dt10$PanTHERIA_species_name)!="factor") dt10$PanTHERIA_species_name<- as.factor(dt10$PanTHERIA_species_name)
+if (class(dt10$habitat)!="factor") dt10$habitat<- as.factor(dt10$habitat)
+if (class(dt10$new_species)!="factor") dt10$new_species<- as.factor(dt10$new_species)
+if (class(dt10$reclassified)!="factor") dt10$reclassified<- as.factor(dt10$reclassified)
+if (class(dt10$spelling_differences)!="factor") dt10$spelling_differences<- as.factor(dt10$spelling_differences)
+
+# Convert Missing Values to NA for non-dates
+
+dt10$habitat <- as.factor(ifelse((trimws(as.character(dt10$habitat))==trimws("NA")),NA,as.character(dt10$habitat)))
+dt10$new_species <- as.factor(ifelse((trimws(as.character(dt10$new_species))==trimws("NA")),NA,as.character(dt10$new_species)))
+dt10$reclassified <- as.factor(ifelse((trimws(as.character(dt10$reclassified))==trimws("NA")),NA,as.character(dt10$reclassified)))
+dt10$spelling_differences <- as.factor(ifelse((trimws(as.character(dt10$spelling_differences))==trimws("NA")),NA,as.character(dt10$spelling_differences)))
+
+# Here is the structure of the input data frame:
+str(dt10)                            
+attach(dt10)                            
+# The analyses below are basic descriptions of the variables. After testing, they should be replaced.
+
+summary(IUCN_species_name)
+summary(elton_species_name)
+summary(PanTHERIA_species_name)
+summary(habitat)
+summary(new_species)
+summary(reclassified)
+summary(spelling_differences) 
+                # Get more details on character variables
+
+summary(as.factor(dt10$IUCN_species_name)) 
+summary(as.factor(dt10$elton_species_name)) 
+summary(as.factor(dt10$PanTHERIA_species_name)) 
+summary(as.factor(dt10$habitat)) 
+summary(as.factor(dt10$new_species)) 
+summary(as.factor(dt10$reclassified)) 
+summary(as.factor(dt10$spelling_differences))
+detach(dt10)
+
+inUrl11  <- "https://pasta.lternet.edu/package/data/eml/edi/1220/6/d1f32cc55b692cfb3d614cb400797299" 
+infile11 <- tempfile()
+try(download.file(inUrl11,infile11,method="curl"))
+if (is.na(file.size(infile11))) download.file(inUrl11,infile11,method="auto")
+
+ dt11 <-read.csv(infile11,header=F 
+          ,skip=1
+            ,sep=","  
+                ,quot='"' 
+        , col.names=c(
+                    "IUCN_species_name",     
+                    "elton_species_name",     
+                    "habitat",     
+                    "new_species",     
+                    "reclassified",     
+                    "spelling_difference"    ), check.names=TRUE)
+
+unlink(infile11)
+
+# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
+
+if (class(dt11$IUCN_species_name)!="factor") dt11$IUCN_species_name<- as.factor(dt11$IUCN_species_name)
+if (class(dt11$elton_species_name)!="factor") dt11$elton_species_name<- as.factor(dt11$elton_species_name)
+if (class(dt11$habitat)!="factor") dt11$habitat<- as.factor(dt11$habitat)
+if (class(dt11$new_species)!="factor") dt11$new_species<- as.factor(dt11$new_species)
+if (class(dt11$reclassified)!="factor") dt11$reclassified<- as.factor(dt11$reclassified)
+if (class(dt11$spelling_difference)!="factor") dt11$spelling_difference<- as.factor(dt11$spelling_difference)
+
+# Convert Missing Values to NA for non-dates
+
+dt11$new_species <- as.factor(ifelse((trimws(as.character(dt11$new_species))==trimws("NA")),NA,as.character(dt11$new_species)))
+dt11$reclassified <- as.factor(ifelse((trimws(as.character(dt11$reclassified))==trimws("NA")),NA,as.character(dt11$reclassified)))
+dt11$spelling_difference <- as.factor(ifelse((trimws(as.character(dt11$spelling_difference))==trimws("NA")),NA,as.character(dt11$spelling_difference)))
+
+# Here is the structure of the input data frame:
+str(dt11)                            
+attach(dt11)                            
+# The analyses below are basic descriptions of the variables. After testing, they should be replaced.
+
+summary(IUCN_species_name)
+summary(elton_species_name)
+summary(habitat)
+summary(new_species)
+summary(reclassified)
+summary(spelling_difference) 
+                # Get more details on character variables
+
+summary(as.factor(dt11$IUCN_species_name)) 
+summary(as.factor(dt11$elton_species_name)) 
+summary(as.factor(dt11$habitat)) 
+summary(as.factor(dt11$new_species)) 
+summary(as.factor(dt11$reclassified)) 
+summary(as.factor(dt11$spelling_difference))
+detach(dt11)
+
+inUrl12  <- "https://pasta.lternet.edu/package/data/eml/edi/1220/6/cf3f2a44b577d0176f08299d4ae9b90d" 
+infile12 <- tempfile()
+try(download.file(inUrl12,infile12,method="curl"))
+if (is.na(file.size(infile12))) download.file(inUrl12,infile12,method="auto")
+
+ dt12 <-read.csv(infile12,header=F 
+          ,skip=1
+            ,sep=","  
+                ,quot='"' 
+        , col.names=c(
+                    "species",     
+                    "code",     
+                    "habitat",     
+                    "suitability",     
+                    "season",     
+                    "majorimportance"    ), check.names=TRUE)
+
+unlink(infile12)
+
+# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
+
+if (class(dt12$species)!="factor") dt12$species<- as.factor(dt12$species)
+if (class(dt12$code)!="factor") dt12$code<- as.factor(dt12$code)
+if (class(dt12$habitat)!="factor") dt12$habitat<- as.factor(dt12$habitat)
+if (class(dt12$suitability)!="factor") dt12$suitability<- as.factor(dt12$suitability)
+if (class(dt12$season)!="factor") dt12$season<- as.factor(dt12$season)
+if (class(dt12$majorimportance)!="factor") dt12$majorimportance<- as.factor(dt12$majorimportance)
+
+# Convert Missing Values to NA for non-dates
+
+dt12$species <- as.factor(ifelse((trimws(as.character(dt12$species))==trimws("NA")),NA,as.character(dt12$species)))
+dt12$habitat <- as.factor(ifelse((trimws(as.character(dt12$habitat))==trimws("NA")),NA,as.character(dt12$habitat)))
+dt12$suitability <- as.factor(ifelse((trimws(as.character(dt12$suitability))==trimws("NA")),NA,as.character(dt12$suitability)))
+dt12$season <- as.factor(ifelse((trimws(as.character(dt12$season))==trimws("NA")),NA,as.character(dt12$season)))
+dt12$majorimportance <- as.factor(ifelse((trimws(as.character(dt12$majorimportance))==trimws("NA")),NA,as.character(dt12$majorimportance)))
+
+# Here is the structure of the input data frame:
+str(dt12)                            
+attach(dt12)                            
+# The analyses below are basic descriptions of the variables. After testing, they should be replaced.
+
+summary(species)
+summary(code)
+summary(habitat)
+summary(suitability)
+summary(season)
+summary(majorimportance) 
+                # Get more details on character variables
+
+summary(as.factor(dt12$species)) 
+summary(as.factor(dt12$code)) 
+summary(as.factor(dt12$habitat)) 
+summary(as.factor(dt12$suitability)) 
+summary(as.factor(dt12$season)) 
+summary(as.factor(dt12$majorimportance))
+detach(dt12)               
 ```

### 2023-07-18 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
update file path

@@
 ```{r}
-TropicalAndes_GBIF_path <- file.path(output_path,"TropAndes_GBIF_plant_occ.csv") 
+TropicalAndes_GBIF_path <- file.path(output_path,"TropicalAndes_GBIF_plant_occ.csv") 
 write.csv(d2, file = TropicalAndes_GBIF_path)

### 2023-07-18 code/L0/L0_TropicalAndes_plants_TRY_traits.Rmd
update input file path add summary info to parallel occurrence

@@
 ```{r}
-
+library(rtry)
 ```
@@
 ```{r}
-try_data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/23650_24112022083128/23650.txt", encoding = "Latin-1")
-try_data <- as.data.frame(try_data)
+try_data <- read.delim("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/25496_17032023010024/25496.txt")
+```
+
+# Summary 
+```{r}
+glimpse(try_data)
 ```
@@
 ```{r}
-
+print("The number of records is")
+nrow(try_data)
+print("The number of species is" )
+length(unique(try_data$species))
+print("The number of families is" )
+length(unique(try_data$family))
 ```

### 2023-07-24 code/L0/L0_TropicalAndes_plants_TRY_traits.Rmd
subset by tropical andes georefferenced records species list code save as csv

--- 2023-07-18 code/L0/L0_TropicalAndes_plants_TRY_traits.Rmd
@@
 data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
-output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
 ```
@@
 ```{r}
-try_data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/23650_24112022083128/23650.txt", encoding = "Latin-1")
-try_data <- as.data.frame(try_data)
+try_data <- rtry_importujkgh("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/25496_17032023010024/25496.txt")
+```
+
+#Subset to Tropical Andes
+```{r}
+head(try_data)
+```
+
+# Summary 
+```{r}
+glimpse(try_data)
 ```
@@
 ```{r}
+print("The number of records is")
+nrow(try_data)
+print("The number of species is" )
+length(unique(try_data$AccSpeciesName))
+```
 
+# Add Lat & Long as columns in database
+
+```{r}
+# Extract the unique value of latitude (DataID 59) and the corresponding ObservationID
+workdata_lat <- rtry_select_anc(try_data, 59)
+
+# Extract the unique value of longitude (DataID 60) and the corresponding ObservationID
+workdata_lon <- rtry_select_anc(try_data, 60)
 ```
+
+```{r}
+# To merge the extracted ancillary data with the numerical traits
+# Merge the relevant data frames based on the ObservationID using rtry_join_left (left join)
+try_data_georef <- rtry_join_left(try_data, workdata_lat, baseOn = ObservationID)
+try_data_georef <- rtry_join_left(try_data_georef, workdata_lon, baseOn = ObservationID)
+glimpse(try_data_georef)
+```
+```{r}
+# remove rows with na in latitude
+try_data_georef_noNA <- try_data_georef[!is.na(try_data_georef$Latitude),]
+glimpse(try_data_georef_noNA)
+```
+
+```{r}
+#subset by bounding box of Tropical Andes
+try_data_georef_lat <- subset(try_data_georef_noNA, Latitude< 12.5 & Latitude> -22.8)
+glimpse(try_data_georef_lat)
+try_data_georef_lon <- subset(try_data_georef_lat, Longitude> -96.6 & Longitude< -57.5)
+glimpse(try_data_georef_lon)
+```
+
+```{r}
+#remove unnecessary column V29
+try_data_TropicalAndes <- subset(try_data_georef_lon, select =-V29)
+```
+
+# Extract species list
+```{r}
+TRY_SpeciesList <- unique(try_data_TropicalAndes$AccSpeciesName)
+```
+
+# Summary
+
+```{r}
+glimpse(try_data_TropicalAndes)
+```
+
+```{r}
+print("The number of records is")
+nrow(try_data_TropicalAndes)
+print("The number of species is" )
+length(unique(try_data_TropicalAndes$AccSpeciesName))
+```
+
+# Write data to csv
+```{r}
+TropicalAndes_TRY_path <- file.path(output_path,"TropicalAndes_TRY_plant_traits.csv") 
+write.csv(try_data_TropicalAndes, file = TropicalAndes_TRY_path)
+TropicalAndes_TRY_species_path <- file.path(output_path,"TropicalAndes_TRY_plant_species.csv") 
+write.csv(TRY_SpeciesList, file = TropicalAndes_TRY_species_path)
+```

### 2023-07-25 code/L0/L0_TropicalAndes_plants_BIEN_traits.Rmd
update summary and species list code

@@
 
+# Extract species list
+```{r}
+BIEN_trait_SpeciesList <- unique(TropicalAndes_BIEN_occ$scrubbed_species_binomial)
+```
+
 # Write data to csv
@@
 write.csv(TropicalAndes_BIEN_traits, file = TropicalAndes_BIEN_traits_path)
+TropicalAndes_BIEN_species_list_path <- file.path(output_path, "TropicalAndes_BIEN_plant_trait_species.csv")
+write.csv(BIEN_trait_SpeciesList, file = TropicalAndes_BIEN_species_list_path)
 ```

### 2023-07-25 code/L0/L0_TropicalAndes_plants_BIEN_occ.Rmd
update summary and species list code

@@
 
-# Read in IUCN data
-Data was obtained from https://www.iucnredlist.org/search?permalink=9621ab77-2341-4bf8-9557-b7f19f7f2a35
+# Get occurrence data for Tropical Andes from BIEN database
 ```{r}
-plants_habitats <- read.csv(file.path(data_path,"IUCN/Tropical Andes Plants - Search Results/habitats.csv"))
+TropicalAndes_BIEN_occ_ec <- BIEN_occurrence_country(country = "Ecuador", all.taxonomy = TRUE, observation.type = TRUE, political.boundaries = TRUE, collection.info = TRUE)
+
+TropicalAndes_BIEN_occ_ve <- BIEN_occurrence_country(country = "Venezuela", all.taxonomy = TRUE, observation.type = TRUE, political.boundaries = TRUE, collection.info = TRUE)
+
+TropicalAndes_BIEN_occ_pe <- BIEN_occurrence_country(country = "Peru", all.taxonomy = TRUE, observation.type = TRUE, political.boundaries = TRUE, collection.info = TRUE)
+
+TropicalAndes_BIEN_occ_co <- BIEN_occurrence_country(country = "Colombia", all.taxonomy = TRUE, observation.type = TRUE, political.boundaries = TRUE, collection.info = TRUE)
+
+TropicalAndes_BIEN_occ_bo <- BIEN_occurrence_country(country = "Bolivia", all.taxonomy = TRUE, observation.type = TRUE, political.boundaries = TRUE, collection.info = TRUE)
 ```
 
-# Create species list from IUCN data
 ```{r}
-plant_species <- unique(plants_habitats$scientificName)
-```
+# combine occ
 
-# List of countries of interest
-```{r}
-country_vector <- c("Ecuador", "Venezuela", "Peru", "Colombia", "Bolivia")
-```
-
-# Get occurrence data for Tropical Andes from BIEN database
-```{r}
-TropicalAndes_BIEN_occ <- BIEN_occurrence_country(country = country_vector, all.taxonomy = TRUE, observation.type = TRUE, political.boundaries = TRUE, collection.info = TRUE)
-
-# species list
-TropicalAndes_BIEN_species_list <- BIEN_list_country(country = country_vector)
 ```
@@
 ```{r}
-glimpse(TropicalAndes_BIEN_species_list)
+print("The number of records is")
+nrow(TropicalAndes_BIEN_occ)
+print("The number of species is")
+length(unique(TropicalAndes_BIEN_occ$scrubbed_species_binomial))
+print("The number of families is")
+length(unique(TropicalAndes_BIEN_occ$family_matched))
+```
+
+# Extract species list
+```{r}
+BIEN_occ_SpeciesList <- unique(TropicalAndes_BIEN_occ$scrubbed_species_binomial)
 ```
@@
 
-TropicalAndes_BIEN_species_list_path <- file.path(output_path, "TropicalAndes_BIEN_plant_species_list.csv")
+TropicalAndes_BIEN_species_list_path <- file.path(output_path, "TropicalAndes_BIEN_plant_occ_species.csv")
 write.csv(TropicalAndes_BIEN_species_list, file = TropicalAndes_BIEN_species_list_path)

### 2023-07-25 code/L0/L0_TropicalAndes_plants_GBIF.Rmd

--- 2023-07-18 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
@@
 
-## Write data to csv
+# Extract species list
 ```{r}
-TropicalAndes_GBIF_path <- file.path(output_path,"TropAndes_GBIF_plant_occ.csv") 
+GBIF_SpeciesList <- unique(d2$species)
+```
+
+# Write data to csv
+```{r}
+TropicalAndes_GBIF_path <- file.path(output_path,"TropicalAndes_GBIF_plant_occ.csv") 
 write.csv(d2, file = TropicalAndes_GBIF_path)
+TropicalAndes_GBIF_species_path <- file.path(output_path,"TropicalAndes_GBIF_plant_species.csv") 
+write.csv(GBIF_SpeciesList, file = TropicalAndes_GBIF_species_path)
 ```

### 2023-07-25 archive/L0_TropicalAndes_plants_BIEN_ranges.Rmd
move to archive

---
title: "Tropical Andes plant species BIEN Ranges"
authors: "Hazel J. Anderson"
collaborators: "None"
data input: "none"
data output: "TBD"
project: "Frugivoria"
date: "2023-02-13"
output: html_document
---
# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(BIEN)
library(tidyr)
library(dplyr)
library(rnaturalearth)
```

# Polygons of countries
```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
#polygon of Tropical Andes
TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
```

# Get data for Tropical Andes from BIEN database
```{r}
# range map data
TropicalAndes_BIEN_ranges <- BIEN_ranges_box(min.lat = -24, max.lat = 14, min.long = -94, max.long = -54, directory = output_path, species.names.only = TRUE)

#BIEN_ranges_sf(TApoly, directory = output_path, return.species.list = TRUE, crop.ranges = TRUE)

```

# Data summary
```{r}
glimpse(TropicalAndes_BIEN_occ)
```

### 2023-07-25 code/L0/L0_TropicalAndes_frugivores_GBIF.Rmd
add additional summary and species list code

@@
 
+```{r}
+print("The number of records is")
+nrow(d)
+print("The number of species is")
+length(unique(d$species))
+print("The number of families is")
+length(unique(d$family))
+```
+
+# Extract species list
+```{r}
+GBIF_occ_SpeciesList <- unique(d$species)
+```
+
 # Write data to csv
@@
 write.csv(d, file = TropicalAndes_GBIF_path)
+TropicalAndes_GBIF_species_path <- file.path(output_path,"TropAndes_GBIF_frugivore_occ_species.csv") 
+write.csv(GBIF_occ_SpeciesList, file = TropicalAndes_GBIF_path)
 ```

### 2023-07-25 archive/MappingGBIF_Richness.RMD
move old code to archive

--- 2023-04-12 MappingGBIF_Richness.RMD
@@
 ```{r}
-NorthAndes_GBIF_plant_occ <- read.csv(file.path(data_path,"TropAndes_GBIF_plant_occ.csv"))
-NorthAndes_GBIF_frugivore_occ <- read.csv(file.path(data_path,"TropAndes_GBIF_frugivore_occ.csv"))
+TropicalAndes_GBIF_plant_occ <- read.csv(file.path(data_path,"TropAndes_GBIF_plant_occ.csv"))
+TropicalAndes_GBIF_frugivore_occ <- read.csv(file.path(data_path,"TropAndes_GBIF_frugivore_occ.csv"))
 ```
@@
 ```{r}
-plants.sf <- st_as_sf(NorthAndes_GBIF_plant_occ, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
+plants.sf <- st_as_sf(TropicalAndes_GBIF_plant_occ, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
 ```
@@
 ```{r}
-frugivores.sf <- st_as_sf(NorthAndes_GBIF_frugivore_occ, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
+frugivores.sf <- st_as_sf(TropicalAndes_GBIF_frugivore_occ, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
 ```

### 2023-07-25 archive/MappingTopicalAndes.Rmd
move old code to archive

Map of Tropical Andes

```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(ggspatial)
```

```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
# country subset. 
ECpoly <- worldMap %>% filter(sovereignt == "Ecuador")
VEpoly <- worldMap %>% filter(sovereignt == "Venezuela")
COpoly <- worldMap %>% filter(sovereignt == "Colombia")
PEpoly <- worldMap %>% filter(sovereignt == "Peru")
BOpoly <- worldMap %>% filter(sovereignt == "Bolivia")
#polygon of Tropical Andes
TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
# set color for Tropical Andes
Tropical_Andes <- "lightgreen"
```

```{r}
MapPlot <-
ggplot() +
geom_sf(data = worldMap, fill = "White") +
geom_sf(data = ECpoly, fill = Tropical_Andes) +
geom_sf(data = VEpoly, fill = Tropical_Andes) +
geom_sf(data = COpoly, fill = Tropical_Andes) +
geom_sf(data = PEpoly, fill = Tropical_Andes) +
geom_sf(data = BOpoly, fill = Tropical_Andes) +
labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
annotation_scale(location = "bl", width_hint = 0.5) +
annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
theme(panel.background = element_rect(fill = "lightblue"))

MapPlot
```

```{r}
ggsave(filename = "Tropical Andes Map.png", dpi = 300, device="png")
```

### 2023-07-25 code/L1/L1_TropicalAndes_plants_GBIF.Rmd
inital commit

---
title: "Cleaning Tropical Andes plant GBIF occurrence records"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: "TropicalAndes_GBIF_plant_occ.csv"
data output: "TropicalAndes_GBIF_plant_occ_cleaned.csv"
date: "2023-07-25"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(countrycode)
library(CoordinateCleaner)
library(dplyr)
```

# Read in data
```{r}
TropicalAndes_GBIF_plant_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ.csv"))
```

# Clean data
Adapted code from https://ropensci.github.io/CoordinateCleaner/articles/Cleaning_GBIF_data_with_CoordinateCleaner.html

## Select columns of interest
```{r}
# Get column names
names(TropicalAndes_GBIF_plant_occ)
```

```{r}
# Subset to columns of interest
TropicalAndes_GBIF_plant_occ <- TropicalAndes_GBIF_plant_occ %>%
  dplyr::select()
# remove records without coordinates
TropicalAndes_GBIF_plant_occ <- TropicalAndes_GBIF_plant_occ %>%
  filter(!is.na(decimalLongitude)) %>%
  filter(!is.na(decimalLatitude))
```

```{r}
 #to avoid specifying it in each function
names(dat)[2:3] <- c("decimallongitude", "decimallatitude")

TropicalAndes_GBIF_plant_occ_clean <- TropicalAndes_GBIF_plant_occ%>%
  cc_val()%>%
  cc_equ()%>%
  cc_cap()%>%
  cc_cen()%>%
  cc_coun(iso3 = "countryCode")%>%
  cc_gbif()%>%
  cc_inst()%>%
  cc_sea()%>%
  cc_zero()%>%
  cc_outl()%>%
  cc_dupl()
```

### 2023-07-27 code/L1/L1_TropicalAndes_plant_lookupTable.Rmd
inital commit

---
title: "Tropical Andes plant Lookup Table"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: ""
data output: ""
date: "2023-07-25"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}

```

# Read in species lists
```{r}
TRY_species <- read.csv(file.path(data_path,"TropicalAndes_TRY_plant_species.csv"))
BIEN_species <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_trait_species.csv"))
```

```{r}
#convert species list into vectors
BIEN_species <- BIEN_species[,2]
TRY_species <- TRY_species[,2]
#what is in both
species <- intersect(TRY_species, BIEN_species)
# what doesn't match
all.equal(TRY_species, BIEN_species)

```

### 2023-07-27 code/L1/L1_TropicalAndes_plants_BIEN_occ.Rmd
inital commit

--- 2023-07-25 code/L1/L1_TropicalAndes_plants_GBIF.Rmd
@@
 ---
-title: "Cleaning Tropical Andes plant GBIF occurrence records"
+title: "Cleaning Tropical Andes plant BIEN occurrence records"
 author: "Hazel J. Anderson"
@@
 ```{r}
-TropicalAndes_GBIF_plant_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ.csv"))
+TropicalAndes_BIEN_plant_occ <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_occ.csv"))
 ```
@@
 # Get column names
-names(TropicalAndes_GBIF_plant_occ)
+names(TropicalAndes_BIEN_plant_occ)
 ```
@@
 # Subset to columns of interest
-TropicalAndes_GBIF_plant_occ <- TropicalAndes_GBIF_plant_occ %>%
+TropicalAndes_BIEN_plant_occ <- TropicalAndes_BIEN_plant_occ %>%
   dplyr::select()
 # remove records without coordinates
-TropicalAndes_GBIF_plant_occ <- TropicalAndes_GBIF_plant_occ %>%
+TropicalAndes_BIEN_plant_occ <- TropicalAndes_BIEN_plant_occ %>%
   filter(!is.na(decimalLongitude)) %>%
@@
 
-TropicalAndes_GBIF_plant_occ_clean <- TropicalAndes_GBIF_plant_occ%>%
+TropicalAndes_BIEN_plant_occ_clean <- TropicalAndes_BIEN_plant_occ%>%
   cc_val()%>%

### 2023-07-27 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
inital commit
change input file reference link

@@
+---
+title: ""
+author: "Hazel J. Anderson"
+project: "Plant-Frugivore Diversity"
+collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
+overview: ""
+data input: "iucn_habitatclassification_composite_lvl2_ver004.tif"
+data output: ""
+date: "2023-07-18"
+output: html_document
+---
+
 # Set file paths
@@
 library(raster)
+library(sf)
 ```
@@
 # Load in Ecosystem Functional Groups
-Data downloaded from www.zendo.org/record/5090419
+Data downloaded from https://zenodo.org/record/5090419                                                                                                 
 ```{r}
@@
 ```
+
+```{r}
+plot(iucn_habitats)
+```
+
+Crop extent
+```{r}
+
+```

### 2023-07-30 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
add clip extent and masking habitat types not forest

--- 2023-07-18 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
@@
+---
+title: "Tropical Andes IUCN Montane and Lowland forest"
+author: "Hazel J. Anderson"
+project: "Plant-Frugivore Diversity"
+collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
+overview: "Create raster of IUCN habitat classification Forest-Subtropical moist montane and Forest - Subtropical moist lowland for Tropical Andes."
+data input: "iucn_habitatclassification_composite_lvl2_ver004.tif"
+data output: "TropicalAndes_IUCNHabitat_Forest.GTiff"
+date: "2023-07-18"
+output: html_document
+---
+
 # Set file paths
@@
 library(raster)
+library(sf)
+library(rnaturalearth)
+library(dplyr)
 ```
@@
 # Load in Ecosystem Functional Groups
-Data downloaded from www.zendo.org/record/5090419
+Data downloaded from https://zenodo.org/record/4058819
+Citation: Martin Jung, Prabhat Raj Dahal, Stuart H.
+M.
+Butchart, Paul F.
+Donald, Xavier De Lamo, Myroslava Lesiv, Valerie Kapos, Carlo Rondinini, & Piero Visconti.
+(2020).
+A global map of terrestrial habitat types (Version 004) [Data set].
+Zenodo.
+https://doi.org/10.5281/zenodo.4058819
+
 ```{r}
-iucn_habitats <- raster(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004.tif"))
+iucn_habitats <- raster(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004/iucn_habitatclassification_composite_lvl2_ver004.tif"))
+iucn_habitats_colors <- read.csv(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004/color_classification.csv"))
 ```
+
+```{r}
+plot(iucn_habitats)
+```
+
+Crop extent - trim to Andean States
+```{r}
+# Pull in world map
+worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
+
+# Subset world map. In this case we are removing the Galapagos by defining the bounding box around the Ecuador polygon.
+study_region <-  worldMap %>% filter(sovereignt == "Bolivia" | sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
+plot(study_region$geometry)
+```
+
+```{r}
+iucn_habitats_crop <- crop(iucn_habitats, study_region)
+TA_iucn_habitats <- mask(iucn_habitats_crop, study_region)
+plot(TA_iucn_habitats)
+```
+
+value 109 is Forest-Subtropical moist montane & value 106 is Forest - Subtropical moist lowland
+see level2.clr or lvl2_style.qml file
+```{r}
+#rename for masking
+TA_iucn_habitats_mask <- TA_iucn_habitats
+TA_iucn_habitats_mask[TA_iucn_habitats$iucn_habitatclassification_composite_lvl2_ver004 != 106 | TA_iucn_habitats$iucn_habitatclassification_composite_lvl2_ver004 != 109] <- NA
+TA_iucn_habitats_forest <- mask(TA_iucn_habitats, TA_iucn_habitats_mask)
+
+plot(TA_iucn_habitats_forest)
+```
+
+# Save raster
+```{r}
+TA_icunHabitat_path <- file.path(output_path, "TropicalAndes_IUCNHabitat_Forest.GTiff")
+writeRaster(TA_iucnHabitats, filename = TA_icunHabitat_path, format = "GTiff", overwrite=TRUE)
+```

### 2023-07-30 code/L0/L0_TropicalAndes_frugivores_frugivoria_traits.Rmd

--- 2023-07-18 code/L0/L0_TropicalAndes_frugivores_frugivoria_traits.Rmd
@@
 output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+output_path2<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
 ```
 
-# Load required packages
+# Retrive Frugivoria data
+Using r code provided at https://portal.edirepository.org/nis/codeGeneration?packageId=edi.1220.6&statisticalFileType=r
 ```{r}
+# Package ID: edi.1220.6 Cataloging System:https://pasta.edirepository.org.
+# Data set title: Frugivoria: A trait database for birds and mammals exhibiting frugivory across contiguous Neotropical moist forests .
+# Data set creator:  Beth Gerstner - Michigan State University 
+# Data set creator:  Phoebe Zarnetske - Michigan State University 
+# Data set creator:  Patrick Bills - Michigan State University 
+# Contact:  Beth Gerstner -  Michigan State University  - gerstn11@msu.edu
+# Stylesheet v2.11 for metadata conversion into program: John H. Porter, Univ. Virginia, jporter@virginia.edu 
+```
 
+## Retrive file Frugivoria: Mammal database simple
+```{r}
+#file 3  
+inUrl3  <- "https://pasta.lternet.edu/package/data/eml/edi/1220/6/3c655f2ab1d525d1b1f05ee78153e875" 
+infile3 <- tempfile()
+try(download.file(inUrl3,infile3,method="curl"))
+if (is.na(file.size(infile3))) download.file(inUrl3,infile3,method="auto")
+
+ Frug_mammal <-read.csv(infile3,header=F 
+          ,skip=1
+            ,sep=","  
+                ,quot='"' 
+        , col.names=c(
+                    "IUCN_species_name",     
+                    "common_name",     
+                    "family",     
+                    "genus",     
+                    "species",     
+                    "subspecies",     
+                    "elton_species_name",     
+                    "diet_cat",     
+                    "diet_source_e",     
+                    "diet_breadth",     
+                    "diet_level",     
+                    "activity_nocturnal_e",     
+                    "activity_source_e",     
+                    "activity_level",     
+                    "for_strat_value_e",     
+                    "for_strat_certainty_e",     
+                    "body_mass_e",     
+                    "body_mass_level_e",     
+                    "body_mass_source_e",     
+                    "body_size_mm",     
+                    "body_size_notes",     
+                    "body_size_level",     
+                    "body_size_source",     
+                    "sexual_dimorphism",     
+                    "sexual_dimorphism_notes",     
+                    "sexual_dimorphism_level",     
+                    "sexual_dimorphism_source",     
+                    "longevity",     
+                    "longevity_notes",     
+                    "longevity_level",     
+                    "longevity_source",     
+                    "home_range_size",     
+                    "home_range_notes",     
+                    "home_range_level",     
+                    "home_range_source",     
+                    "habitat_specialization",     
+                    "habitat_specialization_source",     
+                    "generation_time",     
+                    "generation_time_notes",     
+                    "generation_time_level",     
+                    "generation_time_source",     
+                    "IUCN_category",     
+                    "habitat",     
+                    "habitat_breadth",     
+                    "habitat_breadth_source",     
+                    "habitat_level",     
+                    "mean_CHELSA_bio1_1981.2010_V.2.1",     
+                    "mean_CHELSA_bio12_1981.2010_V.2.1",     
+                    "mean_human_fp_range_2010",     
+                    "mean_human_fp_range_2020",     
+                    "percent_change_hf_2010_2020",     
+                    "inferred_range_sqkm",     
+                    "date_data_obtained",     
+                    "filled_by"    ), check.names=TRUE)
+
+unlink(infile3)
+
+# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
+
+if (class(Frug_mammal$IUCN_species_name)!="factor") Frug_mammal$IUCN_species_name<- as.factor(Frug_mammal$IUCN_species_name)
+if (class(Frug_mammal$common_name)!="factor") Frug_mammal$common_name<- as.factor(Frug_mammal$common_name)
+if (class(Frug_mammal$family)!="factor") Frug_mammal$family<- as.factor(Frug_mammal$family)
+if (class(Frug_mammal$genus)!="factor") Frug_mammal$genus<- as.factor(Frug_mammal$genus)
+if (class(Frug_mammal$species)!="factor") Frug_mammal$species<- as.factor(Frug_mammal$species)
+if (class(Frug_mammal$subspecies)!="factor") Frug_mammal$subspecies<- as.factor(Frug_mammal$subspecies)
+if (class(Frug_mammal$elton_species_name)!="factor") Frug_mammal$elton_species_name<- as.factor(Frug_mammal$elton_species_name)
+if (class(Frug_mammal$diet_cat)!="factor") Frug_mammal$diet_cat<- as.factor(Frug_mammal$diet_cat)
+if (class(Frug_mammal$diet_source_e)!="factor") Frug_mammal$diet_source_e<- as.factor(Frug_mammal$diet_source_e)
+if (class(Frug_mammal$diet_breadth)=="factor") Frug_mammal$diet_breadth <-as.numeric(levels(Frug_mammal$diet_breadth))[as.integer(Frug_mammal$diet_breadth) ]               
+if (class(Frug_mammal$diet_breadth)=="character") Frug_mammal$diet_breadth <-as.numeric(Frug_mammal$diet_breadth)
+if (class(Frug_mammal$diet_level)!="factor") Frug_mammal$diet_level<- as.factor(Frug_mammal$diet_level)
+if (class(Frug_mammal$activity_nocturnal_e)!="factor") Frug_mammal$activity_nocturnal_e<- as.factor(Frug_mammal$activity_nocturnal_e)
+if (class(Frug_mammal$activity_source_e)!="factor") Frug_mammal$activity_source_e<- as.factor(Frug_mammal$activity_source_e)
+if (class(Frug_mammal$activity_level)!="factor") Frug_mammal$activity_level<- as.factor(Frug_mammal$activity_level)
+if (class(Frug_mammal$for_strat_value_e)!="factor") Frug_mammal$for_strat_value_e<- as.factor(Frug_mammal$for_strat_value_e)
+if (class(Frug_mammal$for_strat_certainty_e)!="factor") Frug_mammal$for_strat_certainty_e<- as.factor(Frug_mammal$for_strat_certainty_e)
+if (class(Frug_mammal$body_mass_e)=="factor") Frug_mammal$body_mass_e <-as.numeric(levels(Frug_mammal$body_mass_e))[as.integer(Frug_mammal$body_mass_e) ]               
+if (class(Frug_mammal$body_mass_e)=="character") Frug_mammal$body_mass_e <-as.numeric(Frug_mammal$body_mass_e)
+if (class(Frug_mammal$body_mass_level_e)!="factor") Frug_mammal$body_mass_level_e<- as.factor(Frug_mammal$body_mass_level_e)
+if (class(Frug_mammal$body_mass_source_e)!="factor") Frug_mammal$body_mass_source_e<- as.factor(Frug_mammal$body_mass_source_e)
+if (class(Frug_mammal$body_size_mm)=="factor") Frug_mammal$body_size_mm <-as.numeric(levels(Frug_mammal$body_size_mm))[as.integer(Frug_mammal$body_size_mm) ]               
+if (class(Frug_mammal$body_size_mm)=="character") Frug_mammal$body_size_mm <-as.numeric(Frug_mammal$body_size_mm)
+if (class(Frug_mammal$body_size_notes)!="factor") Frug_mammal$body_size_notes<- as.factor(Frug_mammal$body_size_notes)
+if (class(Frug_mammal$body_size_level)!="factor") Frug_mammal$body_size_level<- as.factor(Frug_mammal$body_size_level)
+if (class(Frug_mammal$body_size_source)!="factor") Frug_mammal$body_size_source<- as.factor(Frug_mammal$body_size_source)
+if (class(Frug_mammal$sexual_dimorphism)!="factor") Frug_mammal$sexual_dimorphism<- as.factor(Frug_mammal$sexual_dimorphism)
+if (class(Frug_mammal$sexual_dimorphism_notes)!="factor") Frug_mammal$sexual_dimorphism_notes<- as.factor(Frug_mammal$sexual_dimorphism_notes)
+if (class(Frug_mammal$sexual_dimorphism_level)!="factor") Frug_mammal$sexual_dimorphism_level<- as.factor(Frug_mammal$sexual_dimorphism_level)
+if (class(Frug_mammal$sexual_dimorphism_source)!="factor") Frug_mammal$sexual_dimorphism_source<- as.factor(Frug_mammal$sexual_dimorphism_source)
+if (class(Frug_mammal$longevity)=="factor") Frug_mammal$longevity <-as.numeric(levels(Frug_mammal$longevity))[as.integer(Frug_mammal$longevity) ]               
+if (class(Frug_mammal$longevity)=="character") Frug_mammal$longevity <-as.numeric(Frug_mammal$longevity)
+if (class(Frug_mammal$longevity_notes)!="factor") Frug_mammal$longevity_notes<- as.factor(Frug_mammal$longevity_notes)
+if (class(Frug_mammal$longevity_level)!="factor") Frug_mammal$longevity_level<- as.factor(Frug_mammal$longevity_level)
+if (class(Frug_mammal$longevity_source)!="factor") Frug_mammal$longevity_source<- as.factor(Frug_mammal$longevity_source)
+if (class(Frug_mammal$home_range_size)=="factor") Frug_mammal$home_range_size <-as.numeric(levels(Frug_mammal$home_range_size))[as.integer(Frug_mammal$home_range_size) ]               
+if (class(Frug_mammal$home_range_size)=="character") Frug_mammal$home_range_size <-as.numeric(Frug_mammal$home_range_size)
+if (class(Frug_mammal$home_range_notes)!="factor") Frug_mammal$home_range_notes<- as.factor(Frug_mammal$home_range_notes)
+if (class(Frug_mammal$home_range_level)!="factor") Frug_mammal$home_range_level<- as.factor(Frug_mammal$home_range_level)
+if (class(Frug_mammal$home_range_source)!="factor") Frug_mammal$home_range_source<- as.factor(Frug_mammal$home_range_source)
+if (class(Frug_mammal$habitat_specialization)!="factor") Frug_mammal$habitat_specialization<- as.factor(Frug_mammal$habitat_specialization)
+if (class(Frug_mammal$habitat_specialization_source)!="factor") Frug_mammal$habitat_specialization_source<- as.factor(Frug_mammal$habitat_specialization_source)
+if (class(Frug_mammal$generation_time)=="factor") Frug_mammal$generation_time <-as.numeric(levels(Frug_mammal$generation_time))[as.integer(Frug_mammal$generation_time) ]               
+if (class(Frug_mammal$generation_time)=="character") Frug_mammal$generation_time <-as.numeric(Frug_mammal$generation_time)
+if (class(Frug_mammal$generation_time_notes)!="factor") Frug_mammal$generation_time_notes<- as.factor(Frug_mammal$generation_time_notes)
+if (class(Frug_mammal$generation_time_level)!="factor") Frug_mammal$generation_time_level<- as.factor(Frug_mammal$generation_time_level)
+if (class(Frug_mammal$generation_time_source)!="factor") Frug_mammal$generation_time_source<- as.factor(Frug_mammal$generation_time_source)
+if (class(Frug_mammal$IUCN_category)!="factor") Frug_mammal$IUCN_category<- as.factor(Frug_mammal$IUCN_category)
+if (class(Frug_mammal$habitat)!="factor") Frug_mammal$habitat<- as.factor(Frug_mammal$habitat)
+if (class(Frug_mammal$habitat_breadth)=="factor") Frug_mammal$habitat_breadth <-as.numeric(levels(Frug_mammal$habitat_breadth))[as.integer(Frug_mammal$habitat_breadth) ]               
+if (class(Frug_mammal$habitat_breadth)=="character") Frug_mammal$habitat_breadth <-as.numeric(Frug_mammal$habitat_breadth)
+if (class(Frug_mammal$habitat_breadth_source)!="factor") Frug_mammal$habitat_breadth_source<- as.factor(Frug_mammal$habitat_breadth_source)
+if (class(Frug_mammal$habitat_level)!="factor") Frug_mammal$habitat_level<- as.factor(Frug_mammal$habitat_level)
+if (class(Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1)=="factor") Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1 <-as.numeric(levels(Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1))[as.integer(Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1) ]
+if (class(Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1)=="character") Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1 <-as.numeric(Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1)
+if (class(Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1)=="factor") Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1 <-as.numeric(levels(Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1))[as.integer(Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1) ]
+if (class(Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1)=="character") Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1 <-as.numeric(Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1)
+if (class(Frug_mammal$mean_human_fp_range_2010)=="factor") Frug_mammal$mean_human_fp_range_2010 <-as.numeric(levels(Frug_mammal$mean_human_fp_range_2010))[as.integer(Frug_mammal$mean_human_fp_range_2010) ]
+if (class(Frug_mammal$mean_human_fp_range_2010)=="character") Frug_mammal$mean_human_fp_range_2010 <-as.numeric(Frug_mammal$mean_human_fp_range_2010)
+if (class(Frug_mammal$mean_human_fp_range_2020)=="factor") Frug_mammal$mean_human_fp_range_2020 <-as.numeric(levels(Frug_mammal$mean_human_fp_range_2020))[as.integer(Frug_mammal$mean_human_fp_range_2020) ]
+if (class(Frug_mammal$mean_human_fp_range_2020)=="character") Frug_mammal$mean_human_fp_range_2020 <-as.numeric(Frug_mammal$mean_human_fp_range_2020)
+if (class(Frug_mammal$percent_change_hf_2010_2020)=="factor") Frug_mammal$percent_change_hf_2010_2020 <-as.numeric(levels(Frug_mammal$percent_change_hf_2010_2020))[as.integer(Frug_mammal$percent_change_hf_2010_2020) ]
+if (class(Frug_mammal$percent_change_hf_2010_2020)=="character") Frug_mammal$percent_change_hf_2010_2020 <-as.numeric(Frug_mammal$percent_change_hf_2010_2020)
+if (class(Frug_mammal$inferred_range_sqkm)=="factor") Frug_mammal$inferred_range_sqkm <-as.numeric(levels(Frug_mammal$inferred_range_sqkm))[as.integer(Frug_mammal$inferred_range_sqkm) ]               
+if (class(Frug_mammal$inferred_range_sqkm)=="character") Frug_mammal$inferred_range_sqkm <-as.numeric(Frug_mammal$inferred_range_sqkm)
+if (class(Frug_mammal$date_data_obtained)!="factor") Frug_mammal$date_data_obtained<- as.factor(Frug_mammal$date_data_obtained)
+if (class(Frug_mammal$filled_by)!="factor") Frug_mammal$filled_by<- as.factor(Frug_mammal$filled_by)
+
+# Convert Missing Values to NA for non-dates
+
+Frug_mammal$common_name <- as.factor(ifelse((trimws(as.character(Frug_mammal$common_name))==trimws("NA")),NA,as.character(Frug_mammal$common_name)))
+Frug_mammal$subspecies <- as.factor(ifelse((trimws(as.character(Frug_mammal$subspecies))==trimws("NA")),NA,as.character(Frug_mammal$subspecies)))
+Frug_mammal$diet_level <- as.factor(ifelse((trimws(as.character(Frug_mammal$diet_level))==trimws("NA")),NA,as.character(Frug_mammal$diet_level)))
+Frug_mammal$activity_level <- as.factor(ifelse((trimws(as.character(Frug_mammal$activity_level))==trimws("NA")),NA,as.character(Frug_mammal$activity_level)))
+Frug_mammal$body_size_mm <- ifelse((trimws(as.character(Frug_mammal$body_size_mm))==trimws("NA")),NA,Frug_mammal$body_size_mm)               
+suppressWarnings(Frug_mammal$body_size_mm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_mammal$body_size_mm))==as.character(as.numeric("NA"))),NA,Frug_mammal$body_size_mm))
+Frug_mammal$body_size_notes <- as.factor(ifelse((trimws(as.character(Frug_mammal$body_size_notes))==trimws("NA")),NA,as.character(Frug_mammal$body_size_notes)))
+Frug_mammal$body_size_level <- as.factor(ifelse((trimws(as.character(Frug_mammal$body_size_level))==trimws("NA")),NA,as.character(Frug_mammal$body_size_level)))
+Frug_mammal$body_size_source <- as.factor(ifelse((trimws(as.character(Frug_mammal$body_size_source))==trimws("NA")),NA,as.character(Frug_mammal$body_size_source)))
+Frug_mammal$sexual_dimorphism <- as.factor(ifelse((trimws(as.character(Frug_mammal$sexual_dimorphism))==trimws("NA")),NA,as.character(Frug_mammal$sexual_dimorphism)))
+Frug_mammal$sexual_dimorphism <- as.factor(ifelse((trimws(as.character(Frug_mammal$sexual_dimorphism))==trimws("NA ")),NA,as.character(Frug_mammal$sexual_dimorphism)))
+Frug_mammal$sexual_dimorphism_notes <- as.factor(ifelse((trimws(as.character(Frug_mammal$sexual_dimorphism_notes))==trimws("NA")),NA,as.character(Frug_mammal$sexual_dimorphism_notes)))
+Frug_mammal$sexual_dimorphism_level <- as.factor(ifelse((trimws(as.character(Frug_mammal$sexual_dimorphism_level))==trimws("NA")),NA,as.character(Frug_mammal$sexual_dimorphism_level)))
+Frug_mammal$sexual_dimorphism_source <- as.factor(ifelse((trimws(as.character(Frug_mammal$sexual_dimorphism_source))==trimws("NA")),NA,as.character(Frug_mammal$sexual_dimorphism_source)))
+Frug_mammal$longevity <- ifelse((trimws(as.character(Frug_mammal$longevity))==trimws("NA")),NA,Frug_mammal$longevity)               
+suppressWarnings(Frug_mammal$longevity <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_mammal$longevity))==as.character(as.numeric("NA"))),NA,Frug_mammal$longevity))
+Frug_mammal$longevity_notes <- as.factor(ifelse((trimws(as.character(Frug_mammal$longevity_notes))==trimws("NA")),NA,as.character(Frug_mammal$longevity_notes)))
+Frug_mammal$longevity_level <- as.factor(ifelse((trimws(as.character(Frug_mammal$longevity_level))==trimws("NA")),NA,as.character(Frug_mammal$longevity_level)))
+Frug_mammal$longevity_source <- as.factor(ifelse((trimws(as.character(Frug_mammal$longevity_source))==trimws("NA")),NA,as.character(Frug_mammal$longevity_source)))
+Frug_mammal$home_range_size <- ifelse((trimws(as.character(Frug_mammal$home_range_size))==trimws("NA")),NA,Frug_mammal$home_range_size)               
+suppressWarnings(Frug_mammal$home_range_size <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_mammal$home_range_size))==as.character(as.numeric("NA"))),NA,Frug_mammal$home_range_size))
+Frug_mammal$home_range_notes <- as.factor(ifelse((trimws(as.character(Frug_mammal$home_range_notes))==trimws("NA")),NA,as.character(Frug_mammal$home_range_notes)))
+Frug_mammal$home_range_level <- as.factor(ifelse((trimws(as.character(Frug_mammal$home_range_level))==trimws("NA")),NA,as.character(Frug_mammal$home_range_level)))
+Frug_mammal$home_range_source <- as.factor(ifelse((trimws(as.character(Frug_mammal$home_range_source))==trimws("NA")),NA,as.character(Frug_mammal$home_range_source)))
+Frug_mammal$habitat_specialization <- as.factor(ifelse((trimws(as.character(Frug_mammal$habitat_specialization))==trimws("NA")),NA,as.character(Frug_mammal$habitat_specialization)))
+Frug_mammal$habitat_specialization_source <- as.factor(ifelse((trimws(as.character(Frug_mammal$habitat_specialization_source))==trimws("NA")),NA,as.character(Frug_mammal$habitat_specialization_source)))
+Frug_mammal$generation_time <- ifelse((trimws(as.character(Frug_mammal$generation_time))==trimws("NA")),NA,Frug_mammal$generation_time)               
+suppressWarnings(Frug_mammal$generation_time <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_mammal$generation_time))==as.character(as.numeric("NA"))),NA,Frug_mammal$generation_time))
+Frug_mammal$generation_time_notes <- as.factor(ifelse((trimws(as.character(Frug_mammal$generation_time_notes))==trimws("NA")),NA,as.character(Frug_mammal$generation_time_notes)))
+Frug_mammal$generation_time_level <- as.factor(ifelse((trimws(as.character(Frug_mammal$generation_time_level))==trimws("NA")),NA,as.character(Frug_mammal$generation_time_level)))
+Frug_mammal$generation_time_source <- as.factor(ifelse((trimws(as.character(Frug_mammal$generation_time_source))==trimws("NA")),NA,as.character(Frug_mammal$generation_time_source)))
+Frug_mammal$habitat_breadth <- ifelse((trimws(as.character(Frug_mammal$habitat_breadth))==trimws("NA")),NA,Frug_mammal$habitat_breadth)               
+suppressWarnings(Frug_mammal$habitat_breadth <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_mammal$habitat_breadth))==as.character(as.numeric("NA"))),NA,Frug_mammal$habitat_breadth))
+Frug_mammal$habitat_level <- as.factor(ifelse((trimws(as.character(Frug_mammal$habitat_level))==trimws("NA")),NA,as.character(Frug_mammal$habitat_level)))
+Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1 <- ifelse((trimws(as.character(Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1))==trimws("NA")),NA,Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1)
+suppressWarnings(Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1))==as.character(as.numeric("NA"))),NA,Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1))
+Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1 <- ifelse((trimws(as.character(Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1))==trimws("NA")),NA,Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1)
+suppressWarnings(Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1))==as.character(as.numeric("NA"))),NA,Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1))
+Frug_mammal$mean_human_fp_range_2010 <- ifelse((trimws(as.character(Frug_mammal$mean_human_fp_range_2010))==trimws("NA")),NA,Frug_mammal$mean_human_fp_range_2010)               
+suppressWarnings(Frug_mammal$mean_human_fp_range_2010 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_mammal$mean_human_fp_range_2010))==as.character(as.numeric("NA"))),NA,Frug_mammal$mean_human_fp_range_2010))
+Frug_mammal$mean_human_fp_range_2020 <- ifelse((trimws(as.character(Frug_mammal$mean_human_fp_range_2020))==trimws("NA")),NA,Frug_mammal$mean_human_fp_range_2020)               
+suppressWarnings(Frug_mammal$mean_human_fp_range_2020 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_mammal$mean_human_fp_range_2020))==as.character(as.numeric("NA"))),NA,Frug_mammal$mean_human_fp_range_2020))
+Frug_mammal$percent_change_hf_2010_2020 <- ifelse((trimws(as.character(Frug_mammal$percent_change_hf_2010_2020))==trimws("NA")),NA,Frug_mammal$percent_change_hf_2010_2020)               
+suppressWarnings(Frug_mammal$percent_change_hf_2010_2020 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_mammal$percent_change_hf_2010_2020))==as.character(as.numeric("NA"))),NA,Frug_mammal$percent_change_hf_2010_2020))
+Frug_mammal$inferred_range_sqkm <- ifelse((trimws(as.character(Frug_mammal$inferred_range_sqkm))==trimws("NA")),NA,Frug_mammal$inferred_range_sqkm)               
+suppressWarnings(Frug_mammal$inferred_range_sqkm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_mammal$inferred_range_sqkm))==as.character(as.numeric("NA"))),NA,Frug_mammal$inferred_range_sqkm))
+Frug_mammal$date_data_obtained <- as.factor(ifelse((trimws(as.character(Frug_mammal$date_data_obtained))==trimws("NA")),NA,as.character(Frug_mammal$date_data_obtained)))
+
+# Here is the structure of the input data frame:
+str(Frug_mammal)                            
+attach(Frug_mammal)                            
+# The analyses below are basic descriptions of the variables. After testing, they should be replaced.
+
+summary(IUCN_species_name)
+summary(common_name)
+summary(family)
+summary(genus)
+summary(species)
+summary(subspecies)
+summary(elton_species_name)
+summary(diet_cat)
+summary(diet_source_e)
+summary(diet_breadth)
+summary(diet_level)
+summary(activity_nocturnal_e)
+summary(activity_source_e)
+summary(activity_level)
+summary(for_strat_value_e)
+summary(for_strat_certainty_e)
+summary(body_mass_e)
+summary(body_mass_level_e)
+summary(body_mass_source_e)
+summary(body_size_mm)
+summary(body_size_notes)
+summary(body_size_level)
+summary(body_size_source)
+summary(sexual_dimorphism)
+summary(sexual_dimorphism_notes)
+summary(sexual_dimorphism_level)
+summary(sexual_dimorphism_source)
+summary(longevity)
+summary(longevity_notes)
+summary(longevity_level)
+summary(longevity_source)
+summary(home_range_size)
+summary(home_range_notes)
+summary(home_range_level)
+summary(home_range_source)
+summary(habitat_specialization)
+summary(habitat_specialization_source)
+summary(generation_time)
+summary(generation_time_notes)
+summary(generation_time_level)
+summary(generation_time_source)
+summary(IUCN_category)
+summary(habitat)
+summary(habitat_breadth)
+summary(habitat_breadth_source)
+summary(habitat_level)
+summary(mean_CHELSA_bio1_1981.2010_V.2.1)
+summary(mean_CHELSA_bio12_1981.2010_V.2.1)
+summary(mean_human_fp_range_2010)
+summary(mean_human_fp_range_2020)
+summary(percent_change_hf_2010_2020)
+summary(inferred_range_sqkm)
+summary(date_data_obtained)
+summary(filled_by) 
+                # Get more details on character variables
+
+summary(as.factor(Frug_mammal$IUCN_species_name)) 
+summary(as.factor(Frug_mammal$common_name)) 
+summary(as.factor(Frug_mammal$family)) 
+summary(as.factor(Frug_mammal$genus)) 
+summary(as.factor(Frug_mammal$species)) 
+summary(as.factor(Frug_mammal$subspecies)) 
+summary(as.factor(Frug_mammal$elton_species_name)) 
+summary(as.factor(Frug_mammal$diet_cat)) 
+summary(as.factor(Frug_mammal$diet_source_e)) 
+summary(as.factor(Frug_mammal$diet_level)) 
+summary(as.factor(Frug_mammal$activity_nocturnal_e)) 
+summary(as.factor(Frug_mammal$activity_source_e)) 
+summary(as.factor(Frug_mammal$activity_level)) 
+summary(as.factor(Frug_mammal$for_strat_value_e)) 
+summary(as.factor(Frug_mammal$for_strat_certainty_e)) 
+summary(as.factor(Frug_mammal$body_mass_level_e)) 
+summary(as.factor(Frug_mammal$body_mass_source_e)) 
+summary(as.factor(Frug_mammal$body_size_notes)) 
+summary(as.factor(Frug_mammal$body_size_level)) 
+summary(as.factor(Frug_mammal$body_size_source)) 
+summary(as.factor(Frug_mammal$sexual_dimorphism)) 
+summary(as.factor(Frug_mammal$sexual_dimorphism_notes)) 
+summary(as.factor(Frug_mammal$sexual_dimorphism_level)) 
+summary(as.factor(Frug_mammal$sexual_dimorphism_source)) 
+summary(as.factor(Frug_mammal$longevity_notes)) 
+summary(as.factor(Frug_mammal$longevity_level)) 
+summary(as.factor(Frug_mammal$longevity_source)) 
+summary(as.factor(Frug_mammal$home_range_notes)) 
+summary(as.factor(Frug_mammal$home_range_level)) 
+summary(as.factor(Frug_mammal$home_range_source)) 
+summary(as.factor(Frug_mammal$habitat_specialization)) 
+summary(as.factor(Frug_mammal$habitat_specialization_source)) 
+summary(as.factor(Frug_mammal$generation_time_notes)) 
+summary(as.factor(Frug_mammal$generation_time_level)) 
+summary(as.factor(Frug_mammal$generation_time_source)) 
+summary(as.factor(Frug_mammal$IUCN_category)) 
+summary(as.factor(Frug_mammal$habitat)) 
+summary(as.factor(Frug_mammal$habitat_breadth_source)) 
+summary(as.factor(Frug_mammal$habitat_level)) 
+summary(as.factor(Frug_mammal$date_data_obtained)) 
+summary(as.factor(Frug_mammal$filled_by))
+detach(Frug_mammal) 
+```
+
+## Retrive file Frugivoria: Bird database simple
+```{r}
+#file 4
+inUrl4  <- "https://pasta.lternet.edu/package/data/eml/edi/1220/6/5a86fde71322a1ff64d94ace0ed1982c" 
+infile4 <- tempfile()
+try(download.file(inUrl4,infile4,method="curl"))
+if (is.na(file.size(infile4))) download.file(inUrl4,infile4,method="auto")
+
+ Frug_bird <-read.csv(infile4,header=F 
+          ,skip=1
+            ,sep=","  
+                ,quot='"' 
+        , col.names=c(
+                    "IUCN_species_name",     
+                    "common_name",     
+                    "family_e",     
+                    "genus",     
+                    "species",     
+                    "elton_species_name",     
+                    "diet_cat_e",     
+                    "diet_source_e",     
+                    "diet_breadth",     
+                    "diet_level",     
+                    "activity_nocturnal_e",     
+                    "for_strat_ground_e",     
+                    "for_strat_understory_e",     
+                    "for_strat_midhigh_e",     
+                    "for_strat_canopy_e",     
+                    "for_strat_aerial_e",     
+                    "for_strat_spec_level",     
+                    "for_strat_source_e",     
+                    "body_mass_e",     
+                    "body_mass_level_e",     
+                    "body_mass_source_e",     
+                    "body_size_mm",     
+                    "body_size_notes",     
+                    "body_size_level",     
+                    "body_size_source",     
+                    "sexual_dimorphism",     
+                    "sexual_dimorphism_notes",     
+                    "sexual_dimorphism_level",     
+                    "sexual_dimorphism_source",     
+                    "longevity",     
+                    "longevity_notes",     
+                    "longevity_level",     
+                    "longevity_source",     
+                    "home_range_size",     
+                    "home_range_notes",     
+                    "home_range_level",     
+                    "home_range_source",     
+                    "habitat_specialization",     
+                    "habitat_specialization_source",     
+                    "generation_time",     
+                    "generation_time_notes",     
+                    "generation_time_level",     
+                    "generation_time_source",     
+                    "IUCN_category",     
+                    "habitat",     
+                    "habitat_breadth",     
+                    "habitat_breadth_source",     
+                    "habitat_level",     
+                    "mean_CHELSA_bio1_1981.2010_V.2.1",     
+                    "mean_CHELSA_bio12_1981.2010_V.2.1",     
+                    "mean_human_fp_range_2010",     
+                    "mean_human_fp_range_2020",     
+                    "percent_change_hf_2010_2020",     
+                    "inferred_range_sqkm",     
+                    "date_data_obtained",     
+                    "filled_by"    ), check.names=TRUE)
+
+unlink(infile4)
+
+# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
+
+if (class(Frug_bird$IUCN_species_name)!="factor") Frug_bird$IUCN_species_name<- as.factor(Frug_bird$IUCN_species_name)
+if (class(Frug_bird$common_name)!="factor") Frug_bird$common_name<- as.factor(Frug_bird$common_name)
+if (class(Frug_bird$family_e)!="factor") Frug_bird$family_e<- as.factor(Frug_bird$family_e)
+if (class(Frug_bird$genus)!="factor") Frug_bird$genus<- as.factor(Frug_bird$genus)
+if (class(Frug_bird$species)!="factor") Frug_bird$species<- as.factor(Frug_bird$species)
+if (class(Frug_bird$elton_species_name)!="factor") Frug_bird$elton_species_name<- as.factor(Frug_bird$elton_species_name)
+if (class(Frug_bird$diet_cat_e)!="factor") Frug_bird$diet_cat_e<- as.factor(Frug_bird$diet_cat_e)
+if (class(Frug_bird$diet_source_e)!="factor") Frug_bird$diet_source_e<- as.factor(Frug_bird$diet_source_e)
+if (class(Frug_bird$diet_breadth)=="factor") Frug_bird$diet_breadth <-as.numeric(levels(Frug_bird$diet_breadth))[as.integer(Frug_bird$diet_breadth) ]               
+if (class(Frug_bird$diet_breadth)=="character") Frug_bird$diet_breadth <-as.numeric(Frug_bird$diet_breadth)
+if (class(Frug_bird$diet_level)!="factor") Frug_bird$diet_level<- as.factor(Frug_bird$diet_level)
+if (class(Frug_bird$activity_nocturnal_e)!="factor") Frug_bird$activity_nocturnal_e<- as.factor(Frug_bird$activity_nocturnal_e)
+if (class(Frug_bird$for_strat_ground_e)=="factor") Frug_bird$for_strat_ground_e <-as.numeric(levels(Frug_bird$for_strat_ground_e))[as.integer(Frug_bird$for_strat_ground_e) ]               
+if (class(Frug_bird$for_strat_ground_e)=="character") Frug_bird$for_strat_ground_e <-as.numeric(Frug_bird$for_strat_ground_e)
+if (class(Frug_bird$for_strat_understory_e)=="factor") Frug_bird$for_strat_understory_e <-as.numeric(levels(Frug_bird$for_strat_understory_e))[as.integer(Frug_bird$for_strat_understory_e) ]
+if (class(Frug_bird$for_strat_understory_e)=="character") Frug_bird$for_strat_understory_e <-as.numeric(Frug_bird$for_strat_understory_e)
+if (class(Frug_bird$for_strat_midhigh_e)=="factor") Frug_bird$for_strat_midhigh_e <-as.numeric(levels(Frug_bird$for_strat_midhigh_e))[as.integer(Frug_bird$for_strat_midhigh_e) ]               
+if (class(Frug_bird$for_strat_midhigh_e)=="character") Frug_bird$for_strat_midhigh_e <-as.numeric(Frug_bird$for_strat_midhigh_e)
+if (class(Frug_bird$for_strat_canopy_e)=="factor") Frug_bird$for_strat_canopy_e <-as.numeric(levels(Frug_bird$for_strat_canopy_e))[as.integer(Frug_bird$for_strat_canopy_e) ]               
+if (class(Frug_bird$for_strat_canopy_e)=="character") Frug_bird$for_strat_canopy_e <-as.numeric(Frug_bird$for_strat_canopy_e)
+if (class(Frug_bird$for_strat_aerial_e)=="factor") Frug_bird$for_strat_aerial_e <-as.numeric(levels(Frug_bird$for_strat_aerial_e))[as.integer(Frug_bird$for_strat_aerial_e) ]               
+if (class(Frug_bird$for_strat_aerial_e)=="character") Frug_bird$for_strat_aerial_e <-as.numeric(Frug_bird$for_strat_aerial_e)
+if (class(Frug_bird$for_strat_spec_level)!="factor") Frug_bird$for_strat_spec_level<- as.factor(Frug_bird$for_strat_spec_level)
+if (class(Frug_bird$for_strat_source_e)!="factor") Frug_bird$for_strat_source_e<- as.factor(Frug_bird$for_strat_source_e)
+if (class(Frug_bird$body_mass_e)=="factor") Frug_bird$body_mass_e <-as.numeric(levels(Frug_bird$body_mass_e))[as.integer(Frug_bird$body_mass_e) ]               
+if (class(Frug_bird$body_mass_e)=="character") Frug_bird$body_mass_e <-as.numeric(Frug_bird$body_mass_e)
+if (class(Frug_bird$body_mass_level_e)!="factor") Frug_bird$body_mass_level_e<- as.factor(Frug_bird$body_mass_level_e)
+if (class(Frug_bird$body_mass_source_e)!="factor") Frug_bird$body_mass_source_e<- as.factor(Frug_bird$body_mass_source_e)
+if (class(Frug_bird$body_size_mm)=="factor") Frug_bird$body_size_mm <-as.numeric(levels(Frug_bird$body_size_mm))[as.integer(Frug_bird$body_size_mm) ]               
+if (class(Frug_bird$body_size_mm)=="character") Frug_bird$body_size_mm <-as.numeric(Frug_bird$body_size_mm)
+if (class(Frug_bird$body_size_notes)!="factor") Frug_bird$body_size_notes<- as.factor(Frug_bird$body_size_notes)
+if (class(Frug_bird$body_size_level)!="factor") Frug_bird$body_size_level<- as.factor(Frug_bird$body_size_level)
+if (class(Frug_bird$body_size_source)!="factor") Frug_bird$body_size_source<- as.factor(Frug_bird$body_size_source)
+if (class(Frug_bird$sexual_dimorphism)!="factor") Frug_bird$sexual_dimorphism<- as.factor(Frug_bird$sexual_dimorphism)
+if (class(Frug_bird$sexual_dimorphism_notes)!="factor") Frug_bird$sexual_dimorphism_notes<- as.factor(Frug_bird$sexual_dimorphism_notes)
+if (class(Frug_bird$sexual_dimorphism_level)!="factor") Frug_bird$sexual_dimorphism_level<- as.factor(Frug_bird$sexual_dimorphism_level)
+if (class(Frug_bird$sexual_dimorphism_source)!="factor") Frug_bird$sexual_dimorphism_source<- as.factor(Frug_bird$sexual_dimorphism_source)
+if (class(Frug_bird$longevity)=="factor") Frug_bird$longevity <-as.numeric(levels(Frug_bird$longevity))[as.integer(Frug_bird$longevity) ]               
+if (class(Frug_bird$longevity)=="character") Frug_bird$longevity <-as.numeric(Frug_bird$longevity)
+if (class(Frug_bird$longevity_notes)!="factor") Frug_bird$longevity_notes<- as.factor(Frug_bird$longevity_notes)
+if (class(Frug_bird$longevity_level)!="factor") Frug_bird$longevity_level<- as.factor(Frug_bird$longevity_level)
+if (class(Frug_bird$longevity_source)!="factor") Frug_bird$longevity_source<- as.factor(Frug_bird$longevity_source)
+if (class(Frug_bird$home_range_size)=="factor") Frug_bird$home_range_size <-as.numeric(levels(Frug_bird$home_range_size))[as.integer(Frug_bird$home_range_size) ]               
+if (class(Frug_bird$home_range_size)=="character") Frug_bird$home_range_size <-as.numeric(Frug_bird$home_range_size)
+if (class(Frug_bird$home_range_notes)!="factor") Frug_bird$home_range_notes<- as.factor(Frug_bird$home_range_notes)
+if (class(Frug_bird$home_range_level)!="factor") Frug_bird$home_range_level<- as.factor(Frug_bird$home_range_level)
+if (class(Frug_bird$home_range_source)!="factor") Frug_bird$home_range_source<- as.factor(Frug_bird$home_range_source)
+if (class(Frug_bird$habitat_specialization)!="factor") Frug_bird$habitat_specialization<- as.factor(Frug_bird$habitat_specialization)
+if (class(Frug_bird$habitat_specialization_source)!="factor") Frug_bird$habitat_specialization_source<- as.factor(Frug_bird$habitat_specialization_source)
+if (class(Frug_bird$generation_time)=="factor") Frug_bird$generation_time <-as.numeric(levels(Frug_bird$generation_time))[as.integer(Frug_bird$generation_time) ]               
+if (class(Frug_bird$generation_time)=="character") Frug_bird$generation_time <-as.numeric(Frug_bird$generation_time)
+if (class(Frug_bird$generation_time_notes)!="factor") Frug_bird$generation_time_notes<- as.factor(Frug_bird$generation_time_notes)
+if (class(Frug_bird$generation_time_level)!="factor") Frug_bird$generation_time_level<- as.factor(Frug_bird$generation_time_level)
+if (class(Frug_bird$generation_time_source)!="factor") Frug_bird$generation_time_source<- as.factor(Frug_bird$generation_time_source)
+if (class(Frug_bird$IUCN_category)!="factor") Frug_bird$IUCN_category<- as.factor(Frug_bird$IUCN_category)
+if (class(Frug_bird$habitat)!="factor") Frug_bird$habitat<- as.factor(Frug_bird$habitat)
+if (class(Frug_bird$habitat_breadth)=="factor") Frug_bird$habitat_breadth <-as.numeric(levels(Frug_bird$habitat_breadth))[as.integer(Frug_bird$habitat_breadth) ]               
+if (class(Frug_bird$habitat_breadth)=="character") Frug_bird$habitat_breadth <-as.numeric(Frug_bird$habitat_breadth)
+if (class(Frug_bird$habitat_breadth_source)!="factor") Frug_bird$habitat_breadth_source<- as.factor(Frug_bird$habitat_breadth_source)
+if (class(Frug_bird$habitat_level)!="factor") Frug_bird$habitat_level<- as.factor(Frug_bird$habitat_level)
+if (class(Frug_bird$mean_CHELSA_bio1_1981.2010_V.2.1)=="factor") Frug_bird$mean_CHELSA_bio1_1981.2010_V.2.1 <-as.numeric(levels(Frug_bird$mean_CHELSA_bio1_1981.2010_V.2.1))[as.integer(Frug_bird$mean_CHELSA_bio1_1981.2010_V.2.1) ]
+if (class(Frug_bird$mean_CHELSA_bio1_1981.2010_V.2.1)=="character") Frug_bird$mean_CHELSA_bio1_1981.2010_V.2.1 <-as.numeric(Frug_bird$mean_CHELSA_bio1_1981.2010_V.2.1)
+if (class(Frug_bird$mean_CHELSA_bio12_1981.2010_V.2.1)=="factor") Frug_bird$mean_CHELSA_bio12_1981.2010_V.2.1 <-as.numeric(levels(Frug_bird$mean_CHELSA_bio12_1981.2010_V.2.1))[as.integer(Frug_bird$mean_CHELSA_bio12_1981.2010_V.2.1) ]
+if (class(Frug_bird$mean_CHELSA_bio12_1981.2010_V.2.1)=="character") Frug_bird$mean_CHELSA_bio12_1981.2010_V.2.1 <-as.numeric(Frug_bird$mean_CHELSA_bio12_1981.2010_V.2.1)
+if (class(Frug_bird$mean_human_fp_range_2010)=="factor") Frug_bird$mean_human_fp_range_2010 <-as.numeric(levels(Frug_bird$mean_human_fp_range_2010))[as.integer(Frug_bird$mean_human_fp_range_2010) ]
+if (class(Frug_bird$mean_human_fp_range_2010)=="character") Frug_bird$mean_human_fp_range_2010 <-as.numeric(Frug_bird$mean_human_fp_range_2010)
+if (class(Frug_bird$mean_human_fp_range_2020)=="factor") Frug_bird$mean_human_fp_range_2020 <-as.numeric(levels(Frug_bird$mean_human_fp_range_2020))[as.integer(Frug_bird$mean_human_fp_range_2020) ]
+if (class(Frug_bird$mean_human_fp_range_2020)=="character") Frug_bird$mean_human_fp_range_2020 <-as.numeric(Frug_bird$mean_human_fp_range_2020)
+if (class(Frug_bird$percent_change_hf_2010_2020)=="factor") Frug_bird$percent_change_hf_2010_2020 <-as.numeric(levels(Frug_bird$percent_change_hf_2010_2020))[as.integer(Frug_bird$percent_change_hf_2010_2020) ]
+if (class(Frug_bird$percent_change_hf_2010_2020)=="character") Frug_bird$percent_change_hf_2010_2020 <-as.numeric(Frug_bird$percent_change_hf_2010_2020)
+if (class(Frug_bird$inferred_range_sqkm)=="factor") Frug_bird$inferred_range_sqkm <-as.numeric(levels(Frug_bird$inferred_range_sqkm))[as.integer(Frug_bird$inferred_range_sqkm) ]               
+if (class(Frug_bird$inferred_range_sqkm)=="character") Frug_bird$inferred_range_sqkm <-as.numeric(Frug_bird$inferred_range_sqkm)
+if (class(Frug_bird$date_data_obtained)!="factor") Frug_bird$date_data_obtained<- as.factor(Frug_bird$date_data_obtained)
+if (class(Frug_bird$filled_by)!="factor") Frug_bird$filled_by<- as.factor(Frug_bird$filled_by)
+
+# Convert Missing Values to NA for non-dates
+
+Frug_bird$IUCN_species_name <- as.factor(ifelse((trimws(as.character(Frug_bird$IUCN_species_name))==trimws("NA")),NA,as.character(Frug_bird$IUCN_species_name)))
+Frug_bird$common_name <- as.factor(ifelse((trimws(as.character(Frug_bird$common_name))==trimws("NA")),NA,as.character(Frug_bird$common_name)))
+Frug_bird$family_e <- as.factor(ifelse((trimws(as.character(Frug_bird$family_e))==trimws("NA")),NA,as.character(Frug_bird$family_e)))
+Frug_bird$genus <- as.factor(ifelse((trimws(as.character(Frug_bird$genus))==trimws("NA")),NA,as.character(Frug_bird$genus)))
+Frug_bird$elton_species_name <- as.factor(ifelse((trimws(as.character(Frug_bird$elton_species_name))==trimws("NA")),NA,as.character(Frug_bird$elton_species_name)))
+Frug_bird$diet_cat_e <- as.factor(ifelse((trimws(as.character(Frug_bird$diet_cat_e))==trimws("NA")),NA,as.character(Frug_bird$diet_cat_e)))
+Frug_bird$diet_level <- as.factor(ifelse((trimws(as.character(Frug_bird$diet_level))==trimws("NA")),NA,as.character(Frug_bird$diet_level)))
+Frug_bird$activity_nocturnal_e <- as.factor(ifelse((trimws(as.character(Frug_bird$activity_nocturnal_e))==trimws("NA")),NA,as.character(Frug_bird$activity_nocturnal_e)))
+Frug_bird$for_strat_spec_level <- as.factor(ifelse((trimws(as.character(Frug_bird$for_strat_spec_level))==trimws("NA")),NA,as.character(Frug_bird$for_strat_spec_level)))
+Frug_bird$for_strat_source_e <- as.factor(ifelse((trimws(as.character(Frug_bird$for_strat_source_e))==trimws("NA")),NA,as.character(Frug_bird$for_strat_source_e)))
+Frug_bird$body_mass_e <- ifelse((trimws(as.character(Frug_bird$body_mass_e))==trimws("NA")),NA,Frug_bird$body_mass_e)               
+suppressWarnings(Frug_bird$body_mass_e <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_bird$body_mass_e))==as.character(as.numeric("NA"))),NA,Frug_bird$body_mass_e))
+Frug_bird$body_mass_level_e <- as.factor(ifelse((trimws(as.character(Frug_bird$body_mass_level_e))==trimws("NA")),NA,as.character(Frug_bird$body_mass_level_e)))
+Frug_bird$body_mass_source_e <- as.factor(ifelse((trimws(as.character(Frug_bird$body_mass_source_e))==trimws("NA")),NA,as.character(Frug_bird$body_mass_source_e)))
+Frug_bird$body_size_mm <- ifelse((trimws(as.character(Frug_bird$body_size_mm))==trimws("NA")),NA,Frug_bird$body_size_mm)               
+suppressWarnings(Frug_bird$body_size_mm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_bird$body_size_mm))==as.character(as.numeric("NA"))),NA,Frug_bird$body_size_mm))
+Frug_bird$body_size_notes <- as.factor(ifelse((trimws(as.character(Frug_bird$body_size_notes))==trimws("NA")),NA,as.character(Frug_bird$body_size_notes)))
+Frug_bird$body_size_level <- as.factor(ifelse((trimws(as.character(Frug_bird$body_size_level))==trimws("NA")),NA,as.character(Frug_bird$body_size_level)))
+Frug_bird$body_size_source <- as.factor(ifelse((trimws(as.character(Frug_bird$body_size_source))==trimws("NA")),NA,as.character(Frug_bird$body_size_source)))
+Frug_bird$sexual_dimorphism <- as.factor(ifelse((trimws(as.character(Frug_bird$sexual_dimorphism))==trimws("NA")),NA,as.character(Frug_bird$sexual_dimorphism)))
+Frug_bird$sexual_dimorphism_notes <- as.factor(ifelse((trimws(as.character(Frug_bird$sexual_dimorphism_notes))==trimws("NA")),NA,as.character(Frug_bird$sexual_dimorphism_notes)))
+Frug_bird$sexual_dimorphism_level <- as.factor(ifelse((trimws(as.character(Frug_bird$sexual_dimorphism_level))==trimws("NA")),NA,as.character(Frug_bird$sexual_dimorphism_level)))
+Frug_bird$sexual_dimorphism_source <- as.factor(ifelse((trimws(as.character(Frug_bird$sexual_dimorphism_source))==trimws("NA")),NA,as.character(Frug_bird$sexual_dimorphism_source)))
+Frug_bird$longevity <- ifelse((trimws(as.character(Frug_bird$longevity))==trimws("NA")),NA,Frug_bird$longevity)               
+suppressWarnings(Frug_bird$longevity <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_bird$longevity))==as.character(as.numeric("NA"))),NA,Frug_bird$longevity))
+Frug_bird$longevity_notes <- as.factor(ifelse((trimws(as.character(Frug_bird$longevity_notes))==trimws("NA")),NA,as.character(Frug_bird$longevity_notes)))
+Frug_bird$longevity_level <- as.factor(ifelse((trimws(as.character(Frug_bird$longevity_level))==trimws("NA")),NA,as.character(Frug_bird$longevity_level)))
+Frug_bird$longevity_source <- as.factor(ifelse((trimws(as.character(Frug_bird$longevity_source))==trimws("NA")),NA,as.character(Frug_bird$longevity_source)))
+Frug_bird$home_range_size <- ifelse((trimws(as.character(Frug_bird$home_range_size))==trimws("NA")),NA,Frug_bird$home_range_size)               
+suppressWarnings(Frug_bird$home_range_size <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_bird$home_range_size))==as.character(as.numeric("NA"))),NA,Frug_bird$home_range_size))
+Frug_bird$home_range_notes <- as.factor(ifelse((trimws(as.character(Frug_bird$home_range_notes))==trimws("NA")),NA,as.character(Frug_bird$home_range_notes)))
+Frug_bird$home_range_level <- as.factor(ifelse((trimws(as.character(Frug_bird$home_range_level))==trimws("NA")),NA,as.character(Frug_bird$home_range_level)))
+Frug_bird$home_range_source <- as.factor(ifelse((trimws(as.character(Frug_bird$home_range_source))==trimws("NA")),NA,as.character(Frug_bird$home_range_source)))
+Frug_bird$habitat_specialization <- as.factor(ifelse((trimws(as.character(Frug_bird$habitat_specialization))==trimws("NA")),NA,as.character(Frug_bird$habitat_specialization)))
+Frug_bird$habitat_specialization_source <- as.factor(ifelse((trimws(as.character(Frug_bird$habitat_specialization_source))==trimws("NA")),NA,as.character(Frug_bird$habitat_specialization_source)))
+Frug_bird$generation_time <- ifelse((trimws(as.character(Frug_bird$generation_time))==trimws("NA")),NA,Frug_bird$generation_time)               
+suppressWarnings(Frug_bird$generation_time <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_bird$generation_time))==as.character(as.numeric("NA"))),NA,Frug_bird$generation_time))
+Frug_bird$generation_time_notes <- as.factor(ifelse((trimws(as.character(Frug_bird$generation_time_notes))==trimws("NA")),NA,as.character(Frug_bird$generation_time_notes)))
+Frug_bird$generation_time_source <- as.factor(ifelse((trimws(as.character(Frug_bird$generation_time_source))==trimws("NA")),NA,as.character(Frug_bird$generation_time_source)))
+Frug_bird$habitat_breadth <- ifelse((trimws(as.character(Frug_bird$habitat_breadth))==trimws("NA")),NA,Frug_bird$habitat_breadth)               
+suppressWarnings(Frug_bird$habitat_breadth <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_bird$habitat_breadth))==as.character(as.numeric("NA"))),NA,Frug_bird$habitat_breadth))
+Frug_bird$habitat_level <- as.factor(ifelse((trimws(as.character(Frug_bird$habitat_level))==trimws("NA")),NA,as.character(Frug_bird$habitat_level)))
+Frug_bird$mean_CHELSA_bio1_1981.2010_V.2.1 <- ifelse((trimws(as.character(Frug_bird$mean_CHELSA_bio1_1981.2010_V.2.1))==trimws("NA")),NA,Frug_bird$mean_CHELSA_bio1_1981.2010_V.2.1)               
+suppressWarnings(Frug_bird$mean_CHELSA_bio1_1981.2010_V.2.1 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_bird$mean_CHELSA_bio1_1981.2010_V.2.1))==as.character(as.numeric("NA"))),NA,Frug_bird$mean_CHELSA_bio1_1981.2010_V.2.1))
+Frug_bird$mean_CHELSA_bio12_1981.2010_V.2.1 <- ifelse((trimws(as.character(Frug_bird$mean_CHELSA_bio12_1981.2010_V.2.1))==trimws("NA")),NA,Frug_bird$mean_CHELSA_bio12_1981.2010_V.2.1)               
+suppressWarnings(Frug_bird$mean_CHELSA_bio12_1981.2010_V.2.1 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_bird$mean_CHELSA_bio12_1981.2010_V.2.1))==as.character(as.numeric("NA"))),NA,Frug_bird$mean_CHELSA_bio12_1981.2010_V.2.1))
+Frug_bird$mean_human_fp_range_2010 <- ifelse((trimws(as.character(Frug_bird$mean_human_fp_range_2010))==trimws("NA")),NA,Frug_bird$mean_human_fp_range_2010)               
+suppressWarnings(Frug_bird$mean_human_fp_range_2010 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_bird$mean_human_fp_range_2010))==as.character(as.numeric("NA"))),NA,Frug_bird$mean_human_fp_range_2010))
+Frug_bird$mean_human_fp_range_2020 <- ifelse((trimws(as.character(Frug_bird$mean_human_fp_range_2020))==trimws("NA")),NA,Frug_bird$mean_human_fp_range_2020)               
+suppressWarnings(Frug_bird$mean_human_fp_range_2020 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_bird$mean_human_fp_range_2020))==as.character(as.numeric("NA"))),NA,Frug_bird$mean_human_fp_range_2020))
+Frug_bird$percent_change_hf_2010_2020 <- ifelse((trimws(as.character(Frug_bird$percent_change_hf_2010_2020))==trimws("NA")),NA,Frug_bird$percent_change_hf_2010_2020)               
+suppressWarnings(Frug_bird$percent_change_hf_2010_2020 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_bird$percent_change_hf_2010_2020))==as.character(as.numeric("NA"))),NA,Frug_bird$percent_change_hf_2010_2020))
+Frug_bird$inferred_range_sqkm <- ifelse((trimws(as.character(Frug_bird$inferred_range_sqkm))==trimws("NA")),NA,Frug_bird$inferred_range_sqkm)               
+suppressWarnings(Frug_bird$inferred_range_sqkm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(Frug_bird$inferred_range_sqkm))==as.character(as.numeric("NA"))),NA,Frug_bird$inferred_range_sqkm))
+Frug_bird$date_data_obtained <- as.factor(ifelse((trimws(as.character(Frug_bird$date_data_obtained))==trimws("NA")),NA,as.character(Frug_bird$date_data_obtained)))
+
+# Here is the structure of the input data frame:
+str(Frug_bird)                            
+attach(Frug_bird)                            
+# The analyses below are basic descriptions of the variables. After testing, they should be replaced.
+
+summary(IUCN_species_name)
+summary(common_name)
+summary(family_e)
+summary(genus)
+summary(species)
+summary(elton_species_name)
+summary(diet_cat_e)
+summary(diet_source_e)
+summary(diet_breadth)
+summary(diet_level)
+summary(activity_nocturnal_e)
+summary(for_strat_ground_e)
+summary(for_strat_understory_e)
+summary(for_strat_midhigh_e)
+summary(for_strat_canopy_e)
+summary(for_strat_aerial_e)
+summary(for_strat_spec_level)
+summary(for_strat_source_e)
+summary(body_mass_e)
+summary(body_mass_level_e)
+summary(body_mass_source_e)
+summary(body_size_mm)
+summary(body_size_notes)
+summary(body_size_level)
+summary(body_size_source)
+summary(sexual_dimorphism)
+summary(sexual_dimorphism_notes)
+summary(sexual_dimorphism_level)
+summary(sexual_dimorphism_source)
+summary(longevity)
+summary(longevity_notes)
+summary(longevity_level)
+summary(longevity_source)
+summary(home_range_size)
+summary(home_range_notes)
+summary(home_range_level)
+summary(home_range_source)
+summary(habitat_specialization)
+summary(habitat_specialization_source)
+summary(generation_time)
+summary(generation_time_notes)
+summary(generation_time_level)
+summary(generation_time_source)
+summary(IUCN_category)
+summary(habitat)
+summary(habitat_breadth)
+summary(habitat_breadth_source)
+summary(habitat_level)
+summary(mean_CHELSA_bio1_1981.2010_V.2.1)
+summary(mean_CHELSA_bio12_1981.2010_V.2.1)
+summary(mean_human_fp_range_2010)
+summary(mean_human_fp_range_2020)
+summary(percent_change_hf_2010_2020)
+summary(inferred_range_sqkm)
+summary(date_data_obtained)
+summary(filled_by) 
+                # Get more details on character variables
+
+summary(as.factor(Frug_bird$IUCN_species_name)) 
+summary(as.factor(Frug_bird$common_name)) 
+summary(as.factor(Frug_bird$family_e)) 
+summary(as.factor(Frug_bird$genus)) 
+summary(as.factor(Frug_bird$species)) 
+summary(as.factor(Frug_bird$elton_species_name)) 
+summary(as.factor(Frug_bird$diet_cat_e)) 
+summary(as.factor(Frug_bird$diet_source_e)) 
+summary(as.factor(Frug_bird$diet_level)) 
+summary(as.factor(Frug_bird$activity_nocturnal_e)) 
+summary(as.factor(Frug_bird$for_strat_spec_level)) 
+summary(as.factor(Frug_bird$for_strat_source_e)) 
+summary(as.factor(Frug_bird$body_mass_level_e)) 
+summary(as.factor(Frug_bird$body_mass_source_e)) 
+summary(as.factor(Frug_bird$body_size_notes)) 
+summary(as.factor(Frug_bird$body_size_level)) 
+summary(as.factor(Frug_bird$body_size_source)) 
+summary(as.factor(Frug_bird$sexual_dimorphism)) 
+summary(as.factor(Frug_bird$sexual_dimorphism_notes)) 
+summary(as.factor(Frug_bird$sexual_dimorphism_level)) 
+summary(as.factor(Frug_bird$sexual_dimorphism_source)) 
+summary(as.factor(Frug_bird$longevity_notes)) 
+summary(as.factor(Frug_bird$longevity_level)) 
+summary(as.factor(Frug_bird$longevity_source)) 
+summary(as.factor(Frug_bird$home_range_notes)) 
+summary(as.factor(Frug_bird$home_range_level)) 
+summary(as.factor(Frug_bird$home_range_source)) 
+summary(as.factor(Frug_bird$habitat_specialization)) 
+summary(as.factor(Frug_bird$habitat_specialization_source)) 
+summary(as.factor(Frug_bird$generation_time_notes)) 
+summary(as.factor(Frug_bird$generation_time_level)) 
+summary(as.factor(Frug_bird$generation_time_source)) 
+summary(as.factor(Frug_bird$IUCN_category)) 
+summary(as.factor(Frug_bird$habitat)) 
+summary(as.factor(Frug_bird$habitat_breadth_source)) 
+summary(as.factor(Frug_bird$habitat_level)) 
+summary(as.factor(Frug_bird$date_data_obtained)) 
+summary(as.factor(Frug_bird$filled_by))
+detach(Frug_bird) 
+```
+
+## Retrive file Lookup table: Mammals exhibiting frugivory
+```{r}
+inUrl10  <- "https://pasta.lternet.edu/package/data/eml/edi/1220/6/f73cd1b16a4ed908dac4cf532690a12b" 
+infile10 <- tempfile()
+try(download.file(inUrl10,infile10,method="curl"))
+if (is.na(file.size(infile10))) download.file(inUrl10,infile10,method="auto")
+
+Frug_mammal_LookupTable <-read.csv(infile10,header=F 
+          ,skip=1
+            ,sep=","  
+                ,quot='"' 
+        , col.names=c(
+                    "IUCN_species_name",     
+                    "elton_species_name",     
+                    "PanTHERIA_species_name",     
+                    "habitat",     
+                    "new_species",     
+                    "reclassified",     
+                    "spelling_differences"    ), check.names=TRUE)
+
+unlink(infile10)
+
+# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
+
+if (class(Frug_mammal_LookupTable$IUCN_species_name)!="factor") Frug_mammal_LookupTable$IUCN_species_name<- as.factor(Frug_mammal_LookupTable$IUCN_species_name)
+if (class(Frug_mammal_LookupTable$elton_species_name)!="factor") Frug_mammal_LookupTable$elton_species_name<- as.factor(Frug_mammal_LookupTable$elton_species_name)
+if (class(Frug_mammal_LookupTable$PanTHERIA_species_name)!="factor") Frug_mammal_LookupTable$PanTHERIA_species_name<- as.factor(Frug_mammal_LookupTable$PanTHERIA_species_name)
+if (class(Frug_mammal_LookupTable$habitat)!="factor") Frug_mammal_LookupTable$habitat<- as.factor(Frug_mammal_LookupTable$habitat)
+if (class(Frug_mammal_LookupTable$new_species)!="factor") Frug_mammal_LookupTable$new_species<- as.factor(Frug_mammal_LookupTable$new_species)
+if (class(Frug_mammal_LookupTable$reclassified)!="factor") Frug_mammal_LookupTable$reclassified<- as.factor(Frug_mammal_LookupTable$reclassified)
+if (class(Frug_mammal_LookupTable$spelling_differences)!="factor") Frug_mammal_LookupTable$spelling_differences<- as.factor(Frug_mammal_LookupTable$spelling_differences)
+
+# Convert Missing Values to NA for non-dates
+
+Frug_mammal_LookupTable$habitat <- as.factor(ifelse((trimws(as.character(Frug_mammal_LookupTable$habitat))==trimws("NA")),NA,as.character(Frug_mammal_LookupTable$habitat)))
+Frug_mammal_LookupTable$new_species <- as.factor(ifelse((trimws(as.character(Frug_mammal_LookupTable$new_species))==trimws("NA")),NA,as.character(Frug_mammal_LookupTable$new_species)))
+Frug_mammal_LookupTable$reclassified <- as.factor(ifelse((trimws(as.character(Frug_mammal_LookupTable$reclassified))==trimws("NA")),NA,as.character(Frug_mammal_LookupTable$reclassified)))
+Frug_mammal_LookupTable$spelling_differences <- as.factor(ifelse((trimws(as.character(Frug_mammal_LookupTable$spelling_differences))==trimws("NA")),NA,as.character(Frug_mammal_LookupTable$spelling_differences)))
+
+# Here is the structure of the input data frame:
+str(Frug_mammal_LookupTable)                            
+attach(Frug_mammal_LookupTable)                            
+# The analyses below are basic descriptions of the variables. After testing, they should be replaced.
+
+summary(IUCN_species_name)
+summary(elton_species_name)
+summary(PanTHERIA_species_name)
+summary(habitat)
+summary(new_species)
+summary(reclassified)
+summary(spelling_differences) 
+                # Get more details on character variables
+
+summary(as.factor(Frug_mammal_LookupTable$IUCN_species_name)) 
+summary(as.factor(Frug_mammal_LookupTable$elton_species_name)) 
+summary(as.factor(Frug_mammal_LookupTable$PanTHERIA_species_name)) 
+summary(as.factor(Frug_mammal_LookupTable$habitat)) 
+summary(as.factor(Frug_mammal_LookupTable$new_species)) 
+summary(as.factor(Frug_mammal_LookupTable$reclassified)) 
+summary(as.factor(Frug_mammal_LookupTable$spelling_differences))
+detach(Frug_mammal_LookupTable)               
+```
+
+## Retive file Lookup table: Birds exhibiting frugivory
+```{r}
+inUrl11  <- "https://pasta.lternet.edu/package/data/eml/edi/1220/6/d1f32cc55b692cfb3d614cb400797299" 
+infile11 <- tempfile()
+try(download.file(inUrl11,infile11,method="curl"))
+if (is.na(file.size(infile11))) download.file(inUrl11,infile11,method="auto")
+
+ Frug_bird_LookupTable <-read.csv(infile11,header=F 
+          ,skip=1
+            ,sep=","  
+                ,quot='"' 
+        , col.names=c(
+                    "IUCN_species_name",     
+                    "elton_species_name",     
+                    "habitat",     
+                    "new_species",     
+                    "reclassified",     
+                    "spelling_difference"    ), check.names=TRUE)
+
+unlink(infile11)
+
+# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
+
+if (class(Frug_bird_LookupTable$IUCN_species_name)!="factor") Frug_bird_LookupTable$IUCN_species_name<- as.factor(Frug_bird_LookupTable$IUCN_species_name)
+if (class(Frug_bird_LookupTable$elton_species_name)!="factor") Frug_bird_LookupTable$elton_species_name<- as.factor(Frug_bird_LookupTable$elton_species_name)
+if (class(Frug_bird_LookupTable$habitat)!="factor") Frug_bird_LookupTable$habitat<- as.factor(Frug_bird_LookupTable$habitat)
+if (class(Frug_bird_LookupTable$new_species)!="factor") Frug_bird_LookupTable$new_species<- as.factor(Frug_bird_LookupTable$new_species)
+if (class(Frug_bird_LookupTable$reclassified)!="factor") Frug_bird_LookupTable$reclassified<- as.factor(Frug_bird_LookupTable$reclassified)
+if (class(Frug_bird_LookupTable$spelling_difference)!="factor") Frug_bird_LookupTable$spelling_difference<- as.factor(Frug_bird_LookupTable$spelling_difference)
+
+# Convert Missing Values to NA for non-dates
+
+Frug_bird_LookupTable$new_species <- as.factor(ifelse((trimws(as.character(Frug_bird_LookupTable$new_species))==trimws("NA")),NA,as.character(Frug_bird_LookupTable$new_species)))
+Frug_bird_LookupTable$reclassified <- as.factor(ifelse((trimws(as.character(Frug_bird_LookupTable$reclassified))==trimws("NA")),NA,as.character(Frug_bird_LookupTable$reclassified)))
+Frug_bird_LookupTable$spelling_difference <- as.factor(ifelse((trimws(as.character(Frug_bird_LookupTable$spelling_difference))==trimws("NA")),NA,as.character(Frug_bird_LookupTable$spelling_difference)))
+
+# Here is the structure of the input data frame:
+str(Frug_bird_LookupTable)                            
+attach(Frug_bird_LookupTable)                            
+# The analyses below are basic descriptions of the variables. After testing, they should be replaced.
+
+summary(IUCN_species_name)
+summary(elton_species_name)
+summary(habitat)
+summary(new_species)
+summary(reclassified)
+summary(spelling_difference) 
+                # Get more details on character variables
+
+summary(as.factor(Frug_bird_LookupTable$IUCN_species_name)) 
+summary(as.factor(Frug_bird_LookupTable$elton_species_name)) 
+summary(as.factor(Frug_bird_LookupTable$habitat)) 
+summary(as.factor(Frug_bird_LookupTable$new_species)) 
+summary(as.factor(Frug_bird_LookupTable$reclassified)) 
+summary(as.factor(Frug_bird_LookupTable$spelling_difference))
+detach(Frug_bird_LookupTable)
+```
+
+# Combine bird and mammal databases
+```{r}
+#rename columns to match
+Frug_bird <- Frug_bird %>% rename(family = family_e, diet_cat = diet_cat_e)
+#remove columns that don't match
+Frug_bird <- Frug_bird %>% select(-one_of("for_strat_ground_e", "for_strat_understory_e", "for_strat_midhigh_e", "for_strat_canopy_e", "for_strat_aerial_e", "for_strat_spec_level", "for_strat_source_e"))
+Frug_mammal <- Frug_mammal %>% select(-one_of("activity_source_e", "activity_level","for_strat_value_e", "for_strat_certainty_e", "subspecies")) 
 ```
@@
 ```{r}
+Frugivoria <- rbind(Frug_mammal, Frug_bird)
+```
 
+```{r}
+Frug_bird_LookupTable <- Frug_bird_LookupTable[,1:2]
+Frug_mammal_LookupTable <- Frug_mammal_LookupTable[,1:3]
+Frugivore_LookupTable <- merge(Frug_bird_LookupTable, Frug_mammal_LookupTable, all = TRUE)
 ```
@@
 ```{r}
-glimpse(d)
+glimpse(Frugivoria)
 ```
@@
 ```{r}
-print("The number of occurrences is")
-nrow(d)
+print("The number of records is")
+nrow(Frugivoria)
 print("The number of species is" )
-length(unique(d$species))
+length(unique(Frugivoria$species))
 print("The number of families is" )
-length(unique(d$family))
+length(unique(Frugivoria$family))
+```
+
+# Extract species list
+```{r}
+Frugivoria_SpeciesList <- unique(Frugivoria$species)
 ```
@@
 TropicalAndes_Frugivoria_path <- file.path(output_path,"TropicalAndes_Frugivoria_frugivore_traits.csv") 
-write.csv(d2, file = TropicalAndes_Frugivoria_path)
+write.csv(Frugivoria, file = TropicalAndes_Frugivoria_path)
+Frugivoria_species_path <- file.path(output_path,"TropicalAndes_Frugivoria_frugivore_traits_species.csv") 
+write.csv(Frugivoria_SpeciesList, file = Frugivoria_species_path)
+Frugivore_LookupTable_path <- file.path(output_path2,"TropicalAndes_frugivore_LookupTable.csv") 
+write.csv(Frugivore_LookupTable, file = Frugivore_LookupTable_path)
 ```

### 2023-07-30 code/L0/L0_TropicalAndes_frugivores_GBIF.Rmd

@@
 ```{r}
-d <- occ_download_get("0154158-230224095556074", path = output_path) %>%
+d <- occ_download_get("0154158-230224095556074", path = output_path, overwrite = TRUE) %>%
   occ_download_import()
@@
 ```{r}
-TropicalAndes_GBIF_path <- file.path(output_path,"TropAndes_GBIF_frugivore_occ.csv") 
+TropicalAndes_GBIF_path <- file.path(output_path,"TropicalAndes_GBIF_frugivore_occ.csv") 
 write.csv(d, file = TropicalAndes_GBIF_path)
-TropicalAndes_GBIF_species_path <- file.path(output_path,"TropAndes_GBIF_frugivore_occ_species.csv") 
+TropicalAndes_GBIF_species_path <- file.path(output_path,"TropicalAndes_GBIF_frugivore_occ_species.csv") 
 write.csv(GBIF_occ_SpeciesList, file = TropicalAndes_GBIF_path)

### 2023-07-31 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
update input files to just moist montane and lowland forest crop each to Tropical Andes combine the two resulting rasters save rasters as output files

@@
 Butchart, Paul F.
-Donald, Xavier De Lamo, Myroslava Lesiv, Valerie Kapos, Carlo Rondinini, & Piero Visconti.
+Donald, Xavier De Lamo, Myroslava Lesiv, … Piero Visconti.
 (2020).
@@
 ```{r}
-iucn_habitats <- raster(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004/iucn_habitatclassification_composite_lvl2_ver004.tif"))
-iucn_habitats_colors <- read.csv(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004/color_classification.csv"))
+iucn_moist_montane_forest <- raster(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004/lvl2_frac_1km_ver004/iucn_habitatclassification_fraction_lvl2__109_Forest – Subtropical-tropical moist montane__ver004.tif"))
+iucn_moist_lowland_forest <- raster(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004/lvl2_frac_1km_ver004/iucn_habitatclassification_fraction_lvl2__106_Forest – Subtropical-tropical moist lowland__ver004.tif"))
 ```
@@
 ```{r}
-plot(iucn_habitats)
+plot(iucn_moist_montane_forest)
+plot(iucn_moist_lowland_forest)
 ```
 
-Crop extent - trim to Andean States
+# Crop extent - trim to Andean States
 ```{r}
@@
 ```{r}
-iucn_habitats_crop <- crop(iucn_habitats, study_region)
-TA_iucn_habitats <- mask(iucn_habitats_crop, study_region)
-plot(TA_iucn_habitats)
+# crop moist montane to tropical andes
+iucn_moist_montane_forest_crop <- crop(iucn_moist_montane_forest, study_region)
+TA_iucn_moist_montane_forest <- mask(iucn_moist_montane_forest_crop, study_region)
+plot(TA_iucn_moist_montane_forest)
+TA_iucn_moist_montane_forest
 ```
 
-value 109 is Forest-Subtropical moist montane & value 106 is Forest - Subtropical moist lowland
-see level2.clr or lvl2_style.qml file
 ```{r}
-#rename for masking
-TA_iucn_habitats_mask <- TA_iucn_habitats
-TA_iucn_habitats_mask[TA_iucn_habitats$iucn_habitatclassification_composite_lvl2_ver004 != 106 | TA_iucn_habitats$iucn_habitatclassification_composite_lvl2_ver004 != 109] <- NA
-TA_iucn_habitats_forest <- mask(TA_iucn_habitats, TA_iucn_habitats_mask)
+#crop moist lowland to tropical andes
+iucn_moist_lowland_forest_crop <- crop(iucn_moist_lowland_forest, study_region)
+TA_iucn_moist_lowland_forest <- mask(iucn_moist_lowland_forest_crop, study_region)
+plot(TA_iucn_moist_lowland_forest)
+TA_iucn_moist_lowland_forest
+```
 
-plot(TA_iucn_habitats_forest)
+```{r}
+# combine TA_iucn_moist_lowland_forest & TA_iucn_moist_montane_forest into one raster
+TA_iucn_habitat_forest <- merge(TA_iucn_moist_lowland_forest, TA_iucn_moist_montane_forest)
+plot(TA_iucn_habitat_forest)
+TA_iucn_habitat_forest
 ```
@@
 ```{r}
-TA_icunHabitat_path <- file.path(output_path, "TropicalAndes_IUCNHabitat_Forest.GTiff")
-writeRaster(TA_iucnHabitats, filename = TA_icunHabitat_path, format = "GTiff", overwrite=TRUE)
+writeRaster(TA_iucn_habitat_forest, filename = file.path(output_path, "TropicalAndes_IUCNHabitat_Forest.GTiff"), format = "GTiff", overwrite=TRUE)
+writeRaster(TA_iucn_moist_montane_forest, filename = file.path(output_path, "TropicalAndes_IUCNHabitat_Forest_Montane.GTiff"), format = "GTiff", overwrite=TRUE)
+writeRaster(TA_iucn_moist_lowland_forest, filename = file.path(output_path, "TropicalAndes_IUCNHabitat_Forest_Lowland.GTiff"), format = "GTiff", overwrite=TRUE)
 ```

### 2023-07-31 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
convert to sf save shapefile

@@
 ```
+
+# Convert to sf
+```{r}
+TA_iucn_habitat_forest_shape <- rasterToPolygons(TA_iucn_habitat_forest, na.rm = TRUE, dissolve = TRUE)
+```
+
+```{r}
+plot(TA_iucn_habitat_forest_shape)
+```
+
+# Save shapefile
+```{r}
+shapefile(TA_iucn_habitat_forest_shape, filename = file.path(output_path, "TropicalAndes_IUCNHabitat_Forest.shp"), overwrite=TRUE)
+```

### 2023-07-31 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
convert to spatRaster, spatVector, and sf write output to file

--- 2023-07-31 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
@@
 library(dplyr)
+library(terra)
 ```
@@
 
-# Subset world map. In this case we are removing the Galapagos by defining the bounding box around the Ecuador polygon.
+# Subset world map to Tropical Andes
 study_region <-  worldMap %>% filter(sovereignt == "Bolivia" | sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
@@
 ```
+
+# Convert to sf
+```{r}
+# convert to SpatRaster
+TA_iucn_habitat_forest_sr <- as(TA_iucn_habitat_forest, "SpatRaster")
+TA_iucn_habitat_forest_sr
+```
+
+```{r}
+TA_iucn_habitat_forest_sv <- as.polygons(TA_iucn_habitat_forest_sr)
+TA_iucn_habitat_forest_sv
+plot(TA_iucn_habitat_forest_sv)
+```
+
+```{r}
+TA_iucn_habitat_forest_shape <- sf::st_as_sf(TA_iucn_habitat_forest_sv)
+TA_iucn_habitat_forest_shape
+plot(TA_iucn_habitat_forest_shape)
+```
+
+# Save shapefile
+```{r}
+st_write(TA_iucn_habitat_forest_shape, file.path(output_path, "TropicalAndes_IUCNHabitat_Forest.shp"), overwrite=TRUE)
+```

### 2023-07-31 code/L0/L0_TropicalAndes_plants_TRY_traits.Rmd
add code to summarize trait count correct spelling error (thanks cats)

@@
 library(rtry)
+library(dplyr)
 ```
@@
 ```{r}
-try_data <- rtry_importujkgh("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/25496_17032023010024/25496.txt")
+try_data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/25496_17032023010024/25496.txt")
 ```
@@
 
+```{r}
+try_data_TropicalAndes %>% count(TraitName)
+```
+
 # Write data to csv

### 2023-08-01 code/L1/L1_TropicalAndes_frugivores_GBIF.Rmd
inital commit

---
title: "Cleaning Tropical Andes frugivore GBIF occurrence records"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Uses CoordinateCleaner to flag and remove problematic records from GBIF frugivore occurrence records."
data input: "TropicalAndes_GBIF_frugivore_occ.csv"
data output: "TropicalAndes_GBIF_frugivore_occ_cleaned.csv, TropicalAndes_GBIF_frugivore_occ_flagged.csv"
date: "2023-07-25"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(countrycode)
library(CoordinateCleaner)
library(dplyr)
```

# Read in data
```{r}
TropicalAndes_GBIF_frugivore_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_frugivore_occ.csv"))
```

# Clean data
Adapted code from https://ropensci.github.io/CoordinateCleaner/articles/Cleaning_GBIF_data_with_CoordinateCleaner.html

```{r}
# remove records without coordinates
TropicalAndes_GBIF_frugivore_occ <- TropicalAndes_GBIF_frugivore_occ %>%
  filter(!is.na(decimalLongitude)) %>%
  filter(!is.na(decimalLatitude))
```

```{r}
flags <- clean_coordinates(x = TropicalAndes_GBIF_frugivore_occ,
                           lon = "decimalLongitude",
                           lat = "decimalLatitude",
                           countries = "countryCode",
                           species = "",
                           tests = c("capitals", "centroids", "countries",
                                     "duplicates", "equal", "gbif",
                                     "institutions", "seas","zeros"))
summary(flags)
plot(flags, lon = "decimalLongitude", lat = "decimalLatitude")
```

```{r}
#exclude problematic records
TropicalAndes_GBIF_frugivore_occ_cleaned <- TropicalAndes_GBIF_frugivore_occ[flags$summary,]
TropicalAndes_GBIF_frugivore_occ_flagged <- TropicalAndes_GBIF_frugivore_occ[!flags$summary,]
```

```{r}
table(TropicalAndes_GBIF_frugivore_occ_cleaned$year)
```

```{r}
# remove records from before second world war
TropicalAndes_GBIF_frugivore_occ_cleaned <- TropicalAndes_GBIF_frugivore_occ_cleaned %>%
  filter(year > 1945) 
```

# Write data to csv
```{r}
write.csv(TropicalAndes_GBIF_frugivore_occ_cleaned, file = file.path(output_path,"TropicalAndes_GBIF_frugivore_occ_cleaned.csv"))

write.csv(TropicalAndes_GBIF_frugivore_occ_flagged, file = file.path(output_path,"TropicalAndes_GBIF_frugivore_occ_flagged.csv"))
```

### 2023-08-01 code/L2/L2_2_TropicalAndes_taxdiv.Rmd
inital commit

--- 2023-04-12 MappingGBIF_Richness.RMD
@@
-Mapping GBIF Richness
-Adapted from https://luisdva.github.io/rstats/richness/
+---
+title: "Tropical Andes Taxonomic Diversity of Plants and Frugivores"
+author: "Hazel J. Anderson"
+project: "Plant-Frugivore Diversity"
+collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
+overview: "Calculating and mapping taxonomic diversity as species richness using occurrence data for plants and frugivores in the Tropical Andes Moist Lowland and Montane forests.
+Note: some code is adapted from https://luisdva.github.io/rstats/richness/."
+data input: "TropicalAndes_plant_occ.csv, TropicalAndes_frugivore_occ.csv"
+data output: "TropicalAndes_frugivore_richness.csv, TropicalAndes_GBIF_plant_richness.csv"
+date: "2023-08-01"
+output: html_document
+---
 
@@
 ```{r}
-data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
-output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
 ```
@@
 ```{r}
-NorthAndes_GBIF_plant_occ <- read.csv(file.path(data_path,"TropAndes_GBIF_plant_occ.csv"))
-NorthAndes_GBIF_frugivore_occ <- read.csv(file.path(data_path,"TropAndes_GBIF_frugivore_occ.csv"))
+TropicalAndes_plant_occ <- read.csv(file.path(data_path,"TropicalAndes_plant_occ.csv"))
+TropicalAndes_frugivore_occ <- read.csv(file.path(data_path,"TropicalAndes_frugivore_occ.csv"))
 ```
@@
 
+# Convert data to spatial data
 ```{r}
-plants.sf <- st_as_sf(NorthAndes_GBIF_plant_occ, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
-```
+plants.sf <- st_as_sf(TropicalAndes_plant_occ, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
 
-```{r}
-frugivores.sf <- st_as_sf(NorthAndes_GBIF_frugivore_occ, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
+frugivores.sf <- st_as_sf(TropicalAndes_frugivore_occ, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
 ```
@@
   pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Bird and Mammal Richness [0.5°]", x = "Latitude", y = "Longitude", fill = "Number of species")
+  labs(title = "Tropical Andes Frugivore Richness [0.5°]", x = "Latitude", y = "Longitude", fill = "Number of species")
 frugivoregridRichTA0.5
@@
 hist(frugivore_cellRichness0.5$num_species,
-     main = "Histogram of Bird & Mammal Richness by cell  [0.5°]",
-     xlab = "Bird & Mammal Richness by cell")
+     main = "Histogram of Frugivore Richness by cell  [0.5°]",
+     xlab = "Frugivore Richness by cell")
 ```
@@
 
-# Make 0.9 Grid
+# Write data to csv
 ```{r}
-# make grid
-TAGrid0.9 <- TApoly %>%
-  st_make_grid(cellsize = c(0.9)) %>%
-  st_intersection(TApoly) %>%
-  st_cast("MULTIPOLYGON") %>%
-  st_sf() %>%
-  mutate(cellid = row_number())
+write.csv(plant_cellRichness0.5, file.path(output_path,"TropicalAndes_GBIF_plant_richness.csv"))
+write.csv(frugivore_cellRichness0.5, file.path(output_path,"TropicalAndes_frugivore_richness.csv"))
 ```
-
-# Plot Empty Grid
-```{r}
-# empty grid
-gridPlot <-
-  ggplot() +
-  geom_sf(data = worldMap, fill = "white")+
-  geom_sf(data = TApoly) +
-  geom_sf(data = TAGrid0.9) +
-  labs(title = "Tropical Andes Grid [0.9°]", x = "Latitude", y = "Longitude") +
-  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  )
-gridPlot
-```
-
-# Cell Richness
-```{r}
-plant_richness_grid0.9 <- TAGrid0.9 %>%
-  st_join(plants_sf_species) %>%
-  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
-  group_by(cellid) %>%
-  summarize(num_species = sum(overlap))
-```
-
-```{r}
-frugivore_richness_grid0.9 <- TAGrid0.9 %>%
-  st_join(frugivores_sf_species) %>%
-  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
-  group_by(cellid) %>%
-  summarize(num_species = sum(overlap))
-```
-
-# Richness plots
-```{r}
-plantgridRichTA0.9 <-
-  ggplot(plant_richness_grid0.9) +
-  geom_sf(data = worldMap, fill = "white")+
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_sf(aes(fill = num_species), color = NA) +
-  scale_fill_scico(palette = "romaO", direction = 1, end = 0.9, midpoint = 1500) +
-  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  ) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Plant Richness [0.9°]", x = "Latitude", y = "Longitude", fill = "Number of species")
-plantgridRichTA0.9
-```
-
-```{r}
-frugivoregridRichTA0.9 <-
-  ggplot(frugivore_richness_grid0.9) +
-  geom_sf(data = worldMap, fill = "white")+
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_sf(aes(fill = num_species), color = NA) +
-  scale_fill_scico(palette = "romaO", direction = 1, end = 0.9, midpoint = 1500) +
-  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  ) +   
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Bird and Mammal Richness [0.9°]", x = "Latitude", y = "Longitude", fill = "Number of species")
-frugivoregridRichTA0.9
-```
-
-# Extract cell values
-```{r}
-plant_cellRichness0.9 <- plantgridRichTA0.9[["data"]]
-frugivore_cellRichness0.9 <- frugivoregridRichTA0.9[["data"]]
-```
-
-```{r}
-plot(plant_cellRichness0.9$num_species, frugivore_cellRichness0.9$num_species,
-     main = "Plant Richness vs Bird & Mammal Richness by cell  [0.9°]",
-     xlab = "Plant Richness by cell", ylab = "Bird & Mammal Richness by cell")
-abline(lm(frugivore_cellRichness0.9$num_species ~ plant_cellRichness0.9$num_species))
-summary(lm(frugivore_cellRichness0.9$num_species ~ plant_cellRichness0.9$num_species))
-```
-
-```{r}
-hist(plant_cellRichness0.9$num_species,
-     main = "Histogram of Plant Richness by cell  [0.9°]",
-     xlab = "Plant Richness by cell")
-hist(frugivore_cellRichness0.9$num_species,
-     main = "Histogram of Bird & Mammal Richness by cell  [0.9°]",
-     xlab = "Bird & Mammal Richness by cell")
-```
-
-```{r}
-summary(plant_cellRichness0.9$num_species)
-summary(frugivore_cellRichness0.9$num_species)
-```
-
-# Make 1.2 Grid
-```{r}
-# make grid
-TAGrid1.4 <- TApoly %>%
-  st_make_grid(cellsize = c(1.4)) %>%
-  st_intersection(TApoly) %>%
-  st_cast("MULTIPOLYGON") %>%
-  st_sf() %>%
-  mutate(cellid = row_number())
-```
-
-# Plot Empty Grid
-```{r}
-# empty grid
-gridPlot <-
-  ggplot() +
-  geom_sf(data = worldMap, fill = "white")+
-  geom_sf(data = TApoly) +
-  geom_sf(data = TAGrid1.4) +
-  labs(title = "Tropical Andes Grid  [1.4°]", x = "Latitude", y = "Longitude") +
-  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  )
-gridPlot
-```
-
-# Cell Richness
-```{r}
-plant_richness_grid1.4 <- TAGrid1.4 %>%
-  st_join(plants_sf_species) %>%
-  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
-  group_by(cellid) %>%
-  summarize(num_species = sum(overlap))
-```
-
-```{r}
-frugivore_richness_grid1.4 <- TAGrid1.4 %>%
-  st_join(frugivores_sf_species) %>%
-  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
-  group_by(cellid) %>%
-  summarize(num_species = sum(overlap))
-```
-
-# Richness plots
-```{r}
-plantgridRichTA1.4 <-
-  ggplot(plant_richness_grid1.4) +
-  geom_sf(data = worldMap, fill = "white")+
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_sf(aes(fill = num_species), color = NA) +
-  scale_fill_scico(palette = "romaO", direction = 1, end = 0.9) +
-  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  ) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Plant Richness  [1.4°]", x = "Latitude", y = "Longitude", fill = "Number of species")
-plantgridRichTA1.4
-```
-
-```{r}
-frugivoregridRichTA1.4 <-
-  ggplot(frugivore_richness_grid1.4) +
-  geom_sf(data = worldMap, fill = "white")+
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_sf(aes(fill = num_species), color = NA) +
-  scale_fill_scico(palette = "romaO", direction = 1, end = 0.9, midpoint = 1500) +
-  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  ) +   
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Bird and Mammal Richness [1.4°]", x = "Latitude", y = "Longitude", fill = "Number of species")
-frugivoregridRichTA1.4
-```
-
-# Extract cell values
-```{r}
-plant_cellRichness1.4 <- plantgridRichTA1.4[["data"]]
-frugivore_cellRichness1.4 <- frugivoregridRichTA1.4[["data"]]
-```
-
-```{r}
-plot(plant_cellRichness1.4$num_species, frugivore_cellRichness1.4$num_species,
-     main = "Plant Richness vs Bird & Mammal Richness by cell  [1.4°]",
-     xlab = "Plant Richness by cell", ylab = "Bird & Mammal Richness by cell")
-abline(lm(frugivore_cellRichness1.4$num_species ~ plant_cellRichness1.4$num_species))
-summary(lm(frugivore_cellRichness1.4$num_species ~ plant_cellRichness1.4$num_species))
-```
-
-```{r}
-hist(plant_cellRichness1.4$num_species,
-     main = "Histogram of Plant Richness by cell  [0.5°]",
-     xlab = "Plant Richness by cell")
-hist(frugivore_cellRichness1.4$num_species,
-     main = "Histogram of Bird & Mammal Richness by cell  [1.4°]",
-     xlab = "Bird & Mammal Richness by cell")
-```
-
-```{r}
-summary(plant_cellRichness1.4$num_species)
-summary(frugivore_cellRichness1.4$num_species)
-```

### 2023-08-01 code/L1/L1_TropicalAndes_plants_GBIF.Rmd
update overview and data output update flag record code add year cleaning add code to export data to csv

@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: ""
+overview: "Uses CoordinateCleaner to flag and remove problematic records from GBIF plant occurrence records."
 data input: "TropicalAndes_GBIF_plant_occ.csv"
-data output: "TropicalAndes_GBIF_plant_occ_cleaned.csv"
+data output: "TropicalAndes_GBIF_plant_occ_cleaned.csv, TropicalAndes_GBIF_plant_occ_flagged.csv"
 date: "2023-07-25"
@@
 
-## Select columns of interest
 ```{r}
-# Get column names
-names(TropicalAndes_GBIF_plant_occ)
-```
-
-```{r}
-# Subset to columns of interest
-TropicalAndes_GBIF_plant_occ <- TropicalAndes_GBIF_plant_occ %>%
-  dplyr::select()
 # remove records without coordinates
@@
 ```{r}
- #to avoid specifying it in each function
-names(dat)[2:3] <- c("decimallongitude", "decimallatitude")
+flags <- clean_coordinates(x = TropicalAndes_GBIF_plant_occ,
+                           lon = "decimalLongitude",
+                           lat = "decimalLatitude",
+                           countries = "countryCode",
+                           species = "",
+                           tests = c("capitals", "centroids", "countries",
+                                     "duplicates", "equal", "gbif",
+                                     "institutions", "seas","zeros"))
+summary(flags)
+plot(flags, lon = "decimalLongitude", lat = "decimalLatitude")
+```
 
-TropicalAndes_GBIF_plant_occ_clean <- TropicalAndes_GBIF_plant_occ%>%
-  cc_val()%>%
-  cc_equ()%>%
-  cc_cap()%>%
-  cc_cen()%>%
-  cc_coun(iso3 = "countryCode")%>%
-  cc_gbif()%>%
-  cc_inst()%>%
-  cc_sea()%>%
-  cc_zero()%>%
-  cc_outl()%>%
-  cc_dupl()
+```{r}
+#exclude problematic records
+TropicalAndes_GBIF_plant_occ_cleaned <- TropicalAndes_GBIF_plant_occ[flags$summary,]
+TropicalAndes_GBIF_plant_occ_flagged <- TropicalAndes_GBIF_plant_occ[!flags$summary,]
 ```
+
+```{r}
+table(TropicalAndes_GBIF_plant_occ_cleaned$year)
+```
+
+```{r}
+# remove records from before second world war
+TropicalAndes_GBIF_plant_occ_cleaned <- TropicalAndes_GBIF_plant_occ_cleaned %>%
+  filter(year > 1945) 
+```
+
+# Write data to csv
+```{r}
+write.csv(TropicalAndes_GBIF_plant_occ_cleaned, file = file.path(output_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
+
+write.csv(TropicalAndes_GBIF_plant_occ_flagged, file = file.path(output_path,"TropicalAndes_GBIF_plant_occ_flagged.csv"))
+```

### 2023-08-01 code/L1/L1_TropicalAndes_frugivores_GBIF.Rmd

@@
                            countries = "countryCode",
-                           species = "",
+                           species = "species",
                            tests = c("capitals", "centroids", "countries",
@@
 
+# Summary
+```{r}
+glimpse(TropicalAndes_GBIF_frugivore_occ_cleaned)
+```
+
+```{r}
+print("The number of records is")
+nrow(TropicalAndes_GBIF_frugivore_occ_cleaned)
+print("The number of species is")
+length(unique(TropicalAndes_GBIF_frugivore_occ_cleaned$species))
+print("The number of families is")
+length(unique(TropicalAndes_GBIF_frugivore_occ_cleaned$family))
+```
+
 # Write data to csv

### 2023-08-01 code/L1/L1_TropicalAndes_plants_GBIF.Rmd

@@
 
+# Summary
+```{r}
+glimpse(TropicalAndes_GBIF_plant_occ_cleaned)
+```
+
+```{r}
+print("The number of records is")
+nrow(TropicalAndes_GBIF_plant_occ_cleaned)
+print("The number of species is")
+length(unique(TropicalAndes_GBIF_plant_occ_cleaned$species))
+print("The number of families is")
+length(unique(TropicalAndes_GBIF_plant_occ_cleaned$family))
+```
+
 # Write data to csv

### 2023-08-01 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
adding code to subset to Tropical Andes lowland and montane forest

@@
 library(rgbif)
+library(raster)
+library(sf)
 ```
@@
 ```
+
+# Subset to Tropical Andes Lowland and Montane Forest
+```{r}
+TropicalAndes_IUCNHabitat_Forest <- st_read(file.path(data_path, "TropicalAndes_IUCNHabitat_Forest.shp"))
+# convert dataframe to spatial object
+coordinates(d) <- ~decimalLatitude+decimalLongitude
+crs(d) <- crs(TropicalAndes_IUCNHabitat_Forest)
+```
+
+```{r}
+GBIF_Forest <- d %>% filter(TropicalAndes_IUCNHabitat_Forest)
+GBIF_Forest
+plot(GBIF_Forest)
+```

### 2023-08-01 code/L0/L0_TropicalAndes_frugivores_frugivoria_traits.Rmd
correct species list to IUCN_species_name not species

@@
 print("The number of species is" )
-length(unique(Frugivoria$species))
+length(unique(Frugivoria$IUCN_species_name))
 print("The number of families is" )
@@
 ```{r}
-Frugivoria_SpeciesList <- unique(Frugivoria$species)
+Frugivoria_SpeciesList <- unique(Frugivoria$IUCN_species_name)
 ```

### 2023-08-02 code/L0/L0_TropicalAndes_frugivores_GBIF.Rmd
update introduction information update reading in Frugivoria species list and download

--- 2023-07-25 code/L0/L0_TropicalAndes_frugivores_GBIF.Rmd
@@
 author: "Hazel J. Anderson"
-collaborators: "None"
-data input: "None"
-data output: "TropAndes_GBIF_frugivore_occ.csv"
-project: "Frugivoria"
-date: "2023-04-06"
+project: "Plant-Frugivore Diversity"
+collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
+overview: "This script retrives plant occurrence data from GBIF using the species list from Frugivoria."
+data input: "TropicalAndes_Frugivoria_frugivore_traits_species.csv"
+data output: "TropicalAndes_GBIF_frugivore_occ.csv and TropicalAndes_GBIF_frugivore_occ_species.csv"
+date: "2023-07-25"
 output: html_document
@@
 
-# Read in IUCN data
-Data was obtained from https://www.iucnredlist.org/search?permalink=f4407d96-c039-413a-8d46-707f262271ee
+# Read in Frugivoria species list
 ```{r}
-frugivore_habitats <- read.csv(file.path(data_path,"IUCN/Tropical Andes Birds and Mammals - Search Results/habitats.csv"))
-```
-
-# Create species list from IUCN data
-```{r}
-frugivore_species <- unique(frugivore_habitats$scientificName)
-write.csv2(frugivore_species, file = file.path(output_path, "TropicalFrugivoreSpecies.csv"))
+frugivore_species <- read.csv(file.path(data_path,"TropicalAndes_Frugivoria_frugivore_traits_species.csv"))
 ```
@@
 ```{r}
-gbif_taxon_keys <- frugivore_species %>% 
+gbif_taxon_keys <- frugivore_species$x %>% 
 name_backbone_checklist() %>% # match to backbone 
@@
 ```{r}
-d <- occ_download_get("0154158-230224095556074", path = output_path) %>%
+d <- occ_download_get("0129743-230530130749713", path = output_path, overwrite = TRUE) %>%
   occ_download_import()
@@
 ```{r}
-TropicalAndes_GBIF_path <- file.path(output_path,"TropAndes_GBIF_frugivore_occ.csv") 
-write.csv(d, file = TropicalAndes_GBIF_path)
-TropicalAndes_GBIF_species_path <- file.path(output_path,"TropAndes_GBIF_frugivore_occ_species.csv") 
-write.csv(GBIF_occ_SpeciesList, file = TropicalAndes_GBIF_path)
+write.csv(d, file = file.path(output_path,"TropicalAndes_GBIF_frugivore_occ.csv"))
+
+write.csv(GBIF_occ_SpeciesList, file = file.path(output_path,"TropicalAndes_GBIF_frugivore_occ_species.csv"))
 ```

### 2023-08-02 code/L1/L1_TropicalAndes_plants_GBIF.Rmd
update tests, remove plot, and update record removal

@@
                            species = "species",
-                           tests = c("capitals", "centroids", "countries",
+                           tests = c("capitals", "centroids",
                                      "duplicates", "equal", "gbif",
@@
 summary(flags)
-plot(flags, lon = "decimalLongitude", lat = "decimalLatitude")
 ```
@@
 #exclude problematic records
-TropicalAndes_GBIF_plant_occ_cleaned <- TropicalAndes_GBIF_plant_occ[flags$summary,]
-TropicalAndes_GBIF_plant_occ_flagged <- TropicalAndes_GBIF_plant_occ[!flags$summary,]
+TropicalAndes_GBIF_plant_occ_cleaned <- TropicalAndes_GBIF_plant_occ[flags$.summary,]
+TropicalAndes_GBIF_plant_occ_flagged <- TropicalAndes_GBIF_plant_occ[!flags$.summary,]
 ```

### 2023-08-02 code/L1/L1_TropicalAndes_subset_occ_forest.Rmd
inital commit

---
title: "Subset occurrence data to IUCN Tropical Andes Lowland and Montane Forest"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script retrives subsets plant and frugivore occurrence data to IUCN Tropical Andes Lowland and Montane Forest."
data input: "
data output: "TropAndes_GBIF_plant_occ.csv"
date: "2023-07-18"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(dplyr)
library(raster)
library(sf)
```

# Read in data
```{r}
TropicalAndes_IUCNHabitat_Forest <- raster(file.path(data_path, "TropicalAndes_IUCNHabitat_Forest.tif"))
TropicalAndes_plant_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
TropicalAndes_frugivore_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_frugivore_occ_cleaned.csv"))
```

```{r}
TropicalAndes_IUCNHabitat_Forest
```

# Rasterize occurrence data
```{r}

```

# Crop occurrence data to forest extent
```{r}

```

# Extract species list for plants and frugivore
```{r}
TropicalAndes_plant_SpeciesList <- unique(TropicalAndes_BIEN_traits$species)
TropicalAndes_frugivore_SpeciesList <- unique(TropicalAndes_BIEN_traits$species)
```

# Write data to csv
```{r}
write.csv(TropicalAndes_BIEN_traits, file = file.path(output_path,"TropicalAndes_BIEN_plant_traits.csv"))
write.csv(TropicalAndes_BIEN_traits, file = file.path(output_path,"TropicalAndes_BIEN_plant_traits.csv"))
write.csv(TropicalAndes_BIEN_traits, file = file.path(output_path,"TropicalAndes_BIEN_plant_traits.csv"))
write.csv(TropicalAndes_BIEN_traits, file = file.path(output_path,"TropicalAndes_BIEN_plant_traits.csv"))
```

### 2023-08-02 code/L0/L0_TropicalAndes_plants_BIEN_occ.Rmd
add combine occ by country code
load forest shape unsuccessfully get occurrence data from BIEN

@@
 library(dplyr)
+library(sf)
+```
+
+# Load Tropical Andes moist lowland and montane forest shape
+```{r}
+TropicalAndes_IUCNHabitat_Forest <- st_read(file.path(data_path, "TropicalAndes_IUCNHabitat_Forest.shp"))
 ```
@@
 ```{r}
-TropicalAndes_BIEN_occ_ec <- BIEN_occurrence_country(country = "Ecuador", all.taxonomy = TRUE, observation.type = TRUE, political.boundaries = TRUE, collection.info = TRUE)
-
-TropicalAndes_BIEN_occ_ve <- BIEN_occurrence_country(country = "Venezuela", all.taxonomy = TRUE, observation.type = TRUE, political.boundaries = TRUE, collection.info = TRUE)
-
-TropicalAndes_BIEN_occ_pe <- BIEN_occurrence_country(country = "Peru", all.taxonomy = TRUE, observation.type = TRUE, political.boundaries = TRUE, collection.info = TRUE)
-
-TropicalAndes_BIEN_occ_co <- BIEN_occurrence_country(country = "Colombia", all.taxonomy = TRUE, observation.type = TRUE, political.boundaries = TRUE, collection.info = TRUE)
-
-TropicalAndes_BIEN_occ_bo <- BIEN_occurrence_country(country = "Bolivia", all.taxonomy = TRUE, observation.type = TRUE, political.boundaries = TRUE, collection.info = TRUE)
-```
-
-```{r}
-# combine occ
-
+TropicalAndes_BIEN_occ <- BIEN_occurrence_sf(TropicalAndes_IUCNHabitat_Forest, all.taxonomy = TRUE, observation.type = TRUE, political.boundaries = TRUE, collection.info = TRUE)
 ```

### 2023-08-02 code/L1/L1_TropicalAndes_subset_occ_forest.Rmd
add code to rasterize data

@@
 data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+data_path2<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
 output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
@@
 ```{r}
-TropicalAndes_IUCNHabitat_Forest <- raster(file.path(data_path, "TropicalAndes_IUCNHabitat_Forest.tif"))
-TropicalAndes_plant_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
+TropicalAndes_IUCNHabitat_Forest <- raster(file.path(data_path2, "TropicalAndes_IUCNHabitat_Forest.tif"))
+#TropicalAndes_plant_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
 TropicalAndes_frugivore_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_frugivore_occ_cleaned.csv"))
@@
 
+```{r}
+TropicalAndes_IUCNHabitat_Forest0.5 <- aggregate(TropicalAndes_IUCNHabitat_Forest, 50, fun=mean, na.rm=TRUE)
+TropicalAndes_IUCNHabitat_Forest0.5
+```
+
 # Rasterize occurrence data
 ```{r}
-
+coordinates(TropicalAndes_GBIF_plant_occ_cleaned) <- ~ decimalLatitude + decimalLongitude
+coordinates(TropicalAndes_frugivore_occ) <- ~ decimalLatitude + decimalLongitude
+# coerce to SpatialPixelsDataFrame
+gridded(TropicalAndes_GBIF_plant_occ_cleaned) <- TRUE
+gridded(TropicalAndes_frugivore_occ) <- TRUE
+# coerce to raster
+plant.raster <- raster(TropicalAndes_GBIF_plant_occ_cleaned)
+frugivore.raster <- raster(TropicalAndes_frugivore_occ)
 ```
@@
 ```{r}
+plant_crop <- crop(plants.raster, TropicalAndes_IUCNHabitat_Forest)
+plant_occ_forest <- mask(plant_crop, TropicalAndes_IUCNHabitat_Forest)
+plot(plant_occ_forest)
+plant_occ_forest
+```
 
+```{r}
+frugivore_crop <- crop(frugivores.raster, TropicalAndes_IUCNHabitat_Forest)
+frugivore_occ_forest <- mask(frugivore_crop, TropicalAndes_IUCNHabitat_Forest)
+plot(frugivore_occ_forest)
+frugivore_occ_forest
 ```
@@
 ```{r}
-TropicalAndes_plant_SpeciesList <- unique(TropicalAndes_BIEN_traits$species)
-TropicalAndes_frugivore_SpeciesList <- unique(TropicalAndes_BIEN_traits$species)
+TropicalAndes_plant_Forest_SpeciesList <- unique(TropicalAndes_BIEN_traits$species)
+TropicalAndes_frugivore_Forest_SpeciesList <- unique(TropicalAndes_BIEN_traits$species)
 ```
@@
 ```{r}
+# subset plant occ, subset frugivore occ, plant species list, frugivore species list
 write.csv(TropicalAndes_BIEN_traits, file = file.path(output_path,"TropicalAndes_BIEN_plant_traits.csv"))
 write.csv(TropicalAndes_BIEN_traits, file = file.path(output_path,"TropicalAndes_BIEN_plant_traits.csv"))
-write.csv(TropicalAndes_BIEN_traits, file = file.path(output_path,"TropicalAndes_BIEN_plant_traits.csv"))
-write.csv(TropicalAndes_BIEN_traits, file = file.path(output_path,"TropicalAndes_BIEN_plant_traits.csv"))
+write.csv(TropicalAndes_plant_SpeciesList, file = file.path(output_path,"TropicalAndes_plant_Forest_SpeciesList.csv"))
+write.csv(TropicalAndes_frugivore_Forest_SpeciesList, file = file.path(output_path,"TropicalAndes_frugivore_Forest_SpeciesList"))
 ```

### 2023-08-02 code/L1/L1_TropicalAndes_subset_occ_forest.Rmd
update file names

@@
 overview: "This script retrives subsets plant and frugivore occurrence data to IUCN Tropical Andes Lowland and Montane Forest."
-data input: "
-data output: "TropAndes_GBIF_plant_occ.csv"
+data input: "TropicalAndes_IUCNHabitat_Forest.tif, TropicalAndes_GBIF_plant_occ_cleaned.csv, TropicalAndes_GBIF_frugivore_occ_cleaned.csv"
+data output: "TropicalAndes_plant_occ_forest.csv, TropicalAndes_frugivore_occ_forest.csv, TropicalAndes_plant_Forest_SpeciesList.csv, TropicalAndes_frugivore_Forest_SpeciesList.csv"
 date: "2023-07-18"
@@
 # subset plant occ, subset frugivore occ, plant species list, frugivore species list
-write.csv(TropicalAndes_BIEN_traits, file = file.path(output_path,"TropicalAndes_BIEN_plant_traits.csv"))
-write.csv(TropicalAndes_BIEN_traits, file = file.path(output_path,"TropicalAndes_BIEN_plant_traits.csv"))
+write.csv(plant_occ_forest, file = file.path(output_path,"TropicalAndes_plant_occ_forest.csv"))
+write.csv(frugivore_occ_forest, file = file.path(output_path,"TropicalAndes_frugivore_occ_forest.csv"))
 write.csv(TropicalAndes_plant_SpeciesList, file = file.path(output_path,"TropicalAndes_plant_Forest_SpeciesList.csv"))

### 2023-08-02 code/L1/L1_TropicalAndes_subset_occ_forest.Rmd
add code to extract values in TropicalAndes_IUCNHabitat_Forest from plant and frugivore occurrence add summary

@@
 ```{r}
+# convert to SpatialPointsDataFrame
 coordinates(TropicalAndes_GBIF_plant_occ_cleaned) <- ~ decimalLatitude + decimalLongitude
 coordinates(TropicalAndes_frugivore_occ) <- ~ decimalLatitude + decimalLongitude
-# coerce to SpatialPixelsDataFrame
-gridded(TropicalAndes_GBIF_plant_occ_cleaned) <- TRUE
-gridded(TropicalAndes_frugivore_occ) <- TRUE
-# coerce to raster
-plant.raster <- raster(TropicalAndes_GBIF_plant_occ_cleaned)
-frugivore.raster <- raster(TropicalAndes_frugivore_occ)
+TropicalAndes_GBIF_plant_occ_cleaned
+TropicalAndes_frugivore_occ
 ```
@@
 ```{r}
-plant_crop <- crop(plants.raster, TropicalAndes_IUCNHabitat_Forest)
-plant_occ_forest <- mask(plant_crop, TropicalAndes_IUCNHabitat_Forest)
-plot(plant_occ_forest)
-plant_occ_forest
+TropicalAndes_GBIF_plant_occ_cleaned$ref<- extract(TropicalAndes_IUCNHabitat_Forest, TropicalAndes_GBIF_plant_occ_cleaned)
+TropicalAndes_frugivore_occ$ref<- extract(TropicalAndes_IUCNHabitat_Forest, TropicalAndes_frugivore_occ)
 ```
@@
 ```{r}
-frugivore_crop <- crop(frugivores.raster, TropicalAndes_IUCNHabitat_Forest)
-frugivore_occ_forest <- mask(frugivore_crop, TropicalAndes_IUCNHabitat_Forest)
-plot(frugivore_occ_forest)
-frugivore_occ_forest
+TropicalAndes_GBIF_plant_occ_cleaned$ref <- as.character(TropicalAndes_GBIF_plant_occ_cleaned$ref)
+TropicalAndes_frugivore_occ$ref <- as.character(TropicalAndes_frugivore_occ$ref)
+```
+
+```{r}
+plant_occ_forest <- TropicalAndes_GBIF_plant_occ_cleaned@data
+frugivore_occ_forest <- TropicalAndes_frugivore_occ@data
+```
+
+```{r}
+#remove row if ref = NA
+plant_occ_forest <- plant_occ_forest[plant_occ_forest$ref != "NA", ]
+frugivore_occ_forest <- frugivore_occ_forest[frugivore_occ_forest$ref != "NA", ]
+```
+
+# Summary
+```{r}
+glimpse(plant_occ_forest)
+glimpse(frugivore_occ_forest)
+```
+
+```{r}
+print("The number of plant records is")
+nrow(plant_occ_forest)
+print("The number of plant species is")
+length(unique(plant_occ_forest$species))
+print("The number of plant families is")
+length(unique(plant_occ_forest$family))
+```
+
+```{r}
+print("The number of frugivore records is")
+nrow(frugivore_occ_forest)
+print("The number of frugivore species is")
+length(unique(frugivore_occ_forest$species))
+print("The number of frugivore families is")
+length(unique(frugivore_occ_forest$family))
 ```
@@
 ```{r}
-TropicalAndes_plant_Forest_SpeciesList <- unique(TropicalAndes_BIEN_traits$species)
-TropicalAndes_frugivore_Forest_SpeciesList <- unique(TropicalAndes_BIEN_traits$species)
+TropicalAndes_plant_Forest_SpeciesList <- unique(plant_occ_forest$species)
+TropicalAndes_frugivore_Forest_SpeciesList <- unique(frugivore_occ_forest$species)
 ```

### 2023-08-03 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
correct spelling errors Reclassify forest raster to keep cells with values over 500 remove shapefile code

@@
 
+# Crop moist montane to Tropical Andes
 ```{r}
-# crop moist montane to tropical andes
 iucn_moist_montane_forest_crop <- crop(iucn_moist_montane_forest, study_region)
@@
 
+# Crop moist lowland to Tropical Andes
 ```{r}
-#crop moist lowland to tropical andes
 iucn_moist_lowland_forest_crop <- crop(iucn_moist_lowland_forest, study_region)
@@
 
+# Combine TA_iucn_moist_lowland_forest & TA_iucn_moist_montane_forest into one raster
 ```{r}
-# combine TA_iucn_moist_lowland_forest & TA_iucn_moist_montane_forest into one raster
 TA_iucn_habitat_forest <- merge(TA_iucn_moist_lowland_forest, TA_iucn_moist_montane_forest)
@@
 TA_iucn_habitat_forest
+```
+
+# Reclassify forest raster with values above 500 to one value and less than 500 to NA
+```{r}
+mat <- c(0, 500, NA,
+        500, 1000, 1)
+rclmat <- matrix(mat, ncol = 3, byrow = TRUE)
+TA_iucn_habitat_forest <- reclassify(TA_iucn_habitat_forest, rclmat)
+TA_iucn_habitat_forest
+plot(TA_iucn_habitat_forest)
 ```
@@
 ```
-
-# Convert to sf
-```{r}
-# convert to SpatRaster
-TA_iucn_habitat_forest_sr <- as(TA_iucn_habitat_forest, "SpatRaster")
-TA_iucn_habitat_forest_sr
-```
-
-```{r}
-TA_iucn_habitat_forest_sv <- as.polygons(TA_iucn_habitat_forest_sr)
-TA_iucn_habitat_forest_sv
-plot(TA_iucn_habitat_forest_sv)
-```
-
-```{r}
-TA_iucn_habitat_forest_shape <- sf::st_as_sf(TA_iucn_habitat_forest_sv)
-TA_iucn_habitat_forest_shape
-plot(TA_iucn_habitat_forest_shape)
-```
-
-# Save shapefile
-```{r}
-st_write(TA_iucn_habitat_forest_shape, file.path(output_path, "TropicalAndes_IUCNHabitat_Forest.shp"), overwrite=TRUE)
-```

### 2023-08-03 code/L1/L1_TropicalAndes_subset_occ_forest.Rmd
fix code converting to SpatialPointsDataframe change plant_occ name fix issue with removing NAs

@@
 ```{r}
-TropicalAndes_IUCNHabitat_Forest0.5 <- aggregate(TropicalAndes_IUCNHabitat_Forest, 50, fun=mean, na.rm=TRUE)
-TropicalAndes_IUCNHabitat_Forest0.5
+TropicalAndes_IUCNHabitat_Forest0.3 <- aggregate(TropicalAndes_IUCNHabitat_Forest, 30, fun=mean, na.rm=TRUE)
+TropicalAndes_IUCNHabitat_Forest0.3
 ```
@@
 # convert to SpatialPointsDataFrame
-coordinates(TropicalAndes_GBIF_plant_occ_cleaned) <- ~ decimalLatitude + decimalLongitude
-coordinates(TropicalAndes_frugivore_occ) <- ~ decimalLatitude + decimalLongitude
-TropicalAndes_GBIF_plant_occ_cleaned
+coordinates(TropicalAndes_plant_occ) <- ~ decimalLongitude + decimalLatitude
+coordinates(TropicalAndes_frugivore_occ) <- ~ decimalLongitude + decimalLatitude
+proj4string(TropicalAndes_plant_occ) <- CRS("+proj=longlat +datum=WGS84 +no_defs")
+proj4string(TropicalAndes_frugivore_occ) <- CRS("+proj=longlat +datum=WGS84 +no_defs")
+TropicalAndes_plant_occ
 TropicalAndes_frugivore_occ
@@
 ```{r}
-TropicalAndes_GBIF_plant_occ_cleaned$ref<- extract(TropicalAndes_IUCNHabitat_Forest, TropicalAndes_GBIF_plant_occ_cleaned)
+TropicalAndes_plant_occ$ref<- extract(TropicalAndes_IUCNHabitat_Forest, TropicalAndes_plant_occ)
 TropicalAndes_frugivore_occ$ref<- extract(TropicalAndes_IUCNHabitat_Forest, TropicalAndes_frugivore_occ)
@@
 ```{r}
-TropicalAndes_GBIF_plant_occ_cleaned$ref <- as.character(TropicalAndes_GBIF_plant_occ_cleaned$ref)
+TropicalAndes_plant_occ$ref <- as.character(TropicalAndes_plant_occ$ref)
 TropicalAndes_frugivore_occ$ref <- as.character(TropicalAndes_frugivore_occ$ref)
@@
 ```{r}
-plant_occ_forest <- TropicalAndes_GBIF_plant_occ_cleaned@data
+plant_occ_forest <- TropicalAndes_plant_occ@data
 frugivore_occ_forest <- TropicalAndes_frugivore_occ@data
@@
 #remove row if ref = NA
-plant_occ_forest <- plant_occ_forest[plant_occ_forest$ref != "NA", ]
-frugivore_occ_forest <- frugivore_occ_forest[frugivore_occ_forest$ref != "NA", ]
+plant_occ_forest <- plant_occ_forest[!is.na(plant_occ_forest$ref), ]
+frugivore_occ_forest <- frugivore_occ_forest[!is.na(frugivore_occ_forest$ref), ]
 ```

### 2023-08-03 code/L1/L1_TropicalAndes_subset_occ_forest.Rmd
add back lat and long to dataframe update export csv names

@@
 TropicalAndes_IUCNHabitat_Forest <- raster(file.path(data_path2, "TropicalAndes_IUCNHabitat_Forest.tif"))
-#TropicalAndes_plant_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
+TropicalAndes_plant_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
 TropicalAndes_frugivore_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_frugivore_occ_cleaned.csv"))
@@
 frugivore_occ_forest <- TropicalAndes_frugivore_occ@data
+```
+
+```{r}
+# add back lat and long
+plant_coords <- TropicalAndes_plant_occ@coords
+frug_coords <- TropicalAndes_frugivore_occ@coords
+plant_coords <- as.data.frame(plant_coords)
+frug_coords <- as.data.frame(frug_coords)
+plant_occ_forest$decimalLatitude <- plant_coords$decimalLatitude
+plant_occ_forest$decimalLongitude <- plant_coords$decimalLongitude
+frugivore_occ_forest$decimalLatitude <- frug_coords$decimalLatitude
+frugivore_occ_forest$decimalLongitude <- frug_coords$decimalLongitude
 ```
@@
 write.csv(frugivore_occ_forest, file = file.path(output_path,"TropicalAndes_frugivore_occ_forest.csv"))
-write.csv(TropicalAndes_plant_SpeciesList, file = file.path(output_path,"TropicalAndes_plant_Forest_SpeciesList.csv"))
-write.csv(TropicalAndes_frugivore_Forest_SpeciesList, file = file.path(output_path,"TropicalAndes_frugivore_Forest_SpeciesList"))
+write.csv(TropicalAndes_plant_Forest_SpeciesList, file = file.path(output_path,"TropicalAndes_plant_Forest_SpeciesList.csv"))
+write.csv(TropicalAndes_frugivore_Forest_SpeciesList, file = file.path(output_path,"TropicalAndes_frugivore_Forest_SpeciesList.csv"))
 ```

### 2023-08-03 code/L2/L2_2_TropicalAndes_taxdiv.Rmd
update input files add code to make grid using forest habitat shape

@@
 data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+data_path2<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
 output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
@@
 ```{r}
-TropicalAndes_plant_occ <- read.csv(file.path(data_path,"TropicalAndes_plant_occ.csv"))
-TropicalAndes_frugivore_occ <- read.csv(file.path(data_path,"TropicalAndes_frugivore_occ.csv"))
+TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_plant_occ_forest.csv"))
+TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_frugivore_occ_forest.csv"))
+TropicalAndes_IUCNHabitat_Forest <- raster(file.path(data_path2, "TropicalAndes_IUCNHabitat_Forest.tif"))
 ```
@@
 ```{r}
-plants.sf <- st_as_sf(TropicalAndes_plant_occ, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
+plants.sf <- st_as_sf(TropicalAndes_plant_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
 
-frugivores.sf <- st_as_sf(TropicalAndes_frugivore_occ, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
+frugivores.sf <- st_as_sf(TropicalAndes_frugivore_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
 ```
@@
 geom_sf(data = TApoly) +
+#geom_raster(data = TropicalAndes_IUCNHabitat_Forest, aes(x = x, y = y, fill = fct_elevation_2), fill = "forestgreen") +
 labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
@@
   geom_sf(data = frugivores_sf_species, pch = 21) +
-  labs(title = "Tropical Andes Bird and Mammal Occurrences", x = "Latitude", y = "Longitude") +
+  labs(title = "Tropical Andes Frugivore Occurrences", x = "Latitude", y = "Longitude") +
   coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
@@
 
-# Make 0.5 Grid
+# Mask grid by forest
+```{r}
+TropicalAndes_IUCNHabitat_Forest <- aggregate(TropicalAndes_IUCNHabitat_Forest, 50, fun=mean, na.rm=TRUE)
+TropicalAndes_IUCNHabitat_Forest
+```
+
+```{r}
+Forest_poly <- rasterToPolygons(TropicalAndes_IUCNHabitat_Forest, na.rm = TRUE, dissolve = TRUE)
+Forest_poly
+plot(Forest_poly)
+```
+
+```{r}
+Forest_sf <- st_as_sf(Forest_poly)
+Forest_sf
+plot(Forest_sf)
+```
+
+# Make 0.75 Grid
 ```{r}
 # make grid
-TAGrid0.5 <- TApoly %>%
-  st_make_grid(cellsize = c(0.5)) %>%
-  st_intersection(TApoly) %>%
+TAGrid0.75 <- TApoly %>%
+  st_make_grid(cellsize = c(0.75)) %>%
+  st_intersection(Forest_sf) %>%
   st_cast("MULTIPOLYGON") %>%
@@
   geom_sf(data = worldMap, fill = "white")+
-  geom_sf(data = TApoly) +
-  geom_sf(data = TAGrid0.5) +
-  labs(title = "Tropical Andes Grid [0.5°]", x = "Latitude", y = "Longitude") +
+  geom_sf(data = Forest_sf) +
+  geom_sf(data = TAGrid0.75) +
+  labs(title = "Tropical Andes Grid [0.75°]", x = "Latitude", y = "Longitude") +
   coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
@@
 ```{r}
-plant_richness_grid0.5 <- TAGrid0.5 %>%
+plant_richness_grid0.75 <- TAGrid0.75 %>%
   st_join(plants_sf_species) %>%
@@
 ```{r}
-frugivore_richness_grid0.5 <- TAGrid0.5 %>%
+frugivore_richness_grid0.75 <- TAGrid0.75 %>%
   st_join(frugivores_sf_species) %>%
@@
 ```{r}
-plantgridRichTA0.5 <-
-  ggplot(plant_richness_grid0.5) +
+plantgridRichTA0.75 <-
+  ggplot(plant_richness_grid0.75) +
   geom_sf(data = worldMap, fill = "white")+
@@
   pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Plant Richness [0.5°]", x = "Latitude", y = "Longitude", fill = "Number of species")
-plantgridRichTA0.5
+  labs(title = "Tropical Andes Plant Richness", x = "Latitude", y = "Longitude", fill = "Number of species")
+plantgridRichTA0.75
 ```
@@
 ```{r}
-frugivoregridRichTA0.5 <-
-  ggplot(frugivore_richness_grid0.5) +
+frugivoregridRichTA0.75 <-
+  ggplot(frugivore_richness_grid0.75) +
   geom_sf(data = worldMap, fill = "white")+
@@
   pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Frugivore Richness [0.5°]", x = "Latitude", y = "Longitude", fill = "Number of species")
-frugivoregridRichTA0.5
+  labs(title = "Tropical Andes Frugivore Richness", x = "Latitude", y = "Longitude", fill = "Number of species")
+frugivoregridRichTA0.75
 ```
@@
 ```{r}
-plant_cellRichness0.5 <- plantgridRichTA0.5[["data"]]
-frugivore_cellRichness0.5 <- frugivoregridRichTA0.5[["data"]]
+plant_cellRichness0.75 <- plantgridRichTA0.75[["data"]]
+frugivore_cellRichness0.75 <- frugivoregridRichTA0.75[["data"]]
 ```
@@
 ```{r}
-plot(plant_cellRichness0.5$num_species, frugivore_cellRichness0.5$num_species,
-     main = "Plant Richness vs Bird & Mammal Richness by cell  [0.5°]",
-     xlab = "Plant Richness by cell", ylab = "Bird & Mammal Richness by cell")
-abline(lm(frugivore_cellRichness0.5$num_species ~ plant_cellRichness0.5$num_species))
-summary(lm(frugivore_cellRichness0.5$num_species ~ plant_cellRichness0.5$num_species))
+plot(plant_cellRichness0.75$num_species, frugivore_cellRichness0.75$num_species,
+     main = "Plant Richness vs Frugivore Richness by cell",
+     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
+abline(lm(frugivore_cellRichness0.75$num_species ~ plant_cellRichness0.75$num_species))
+summary(lm(frugivore_cellRichness0.75$num_species ~ plant_cellRichness0.75$num_species))
 ```
@@
 ```{r}
-hist(plant_cellRichness0.5$num_species,
-     main = "Histogram of Plant Richness by cell  [0.5°]",
+hist(plant_cellRichness0.75$num_species,
+     main = "Histogram of Plant Richness by cell",
      xlab = "Plant Richness by cell")
-hist(frugivore_cellRichness0.5$num_species,
-     main = "Histogram of Frugivore Richness by cell  [0.5°]",
+hist(frugivore_cellRichness0.75$num_species,
+     main = "Histogram of Frugivore Richness by cell",
      xlab = "Frugivore Richness by cell")
@@
 ```{r}
-summary(plant_cellRichness0.5$num_species)
-summary(frugivore_cellRichness0.5$num_species)
+summary(plant_cellRichness0.75$num_species)
+summary(frugivore_cellRichness0.75$num_species)
 ```
@@
 ```{r}
-write.csv(plant_cellRichness0.5, file.path(output_path,"TropicalAndes_GBIF_plant_richness.csv"))
-write.csv(frugivore_cellRichness0.5, file.path(output_path,"TropicalAndes_frugivore_richness.csv"))
+write.csv(plant_cellRichness0.75, file.path(output_path,"TropicalAndes_GBIF_plant_richness.csv"))
+write.csv(frugivore_cellRichness0.75, file.path(output_path,"TropicalAndes_frugivore_richness.csv"))
 ```

### 2023-08-03 code/L0/L0_TropicalAndes_plants_BIEN_traits.Rmd
update species list code
add code to summarize trait count
update for species list remove old code

--- 2023-07-18 code/L0/L0_TropicalAndes_plants_BIEN_traits.Rmd
@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: ""
-data input: ""
-data output: ""
+overview: "This script retrieves plant trait data from the BIEN database for plant species list."
+data input: "none"
+data output: "TropicalAndes_BIEN_traits.csv"
 date: "2023-07-18"
@@
 
-# List of countries of interest
 ```{r}
-country_vector <- c("Ecuador", "Venezuela", "Peru", "Colombia", "Bolivia")
+#filter by species list
+TropicalAndes_plant_Forest_SpeciesList <- read.csv("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1/TropicalAndes_plant_Forest_SpeciesList.csv")
+species_list <- TropicalAndes_plant_Forest_SpeciesList$x
 ```
@@
 ```{r}
-TropicalAndes_BIEN_traits <- BIEN_trait_country(country_vector, all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
+trait_list <- c("whole plant dispersal syndrome", "whole plant height", "maximum whole plant longevity")
+TropicalAndes_BIEN_plant_traits_forest <- BIEN_trait_traitbyspecies(species_list, trait_list, all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
+```
+
+```{r}
+dispersal_mean <- BIEN_trait_mean(species_list, "whole plant dispersal syndrome")
+height_mean <- BIEN_trait_mean(species_list, "whole plant height")
+longevity_mean <- BIEN_trait_mean(species_list, "maximum whole plant longevity")
 ```
@@
 ```{r}
-glimpse(TropicalAndes_BIEN_traits)
+glimpse(TropicalAndes_BIEN_plant_traits_forest)
 ```
@@
 ```{r}
+print("The number of records is")
+nrow(TropicalAndes_BIEN_plant_traits_forest)
+print("The number of species is")
+length(unique(TropicalAndes_BIEN_plant_traits_forest$scrubbed_species_binomial))
+print("The number of families is")
+length(unique(TropicalAndes_BIEN_plant_traits_forest$family_matched))
+```
 
+```{r}
+TropicalAndes_BIEN_plant_traits_forest %>% count(trait_name)
 ```
@@
 ```{r}
-TropicalAndes_BIEN_traits_path <- file.path(output_path,"TropicalAndes_BIEN_traits.csv") 
-write.csv(TropicalAndes_BIEN_traits, file = TropicalAndes_BIEN_traits_path)
+TropicalAndes_BIEN_traits_path <- file.path(output_path,"TropicalAndes_BIEN_plant_traits_forest.csv") 
+write.csv(TropicalAndes_BIEN_plant_traits_forest, file = TropicalAndes_BIEN_traits_path)
 ```

### 2023-08-03 code/L2/L2_1_TropicalAndes_plant_traits.Rmd
inital commit

---
title: "Tropical Andes plant traits"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Combinging TRY and BIEN plant trait data"
data input: ""
data output: ""
date: "2023-08-03"
output: html_document
---

# Load required packages
```{r}
library(dplyr)
library(tidyr)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Read in Data
```{r}
BIEN_traits <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_traits_forest.csv"))
TRY_traits <- read.csv(file.path(data_path,"TropicalAndes_TRY_plant_traits_forestSpecies.csv"))
```

# Subset to columns of interest
```{r}
# remove TRY columns without value in TraitID
TRY_traits_2 <- TRY_traits[!is.na(TRY_traits$TraitID), ]
```

```{r}
BIEN_traits_sub <- BIEN_traits[ , c("scrubbed_species_binomial", "trait_name", "trait_value", "unit")]
TRY_traits_sub <- TRY_traits_2[ , c("AccSpeciesName", "TraitName", "OrigValueStr", "OrigUnitStr")]
```

# Standardize column names
```{r}
BIEN_traits_stan <- BIEN_traits_sub %>% rename(species = scrubbed_species_binomial)
TRY_traits_stan <- TRY_traits_sub %>% rename(species = AccSpeciesName, trait_name = TraitName, trait_value = OrigValueStr, unit = OrigUnitStr)
```

```{r}
plant_traits <- dplyr::bind_rows(BIEN_traits_stan, TRY_traits_stan, .id = 'source')
```

```{r}
#replace source 1 or 2 with BIEN or TRY
plant_traits['source'][plant_traits['source'] == 1] <- "BIEN"
plant_traits['source'][plant_traits['source'] == 2] <- "TRY"
```

# Summary
```{r}
glimpse(plant_traits)
```

```{r}
plant_traits %>%
  count(trait_name)
```

# Dispersal Traits
```{r}
# subset plant traits to just dispersal traits
plant_traits_dispersal <- plant_traits[plant_traits$trait_name == "Dispersal syndrome" | plant_traits$trait_name == "whole plant dispersal syndrome",]
# look at possible trait values
plant_traits_dispersal_values <- unique(plant_traits_dispersal$trait_value)
# save list to manually classify biotic, abiotic, none, and NA
#write.csv(plant_traits_dispersal_values, file.path(data_path,"plant_trait_dispersal_values.csv"))
# read in classified table
dispersal_classified <- read.csv(file.path(data_path,"plant_trait_dispersal_values_classified.csv"))
# reclassify dispersal trait value to table
plant_traits_dispersal$trait_value <- 
dispersal_classified$Classification[match(plant_traits_dispersal$trait_value,dispersal_classified$OrigValue )]
# check to make sure it worked
unique(plant_traits_dispersal$trait_value)
```

```{r}
#need to remove none & NA
plant_traits_dispersal <- plant_traits_dispersal[plant_traits_dispersal$trait_value == "biotic" | plant_traits_dispersal$trait_value == "abiotic",]
#need one value per species

plant_traits_dispersal_wide <- pivot_wider(data = plant_traits_dispersal, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = list)
```

# Plant fruit mass traits
```{r}
plant_traits_mass <- plant_traits[plant_traits$trait_name == "Fruit dry mass",]
plant_traits_mass <- plant_traits_mass[plant_traits_mass$unit == "mg",]
plant_traits_mass$trait_value <- as.numeric(plant_traits_mass$trait_value)
plant_traits_mass_wide <- pivot_wider(data = plant_traits_mass, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = mean)
# ready to combine
```

# Plant height traits
```{r}
plant_traits_height <- plant_traits[plant_traits$trait_name == "Plant height vegetative" |plant_traits$trait_name == "whole plant height" ,]
plant_traits_height$trait_value <- as.numeric(plant_traits_height$trait_value)
plant_traits_height_wide <- pivot_wider(data = plant_traits_height, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = mean)
plant_traits_height_wide$plant_height <- rowMeans(plant_traits_height_wide[,2:3], na.rm = TRUE)
# ready to combine
```

# Plant longevity traits
```{r}
plant_traits_longevity <- plant_traits[plant_traits$trait_name == "Plant lifespan (longevity)" | plant_traits$trait_name == "maximum whole plant longevity",]
#trait values are a mix of numeric and characters
```

# Combine all traits
```{r}
# combine all trait wide tables
plant_traits_wide <- bind_cols(plant_traits_height_wide, plant_traits_mass_wide)
# remove species with NAs for one or more traits

```

# Write data to csv
```{r}
write.csv(plant_traits, file.path(output_path,"TropicalAndes_plant_traits_forest.csv"))
```

### 2023-08-03 code/L2/L2_2_TropicalAndes_taxdiv.Rmd
crop to remove galapagos islands fix lat and long errors update figure names

@@
 geom_sf(data = TApoly) +
-#geom_raster(data = TropicalAndes_IUCNHabitat_Forest, aes(x = x, y = y, fill = fct_elevation_2), fill = "forestgreen") +
 labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
-coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
+coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
 annotation_scale(location = "bl", width_hint = 0.5) +
@@
   labs(title = "Tropical Andes Plant Occurrences", x = "Latitude", y = "Longitude") +
-  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
   theme(
@@
   labs(title = "Tropical Andes Frugivore Occurrences", x = "Latitude", y = "Longitude") +
-  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
   theme(
@@
   labs(title = "Tropical Andes Grid [0.75°]", x = "Latitude", y = "Longitude") +
-  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
   theme(
@@
   scale_fill_scico(palette = "romaO", direction = 1, end = 0.9, midpoint = 1500) +
-  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
   theme(
@@
   annotation_north_arrow(location = "bl", which_north = "true", 
-  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Plant Richness", x = "Latitude", y = "Longitude", fill = "Number of species")
+  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  labs(title = "Tropical Andes Plant Raw Richness", x = "Longitude", y = "Latitude", fill = "Number of species")
 plantgridRichTA0.75
@@
   geom_sf(aes(fill = num_species), color = NA) +
-  scale_fill_scico(palette = "romaO", direction = 1, end = 0.9, midpoint = 1500) +
-  coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
+  scale_fill_scico(palette = "romaO", direction = 1, end = 0.9, midpoint = 225) +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
   theme(
@@
   annotation_north_arrow(location = "bl", which_north = "true", 
-  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Frugivore Richness", x = "Latitude", y = "Longitude", fill = "Number of species")
+  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  labs(title = "Tropical Andes Frugivore Raw Richness", x = "Longitude", y = "Latitude", fill = "Number of species")
 frugivoregridRichTA0.75
@@
 plot(plant_cellRichness0.75$num_species, frugivore_cellRichness0.75$num_species,
-     main = "Plant Richness vs Frugivore Richness by cell",
+     main = "Plant Richness vs Frugivore Richness",
      xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")

### 2023-08-04 code/L0/L0_TropicalAndes_plants_GIFT_traits.Rmd
inital commit

---
title: "Tropical Andes plant trait data GIFT"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script retrieves plant trait data from the GIFT database for plant species list."
data input: ""
data output: ""
date: "2023-08-03"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(GIFT)
library(dplyr)
```

# Load in species list
```{r}
species_list <- read.csv("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1/TropicalAndes_plant_Forest_SpeciesList.csv")
species_list <- species_list$x
```

# Retrive GIFT data
```{r}
trait_meta <- GIFT_traits_meta()
```
Traits of interest: plant height (ID = 1.6.3), lifespan (ID = 2.2.1), dispersal syndrome (ID = 3.3)

```{r}
# 
height <- GIFT_traits(trait_IDs = c("1.6.2"))
lifespan <- GIFT_traits(trait_IDs = c("2.2.1"))
dispersal_syndrome <- GIFT_traits(trait_IDs = c("3.3.1", "3.3.2"))
```

```{r}
# filter by species list
height_sub <- height %>%
  filter(work_species %in% species_list)
dispersal_syndrome_sub <- dispersal_syndrome %>%
  filter(work_species %in% species_list)
lifespan_sub <- lifespan %>%
  filter(work_species %in% species_list)
```

```{r}
#rename columns to trait name
height_sub <- height_sub %>% rename(trait_value = trait_value_1.6.2, species = work_species)
dispersal_syndrome_sub <- dispersal_syndrome_sub %>% rename(trait_value = trait_value_3.3.1, species = work_species)
lifespan_sub <- lifespan_sub %>% rename(trait_value = trait_value_2.2.1, species = work_species)
```

```{r}
#keep just species and trait columns
height_sub <- height_sub %>% select(species, trait_value)
dispersal_syndrome_sub <- dispersal_syndrome_sub %>% select(species, trait_value)
lifespan_sub <- lifespan_sub %>% select(species, trait_value)
```

```{r}
# add trait name column
height_sub$trait_name <- "plant_height"
dispersal_syndrome_sub$trait_name <- "dispersal_syndrome"
lifespan_sub$trait_name <- "plant_lifespan"
```

```{r}
#convert numeric trait values to character
height_sub$trait_value <- as.character(height_sub$trait_value)
lifespan_sub$trait_value <- as.character(lifespan_sub$trait_value)
```

```{r}
#combine all trait dataframes into one long table
GIFT_traits <- dplyr::bind_rows(height_sub, dispersal_syndrome_sub, lifespan_sub)
```

# Summary
```{r}
glimpse(GIFT_traits)
```

```{r}
print("The number of records is")
nrow(GIFT_traits)
print("The number of species is")
length(unique(GIFT_traits$species))
```

# Write data to csv
```{r}
write.csv(GIFT_traits, file = file.path(output_path,"TropicalAndes_GIFT_plant_traits_forest.csv"))
```

### 2023-08-04 code/L2/L2_1_TropicalAndes_plant_traits.Rmd

@@
 library(tidyr)
+library(data.table)
 ```
@@
 # save list to manually classify biotic, abiotic, none, and NA
-#write.csv(plant_traits_dispersal_values, file.path(data_path,"plant_trait_dispersal_values.csv"))
+write.csv(plant_traits_dispersal_values, file.path(data_path,"plant_trait_dispersal_values.csv"))
 # read in classified table
@@
 ```{r}
+# change all trait_name to dispersal_syndrome
+plant_traits_dispersal['trait_name'][plant_traits_dispersal['trait_name'] == 'Dispersal syndrome'] <- 'dispersal_syndrome'
+plant_traits_dispersal['trait_name'][plant_traits_dispersal['trait_name'] == 'whole plant dispersal syndrome'] <- 'dispersal_syndrome'
 #need to remove none & NA
@@
 #need one value per species
-
+plant_traits_dispersal <- plant_traits_dispersal %>% distinct(species, .keep_all = TRUE)
 plant_traits_dispersal_wide <- pivot_wider(data = plant_traits_dispersal, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = list)
+#ready to combine
 ```
@@
 plant_traits_longevity <- plant_traits[plant_traits$trait_name == "Plant lifespan (longevity)" | plant_traits$trait_name == "maximum whole plant longevity",]
-#trait values are a mix of numeric and characters
+# change all trait_name to plant lifespan
+plant_traits_longevity['trait_name'][plant_traits_longevity['trait_name'] == 'Plant lifespan (longevity)'] <- 'plant_lifespan'
+plant_traits_longevity['trait_name'][plant_traits_longevity['trait_name'] == 'maximum whole plant longevity'] <- 'plant_lifespan'
+
+#remove non-numeric values
+plant_traits_longevity <- plant_traits_longevity[!is.na(as.numeric(as.character(plant_traits_longevity$trait_value))),]
+```
+
+```{r}
+plant_traits_longevity$trait_value <- as.numeric(plant_traits_longevity$trait_value)
+plant_traits_longevity_wide <- pivot_wider(data = plant_traits_longevity, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = mean)
 ```
@@
 ```{r}
-# combine all trait wide tables
-plant_traits_wide <- bind_cols(plant_traits_height_wide, plant_traits_mass_wide)
+# convert all to dataframes
+plant_traits_height_wide <- as.data.frame(plant_traits_height_wide)
+plant_traits_dispersal_wide <- as.data.frame(plant_traits_dispersal_wide)
+plant_traits_mass_wide <- as.data.frame(plant_traits_mass_wide)
+plant_traits_longevity_wide <- as.data.frame(plant_traits_longevity_wide)
+
+plant_traits_wide <- merge(x = plant_traits_height_wide, y = plant_traits_dispersal_wide, by = "species", all = TRUE)
+plant_traits_wide <- merge(x = plant_traits_wide, y = plant_traits_mass_wide, by = "species", all = TRUE)
+plant_traits_wide <- merge(x = plant_traits_wide, y = plant_traits_longevity_wide, by = "species", all = TRUE)
+```
+
+```{r}
+plant_traits_wide <- plant_traits_wide [, c("species", "plant_height", "dispersal_syndrome", "Fruit dry mass", "plant_lifespan")]
 # remove species with NAs for one or more traits
+plant_traits_wide_complete <- na.omit(plant_traits_wide)
+```
 
+#Summary
+```{r}
+glimpse(plant_traits_wide_complete)
 ```

### 2023-08-04 code/L2/L2_1_TropicalAndes_plant_traits.Rmd
add GIFT trait data

@@
 TRY_traits <- read.csv(file.path(data_path,"TropicalAndes_TRY_plant_traits_forestSpecies.csv"))
+GIFT_traits <- read.csv(file.path(data_path,"TropicalAndes_GIFT_plant_traits_forest.csv"))
 ```
@@
 ```{r}
-plant_traits <- dplyr::bind_rows(BIEN_traits_stan, TRY_traits_stan, .id = 'source')
+plant_traits <- dplyr::bind_rows(BIEN_traits_stan, TRY_traits_stan, GIFT_traits, .id = 'source')
 ```
@@
 plant_traits['source'][plant_traits['source'] == 2] <- "TRY"
+plant_traits['source'][plant_traits['source'] == 3] <- "GIFT"
 ```
@@
 # subset plant traits to just dispersal traits
-plant_traits_dispersal <- plant_traits[plant_traits$trait_name == "Dispersal syndrome" | plant_traits$trait_name == "whole plant dispersal syndrome",]
+plant_traits_dispersal <- plant_traits[plant_traits$trait_name == "Dispersal syndrome" | plant_traits$trait_name == "whole plant dispersal syndrome" | plant_traits$trait_name == "dispersal_syndrome",]
 # look at possible trait values
@@
 write.csv(plant_traits_dispersal_values, file.path(data_path,"plant_trait_dispersal_values.csv"))
+plant_traits_dispersal %>%
+  count(trait_value)
+```
+
+```{r}
 # read in classified table
@@
 ```{r}
-plant_traits_height <- plant_traits[plant_traits$trait_name == "Plant height vegetative" |plant_traits$trait_name == "whole plant height" ,]
+plant_traits_height <- plant_traits[plant_traits$trait_name == "Plant height vegetative" |plant_traits$trait_name == "whole plant height" |plant_traits$trait_name == "plant_height" ,]
 plant_traits_height$trait_value <- as.numeric(plant_traits_height$trait_value)
@@
 ```{r}
-plant_traits_longevity <- plant_traits[plant_traits$trait_name == "Plant lifespan (longevity)" | plant_traits$trait_name == "maximum whole plant longevity",]
+plant_traits_longevity <- plant_traits[plant_traits$trait_name == "Plant lifespan (longevity)" | plant_traits$trait_name == "maximum whole plant longevity" | plant_traits$trait_name == "plant_lifespan",]
 # change all trait_name to plant lifespan
@@
 ```{r}
-plant_traits_wide <- plant_traits_wide [, c("species", "plant_height", "dispersal_syndrome", "Fruit dry mass", "plant_lifespan")]
+plant_traits_wide <- plant_traits_wide [, c("species", "plant_height", "dispersal_syndrome", "plant_lifespan")]
 # remove species with NAs for one or more traits

### 2023-08-04 code/L2/L2_2_TropicalAndes_funcdiv.Rmd
inital commit

---
title: "Tropical Andes Functional diversity for plants and Frugivores"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: ""
data output: ""
date: "2023-08-03"
output: html_document
---

# Load required packages
```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(smoothr)
library(purrr)
library(raster)
library(scico)
library(ggspatial)
library(letsR)
library(mFD)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
data_path2<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
```

# Read in Data
```{r}
TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_plant_occ_forest.csv"))
TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_frugivore_occ_forest.csv"))
TropicalAndes_IUCNHabitat_Forest <- raster(file.path(data_path2, "TropicalAndes_IUCNHabitat_Forest.tif"))
frugivore_traits <- read.csv(file.path(data_path,"TropicalAndes_Frugivoria_traits_Forest.csv"))
plant_traits <- 
```

```{r}
lets.presab.points
```

### 2023-08-04 code/L2/L2_2_TropicalAndes_funcdiv.Rmd
add PAM code

@@
 frugivore_traits <- read.csv(file.path(data_path,"TropicalAndes_Frugivoria_traits_Forest.csv"))
-plant_traits <- 
+plant_traits <- file.path(data_path,"TropicalAndes_plant_traits_forest.csv")
+```
+
+# Remove species with NAs for one or more traits
+```{r}
+frugivore_traits <- na.omit(frugivore_traits)
+```
+
+# Forest shape conversion
+```{r}
+TropicalAndes_IUCNHabitat_Forest <- aggregate(TropicalAndes_IUCNHabitat_Forest, 50, fun=mean, na.rm=TRUE)
+TropicalAndes_IUCNHabitat_Forest
 ```
@@
 ```{r}
-lets.presab.points
+Forest_poly <- rasterToPolygons(TropicalAndes_IUCNHabitat_Forest, na.rm = TRUE, dissolve = TRUE)
+Forest_poly
+plot(Forest_poly)
 ```
+
+```{r}
+Forest_sf <- st_as_sf(Forest_poly)
+Forest_sf
+plot(Forest_sf)
+st_write(Forest_sf, file.path(data_path, "Forest_sf.shp"))
+```
+
+# Create Presence-Absense matrix
+```{r}
+x <- TropicalAndes_plant_occ_forest$decimalLongitude
+y <- TropicalAndes_plant_occ_forest$decimalLatitude
+plant_xy <- cbind(x, y)
+plant_species <- TropicalAndes_plant_occ_forest$species
+plants_PAM <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14)
+summary(plants_PAM)
+plot(plants_PAM, xlab = "Longitude", ylab = "Latitude", main = "Plant richness map")
+```
+
+```{r}
+x <- TropicalAndes_frugivore_occ_forest$decimalLongitude
+y <- TropicalAndes_frugivore_occ_forest$decimalLatitude
+frugivore_xy <- cbind(x, y)
+frugivore_species <- TropicalAndes_frugivore_occ_forest$species
+frugivore_PAM <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14)
+summary(frugivore_PAM)
+plot(frugivore_PAM, xlab = "Longitude", ylab = "Latitude", main = "Frugivore richness map")
+```
+
+# Create trait type table
+```{r}
+frug_trait_name <- c("diet_cat", "body_mass", "body_size", "generation_time")
+frug_trait_type <- c("N", "Q", "Q", "Q")
+frug_trait_cat <- as.data.frame(cbind(frug_trait_name, frug_trait_type))
+plant_trait_name <- c("dispersal_syndrome", "plant_height", "plant_lifespan")
+frug_trait_type <- c("N", "Q", "Q") 
+```

### 2023-08-04 code/L2/L2_1_TropicalAndes_plant_traits.Rmd
remove unnecessary comments extract species list

@@
 plant_traits_dispersal_wide <- pivot_wider(data = plant_traits_dispersal, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = list)
-#ready to combine
 ```
@@
 plant_traits_mass_wide <- pivot_wider(data = plant_traits_mass, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = mean)
-# ready to combine
 ```
@@
 plant_traits_height_wide$plant_height <- rowMeans(plant_traits_height_wide[,2:3], na.rm = TRUE)
-# ready to combine
 ```
@@
 
+# Extract species list
+```{r}
+species_list <- unique(plant_traits_wide_complete$species)
+```
+
 # Write data to csv
@@
 write.csv(plant_traits, file.path(output_path,"TropicalAndes_plant_traits_forest.csv"))
+write.csv(species_list, file.path(output_path,"TropicalAndes_plant_traits_forest_species_list.csv"))
 ```

### 2023-08-04 code/L1/L1_TropicalAndes_subset_occ_forest_taxResolution.Rmd
inital commit

---
title: "Tropical Andes plant species list"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script retrives updated names for GBIF species list."
data input: "none"
data output: ""
date: "2023-08-03"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(TNRS)
devtools::install_github("ecoinfor/U.Taxonstand")
library(U.Taxonstand)
library(tidyr)
library(dplyr)
```

# Load species list
```{r}
#species list
TropicalAndes_plant_Forest_SpeciesList <- read.csv("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1/TropicalAndes_plant_Forest_SpeciesList.csv")
TropicalAndes_plant_Forest_SpeciesList$NAME <- TropicalAndes_plant_Forest_SpeciesList$x
species_list <- TropicalAndes_plant_Forest_SpeciesList$NAME
```

# Use TRNS package
```{r}
TNRS_results <- TNRS(TropicalAndes_plant_Forest_SpeciesList)
```

```{r}
write.csv(TNRS_results, file.path(output_path,"TropicalAndes_plant_Forest_SpeciesList_TNRS_results.csv"))
```

# Using U.Taxonstand package
```{r}
U.Taxonstand_results <- nameMatch(spList = species_list, spSource = TropicalAndes_plant_Forest_SpeciesList, author = FALSE, max.distance = 1)
```

```{r}
write.csv(U.Taxonstand_results, file.path(output_path,"TropicalAndes_plant_Forest_SpeciesList_U.Taxonstand_results.csv"))
```

### 2023-08-04 code/L2/L2_2_TropicalAndes_taxdiv.Rmd
add filtering occ records by trait species lists

@@
 
+# Filter plant occ by plant trait species list
+```{r}
+plant_trait_species_list <- read.csv(file.path(data_path,"TropicalAndes_plant_traits_forest_species_list.csv"))
+TropicalAndes_plant_occ_forest_sub <- TropicalAndes_plant_occ_forest %>% filter(species %in% plant_trait_species_list$x)
+write.csv(TropicalAndes_plant_occ_forest_sub, file.path(data_path,"TropicalAndes_plant_occ_forest_sub.csv"))
+```
+
+# Filter frugivore occ by frugivore trait species list
+```{r}
+frugivore_trait_species_list <- read.csv(file.path(data_path,"TropicalAndes_frugivore_traits_forest_species_list.csv"))
+TropicalAndes_frugivore_occ_forest_sub <- TropicalAndes_frugivore_occ_forest %>% filter(species %in% frugivore_trait_species_list$x)
+write.csv(TropicalAndes_frugivore_occ_forest_sub, file.path(data_path,"TropicalAndes_frugivore_occ_forest_sub.csv"))
+```
+
 # Polygons of countries
@@
 ```{r}
-plants.sf <- st_as_sf(TropicalAndes_plant_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
+plants.sf <- st_as_sf(TropicalAndes_plant_occ_forest_sub, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
 
-frugivores.sf <- st_as_sf(TropicalAndes_frugivore_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
+frugivores.sf <- st_as_sf(TropicalAndes_frugivore_occ_forest_sub, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
 ```
@@
 plot(Forest_sf)
+st_write(Forest_sf, file.path(data_path, "Forest_sf.shp"))
 ```
@@
 ```{r}
-write.csv(plant_cellRichness0.75, file.path(output_path,"TropicalAndes_GBIF_plant_richness.csv"))
+write.csv(plant_cellRichness0.75, file.path(output_path,"TropicalAndes_GBIF_plant_richness_sub.csv"))
 write.csv(frugivore_cellRichness0.75, file.path(output_path,"TropicalAndes_frugivore_richness.csv"))

### 2023-08-04 code/L2/L2_2_TropicalAndes_taxdiv.Rmd
fix figure save data as new name

@@
   geom_sf(aes(fill = num_species), color = NA) +
-  scale_fill_scico(palette = "romaO", direction = 1, end = 0.9, midpoint = 1500) +
+  scale_fill_scico(palette = "romaO", direction = 1, end = 0.9, midpoint = 50) +
   coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
@@
 write.csv(plant_cellRichness0.75, file.path(output_path,"TropicalAndes_GBIF_plant_richness_sub.csv"))
-write.csv(frugivore_cellRichness0.75, file.path(output_path,"TropicalAndes_frugivore_richness.csv"))
+write.csv(frugivore_cellRichness0.75, file.path(output_path,"TropicalAndes_frugivore_richness_sub.csv"))
 ```

### 2023-08-04 code/L2/L2_2_TropicalAndes_funcdiv.Rmd
add functional diversity code

--- 2023-08-04 code/L2/L2_2_TropicalAndes_funcdiv.Rmd
@@
 ```{r}
-TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_plant_occ_forest.csv"))
-TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_frugivore_occ_forest.csv"))
+TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_plant_occ_forest_sub.csv"))
+TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_frugivore_occ_forest_sub.csv"))
 TropicalAndes_IUCNHabitat_Forest <- raster(file.path(data_path2, "TropicalAndes_IUCNHabitat_Forest.tif"))
 frugivore_traits <- read.csv(file.path(data_path,"TropicalAndes_Frugivoria_traits_Forest.csv"))
-plant_traits <- 
+plant_traits <- read.csv(file.path(data_path,"TropicalAndes_plant_traits_forest.csv"))
+```
+
+# Forest shape conversion
+```{r}
+TropicalAndes_IUCNHabitat_Forest <- aggregate(TropicalAndes_IUCNHabitat_Forest, 50, fun=mean, na.rm=TRUE)
+TropicalAndes_IUCNHabitat_Forest
 ```
@@
 ```{r}
-lets.presab.points
+Forest_poly <- rasterToPolygons(TropicalAndes_IUCNHabitat_Forest, na.rm = TRUE, dissolve = TRUE)
+Forest_poly
+plot(Forest_poly)
 ```
+
+```{r}
+Forest_sf <- st_as_sf(Forest_poly)
+Forest_sf
+plot(Forest_sf)
+st_write(Forest_sf, file.path(data_path, "Forest_sf.shp"))
+```
+
+# Create Presence-Absense matrix
+```{r}
+x <- TropicalAndes_plant_occ_forest$decimalLongitude
+y <- TropicalAndes_plant_occ_forest$decimalLatitude
+plant_xy <- cbind(x, y)
+plant_species <- TropicalAndes_plant_occ_forest$species
+plants_PAM <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14)
+summary(plants_PAM)
+plot(plants_PAM, xlab = "Longitude", ylab = "Latitude", main = "Plant richness map")
+plants_PAM_matrix <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE)
+```
+
+```{r}
+x <- TropicalAndes_frugivore_occ_forest$decimalLongitude
+y <- TropicalAndes_frugivore_occ_forest$decimalLatitude
+frugivore_xy <- cbind(x, y)
+frugivore_species <- TropicalAndes_frugivore_occ_forest$species
+frugivore_PAM <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14)
+summary(frugivore_PAM)
+plot(frugivore_PAM, xlab = "Longitude", ylab = "Latitude", main = "Frugivore richness map")
+frugivore_PAM_matrix <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE)
+```
+
+# Prepping data for functional diversity calculations
+
+```{r}
+plant_traits <- na.omit(plant_traits)
+frugivore_traits <- na.omit(frugivore_traits)
+frugivore_traits$body_mass_e <- as.numeric(frugivore_traits$body_mass_e)
+frugivore_traits$body_size_mm <- as.numeric(frugivore_traits$body_size_mm)
+plant_traits <- plant_traits[, c("species", "plant_height", "dispersal_syndrome", "plant_lifespan")]
+frugivore_traits <- frugivore_traits[, c("IUCN_species_name", "diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
+```
+
+```{r}
+coordinates_sf_plant <- st_as_sf(TropicalAndes_plant_occ_forest_sub, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(Forest_sf))
+coordinates_sf_frugivore <- st_as_sf(TropicalAndes_frugivore_occ_forest_sub, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(Forest_sf))
+```
+
+```{r}
+# Perform the spatial intersection
+intersections_plant_TA <- st_intersection(coordinates_sf_plant, Forest_sf)
+intersections_frugivore_TA <- st_intersection(coordinates_sf_frugivore, Forest_sf)
+# Convert the intersected sf object back to a dataframe
+coordinates_df_subset_plant <- as.data.frame(intersections_plant_TA)
+coordinates_df_subset_frugivore <- as.data.frame(intersections_frugivore_TA)
+```
+
+```{r}
+# Print the resulting subset of coordinates
+print(coordinates_df_subset_plant)
+print(coordinates_df_subset_frugivore)
+# Convert the point geometry to a dataframe of longitude and latitude
+subset_plant <- as.data.frame(st_coordinates(intersections_plant_TA))
+subset_frugivore <- as.data.frame(st_coordinates(intersections_frugivore_TA))
+
+# Match the coordinates to the original dataframe
+#merge the subset and full dataframe together to get final TA dataset
+plant_PAM_filter <- merge(plants_PAM_matrix, subset_plant, by.x = c("Longitude(x)", "Latitude(y)"), by.y = c("X", "Y"))
+frugivore_PAM_filter <- merge(frugivore_PAM_matrix, subset_frugivore, by.x = c("Longitude(x)", "Latitude(y)"), by.y = c("X", "Y"))
+```
+
+```{r}
+# Turn PAM into matrix
+plant_PAM <- as.matrix(plant_PAM_filter)
+frugivore_PAM <- as.matrix(frugivore_PAM_filter)
+
+# Add rownames
+# Get the number of rows in the matrix
+num_rows_plant <- nrow(plant_PAM)
+num_rows_frugivore <- nrow(frugivore_PAM)
+# Generate unique row names
+row_names_plant <- paste0("cell", 1:num_rows_plant)
+row_names_frugivore <- paste0("cell", 1:num_rows_frugivore)
+# Assign row names to the matrix
+rownames(plant_PAM) <- row_names_plant
+rownames(frugivore_PAM) <- row_names_frugivore
+# Print the matrix with row names
+print(plant_PAM)
+column_names_plant <- colnames(plant_PAM)
+clean_column_names_plant <- gsub("_", " ", column_names_plant)
+print(frugivore_PAM)
+column_names_frugivore <- colnames(frugivore_PAM)
+clean_column_names_frugivore <- gsub("_", " ", column_names_frugivore)
+#Insert clean column names
+colnames(plant_PAM) <- clean_column_names_plant
+colnames(frugivore_PAM) <- clean_column_names_frugivore
+```
+
+```{r}
+#turn data into correct data types for inputs into the trait categories dataframe
+frugivore_traits$diet_cat <- as.factor(frugivore_traits$diet_cat)
+frugivore_traits$body_mass_e <- as.numeric(frugivore_traits$body_mass_e)
+frugivore_traits$body_size_mm <- as.numeric(frugivore_traits$body_size_mm)
+frugivore_traits$generation_time <- as.numeric(frugivore_traits$generation_time)
+
+plant_traits$plant_height <- as.numeric(plant_traits$plant_height)
+plant_traits$plant_lifespan<- as.numeric(plant_traits$plant_lifespan)
+plant_traits$dispersal_syndrome <- as.factor(plant_traits$dispersal_syndrome)
+
+# Remove the species from PAM that have no occurrences anymore after subsetting to TA
+# Remove columns with sum equal to zero
+PAM_plant_site_final <- plant_PAM[, colSums(plant_PAM) != 0]
+PAM_frugivore_site_final <- frugivore_PAM[, colSums(frugivore_PAM) != 0]
+#Save coordinates for later
+site_loc_key_plant <- PAM_plant_site_final[,1:2]
+site_loc_key_frugivore <- PAM_frugivore_site_final[,1:2]
+
+columns_to_remove <- c(1,2)
+PAM_plant_site_final <- PAM_plant_site_final[,-columns_to_remove]
+PAM_frugivore_site_final <- PAM_frugivore_site_final[,-columns_to_remove]
+colnames_plant <- colnames(PAM_plant_site_final)
+colnames_frugivore <- colnames(PAM_frugivore_site_final)
+```
+
+```{r}
+# Remove species names from trait matrix not in the PAM
+plant_traits_df_subset <- plant_traits %>% filter(species %in% colnames_plant)
+frugivore_traits_df_subset <- frugivore_traits %>% filter(IUCN_species_name %in% colnames_frugivore)
+
+# Turn trait dataframe into a matrix
+class(plant_traits)
+class(frugivore_traits)
+plant_traits_matrix <- as.matrix(plant_traits_df_subset)
+frugivore_traits_matrix <- as.matrix(frugivore_traits_df_subset)
+
+# Define row names as species names
+row_names_plant <- plant_traits_matrix[,1]
+row_names_frugivore <- frugivore_traits_matrix[,1]
+
+# Assign row names to the matrix
+rownames(plant_traits_matrix) <- row_names_plant
+rownames(frugivore_traits_matrix) <- row_names_frugivore
+# Turn back into dataframe
+plant_traits_df_final <-as.data.frame(plant_traits_matrix)
+frugivore_traits_df_final <-as.data.frame(frugivore_traits_matrix)
+plant_traits_df$X <-NULL
+frugivore_traits_df$X <-NULL
+```
+
+```{r}
+#fix types
+frugivore_traits_df_final$diet_cat <- as.factor(frugivore_traits_df_final$diet_cat)
+frugivore_traits_df_final$body_mass_e <- as.numeric(frugivore_traits_df_final$body_mass_e)
+frugivore_traits_df_final$body_size_mm<- as.numeric(frugivore_traits_df_final$body_size_mm)
+frugivore_traits_df_final$generation_time <- as.numeric(frugivore_traits_df_final$generation_time)
+
+plant_traits_df_final$plant_height <- as.numeric(plant_traits_df_final$plant_height)
+plant_traits_df_final$plant_lifespan<- as.numeric(plant_traits_df_final$plant_lifespan)
+plant_traits_df_final$dispersal_syndrome <- as.factor(plant_traits_df_final$dispersal_syndrome)
+
+# remove duplicate species name column
+plant_traits_df_final <- plant_traits_df_final[, c("plant_height", "dispersal_syndrome", "plant_lifespan")]
+frugivore_traits_df_final <- frugivore_traits_df_final[, c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
+```
+
+# Create trait type table
+```{r}
+trait_name <- c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")
+trait_type <- c("N", "Q", "Q", "Q")
+frug_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
+trait_name <- c("dispersal_syndrome", "plant_height", "plant_lifespan")
+trait_type <- c("N", "Q", "Q") 
+plant_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
+```
+
+```{r}
+# Summary of the assemblages * species dataframe:
+asb_sp_plant_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_plant_site_final)
+asb_sp_frugivore_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_frugivore_site_final)
+```
+
+```{r}
+# Species traits summary:
+plant_traits_summ <- mFD::sp.tr.summary(
+  tr_cat     = plant_trait_cat,   
+  sp_tr      = plant_traits_df_final, 
+  stop_if_NA = TRUE)
+frugivore_traits_summ <- mFD::sp.tr.summary(
+  tr_cat     = frug_trait_cat,   
+  sp_tr      = frugivore_traits_df_final, 
+  stop_if_NA = TRUE)
+```
+
+# Estimate functional trait-based distances between species
+```{r}
+
+sp_dist_plants <- mFD::funct.dist(
+  sp_tr         = plant_traits_df_final,
+  tr_cat        = plant_trait_cat,
+  metric        = "gower",
+  scale_euclid  = "scale_center",
+  ordinal_var   = "classic",
+  weight_type   = "equal",
+  stop_if_NA    = TRUE)
+```
+
+```{r}
+sp_dist_frugivore <- mFD::funct.dist(
+  sp_tr         = frugivore_traits_df_final,
+  tr_cat        = frug_trait_cat,
+  metric        = "gower",
+  scale_euclid  = "scale_center",
+  ordinal_var   = "classic",
+  weight_type   = "equal",
+  stop_if_NA    = TRUE)
+```
+
+# Generate a multidimensional space
+```{r}
+fspaces_quality_plants <- mFD::quality.fspaces(
+  sp_dist             = sp_dist_plants,
+  maxdim_pcoa         = 10,
+  deviation_weighting = "absolute",
+  fdist_scaling       = FALSE,
+  fdendro             = "average")
+```
+
+```{r}
+fspaces_quality_frugivore <- mFD::quality.fspaces(
+  sp_dist             = sp_dist_frugivore,
+  maxdim_pcoa         = 10,
+  deviation_weighting = "absolute",
+  fdist_scaling       = FALSE,
+  fdendro             = "average")
+```
+
+```{r}
+# Look at the quality spaces only (MAD index looks at mean absolute deviation from the dissimilary matrix; want the deviation
+# to be low, meaning that the true distances have been retained in the PCA)
+round(fspaces_quality_plants$"quality_fspaces", 3)            
+round(fspaces_quality_frugivore$"quality_fspaces", 3)     
+# Plot the quality spaces (chose to look at 3D, 4D, and 5D since they had the lowest MAD). Will go with 4 dimensions.
+mFD::quality.fspaces.plot(
+  fspaces_quality            = fspaces_quality_plants,
+  quality_metric             = "mad",
+  fspaces_plot               = c("pcoa_3d", "pcoa_4d", 
+                                 "pcoa_5d"),
+  name_file                  = NULL,
+  range_dist                 = NULL,
+  range_dev                  = NULL,
+  range_qdev                 = NULL,
+  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
+  gradient_deviation_quality = c(low = "yellow", high = "red"),
+  x_lab                      = "Trait-based distance")
+mFD::quality.fspaces.plot(
+  fspaces_quality            = fspaces_quality_frugivore,
+  quality_metric             = "mad",
+  fspaces_plot               = c("pcoa_3d", "pcoa_4d", 
+                                 "pcoa_5d"),
+  name_file                  = NULL,
+  range_dist                 = NULL,
+  range_dev                  = NULL,
+  range_qdev                 = NULL,
+  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
+  gradient_deviation_quality = c(low = "yellow", high = "red"),
+  x_lab                      = "Trait-based distance")
+```
+
+```{r}
+#testing correlation between functional axes and traits
+sp_faxes_coord_plants <- fspaces_quality_plants$"details_fspaces"$"sp_pc_coord"
+sp_faxes_coord_frugivore <- fspaces_quality_frugivore$"details_fspaces"$"sp_pc_coord"
+# View the components of the PCA axes *Remember the firt few components explain the most variation in dissimilarity. Clusters into groups
+# Computes linear model for continuous traits and Kruskall-Wallis tests for other types. 
+plant_tr_faxes <- mFD::traits.faxes.cor(
+  sp_tr          = plant_traits_df_final, 
+  sp_faxes_coord = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")], 
+  plot           = TRUE)
+frugivore_tr_faxes <- mFD::traits.faxes.cor(
+  sp_tr          = frugivore_traits_df_final, 
+  sp_faxes_coord = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")], 
+  plot           = TRUE)
+# Print traits with significant effect:
+plant_tr_faxes$"tr_faxes_stat"[which(plant_tr_faxes$"tr_faxes_stat"$"p.value" < 0.05), ]
+frugivore_tr_faxes$"tr_faxes_stat"[which(frugivore_tr_faxes$"tr_faxes_stat"$"p.value" < 0.05), ]
+# Return plots:
+plant_tr_faxes$"tr_faxes_plot"
+frugivore_tr_faxes$"tr_faxes_plot"
+```
+
+```{r}
+#plotting functional space
+sp_faxes_coord_plants <- fspaces_quality_plants$"details_fspaces"$"sp_pc_coord"
+sp_faxes_coord_frugivore <- fspaces_quality_frugivore$"details_fspaces"$"sp_pc_coord"
+big_plot_plants <- mFD::funct.space.plot(
+  sp_faxes_coord  = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")],
+  faxes           = c("PC1", "PC2", "PC3", "PC4"),
+  name_file       = NULL,
+  faxes_nm        = NULL,
+  range_faxes     = c(NA, NA),
+  color_bg        = "grey95",
+  color_pool      = "darkgreen",
+  fill_pool       = "white",
+  shape_pool      = 21,
+  size_pool       = 1,
+  plot_ch         = TRUE,
+  color_ch        = "black",
+  fill_ch         = "white",
+  alpha_ch        = 0.5,
+  plot_vertices   = TRUE,
+  color_vert      = "blueviolet",
+  fill_vert       = "blueviolet",
+  shape_vert      = 6,
+  size_vert       = 1,
+  plot_sp_nm      = NULL,
+  nm_size         = 3,
+  nm_color        = "black",
+  nm_fontface     = "plain",
+  check_input     = TRUE)
+big_plot_frugivore <- mFD::funct.space.plot(
+  sp_faxes_coord  = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")],
+  faxes           = c("PC1", "PC2", "PC3", "PC4"),
+  name_file       = NULL,
+  faxes_nm        = NULL,
+  range_faxes     = c(NA, NA),
+  color_bg        = "grey95",
+  color_pool      = "darkgreen",
+  fill_pool       = "white",
+  shape_pool      = 21,
+  size_pool       = 1,
+  plot_ch         = TRUE,
+  color_ch        = "black",
+  fill_ch         = "white",
+  alpha_ch        = 0.5,
+  plot_vertices   = TRUE,
+  color_vert      = "blueviolet",
+  fill_vert       = "blueviolet",
+  shape_vert      = 6,
+  size_vert       = 1,
+  plot_sp_nm      = NULL,
+  nm_size         = 3,
+  nm_color        = "black",
+  nm_fontface     = "plain",
+  check_input     = TRUE)
+#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
+# Calculate row sums
+row_sums_plant <- rowSums(PAM_plant_site_final)
+subset_matrix_plant <- PAM_plant_site_final[row_sums_plant >= 4, ]
+row_sums_frugivore <- rowSums(PAM_frugivore_site_final)
+subset_matrix_frugivore <- PAM_frugivore_site_final[row_sums_frugivore >= 4, ]
+
+```
+
+```{r}
+big_plot_plants
+big_plot_frugivore
+```
+
+```{r}
+# getting sp_faxes_coord_plants and subset_matrix_plants names to match
+sp_faxes_coord_plants_sub <- sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")]
+sp_faxes_coord_plants_sub <- as.data.frame(sp_faxes_coord_plants_sub)
+sp_faxes_coord_plants_sub_names <- row.names(sp_faxes_coord_plants_sub)
+subset_matrix_plant <- as.data.frame(subset_matrix_plant)
+subset_matrix_plant_names <- colnames(subset_matrix_plant)
+subset_matrix_plant_sub <- subset_matrix_plant %>% filter(names(subset_matrix_plant) %in% sp_faxes_coord_plants_sub_names)
+```
+
+```{r}
+#computing FD
+# The number of species per assemblage has to be higher or equal to the number of traits
+alpha_fd_indices_plant <- mFD::alpha.fd.multidim(
+  sp_faxes_coord   = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")],
+  asb_sp_w         = subset_matrix_plant,
+  ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
+  scaling          = TRUE,
+  check_input      = TRUE,
+  details_returned = TRUE)
+
+alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(
+  sp_faxes_coord   = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")],
+  asb_sp_w         = subset_matrix_frugivore,
+  ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
+  scaling          = TRUE,
+  check_input      = TRUE,
+  details_returned = TRUE)
+
+details_list_plant <- alpha_fd_indices_plant$"details"
+details_list_frugivore <- alpha_fd_indices_frugivore$"details"
+
+#plot
+plots_alpha_plant <- mFD::alpha.multidim.plot(
+  output_alpha_fd_multidim = alpha_fd_indices_plant,
+  plot_asb_nm              = c("cell40", "cell200"),
+  ind_nm                   = c("fdis", "fide", "fnnd"
+                        , "fori", "fspe"),
+  faxes                    = NULL,
+  faxes_nm                 = NULL,
+  range_faxes              = c(NA, NA),
+  color_bg                 = "grey95",
+  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
+  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
+  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  color_ch          
+  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
+  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
+  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
+  shape_centroid_fspe      = 23,
+  color_centroid_fspe      = "black",
+  size_sp_nm               = 3, 
+  color_sp_nm              = "black",
+  plot_sp_nm               = NULL,
+  fontface_sp_nm           = "plain",
+  save_file                = FALSE,
+  check_input              = TRUE) 
+plots_alpha_frugivore <- mFD::alpha.multidim.plot(
+  output_alpha_fd_multidim = alpha_fd_indices_frugivore,
+  plot_asb_nm              = c("cell40", "cell200"),
+  ind_nm                   = c("fdis", "fide", "fnnd"
+                        , "fori", "fspe"),
+  faxes                    = NULL,
+  faxes_nm                 = NULL,
+  range_faxes              = c(NA, NA),
+  color_bg                 = "grey95",
+  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
+  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
+  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  color_ch          
+  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
+  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
+  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
+  shape_centroid_fspe      = 23,
+  color_centroid_fspe      = "black",
+  size_sp_nm               = 3, 
+  color_sp_nm              = "black",
+  plot_sp_nm               = NULL,
+  fontface_sp_nm           = "plain",
+  save_file                = FALSE,
+  check_input              = TRUE)
+
+plots_alpha_plant$"fric"$"patchwork"  #comparing convex hulls of different hulls
+plots_alpha_frugivore$"fric"$"patchwork"
+plots_alpha_plant$"fdiv"$"patchwork"
+plots_alpha_frugivore$"fdiv"$"patchwork"
+plots_alpha_plant$"fspe"$"patchwork"
+plots_alpha_frugivore$"fspe"$"patchwork"
+plots_alpha_plant$"fdis"$"patchwork"
+plots_alpha_frugivore$"fdis"$"patchwork"
+plots_alpha_plant$"fori"$"patchwork"
+plots_alpha_frugivore$"fori"$"patchwork"
+```
+
+```{r}
+## Turning the FD values into a raster
+
+#Need to subset the coordinates made at the top to the same set of the assemblages (removed values less than 5 so these coords need to be removed to do this correctly)
+
+#Have coordinates and associated values
+
+library(raster)
+
+# Create an empty raster with desired resolution and extent
+resolution <- c(0.75, 0.75)  # Set the desired resolution
+extent <- c(-85, -54, -24, 14)  # Set the desired extent
+empty_raster <- raster(resolution = resolution, xmn = extent[1], xmx = extent[2], ymn = extent[3], ymx = extent[4])
+
+# Generate coordinates
+subset_coords_plant <- site_loc_key_plant[rowSums(PAM_plant_site_final) >= 5, ]
+subset_coords_plant_sp <-subset_coords_plant[,1:2 ]
+subset_coords_frugivore <- site_loc_key_frugivore[rowSums(PAM_frugivore_site_final) >= 5, ]
+subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]
+
+#Get functional dispersion
+fdis_plant <- alpha_fd_indices_plant$functional_diversity_indices$fdis
+fdis_frugivore <- alpha_fd_indices_frugivore$functional_diversity_indices$fdis
+
+plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
+plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
+
+# Convert the dataframe to sf format
+spatial_fdis_bird <- st_as_sf(bird_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
+
+# Rasterize the sf data to create the FD raster
+fd_raster_bird <- rasterize(spatial_fdis_bird, empty_raster)
+writeRaster(fd_raster_bird$fdis_bird, "FD_birds_TA.tif", format="GTiff",overwrite=T)
+writeRaster(fd_raster_bird$ID, "FD_birds_TA_ID.tif", format="GTiff")
+
+```

### 2023-08-05 code/L2/L2_2_TropicalAndes_funcdiv.Rmd
add plotting code fix mismatched name issue

@@
 sp_faxes_coord_plants_sub <- sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")]
-sp_faxes_coord_plants_sub <- as.data.frame(sp_faxes_coord_plants_sub)
+## check number of species names
+nrow(sp_faxes_coord_plants_sub) # 54
+ncol(subset_matrix_plant) #82
+
 sp_faxes_coord_plants_sub_names <- row.names(sp_faxes_coord_plants_sub)
-subset_matrix_plant <- as.data.frame(subset_matrix_plant)
 subset_matrix_plant_names <- colnames(subset_matrix_plant)
-subset_matrix_plant_sub <- subset_matrix_plant %>% filter(names(subset_matrix_plant) %in% sp_faxes_coord_plants_sub_names)
+
+sp_faxes_coord_plants_sub_df <- as.data.frame(sp_faxes_coord_plants_sub)
+subset_matrix_plant_df <- as.data.frame(subset_matrix_plant)
+
+subset_matrix_plant <- subset_matrix_plant_df[, which((names(subset_matrix_plant_df) %in% sp_faxes_coord_plants_sub_names)==TRUE)]
+
+ncol(subset_matrix_plant)
+subset_matrix_plant <- as.matrix(subset_matrix_plant)
+```
+
+```{r}
+# match frugivore names
+sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")]
+
+## check number of species names
+nrow(sp_faxes_coord_frugivore_sub) #718
+ncol(subset_matrix_frugivore) #712
+
+sp_faxes_coord_frugivore_sub_names <- row.names(sp_faxes_coord_frugivore_sub)
+subset_matrix_frugivore_names <- colnames(subset_matrix_frugivore)
+
+sp_faxes_coord_frugivore_sub_names <- chartr(".", " ", sp_faxes_coord_frugivore_sub_names)
+sp_faxes_coord_frugivore_sub <- as.data.frame(sp_faxes_coord_frugivore_sub)
+row.names(sp_faxes_coord_frugivore_sub) <- sp_faxes_coord_frugivore_sub_names
+
+sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore_sub[ which((row.names(sp_faxes_coord_frugivore_sub) %in% subset_matrix_frugivore_names)==TRUE), ]
+
+nrow(sp_faxes_coord_frugivore_sub)
+sp_faxes_coord_frugivore_sub <- as.matrix(sp_faxes_coord_frugivore_sub)
 ```
@@
 alpha_fd_indices_plant <- mFD::alpha.fd.multidim(
-  sp_faxes_coord   = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")],
+  sp_faxes_coord   = sp_faxes_coord_plants_sub,
   asb_sp_w         = subset_matrix_plant,
@@
 alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(
-  sp_faxes_coord   = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")],
+  sp_faxes_coord   = sp_faxes_coord_frugivore_sub,
   asb_sp_w         = subset_matrix_frugivore,
@@
   output_alpha_fd_multidim = alpha_fd_indices_plant,
-  plot_asb_nm              = c("cell40", "cell200"),
+  plot_asb_nm              = c("cell4", "cell20"),
   ind_nm                   = c("fdis", "fide", "fnnd"
@@
   check_input              = TRUE)
+
 plots_alpha_frugivore <- mFD::alpha.multidim.plot(
   output_alpha_fd_multidim = alpha_fd_indices_frugivore,
-  plot_asb_nm              = c("cell40", "cell200"),
+  plot_asb_nm              = c("cell4", "cell20"),
   ind_nm                   = c("fdis", "fide", "fnnd"
@@
 
+# Turning the FD values into a raster
 ```{r}
-## Turning the FD values into a raster
 
@@
 plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
-plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
+frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)
 
 # Convert the dataframe to sf format
-spatial_fdis_bird <- st_as_sf(bird_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
-
-# Rasterize the sf data to create the FD raster
-fd_raster_bird <- rasterize(spatial_fdis_bird, empty_raster)
-writeRaster(fd_raster_bird$fdis_bird, "FD_birds_TA.tif", format="GTiff",overwrite=T)
-writeRaster(fd_raster_bird$ID, "FD_birds_TA_ID.tif", format="GTiff")
+spatial_fdis_plant <- st_as_sf(plant_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
+spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
 
 ```
+
+```{r}
+# Rasterize the sf data to create the FD raster
+fd_raster_plant <- rasterize(spatial_fdis_plant, empty_raster)
+writeRaster(fd_raster_plant$fdis_plant, filename = file.path(output_path,"FD_plant_TA.tif"), format="GTiff",overwrite=T)
+writeRaster(fd_raster_plant$ID, filename = file.path(output_path,"FD_plant_TA_ID.tif"), format="GTiff")
+
+fd_raster_frugivore <- rasterize(spatial_fdis_frugivore, empty_raster)
+writeRaster(fd_raster_frugivore$fdis_frugivore, filename = file.path(output_path,"FD_frugivore_TA.tif"), format="GTiff",overwrite=T)
+writeRaster(fd_raster_frugivore$ID, filename = file.path(output_path,"FD_frugivore_TA_ID.tif"), format="GTiff")
+```
+
+# Functional diversity plots
+```{r}
+# set crs of sf objects
+spatial_fdis_plant <- spatial_fdis_plant %>% st_set_crs(4326)
+spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(4326)
+```
+
+```{r}
+# Raster to raster point
+spec_FD_raster_plant_point <- rasterToPoints(fd_raster_plant, spatial = TRUE)
+spec_FD_raster_frugivore_point <- rasterToPoints(fd_raster_frugivore, spatial = TRUE)
+# Convert to a 'conventional' dataframe
+fd_raster_plant_point <- data.frame(spec_FD_raster_plant_point)
+fd_raster_frugivore_point <- data.frame(spec_FD_raster_frugivore_point)
+```
+
+```{r}
+plantgridFDTA <-
+  ggplot() +
+  geom_sf(data = worldMap, fill = "white")+
+  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
+  geom_tile(data= fd_raster_plant_point, aes(x=x, y=y, fill=fdis_plant)) +
+  scale_fill_scico(palette = "romaO", direction = 1) +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
+  theme(
+    plot.background = element_rect(fill = "#f1f2f3"),
+    panel.background = element_rect(fill = "lightblue"),
+    panel.grid = element_blank(),
+    line = element_blank(),
+    rect = element_blank()
+  ) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  labs(title = "Tropical Andes Plant Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
+plantgridFDTA
+```
+
+```{r}
+frugivoregridFDTA <-
+  ggplot() +
+  geom_sf(data = worldMap, fill = "white")+
+  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
+  geom_tile(data= fd_raster_frugivore_point, aes(x=x, y=y, fill=fdis_frugivore)) +
+  scale_fill_scico(palette = "romaO", direction = 1) +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
+  theme(
+    plot.background = element_rect(fill = "#f1f2f3"),
+    panel.background = element_rect(fill = "lightblue"),
+    panel.grid = element_blank(),
+    line = element_blank(),
+    rect = element_blank()
+  ) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  labs(title = "Tropical Andes Frugivore Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
+frugivoregridFDTA
+```
+
+# Extract cell values
+```{r}
+plantgridFDTA <- plantgridFDTA[["data"]]
+frugivoregridFDTA <- frugivoregridFDTA[["data"]]
+```
+
+```{r}
+plot(spatial_fdis_plant$fdis_plant, spatial_fdis_frugivore$fdis_frugivore,
+     main = "Plant FD vs Frugivore FD",
+     xlab = "Plant Functional Dispersion by cell", ylab = "Frugivore Functional Dispersion by cell")
+abline(lm(spatial_fdis_frugivore$fdis_frugivore ~ spatial_fdis_plant$fdis_plant))
+summary(lm(spatial_fdis_frugivore$fdis_frugivore ~ spatial_fdis_plant$fdis_plant))
+```

### 2023-08-05 code/L2/L2_2_TropicalAndes_funcdiv.Rmd
update plots

@@
   quality_metric             = "mad",
-  fspaces_plot               = c("pcoa_3d", "pcoa_4d", 
-                                 "pcoa_5d"),
+  fspaces_plot               = c("pcoa_2d", "pcoa_3d", 
+                                 "pcoa_4d"),
   name_file                  = NULL,
@@
 alpha_fd_indices_plant <- mFD::alpha.fd.multidim(
-  sp_faxes_coord   = sp_faxes_coord_plants_sub,
-  asb_sp_w         = subset_matrix_plant,
+  sp_faxes_coord   = sp_faxes_coord_plants,
+  asb_sp_w         = PAM_plant_site_final,
   ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
@@
 alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(
-  sp_faxes_coord   = sp_faxes_coord_frugivore_sub,
-  asb_sp_w         = subset_matrix_frugivore,
+  sp_faxes_coord   = sp_faxes_coord_frugivore,
+  asb_sp_w         = PAM_plant_site_final,
   ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
@@
 fdis_frugivore <- alpha_fd_indices_frugivore$functional_diversity_indices$fdis
+## subset frugivore to compare for plotting
+fdis_frugivore_sub <- fdis_frugivore[1:25]
 
 plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
+frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)
+## subset frugivore to compare for plotting
 frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)

### 2023-09-05 code/L1/L1_TropicalAndes_plant_lookupTable.Rmd
add TRNS package code add taxize code

@@
 ```{r}
-
+library(taxize)
+library(knitr)
+library(TNRS)
 ```
@@
 BIEN_species <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_trait_species.csv"))
+GBIF_species <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_species.csv"))
 ```
@@
 ```{r}
-#convert species list into vectors
-BIEN_species <- BIEN_species[,2]
-TRY_species <- TRY_species[,2]
-#what is in both
-species <- intersect(TRY_species, BIEN_species)
-# what doesn't match
-all.equal(TRY_species, BIEN_species)
+# convert species list to character
+TRY_species_chr <- as.character(TRY_species$x)
+BIEN_species_chr <- as.character(BIEN_species$x)
+GBIF_species_chr <- as.character(GBIF_species$x)
+```
 
+# Using TNRS
+```{r}
+BIEN_results <- TNRS(taxonomic_names = BIEN_species)
+TRY_results <- TNRS(taxonomic_names = TRY_species)
+GBIF_results <- TNRS(taxonomic_names = GBIF_species)
 ```
+
+# Using taxsize
+```{r}
+GBIF_species_resolved <- resolve(GBIF_species)
+TRY_species_resolved <- resolve(TRY_species)
+BIEN_species_resolved <- resolve(BIEN_species)
+```

### 2023-09-10 code/L1/L1_TropicalAndes_plant_lookupTable.Rmd
split up gbif call to speed up api call combine gbif results save results

@@
 BIEN_results <- TNRS(taxonomic_names = BIEN_species)
+```
+
+```{r}
 TRY_results <- TNRS(taxonomic_names = TRY_species)
-GBIF_results <- TNRS(taxonomic_names = GBIF_species)
+```
+
+```{r}
+#split up gbif species list to speed up api call
+GBIF_species1 <-GBIF_species[1:33107,]
+GBIF_species2 <-GBIF_species[33107:66214,]
+GBIF_species3 <-GBIF_species[66214:99321,]
+GBIF_species4 <-GBIF_species[99321:132428,]
+
+GBIF_results1 <- TNRS(taxonomic_names = GBIF_species1)
+GBIF_results2 <- TNRS(taxonomic_names = GBIF_species2)
+GBIF_results3 <- TNRS(taxonomic_names = GBIF_species3)
+GBIF_results4 <- TNRS(taxonomic_names = GBIF_species4)
+```
+
+```{r}
+#combine gbif results
+GBIF_results <- rbind(GBIF_results1, GBIF_results2, GBIF_results3, GBIF_results4)
+```
+
+```{r}
+#save full results
+write.csv(BIEN_results, file.path(output_path,"BIEN_plant_species_resolved_output_full.csv"))
+write.csv(TRY_results, file.path(output_path,"TRY_plant_species_resolved_output_full.csv"))
+write.csv(GBIF_results, file.path(output_path,"GBIF_plant_species_resolved_output_full.csv"))
+```
+
+```{r}
+# subset results to rows with Name_submitted values not the same as Accepted_name
+
 ```

### 2023-09-12 code/L1/L1_TropicalAndes_plant_lookupTable.Rmd
split gbif calls up even more seperate code chunks

@@
 #split up gbif species list to speed up api call
-GBIF_species1 <-GBIF_species[1:33107,]
-GBIF_species2 <-GBIF_species[33107:66214,]
-GBIF_species3 <-GBIF_species[66214:99321,]
-GBIF_species4 <-GBIF_species[99321:132428,]
+GBIF_species1 <-GBIF_species[1:16553,]
+GBIF_species2 <-GBIF_species[16553:33107,]
+GBIF_species3 <-GBIF_species[33107:49660,]
+GBIF_species4 <-GBIF_species[49660:66214,]
+GBIF_species5 <-GBIF_species[66214:82767,]
+GBIF_species6 <-GBIF_species[82767:99321,]
+GBIF_species7 <-GBIF_species[99321:115874,]
+GBIF_species8 <-GBIF_species[115874:132428,]
+```
 
+```{r}
+#run one line at a time for best results
 GBIF_results1 <- TNRS(taxonomic_names = GBIF_species1)
@@
 GBIF_results4 <- TNRS(taxonomic_names = GBIF_species4)
+GBIF_results5 <- TNRS(taxonomic_names = GBIF_species5)
+GBIF_results6 <- TNRS(taxonomic_names = GBIF_species6)
+GBIF_results7 <- TNRS(taxonomic_names = GBIF_species7)
+GBIF_results8 <- TNRS(taxonomic_names = GBIF_species8)
 ```
@@
 #combine gbif results
-GBIF_results <- rbind(GBIF_results1, GBIF_results2, GBIF_results3, GBIF_results4)
+GBIF_results <- rbind(GBIF_results1, GBIF_results2, GBIF_results3, GBIF_results4, GBIF_results5, GBIF_results6, GBIF_results7, GBIF_results8)
 ```

### 2023-09-13 archive/MappingTopicalAndes.Rmd
update for forest

@@
 library(ggspatial)
+```
+
+# Load Tropical Andes moist lowland and montane forest shape
+```{r}
+TropicalAndes_IUCNHabitat_Forest <- st_read(file.path(data_path, "TropicalAndes_IUCNHabitat_Forest.shp"))
+```
+
+```{r}
+# set projection to WGS84
+st_crs(TropicalAndes_IUCNHabitat_Forest) = 4326
+TropicalAndes_IUCNHabitat_Forest
 ```
@@
 # set color for Tropical Andes
-Tropical_Andes <- "lightgreen"
+Tropical_Andes <- "grey"
+Forest <- "darkgreen"
 ```
@@
 geom_sf(data = BOpoly, fill = Tropical_Andes) +
+geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = Forest) +
 labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
@@
 ```{r}
-ggsave(filename = "Tropical Andes Map.png", dpi = 300, device="png")
+ggsave(filename = "Tropical Andes Forest Map.png", dpi = 300, device="png")
 ```

### 2023-09-13 code/L1/L1_TropicalAndes_plant_taxonomiccoverage.Rmd
inital commit

```{r}
#input trait datasets
#compare coverage of species, families, genera by trait coverage (plot)
#do the same for GBIF occurrence 
library(WorldFlora)
library(taxize)
library(BIEN)
library(GIFT)
library(dplyr)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Read in Data
```{r}
BIEN_traits <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_traits_forest.csv"))
TRY_traits <- read.csv(file.path(data_path,"TropicalAndes_TRY_plant_traits_forestSpecies.csv"))
GIFT_traits <- read.csv(file.path(data_path,"TropicalAndes_GIFT_plant_traits_forest.csv"))
GBIF_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
```

```{r}
#species list
BIEN_species <- unique(BIEN_traits$scrubbed_species_binomial)
TRY_species <- unique(TRY_traits$SpeciesName)
GIFT_species <- unique(GIFT_traits$species)
GBIF_species <- unique(GBIF_occ$species)
```

# Get higher taxonomy for species lists
```{r}
BIEN_tax<- BIEN_taxonomy_species(BIEN_species)
```

```{r}
TRY_tax <- tax_name(TRY_species, get = c('family', 'genus'), db= "itis")
```

```{r}
gift_sp <- GIFT_species()
sapply(GIFT_species, function(x) grep(x, gift_sp$work_species))
gift_sp[sapply(GIFT_species, function(x) grep(x, gift_sp$work_species)), ]

#GIFT_tax <- GIFT_taxgroup(taxon_lvl = c("family", "genus")[1], species = GIFT_species)
```

# Plot coverage
```{r}
BIEN_family_coverage <- BIEN_tax %>% count(scrubbed_family)
BIEN_family_coverage 
BIEN_genus_coverage <- BIEN_tax %>% count(scrubbed_genus) 
BIEN_genus_coverage
barplot(height = BIEN_family_coverage$n, names = BIEN_family_coverage$scrubbed_family, horiz=T, las=1)
barplot(height = BIEN_genus_coverage$n, names = BIEN_genus_coverage$scrubbed_genus, horiz=T, las=1)
```

### 2023-09-19 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
remove forest subset code
clean up unnecessary code

--- 2023-07-25 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
@@
 library(rgbif)
+library(raster)
+library(sf)
 ```
 
-# Plant occurrences using IUCN habitat subset
-## Read in IUCN data
-Data was obtained from https://www.iucnredlist.org/search?permalink=9621ab77-2341-4bf8-9557-b7f19f7f2a35
-```{r}
-plants_habitats <- read.csv(file.path(data_path,"IUCN/Tropical Andes Plants - Search Results/habitats.csv"))
-```
-
-## Create species list from IUCN data
-```{r}
-plant_species <- unique(plants_habitats$scientificName)
-print("The number of species on the IUCN is")
-length(plant_species)
-```
-
-## Match species names to GBIF
-```{r}
-gbif_taxon_keys <- plant_species %>% 
-name_backbone_checklist() %>% # match to backbone 
-filter(!matchType == "NONE") %>% # get matched names
-pull(usageKey) 
-```
-
-## Download data
-```{r}
-occ_download(
-pred_in("taxonKey", gbif_taxon_keys),
-pred("hasCoordinate", TRUE),
-pred("hasGeospatialIssue", FALSE),
-pred_in("country",c("EC","CO","VE", "PE", "BO")),
-format = "SIMPLE_CSV"
-)
-```
-
-## Retrive GBIF download, save file, and load into r
-```{r}
-d <- occ_download_get("0097296-230530130749713", path = output_path) %>%
-  occ_download_import()
-```
-
-## Summary 
-```{r}
-glimpse(d)
-```
-
-```{r}
-print("The number of occurrences is")
-nrow(d)
-print("The number of species is" )
-length(unique(d$species))
-print("The number of families is" )
-length(unique(d$family))
-```
-
-# Plant occurrences without habitat subset
-
-## Download data
+# Download plant occurrence data from GBIF
 ```{r}
@@
 ```{r}
-d2 <- occ_download_get("0097297-230530130749713", path = output_path, overwrite = TRUE) %>%
+d1 <- occ_download_get("0097297-230530130749713", path = output_path, overwrite = TRUE) %>%
   occ_download_import()
@@
 
-## Summary 
+# Summary 
 ```{r}
-glimpse(d2)
+glimpse(d1)
 ```
@@
 print("The number of occurrences is")
-nrow(d2)
+nrow(d1)
 print("The number of species is" )
-length(unique(d2$species))
+length(unique(d1$species))
 print("The number of families is" )
-length(unique(d2$family))
+length(unique(d1$family))
 ```
@@
 ```{r}
-GBIF_SpeciesList <- unique(d2$species)
+GBIF_SpeciesList <- unique(d1$species)
 ```

### 2023-09-20 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
update GBIF call to correct not pulling for plants

@@
 occ_download(
+pred("taxonKey", 6),
 pred("hasCoordinate", TRUE),
@@
 ```{r}
-d1 <- occ_download_get("0097297-230530130749713", path = output_path, overwrite = TRUE) %>%
+d1 <- occ_download_get("0001452-230918134249559", path = output_path, overwrite = TRUE) %>%
   occ_download_import()
@@
 length(unique(d1$species))
+print("The number of genera is" )
+length(unique(d1$genus))
 print("The number of families is" )
@@
 TropicalAndes_GBIF_path <- file.path(output_path,"TropicalAndes_GBIF_plant_occ.csv") 
-write.csv(d2, file = TropicalAndes_GBIF_path)
+write.csv(d1, file = TropicalAndes_GBIF_path)
 TropicalAndes_GBIF_species_path <- file.path(output_path,"TropicalAndes_GBIF_plant_species.csv") 

### 2023-09-27 code/L2/L2_2_TropicalAndes_taxdiv.Rmd
add forest shape to blank map fix north arrow

@@
 geom_sf(data = TApoly) +
+geom_sf(data = Forest_sf, fill = "forestgreen")+
 labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
@@
 annotation_north_arrow(location = "bl", which_north = "true", 
-  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
+  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
 theme(panel.background = element_rect(fill = "lightblue"))

### 2023-09-27 code/L1/L1_TropicalAndes_plant_lookupTable.Rmd
update file path, gbif species list spilt; add code to subset results

--- 2023-09-05 code/L1/L1_TropicalAndes_plant_lookupTable.Rmd
@@
 data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+data_path2 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
 output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
@@
 BIEN_species <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_trait_species.csv"))
-GBIF_species <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_species.csv"))
-```
-
-```{r}
-# convert species list to character
-TRY_species_chr <- as.character(TRY_species$x)
-BIEN_species_chr <- as.character(BIEN_species$x)
-GBIF_species_chr <- as.character(GBIF_species$x)
+GBIF_species <- read.csv(file.path(data_path2,"TropicalAndes_plant_Forest_SpeciesList.csv"))
 ```
@@
 BIEN_results <- TNRS(taxonomic_names = BIEN_species)
-TRY_results <- TNRS(taxonomic_names = TRY_species)
-GBIF_results <- TNRS(taxonomic_names = GBIF_species)
 ```
 
-# Using taxsize
 ```{r}
-GBIF_species_resolved <- resolve(GBIF_species)
-TRY_species_resolved <- resolve(TRY_species)
-BIEN_species_resolved <- resolve(BIEN_species)
+TRY_results <- TNRS(taxonomic_names = TRY_species)
 ```
+
+```{r}
+#split up gbif species list to speed up api call
+GBIF_species1 <-GBIF_species[1:2245,]
+GBIF_species2 <-GBIF_species[2245:6745,]
+GBIF_species3 <-GBIF_species[6745:11245,]
+GBIF_species4 <-GBIF_species[11245:15745,]
+GBIF_species5 <-GBIF_species[15745:20245,]
+GBIF_species6 <-GBIF_species[16553:20245,]
+GBIF_species7 <-GBIF_species[20245:24745,]
+GBIF_species8 <-GBIF_species[24745:29245,]
+GBIF_species9 <-GBIF_species[29245:33745,]
+GBIF_species10 <-GBIF_species[33745:38245,]
+GBIF_species11 <-GBIF_species[38245:42745,]
+GBIF_species12 <-GBIF_species[42745:47245,]
+```
+
+```{r}
+GBIF_results1 <- TNRS(taxonomic_names = GBIF_species1)
+GBIF_results2 <- TNRS(taxonomic_names = GBIF_species2)
+GBIF_results3 <- TNRS(taxonomic_names = GBIF_species3)
+GBIF_results4 <- TNRS(taxonomic_names = GBIF_species4)
+GBIF_results5 <- TNRS(taxonomic_names = GBIF_species5)
+GBIF_results6 <- TNRS(taxonomic_names = GBIF_species6)
+GBIF_results7 <- TNRS(taxonomic_names = GBIF_species7)
+GBIF_results8 <- TNRS(taxonomic_names = GBIF_species8)
+GBIF_results9 <- TNRS(taxonomic_names = GBIF_species9)
+GBIF_results10 <- TNRS(taxonomic_names = GBIF_species10)
+GBIF_results11 <- TNRS(taxonomic_names = GBIF_species11)
+GBIF_results12 <- TNRS(taxonomic_names = GBIF_species12)
+```
+
+```{r}
+#combine gbif results
+GBIF_results <- rbind(GBIF_results1, GBIF_results2, GBIF_results3, GBIF_results4, GBIF_results5, GBIF_results6, GBIF_results7, GBIF_results8, GBIF_results9, GBIF_results10, GBIF_results11, GBIF_results12)
+```
+
+# Save full results
+```{r}
+write.csv(BIEN_results, file.path(output_path,"BIEN_plant_species_resolved_output_full.csv"))
+write.csv(TRY_results, file.path(output_path,"TRY_plant_species_resolved_output_full.csv"))
+write.csv(GBIF_results, file.path(output_path,"GBIF_plant_species_resolved_output_full.csv"))
+```
+
+# Subset results to species with matches, conflicts, and disparties
+Name_submitted values not the same as Accepted_name
+```{r}
+BIEN_matches <- subset(BIEN_results, Taxonomic_status == "Accepted")
+#Results in 3799 species
+BIEN_conflicts <- subset(BIEN_results, Taxonomic_status == "Synonym")
+#Results in 646 species
+BIEN_disparities <- subset(BIEN_results, Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", ""))
+#Results in 44 species
+```
+
+```{r}
+TRY_matches <- subset(TRY_results, Taxonomic_status == "Accepted")
+#Results in 4043 species
+TRY_conflicts <- subset(TRY_results, Taxonomic_status == "Synonym")
+#Results in 307 species
+TRY_disparities <- subset(TRY_results, Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", ""))
+#Results in 37 species
+```
+
+```{r}
+GBIF_matches <- subset(GBIF_results, Taxonomic_status == "Accepted")
+#Results in 47554 species
+GBIF_conflicts <- subset(GBIF_results, Taxonomic_status == "Synonym")
+#Results in 1572 species
+GBIF_disparities <- subset(GBIF_results, Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", ""))
+#Results in 1821 species
+```
+
+# Checking 
+```{r}
+
+```

### 2023-09-27 code/L1/L1_TropicalAndes_plant_species_synonyms.Rmd
inital commit

Get synonmyms for species list

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path2 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

```{r}
library(taxadb)
```

```{r}
#create local itis database
td_create("itis",overwrite=FALSE)
```

# Read in species lists
```{r}
TRY_species <- read.csv(file.path(data_path,"TropicalAndes_TRY_plant_species.csv"))
BIEN_species <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_trait_species.csv"))
GBIF_species <- read.csv(file.path(data_path2,"TropicalAndes_plant_Forest_SpeciesList.csv"))
```

# Combine species lists
```{r}
allnames <- rbind(TRY_species, BIEN_species, GBIF_species)
```

```{r}
#get  IDS for each scientific name
syn1<-allnames %>%
  select(Scientific.Name) %>%
  mutate(ID=get_ids(Scientific.Name,"itis"))

#Deal with NAs (one name corresponds to more than 1 ITIS code) (~10k names)

syn1_NA<-as.data.frame(syn1$Scientific.Name[is.na(syn1$ID)])
colnames(syn1_NA)<-c("name")

NA_IDS<-NULL
for(i in unique(syn1_NA$name)){
  tmp<-as.data.frame(filter_name(i, 'itis')[5])
  tmp$name<-paste0(i)
  NA_IDS<-rbind(NA_IDS,tmp)
}

#join with original names
colnames(syn1)<-c("name","ID")
IDS<-left_join(syn1,NA_IDS,by="name") #I think its a left join double check this

#extract just the unique IDs
IDS<-data.frame(ID=c(IDS[,"ID"],IDS[,"acceptedNameUsageID"]))
IDS<-as.data.frame(unique(IDS$ID))
IDS<-as.data.frame(IDS[-is.na(IDS)])
colnames(IDS)<-"ID"
#extract all names with synonyms in ITIS that are at the species level [literally all of them]
#set query
ITIS<-taxa_tbl("itis") %>%
  select(scientificName,taxonRank,acceptedNameUsageID,taxonomicStatus) %>%
  filter(taxonRank == "species")

#see query
ITIS %>% show_query()
#retrieve results
ITIS_names<-ITIS %>% collect()

#filter to only those that match ITIS codes for all my species
ITIS_names<-ITIS_names %>%
  filter(acceptedNameUsageID %in% IDS$ID)
```

### 2023-09-27 code/L2/L2_TropicalAndes_plant_taxonomiccoverage.Rmd
moved from L1 to L2

--- 2023-09-13 code/L1/L1_TropicalAndes_plant_taxonomiccoverage.Rmd
@@
 GIFT_traits <- read.csv(file.path(data_path,"TropicalAndes_GIFT_plant_traits_forest.csv"))
-GBIF_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
+GBIF_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ.csv"))
 ```

### 2023-09-27 code/L1/L1_TropicalAndes_subset_occ_forest.Rmd
add genera count and save raster file

@@
 length(unique(plant_occ_forest$species))
+print("The number of plant genera is")
+length(unique(plant_occ_forest$genus))
 print("The number of plant families is")
@@
 length(unique(frugivore_occ_forest$species))
+print("The number of frugivore genera is")
+length(unique(frugivore_occ_forest$genus))
 print("The number of frugivore families is")
@@
 write.csv(TropicalAndes_frugivore_Forest_SpeciesList, file = file.path(output_path,"TropicalAndes_frugivore_Forest_SpeciesList.csv"))
+writeRaster(TropicalAndes_IUCNHabitat_Forest0.3,file.path(output_path, "TropicalAndes_IUCNHabitat_Forest0.3.tif"), overwrite = TRUE)
 ```

### 2023-09-27 code/L0/L0_TropicalAndes_plants_TRY.Rmd
standardizing code file naming

---
title: "Tropical Andes plant trait data TRY"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script retrieves plant trait data from the TRY database for plant species list."
data input: "none"
data output: "TropicalAndes_TRY_traits.csv"
date: "2023-07-18"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(rtry)
library(dplyr)
```

# Load TRY data
Data was obtained from https://www.try-db.org/TryWeb/Prop0.php 
```{r}
try_data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/28165_02082023081407/28165.txt")
```

```{r}
glimpse(try_data)
```

#Subset to species list
```{r}
species_list <- read.csv("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1/TropicalAndes_plant_Forest_SpeciesList.csv")
species_list <- species_list$x
```

```{r}
try_data_filtered <- try_data %>%
  filter(AccSpeciesName %in% species_list)
```

# Summary 
```{r}
glimpse(try_data_filtered)
```

```{r}
print("The number of records is")
nrow(try_data_filtered)
print("The number of species is" )
length(unique(try_data_filtered$AccSpeciesName))
```

```{r}
try_data_filtered %>% count(TraitName)
```

# Write data to csv
```{r}
TropicalAndes_TRY_path <- file.path(output_path,"TropicalAndes_TRY_plant_traits_forestSpecies.csv") 
write.csv(try_data_filtered, file = TropicalAndes_TRY_path)
```

### 2023-10-04 code/L2/L2_2_TropicalAndes_funcdiv.Rmd
remove forest habitat code to simplify

@@
 plant_traits <- read.csv(file.path(data_path,"TropicalAndes_plant_traits_forest.csv"))
-```
-
-# Forest shape conversion
-```{r}
-TropicalAndes_IUCNHabitat_Forest <- aggregate(TropicalAndes_IUCNHabitat_Forest, 50, fun=mean, na.rm=TRUE)
-TropicalAndes_IUCNHabitat_Forest
-```
-
-```{r}
-Forest_poly <- rasterToPolygons(TropicalAndes_IUCNHabitat_Forest, na.rm = TRUE, dissolve = TRUE)
-Forest_poly
-plot(Forest_poly)
-```
-
-```{r}
-Forest_sf <- st_as_sf(Forest_poly)
-Forest_sf
-plot(Forest_sf)
-st_write(Forest_sf, file.path(data_path, "Forest_sf.shp"))
 ```

### 2023-10-04 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
add aggregation and type conversion code

@@
 
-# Crop moist montane to Tropical Andes
+## Crop moist montane
 ```{r}
@@
 
-# Crop moist lowland to Tropical Andes
+## Crop moist lowland
 ```{r}
@@
 ```
+# Aggregate to 0.3
+```{r}
+TropicalAndes_IUCNHabitat_Forest0.3 <- aggregate(TropicalAndes_IUCNHabitat_Forest, 30, fun=mean, na.rm=TRUE)
+TropicalAndes_IUCNHabitat_Forest0.3
+```
+
+```{r}
+writeRaster(TropicalAndes_IUCNHabitat_Forest0.3,file.path(output_path, "TropicalAndes_IUCNHabitat_Forest0.3.tif"), overwrite = TRUE)
+```
+
+# Forest shape conversion
+```{r}
+TropicalAndes_IUCNHabitat_Forest <- aggregate(TropicalAndes_IUCNHabitat_Forest, 50, fun=mean, na.rm=TRUE)
+TropicalAndes_IUCNHabitat_Forest
+```
+
+```{r}
+Forest_poly <- rasterToPolygons(TropicalAndes_IUCNHabitat_Forest, na.rm = TRUE, dissolve = TRUE)
+Forest_poly
+plot(Forest_poly)
+```
+
+```{r}
+Forest_sf <- st_as_sf(Forest_poly)
+Forest_sf
+plot(Forest_sf)
+st_write(Forest_sf, file.path(data_path, "Forest_sf.shp"))
+```

### 2023-10-04 archive/L1_TropicalAndes_subset_occ_forest.Rmd
clean repo: update names and move old code to archive

--- 2023-08-03 code/L1/L1_TropicalAndes_subset_occ_forest.Rmd
@@
 TropicalAndes_IUCNHabitat_Forest
-```
-
-```{r}
-TropicalAndes_IUCNHabitat_Forest0.3 <- aggregate(TropicalAndes_IUCNHabitat_Forest, 30, fun=mean, na.rm=TRUE)
-TropicalAndes_IUCNHabitat_Forest0.3
 ```

### 2023-10-04 code/L1/L1_F1_TropicalAndes_frugivores_GBIF.Rmd
clean repo: update names and move old code to archive

--- 2023-08-01 code/L1/L1_TropicalAndes_frugivores_GBIF.Rmd
@@
 #exclude problematic records
-TropicalAndes_GBIF_frugivore_occ_cleaned <- TropicalAndes_GBIF_frugivore_occ[flags$summary,]
-TropicalAndes_GBIF_frugivore_occ_flagged <- TropicalAndes_GBIF_frugivore_occ[!flags$summary,]
+TropicalAndes_GBIF_frugivore_occ_cleaned <- TropicalAndes_GBIF_frugivore_occ[flags$.summary,]
+TropicalAndes_GBIF_frugivore_occ_flagged <- TropicalAndes_GBIF_frugivore_occ[!flags$.summary,]
 ```

### 2023-10-04 code/L1/L1_F2_TropicalAndes_frugivore_traits.Rmd
clean repo: update names and move old code to archive

---
title: "Frugivoria frugivore trait subset"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Subsetting Frugivoria to the species and traits of interest"
data input: ""
data output: ""
date: "2023-08-03"
output: html_document
---

# Load required packages
```{r}
library(dplyr)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path2<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Read in data
```{r}
Frugivoria <- read.csv(file.path(data_path, file = "TropicalAndes_Frugivoria_frugivore_traits.csv"))
frugivore_species <- read.csv(file.path(data_path2, file = "TropicalAndes_frugivore_Forest_SpeciesList.csv"))
```

```{r}
frugivore_species <- frugivore_species$x
```

```{r}
# subset Frugivoria by species list
Frugivoria_filtered <- Frugivoria %>%
  filter(IUCN_species_name %in% frugivore_species)
```

```{r}
# keep columns with traits of diet category, body size, body mass, generation time
Frugivoria_subset <- Frugivoria_filtered[ , c("IUCN_species_name", "family", "genus", "species", "diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
```

# Summary
```{r}
glimpse(Frugivoria_subset)
```

# Write data to csv
```{r}
write.csv(Frugivoria_subset, file.path(output_path,"TropicalAndes_Frugivoria_traits_Forest.csv"))
```

### 2023-10-04 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
clean repo - update names and move old code to archive

--- 2023-10-04 code/L2/L2_2_TropicalAndes_funcdiv.Rmd
@@
 ```{r}
-TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_plant_occ_forest_sub.csv"))
-TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_frugivore_occ_forest_sub.csv"))
+TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_plant_occ_forest.csv"))
+TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_frugivore_occ_forest.csv"))
 TropicalAndes_IUCNHabitat_Forest <- raster(file.path(data_path2, "TropicalAndes_IUCNHabitat_Forest.tif"))

### 2023-10-04 archive/L2_1_TropicalAndes_plant_traits.Rmd
clean repo - update names and move old code to archive

--- 2023-08-04 code/L2/L2_1_TropicalAndes_plant_traits.Rmd
@@
 
+```{r}
+plant_traits_wide_complete$dispersal_syndrome <- as.character(plant_traits_wide_complete$dispersal_syndrome)
+plant_traits_wide_complete$plant_lifespan <- as.character(plant_traits_wide_complete$plant_lifespan)
+```
+
 # Write data to csv
 ```{r}
-write.csv(plant_traits, file.path(output_path,"TropicalAndes_plant_traits_forest.csv"))
+write.csv(plant_traits_wide_complete, file.path(output_path,"TropicalAndes_plant_traits_forest.csv"))
 write.csv(species_list, file.path(output_path,"TropicalAndes_plant_traits_forest_species_list.csv"))

### 2023-10-04 code/L2/L2_TropicalAndes_plant_taxcoverage.Rmd
update name

--- 2023-09-27 code/L2/L2_TropicalAndes_plant_taxonomiccoverage.Rmd
@@
+---
+title: "Taxonomic coverage for plant data"
+author: "Hazel J. Anderson"
+project: "Plant-Frugivore Diversity"
+collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
+overview: "This script summarize taxonomic coverage and gaps of plant data used."
+data input: "ADD"
+data output: "ADD"
+date: "2023-10-04"
+output: html_document
+---
+
+# Set file paths
 ```{r}
-#input trait datasets
-#compare coverage of species, families, genera by trait coverage (plot)
-#do the same for GBIF occurrence 
+data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
+```
+
+# Load required packages
+```{r}
 library(WorldFlora)
@@
 library(dplyr)
-```
-
-# Set file paths
-```{r}
-data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
-output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
 ```

### 2023-10-04 code/L1/L1_F3_TropicalAndes_frug_occ.Rmd
inital commit

---
title: "TITLE"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script subsets frugivore occurrence data to species with complete trait coverage."
data input: "ADD"
data output: "ADD"
date: "2023-10-04"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}

```

### 2023-10-04 code/L1/L1_P3_TropicalAndes_plant_traits.Rmd
inital commit

--- 2023-10-04 code/L1/L1_F3_TropicalAndes_frug_occ.Rmd
@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: "This script subsets frugivore occurrence data to species with complete trait coverage."
+overview: "This script subsets plant trait data by synonyms species list and combine."
 data input: "ADD"

### 2023-10-04 code/L2/L2_2_TropicalAndes_combineddiv.Rmd
inital commit

---
title: "Combining taxonomic and functional diversity for plants and frugivores at various spatial grains"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script combines taxonomic and functional diversity per cell for various spatial grains for plants and frugivores."
data input: "ADD"
data output: "ADD"
date: "2023-10-04"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
```

# Load required packages
```{r}

```

# Read in Data
```{r}

```

### 2023-10-04 code/L2/L2_TropicalAndes_plant_frugivore_trait_alignment.Rmd
inital commit

--- 2023-10-04 code/L2/L2_TropicalAndes_plant_taxcoverage.Rmd
@@
 ---
-title: "Taxonomic coverage for plant data"
+title: "Plant-frugivore trait alignment"
 author: "Hazel J. Anderson"
@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: "This script summarize taxonomic coverage and gaps of plant data used."
+overview: "This script looks at pairs of traits between plants and frugivores to test alignment."
 data input: "ADD"
@@
 ```{r}
-library(WorldFlora)
-library(taxize)
-library(BIEN)
-library(GIFT)
-library(dplyr)
+
 ```
@@
 ```{r}
-BIEN_traits <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_traits_forest.csv"))
-TRY_traits <- read.csv(file.path(data_path,"TropicalAndes_TRY_plant_traits_forestSpecies.csv"))
-GIFT_traits <- read.csv(file.path(data_path,"TropicalAndes_GIFT_plant_traits_forest.csv"))
-GBIF_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ.csv"))
+
 ```
-
-```{r}
-#species list
-BIEN_species <- unique(BIEN_traits$scrubbed_species_binomial)
-TRY_species <- unique(TRY_traits$SpeciesName)
-GIFT_species <- unique(GIFT_traits$species)
-GBIF_species <- unique(GBIF_occ$species)
-```
-
-# Get higher taxonomy for species lists
-```{r}
-BIEN_tax<- BIEN_taxonomy_species(BIEN_species)
-```
-
-```{r}
-TRY_tax <- tax_name(TRY_species, get = c('family', 'genus'), db= "itis")
-```
-
-```{r}
-gift_sp <- GIFT_species()
-sapply(GIFT_species, function(x) grep(x, gift_sp$work_species))
-gift_sp[sapply(GIFT_species, function(x) grep(x, gift_sp$work_species)), ]
-
-#GIFT_tax <- GIFT_taxgroup(taxon_lvl = c("family", "genus")[1], species = GIFT_species)
-```
-
-# Plot coverage
-```{r}
-BIEN_family_coverage <- BIEN_tax %>% count(scrubbed_family)
-BIEN_family_coverage 
-BIEN_genus_coverage <- BIEN_tax %>% count(scrubbed_genus) 
-BIEN_genus_coverage
-barplot(height = BIEN_family_coverage$n, names = BIEN_family_coverage$scrubbed_family, horiz=T, las=1)
-barplot(height = BIEN_genus_coverage$n, names = BIEN_genus_coverage$scrubbed_genus, horiz=T, las=1)
-```

### 2023-10-11 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
change crop to Tropical Andes hotspot, add output file type conversion - sf

--- 2023-08-03 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
@@
 ```
+## Crop to Tropical Andes hotspot
+```{r}
+# Load in shape
+TropicalAndes <- read_sf(file.path(data_path, "TA_refined.shp"), layer = "TA_refined")
+TropicalAndes
+plot(TropicalAndes$geometry)
+```
 
-# Crop moist montane to Tropical Andes
+## Crop moist montane
 ```{r}
-iucn_moist_montane_forest_crop <- crop(iucn_moist_montane_forest, study_region)
-TA_iucn_moist_montane_forest <- mask(iucn_moist_montane_forest_crop, study_region)
+iucn_moist_montane_forest_crop <- crop(iucn_moist_montane_forest, TropicalAndes)
+TA_iucn_moist_montane_forest <- mask(iucn_moist_montane_forest_crop, TropicalAndes)
 plot(TA_iucn_moist_montane_forest)
@@
 
-# Crop moist lowland to Tropical Andes
+## Crop moist lowland
 ```{r}
-iucn_moist_lowland_forest_crop <- crop(iucn_moist_lowland_forest, study_region)
-TA_iucn_moist_lowland_forest <- mask(iucn_moist_lowland_forest_crop, study_region)
+iucn_moist_lowland_forest_crop <- crop(iucn_moist_lowland_forest, TropicalAndes)
+TA_iucn_moist_lowland_forest <- mask(iucn_moist_lowland_forest_crop, TropicalAndes)
 plot(TA_iucn_moist_lowland_forest)
@@
 ```
+
+# Aggregate to 0.3
+```{r}
+TropicalAndes_IUCNHabitat_Forest0.1 <- aggregate(TA_iucn_habitat_forest, 10, fun=mean, na.rm=TRUE)
+TropicalAndes_IUCNHabitat_Forest0.1
+```
+
+```{r}
+writeRaster(TropicalAndes_IUCNHabitat_Forest0.1,file.path(output_path, "TropicalAndes_IUCNHabitat_Forest0.1.tif"), overwrite = TRUE)
+```
+
+# Forest shape conversion
+
+```{r}
+Forest_poly <- rasterToPolygons(TropicalAndes_IUCNHabitat_Forest0.1, na.rm = TRUE, dissolve = TRUE)
+Forest_poly
+plot(Forest_poly)
+```
+
+```{r}
+Forest_sf <- st_as_sf(Forest_poly)
+Forest_sf
+plot(Forest_sf)
+st_write(Forest_sf, file.path(data_path, "Forest_sf.shp"), append = TRUE)
+```

### 2023-10-11 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
add mapping code to generate study region image

@@
 
-# Crop extent - trim to Andean States
-```{r}
-# Pull in world map
-worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
-
-# Subset world map to Tropical Andes
-study_region <-  worldMap %>% filter(sovereignt == "Bolivia" | sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
-plot(study_region$geometry)
-```
 ## Crop to Tropical Andes hotspot
@@
 
-# Aggregate to 0.3
+# Aggregate to 0.1
 ```{r}
@@
 ```
+
+# Map of Tropical Andes Forested region
+```{r}
+worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
+# country subset. 
+ECpoly <- worldMap %>% filter(sovereignt == "Ecuador")
+VEpoly <- worldMap %>% filter(sovereignt == "Venezuela")
+COpoly <- worldMap %>% filter(sovereignt == "Colombia")
+PEpoly <- worldMap %>% filter(sovereignt == "Peru")
+BOpoly <- worldMap %>% filter(sovereignt == "Bolivia")
+#polygon of Tropical Andes
+TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
+# set colors
+Andean_states <- "lightgrey"
+Tropical_Andes <- "grey"
+Forest <- "forestgreen"
+```
+
+```{r}
+MapPlot <-
+ggplot() +
+geom_sf(data = worldMap, fill = "white") +
+geom_sf(data = ECpoly, fill = Andean_states) +
+geom_sf(data = VEpoly, fill = Andean_states) +
+geom_sf(data = COpoly, fill = Andean_states) +
+geom_sf(data = PEpoly, fill = Andean_states) +
+geom_sf(data = BOpoly, fill = Andean_states) +
+geom_sf(data = TropicalAndes, fill = Tropical_Andes) +
+geom_sf(data = Forest_sf, fill = Forest) +
+labs(title = "Tropical Andes IUCN Montane and Lowland forest", x = "Latitude", y = "Longitude") +
+coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
+annotation_scale(location = "bl", width_hint = 0.5) +
+annotation_north_arrow(location = "bl", which_north = "true", 
+  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
+theme(panel.background = element_rect(fill = "lightblue"))
+
+MapPlot
+```
+
+```{r}
+ggsave(filename = "Tropical Andes Forest Map.png", dpi = 300, device="png")
+```

### 2023-10-17 code/L0/L0_TropicalAndes_frugivores_GBIF.Rmd
add subset to Tropical Andes habitat shape code, update object names

@@
 ```{r}
-d <- occ_download_get("0129743-230530130749713", path = output_path, overwrite = TRUE) %>%
+d <- occ_download_get("0009771-231002084531237", path = output_path, overwrite = TRUE) %>%
   occ_download_import()
+```
+
+# Subset occurrence to Tropical Andes IUCN Montane and Lowland forest
+
+## Read in Forest shape
+```{r}
+TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path, "Forest_sf.shp"), layer = "Forest_sf")
+```
+
+## Convert data frame to sf object
+```{r}
+TropicalAndes_frugivore_occ <- st_as_sf(x = d, 
+                        coords = c("decimalLongitude", "decimalLatitude"),
+                        crs = "+proj=longlat +datum=WGS84 +no_defs")
+```
+
+# Crop occurrence data to forest extent
+```{r}
+TropicalAndes_frugivore_occ_forest <- TropicalAndes_frugivore_occ[Forest_sf,]
 ```
@@
 ```{r}
-glimpse(d)
+glimpse(TropicalAndes_frugivore_occ_forest)
 ```
@@
 print("The number of records is")
-nrow(d)
+nrow(TropicalAndes_frugivore_occ_forest)
 print("The number of species is")
-length(unique(d$species))
+length(unique(TropicalAndes_frugivore_occ_forest$species))
+print("The number of genera is" )
+length(unique(TropicalAndes_frugivore_occ_forest$genus))
 print("The number of families is")
-length(unique(d$family))
+length(unique(TropicalAndes_frugivore_occ_forest$family))
 ```
@@
 ```{r}
-GBIF_occ_SpeciesList <- unique(d$species)
+GBIF_occ_SpeciesList <- unique(TropicalAndes_frugivore_occ_forest$species)
 ```
@@
 ```{r}
-write.csv(d, file = file.path(output_path,"TropicalAndes_GBIF_frugivore_occ.csv"))
+write.csv(TropicalAndes_frugivore_occ_forest, file = file.path(output_path,"TropicalAndes_GBIF_frugivore_occ.csv"))
 

### 2023-10-17 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
add subset to Tropical Andes habitat shape code, update object names

@@
 ```{r}
-d1 <- occ_download_get("0001452-230918134249559", path = output_path, overwrite = TRUE) %>%
+d1 <- occ_download_get("0009835-231002084531237", path = output_path, overwrite = TRUE) %>%
   occ_download_import()
+```
+
+# Subset occurrence to Tropical Andes IUCN Montane and Lowland forest
+
+## Read in Forest shape
+```{r}
+TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path, "Forest_sf.shp"), layer = "Forest_sf")
+```
+
+## Convert data frame to sf object
+```{r}
+TropicalAndes_plant_occ <- st_as_sf(x = d1, 
+                        coords = c("decimalLongitude", "decimalLatitude"),
+                        crs = "+proj=longlat +datum=WGS84 +no_defs")
+```
+
+# Crop occurrence data to forest extent
+```{r}
+TropicalAndes_plant_occ_forest <- TropicalAndes_plant_occ[Forest_sf,]
 ```
@@
 ```{r}
-glimpse(d1)
+glimpse(TropicalAndes_plant_occ_forest)
 ```
@@
 print("The number of occurrences is")
-nrow(d1)
+nrow(TropicalAndes_plant_occ_forest)
 print("The number of species is" )
-length(unique(d1$species))
+length(unique(TropicalAndes_plant_occ_forest$species))
 print("The number of genera is" )
-length(unique(d1$genus))
+length(unique(TropicalAndes_plant_occ_forest$genus))
 print("The number of families is" )
-length(unique(d1$family))
+length(unique(TropicalAndes_plant_occ_forest$family))
 ```
@@
 ```{r}
-GBIF_SpeciesList <- unique(d1$species)
+GBIF_SpeciesList <- unique(TropicalAndes_plant_occ_forest$species)
 ```
@@
 ```{r}
-TropicalAndes_GBIF_path <- file.path(output_path,"TropicalAndes_GBIF_plant_occ.csv") 
-write.csv(d1, file = TropicalAndes_GBIF_path)
-TropicalAndes_GBIF_species_path <- file.path(output_path,"TropicalAndes_GBIF_plant_species.csv") 
-write.csv(GBIF_SpeciesList, file = TropicalAndes_GBIF_species_path)
+write.csv(TropicalAndes_plant_occ_forest, file.path(output_path,"TropicalAndes_GBIF_plant_occ.csv"))
+
+write.csv(GBIF_SpeciesList, file.path(output_path,"TropicalAndes_GBIF_plant_species.csv"))
 ```

### 2023-10-17 code/L0/L0_TropicalAndes_plants_BIEN_traits.Rmd
remove species list and aquire all records for desired traits

--- 2023-07-18 code/L0/L0_TropicalAndes_plants_BIEN_traits.Rmd
@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: "This script retrieves plant trait data from the BIEN database for plant species in the countries that make up the Tropical Andes."
+overview: "This script retrieves plant trait data from the BIEN database for Tropical Andean countries."
 data input: "none"
-data output: "TropicalAndes_BIEN_traits.csv"
-date: "2023-07-18"
+data output: "TropicalAndes_BIEN_traits.csv, AllDesired_BIEN_plant_traits.csv"
+date: "2023-10-17"
 output: html_document
@@
 
-# List of countries of interest
+# Retrived desired trait data from BIEN
+
+## List of desired traits
 ```{r}
-country_vector <- c("Ecuador", "Venezuela", "Peru", "Colombia", "Bolivia")
+trait_list <- c("whole plant dispersal syndrome", "whole plant height", "maximum whole plant longevity", "maximum whole plant height", "minimum whole plant height", "longest whole plant longevity", "fruit type", "maximum fruit length", "minimum fruit length", "seed mass")
 ```
 
+## List of desired countries
 ```{r}
-TropicalAndes_BIEN_traits <- BIEN_trait_country(country_vector, all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
+TA_countries <- c("Bolivia", "Ecuador", "Venezuela", "Colombia", "Peru")
+```
+
+## Retrive all BIEN records for desired traits
+```{r}
+AllDesired_BIEN_plant_traits <- BIEN_trait_trait(trait_list, all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
+```
+
+## Retrive all BIEN records for desired traits in Tropical Andes countries
+```{r}
+TropicalAndes_BIEN_plant_traits <- BIEN_trait_country(TA_countries, trait_list, all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
 ```
@@
 # Data summary
+
+## Desired traits
 ```{r}
-glimpse(TropicalAndes_BIEN_traits)
+glimpse(AllDesired_BIEN_plant_traits)
 ```
@@
 print("The number of records is")
-nrow(TropicalAndes_BIEN_traits)
+nrow(AllDesired_BIEN_plant_traits)
 print("The number of species is")
-length(unique(TropicalAndes_BIEN_traits$scrubbed_species_binomial))
+length(unique(AllDesired_BIEN_plant_traits$scrubbed_species_binomial))
+print("The number of genera is" )
+length(unique(AllDesired_BIEN_plant_traits$scrubbed_genus))
 print("The number of families is")
-length(unique(TropicalAndes_BIEN_traits$family_matched))
+length(unique(AllDesired_BIEN_plant_traits$scrubbed_family))
+```
+
+```{r}
+AllDesired_BIEN_plant_traits %>% count(trait_name)
+```
+
+## Desired traits in Tropical Andes countries
+```{r}
+glimpse(TropicalAndes_BIEN_plant_traits)
+```
+
+```{r}
+print("The number of records is")
+nrow(TropicalAndes_BIEN_plant_traits)
+print("The number of species is")
+length(unique(TropicalAndes_BIEN_plant_traits$scrubbed_species_binomial))
+print("The number of genera is" )
+length(unique(TropicalAndes_BIEN_plant_traits$scrubbed_genus))
+print("The number of families is")
+length(unique(TropicalAndes_BIEN_plant_traits$scrubbed_family))
+```
+
+```{r}
+TropicalAndes_BIEN_plant_traits %>% count(trait_name)
 ```
@@
 ```{r}
-TropicalAndes_BIEN_traits_path <- file.path(output_path,"TropicalAndes_BIEN_plant_traits.csv") 
-write.csv(TropicalAndes_BIEN_traits, file = TropicalAndes_BIEN_traits_path)
+write.csv(AllDesired_BIEN_plant_traits, file.path(output_path,"AllDesired_BIEN_plant_traits.csv"))
+write.csv(TropicalAndes_BIEN_plant_traits, file.path(output_path,"TropicalAndes_BIEN_plant_traits.csv"))
 ```

### 2023-10-18 code/L0/L0_TropicalAndes_plants_GIFT.Rmd
standardizing code file naming
remove species list, add more traits

--- 2023-08-04 code/L0/L0_TropicalAndes_plants_GIFT_traits.Rmd
@@
 
-# Load in species list
-```{r}
-species_list <- read.csv("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1/TropicalAndes_plant_Forest_SpeciesList.csv")
-species_list <- species_list$x
-```
-
 # Retrive GIFT data
@@
 ```
-Traits of interest: plant height (ID = 1.6.3), lifespan (ID = 2.2.1), dispersal syndrome (ID = 3.3)
+Traits of interest: plant height (ID = 1.6.3), lifespan (ID = 2.2.1), dispersal syndrome (ID = 3.3), Plant_height_min (1.6.1), Plant_height_max (1.6.2), Fruit_type_1 (3.16.1), fruit_dryness_1 (3.18.1), Fruit_length_min (3.13.1), Fruit_length_max (3.13.2), Fruit_length_mean (3.13.3), Seed_mass_min (3.2.1), Seed_mass_max (3.2.2), Seed_mass_mean (3.2.3), Seed_length_min (3.10.1), Seed_length_max (3.10.2), Seed_length_mean (3.10.3), Seed_width_min (3.11.1), Seed_width_max (3.11.2), Seed_width_mean (3.11.3)
 
@@
 # 
-height <- GIFT_traits(trait_IDs = c("1.6.2"))
+height <- GIFT_traits(trait_IDs = c("1.6.2", "1.6.1", "1.6.3"))
 lifespan <- GIFT_traits(trait_IDs = c("2.2.1"))
 dispersal_syndrome <- GIFT_traits(trait_IDs = c("3.3.1", "3.3.2"))
+fruit_type <- GIFT_traits(trait_IDs = c("3.16.1", "3.18.1"))
+fruit_length <- GIFT_traits(trait_IDs = c("3.13.1", "3.13.2", "3.13.3"))
+seed_mass <- GIFT_traits(trait_IDs = c("3.2.1", "3.2.2", "3.2.3"))
+seed_length <- GIFT_traits(trait_IDs = c("3.10.1", "3.10.2", "3.10.3"))
+seed_width <- GIFT_traits(trait_IDs = c("3.11.1", "3.11.2", "3.11.3"))
 ```
@@
 
+```{r}
+GIFT_traits %>% count(trait_name)
+```
+
 # Write data to csv

### 2023-10-18 code/L0/L0_TropicalAndes_plants_checklist.Rmd
inital commit, add read in shapefile, and checklist code

---
title: "Tropical Andes plant checklist using GIFT"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script retrieves plant checklist data from the GIFT database for Tropical Andes Forest region."
data input: "Forest_sf.sph"
data output: ""
date: "2023-10-17"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(GIFT)
library(dplyr)
library(sf)
```

# Load Tropical Andes Forest shape
```{r}
TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path, "Forest_sf.shp"), layer = "Forest_sf")
```

```{r}
checklist <- GIFT_checklists(taxon_name = "Angiospermae", taxonomic_group = TRUE, floristic_group = "all", shp = TropicalAndes_IUCNHabitat_Forest, overlap = "extent_intersect")
```

### 2023-10-18 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
update forest subset file type -> switch to sf

@@
 TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_frugivore_occ_forest.csv"))
-TropicalAndes_IUCNHabitat_Forest <- raster(file.path(data_path2, "TropicalAndes_IUCNHabitat_Forest.tif"))
+TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path2, "Forest_sf.shp"), layer = "Forest_sf")
 frugivore_traits <- read.csv(file.path(data_path,"TropicalAndes_Frugivoria_traits_Forest.csv"))
@@
 ```{r}
-coordinates_sf_plant <- st_as_sf(TropicalAndes_plant_occ_forest_sub, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(Forest_sf))
-coordinates_sf_frugivore <- st_as_sf(TropicalAndes_frugivore_occ_forest_sub, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(Forest_sf))
+coordinates_sf_plant <- st_as_sf(TropicalAndes_plant_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(TropicalAndes_IUCNHabitat_Forest))
+coordinates_sf_frugivore <- st_as_sf(TropicalAndes_frugivore_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(TropicalAndes_IUCNHabitat_Forest))
 ```
@@
 # Perform the spatial intersection
-intersections_plant_TA <- st_intersection(coordinates_sf_plant, Forest_sf)
-intersections_frugivore_TA <- st_intersection(coordinates_sf_frugivore, Forest_sf)
+intersections_plant_TA <- st_intersection(coordinates_sf_plant, TropicalAndes_IUCNHabitat_Forest)
+intersections_frugivore_TA <- st_intersection(coordinates_sf_frugivore, TropicalAndes_IUCNHabitat_Forest)
 # Convert the intersected sf object back to a dataframe
@@
 frugivore_traits_df_final <-as.data.frame(frugivore_traits_matrix)
-plant_traits_df$X <-NULL
-frugivore_traits_df$X <-NULL
+plant_traits_df_final$X <-NULL
+frugivore_traits_df_final$X <-NULL
 ```

### 2023-10-18 code/L0/L0_TropicalAndes_plants_GIFT.Rmd
seperate traits, update combine code

@@
 # 
-height <- GIFT_traits(trait_IDs = c("1.6.2", "1.6.1", "1.6.3"))
+height_mean <- GIFT_traits(trait_IDs = c("1.6.3"))
+height_min <- GIFT_traits(trait_IDs = c("1.6.1"))
+height_max <- GIFT_traits(trait_IDs = c("1.6.2"))
+
 lifespan <- GIFT_traits(trait_IDs = c("2.2.1"))
-dispersal_syndrome <- GIFT_traits(trait_IDs = c("3.3.1", "3.3.2"))
-fruit_type <- GIFT_traits(trait_IDs = c("3.16.1", "3.18.1"))
-fruit_length <- GIFT_traits(trait_IDs = c("3.13.1", "3.13.2", "3.13.3"))
-seed_mass <- GIFT_traits(trait_IDs = c("3.2.1", "3.2.2", "3.2.3"))
-seed_length <- GIFT_traits(trait_IDs = c("3.10.1", "3.10.2", "3.10.3"))
-seed_width <- GIFT_traits(trait_IDs = c("3.11.1", "3.11.2", "3.11.3"))
-```
 
-```{r}
-# filter by species list
-height_sub <- height %>%
-  filter(work_species %in% species_list)
-dispersal_syndrome_sub <- dispersal_syndrome %>%
-  filter(work_species %in% species_list)
-lifespan_sub <- lifespan %>%
-  filter(work_species %in% species_list)
+dispersal_syndrome_1 <- GIFT_traits(trait_IDs = c("3.3.1"))
+dispersal_syndrome_2 <- GIFT_traits(trait_IDs = c("3.3.2"))
+
+fruit_type <- GIFT_traits(trait_IDs = c("3.16.1"))
+fruit_dryness <- GIFT_traits(trait_IDs = c("3.18.1"))
+
+fruit_length_mean <- GIFT_traits(trait_IDs = c("3.13.3"))
+fruit_length_min <- GIFT_traits(trait_IDs = c("3.13.1"))
+fruit_length_max <- GIFT_traits(trait_IDs = c("3.13.2"))
+
+seed_mass_mean <- GIFT_traits(trait_IDs = c("3.2.3"))
+seed_mass_min <- GIFT_traits(trait_IDs = c("3.2.1"))
+seed_mass_max <- GIFT_traits(trait_IDs = c("3.2.2"))
+
+seed_length_mean <- GIFT_traits(trait_IDs = c("3.10.3"))
+seed_length_min <- GIFT_traits(trait_IDs = c("3.10.1"))
+seed_length_max <- GIFT_traits(trait_IDs = c("3.10.2"))
+
+seed_width_mean <- GIFT_traits(trait_IDs = c("3.11.3"))
+seed_width_min <- GIFT_traits(trait_IDs = c("3.11.1"))
+seed_width_max <- GIFT_traits(trait_IDs = c("3.11.2"))
 ```
@@
 #rename columns to trait name
-height_sub <- height_sub %>% rename(trait_value = trait_value_1.6.2, species = work_species)
-dispersal_syndrome_sub <- dispersal_syndrome_sub %>% rename(trait_value = trait_value_3.3.1, species = work_species)
-lifespan_sub <- lifespan_sub %>% rename(trait_value = trait_value_2.2.1, species = work_species)
+height_mean <- height_mean %>% rename(trait_value = trait_value_1.6.3, species = work_species)
+height_min <- height_min %>% rename(trait_value = trait_value_1.6.1, species = work_species)
+height_max <- height_max %>% rename(trait_value = trait_value_1.6.2, species = work_species)
+
+lifespan <- lifespan %>% rename(trait_value = trait_value_2.2.1, species = work_species)
+
+dispersal_syndrome_1 <- dispersal_syndrome_1 %>% rename(trait_value = trait_value_3.3.1, species = work_species)
+dispersal_syndrome_2 <- dispersal_syndrome_2 %>% rename(trait_value = trait_value_3.3.2, species = work_species)
+
+fruit_type <- fruit_type %>% rename(trait_value = trait_value_3.16.1, species = work_species)
+fruit_dryness <- fruit_dryness %>% rename(trait_value = trait_value_3.18.1, species = work_species)
+
+fruit_length_mean <- fruit_length_mean %>% rename(trait_value = trait_value_3.13.3, species = work_species)
+fruit_length_min <- fruit_length_min %>% rename(trait_value = trait_value_3.13.1, species = work_species)
+fruit_length_max <- fruit_length_max %>% rename(trait_value = trait_value_3.13.2, species = work_species)
+
+seed_mass_mean <- seed_mass_mean %>% rename(trait_value = trait_value_3.2.3, species = work_species)
+seed_mass_min <- seed_mass_min %>% rename(trait_value = trait_value_3.2.1, species = work_species)
+seed_mass_max <- seed_mass_max %>% rename(trait_value = trait_value_3.2.2, species = work_species)
+
+seed_length_mean <- seed_length_mean %>% rename(trait_value = trait_value_3.10.3, species = work_species)
+seed_length_min <- seed_length_min %>% rename(trait_value = trait_value_3.10.1, species = work_species)
+seed_length_max <- seed_length_max %>% rename(trait_value = trait_value_3.10.2, species = work_species)
+
+seed_width_mean <- seed_width_mean %>% rename(trait_value = trait_value_3.11.3, species = work_species)
+seed_width_min <- seed_width_min %>% rename(trait_value = trait_value_3.11.1, species = work_species)
+seed_width_max <- seed_width_max %>% rename(trait_value = trait_value_3.11.2, species = work_species)
 ```
@@
 #keep just species and trait columns
-height_sub <- height_sub %>% select(species, trait_value)
-dispersal_syndrome_sub <- dispersal_syndrome_sub %>% select(species, trait_value)
-lifespan_sub <- lifespan_sub %>% select(species, trait_value)
+height_mean <- height_mean %>% select(species, trait_value)
+height_min <- height_min %>% select(species, trait_value)
+height_max <- height_max %>% select(species, trait_value)
+
+lifespan <- lifespan %>% select(species, trait_value)
+
+dispersal_syndrome_1 <- dispersal_syndrome_1 %>% select(species, trait_value)
+dispersal_syndrome_2 <- dispersal_syndrome_2 %>% select(species, trait_value)
+
+fruit_type <- fruit_type %>% select(species, trait_value)
+fruit_dryness <- fruit_dryness %>% select(species, trait_value)
+
+fruit_length_mean <- fruit_length_mean %>% select(species, trait_value)
+fruit_length_min <- fruit_length_min %>% select(species, trait_value)
+fruit_length_max <- fruit_length_max %>% select(species, trait_value)
+
+seed_mass_mean <- seed_mass_mean %>% select(species, trait_value)
+seed_mass_min <- seed_mass_min %>% select(species, trait_value)
+seed_mass_max <- seed_mass_max %>% select(species, trait_value)
+
+seed_length_mean <- seed_length_mean %>% select(species, trait_value)
+seed_length_min <- seed_length_min %>% select(species, trait_value)
+seed_length_max <- seed_length_max %>% select(species, trait_value)
+
+seed_width_mean <- seed_width_mean %>% select(species, trait_value)
+seed_width_min <- seed_width_min %>% select(species, trait_value)
+seed_width_max <- seed_width_max %>% select(species, trait_value)
 ```
@@
 # add trait name column
-height_sub$trait_name <- "plant_height"
-dispersal_syndrome_sub$trait_name <- "dispersal_syndrome"
-lifespan_sub$trait_name <- "plant_lifespan"
+height_mean$trait_name <- "plant_height"
+height_min$trait_name <- "plant_height_min"
+height_max$trait_name <- "plant_height_max"
+lifespan$trait_name <- "plant_lifespan"
+dispersal_syndrome_1$trait_name <- "dispersal_syndrome_1"
+dispersal_syndrome_2$trait_name <- "dispersal_syndrome_2"
+fruit_type$trait_name <- "fruit_type"
+fruit_dryness$trait_name <- "fruit_dryness"
+fruit_length_mean$trait_name <- "fruit_length_mean"
+fruit_length_min$trait_name <- "fruit_length_min"
+fruit_length_max$trait_name <- "fruit_length_max"
+seed_mass_mean$trait_name <- "seed_mass_mean"
+seed_mass_min$trait_name <- "seed_mass_min"
+seed_mass_max$trait_name <- "seed_mass_max"
+seed_length_mean$trait_name <- "seed_length_mean"
+seed_length_min$trait_name <- "seed_length_min"
+seed_length_max$trait_name <- "seed_length_max"
+seed_width_mean$trait_name <- "seed_width_mean"
+seed_width_min$trait_name <- "seed_width_min"
+seed_width_max$trait_name <- "seed_width_max"
+
 ```
@@
 ```{r}
-#convert numeric trait values to character
-height_sub$trait_value <- as.character(height_sub$trait_value)
-lifespan_sub$trait_value <- as.character(lifespan_sub$trait_value)
+# convert numberic traits into characters so we can combine into one dataframe
+height_mean$trait_value <- as.character(height_mean$trait_value)
+height_min$trait_value <- as.character(height_min$trait_value)
+height_max$trait_value <- as.character(height_max$trait_value)
+lifespan$trait_value <- as.character(lifespan$trait_value)
+fruit_length_mean$trait_value <- as.character(fruit_length_mean$trait_value)
+fruit_length_min$trait_value <- as.character(fruit_length_min$trait_value)
+fruit_length_max$trait_value <- as.character(fruit_length_max$trait_value)
+seed_mass_mean$trait_value <- as.character(seed_mass_mean$trait_value)
+seed_mass_min$trait_value <- as.character(seed_mass_min$trait_value)
+seed_mass_max$trait_value <- as.character(seed_mass_max$trait_value)
+seed_length_mean$trait_value <- as.character(seed_length_mean$trait_value)
+seed_length_min$trait_value <- as.character(seed_length_min$trait_value)
+seed_length_max$trait_value <- as.character(seed_length_max$trait_value)
+seed_width_mean$trait_value <- as.character(seed_width_mean$trait_value)
+seed_width_min$trait_value <- as.character(seed_width_min$trait_value)
+seed_width_max$trait_value <- as.character(seed_width_max$trait_value)
 ```
@@
 #combine all trait dataframes into one long table
-GIFT_traits <- dplyr::bind_rows(height_sub, dispersal_syndrome_sub, lifespan_sub)
+GIFT_traits <- dplyr::bind_rows(height_mean, height_min, height_max, lifespan, dispersal_syndrome_1, dispersal_syndrome_2, fruit_type, fruit_dryness, fruit_length_mean, fruit_length_min, fruit_length_max, seed_mass_mean, seed_mass_min, seed_mass_max, seed_length_mean, seed_length_min, seed_length_max, seed_width_mean, seed_width_min, seed_width_max)
 ```
@@
 ```{r}
-write.csv(GIFT_traits, file = file.path(output_path,"TropicalAndes_GIFT_plant_traits_forest.csv"))
+write.csv(GIFT_traits, file = file.path(output_path,"TropicalAndes_GIFT_plant_traits.csv"))
 ```

### 2023-10-18 code/L0/L0_TropicalAndes_plants_checklist.Rmd
add summary and save to csv code

@@
 ```
+
+```{r}
+print("The number of species is")
+length(unique(checklist$checklists$work_species))
+print("The number of families is")
+length(unique(checklist$checklists$family))
+```
+
+# Write data to csv
+```{r}
+write.csv(checklist$checklists, file = file.path(output_path,"TropicalAndes_plant_checklist.csv"))
+```

### 2023-10-18 code/L2/funcDiv_troubleshooting.rmd
inital commit

--- 2023-10-04 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
@@
----
-title: "Tropical Andes Functional diversity for plants and Frugivores"
-author: "Hazel J. Anderson"
-project: "Plant-Frugivore Diversity"
-collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: ""
-data input: ""
-data output: ""
-date: "2023-08-03"
-output: html_document
----
-
-# Load required packages
-```{r}
-library(sf)
-library(dplyr)
-library(ggplot2)
-library(rnaturalearth)
-library(smoothr)
-library(purrr)
-library(raster)
-library(scico)
-library(ggspatial)
-library(letsR)
-library(mFD)
-```
-
-# Set file paths
-```{r}
-data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
-data_path2<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
-output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
-```
-
-# Read in Data
-```{r}
-TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_plant_occ_forest.csv"))
-TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_frugivore_occ_forest.csv"))
-TropicalAndes_IUCNHabitat_Forest <- raster(file.path(data_path2, "TropicalAndes_IUCNHabitat_Forest.tif"))
-frugivore_traits <- read.csv(file.path(data_path,"TropicalAndes_Frugivoria_traits_Forest.csv"))
-plant_traits <- read.csv(file.path(data_path,"TropicalAndes_plant_traits_forest.csv"))
-```
-
-# Create Presence-Absense matrix
-```{r}
-x <- TropicalAndes_plant_occ_forest$decimalLongitude
-y <- TropicalAndes_plant_occ_forest$decimalLatitude
-plant_xy <- cbind(x, y)
-plant_species <- TropicalAndes_plant_occ_forest$species
-plants_PAM <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14)
-summary(plants_PAM)
-plot(plants_PAM, xlab = "Longitude", ylab = "Latitude", main = "Plant richness map")
-plants_PAM_matrix <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE)
-```
-
-```{r}
-x <- TropicalAndes_frugivore_occ_forest$decimalLongitude
-y <- TropicalAndes_frugivore_occ_forest$decimalLatitude
-frugivore_xy <- cbind(x, y)
-frugivore_species <- TropicalAndes_frugivore_occ_forest$species
-frugivore_PAM <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14)
-summary(frugivore_PAM)
-plot(frugivore_PAM, xlab = "Longitude", ylab = "Latitude", main = "Frugivore richness map")
-frugivore_PAM_matrix <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE)
-```
-
-# Prepping data for functional diversity calculations
-
-```{r}
-plant_traits <- na.omit(plant_traits)
-frugivore_traits <- na.omit(frugivore_traits)
-frugivore_traits$body_mass_e <- as.numeric(frugivore_traits$body_mass_e)
-frugivore_traits$body_size_mm <- as.numeric(frugivore_traits$body_size_mm)
-plant_traits <- plant_traits[, c("species", "plant_height", "dispersal_syndrome", "plant_lifespan")]
-frugivore_traits <- frugivore_traits[, c("IUCN_species_name", "diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
-```
-
-```{r}
-coordinates_sf_plant <- st_as_sf(TropicalAndes_plant_occ_forest_sub, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(Forest_sf))
-coordinates_sf_frugivore <- st_as_sf(TropicalAndes_frugivore_occ_forest_sub, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(Forest_sf))
-```
-
-```{r}
-# Perform the spatial intersection
-intersections_plant_TA <- st_intersection(coordinates_sf_plant, Forest_sf)
-intersections_frugivore_TA <- st_intersection(coordinates_sf_frugivore, Forest_sf)
-# Convert the intersected sf object back to a dataframe
-coordinates_df_subset_plant <- as.data.frame(intersections_plant_TA)
-coordinates_df_subset_frugivore <- as.data.frame(intersections_frugivore_TA)
-```
-
-```{r}
-# Print the resulting subset of coordinates
-print(coordinates_df_subset_plant)
-print(coordinates_df_subset_frugivore)
-# Convert the point geometry to a dataframe of longitude and latitude
-subset_plant <- as.data.frame(st_coordinates(intersections_plant_TA))
-subset_frugivore <- as.data.frame(st_coordinates(intersections_frugivore_TA))
-
-# Match the coordinates to the original dataframe
-#merge the subset and full dataframe together to get final TA dataset
-plant_PAM_filter <- merge(plants_PAM_matrix, subset_plant, by.x = c("Longitude(x)", "Latitude(y)"), by.y = c("X", "Y"))
-frugivore_PAM_filter <- merge(frugivore_PAM_matrix, subset_frugivore, by.x = c("Longitude(x)", "Latitude(y)"), by.y = c("X", "Y"))
-```
-
-```{r}
-# Turn PAM into matrix
-plant_PAM <- as.matrix(plant_PAM_filter)
-frugivore_PAM <- as.matrix(frugivore_PAM_filter)
-
-# Add rownames
-# Get the number of rows in the matrix
-num_rows_plant <- nrow(plant_PAM)
-num_rows_frugivore <- nrow(frugivore_PAM)
-# Generate unique row names
-row_names_plant <- paste0("cell", 1:num_rows_plant)
-row_names_frugivore <- paste0("cell", 1:num_rows_frugivore)
-# Assign row names to the matrix
-rownames(plant_PAM) <- row_names_plant
-rownames(frugivore_PAM) <- row_names_frugivore
-# Print the matrix with row names
-print(plant_PAM)
-column_names_plant <- colnames(plant_PAM)
-clean_column_names_plant <- gsub("_", " ", column_names_plant)
-print(frugivore_PAM)
-column_names_frugivore <- colnames(frugivore_PAM)
-clean_column_names_frugivore <- gsub("_", " ", column_names_frugivore)
-#Insert clean column names
-colnames(plant_PAM) <- clean_column_names_plant
-colnames(frugivore_PAM) <- clean_column_names_frugivore
-```
-
-```{r}
-#turn data into correct data types for inputs into the trait categories dataframe
-frugivore_traits$diet_cat <- as.factor(frugivore_traits$diet_cat)
-frugivore_traits$body_mass_e <- as.numeric(frugivore_traits$body_mass_e)
-frugivore_traits$body_size_mm <- as.numeric(frugivore_traits$body_size_mm)
-frugivore_traits$generation_time <- as.numeric(frugivore_traits$generation_time)
-
-plant_traits$plant_height <- as.numeric(plant_traits$plant_height)
-plant_traits$plant_lifespan<- as.numeric(plant_traits$plant_lifespan)
-plant_traits$dispersal_syndrome <- as.factor(plant_traits$dispersal_syndrome)
-
-# Remove the species from PAM that have no occurrences anymore after subsetting to TA
-# Remove columns with sum equal to zero
-PAM_plant_site_final <- plant_PAM[, colSums(plant_PAM) != 0]
-PAM_frugivore_site_final <- frugivore_PAM[, colSums(frugivore_PAM) != 0]
-#Save coordinates for later
-site_loc_key_plant <- PAM_plant_site_final[,1:2]
-site_loc_key_frugivore <- PAM_frugivore_site_final[,1:2]
-
-columns_to_remove <- c(1,2)
-PAM_plant_site_final <- PAM_plant_site_final[,-columns_to_remove]
-PAM_frugivore_site_final <- PAM_frugivore_site_final[,-columns_to_remove]
-colnames_plant <- colnames(PAM_plant_site_final)
-colnames_frugivore <- colnames(PAM_frugivore_site_final)
-```
-
-```{r}
-# Remove species names from trait matrix not in the PAM
-plant_traits_df_subset <- plant_traits %>% filter(species %in% colnames_plant)
-frugivore_traits_df_subset <- frugivore_traits %>% filter(IUCN_species_name %in% colnames_frugivore)
-
-# Turn trait dataframe into a matrix
-class(plant_traits)
-class(frugivore_traits)
-plant_traits_matrix <- as.matrix(plant_traits_df_subset)
-frugivore_traits_matrix <- as.matrix(frugivore_traits_df_subset)
-
-# Define row names as species names
-row_names_plant <- plant_traits_matrix[,1]
-row_names_frugivore <- frugivore_traits_matrix[,1]
-
-# Assign row names to the matrix
-rownames(plant_traits_matrix) <- row_names_plant
-rownames(frugivore_traits_matrix) <- row_names_frugivore
-# Turn back into dataframe
-plant_traits_df_final <-as.data.frame(plant_traits_matrix)
-frugivore_traits_df_final <-as.data.frame(frugivore_traits_matrix)
-plant_traits_df$X <-NULL
-frugivore_traits_df$X <-NULL
-```
-
-```{r}
-#fix types
-frugivore_traits_df_final$diet_cat <- as.factor(frugivore_traits_df_final$diet_cat)
-frugivore_traits_df_final$body_mass_e <- as.numeric(frugivore_traits_df_final$body_mass_e)
-frugivore_traits_df_final$body_size_mm<- as.numeric(frugivore_traits_df_final$body_size_mm)
-frugivore_traits_df_final$generation_time <- as.numeric(frugivore_traits_df_final$generation_time)
-
-plant_traits_df_final$plant_height <- as.numeric(plant_traits_df_final$plant_height)
-plant_traits_df_final$plant_lifespan<- as.numeric(plant_traits_df_final$plant_lifespan)
-plant_traits_df_final$dispersal_syndrome <- as.factor(plant_traits_df_final$dispersal_syndrome)
-
-# remove duplicate species name column
-plant_traits_df_final <- plant_traits_df_final[, c("plant_height", "dispersal_syndrome", "plant_lifespan")]
-frugivore_traits_df_final <- frugivore_traits_df_final[, c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
-```
-
-# Create trait type table
-```{r}
-trait_name <- c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")
-trait_type <- c("N", "Q", "Q", "Q")
-frug_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
-trait_name <- c("dispersal_syndrome", "plant_height", "plant_lifespan")
-trait_type <- c("N", "Q", "Q") 
-plant_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
-```
-
-```{r}
-# Summary of the assemblages * species dataframe:
-asb_sp_plant_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_plant_site_final)
-asb_sp_frugivore_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_frugivore_site_final)
-```
-
-```{r}
-# Species traits summary:
-plant_traits_summ <- mFD::sp.tr.summary(
-  tr_cat     = plant_trait_cat,   
-  sp_tr      = plant_traits_df_final, 
-  stop_if_NA = TRUE)
-frugivore_traits_summ <- mFD::sp.tr.summary(
-  tr_cat     = frug_trait_cat,   
-  sp_tr      = frugivore_traits_df_final, 
-  stop_if_NA = TRUE)
-```
-
-# Estimate functional trait-based distances between species
-```{r}
-
-sp_dist_plants <- mFD::funct.dist(
-  sp_tr         = plant_traits_df_final,
-  tr_cat        = plant_trait_cat,
-  metric        = "gower",
-  scale_euclid  = "scale_center",
-  ordinal_var   = "classic",
-  weight_type   = "equal",
-  stop_if_NA    = TRUE)
-```
-
-```{r}
-sp_dist_frugivore <- mFD::funct.dist(
-  sp_tr         = frugivore_traits_df_final,
-  tr_cat        = frug_trait_cat,
-  metric        = "gower",
-  scale_euclid  = "scale_center",
-  ordinal_var   = "classic",
-  weight_type   = "equal",
-  stop_if_NA    = TRUE)
-```
-
-# Generate a multidimensional space
-```{r}
-fspaces_quality_plants <- mFD::quality.fspaces(
-  sp_dist             = sp_dist_plants,
-  maxdim_pcoa         = 10,
-  deviation_weighting = "absolute",
-  fdist_scaling       = FALSE,
-  fdendro             = "average")
-```
-
-```{r}
-fspaces_quality_frugivore <- mFD::quality.fspaces(
-  sp_dist             = sp_dist_frugivore,
-  maxdim_pcoa         = 10,
-  deviation_weighting = "absolute",
-  fdist_scaling       = FALSE,
-  fdendro             = "average")
-```
-
-```{r}
-# Look at the quality spaces only (MAD index looks at mean absolute deviation from the dissimilary matrix; want the deviation
-# to be low, meaning that the true distances have been retained in the PCA)
-round(fspaces_quality_plants$"quality_fspaces", 3)            
-round(fspaces_quality_frugivore$"quality_fspaces", 3)     
-# Plot the quality spaces (chose to look at 3D, 4D, and 5D since they had the lowest MAD). Will go with 4 dimensions.
-mFD::quality.fspaces.plot(
-  fspaces_quality            = fspaces_quality_plants,
-  quality_metric             = "mad",
-  fspaces_plot               = c("pcoa_2d", "pcoa_3d", 
-                                 "pcoa_4d"),
-  name_file                  = NULL,
-  range_dist                 = NULL,
-  range_dev                  = NULL,
-  range_qdev                 = NULL,
-  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
-  gradient_deviation_quality = c(low = "yellow", high = "red"),
-  x_lab                      = "Trait-based distance")
-mFD::quality.fspaces.plot(
-  fspaces_quality            = fspaces_quality_frugivore,
-  quality_metric             = "mad",
-  fspaces_plot               = c("pcoa_3d", "pcoa_4d", 
-                                 "pcoa_5d"),
-  name_file                  = NULL,
-  range_dist                 = NULL,
-  range_dev                  = NULL,
-  range_qdev                 = NULL,
-  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
-  gradient_deviation_quality = c(low = "yellow", high = "red"),
-  x_lab                      = "Trait-based distance")
-```
-
-```{r}
-#testing correlation between functional axes and traits
-sp_faxes_coord_plants <- fspaces_quality_plants$"details_fspaces"$"sp_pc_coord"
-sp_faxes_coord_frugivore <- fspaces_quality_frugivore$"details_fspaces"$"sp_pc_coord"
-# View the components of the PCA axes *Remember the firt few components explain the most variation in dissimilarity. Clusters into groups
-# Computes linear model for continuous traits and Kruskall-Wallis tests for other types. 
-plant_tr_faxes <- mFD::traits.faxes.cor(
-  sp_tr          = plant_traits_df_final, 
-  sp_faxes_coord = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")], 
-  plot           = TRUE)
-frugivore_tr_faxes <- mFD::traits.faxes.cor(
-  sp_tr          = frugivore_traits_df_final, 
-  sp_faxes_coord = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")], 
-  plot           = TRUE)
-# Print traits with significant effect:
-plant_tr_faxes$"tr_faxes_stat"[which(plant_tr_faxes$"tr_faxes_stat"$"p.value" < 0.05), ]
-frugivore_tr_faxes$"tr_faxes_stat"[which(frugivore_tr_faxes$"tr_faxes_stat"$"p.value" < 0.05), ]
-# Return plots:
-plant_tr_faxes$"tr_faxes_plot"
-frugivore_tr_faxes$"tr_faxes_plot"
-```
-
-```{r}
-#plotting functional space
-sp_faxes_coord_plants <- fspaces_quality_plants$"details_fspaces"$"sp_pc_coord"
-sp_faxes_coord_frugivore <- fspaces_quality_frugivore$"details_fspaces"$"sp_pc_coord"
-big_plot_plants <- mFD::funct.space.plot(
-  sp_faxes_coord  = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")],
-  faxes           = c("PC1", "PC2", "PC3", "PC4"),
-  name_file       = NULL,
-  faxes_nm        = NULL,
-  range_faxes     = c(NA, NA),
-  color_bg        = "grey95",
-  color_pool      = "darkgreen",
-  fill_pool       = "white",
-  shape_pool      = 21,
-  size_pool       = 1,
-  plot_ch         = TRUE,
-  color_ch        = "black",
-  fill_ch         = "white",
-  alpha_ch        = 0.5,
-  plot_vertices   = TRUE,
-  color_vert      = "blueviolet",
-  fill_vert       = "blueviolet",
-  shape_vert      = 6,
-  size_vert       = 1,
-  plot_sp_nm      = NULL,
-  nm_size         = 3,
-  nm_color        = "black",
-  nm_fontface     = "plain",
-  check_input     = TRUE)
-big_plot_frugivore <- mFD::funct.space.plot(
-  sp_faxes_coord  = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")],
-  faxes           = c("PC1", "PC2", "PC3", "PC4"),
-  name_file       = NULL,
-  faxes_nm        = NULL,
-  range_faxes     = c(NA, NA),
-  color_bg        = "grey95",
-  color_pool      = "darkgreen",
-  fill_pool       = "white",
-  shape_pool      = 21,
-  size_pool       = 1,
-  plot_ch         = TRUE,
-  color_ch        = "black",
-  fill_ch         = "white",
-  alpha_ch        = 0.5,
-  plot_vertices   = TRUE,
-  color_vert      = "blueviolet",
-  fill_vert       = "blueviolet",
-  shape_vert      = 6,
-  size_vert       = 1,
-  plot_sp_nm      = NULL,
-  nm_size         = 3,
-  nm_color        = "black",
-  nm_fontface     = "plain",
-  check_input     = TRUE)
-#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
-# Calculate row sums
-row_sums_plant <- rowSums(PAM_plant_site_final)
-subset_matrix_plant <- PAM_plant_site_final[row_sums_plant >= 4, ]
-row_sums_frugivore <- rowSums(PAM_frugivore_site_final)
-subset_matrix_frugivore <- PAM_frugivore_site_final[row_sums_frugivore >= 4, ]
-
-```
-
-```{r}
-big_plot_plants
-big_plot_frugivore
-```
+Func Div troubleshooting
 
@@
 subset_matrix_plant <- as.matrix(subset_matrix_plant)
-```
-
-```{r}
-# match frugivore names
-sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")]
-
-## check number of species names
-nrow(sp_faxes_coord_frugivore_sub) #718
-ncol(subset_matrix_frugivore) #712
-
-sp_faxes_coord_frugivore_sub_names <- row.names(sp_faxes_coord_frugivore_sub)
-subset_matrix_frugivore_names <- colnames(subset_matrix_frugivore)
-
-sp_faxes_coord_frugivore_sub_names <- chartr(".", " ", sp_faxes_coord_frugivore_sub_names)
-sp_faxes_coord_frugivore_sub <- as.data.frame(sp_faxes_coord_frugivore_sub)
-row.names(sp_faxes_coord_frugivore_sub) <- sp_faxes_coord_frugivore_sub_names
-
-sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore_sub[ which((row.names(sp_faxes_coord_frugivore_sub) %in% subset_matrix_frugivore_names)==TRUE), ]
-
-nrow(sp_faxes_coord_frugivore_sub)
-sp_faxes_coord_frugivore_sub <- as.matrix(sp_faxes_coord_frugivore_sub)
 ```
@@
 ```
-
-# Turning the FD values into a raster
-```{r}
-
-#Need to subset the coordinates made at the top to the same set of the assemblages (removed values less than 5 so these coords need to be removed to do this correctly)
-
-#Have coordinates and associated values
-
-library(raster)
-
-# Create an empty raster with desired resolution and extent
-resolution <- c(0.75, 0.75)  # Set the desired resolution
-extent <- c(-85, -54, -24, 14)  # Set the desired extent
-empty_raster <- raster(resolution = resolution, xmn = extent[1], xmx = extent[2], ymn = extent[3], ymx = extent[4])
-
-# Generate coordinates
-subset_coords_plant <- site_loc_key_plant[rowSums(PAM_plant_site_final) >= 5, ]
-subset_coords_plant_sp <-subset_coords_plant[,1:2 ]
-subset_coords_frugivore <- site_loc_key_frugivore[rowSums(PAM_frugivore_site_final) >= 5, ]
-subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]
-
-#Get functional dispersion
-fdis_plant <- alpha_fd_indices_plant$functional_diversity_indices$fdis
-fdis_frugivore <- alpha_fd_indices_frugivore$functional_diversity_indices$fdis
-## subset frugivore to compare for plotting
-fdis_frugivore_sub <- fdis_frugivore[1:25]
-
-plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
-frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)
-## subset frugivore to compare for plotting
-frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)
-
-# Convert the dataframe to sf format
-spatial_fdis_plant <- st_as_sf(plant_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
-spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
-
-```
-
-```{r}
-# Rasterize the sf data to create the FD raster
-fd_raster_plant <- rasterize(spatial_fdis_plant, empty_raster)
-writeRaster(fd_raster_plant$fdis_plant, filename = file.path(output_path,"FD_plant_TA.tif"), format="GTiff",overwrite=T)
-writeRaster(fd_raster_plant$ID, filename = file.path(output_path,"FD_plant_TA_ID.tif"), format="GTiff")
-
-fd_raster_frugivore <- rasterize(spatial_fdis_frugivore, empty_raster)
-writeRaster(fd_raster_frugivore$fdis_frugivore, filename = file.path(output_path,"FD_frugivore_TA.tif"), format="GTiff",overwrite=T)
-writeRaster(fd_raster_frugivore$ID, filename = file.path(output_path,"FD_frugivore_TA_ID.tif"), format="GTiff")
-```
-
-# Functional diversity plots
-```{r}
-# set crs of sf objects
-spatial_fdis_plant <- spatial_fdis_plant %>% st_set_crs(4326)
-spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(4326)
-```
-
-```{r}
-# Raster to raster point
-spec_FD_raster_plant_point <- rasterToPoints(fd_raster_plant, spatial = TRUE)
-spec_FD_raster_frugivore_point <- rasterToPoints(fd_raster_frugivore, spatial = TRUE)
-# Convert to a 'conventional' dataframe
-fd_raster_plant_point <- data.frame(spec_FD_raster_plant_point)
-fd_raster_frugivore_point <- data.frame(spec_FD_raster_frugivore_point)
-```
-
-```{r}
-plantgridFDTA <-
-  ggplot() +
-  geom_sf(data = worldMap, fill = "white")+
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_tile(data= fd_raster_plant_point, aes(x=x, y=y, fill=fdis_plant)) +
-  scale_fill_scico(palette = "romaO", direction = 1) +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  ) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Plant Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
-plantgridFDTA
-```
-
-```{r}
-frugivoregridFDTA <-
-  ggplot() +
-  geom_sf(data = worldMap, fill = "white")+
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_tile(data= fd_raster_frugivore_point, aes(x=x, y=y, fill=fdis_frugivore)) +
-  scale_fill_scico(palette = "romaO", direction = 1) +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  ) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Frugivore Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
-frugivoregridFDTA
-```
-
-# Extract cell values
-```{r}
-plantgridFDTA <- plantgridFDTA[["data"]]
-frugivoregridFDTA <- frugivoregridFDTA[["data"]]
-```
-
-```{r}
-plot(spatial_fdis_plant$fdis_plant, spatial_fdis_frugivore$fdis_frugivore,
-     main = "Plant FD vs Frugivore FD",
-     xlab = "Plant Functional Dispersion by cell", ylab = "Frugivore Functional Dispersion by cell")
-abline(lm(spatial_fdis_frugivore$fdis_frugivore ~ spatial_fdis_plant$fdis_plant))
-summary(lm(spatial_fdis_frugivore$fdis_frugivore ~ spatial_fdis_plant$fdis_plant))
-```

### 2023-10-18 code/L2/funcDiv_troubleshooting.rmd
add more code from other script

--- 2023-10-18 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
@@
----
-title: "Tropical Andes Functional diversity for plants and Frugivores"
-author: "Hazel J. Anderson"
-project: "Plant-Frugivore Diversity"
-collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: ""
-data input: ""
-data output: ""
-date: "2023-08-03"
-output: html_document
----
+Func Div troubleshooting
 
@@
 ```{r}
-# Look at the quality spaces only (MAD index looks at mean absolute deviation from the dissimilary matrix; want the deviation
-# to be low, meaning that the true distances have been retained in the PCA)
-round(fspaces_quality_plants$"quality_fspaces", 3)            
-round(fspaces_quality_frugivore$"quality_fspaces", 3)     
-# Plot the quality spaces (chose to look at 3D, 4D, and 5D since they had the lowest MAD). Will go with 4 dimensions.
-mFD::quality.fspaces.plot(
-  fspaces_quality            = fspaces_quality_plants,
-  quality_metric             = "mad",
-  fspaces_plot               = c("pcoa_2d", "pcoa_3d", 
-                                 "pcoa_4d"),
-  name_file                  = NULL,
-  range_dist                 = NULL,
-  range_dev                  = NULL,
-  range_qdev                 = NULL,
-  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
-  gradient_deviation_quality = c(low = "yellow", high = "red"),
-  x_lab                      = "Trait-based distance")
-mFD::quality.fspaces.plot(
-  fspaces_quality            = fspaces_quality_frugivore,
-  quality_metric             = "mad",
-  fspaces_plot               = c("pcoa_3d", "pcoa_4d", 
-                                 "pcoa_5d"),
-  name_file                  = NULL,
-  range_dist                 = NULL,
-  range_dev                  = NULL,
-  range_qdev                 = NULL,
-  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
-  gradient_deviation_quality = c(low = "yellow", high = "red"),
-  x_lab                      = "Trait-based distance")
-```
-
-```{r}
 #testing correlation between functional axes and traits
@@
 sp_faxes_coord_frugivore <- fspaces_quality_frugivore$"details_fspaces"$"sp_pc_coord"
-# View the components of the PCA axes *Remember the firt few components explain the most variation in dissimilarity. Clusters into groups
-# Computes linear model for continuous traits and Kruskall-Wallis tests for other types. 
-plant_tr_faxes <- mFD::traits.faxes.cor(
-  sp_tr          = plant_traits_df_final, 
-  sp_faxes_coord = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")], 
-  plot           = TRUE)
-frugivore_tr_faxes <- mFD::traits.faxes.cor(
-  sp_tr          = frugivore_traits_df_final, 
-  sp_faxes_coord = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")], 
-  plot           = TRUE)
-# Print traits with significant effect:
-plant_tr_faxes$"tr_faxes_stat"[which(plant_tr_faxes$"tr_faxes_stat"$"p.value" < 0.05), ]
-frugivore_tr_faxes$"tr_faxes_stat"[which(frugivore_tr_faxes$"tr_faxes_stat"$"p.value" < 0.05), ]
-# Return plots:
-plant_tr_faxes$"tr_faxes_plot"
-frugivore_tr_faxes$"tr_faxes_plot"
+
 ```
@@
 sp_faxes_coord_frugivore <- fspaces_quality_frugivore$"details_fspaces"$"sp_pc_coord"
-big_plot_plants <- mFD::funct.space.plot(
-  sp_faxes_coord  = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")],
-  faxes           = c("PC1", "PC2", "PC3", "PC4"),
-  name_file       = NULL,
-  faxes_nm        = NULL,
-  range_faxes     = c(NA, NA),
-  color_bg        = "grey95",
-  color_pool      = "darkgreen",
-  fill_pool       = "white",
-  shape_pool      = 21,
-  size_pool       = 1,
-  plot_ch         = TRUE,
-  color_ch        = "black",
-  fill_ch         = "white",
-  alpha_ch        = 0.5,
-  plot_vertices   = TRUE,
-  color_vert      = "blueviolet",
-  fill_vert       = "blueviolet",
-  shape_vert      = 6,
-  size_vert       = 1,
-  plot_sp_nm      = NULL,
-  nm_size         = 3,
-  nm_color        = "black",
-  nm_fontface     = "plain",
-  check_input     = TRUE)
-big_plot_frugivore <- mFD::funct.space.plot(
-  sp_faxes_coord  = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")],
-  faxes           = c("PC1", "PC2", "PC3", "PC4"),
-  name_file       = NULL,
-  faxes_nm        = NULL,
-  range_faxes     = c(NA, NA),
-  color_bg        = "grey95",
-  color_pool      = "darkgreen",
-  fill_pool       = "white",
-  shape_pool      = 21,
-  size_pool       = 1,
-  plot_ch         = TRUE,
-  color_ch        = "black",
-  fill_ch         = "white",
-  alpha_ch        = 0.5,
-  plot_vertices   = TRUE,
-  color_vert      = "blueviolet",
-  fill_vert       = "blueviolet",
-  shape_vert      = 6,
-  size_vert       = 1,
-  plot_sp_nm      = NULL,
-  nm_size         = 3,
-  nm_color        = "black",
-  nm_fontface     = "plain",
-  check_input     = TRUE)
+
 #Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
@@
 
-```
-
-```{r}
-big_plot_plants
-big_plot_frugivore
 ```
@@
   details_returned = TRUE)
-
-details_list_plant <- alpha_fd_indices_plant$"details"
-details_list_frugivore <- alpha_fd_indices_frugivore$"details"
-
-#plot
-plots_alpha_plant <- mFD::alpha.multidim.plot(
-  output_alpha_fd_multidim = alpha_fd_indices_plant,
-  plot_asb_nm              = c("cell4", "cell20"),
-  ind_nm                   = c("fdis", "fide", "fnnd"
-                        , "fori", "fspe"),
-  faxes                    = NULL,
-  faxes_nm                 = NULL,
-  range_faxes              = c(NA, NA),
-  color_bg                 = "grey95",
-  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
-  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
-  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  color_ch          
-  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
-  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
-  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
-  shape_centroid_fspe      = 23,
-  color_centroid_fspe      = "black",
-  size_sp_nm               = 3, 
-  color_sp_nm              = "black",
-  plot_sp_nm               = NULL,
-  fontface_sp_nm           = "plain",
-  save_file                = FALSE,
-  check_input              = TRUE)
-
-plots_alpha_frugivore <- mFD::alpha.multidim.plot(
-  output_alpha_fd_multidim = alpha_fd_indices_frugivore,
-  plot_asb_nm              = c("cell4", "cell20"),
-  ind_nm                   = c("fdis", "fide", "fnnd"
-                        , "fori", "fspe"),
-  faxes                    = NULL,
-  faxes_nm                 = NULL,
-  range_faxes              = c(NA, NA),
-  color_bg                 = "grey95",
-  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
-  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
-  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  color_ch          
-  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
-  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
-  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
-  shape_centroid_fspe      = 23,
-  color_centroid_fspe      = "black",
-  size_sp_nm               = 3, 
-  color_sp_nm              = "black",
-  plot_sp_nm               = NULL,
-  fontface_sp_nm           = "plain",
-  save_file                = FALSE,
-  check_input              = TRUE)
-
-plots_alpha_plant$"fric"$"patchwork"  #comparing convex hulls of different hulls
-plots_alpha_frugivore$"fric"$"patchwork"
-plots_alpha_plant$"fdiv"$"patchwork"
-plots_alpha_frugivore$"fdiv"$"patchwork"
-plots_alpha_plant$"fspe"$"patchwork"
-plots_alpha_frugivore$"fspe"$"patchwork"
-plots_alpha_plant$"fdis"$"patchwork"
-plots_alpha_frugivore$"fdis"$"patchwork"
-plots_alpha_plant$"fori"$"patchwork"
-plots_alpha_frugivore$"fori"$"patchwork"
 ```
-
-# Turning the FD values into a raster
-```{r}
-
-#Need to subset the coordinates made at the top to the same set of the assemblages (removed values less than 5 so these coords need to be removed to do this correctly)
-
-#Have coordinates and associated values
-
-library(raster)
-
-# Create an empty raster with desired resolution and extent
-resolution <- c(0.75, 0.75)  # Set the desired resolution
-extent <- c(-85, -54, -24, 14)  # Set the desired extent
-empty_raster <- raster(resolution = resolution, xmn = extent[1], xmx = extent[2], ymn = extent[3], ymx = extent[4])
-
-# Generate coordinates
-subset_coords_plant <- site_loc_key_plant[rowSums(PAM_plant_site_final) >= 5, ]
-subset_coords_plant_sp <-subset_coords_plant[,1:2 ]
-subset_coords_frugivore <- site_loc_key_frugivore[rowSums(PAM_frugivore_site_final) >= 5, ]
-subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]
-
-#Get functional dispersion
-fdis_plant <- alpha_fd_indices_plant$functional_diversity_indices$fdis
-fdis_frugivore <- alpha_fd_indices_frugivore$functional_diversity_indices$fdis
-## subset frugivore to compare for plotting
-fdis_frugivore_sub <- fdis_frugivore[1:25]
-
-plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
-frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)
-## subset frugivore to compare for plotting
-frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)
-
-# Convert the dataframe to sf format
-spatial_fdis_plant <- st_as_sf(plant_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
-spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
-
-```
-
-```{r}
-# Rasterize the sf data to create the FD raster
-fd_raster_plant <- rasterize(spatial_fdis_plant, empty_raster)
-writeRaster(fd_raster_plant$fdis_plant, filename = file.path(output_path,"FD_plant_TA.tif"), format="GTiff",overwrite=T)
-writeRaster(fd_raster_plant$ID, filename = file.path(output_path,"FD_plant_TA_ID.tif"), format="GTiff")
-
-fd_raster_frugivore <- rasterize(spatial_fdis_frugivore, empty_raster)
-writeRaster(fd_raster_frugivore$fdis_frugivore, filename = file.path(output_path,"FD_frugivore_TA.tif"), format="GTiff",overwrite=T)
-writeRaster(fd_raster_frugivore$ID, filename = file.path(output_path,"FD_frugivore_TA_ID.tif"), format="GTiff")
-```
-
-# Functional diversity plots
-```{r}
-# set crs of sf objects
-spatial_fdis_plant <- spatial_fdis_plant %>% st_set_crs(4326)
-spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(4326)
-```
-
-```{r}
-# Raster to raster point
-spec_FD_raster_plant_point <- rasterToPoints(fd_raster_plant, spatial = TRUE)
-spec_FD_raster_frugivore_point <- rasterToPoints(fd_raster_frugivore, spatial = TRUE)
-# Convert to a 'conventional' dataframe
-fd_raster_plant_point <- data.frame(spec_FD_raster_plant_point)
-fd_raster_frugivore_point <- data.frame(spec_FD_raster_frugivore_point)
-```
-
-```{r}
-plantgridFDTA <-
-  ggplot() +
-  geom_sf(data = worldMap, fill = "white")+
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_tile(data= fd_raster_plant_point, aes(x=x, y=y, fill=fdis_plant)) +
-  scale_fill_scico(palette = "romaO", direction = 1) +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  ) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Plant Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
-plantgridFDTA
-```
-
-```{r}
-frugivoregridFDTA <-
-  ggplot() +
-  geom_sf(data = worldMap, fill = "white")+
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_tile(data= fd_raster_frugivore_point, aes(x=x, y=y, fill=fdis_frugivore)) +
-  scale_fill_scico(palette = "romaO", direction = 1) +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  ) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Frugivore Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
-frugivoregridFDTA
-```
-
-# Extract cell values
-```{r}
-plantgridFDTA <- plantgridFDTA[["data"]]
-frugivoregridFDTA <- frugivoregridFDTA[["data"]]
-```
-
-```{r}
-plot(spatial_fdis_plant$fdis_plant, spatial_fdis_frugivore$fdis_frugivore,
-     main = "Plant FD vs Frugivore FD",
-     xlab = "Plant Functional Dispersion by cell", ylab = "Frugivore Functional Dispersion by cell")
-abline(lm(spatial_fdis_frugivore$fdis_frugivore ~ spatial_fdis_plant$fdis_plant))
-summary(lm(spatial_fdis_frugivore$fdis_frugivore ~ spatial_fdis_plant$fdis_plant))
-```

### 2023-10-18 code/L0/L0_TropicalAndes_frugivores_GBIF.Rmd
add code to extract lat and long from geometry and convert to dataframe

@@
 library(rgbif) 
+library(sf)
 ```
@@
 ```{r}
-TropicalAndes_frugivore_occ_forest <- TropicalAndes_frugivore_occ[Forest_sf,]
+TropicalAndes_frugivore_occ_forest <- TropicalAndes_frugivore_occ[TropicalAndes_IUCNHabitat_Forest,]
+```
+
+## Extract lat and long to seperate columns
+```{r}
+TropicalAndes_frugivore_occ_forest <- TropicalAndes_frugivore_occ_forest %>%
+    mutate(decimalLongitude = sf::st_coordinates(.)[,1],
+           decimalLatitude = sf::st_coordinates(.)[,2])
+```
+
+```{r}
+TropicalAndes_frugivore_occ_forest <- as.data.frame(TropicalAndes_frugivore_occ_forest)
 ```

### 2023-10-18 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
add code to extract lat and long from geometry and convert to dataframe

@@
 ```{r}
-TropicalAndes_plant_occ_forest <- TropicalAndes_plant_occ[Forest_sf,]
+TropicalAndes_plant_occ_forest <- TropicalAndes_plant_occ[TropicalAndes_IUCNHabitat_Forest,]
+```
+
+## Extract lat and long to seperate columns
+```{r}
+TropicalAndes_plant_occ_forest <- TropicalAndes_plant_occ_forest %>%
+    mutate(decimalLongitude = sf::st_coordinates(.)[,1],
+           decimalLatitude = sf::st_coordinates(.)[,2])
+```
+
+```{r}
+TropicalAndes_plant_occ_forest <- as.data.frame(TropicalAndes_plant_occ_forest)
 ```

### 2023-10-19 code/L2/FunDiv_troubleshooting.R
change to R script not markdown

@@ -1,8 +1,7 @@
-Func Div troubleshooting
+#Func Div troubleshooting
 
 
 # Load required packages
-```{r}
 library(sf)
 library(dplyr)
 library(ggplot2)
@@ -14,26 +13,24 @@ library(scico)
 library(ggspatial)
 library(letsR)
 library(mFD)
-```
 
 # Set file paths
-```{r}
-data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
-data_path2<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
-output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
-```
+data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+data_path_L0 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+output_path_L2 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
+
 
 # Read in Data
-```{r}
-TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_plant_occ_forest.csv"))
-TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_frugivore_occ_forest.csv"))
-TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path2, "Forest_sf.shp"), layer = "Forest_sf")
-frugivore_traits <- read.csv(file.path(data_path,"TropicalAndes_Frugivoria_traits_Forest.csv"))
-plant_traits <- read.csv(file.path(data_path,"TropicalAndes_plant_traits_forest.csv"))
-```
+
+TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L0,"TropicalAndes_GBIF_plant_occ.csv"))
+TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path_L0,"TropicalAndes_GBIF_frugivore_occ.csv"))
+TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
+frugivore_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_Frugivoria_traits_Forest.csv"))
[…]

### 2023-10-19 code/L2/FunDiv_troubleshooting.R
remove unnecessary spaces

@@ -41,7 +41,6 @@ plot(plants_PAM, xlab = "Longitude", ylab = "Latitude", main = "Plant richness m
 plants_PAM_matrix <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE)
 
 
-
 x <- TropicalAndes_frugivore_occ_forest$decimalLongitude
 y <- TropicalAndes_frugivore_occ_forest$decimalLatitude
 frugivore_xy <- cbind(x, y)
@@ -63,10 +62,8 @@ plant_traits <- plant_traits[, c("species", "plant_height", "dispersal_syndrome"
 frugivore_traits <- frugivore_traits[, c("IUCN_species_name", "diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
 
 
-
 coordinates_sf_plant <- st_as_sf(TropicalAndes_plant_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(TropicalAndes_IUCNHabitat_Forest))
 coordinates_sf_frugivore <- st_as_sf(TropicalAndes_frugivore_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(TropicalAndes_IUCNHabitat_Forest))
-```
 
 
 # Perform the spatial intersection
@@ -75,7 +72,7 @@ intersections_frugivore_TA <- st_intersection(coordinates_sf_frugivore, Tropical
 # Convert the intersected sf object back to a dataframe
 coordinates_df_subset_plant <- as.data.frame(intersections_plant_TA)
 coordinates_df_subset_frugivore <- as.data.frame(intersections_frugivore_TA)
-```
+
 
 
 # Print the resulting subset of coordinates

### 2023-10-25 code/L0/L0_TropicalAndes_plants_GIFT.Rmd
update to include plant growth form

@@
 ```{r}
-trait_meta <- GIFT_traits_meta()
+GIFT_traits_meta()
 ```
-Traits of interest: plant height (ID = 1.6.3), lifespan (ID = 2.2.1), dispersal syndrome (ID = 3.3), Plant_height_min (1.6.1), Plant_height_max (1.6.2), Fruit_type_1 (3.16.1), fruit_dryness_1 (3.18.1), Fruit_length_min (3.13.1), Fruit_length_max (3.13.2), Fruit_length_mean (3.13.3), Seed_mass_min (3.2.1), Seed_mass_max (3.2.2), Seed_mass_mean (3.2.3), Seed_length_min (3.10.1), Seed_length_max (3.10.2), Seed_length_mean (3.10.3), Seed_width_min (3.11.1), Seed_width_max (3.11.2), Seed_width_mean (3.11.3)
+Traits of interest: plant height (ID = 1.6.3), lifespan (ID = 2.2.1), dispersal syndrome (ID = 3.3), Plant_height_min (1.6.1), Plant_height_max (1.6.2), Fruit_type_1 (3.16.1), fruit_dryness_1 (3.18.1), Fruit_length_min (3.13.1), Fruit_length_max (3.13.2), Fruit_length_mean (3.13.3), Seed_mass_min (3.2.1), Seed_mass_max (3.2.2), Seed_mass_mean (3.2.3), Seed_length_min (3.10.1), Seed_length_max (3.10.2), Seed_length_mean (3.10.3), Seed_width_min (3.11.1), Seed_width_max (3.11.2), Seed_width_mean (3.11.3), Growth_form (1.2.1)
 
@@
 seed_width_max <- GIFT_traits(trait_IDs = c("3.11.2"))
+
+growth_form <- GIFT_traits(trait_IDs = c("1.2.1"))
 ```
@@
 seed_width_max <- seed_width_max %>% rename(trait_value = trait_value_3.11.2, species = work_species)
+
+growth_form <- growth_form %>% rename(trait_value = trait_value_1.2.1, species = work_species)
 ```
@@
 seed_width_max <- seed_width_max %>% select(species, trait_value)
+
+growth_form <- growth_form %>% select(species, trait_value)
 ```
@@
 seed_width_max$trait_name <- "seed_width_max"
-
+growth_form$trait_name <- "growth_form"
 ```
@@
 #combine all trait dataframes into one long table
-GIFT_traits <- dplyr::bind_rows(height_mean, height_min, height_max, lifespan, dispersal_syndrome_1, dispersal_syndrome_2, fruit_type, fruit_dryness, fruit_length_mean, fruit_length_min, fruit_length_max, seed_mass_mean, seed_mass_min, seed_mass_max, seed_length_mean, seed_length_min, seed_length_max, seed_width_mean, seed_width_min, seed_width_max)
+GIFT_traits <- dplyr::bind_rows(height_mean, height_min, height_max, lifespan, dispersal_syndrome_1, dispersal_syndrome_2, fruit_type, fruit_dryness, fruit_length_mean, fruit_length_min, fruit_length_max, seed_mass_mean, seed_mass_min, seed_mass_max, seed_length_mean, seed_length_min, seed_length_max, seed_width_mean, seed_width_min, seed_width_max, growth_form)
 ```

### 2023-10-25 code/L0/L0_TropicalAndes_plants_TRY.Rmd
update to include plant growth form

@@
 ```{r}
-try_data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/28165_02082023081407/28165.txt")
+try_data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/29548_24102023041242/29548.txt")
 ```
 
+# Summary
 ```{r}
@@
 
-#Subset to species list
 ```{r}
-species_list <- read.csv("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1/TropicalAndes_plant_Forest_SpeciesList.csv")
-species_list <- species_list$x
+print("The number of records is")
+nrow(try_data)
+print("The number of species is" )
+length(unique(try_data$AccSpeciesName))
 ```
@@
 ```{r}
-try_data_filtered <- try_data %>%
-  filter(AccSpeciesName %in% species_list)
-```
-
-# Summary 
-```{r}
-glimpse(try_data_filtered)
-```
-
-```{r}
-print("The number of records is")
-nrow(try_data_filtered)
-print("The number of species is" )
-length(unique(try_data_filtered$AccSpeciesName))
-```
-
-```{r}
-try_data_filtered %>% count(TraitName)
+try_data %>% count(TraitName)
 ```
@@
 ```{r}
-TropicalAndes_TRY_path <- file.path(output_path,"TropicalAndes_TRY_plant_traits_forestSpecies.csv") 
-write.csv(try_data_filtered, file = TropicalAndes_TRY_path)
+write.csv(try_data, file = file.path(output_path,"TropicalAndes_TRY_plant_traits.csv") )
 ```

### 2023-10-25 code/L0/L0_TropicalAndes_plants_BIEN_traits.Rmd
add full trait list code
update to include plant growth form

@@
 
+```{r}
+BIEN_trait_list()
+```
+
 ## List of desired traits
 ```{r}
-trait_list <- c("whole plant dispersal syndrome", "whole plant height", "maximum whole plant longevity", "maximum whole plant height", "minimum whole plant height", "longest whole plant longevity", "fruit type", "maximum fruit length", "minimum fruit length", "seed mass")
+trait_list <- c("whole plant dispersal syndrome", "whole plant height", "maximum whole plant longevity", "maximum whole plant height", "minimum whole plant height", "longest whole plant longevity", "fruit type", "maximum fruit length", "minimum fruit length", "seed mass", "whole plant growth form", "whole plant growth form diversity")
 ```

### 2023-10-31 code/L0/L0_TropicalAndes_plants_checklist.Rmd
add genus based on taxon_id from checklist

--- 2023-10-18 code/L0/L0_TropicalAndes_plants_checklist.Rmd
@@
 ```
+```{r}
+glimpse(checklist$checklists)
+```
+```{r}
+#add genus based on work_ID
+full <- GIFT_taxonomy()
+```
+
+```{r}
+genus_ID <- checklist$checklists$genus_ID
+genus_list <- full %>% filter(taxon_ID %in% genus_ID) %>% select(taxon_ID, taxon_name)
+names(genus_list) <- c("genus_ID", "genus")
+```
+
+```{r}
+checklist_tax <- dplyr::inner_join(checklist$checklists, genus_list)
+```
+
+```{r}
+print("The number of species is")
+length(unique(checklist_tax$work_species))
+print("The number of genera is")
+length(unique(checklist_tax$genus))
+print("The number of families is")
+length(unique(checklist_tax$family))
+```
+
+# Write data to csv
+```{r}
+write.csv(checklist_tax, file = file.path(output_path,"TropicalAndes_plant_checklist.csv"))
+```

### 2023-10-31 code/L1/L1_F2_TropicalAndes_frugivore_traits.Rmd
update input files, file paths, and minor headings

@@
 overview: "Subsetting Frugivoria to the species and traits of interest"
-data input: ""
+data input: "TropicalAndes_Frugivoria_frugivore_traits.csv", "TropicalAndes_GBIF_frugivore_occ_cleaned.csv"
 data output: ""
@@
 ```{r}
-data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
-data_path2<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+data_path_L1<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
 output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
@@
 ```{r}
-Frugivoria <- read.csv(file.path(data_path, file = "TropicalAndes_Frugivoria_frugivore_traits.csv"))
-frugivore_species <- read.csv(file.path(data_path2, file = "TropicalAndes_frugivore_Forest_SpeciesList.csv"))
+Frugivoria <- read.csv(file.path(data_path_L0, file = "TropicalAndes_Frugivoria_frugivore_traits.csv"))
+frugivore_occ <- read.csv(file.path(data_path_L1, file = "TropicalAndes_GBIF_frugivore_occ_cleaned.csv"))
 ```
 
+# Get species list from frugivore_occ data
 ```{r}
-frugivore_species <- frugivore_species$x
+frugivore_species <- unique(frugivore_occ$species)
 ```
 
+# Subset Frugivoria by species list
 ```{r}
-# subset Frugivoria by species list
 Frugivoria_filtered <- Frugivoria %>%
@@
 
+# Keep only traits of interest
 ```{r}
@@
 
+```{r}
+source(file.path(data_path_L0, "data_summary.R"))
+```
+
+```{r}
+data_summary(Frugivoria_subset, Frugivoria_subset$species, Frugivoria_subset$genus, Frugivoria_subset$family)
+```
+
 # Write data to csv

### 2023-10-31 code/L1/L1_P2_TropicalAndes_plant_species_synonyms.Rmd
add header information, update file path, remove unnecessary/old code, clean up unnecessary text

---
title: "Plant species synonym list"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Get synonmys for the plant species list of GBIF occurrence data"
data input: "GBIF_species_list.csv"
data output: ""
date: "2023-08-03"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

```{r}
library(taxadb)
#library(taxize)
library(dplyr)
```

```{r}
#create local itis database
td_create("itis",overwrite=FALSE)
```

# Read in species list
```{r}
GBIF_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
GBIF_species_list <- unique(GBIF_occ$species)
```

```{r}
GBIF_species_list <- as.data.frame(GBIF_species_list)
names(GBIF_species_list) <- c("Scientific.Name")
```

#Using taxize
```{r}
species_synonyms <- synonyms(GBIF_species_list$Scientific.Name, db = "pow", rows = 1)
```

# Using taxadb
```{r}
#get IDS for each scientific name
syn1<-GBIF_species_list %>%
  mutate(ID=get_ids(Scientific.Name,"itis"))

#Deal with NAs (one name corresponds to more than 1 ITIS code)
syn1_NA<-as.data.frame(syn1$Scientific.Name[is.na(syn1$ID)])
colnames(syn1_NA)<-c("name")

NA_IDS<-NULL
tmp <- NULL
tmp$name <- NA
for(i in unique(syn1_NA$name)){
  tmp <-as.data.frame(filter_name(i, 'itis')[5])
  tmp$name<-paste0(i)
  NA_IDS<-rbind(NA_IDS,tmp)
}

#join with original names
colnames(syn1)<-c("name","ID")
IDS<-left_join(syn1,NA_IDS,by="name")

#extract just the unique IDs
IDS<-data.frame(ID=c(IDS[,"ID"],IDS[,"acceptedNameUsageID"]))
IDS<-as.data.frame(unique(IDS$ID))
IDS<-as.data.frame(IDS[-is.na(IDS)])
colnames(IDS)<-"ID"
#extract all names with synonyms in ITIS that are at the species level
#set query
ITIS<-taxa_tbl("itis") %>%
  select(scientificName,taxonRank,acceptedNameUsageID,taxonomicStatus) %>%
  filter(taxonRank == "species")

#see query
ITIS %>% show_query()
#retrieve results
ITIS_names<-ITIS %>% collect()

#filter to only those that match ITIS codes for all my species
ITIS_names<-ITIS_names %>%
  filter(acceptedNameUsageID %in% IDS$ID)
```

### 2023-10-31 code/L2/FunDiv_troubleshooting.R
remove frugivore code to simplify

@@ -1,5 +1,5 @@
 #Func Div troubleshooting
-
+## Focus on plants
 
 # Load required packages
 library(sf)
@@ -22,10 +22,8 @@ output_path_L2 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L
 
 # Read in Data
 
-TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L0,"TropicalAndes_GBIF_plant_occ.csv"))
-TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path_L0,"TropicalAndes_GBIF_frugivore_occ.csv"))
+TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
 TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
-frugivore_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_Frugivoria_traits_Forest.csv"))
 plant_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_plant_traits_forest.csv"))
 
 
@@ -41,85 +39,50 @@ plot(plants_PAM, xlab = "Longitude", ylab = "Latitude", main = "Plant richness m
 plants_PAM_matrix <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE)
 
 
-x <- TropicalAndes_frugivore_occ_forest$decimalLongitude
-y <- TropicalAndes_frugivore_occ_forest$decimalLatitude
-frugivore_xy <- cbind(x, y)
-frugivore_species <- TropicalAndes_frugivore_occ_forest$species
-frugivore_PAM <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14)
-summary(frugivore_PAM)
-plot(frugivore_PAM, xlab = "Longitude", ylab = "Latitude", main = "Frugivore richness map")
-frugivore_PAM_matrix <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE)
-
-
[…]

### 2023-10-31 code/L2/L2_TropicalAndes_data_citizenscience.Rmd
inital commit, read in data, plot year vs number of records for occurrence data, subset cleaned data for citizen science

Citizen science assessment

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
```

# Load required packages
```{r}
library(dplyr)
library(ggplot2)
library(rgbif)
```

# Read in Data
```{r}
plant_GBIF_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
```

# Time (year) vs records plot
```{r}
records_time <- GBIF_occ %>% count(year)
names(records_time) <- c("year", "count")
```

```{r}
ggplot(records_time, aes(x=year, y=count)) +
  geom_line() +
  xlim(1950, 2023) +
  geom_area(fill= "darkgreen")+
  ggtitle("Plant occurrence records in GBIF by year since 1950")
```

# Subset citizen science data
```{r}
dataset_keys <- unique(GBIF_occ$datasetKey)
# 50c9509d-22c7-4a22-a47d-8c48425ef4a7 = iNaturalist
GBIF_occ_iNaturalist <- GBIF_occ %>% filter(datasetKey == "50c9509d-22c7-4a22-a47d-8c48425ef4a7")
```

```{r}
records_time_iNaturalist <- GBIF_occ_iNaturalist %>% count(year)
names(records_time_iNaturalist) <- c("year", "count")
```

```{r}
ggplot(records_time_iNaturalist, aes(x=year, y=count)) +
  geom_line() +
  xlim(1950, 2023) +
  geom_area(fill= "lightgreen") +
  ggtitle("iNaturalist Plant occurrence records in GBIF by year since 1950")
```
# iNaturalist proportion of total records
```{r}
records_time$group <- c("all")
records_time_iNaturalist$group <- c("iNaturalist")
records_time_both <- rbind(records_time, records_time_iNaturalist)
```

```{r}
ggplot(records_time_both, aes(x=year, y=count, color=group)) + 
    geom_line() +
  xlim(1970, 2023) +
  ggtitle("Plant occurrence records in GBIF by year since 1970")
```

### 2023-10-31 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
update with fixes from troubleshooting

--- 2023-10-04 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
@@
 ```{r}
-data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
-data_path2<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
-output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
+data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
 ```
@@
 ```{r}
-TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_plant_occ_forest.csv"))
-TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_frugivore_occ_forest.csv"))
-TropicalAndes_IUCNHabitat_Forest <- raster(file.path(data_path2, "TropicalAndes_IUCNHabitat_Forest.tif"))
-frugivore_traits <- read.csv(file.path(data_path,"TropicalAndes_Frugivoria_traits_Forest.csv"))
-plant_traits <- read.csv(file.path(data_path,"TropicalAndes_plant_traits_forest.csv"))
+TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
+TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_frugivore_occ_cleaned.csv"))
+TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
+frugivore_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_Frugivoria_traits_Forest.csv"))
+plant_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_plant_traits_forest.csv"))
 ```
@@
 # Prepping data for functional diversity calculations
+```{r}
+plant_traits <- na.omit(plant_traits)
+plant_traits <- plant_traits[, c("species", "plant_height", "dispersal_syndrome", "plant_lifespan")]
+```
 
 ```{r}
-plant_traits <- na.omit(plant_traits)
 frugivore_traits <- na.omit(frugivore_traits)
@@
 frugivore_traits$body_size_mm <- as.numeric(frugivore_traits$body_size_mm)
-plant_traits <- plant_traits[, c("species", "plant_height", "dispersal_syndrome", "plant_lifespan")]
 frugivore_traits <- frugivore_traits[, c("IUCN_species_name", "diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
@@
 ```{r}
-coordinates_sf_plant <- st_as_sf(TropicalAndes_plant_occ_forest_sub, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(Forest_sf))
-coordinates_sf_frugivore <- st_as_sf(TropicalAndes_frugivore_occ_forest_sub, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(Forest_sf))
+coordinates_sf_plant <- st_as_sf(TropicalAndes_plant_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(TropicalAndes_IUCNHabitat_Forest))
 ```
@@
 ```{r}
-# Perform the spatial intersection
-intersections_plant_TA <- st_intersection(coordinates_sf_plant, Forest_sf)
-intersections_frugivore_TA <- st_intersection(coordinates_sf_frugivore, Forest_sf)
+coordinates_sf_frugivore <- st_as_sf(TropicalAndes_frugivore_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(TropicalAndes_IUCNHabitat_Forest))
+```
+
+## Perform the spatial intersection
+```{r}
+
+intersections_plant_TA <- st_intersection(coordinates_sf_plant, TropicalAndes_IUCNHabitat_Forest)
+intersections_frugivore_TA <- st_intersection(coordinates_sf_frugivore, TropicalAndes_IUCNHabitat_Forest)
+
 # Convert the intersected sf object back to a dataframe
@@
 print(coordinates_df_subset_frugivore)
+
 # Convert the point geometry to a dataframe of longitude and latitude
@@
 subset_frugivore <- as.data.frame(st_coordinates(intersections_frugivore_TA))
+```
 
-# Match the coordinates to the original dataframe
+## Match the coordinates to the original dataframe
+```{r}
 #merge the subset and full dataframe together to get final TA dataset
@@
 
+## Turn PAM into matrix
 ```{r}
-# Turn PAM into matrix
 plant_PAM <- as.matrix(plant_PAM_filter)
 frugivore_PAM <- as.matrix(frugivore_PAM_filter)
+```
 
-# Add rownames
+## Add rownames
+```{r}
 # Get the number of rows in the matrix
 num_rows_plant <- nrow(plant_PAM)
-num_rows_frugivore <- nrow(frugivore_PAM)
+
 # Generate unique row names
 row_names_plant <- paste0("cell", 1:num_rows_plant)
-row_names_frugivore <- paste0("cell", 1:num_rows_frugivore)
+
 # Assign row names to the matrix
 rownames(plant_PAM) <- row_names_plant
-rownames(frugivore_PAM) <- row_names_frugivore
+
 # Print the matrix with row names
@@
 clean_column_names_plant <- gsub("_", " ", column_names_plant)
+
+#Insert clean column names
+colnames(plant_PAM) <- clean_column_names_plant
+```
+
+```{r}
+# Get the number of rows in the matrix
+num_rows_frugivore <- nrow(frugivore_PAM)
+
+# Generate unique row names
+row_names_frugivore <- paste0("cell", 1:num_rows_frugivore)
+
+# Assign row names to the matrix
+rownames(frugivore_PAM) <- row_names_frugivore
+
+# Print the matrix with row names
 print(frugivore_PAM)
@@
 clean_column_names_frugivore <- gsub("_", " ", column_names_frugivore)
+
 #Insert clean column names
-colnames(plant_PAM) <- clean_column_names_plant
 colnames(frugivore_PAM) <- clean_column_names_frugivore
@@
 
+## Turn data into correct data types for inputs into the trait categories dataframe
 ```{r}
-#turn data into correct data types for inputs into the trait categories dataframe
+
+plant_traits$plant_height <- as.numeric(plant_traits$plant_height)
+plant_traits$plant_lifespan<- as.numeric(plant_traits$plant_lifespan)
+plant_traits$dispersal_syndrome <- as.factor(plant_traits$dispersal_syndrome)
+```
+
+```{r}
 frugivore_traits$diet_cat <- as.factor(frugivore_traits$diet_cat)
@@
 frugivore_traits$generation_time <- as.numeric(frugivore_traits$generation_time)
+```
 
-plant_traits$plant_height <- as.numeric(plant_traits$plant_height)
-plant_traits$plant_lifespan<- as.numeric(plant_traits$plant_lifespan)
-plant_traits$dispersal_syndrome <- as.factor(plant_traits$dispersal_syndrome)
-
-# Remove the species from PAM that have no occurrences anymore after subsetting to TA
+## Remove the species from PAM that have no occurrences anymore after subsetting to TA
+```{r}
 # Remove columns with sum equal to zero
@@
 PAM_frugivore_site_final <- frugivore_PAM[, colSums(frugivore_PAM) != 0]
+
 #Save coordinates for later
@@
 PAM_frugivore_site_final <- PAM_frugivore_site_final[,-columns_to_remove]
+
 colnames_plant <- colnames(PAM_plant_site_final)
@@
 
+## Remove species names from trait matrix not in the PAM
 ```{r}
-# Remove species names from trait matrix not in the PAM
+
 plant_traits_df_subset <- plant_traits %>% filter(species %in% colnames_plant)
@@
 class(frugivore_traits)
+
 plant_traits_matrix <- as.matrix(plant_traits_df_subset)
@@
 rownames(frugivore_traits_matrix) <- row_names_frugivore
+
 # Turn back into dataframe
@@
 frugivore_traits_df_final <-as.data.frame(frugivore_traits_matrix)
-plant_traits_df$X <-NULL
-frugivore_traits_df$X <-NULL
+
+plant_traits_df_final$X <-NULL
+frugivore_traits_df_final$X <-NULL
 ```
@@
 #fix types
+plant_traits_df_final$plant_height <- as.numeric(plant_traits_df_final$plant_height)
+plant_traits_df_final$plant_lifespan<- as.numeric(plant_traits_df_final$plant_lifespan)
+plant_traits_df_final$dispersal_syndrome <- as.factor(plant_traits_df_final$dispersal_syndrome)
+```
+
+```{r}
 frugivore_traits_df_final$diet_cat <- as.factor(frugivore_traits_df_final$diet_cat)
@@
 frugivore_traits_df_final$generation_time <- as.numeric(frugivore_traits_df_final$generation_time)
+```
 
-plant_traits_df_final$plant_height <- as.numeric(plant_traits_df_final$plant_height)
-plant_traits_df_final$plant_lifespan<- as.numeric(plant_traits_df_final$plant_lifespan)
-plant_traits_df_final$dispersal_syndrome <- as.factor(plant_traits_df_final$dispersal_syndrome)
-
+```{r}
 # remove duplicate species name column
@@
 ```{r}
-trait_name <- c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")
-trait_type <- c("N", "Q", "Q", "Q")
-frug_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
-trait_name <- c("dispersal_syndrome", "plant_height", "plant_lifespan")
-trait_type <- c("N", "Q", "Q") 
-plant_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
+plant_trait_name <- c("dispersal_syndrome", "plant_height", "plant_lifespan")
+plant_trait_type <- c("N", "Q", "Q") 
+plant_trait_cat <- as.data.frame(cbind(plant_trait_name, plant_trait_type))
 ```
@@
 ```{r}
-# Summary of the assemblages * species dataframe:
+frug_trait_name <- c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")
+frug_trait_type <- c("N", "Q", "Q", "Q")
+frug_trait_cat <- as.data.frame(cbind(frug_trait_name, frug_trait_type))
+```
+
+# Summary of the assemblages * species dataframe
+```{r}
 asb_sp_plant_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_plant_site_final)
+
 asb_sp_frugivore_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_frugivore_site_final)
@@
 
+# Species traits summary
 ```{r}
-# Species traits summary:
 plant_traits_summ <- mFD::sp.tr.summary(
@@
   stop_if_NA = TRUE)
+
 frugivore_traits_summ <- mFD::sp.tr.summary(
@@
 ```{r}
-
 sp_dist_plants <- mFD::funct.dist(
@@
 ```{r}
-# Look at the quality spaces only (MAD index looks at mean absolute deviation from the dissimilary matrix; want the deviation
-# to be low, meaning that the true distances have been retained in the PCA)
+# Look at the quality spaces only (MAD index looks at mean absolute deviation from the dissimilary matrix; want the deviation to be low, meaning that the true distances have been retained in the PCA)
 round(fspaces_quality_plants$"quality_fspaces", 3)            
@@
 sp_faxes_coord_frugivore <- fspaces_quality_frugivore$"details_fspaces"$"sp_pc_coord"
+```
+
+```{r}
 # View the components of the PCA axes *Remember the firt few components explain the most variation in dissimilarity. Clusters into groups
+
 # Computes linear model for continuous traits and Kruskall-Wallis tests for other types. 
@@
   plot           = TRUE)
+
 frugivore_tr_faxes <- mFD::traits.faxes.cor(
@@
   plot           = TRUE)
+
 # Print traits with significant effect:
@@
   check_input     = TRUE)
+
 big_plot_frugivore <- mFD::funct.space.plot(
@@
   check_input     = TRUE)
+
 #Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
@@
 alpha_fd_indices_plant <- mFD::alpha.fd.multidim(
-  sp_faxes_coord   = sp_faxes_coord_plants,
+  sp_faxes_coord   = sp_faxes_coord_plants_sub,
   asb_sp_w         = PAM_plant_site_final,
@@
 alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(
-  sp_faxes_coord   = sp_faxes_coord_frugivore,
+  sp_faxes_coord   = sp_faxes_coord_frugivore_sub,
   asb_sp_w         = PAM_plant_site_final,

### 2023-11-01 code/L1/L1_P5_TropicalAndes_plant_occ.Rmd
update file names for adjusted workflow

--- 2023-10-04 code/L1/L1_P3_TropicalAndes_plant_traits.Rmd
@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: "This script subsets plant trait data by synonyms species list and combine."
+overview: "This script subsets plant occurrence data to species with complete trait coverage."
 data input: "ADD"

### 2023-11-01 code/L0/L0_TropicalAndes_plants_speciesLink_occ.Rmd

---
title: "Tropical Andes plant speciesLink data"
authors: "Hazel J. Anderson"
collaborators: "None"
data input: "none"
data output: "TBD"
project: "Frugivoria"
date: "2023-10-18"
output: html_document
---
# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(Rocc)
library(tidyr)
library(dplyr)
library(sf)
```

## List of desired countries
```{r}
TA_countries <- c("Bolivia", "Ecuador", "Venezuela", "Colombia", "Peru")
```

```{r}
TropicalAndes_speciesLink_occ <- rspeciesLink(dir = output_path, filename = "TropicalAndes_speciesLink_occ", basisOfRecord = c('PreservedSpecimen', 'HumanObservation', 'MachineObservation'), country = TA_countries, Coordinates = c("Yes", "Original", "Automatic"), Scope = "plants")
```

### 2023-11-01 archive/L1_P2_TropicalAndes_plant_species_synonyms.Rmd

--- 2023-10-31 code/L1/L1_P2_TropicalAndes_plant_species_synonyms.Rmd
@@
 ```{r}
-species_synonyms <- synonyms(GBIF_species_list$Scientific.Name, db = "pow", rows = 1)
+species_synonyms <- synonyms(GBIF_species_list$Scientific.Name, db = "pow")
 ```
@@
 syn1<-GBIF_species_list %>%
+    select(Scientific.Name) %>%
     mutate(ID=get_ids(Scientific.Name,"itis"))
+```
 
+```{r}
 #Deal with NAs (one name corresponds to more than 1 ITIS code)

### 2023-11-01 code/L0/L0_TropicalAndes_plants_BIEN_occ.Rmd

@@
 
-# Load Tropical Andes moist lowland and montane forest shape
+# Retrive occurrence data for Tropical Andes countries from BIEN database
+
+## List of desired countries
 ```{r}
-TropicalAndes_IUCNHabitat_Forest <- st_read(file.path(data_path, "TropicalAndes_IUCNHabitat_Forest.shp"))
+TA_countries <- c("Bolivia", "Ecuador", "Venezuela", "Colombia", "Peru")
 ```
 
-# Get occurrence data for Tropical Andes from BIEN database
+## Get occurrence data from BIEN
 ```{r}

### 2023-11-01 code/L0/L0_TropicalAndes_plants_GIFT.Rmd

@@
 ```{r}
-# convert numberic traits into characters so we can combine into one dataframe
+# convert numeric traits into characters so we can combine into one dataframe
 height_mean$trait_value <- as.character(height_mean$trait_value)

### 2023-11-01 code/L1/L1_P4_TropicalAndes_plant_imputetraits.Rmd

--- 2023-10-04 code/L1/L1_P3_TropicalAndes_plant_traits.Rmd
@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: "This script subsets plant trait data by synonyms species list and combine."
+overview: "This script fill plant trait gaps with imputation."
 data input: "ADD"

### 2023-11-01 archive/L0_NorthAndes_plants_BIEN.Rmd

---
title: "Northern Andes plant species BIEN data"
authors: "Hazel J. Anderson"
collaborators: "None"
data input: "none"
data output: "NorthAndes_BIEN_species_list.csv, NorthAndes_BIEN_occurrence.csv, NorthAndes_BIEN_traits.csv"
project: "Frugivoria"
date: "2023-02-13"
output: html_document
---
# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(BIEN)
library(tidyr)
library(dplyr)
library(DataExplorer)
library(summarytools)
library(bdvis)
library(visdat)
```

# Read in IUCN data
Data was obtained from https://www.iucnredlist.org/search?permalink=9621ab77-2341-4bf8-9557-b7f19f7f2a35
```{r}
plants_habitats <- read.csv(file.path(data_path,"IUCN/Tropical Andes Plants - Search Results/habitats.csv"))
```

# Create species list from IUCN data
```{r}
plant_species <- unique(plants_habitats$scientificName)
```

# List of countries of interest
```{r}
country_vector <- c("Ecuador", "Venezuela", "Peru", "Colombia", "Bolivia")
```

# Get data for Tropical Andes from BIEN database
```{r}
# occurrence data
# old code -- TropicalAndes_BIEN_occ <- BIEN_occurrence_country(country_vector, all.taxonomy = TRUE, collection.info = TRUE, political.boundaries = TRUE, native.status = TRUE)
TropicalAndes_BIEN_occ <- BIEN_occurrence_box(min.lat = -94, max.lat = -54, min.long = -24, max.long = 14,species = plant_species, all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE, observation.type = TRUE)

# species list
TropicalAndes_BIEN_species_list <- BIEN_list_country(country = country_vector)

# trait data
TropicalAndes_BIEN_traits <- BIEN_trait_country(country_vector, all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
```

# Data summary
```{r}
glimpse(TropicalAndes_BIEN_occ)
```

```{r}
glimpse(TropicalAndes_BIEN_species_list)
```

```{r}
glimpse(TropicalAndes_BIEN_traits)
```

# Write data to csv
```{r}
TropicalAndes_BIEN_occurrence_path <- file.path(output_path,"TropicalAndes_BIEN_occurrence.csv") 
write.csv(TropicalAndes_BIEN_occ, file = TropicalAndes_BIEN_occurrence_path)

TropicalAndes_BIEN_species_list_path <- file.path(output_path, "TropicalAndes_BIEN_species_list.csv")
write.csv(TropicalAndes_BIEN_species_list, file = TropicalAndes_BIEN_species_list_path)

TropicalAndes_BIEN_traits_path <- file.path(output_path, "TropicalAndes_BIEN_traits.csv")
write.csv(TropicalAndes_BIEN_traits, file = TropicalAndes_BIEN_traits_path)
```

### 2023-11-01 code/L1/L1_F2_TropicalAndes_frugivore_traits.Rmd

@@
 data input: "TropicalAndes_Frugivoria_frugivore_traits.csv", "TropicalAndes_GBIF_frugivore_occ_cleaned.csv"
-data output: ""
-date: "2023-08-03"
+data output: "TropicalAndes_Frugivoria_traits_subset.csv"
+date: "2023-10-31"
 output: html_document
@@
 ```{r}
-source(file.path(data_path_L0, "data_summary.R"))
+source("~/GitHub/neotropical_plants/code/L0/data_summary.R")
 ```
@@
 ```{r}
-write.csv(Frugivoria_subset, file.path(output_path,"TropicalAndes_Frugivoria_traits_Forest.csv"))
+write.csv(Frugivoria_subset, file.path(output_path,"TropicalAndes_Frugivoria_traits_subset.csv"))
 ```

### 2023-11-01 code/L1/L1_P1_TropicalAndes_plants_GBIF.Rmd

--- 2023-08-02 code/L1/L1_TropicalAndes_plants_GBIF.Rmd
@@
 
-```{r}
-table(TropicalAndes_GBIF_plant_occ_cleaned$year)
-```
-
-```{r}
-# remove records from before second world war
-TropicalAndes_GBIF_plant_occ_cleaned <- TropicalAndes_GBIF_plant_occ_cleaned %>%
-  filter(year > 1945) 
-```
-
 # Summary
@@
 length(unique(TropicalAndes_GBIF_plant_occ_cleaned$species))
+print("The number of genera is" )
+length(unique(TropicalAndes_GBIF_plant_occ_cleaned$genus))
 print("The number of families is")

### 2023-11-01 code/L2/L2_1_TropicalAndes_taxdiv.Rmd

---
title: "Tropical Andes Taxonomic Diversity of Plants and Frugivores"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Calculating and mapping taxonomic diversity as species richness using occurrence data for plants and frugivores in the Tropical Andes Moist Lowland and Montane forests. Note: some code is adapted from https://luisdva.github.io/rstats/richness/."
data input: "TropicalAndes_plant_occ.csv, TropicalAndes_frugivore_occ.csv"
data output: "TropicalAndes_frugivore_richness.csv, TropicalAndes_GBIF_plant_richness.csv"
date: "2023-08-01"
output: html_document
---

# Load required packages
```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(smoothr)
library(purrr)
library(raster)
library(scico)
library(ggspatial)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
```

# Read in Data
```{r}
TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_frugivore_occ_cleaned.csv"))
TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
```

# Polygons of countries
```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
#polygon of Tropical Andes
TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
```

```{r}
TApoly <- st_transform(TApoly, crs = )
```

# Convert data to spatial data
```{r}
plants.sf <- st_as_sf(TropicalAndes_plant_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)

frugivores.sf <- st_as_sf(TropicalAndes_frugivore_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
```

```{r}
plants_sf_species <- plants.sf %>%
  group_by(species) %>%
  summarise()

frugivores_sf_species <- frugivores.sf %>%
  group_by(species) %>%
  summarise()
```

# Plot Base Map
```{r}
basePlot <-
ggplot() +
geom_sf(data = worldMap, fill = "white") +
geom_sf(data = TApoly) +
geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "forestgreen")+
labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
annotation_scale(location = "bl", width_hint = 0.5) +
annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
theme(panel.background = element_rect(fill = "lightblue"))

basePlot
```

# Plot points
```{r}
plantsPointsPlot <-
  ggplot() +
  geom_sf(data = worldMap, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = plants_sf_species, pch = 21) +
  labs(title = "Tropical Andes Plant Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  )
plantsPointsPlot
```

```{r}
frugivoresPointsPlot <-
  ggplot() +
  geom_sf(data = worldMap, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = frugivores_sf_species, pch = 21) +
  labs(title = "Tropical Andes Frugivore Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  )
frugivoresPointsPlot
```

# Make 0.5 Grid
```{r}
# make grid
TAGrid0.5 <- TApoly %>%
  st_make_grid(cellsize = c(0.5)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())
```

# Plot Empty Grid
```{r}
# empty grid
gridPlot <-
  ggplot() +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
  geom_sf(data = TAGrid0.5) +
  labs(title = "Tropical Andes Grid [0.5°]", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  )
gridPlot
```

# Cell Richness
```{r}
plant_richness_grid0.5 <- TAGrid0.5 %>%
  st_join(plants_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

```{r}
frugivore_richness_grid0.5 <- TAGrid0.5 %>%
  st_join(frugivores_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

# Richness plots
```{r}
plantgridRichTA0.5 <-
  ggplot(plant_richness_grid0.5) +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  labs(title = "Tropical Andes Plant Raw Richness", x = "Longitude", y = "Latitude", fill = "Number of species")
plantgridRichTA0.5
```

```{r}
frugivoregridRichTA0.5 <-
  ggplot(frugivore_richness_grid0.5) +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  ) +   
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  labs(title = "Tropical Andes Frugivore Raw Richness", x = "Longitude", y = "Latitude", fill = "Number of species")
frugivoregridRichTA0.5
```

# Extract cell values
```{r}
plant_cellRichness0.5 <- plantgridRichTA0.5[["data"]]
frugivore_cellRichness0.5 <- frugivoregridRichTA0.5[["data"]]
```

```{r}
plot(plant_cellRichness0.5$num_species, frugivore_cellRichness0.5$num_species,
     main = "Plant Richness vs Frugivore Richness",
     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
abline(lm(frugivore_cellRichness0.5$num_species ~ plant_cellRichness0.5$num_species))
summary(lm(frugivore_cellRichness0.5$num_species ~ plant_cellRichness0.5$num_species))
```

```{r}
hist(plant_cellRichness0.5$num_species,
     main = "Histogram of Plant Richness by cell",
     xlab = "Plant Richness by cell")
hist(frugivore_cellRichness0.5$num_species,
     main = "Histogram of Frugivore Richness by cell",
     xlab = "Frugivore Richness by cell")
```

```{r}
summary(plant_cellRichness0.5$num_species)
summary(frugivore_cellRichness0.5$num_species)
```

# Write data to csv
```{r}
write.csv(plant_cellRichness0.75, file.path(output_path,"TropicalAndes_GBIF_plant_richness_sub.csv"))
write.csv(frugivore_cellRichness0.75, file.path(output_path,"TropicalAndes_frugivore_richness_sub.csv"))
```

### 2023-11-01 code/L2/L2_TropicalAndes_plant_taxcoverage.Rmd

@@
 ```{r}
-library(WorldFlora)
-library(taxize)
-library(BIEN)
-library(GIFT)
 library(dplyr)
@@
 ```{r}
-BIEN_traits <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_traits_forest.csv"))
-TRY_traits <- read.csv(file.path(data_path,"TropicalAndes_TRY_plant_traits_forestSpecies.csv"))
-GIFT_traits <- read.csv(file.path(data_path,"TropicalAndes_GIFT_plant_traits_forest.csv"))
-GBIF_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ.csv"))
+GBIF_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
+GIFT_checklist <- read.csv(file.path(data_path,"TropicalAndes_plant_checklist.csv"))
+```
+
+# Plot coverage
+GIFT checklist 
+species 70376
+genera  4381
+families 273
+
+GBIF occ
+species 46,073
+genera 4,885
+families 581
+
+```{r}
+checklist <- c(70376, 4381, 273)
+occ <- c(46073, 4885, 581)
+compare_coverage <- data.frame(checklist, occ, row.names = c("species", "genera", "families"))
 ```
@@
 ```{r}
-#species list
-BIEN_species <- unique(BIEN_traits$scrubbed_species_binomial)
-TRY_species <- unique(TRY_traits$SpeciesName)
-GIFT_species <- unique(GIFT_traits$species)
-GBIF_species <- unique(GBIF_occ$species)
+barplot(species, df = compare_coverage)
 ```
-
-# Get higher taxonomy for species lists
-```{r}
-BIEN_tax<- BIEN_taxonomy_species(BIEN_species)
-```
-
-```{r}
-TRY_tax <- tax_name(TRY_species, get = c('family', 'genus'), db= "itis")
-```
-
-```{r}
-gift_sp <- GIFT_species()
-sapply(GIFT_species, function(x) grep(x, gift_sp$work_species))
-gift_sp[sapply(GIFT_species, function(x) grep(x, gift_sp$work_species)), ]
-
-#GIFT_tax <- GIFT_taxgroup(taxon_lvl = c("family", "genus")[1], species = GIFT_species)
-```
-
-# Plot coverage
-```{r}
-BIEN_family_coverage <- BIEN_tax %>% count(scrubbed_family)
-BIEN_family_coverage 
-BIEN_genus_coverage <- BIEN_tax %>% count(scrubbed_genus) 
-BIEN_genus_coverage
-barplot(height = BIEN_family_coverage$n, names = BIEN_family_coverage$scrubbed_family, horiz=T, las=1)
-barplot(height = BIEN_genus_coverage$n, names = BIEN_genus_coverage$scrubbed_genus, horiz=T, las=1)
-```

### 2023-11-01 code/L2/L2_TropicalAndes_plant_traitcoverage.Rmd

--- 2023-08-04 code/L2/L2_1_TropicalAndes_plant_traits.Rmd
@@
 ---
-title: "Tropical Andes plant traits"
+title: "Trait coverage for plant data"
 author: "Hazel J. Anderson"
@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: "Combinging TRY and BIEN plant trait data"
-data input: ""
-data output: ""
-date: "2023-08-03"
+overview: "This script summarize trait coverage and gaps of plant data used."
+data input: "ADD"
+data output: "ADD"
+date: "2023-10-04"
 output: html_document
 ---
-
-# Load required packages
-```{r}
-library(dplyr)
-library(tidyr)
-library(data.table)
-```
 
@@
 
+# Load required packages
+```{r}
+
+```
+
 # Read in Data
 ```{r}
-BIEN_traits <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_traits_forest.csv"))
-TRY_traits <- read.csv(file.path(data_path,"TropicalAndes_TRY_plant_traits_forestSpecies.csv"))
-GIFT_traits <- read.csv(file.path(data_path,"TropicalAndes_GIFT_plant_traits_forest.csv"))
+
 ```
-
-# Subset to columns of interest
-```{r}
-# remove TRY columns without value in TraitID
-TRY_traits_2 <- TRY_traits[!is.na(TRY_traits$TraitID), ]
-```
-
-```{r}
-BIEN_traits_sub <- BIEN_traits[ , c("scrubbed_species_binomial", "trait_name", "trait_value", "unit")]
-TRY_traits_sub <- TRY_traits_2[ , c("AccSpeciesName", "TraitName", "OrigValueStr", "OrigUnitStr")]
-```
-
-# Standardize column names
-```{r}
-BIEN_traits_stan <- BIEN_traits_sub %>% rename(species = scrubbed_species_binomial)
-TRY_traits_stan <- TRY_traits_sub %>% rename(species = AccSpeciesName, trait_name = TraitName, trait_value = OrigValueStr, unit = OrigUnitStr)
-```
-
-```{r}
-plant_traits <- dplyr::bind_rows(BIEN_traits_stan, TRY_traits_stan, GIFT_traits, .id = 'source')
-```
-
-```{r}
-#replace source 1 or 2 with BIEN or TRY
-plant_traits['source'][plant_traits['source'] == 1] <- "BIEN"
-plant_traits['source'][plant_traits['source'] == 2] <- "TRY"
-plant_traits['source'][plant_traits['source'] == 3] <- "GIFT"
-```
-
-# Summary
-```{r}
-glimpse(plant_traits)
-```
-
-```{r}
-plant_traits %>%
-  count(trait_name)
-```
-
-# Dispersal Traits
-```{r}
-# subset plant traits to just dispersal traits
-plant_traits_dispersal <- plant_traits[plant_traits$trait_name == "Dispersal syndrome" | plant_traits$trait_name == "whole plant dispersal syndrome" | plant_traits$trait_name == "dispersal_syndrome",]
-# look at possible trait values
-plant_traits_dispersal_values <- unique(plant_traits_dispersal$trait_value)
-# save list to manually classify biotic, abiotic, none, and NA
-write.csv(plant_traits_dispersal_values, file.path(data_path,"plant_trait_dispersal_values.csv"))
-plant_traits_dispersal %>%
-  count(trait_value)
-```
-
-```{r}
-# read in classified table
-dispersal_classified <- read.csv(file.path(data_path,"plant_trait_dispersal_values_classified.csv"))
-# reclassify dispersal trait value to table
-plant_traits_dispersal$trait_value <- 
-dispersal_classified$Classification[match(plant_traits_dispersal$trait_value,dispersal_classified$OrigValue )]
-# check to make sure it worked
-unique(plant_traits_dispersal$trait_value)
-```
-
-```{r}
-# change all trait_name to dispersal_syndrome
-plant_traits_dispersal['trait_name'][plant_traits_dispersal['trait_name'] == 'Dispersal syndrome'] <- 'dispersal_syndrome'
-plant_traits_dispersal['trait_name'][plant_traits_dispersal['trait_name'] == 'whole plant dispersal syndrome'] <- 'dispersal_syndrome'
-#need to remove none & NA
-plant_traits_dispersal <- plant_traits_dispersal[plant_traits_dispersal$trait_value == "biotic" | plant_traits_dispersal$trait_value == "abiotic",]
-#need one value per species
-plant_traits_dispersal <- plant_traits_dispersal %>% distinct(species, .keep_all = TRUE)
-plant_traits_dispersal_wide <- pivot_wider(data = plant_traits_dispersal, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = list)
-#ready to combine
-```
-
-# Plant fruit mass traits
-```{r}
-plant_traits_mass <- plant_traits[plant_traits$trait_name == "Fruit dry mass",]
-plant_traits_mass <- plant_traits_mass[plant_traits_mass$unit == "mg",]
-plant_traits_mass$trait_value <- as.numeric(plant_traits_mass$trait_value)
-plant_traits_mass_wide <- pivot_wider(data = plant_traits_mass, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = mean)
-# ready to combine
-```
-
-# Plant height traits
-```{r}
-plant_traits_height <- plant_traits[plant_traits$trait_name == "Plant height vegetative" |plant_traits$trait_name == "whole plant height" |plant_traits$trait_name == "plant_height" ,]
-plant_traits_height$trait_value <- as.numeric(plant_traits_height$trait_value)
-plant_traits_height_wide <- pivot_wider(data = plant_traits_height, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = mean)
-plant_traits_height_wide$plant_height <- rowMeans(plant_traits_height_wide[,2:3], na.rm = TRUE)
-# ready to combine
-```
-
-# Plant longevity traits
-```{r}
-plant_traits_longevity <- plant_traits[plant_traits$trait_name == "Plant lifespan (longevity)" | plant_traits$trait_name == "maximum whole plant longevity" | plant_traits$trait_name == "plant_lifespan",]
-# change all trait_name to plant lifespan
-plant_traits_longevity['trait_name'][plant_traits_longevity['trait_name'] == 'Plant lifespan (longevity)'] <- 'plant_lifespan'
-plant_traits_longevity['trait_name'][plant_traits_longevity['trait_name'] == 'maximum whole plant longevity'] <- 'plant_lifespan'
-
-#remove non-numeric values
-plant_traits_longevity <- plant_traits_longevity[!is.na(as.numeric(as.character(plant_traits_longevity$trait_value))),]
-```
-
-```{r}
-plant_traits_longevity$trait_value <- as.numeric(plant_traits_longevity$trait_value)
-plant_traits_longevity_wide <- pivot_wider(data = plant_traits_longevity, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = mean)
-```
-
-# Combine all traits
-```{r}
-# convert all to dataframes
-plant_traits_height_wide <- as.data.frame(plant_traits_height_wide)
-plant_traits_dispersal_wide <- as.data.frame(plant_traits_dispersal_wide)
-plant_traits_mass_wide <- as.data.frame(plant_traits_mass_wide)
-plant_traits_longevity_wide <- as.data.frame(plant_traits_longevity_wide)
-
-plant_traits_wide <- merge(x = plant_traits_height_wide, y = plant_traits_dispersal_wide, by = "species", all = TRUE)
-plant_traits_wide <- merge(x = plant_traits_wide, y = plant_traits_mass_wide, by = "species", all = TRUE)
-plant_traits_wide <- merge(x = plant_traits_wide, y = plant_traits_longevity_wide, by = "species", all = TRUE)
-```
-
-```{r}
-plant_traits_wide <- plant_traits_wide [, c("species", "plant_height", "dispersal_syndrome", "plant_lifespan")]
-# remove species with NAs for one or more traits
-plant_traits_wide_complete <- na.omit(plant_traits_wide)
-```
-
-#Summary
-```{r}
-glimpse(plant_traits_wide_complete)
-```
-
-# Write data to csv
-```{r}
-write.csv(plant_traits, file.path(output_path,"TropicalAndes_plant_traits_forest.csv"))
-```

### 2023-11-01 archive/L1_NorthAndes_plants_BIEN_traits.Rmd

--- 2022-12-02 code/L1/L1_Ecuador_plants_BIEN_traits.Rmd
@@
 # darwinizing a dataset
-config <- read.csv(file.path(output_path,"configuration table.csv"))
+config <- read.csv(file.path(output_path,"Ecuador_plant_BIEN_trait_ConfigurationTable.csv"))
 

### 2023-11-01 archive/L2_NorthAndes_plant_traits.Rmd

---
title: "Ecuador plant trait data merging"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "TBD"
data output: "TBD"
project: "Frugivoria"
date: "2022-11-30"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
```

# Load required packages
```{r}
library(bdc)
```

# Read in the configuration table
```{r}
metadata <-
  readr::read_csv(system.file("ADD FILE PATH HERE",
                              package = "bdc"),
                  show_col_types = FALSE)
```

# Read in BIEN and TRY trait data
```{r}
# Path to the folder containing the example datasets. For instance:
path <- data_path

# Change in the Configuration table the path to the folder in your computer containing the example datasets
metadata$fileName <-
  gsub(pattern = "https://raw.githubusercontent.com/brunobrr/bdc/master/inst/extdata/input_files/",
       replacement = path,
       x = metadata$fileName)
```

```{r}
database <-
bdc_standardize_datasets(metadata = metadata,
                         format = "csv",
                         overwrite = TRUE,
                         save_database = TRUE)
```

### 2023-11-01 code/L1/L1_F1_TropicalAndes_frugivores_GBIF.Rmd

@@
 
-```{r}
-table(TropicalAndes_GBIF_frugivore_occ_cleaned$year)
-```
-
-```{r}
-# remove records from before second world war
-TropicalAndes_GBIF_frugivore_occ_cleaned <- TropicalAndes_GBIF_frugivore_occ_cleaned %>%
-  filter(year > 1945) 
-```
-
 # Summary
@@
 length(unique(TropicalAndes_GBIF_frugivore_occ_cleaned$species))
+print("The number of genera is" )
+length(unique(TropicalAndes_GBIF_frugivore_occ_cleaned$genus))
 print("The number of families is")

### 2023-11-01 code/L1/L1_F3_TropicalAndes_frug_occ.Rmd

@@
 ---
-title: "TITLE"
+title: "Frugivore occurrence subset by species with complete trait records"
 author: "Hazel J. Anderson"
@@
 ```{r}
-data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+data_path_L1<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
 output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
@@
 ```{r}
+library(dplyr)
+#run if not installed remotes::install_github("FRBCesab/funbiogeo")
+library(funbiogeo)
+```
 
+# Read in data
+```{r}
+Frugivoria_subset <- read.csv(file.path(data_path_L1, file = "TropicalAndes_Frugivoria_traits_subset.csv"))
+frugivore_occ <- read.csv(file.path(data_path_L1, file = "TropicalAndes_GBIF_frugivore_occ_cleaned.csv"))
 ```
+
+# Check trait coverage
+```{r}
+summary(Frugivoria_subset)
+```
+```{r}
+fb_plot_number_species_by_trait(Frugivoria_subset)
+```
+```{r}
+Frugivoria_complete <- fb_filter_species_by_trait_coverage(Frugivoria_subset, threshold_traits_proportion =  1)
+```
+
+```{r}
+summary(Frugivoria_complete)
+```
+
+```{r}
+fb_plot_number_species_by_trait(Frugivoria_complete)
+```
+
+# Create list of species with complete trait records
+```{r}
+frugivoria_species <- unique(Frugivoria_complete$IUCN_species_name)
+```
+
+# Subset frugivore occurrence data by species list
+```{r}
+frugivore_occ_subset <- frugivore_occ %>%
+  filter(species %in% frugivoria_species)
+```
+
+# Summary
+```{r}
+glimpse(frugivore_occ_subset)
+```
+
+```{r}
+source("~/GitHub/neotropical_plants/code/L0/data_summary.R")
+```
+
+```{r}
+data_summary(frugivore_occ_subset, frugivore_occ_subset$species, frugivore_occ_subset$genus, frugivore_occ_subset$family)
+```
+
+# Write data to csv
+```{r}
+write.csv(frugivor_occ_subset, file.path(output_path,"TropicalAndes_GBIF_frugivore_occ_cleaned_subset.csv"))
+```

### 2023-11-01 code/L2/L2_3_TropicalAndes_divRelationships.Rmd

--- 2023-10-04 code/L2/L2_2_TropicalAndes_combineddiv.Rmd
@@
 ---
-title: "Combining taxonomic and functional diversity for plants and frugivores at various spatial grains"
+title: "Diversity relationships for plants and frugivores"
 author: "Hazel J. Anderson"
@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: "This script combines taxonomic and functional diversity per cell for various spatial grains for plants and frugivores."
+overview: "This script compares taxonomic and functional diversity relationships for plants and frugivores for various spatial grains and combined diversity for each grain."
 data input: "ADD"

### 2023-11-01 code/L2/L2_TropicalAndes_data_bias.Rmd

--- 2023-10-04 code/L2/L2_TropicalAndes_plant_frugivore_trait_alignment.Rmd
@@
 ---
-title: "Plant-frugivore trait alignment"
+title: "Bias Analysis"
 author: "Hazel J. Anderson"
@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: "This script looks at pairs of traits between plants and frugivores to test alignment."
+overview: "This script combines quantifies bias in the data for plants and frugivores using various methods."
 data input: "ADD"

### 2023-11-02 code/L1/L1_F3_TropicalAndes_frug_occ.Rmd
update header info

@@
 overview: "This script subsets frugivore occurrence data to species with complete trait coverage."
-data input: "ADD"
-data output: "ADD"
+data input: "TropicalAndes_Frugivoria_traits_subset.csv", "TropicalAndes_GBIF_frugivore_occ_cleaned.csv"
+data output: "TropicalAndes_GBIF_frugivore_occ_cleaned_subset.csv"
 date: "2023-10-04"

### 2023-11-02 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
update input filepath

@@
 TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
-TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_frugivore_occ_cleaned.csv"))
+TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_frugivore_occ_cleaned_subset.csv"))
 TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")

### 2023-11-02 code/L1/L1_P2_TropicalAndes_taxcleaning.Rmd
update file paths and input files, add new count code, update for updated files

---
title: "Tropical Andes plant Lookup Table"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: ""
data output: ""
date: "2023-07-25"
output: html_document
---

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(knitr)
library(TNRS)
```

# Read in data
```{r}
GBIF_occ <- read.csv(file.path(data_path_L1, "TropicalAndes_GBIF_plant_occ_cleaned.csv"))
TRY_traits <- read.csv(file.path(data_path_L0,"TropicalAndes_TRY_plant_traits.csv"))
BIEN_traits <- read.csv(file.path(data_path_L0,"AllDesired_BIEN_plant_traits.csv"))
GIFT_traits <- read.csv(file.path(data_path_L0,"TropicalAndes_GIFT_plant_traits.csv"))
```

# Generate species lists
```{r}
GBIF_species <- unique(GBIF_occ$species)
TRY_species <- unique(TRY_traits$SpeciesName)
BIEN_species <- unique(BIEN_traits$scrubbed_species_binomial)
GIFT_species <- unique(GIFT_traits$species)
```

```{r}
length(GBIF_species)
length(TRY_species)
length(BIEN_species)
length(GIFT_species)
```

# Using TNRS to get harmonize taxonomy
## GBIF species
```{r}
#split up to speed up api call
GBIF_species1 <-GBIF_species[1:2245]
GBIF_species2 <-GBIF_species[2245:6745]
GBIF_species3 <-GBIF_species[6745:11245]
GBIF_species4 <-GBIF_species[11245:15745]
GBIF_species5 <-GBIF_species[15745:20245]
GBIF_species6 <-GBIF_species[16553:20245]
GBIF_species7 <-GBIF_species[20245:24745]
GBIF_species8 <-GBIF_species[24745:29245]
GBIF_species9 <-GBIF_species[29245:33745]
GBIF_species10 <-GBIF_species[33745:38245]
GBIF_species11 <-GBIF_species[38245:42745]
GBIF_species12 <-GBIF_species[42745:47245]
```

```{r}
GBIF_results1 <- TNRS(taxonomic_names = GBIF_species1)
GBIF_results2 <- TNRS(taxonomic_names = GBIF_species2)
GBIF_results3 <- TNRS(taxonomic_names = GBIF_species3)
GBIF_results4 <- TNRS(taxonomic_names = GBIF_species4)
GBIF_results5 <- TNRS(taxonomic_names = GBIF_species5)
GBIF_results6 <- TNRS(taxonomic_names = GBIF_species6)
GBIF_results7 <- TNRS(taxonomic_names = GBIF_species7)
GBIF_results8 <- TNRS(taxonomic_names = GBIF_species8)
GBIF_results9 <- TNRS(taxonomic_names = GBIF_species9)
GBIF_results10 <- TNRS(taxonomic_names = GBIF_species10)
GBIF_results11 <- TNRS(taxonomic_names = GBIF_species11)
GBIF_results12 <- TNRS(taxonomic_names = GBIF_species12)
```

```{r}
#combine results
GBIF_results <- rbind(GBIF_results1, GBIF_results2, GBIF_results3, GBIF_results4, GBIF_results5, GBIF_results6, GBIF_results7, GBIF_results8, GBIF_results9, GBIF_results10, GBIF_results11, GBIF_results12)
```

## TRY species
```{r}
#split up to speed up api call
TRY_species1 <-TRY_species[1:5000]
TRY_species2 <-TRY_species[5001:10000]
TRY_species3 <-TRY_species[10001:15000]
TRY_species4 <-TRY_species[15001:20000]
TRY_species5 <-TRY_species[20001:25000]
TRY_species6 <-TRY_species[25001:30000]
TRY_species7 <-TRY_species[30001:35000]
TRY_species8 <-TRY_species[35001:40000]
TRY_species9 <-TRY_species[40001:45000]
TRY_species10 <-TRY_species[45001:50000]
TRY_species11 <-TRY_species[50001:55000]
TRY_species12 <-TRY_species[55001:60000]
TRY_species13 <-TRY_species[60001:65000]
TRY_species14 <-TRY_species[65001:70000]
TRY_species15 <-TRY_species[70001:75000]
TRY_species16 <-TRY_species[75001:80000]
TRY_species17 <-TRY_species[80001:85000]
TRY_species18 <-TRY_species[85001:90000]
TRY_species19 <-TRY_species[90001:95000]
TRY_species20 <-TRY_species[95001:100000]
TRY_species21 <-TRY_species[100001:105000]
TRY_species22 <-TRY_species[105001:110000]
TRY_species23 <-TRY_species[110001:115000]
TRY_species24 <-TRY_species[115001:120000]
TRY_species25 <-TRY_species[120001:125000]
TRY_species26 <-TRY_species[125001:130000]
TRY_species27 <-TRY_species[130001:135000]
TRY_species28 <-TRY_species[135001:140000]
TRY_species29 <-TRY_species[140001:145000]
TRY_species30 <-TRY_species[145001:150000]
TRY_species31 <-TRY_species[150001:155000]
TRY_species32 <-TRY_species[155001:160000]
TRY_species33 <-TRY_species[160001:165000]
TRY_species34 <-TRY_species[165001:170000]
TRY_species35 <-TRY_species[170001:175000]
TRY_species36 <-TRY_species[175001:180000]
TRY_species37 <-TRY_species[180001:185000]
TRY_species38 <-TRY_species[185001:190000]
TRY_species39 <-TRY_species[190001:195000]
TRY_species40 <-TRY_species[195001:200000]
TRY_species41 <-TRY_species[200001:205000]
TRY_species42 <-TRY_species[205001:210000]
TRY_species43 <-TRY_species[210001:215000]
TRY_species44 <-TRY_species[215001:220000]
TRY_species45 <-TRY_species[220001:225000]
TRY_species46 <-TRY_species[225001:230000]
TRY_species47 <-TRY_species[230001:235000]
TRY_species48 <-TRY_species[235001:240000]
TRY_species49 <-TRY_species[240001:245000]
TRY_species50 <-TRY_species[245001:250000]
TRY_species51 <-TRY_species[250001:255000]
TRY_species52 <-TRY_species[255001:260000]
TRY_species53 <-TRY_species[260001:265000]
TRY_species54 <-TRY_species[265001:270000]
TRY_species55 <-TRY_species[270001:275000]
TRY_species56 <-TRY_species[275001:280000]
TRY_species57 <-TRY_species[280001:285000]
TRY_species58 <-TRY_species[285001:290000]
TRY_species59 <-TRY_species[290001:295000]
TRY_species60 <-TRY_species[295001:300000]
TRY_species61 <-TRY_species[300001:305000]
TRY_species62 <-TRY_species[305001:310000]
TRY_species63 <-TRY_species[310001:315000]
TRY_species64 <-TRY_species[315001:320000]
TRY_species65 <-TRY_species[320001:325000]
TRY_species66 <-TRY_species[325001:330000]
TRY_species67 <-TRY_species[330001:335000]
TRY_species68 <-TRY_species[335001:340000]
TRY_species69 <-TRY_species[340001:345000]
TRY_species70 <-TRY_species[345001:350000]
TRY_species71 <-TRY_species[350001:355000]
TRY_species72 <-TRY_species[355001:360000]
TRY_species73 <-TRY_species[360001:365000]
TRY_species74 <-TRY_species[365001:370000]
TRY_species75 <-TRY_species[370001:375000]
TRY_species76 <-TRY_species[375001:380000]
TRY_species77 <-TRY_species[380001:385000]
TRY_species78 <-TRY_species[385001:390000]
TRY_species79 <-TRY_species[390001:395000]
TRY_species80 <-TRY_species[395001:400000]
```

```{r}
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results2 <- TNRS(taxonomic_names = TRY_species2)
TRY_results3 <- TNRS(taxonomic_names = TRY_species3)
TRY_results4 <- TNRS(taxonomic_names = TRY_species4)
TRY_results5 <- TNRS(taxonomic_names = TRY_species5)
TRY_results6 <- TNRS(taxonomic_names = TRY_species6)
TRY_results7 <- TNRS(taxonomic_names = TRY_species7)
TRY_results8 <- TNRS(taxonomic_names = TRY_species8)
TRY_results9 <- TNRS(taxonomic_names = TRY_species9)
TRY_results10 <- TNRS(taxonomic_names = TRY_species10)
TRY_results11 <- TNRS(taxonomic_names = TRY_species11)
TRY_results12 <- TNRS(taxonomic_names = TRY_species12)
TRY_results13 <- TNRS(taxonomic_names = TRY_species13)
TRY_results14 <- TNRS(taxonomic_names = TRY_species14)
TRY_results15 <- TNRS(taxonomic_names = TRY_species15)
TRY_results16 <- TNRS(taxonomic_names = TRY_species16)
TRY_results17 <- TNRS(taxonomic_names = TRY_species17)
TRY_results18 <- TNRS(taxonomic_names = TRY_species18)
TRY_results19 <- TNRS(taxonomic_names = TRY_species19)
TRY_results20 <- TNRS(taxonomic_names = TRY_species20)
TRY_results21 <- TNRS(taxonomic_names = TRY_species21)
TRY_results22 <- TNRS(taxonomic_names = TRY_species22)
TRY_results23 <- TNRS(taxonomic_names = TRY_species23)
TRY_results24 <- TNRS(taxonomic_names = TRY_species24)
TRY_results25 <- TNRS(taxonomic_names = TRY_species25)
TRY_results26 <- TNRS(taxonomic_names = TRY_species26)
TRY_results27 <- TNRS(taxonomic_names = TRY_species27)
TRY_results28 <- TNRS(taxonomic_names = TRY_species28)
TRY_results29 <- TNRS(taxonomic_names = TRY_species1)
TRY_results30 <- TNRS(taxonomic_names = TRY_species1)
TRY_results31 <- TNRS(taxonomic_names = TRY_species1)
TRY_results32 <- TNRS(taxonomic_names = TRY_species1)
TRY_results33 <- TNRS(taxonomic_names = TRY_species1)
TRY_results34 <- TNRS(taxonomic_names = TRY_species1)
TRY_results35 <- TNRS(taxonomic_names = TRY_species1)
TRY_results36 <- TNRS(taxonomic_names = TRY_species1)
TRY_results37 <- TNRS(taxonomic_names = TRY_species1)
TRY_results38 <- TNRS(taxonomic_names = TRY_species1)
TRY_results39 <- TNRS(taxonomic_names = TRY_species1)
TRY_results40 <- TNRS(taxonomic_names = TRY_species1)
TRY_results41 <- TNRS(taxonomic_names = TRY_species1)
TRY_results42 <- TNRS(taxonomic_names = TRY_species1)
TRY_results43 <- TNRS(taxonomic_names = TRY_species1)
TRY_results44 <- TNRS(taxonomic_names = TRY_species1)
TRY_results45 <- TNRS(taxonomic_names = TRY_species1)
TRY_results46 <- TNRS(taxonomic_names = TRY_species1)
TRY_results47 <- TNRS(taxonomic_names = TRY_species1)
TRY_results48 <- TNRS(taxonomic_names = TRY_species1)
TRY_results49 <- TNRS(taxonomic_names = TRY_species1)
TRY_results50 <- TNRS(taxonomic_names = TRY_species1)
TRY_results51 <- TNRS(taxonomic_names = TRY_species1)
TRY_results52 <- TNRS(taxonomic_names = TRY_species1)
TRY_results53 <- TNRS(taxonomic_names = TRY_species1)
TRY_results54 <- TNRS(taxonomic_names = TRY_species1)
TRY_results55 <- TNRS(taxonomic_names = TRY_species1)
TRY_results56 <- TNRS(taxonomic_names = TRY_species1)
TRY_results57 <- TNRS(taxonomic_names = TRY_species1)
TRY_results58 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
```

```{r}
#combine results
GBIF_results <- rbind(GBIF_results1, GBIF_results2, GBIF_results3, GBIF_results4, GBIF_results5, GBIF_results6, GBIF_results7, GBIF_results8, GBIF_results9, GBIF_results10, GBIF_results11, GBIF_results12)
```

## BIEN species
```{r}
BIEN_results <- TNRS(taxonomic_names = BIEN_species)
```

## GIFT species
```{r}
GIFT_results <- TNRS(taxonomic_names = GIFT_species)
```

# Save full results
```{r}
write.csv(BIEN_results, file.path(output_path,"BIEN_plant_species_resolved_output_full.csv"))
write.csv(TRY_results, file.path(output_path,"TRY_plant_species_resolved_output_full.csv"))
write.csv(GBIF_results, file.path(output_path,"GBIF_plant_species_resolved_output_full.csv"))
write.csv(GIFT_results, file.path(output_path,"GIFT_plant_species_resolved_output_full.csv"))
```

# Subset results to species with matches, conflicts, and disparties
Name_submitted values not the same as Accepted_name
```{r}
GBIF_matches <- subset(GBIF_results, Taxonomic_status == "Accepted")
nrow(GBIF_matches)
GBIF_conflicts <- subset(GBIF_results, Taxonomic_status == "Synonym")
nrow(GBIF_conflicts)
GBIF_disparities <- subset(GBIF_results, Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", ""))
nrow(GBIF_disparities)
```

```{r}
TRY_matches <- subset(TRY_results, Taxonomic_status == "Accepted")
nrow(TRY_matches)
TRY_conflicts <- subset(TRY_results, Taxonomic_status == "Synonym")
nrow(TRY_conflicts)
TRY_disparities <- subset(TRY_results, Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", ""))
nrow(TRY_disparities)
```

```{r}
BIEN_matches <- subset(BIEN_results, Taxonomic_status == "Accepted")
nrow(BIEN_matches)
BIEN_conflicts <- subset(BIEN_results, Taxonomic_status == "Synonym")
nrow(BIEN_conflicts)
BIEN_disparities <- subset(BIEN_results, Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", ""))
nrow(BIEN_disparities)
```

```{r}
GIFT_matches <- subset(GIFT_results, Taxonomic_status == "Accepted")
nrow(GIFT_matches)
GIFT_conflicts <- subset(GIFT_results, Taxonomic_status == "Synonym")
nrow(GIFT_conflicts)
GIFT_disparities <- subset(GIFT_results, Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", ""))
nrow(GIFT_disparities)
```

# Checking 
```{r}

```

### 2023-11-08 code/L1/L1_P2_TropicalAndes_taxcleaning.Rmd
update packages, combine species list to simplify harmonization, remove NAs from species list, add code to check common scientific name issues, add bdc to harmonize species that TNRS had disparties with, create lookup table

@@
 
+# Load required packages
+```{r}
+library(knitr)
+library(TNRS)
+library(bdc)
+library(dplyr)
+```
+
 # Set file paths
@@
 output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
-```
-
-# Load required packages
-```{r}
-library(knitr)
-library(TNRS)
 ```
@@
 ```
+## combine species list together
+```{r}
+all_species <- c(GBIF_species, TRY_species, BIEN_species, GIFT_species)
+all_species <- unique(all_species)
+length(all_species)
+```
+# Remove NAs
+```{r}
+all_species <- all_species[!is.na(all_species)]
+length(all_species)
+```
+
+# Use Rocc to check scientific names 
+Adapted from tutorial https://liibre.github.io/Rocc/articles/articles/check_taxa.html
+```{r}
+name_check_results <- bdc_clean_names(sci_names = all_species, save_outputs = FALSE)
+```
+```{r}
+all_species_cleaned <- name_check_results$names_clean
+```
 
 # Using TNRS to get harmonize taxonomy
-## GBIF species
+
 ```{r}
-#split up to speed up api call
-GBIF_species1 <-GBIF_species[1:2245]
-GBIF_species2 <-GBIF_species[2245:6745]
-GBIF_species3 <-GBIF_species[6745:11245]
-GBIF_species4 <-GBIF_species[11245:15745]
-GBIF_species5 <-GBIF_species[15745:20245]
-GBIF_species6 <-GBIF_species[16553:20245]
-GBIF_species7 <-GBIF_species[20245:24745]
-GBIF_species8 <-GBIF_species[24745:29245]
-GBIF_species9 <-GBIF_species[29245:33745]
-GBIF_species10 <-GBIF_species[33745:38245]
-GBIF_species11 <-GBIF_species[38245:42745]
-GBIF_species12 <-GBIF_species[42745:47245]
+all_species_hamonized_results <- TNRS(taxonomic_names = all_species_cleaned)
+```
+
+# Subset TNRS results to species with matches, conflicts, and disparties
+```{r}
+unique(all_species_hamonized_results$Taxonomic_status)
 ```
@@
 ```{r}
-GBIF_results1 <- TNRS(taxonomic_names = GBIF_species1)
-GBIF_results2 <- TNRS(taxonomic_names = GBIF_species2)
-GBIF_results3 <- TNRS(taxonomic_names = GBIF_species3)
-GBIF_results4 <- TNRS(taxonomic_names = GBIF_species4)
-GBIF_results5 <- TNRS(taxonomic_names = GBIF_species5)
-GBIF_results6 <- TNRS(taxonomic_names = GBIF_species6)
-GBIF_results7 <- TNRS(taxonomic_names = GBIF_species7)
-GBIF_results8 <- TNRS(taxonomic_names = GBIF_species8)
-GBIF_results9 <- TNRS(taxonomic_names = GBIF_species9)
-GBIF_results10 <- TNRS(taxonomic_names = GBIF_species10)
-GBIF_results11 <- TNRS(taxonomic_names = GBIF_species11)
-GBIF_results12 <- TNRS(taxonomic_names = GBIF_species12)
+all_species_matches <- subset(all_species_hamonized_results, Taxonomic_status == "Accepted")
+nrow(all_species_matches)
+all_species_conflicts <- subset(all_species_hamonized_results, Taxonomic_status == "Synonym")
+nrow(all_species_conflicts)
+
+# Disparties
+all_species_no.opinion <- subset(all_species_hamonized_results, Taxonomic_status == "No opinion")
+nrow(all_species_no.opinion)
+all_species_illegitimate <- subset(all_species_hamonized_results, Taxonomic_status == "Illegitimate")
+nrow(all_species_illegitimate)
+all_species_invalid <- subset(all_species_hamonized_results, Taxonomic_status == "Invalid")
+nrow(all_species_invalid)
+all_species_other <- subset(all_species_hamonized_results, Taxonomic_status == "")
+nrow(all_species_other)
+```
+# Create species list of names with disparities
+```{r}
+all_species_disparities <- all_species_hamonized_results %>%
+                            filter(Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", "")) %>%
+                            select(Name_submitted)
+```
+
+# Use bdc to harmonize disparities
+## Using ITIS database
+```{r}
+all_species_disparities_col_results <- bdc_query_names_taxadb(sci_name = all_species_disparities$Name_submitted, replace_synonyms = TRUE,suggest_names = TRUE, db = "col", rank_name = "Plantae", rank = "kingdom")
+```
+
+## Subset COL results to accepted, synonym, and NA
+
+```{r}
+unique(all_species_disparities_col_results$taxonomicStatus)
 ```
@@
 ```{r}
-#combine results
-GBIF_results <- rbind(GBIF_results1, GBIF_results2, GBIF_results3, GBIF_results4, GBIF_results5, GBIF_results6, GBIF_results7, GBIF_results8, GBIF_results9, GBIF_results10, GBIF_results11, GBIF_results12)
+all_species_col_accepted <- subset(all_species_disparities_col_results, taxonomicStatus == "accepted")
+nrow(all_species_col_accepted)
+all_species_col_synonym <- subset(all_species_disparities_col_results, taxonomicStatus == "synonym")
+nrow(all_species_col_synonym)
+all_species_col_NA <- subset(all_species_disparities_col_results, is.na(taxonomicStatus))
+nrow(all_species_col_NA)
 ```
 
-## TRY species
+# Create lookup table
+## From TNRS
 ```{r}
-#split up to speed up api call
-TRY_species1 <-TRY_species[1:5000]
-TRY_species2 <-TRY_species[5001:10000]
-TRY_species3 <-TRY_species[10001:15000]
-TRY_species4 <-TRY_species[15001:20000]
-TRY_species5 <-TRY_species[20001:25000]
-TRY_species6 <-TRY_species[25001:30000]
-TRY_species7 <-TRY_species[30001:35000]
-TRY_species8 <-TRY_species[35001:40000]
-TRY_species9 <-TRY_species[40001:45000]
-TRY_species10 <-TRY_species[45001:50000]
-TRY_species11 <-TRY_species[50001:55000]
-TRY_species12 <-TRY_species[55001:60000]
-TRY_species13 <-TRY_species[60001:65000]
-TRY_species14 <-TRY_species[65001:70000]
-TRY_species15 <-TRY_species[70001:75000]
-TRY_species16 <-TRY_species[75001:80000]
-TRY_species17 <-TRY_species[80001:85000]
-TRY_species18 <-TRY_species[85001:90000]
-TRY_species19 <-TRY_species[90001:95000]
-TRY_species20 <-TRY_species[95001:100000]
-TRY_species21 <-TRY_species[100001:105000]
-TRY_species22 <-TRY_species[105001:110000]
-TRY_species23 <-TRY_species[110001:115000]
-TRY_species24 <-TRY_species[115001:120000]
-TRY_species25 <-TRY_species[120001:125000]
-TRY_species26 <-TRY_species[125001:130000]
-TRY_species27 <-TRY_species[130001:135000]
-TRY_species28 <-TRY_species[135001:140000]
-TRY_species29 <-TRY_species[140001:145000]
-TRY_species30 <-TRY_species[145001:150000]
-TRY_species31 <-TRY_species[150001:155000]
-TRY_species32 <-TRY_species[155001:160000]
-TRY_species33 <-TRY_species[160001:165000]
-TRY_species34 <-TRY_species[165001:170000]
-TRY_species35 <-TRY_species[170001:175000]
-TRY_species36 <-TRY_species[175001:180000]
-TRY_species37 <-TRY_species[180001:185000]
-TRY_species38 <-TRY_species[185001:190000]
-TRY_species39 <-TRY_species[190001:195000]
-TRY_species40 <-TRY_species[195001:200000]
-TRY_species41 <-TRY_species[200001:205000]
-TRY_species42 <-TRY_species[205001:210000]
-TRY_species43 <-TRY_species[210001:215000]
-TRY_species44 <-TRY_species[215001:220000]
-TRY_species45 <-TRY_species[220001:225000]
-TRY_species46 <-TRY_species[225001:230000]
-TRY_species47 <-TRY_species[230001:235000]
-TRY_species48 <-TRY_species[235001:240000]
-TRY_species49 <-TRY_species[240001:245000]
-TRY_species50 <-TRY_species[245001:250000]
-TRY_species51 <-TRY_species[250001:255000]
-TRY_species52 <-TRY_species[255001:260000]
-TRY_species53 <-TRY_species[260001:265000]
-TRY_species54 <-TRY_species[265001:270000]
-TRY_species55 <-TRY_species[270001:275000]
-TRY_species56 <-TRY_species[275001:280000]
-TRY_species57 <-TRY_species[280001:285000]
-TRY_species58 <-TRY_species[285001:290000]
-TRY_species59 <-TRY_species[290001:295000]
-TRY_species60 <-TRY_species[295001:300000]
-TRY_species61 <-TRY_species[300001:305000]
-TRY_species62 <-TRY_species[305001:310000]
-TRY_species63 <-TRY_species[310001:315000]
-TRY_species64 <-TRY_species[315001:320000]
-TRY_species65 <-TRY_species[320001:325000]
-TRY_species66 <-TRY_species[325001:330000]
-TRY_species67 <-TRY_species[330001:335000]
-TRY_species68 <-TRY_species[335001:340000]
-TRY_species69 <-TRY_species[340001:345000]
-TRY_species70 <-TRY_species[345001:350000]
-TRY_species71 <-TRY_species[350001:355000]
-TRY_species72 <-TRY_species[355001:360000]
-TRY_species73 <-TRY_species[360001:365000]
-TRY_species74 <-TRY_species[365001:370000]
-TRY_species75 <-TRY_species[370001:375000]
-TRY_species76 <-TRY_species[375001:380000]
-TRY_species77 <-TRY_species[380001:385000]
-TRY_species78 <-TRY_species[385001:390000]
-TRY_species79 <-TRY_species[390001:395000]
-TRY_species80 <-TRY_species[395001:400000]
+species_matches <- all_species_matches[ , c("Name_submitted", "Name_matched", "Source")]
+species_corrected <- all_species_conflicts[ , c("Name_submitted", "Name_matched", "Source")]
+```
+
+## From bdc using COL
+```{r}
+species_col_accepted <- all_species_col_accepted[ , c("original_search", "suggested_name")]
+species_col_accepted$source <- c("col")
+species_col_synonmyn <- all_species_col_synonym[ , c("original_search", "suggested_name")]
+species_col_synonmyn$source <- c("col")
 ```
@@
 ```{r}
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results2 <- TNRS(taxonomic_names = TRY_species2)
-TRY_results3 <- TNRS(taxonomic_names = TRY_species3)
-TRY_results4 <- TNRS(taxonomic_names = TRY_species4)
-TRY_results5 <- TNRS(taxonomic_names = TRY_species5)
-TRY_results6 <- TNRS(taxonomic_names = TRY_species6)
-TRY_results7 <- TNRS(taxonomic_names = TRY_species7)
-TRY_results8 <- TNRS(taxonomic_names = TRY_species8)
-TRY_results9 <- TNRS(taxonomic_names = TRY_species9)
-TRY_results10 <- TNRS(taxonomic_names = TRY_species10)
-TRY_results11 <- TNRS(taxonomic_names = TRY_species11)
-TRY_results12 <- TNRS(taxonomic_names = TRY_species12)
-TRY_results13 <- TNRS(taxonomic_names = TRY_species13)
-TRY_results14 <- TNRS(taxonomic_names = TRY_species14)
-TRY_results15 <- TNRS(taxonomic_names = TRY_species15)
-TRY_results16 <- TNRS(taxonomic_names = TRY_species16)
-TRY_results17 <- TNRS(taxonomic_names = TRY_species17)
-TRY_results18 <- TNRS(taxonomic_names = TRY_species18)
-TRY_results19 <- TNRS(taxonomic_names = TRY_species19)
-TRY_results20 <- TNRS(taxonomic_names = TRY_species20)
-TRY_results21 <- TNRS(taxonomic_names = TRY_species21)
-TRY_results22 <- TNRS(taxonomic_names = TRY_species22)
-TRY_results23 <- TNRS(taxonomic_names = TRY_species23)
-TRY_results24 <- TNRS(taxonomic_names = TRY_species24)
-TRY_results25 <- TNRS(taxonomic_names = TRY_species25)
-TRY_results26 <- TNRS(taxonomic_names = TRY_species26)
-TRY_results27 <- TNRS(taxonomic_names = TRY_species27)
-TRY_results28 <- TNRS(taxonomic_names = TRY_species28)
-TRY_results29 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results30 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results31 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results32 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results33 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results34 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results35 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results36 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results37 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results38 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results39 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results40 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results41 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results42 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results43 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results44 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results45 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results46 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results47 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results48 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results49 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results50 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results51 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results52 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results53 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results54 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results55 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results56 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results57 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results58 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
-TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
+# combine to create lookup table
+lookup_table <- rbind(species_matches, species_corrected)
 ```
 
-```{r}
-#combine results
-GBIF_results <- rbind(GBIF_results1, GBIF_results2, GBIF_results3, GBIF_results4, GBIF_results5, GBIF_results6, GBIF_results7, GBIF_results8, GBIF_results9, GBIF_results10, GBIF_results11, GBIF_results12)
-```
-
-## BIEN species
-```{r}
-BIEN_results <- TNRS(taxonomic_names = BIEN_species)
-```
-
-## GIFT species
-```{r}
-GIFT_results <- TNRS(taxonomic_names = GIFT_species)
-```
-
-# Save full results
-```{r}
-write.csv(BIEN_results, file.path(output_path,"BIEN_plant_species_resolved_output_full.csv"))
-write.csv(TRY_results, file.path(output_path,"TRY_plant_species_resolved_output_full.csv"))
-write.csv(GBIF_results, file.path(output_path,"GBIF_plant_species_resolved_output_full.csv"))
-write.csv(GIFT_results, file.path(output_path,"GIFT_plant_species_resolved_output_full.csv"))
-```
-
-# Subset results to species with matches, conflicts, and disparties
-Name_submitted values not the same as Accepted_name
-```{r}
-GBIF_matches <- subset(GBIF_results, Taxonomic_status == "Accepted")
-nrow(GBIF_matches)
-GBIF_conflicts <- subset(GBIF_results, Taxonomic_status == "Synonym")
-nrow(GBIF_conflicts)
-GBIF_disparities <- subset(GBIF_results, Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", ""))
-nrow(GBIF_disparities)
-```
-
-```{r}
-TRY_matches <- subset(TRY_results, Taxonomic_status == "Accepted")
-nrow(TRY_matches)
-TRY_conflicts <- subset(TRY_results, Taxonomic_status == "Synonym")
-nrow(TRY_conflicts)
-TRY_disparities <- subset(TRY_results, Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", ""))
-nrow(TRY_disparities)
-```
-
-```{r}
-BIEN_matches <- subset(BIEN_results, Taxonomic_status == "Accepted")
-nrow(BIEN_matches)
-BIEN_conflicts <- subset(BIEN_results, Taxonomic_status == "Synonym")
-nrow(BIEN_conflicts)
-BIEN_disparities <- subset(BIEN_results, Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", ""))
-nrow(BIEN_disparities)
-```
-
-```{r}
-GIFT_matches <- subset(GIFT_results, Taxonomic_status == "Accepted")
-nrow(GIFT_matches)
-GIFT_conflicts <- subset(GIFT_results, Taxonomic_status == "Synonym")
-nrow(GIFT_conflicts)
-GIFT_disparities <- subset(GIFT_results, Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", ""))
-nrow(GIFT_disparities)
-```
-
-# Checking 
+# Save lookup table as csv
 ```{r}
@@
 ```
+
+# Update data files species names using lookup table
+```{r}
+
+```
+
+# Write data to csv
+```{r}
+
+```

### 2023-11-08 code/L1/L1_P2_TropicalAndes_taxcleaning.Rmd
combine harmonization results to create lookup table, apply lookup table to GBIF data

@@
 
-# Use Rocc to check scientific names 
-Adapted from tutorial https://liibre.github.io/Rocc/articles/articles/check_taxa.html
+# Use bdc to check scientific names 
 ```{r}
@@
 
+# Create list of species still unharmonized
+```{r}
+unresolved_species <- all_species_col_NA$original_search
+```
+
 # Create lookup table
@@
 species_corrected <- all_species_conflicts[ , c("Name_submitted", "Name_matched", "Source")]
+```
+
+```{r}
+#combine
+lookup_table_TNRS <- rbind(species_matches, species_corrected)
+lookup_table_TNRS$harmonization_package <- c("TNRS")
 ```
@@
 ```{r}
-# combine to create lookup table
-lookup_table <- rbind(species_matches, species_corrected)
+#update col names
+colnames(species_col_accepted) <- c("Name_submitted", "Name_matched", "Source")
+colnames(species_col_synonmyn) <- c("Name_submitted", "Name_matched", "Source")
+```
+
+```{r}
+#combine
+lookup_table_bdc <- rbind(species_col_accepted, species_col_synonmyn)
+lookup_table_bdc$harmonization_package <- c("bdc")
+```
+
+## Combine to create lookup table
+```{r}
+lookup_table <- rbind(lookup_table_TNRS, lookup_table_bdc)
+nrow(lookup_table)
 ```
@@
 ```{r}
-
+write.csv(lookup_table, file.path(output_path,"plant_lookup_table.csv"))
 ```
@@
 # Update data files species names using lookup table
+## GBIF
+```{r}
+# filter data to species in lookup_table$Names_submitted
+GBIF_occ_subset <- GBIF_occ %>%
+  filter(species %in% lookup_table$Name_submitted)
+nrow(GBIF_occ_subset)
+# replace species with lookup_table$Name_matched
+GBIF_occ_subset_harmonized <- GBIF_occ_subset %>% mutate(species = replace(species, species %in% lookup_table$Name_submitted, lookup_table$Name_matched))
+nrow(GBIF_occ_subset_harmonized)
+```
+
+# Summary
 ```{r}

### 2023-11-15 code/L1/L1_P2_TropicalAndes_taxcleaning.Rmd
add nrow check code and update column names

@@
 date: "2023-07-25"
-output: html_document
+output: 
+  html_document: 
+    toc: yes
+    df_print: tibble
 ---
@@
 all_species_cleaned <- name_check_results$names_clean
+length(all_species_cleaned)
 ```
@@
 ```{r}
+nrow(all_species_hamonized_results)
 unique(all_species_hamonized_results$Taxonomic_status)
@@
                             select(Name_submitted)
+nrow(all_species_disparities)
 ```
@@
 # Use bdc to harmonize disparities
-## Using ITIS database
+## Using COL database
 ```{r}
@@
 ## Subset COL results to accepted, synonym, and NA
-
 ```{r}
@@
 ```{r}
-species_matches <- all_species_matches[ , c("Name_submitted", "Name_matched", "Source")]
-species_corrected <- all_species_conflicts[ , c("Name_submitted", "Name_matched", "Source")]
+species_matches <- all_species_matches[ , c("Name_submitted", "Accepted_species", "Source")]
+species_corrected <- all_species_conflicts[ , c("Name_submitted", "Accepted_species", "Source")]
 ```
@@
 ```{r}
-species_col_accepted <- all_species_col_accepted[ , c("original_search", "suggested_name")]
+species_col_accepted <- all_species_col_accepted[ , c("original_search", "scientificName")]
 species_col_accepted$source <- c("col")
-species_col_synonmyn <- all_species_col_synonym[ , c("original_search", "suggested_name")]
+species_col_synonmyn <- all_species_col_synonym[ , c("original_search", "scientificName")]
 species_col_synonmyn$source <- c("col")
@@
 #update col names
-colnames(species_col_accepted) <- c("Name_submitted", "Name_matched", "Source")
-colnames(species_col_synonmyn) <- c("Name_submitted", "Name_matched", "Source")
+colnames(species_col_accepted) <- c("Name_submitted", "Accepted_species", "Source")
+colnames(species_col_synonmyn) <- c("Name_submitted", "Accepted_species", "Source")
 ```
@@
 nrow(GBIF_occ_subset)
-# replace species with lookup_table$Name_matched
-GBIF_occ_subset_harmonized <- GBIF_occ_subset %>% mutate(species = replace(species, species %in% lookup_table$Name_submitted, lookup_table$Name_matched))
+# add accepted species column from lookup table based on species column
+GBIF_occ_subset_harmonized <- merge(GBIF_occ_subset, lookup_table, by.x = "species", by.y = "Name_submitted")
 nrow(GBIF_occ_subset_harmonized)
+```
+
+## TRY
+```{r}
+
+```
+
+## BIEN
+```{r}
+
+```
+
+## GIFT
+```{r}
+
 ```
@@
 ```{r}
+source("~/GitHub/neotropical_plants/functions/data_summary.R")
+```
 
+```{r}
+data_summary(GBIF_occ_subset_harmonized, GBIF_occ_subset_harmonized$species, GBIF_occ_subset_harmonized$genus, GBIF_occ_subset_harmonized$family)
 ```

### 2023-11-15 code/L1/L1_P3_TropicalAndes_plant_traits.Rmd
add code that parrallels L1_F2_TropicalAndes_frugivore_traits

@@
 data output: "ADD"
-date: "2023-10-04"
+date: "2023-11-04"
 output: html_document
@@
 
+# Load required packages
+```{r}
+library(dplyr)
+```
+
 # Set file paths
 ```{r}
-data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
-output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
 ```
 
-# Load required packages
+# Read in data
 ```{r}
+plant_occ <- read.csv(file.path(data_path_L1, "TropicalAndes_GBIF_plant_occ_cleaned.csv"))
+TRY_traits <- read.csv(file.path(data_path_L0,"TropicalAndes_TRY_plant_traits.csv"))
+BIEN_traits <- read.csv(file.path(data_path_L0,"AllDesired_BIEN_plant_traits.csv"))
+GIFT_traits <- read.csv(file.path(data_path_L0,"TropicalAndes_GIFT_plant_traits.csv"))
+```
 
+# Get species list from plant_occ data
+```{r}
+plant_species <- unique(plant_occ$species)
 ```
+
+# Subset TRY, BIEN, and GIFT by species list
+## TRY
+```{r}
+TRY_filtered <- TRY_traits %>%
+  filter(SpeciesName %in% plant_species)
+```
+
+## BIEN
+```{r}
+BIEN_filtered <- BIEN_traits %>%
+  filter(scrubbed_species_binomial %in% plant_species)
+```
+
+## GIFT
+```{r}
+GIFT_filtered <- GIFT_traits %>%
+  filter(species %in% plant_species)
+```

### 2023-11-22 code/L1/L1_P2_TropicalAndes_taxcleaning.Rmd
remove NAs and duplicates from all_species list, check lookup table for duplicates, remove duplicates and NA from lookup table, update adding accepted species column to data

@@
 ```
-# Remove NAs
-```{r}
-all_species <- all_species[!is.na(all_species)]
-length(all_species)
-```
 
@@
 all_species_cleaned <- name_check_results$names_clean
+length(all_species_cleaned)
+```
+
+# Remove NAs and duplicates
+```{r}
+all_species_cleaned <- all_species[!is.na(all_species_cleaned)]
+length(all_species_cleaned)
+```
+```{r}
+all_species_cleaned <- unique(all_species_cleaned)
 length(all_species_cleaned)
@@
 nrow(all_species_other)
+all_species_NA <- subset(all_species_hamonized_results, Taxonomic_status == NA)
+nrow(all_species_NA)
 ```
@@
 
+# Check lookup table for duplicates
+```{r}
+n_occur <- data.frame(table(lookup_table$Name_submitted))
+nrow(n_occur)
+duplicate_counts <- n_occur[n_occur$Freq >1,]
+nrow(duplicate_counts)
+duplicates <- lookup_table[lookup_table$Name_submitted %in% n_occur$Var1[n_occur$Freq > 1],]
+nrow(duplicates)
+duplicates_pairs <- duplicates[order(duplicates$Name_submitted),]
+nrow(duplicates_pairs)
+```
+
+# Remove duplicates and accepted names that are NA
+```{r}
+lookup_table <- lookup_table[!duplicated(lookup_table), ]
+nrow(lookup_table)
+lookup_table <- subset(lookup_table, !is.na(Accepted_species))
+nrow(lookup_table)
+```
+
 # Save lookup table as csv
 ```{r}
-write.csv(lookup_table, file.path(output_path,"plant_lookup_table.csv"))
+write.csv(lookup_table, file.path(output_path,"plant_lookup_table.csv"), row.names = FALSE)
 ```
@@
 nrow(GBIF_occ_subset)
-# add accepted species column from lookup table based on species column
-GBIF_occ_subset_harmonized <- merge(GBIF_occ_subset, lookup_table, by.x = "species", by.y = "Name_submitted")
+# filter lookup_table to species in data
+lookup_table_subset <- lookup_table %>%
+  filter(Name_submitted %in% GBIF_occ_subset$species)
+nrow(lookup_table_subset)
+
+## add accepted species column from lookup table based on species column
+GBIF_occ_subset_harmonized <- right_join(lookup_table_subset, GBIF_occ_subset, join_by(Name_submitted == species), keep = TRUE, relationship = "one-to-many")
 nrow(GBIF_occ_subset_harmonized)
@@
 ```{r}
+# filter data to species in lookup_table$Names_submitted
+TRY_traits_subset <- TRY_traits %>%
+  filter(SpeciesName %in% lookup_table$Name_submitted)
+nrow(TRY_traits_subset)
+# filter lookup_table to species in data
+lookup_table_subset <- lookup_table %>%
+  filter(Name_submitted %in% TRY_traits_subset$SpeciesName)
+nrow(lookup_table_subset)
 
+## add accepted species column from lookup table based on species column
+TRY_traits_subset_harmonized <- right_join(lookup_table_subset, TRY_traits_subset, join_by(Name_submitted == SpeciesName), keep = TRUE, relationship = "one-to-many")
+nrow(TRY_traits_subset_harmonized)
 ```

### 2023-12-06 code/L0/L0_TropicalAndes_plants_TRY.Rmd
Add keep only records with non-NA trait IDs

@@
 library(dplyr)
+library(tidyr)
 ```
@@
 
+# Keep only record with non-NA trait IDs
+```{r}
+try_data <- try_data %>%
+  drop_na(TraitID)
+nrow(try_data)
+```
+
 ```{r}
@@
 
+# Get units for each trait
+```{r}
+try_data %>% distinct(TraitName, UnitName)
+```
+
 # Write data to csv

### 2023-12-06 code/L1/L1_P2_TropicalAndes_taxcleaning.Rmd
update NA removal code so it works, filter trait data to species in lookup table, export filtered data

@@
 nrow(all_species_other)
-all_species_NA <- subset(all_species_hamonized_results, Taxonomic_status == NA)
+all_species_NA <- subset(all_species_hamonized_results, is.na(all_species_hamonized_results$Taxonomic_status))
 nrow(all_species_NA)
@@
 all_species_disparities <- all_species_hamonized_results %>%
-                            filter(Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", "")) %>%
+                            filter(Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", "", NA)) %>%
                             select(Name_submitted)
@@
 
-# Save lookup table as csv
+# Save lookup table and unresolved species list as csv
 ```{r}
 write.csv(lookup_table, file.path(output_path,"plant_lookup_table.csv"), row.names = FALSE)
+write.csv(unresolved_species, file.path(output_path,"plant_unresolved_species.csv"), row.names = FALSE)
 ```
@@
 ```{r}
+# filter data to species in lookup_table$Names_submitted
+BIEN_traits_subset <- BIEN_traits %>%
+  filter(scrubbed_species_binomial %in% lookup_table$Name_submitted)
+nrow(BIEN_traits_subset)
+# filter lookup_table to species in data
+lookup_table_subset <- lookup_table %>%
+  filter(Name_submitted %in% BIEN_traits_subset$scrubbed_species_binomial)
+nrow(lookup_table_subset)
 
+## add accepted species column from lookup table based on species column
+BIEN_traits_subset_harmonized <- right_join(lookup_table_subset, BIEN_traits_subset, join_by(Name_submitted == scrubbed_species_binomial), keep = TRUE, relationship = "one-to-many")
+nrow(BIEN_traits_subset_harmonized)
 ```
@@
 ```{r}
+GIFT_traits_subset <- GIFT_traits %>%
+  filter(species %in% lookup_table$Name_submitted)
+nrow(GIFT_traits_subset)
+# filter lookup_table to species in data
+lookup_table_subset <- lookup_table %>%
+  filter(Name_submitted %in% GIFT_traits_subset$species)
+nrow(lookup_table_subset)
 
+## add accepted species column from lookup table based on species column
+GIFT_traits_subset_harmonized <- right_join(lookup_table_subset, GIFT_traits_subset, join_by(Name_submitted == species), keep = TRUE, relationship = "one-to-many")
+nrow(GIFT_traits_subset_harmonized)
 ```
@@
 # Summary
+## GBIF
 ```{r}
-source("~/GitHub/neotropical_plants/functions/data_summary.R")
+print("Number of records: ")
+nrow(GBIF_occ_subset_harmonized)
+print("Number of species: ")
+length(unique(GBIF_occ_subset_harmonized$Accepted_species))
 ```
 
+## TRY
 ```{r}
-data_summary(GBIF_occ_subset_harmonized, GBIF_occ_subset_harmonized$species, GBIF_occ_subset_harmonized$genus, GBIF_occ_subset_harmonized$family)
+print("Number of records: ")
+nrow(TRY_traits_subset_harmonized)
+print("Number of species: ")
+length(unique(TRY_traits_subset_harmonized$Accepted_species))
+```
+
+## BIEN
+```{r}
+print("Number of records: ")
+nrow(BIEN_traits_subset_harmonized)
+print("Number of species: ")
+length(unique(BIEN_traits_subset_harmonized$Accepted_species))
+```
+
+## GIFT
+```{r}
+print("Number of records: ")
+nrow(GIFT_traits_subset_harmonized)
+print("Number of species: ")
+length(unique(GIFT_traits_subset_harmonized$Accepted_species))
 ```
@@
 ```{r}
-
+write.csv(GBIF_occ_subset_harmonized, file.path(output_path,"TropicalAndes_GBIF_plant_occ_harmonized.csv"), row.names = FALSE)
+write.csv(TRY_traits_subset_harmonized, file.path(output_path,"TropicalAndes_TRY_plant_traits_harmonized.csv"), row.names = FALSE)
+write.csv(BIEN_traits_subset_harmonized, file.path(output_path,"TropicalAndes_BIEN_plant_traits_harmonized.csv"), row.names = FALSE)
+write.csv(GIFT_traits_subset_harmonized, file.path(output_path,"TropicalAndes_GIFT_plant_traits_harmonized.csv"), row.names = FALSE)
 ```

### 2023-12-06 code/L1/L1_P3_TropicalAndes_plant_traits.Rmd
update input filepaths, add row count code, add code to combine trait data, start reclassification code

@@
 ```{r}
-plant_occ <- read.csv(file.path(data_path_L1, "TropicalAndes_GBIF_plant_occ_cleaned.csv"))
-TRY_traits <- read.csv(file.path(data_path_L0,"TropicalAndes_TRY_plant_traits.csv"))
-BIEN_traits <- read.csv(file.path(data_path_L0,"AllDesired_BIEN_plant_traits.csv"))
-GIFT_traits <- read.csv(file.path(data_path_L0,"TropicalAndes_GIFT_plant_traits.csv"))
+plant_occ <- read.csv(file.path(data_path_L1, "TropicalAndes_GBIF_plant_occ_harmonized.csv"))
+TRY_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_TRY_plant_traits_harmonized.csv"))
+BIEN_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_BIEN_plant_traits_harmonized.csv"))
+GIFT_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_GIFT_plant_traits_harmonized.csv"))
 ```
@@
 ```{r}
-plant_species <- unique(plant_occ$species)
+plant_species <- unique(plant_occ$Accepted_species)
+length(plant_species)
 ```
@@
 ```{r}
+nrow(TRY_traits)
 TRY_filtered <- TRY_traits %>%
-  filter(SpeciesName %in% plant_species)
+  filter(Accepted_species %in% plant_species)
+nrow(TRY_filtered)
 ```
@@
 ```{r}
+nrow(BIEN_traits)
 BIEN_filtered <- BIEN_traits %>%
-  filter(scrubbed_species_binomial %in% plant_species)
+  filter(Accepted_species %in% plant_species)
+nrow(BIEN_filtered)
 ```
@@
 ```{r}
+nrow(GIFT_traits)
 GIFT_filtered <- GIFT_traits %>%
-  filter(species %in% plant_species)
+  filter(Accepted_species %in% plant_species)
+nrow(GIFT_filtered)
 ```
+
+# Combine trait information from TRY, BIEN, GIFT
+
+```{r}
+TRYTraitTypes <- TRY_filtered %>%
+  group_by(TraitName) %>%
+  distinct(DataName)
+```
+
+```{r}
+colnames(TRY_filtered)
+```
+columns to keep : Accepted_species, TraitName, OrigValueStr
+
+```{r}
+TRY_filtered_subset <- TRY_filtered[ , c("Accepted_species", "TraitName", "OrigValueStr", "UnitName")]
+colnames(TRY_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue", "Unit")
+```
+
+```{r}
+colnames(BIEN_filtered)
+```
+columns to keep : Accepted_species, trait_name, trait_value, unit
+
+```{r}
+BIEN_filtered_subset <- BIEN_filtered[ , c("Accepted_species", "trait_name", "trait_value", "unit")]
+colnames(BIEN_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue", "Unit")
+```
+
+```{r}
+colnames(GIFT_filtered)
+```
+columns to keep : Accepted_species, trait_value, trait_name
+
+```{r}
+GIFT_filtered_subset <- GIFT_filtered[ , c("Accepted_species", "trait_name", "trait_value")]
+colnames(GIFT_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue")
+```
+
+Add Unit column to GIFT dataframe
+```{r}
+GIFT_filtered_subset <- GIFT_filtered_subset %>%
+  mutate(Unit = case_when(
+  TraitName == "plant_height" ~ "m",
+  TraitName == "plant_height_max" ~ "m",
+  TraitName == "plant_height_min" ~ "m",
+  TraitName == "seed_mass_mean" ~ "g",
+  TraitName == "seed_mass_min" ~ "g",
+  TraitName == "seed_mass_max" ~ "g",
+  TraitName == "fruit_length_min" ~ "cm",
+  TraitName == "fruit_length_max" ~ "cm",
+  TraitName == "fruit_length_mean" ~ "cm",
+  TraitName == "seed_length_max" ~ "mm",
+  TraitName == "seed_length_min" ~ "mm",
+  TraitName == "seed_length_mean" ~ "mm",
+  TraitName == "seed_width_max" ~ "mm",
+  TraitName == "seed_width_min" ~ "mm",
+  TraitName == "seed_width_mean" ~ "mm",
+  TraitName == "plant_lifespan" ~ "years",
+  .default = ""
+  ))
+```
+
+```{r}
+traits <- rbind(TRY_filtered_subset, BIEN_filtered_subset, GIFT_filtered_subset)
+```
+
+```{r}
+traits %>%
+  count(TraitName)
+```
+
+# Standardize trait measurements
+## Dispersal syndrome
+```{r}
+traits_dispersal <- traits %>%
+  filter(grepl('Dispersal|dispersal', TraitName))
+traits_dispersal$OrigTraitValue <- traits_dispersal$TraitValue
+```
+
+Reclassify to the following:
+anemochorous - wind dispersal
+anthropochorous - human dispersal
+autochorous - non-aided dispersal
+hydrochorous - water dispersal
+myrmecochorous - ant dispersal
+zoochorous - animal dispersal
+unspecialized - multiple
+```{r}
+traits_dispersal <- traits_dispersal %>%
+  filter() %>%
+  mutate(TraitValue= case_when(
+    grepl('Eaten|Animal|Bird|Mammal|Monkey', OrigTraitValue, ignore.case = TRUE) ~ "zoochorous",
+    grepl('Ants', OrigTraitValue, ignore.case = TRUE) ~ "myrmecochorous",
+    grepl('wind|Anemochory', OrigTraitValue, ignore.case = TRUE) ~ "anemochorous",
+    grepl('human', OrigTraitValue, ignore.case = TRUE) ~ "anthropochorous",
+    grepl('No', OrigTraitValue, ignore.case = TRUE) ~ "autochorous"
+  ))
+```
+
+```{r}
+print(unique(traits_dispersal$TraitValue))
+```

### 2023-12-13 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
add code to tranform sf to projected coordinate reference system, update plotting code to work with new crs, add more resolutions for mapping/plotting

@@
 worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
+Americas <- ne_countries(continent = c("North America", "South America"), returnclass = "sf")
 #polygon of Tropical Andes
 TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
-```
-
-```{r}
-TApoly <- st_transform(TApoly, crs = )
 ```
@@
 
+# Transform to projected coordinate reference system (units from degrees to meters)
+```{r}
+Americas <- st_transform(Americas, 5389)
+TApoly <- st_transform(TApoly, 5389)
+TropicalAndes_IUCNHabitat_Forest <- st_transform(TropicalAndes_IUCNHabitat_Forest, 5389)
+plants.sf <- st_transform(plants.sf, 5389)
+frugivores.sf <- st_transform(frugivores.sf, 5389)
+```
+
+# Group by species
 ```{r}
@@
 ggplot() +
-geom_sf(data = worldMap, fill = "white") +
+  geom_sf(data = Americas, fill = "white") +
   geom_sf(data = TApoly) +
@@
   labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
-coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
   annotation_scale(location = "bl", width_hint = 0.5) +
@@
 ggplot() +
-  geom_sf(data = worldMap, fill = "white") +
+  geom_sf(data = Americas, fill = "white") +
   geom_sf(data = TApoly) +
@@
   labs(title = "Tropical Andes Plant Occurrences", x = "Latitude", y = "Longitude") +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  )
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  theme(panel.background = element_rect(fill = "lightblue"))
+
 plantsPointsPlot
@@
   labs(title = "Tropical Andes Frugivore Occurrences", x = "Latitude", y = "Longitude") +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  )
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  theme(panel.background = element_rect(fill = "lightblue"))
+
 frugivoresPointsPlot
@@
 
-# Make 0.5 Grid
+# Plot TD/Richness
+
+## 5 km resolution
+### Make Grid
 ```{r}
-# make grid
-TAGrid0.5 <- TApoly %>%
-  st_make_grid(cellsize = c(0.5)) %>%
+#check units
+st_crs(TApoly, parameters = TRUE)$units_gdal
+st_crs(TApoly)
+st_crs(TropicalAndes_IUCNHabitat_Forest)
+```
+
+```{r}
+TAGrid_5 <- TApoly %>%
+  st_make_grid(cellsize = c(5000)) %>%
   st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
@@
 
-# Plot Empty Grid
+### Plot Empty Grid
 ```{r}
-# empty grid
 gridPlot <-
 ggplot() +
-  geom_sf(data = worldMap, fill = "white")+
+  geom_sf(data = Americas, fill = "white") +
   geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
-  geom_sf(data = TAGrid0.5) +
-  labs(title = "Tropical Andes Grid [0.5°]", x = "Latitude", y = "Longitude") +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  )
+  geom_sf(data = TAGrid_5) +
+  labs(title = "Tropical Andes Grid [5 km]", x = "Latitude", y = "Longitude") +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  theme(panel.background = element_rect(fill = "lightblue"))
+
 gridPlot
@@
 
-# Cell Richness
+### Cell Richness
 ```{r}
-plant_richness_grid0.5 <- TAGrid0.5 %>%
+plant_richness_grid_5 <- TAGrid_5 %>%
   st_join(plants_sf_species) %>%
@@
 ```{r}
-frugivore_richness_grid0.5 <- TAGrid0.5 %>%
+frugivore_richness_grid_5 <- TAGrid_5 %>%
   st_join(frugivores_sf_species) %>%
@@
 
-# Richness plots
+### Richness plots
 ```{r}
-plantgridRichTA0.5 <-
-  ggplot(plant_richness_grid0.5) +
-  geom_sf(data = worldMap, fill = "white")+
+plantgridRichTA_5km <-
+ggplot(plant_richness_grid_5) +
+  geom_sf(data = Americas, fill = "white") +
   geom_sf(data = TApoly, fill = "grey", size = 0.1) +
@@
   scale_fill_scico(palette = "romaO", direction = 1) +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  ) +
+  labs(title = "Tropical Andes Plant Richness [5 km]", x = "Longitude", y = "Latitude",
+       fill = "Number of species") +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
   annotation_scale(location = "bl", width_hint = 0.5) +
@@
     pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Plant Raw Richness", x = "Longitude", y = "Latitude", fill = "Number of species")
-plantgridRichTA0.5
+  theme(panel.background = element_rect(fill = "lightblue"))
+
+plantgridRichTA_5km
 ```
@@
 ```{r}
-frugivoregridRichTA0.5 <-
-  ggplot(frugivore_richness_grid0.5) +
-  geom_sf(data = worldMap, fill = "white")+
+frugivoregridRichTA_5km <-
+ggplot(frugivore_richness_grid_5) +
+  geom_sf(data = Americas, fill = "white") +
   geom_sf(data = TApoly, fill = "grey", size = 0.1) +
@@
   scale_fill_scico(palette = "romaO", direction = 1) +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  ) +   
+  labs(title = "Tropical Andes Frugivore Richness [5 km]", x = "Longitude", y = "Latitude",
+       fill = "Number of species") +
+    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
   annotation_scale(location = "bl", width_hint = 0.5) +
@@
     pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Frugivore Raw Richness", x = "Longitude", y = "Latitude", fill = "Number of species")
-frugivoregridRichTA0.5
+  theme(panel.background = element_rect(fill = "lightblue"))
+
+frugivoregridRichTA_5km
 ```
 
-# Extract cell values
+## Extract cell values
 ```{r}
-plant_cellRichness0.5 <- plantgridRichTA0.5[["data"]]
-frugivore_cellRichness0.5 <- frugivoregridRichTA0.5[["data"]]
+plant_cellRichness_5km <- plantgridRichTA_5km[["data"]]
+frugivore_cellRichness_5km <- frugivoregridRichTA_5km[["data"]]
+```
+
+### Plant vs frugivore richness plot
+```{r}
+compare_richness_summary_5km <- summary(lm(frugivore_cellRichness_5km$num_species ~ plant_cellRichness_5km$num_species))
+compare_richness_summary_5km
+
+# save r-squared value
+compare_richness_r2_5km <- compare_richness_summary_5km$r.squared
 ```
@@
 ```{r}
-plot(plant_cellRichness0.5$num_species, frugivore_cellRichness0.5$num_species,
-     main = "Plant Richness vs Frugivore Richness",
+plot(plant_cellRichness_5km$num_species, frugivore_cellRichness_5km$num_species,
+     main = "Plant Richness vs Frugivore Richness [5 km]",
      xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
-abline(lm(frugivore_cellRichness0.5$num_species ~ plant_cellRichness0.5$num_species))
-summary(lm(frugivore_cellRichness0.5$num_species ~ plant_cellRichness0.5$num_species))
+abline(lm(frugivore_cellRichness_5km$num_species ~ plant_cellRichness_5km$num_species))
 ```
@@
 ```{r}
-hist(plant_cellRichness0.5$num_species,
+hist(plant_cellRichness_5km$num_species,
      main = "Histogram of Plant Richness by cell",
      xlab = "Plant Richness by cell")
-hist(frugivore_cellRichness0.5$num_species,
+hist(frugivore_cellRichness_5km$num_species,
      main = "Histogram of Frugivore Richness by cell",
@@
 
+## 25 km 
+### Make Grid
 ```{r}
-summary(plant_cellRichness0.5$num_species)
-summary(frugivore_cellRichness0.5$num_species)
+TAGrid_25 <- TApoly %>%
+  st_make_grid(cellsize = c(25000)) %>%
+  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+  st_cast("MULTIPOLYGON") %>%
+  st_sf() %>%
+  mutate(cellid = row_number())
 ```
 
+### Plot Empty Grid
+```{r}
+gridPlot <-
+ggplot() +
+  geom_sf(data = Americas, fill = "white") +
+  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
+  geom_sf(data = TAGrid_25) +
+  labs(title = "Tropical Andes Grid [25 km]", x = "Latitude", y = "Longitude") +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  theme(panel.background = element_rect(fill = "lightblue"))
+
+gridPlot
+```
+
+### Cell Richness
+```{r}
+plant_richness_grid_25 <- TAGrid_25 %>%
+  st_join(plants_sf_species) %>%
+  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
+  group_by(cellid) %>%
+  summarize(num_species = sum(overlap))
+```
+
+```{r}
+frugivore_richness_grid_25 <- TAGrid_25 %>%
+  st_join(frugivores_sf_species) %>%
+  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
+  group_by(cellid) %>%
+  summarize(num_species = sum(overlap))
+```
+
+### Richness plots
+```{r}
+plantgridRichTA_25km <-
+ggplot(plant_richness_grid_25) +
+  geom_sf(data = Americas, fill = "white") +
+  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
+  geom_sf(aes(fill = num_species), color = NA) +
+  scale_fill_scico(palette = "romaO", direction = 1) +
+  labs(title = "Tropical Andes Plant Richness [25 km]", x = "Longitude", y = "Latitude",
+       fill = "Number of species") +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  theme(panel.background = element_rect(fill = "lightblue"))
+
+plantgridRichTA_25km
+```
+
+```{r}
+frugivoregridRichTA_25km <-
+ggplot(frugivore_richness_grid_25) +
+  geom_sf(data = Americas, fill = "white") +
+  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
+  geom_sf(aes(fill = num_species), color = NA) +
+  scale_fill_scico(palette = "romaO", direction = 1) +
+  labs(title = "Tropical Andes Frugivore Richness [25 km]", x = "Longitude", y = "Latitude",
+       fill = "Number of species") +
+    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  theme(panel.background = element_rect(fill = "lightblue"))
+
+frugivoregridRichTA_25km
+```
+
+## Extract cell values
+```{r}
+plant_cellRichness_25km <- plantgridRichTA_25km[["data"]]
+frugivore_cellRichness_25km <- frugivoregridRichTA_25km[["data"]]
+```
+
+### Plant vs frugivore richness plot
+```{r}
+compare_richness_summary_25km <- summary(lm(frugivore_cellRichness_25km$num_species ~ plant_cellRichness_25km$num_species))
+compare_richness_summary_25km
+
+# save r-squared value
+compare_richness_r2_25km <- compare_richness_summary_25km$r.squared
+```
+
+```{r}
+plot(plant_cellRichness_25km$num_species, frugivore_cellRichness_25km$num_species,
+     main = "Plant Richness vs Frugivore Richness [25 km]",
+     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
+abline(lm(frugivore_cellRichness_25km$num_species ~ plant_cellRichness_25km$num_species))
+```
+
+```{r}
+hist(plant_cellRichness_25km$num_species,
+     main = "Histogram of Plant Richness by cell",
+     xlab = "Plant Richness by cell")
+hist(frugivore_cellRichness_25km$num_species,
+     main = "Histogram of Frugivore Richness by cell",
+     xlab = "Frugivore Richness by cell")
+```
+
+## 50 km 
+### Make Grid
+```{r}
+TAGrid_50 <- TApoly %>%
+  st_make_grid(cellsize = c(50000)) %>%
+  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+  st_cast("MULTIPOLYGON") %>%
+  st_sf() %>%
+  mutate(cellid = row_number())
+```
+
+### Plot Empty Grid
+```{r}
+gridPlot <-
+ggplot() +
+  geom_sf(data = Americas, fill = "white") +
+  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
+  geom_sf(data = TAGrid_50) +
+  labs(title = "Tropical Andes Grid [50 km]", x = "Latitude", y = "Longitude") +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  theme(panel.background = element_rect(fill = "lightblue"))
+
+gridPlot
+```
+
+### Cell Richness
+```{r}
+plant_richness_grid_50 <- TAGrid_50 %>%
+  st_join(plants_sf_species) %>%
+  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
+  group_by(cellid) %>%
+  summarize(num_species = sum(overlap))
+```
+
+```{r}
+frugivore_richness_grid_50 <- TAGrid_50 %>%
+  st_join(frugivores_sf_species) %>%
+  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
+  group_by(cellid) %>%
+  summarize(num_species = sum(overlap))
+```
+
+### Richness plots
+```{r}
+plantgridRichTA_50km <-
+ggplot(plant_richness_grid_50) +
+  geom_sf(data = Americas, fill = "white") +
+  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
+  geom_sf(aes(fill = num_species), color = NA) +
+  scale_fill_scico(palette = "romaO", direction = 1) +
+  labs(title = "Tropical Andes Plant Richness [50 km]", x = "Longitude", y = "Latitude",
+       fill = "Number of species") +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  theme(panel.background = element_rect(fill = "lightblue"))
+
+plantgridRichTA_50km
+```
+
+```{r}
+frugivoregridRichTA_50km <-
+ggplot(frugivore_richness_grid_50) +
+  geom_sf(data = Americas, fill = "white") +
+  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
+  geom_sf(aes(fill = num_species), color = NA) +
+  scale_fill_scico(palette = "romaO", direction = 1) +
+  labs(title = "Tropical Andes Frugivore Richness [50 km]", x = "Longitude", y = "Latitude",
+       fill = "Number of species") +
+    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  theme(panel.background = element_rect(fill = "lightblue"))
+
+frugivoregridRichTA_50km
+```
+
+## Extract cell values
+```{r}
+plant_cellRichness_50km <- plantgridRichTA_50km[["data"]]
+frugivore_cellRichness_50km <- frugivoregridRichTA_50km[["data"]]
+```
+
+### Plant vs frugivore richness plot
+```{r}
+compare_richness_summary_50km <- summary(lm(frugivore_cellRichness_50km$num_species ~ plant_cellRichness_50km$num_species))
+compare_richness_summary_50km
+
+# save r-squared value
+compare_richness_r2_50km <- compare_richness_summary_50km$r.squared
+```
+
+```{r}
+plot(plant_cellRichness_50km$num_species, frugivore_cellRichness_50km$num_species,
+     main = "Plant Richness vs Frugivore Richness [50 km]",
+     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
+abline(lm(frugivore_cellRichness_50km$num_species ~ plant_cellRichness_50km$num_species))
+```
+
+```{r}
+hist(plant_cellRichness_50km$num_species,
+     main = "Histogram of Plant Richness by cell",
+     xlab = "Plant Richness by cell")
+hist(frugivore_cellRichness_50km$num_species,
+     main = "Histogram of Frugivore Richness by cell",
+     xlab = "Frugivore Richness by cell")
+```
+
+## 75 km 
+### Make Grid
+```{r}
+TAGrid_75 <- TApoly %>%
+  st_make_grid(cellsize = c(75000)) %>%
+  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+  st_cast("MULTIPOLYGON") %>%
+  st_sf() %>%
+  mutate(cellid = row_number())
+```
+
+### Plot Empty Grid
+```{r}
+gridPlot <-
+ggplot() +
+  geom_sf(data = Americas, fill = "white") +
+  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
+  geom_sf(data = TAGrid_75) +
+  labs(title = "Tropical Andes Grid [75 km]", x = "Latitude", y = "Longitude") +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  theme(panel.background = element_rect(fill = "lightblue"))
+
+gridPlot
+```
+
+### Cell Richness
+```{r}
+plant_richness_grid_75 <- TAGrid_75 %>%
+  st_join(plants_sf_species) %>%
+  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
+  group_by(cellid) %>%
+  summarize(num_species = sum(overlap))
+```
+
+```{r}
+frugivore_richness_grid_75 <- TAGrid_75 %>%
+  st_join(frugivores_sf_species) %>%
+  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
+  group_by(cellid) %>%
+  summarize(num_species = sum(overlap))
+```
+
+### Richness plots
+```{r}
+plantgridRichTA_75km <-
+ggplot(plant_richness_grid_75) +
+  geom_sf(data = Americas, fill = "white") +
+  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
+  geom_sf(aes(fill = num_species), color = NA) +
+  scale_fill_scico(palette = "romaO", direction = 1) +
+  labs(title = "Tropical Andes Plant Richness [75 km]", x = "Longitude", y = "Latitude",
+       fill = "Number of species") +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  theme(panel.background = element_rect(fill = "lightblue"))
+
+plantgridRichTA_75km
+```
+
+```{r}
+frugivoregridRichTA_75km <-
+ggplot(frugivore_richness_grid_75) +
+  geom_sf(data = Americas, fill = "white") +
+  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
+  geom_sf(aes(fill = num_species), color = NA) +
+  scale_fill_scico(palette = "romaO", direction = 1) +
+  labs(title = "Tropical Andes Frugivore Richness [75 km]", x = "Longitude", y = "Latitude",
+       fill = "Number of species") +
+    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  theme(panel.background = element_rect(fill = "lightblue"))
+
+frugivoregridRichTA_75km
+```
+
+## Extract cell values
+```{r}
+plant_cellRichness_75km <- plantgridRichTA_75km[["data"]]
+frugivore_cellRichness_75km <- frugivoregridRichTA_75km[["data"]]
+```
+
+### Plant vs frugivore richness plot
+```{r}
+compare_richness_summary_75km <- summary(lm(frugivore_cellRichness_75km$num_species ~ plant_cellRichness_75km$num_species))
+compare_richness_summary_75km
+
+# save r-squared value
+compare_richness_r2_75km <- compare_richness_summary_75km$r.squared
+```
+
+# Compare r-squared by resolution
+Plot
+# Multi panel plots
+Richness maps
 # Write data to csv

### 2023-12-13 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
add 100 km resolution and code to plot r-squared values by resolution

@@
 
+```{r}
+plot(plant_cellRichness_75km$num_species, frugivore_cellRichness_75km$num_species,
+     main = "Plant Richness vs Frugivore Richness [75 km]",
+     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
+abline(lm(frugivore_cellRichness_75km$num_species ~ plant_cellRichness_75km$num_species))
+```
+
+```{r}
+hist(plant_cellRichness_75km$num_species,
+     main = "Histogram of Plant Richness by cell",
+     xlab = "Plant Richness by cell")
+hist(frugivore_cellRichness_75km$num_species,
+     main = "Histogram of Frugivore Richness by cell",
+     xlab = "Frugivore Richness by cell")
+```
+
+## 100 km 
+### Make Grid
+```{r}
+TAGrid_100 <- TApoly %>%
+  st_make_grid(cellsize = c(100000)) %>%
+  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+  st_cast("MULTIPOLYGON") %>%
+  st_sf() %>%
+  mutate(cellid = row_number())
+```
+
+### Plot Empty Grid
+```{r}
+gridPlot <-
+ggplot() +
+  geom_sf(data = Americas, fill = "white") +
+  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
+  geom_sf(data = TAGrid_100) +
+  labs(title = "Tropical Andes Grid [100 km]", x = "Latitude", y = "Longitude") +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  theme(panel.background = element_rect(fill = "lightblue"))
+
+gridPlot
+```
+
+### Cell Richness
+```{r}
+plant_richness_grid_100 <- TAGrid_100 %>%
+  st_join(plants_sf_species) %>%
+  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
+  group_by(cellid) %>%
+  summarize(num_species = sum(overlap))
+```
+
+```{r}
+frugivore_richness_grid_100 <- TAGrid_100 %>%
+  st_join(frugivores_sf_species) %>%
+  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
+  group_by(cellid) %>%
+  summarize(num_species = sum(overlap))
+```
+
+### Richness plots
+```{r}
+plantgridRichTA_100km <-
+ggplot(plant_richness_grid_100) +
+  geom_sf(data = Americas, fill = "white") +
+  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
+  geom_sf(aes(fill = num_species), color = NA) +
+  scale_fill_scico(palette = "romaO", direction = 1) +
+  labs(title = "Tropical Andes Plant Richness [100 km]", x = "Longitude", y = "Latitude",
+       fill = "Number of species") +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  theme(panel.background = element_rect(fill = "lightblue"))
+
+plantgridRichTA_100km
+```
+
+```{r}
+frugivoregridRichTA_100km <-
+ggplot(frugivore_richness_grid_100) +
+  geom_sf(data = Americas, fill = "white") +
+  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
+  geom_sf(aes(fill = num_species), color = NA) +
+  scale_fill_scico(palette = "romaO", direction = 1) +
+  labs(title = "Tropical Andes Frugivore Richness [100 km]", x = "Longitude", y = "Latitude",
+       fill = "Number of species") +
+    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  theme(panel.background = element_rect(fill = "lightblue"))
+
+frugivoregridRichTA_100km
+```
+
+## Extract cell values
+```{r}
+plant_cellRichness_100km <- plantgridRichTA_100km[["data"]]
+frugivore_cellRichness_100km <- frugivoregridRichTA_100km[["data"]]
+```
+
+### Plant vs frugivore richness plot
+```{r}
+compare_richness_summary_100km <- summary(lm(frugivore_cellRichness_100km$num_species ~ plant_cellRichness_100km$num_species))
+compare_richness_summary_100km
+
+# save r-squared value
+compare_richness_r2_100km <- compare_richness_summary_100km$r.squared
+```
+
+```{r}
+plot(plant_cellRichness_100km$num_species, frugivore_cellRichness_100km$num_species,
+     main = "Plant Richness vs Frugivore Richness [100 km]",
+     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
+abline(lm(frugivore_cellRichness_100km$num_species ~ plant_cellRichness_100km$num_species))
+```
+
+```{r}
+hist(plant_cellRichness_100km$num_species,
+     main = "Histogram of Plant Richness by cell",
+     xlab = "Plant Richness by cell")
+hist(frugivore_cellRichness_100km$num_species,
+     main = "Histogram of Frugivore Richness by cell",
+     xlab = "Frugivore Richness by cell")
+```
+
 # Compare r-squared by resolution
 Plot
+```{r}
+# create dataframe of resolutions and r-squared values
+resolution <- c("5", "25", "50", "75", "100")
+r_squared <- c(compare_richness_r2_5km, compare_richness_r2_25km, compare_richness_r2_50km, compare_richness_r2_75km, compare_richness_r2_100km)
+r_squared_compare <- data.frame(resolution, r-squared)
+```
+
+```{r}
+ggplot(r_squared_compare, aes(x=resolution, y=r_squared)) +
+  geom_point()+
+  labs(title = "R-squared values of plant-frugivore richness relationships by resolution", x = "Resolution (km)", y = "R-squared",
+       fill = "Number of species") 
+```
+
 # Multi panel plots

### 2024-01-18 code/L1/taxharmonization_troubleshooting.R
update to add troubleshooting file

@@ -1,0 +1,32 @@
+# Trouble shooting adding accepted names from lookup table to occurrence data
+
+library(dplyr)
+
+## Set file paths
+data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+
+## Load files
+GBIF_occ <- read.csv(file.path(data_path_L1, "TropicalAndes_GBIF_plant_occ_cleaned.csv"))
+lookup_table <- read.csv(file.path(output_path,"plant_lookup_table.csv"))
+
+#small subset for testing
+lookup_table_5 <- lookup_table[1:5, 2:3]
+
+GBIF_occ_subset_5 <- GBIF_occ %>%
+  filter(species %in% lookup_table_5$Name_submitted)
+nrow(GBIF_occ_subset_5)
+
+# x is lookuptable  y is occurrences
+GBIF_occ_subset_harmonized_5 <- right_join(lookup_table_5, GBIF_occ_subset_5, join_by(Name_submitted == species), keep = TRUE)
+nrow(GBIF_occ_subset_harmonized_5)
+
+## filter data to species in lookup_table$Names_submitted
+GBIF_occ_subset <- GBIF_occ %>%
+  filter(species %in% lookup_table$Name_submitted)
+nrow(GBIF_occ_subset)
+
+## add accepted species column from lookup table based on species column
+GBIF_occ_subset_harmonized <- right_join(lookup_table, GBIF_occ_subset, join_by(Name_submitted == species), keep = TRUE)
+nrow(GBIF_occ_subset_harmonized)

### 2024-01-18 ChoosingCRS.Rmd
Create ChoosingCRS.Rmd

# Choosing CRS
Compare richness maps using degree approximation and various reprojection options

# 
https://www.pmassicotte.com/posts/2021-11-12-choosing-the-right-projection/
```{r}
library(rnaturalearth)
library(sf)
library(patchwork)
#install.packages("crsuggest")
library(crsuggest)
```

```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
# load in shapefile
TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
```

```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
#polygon of Tropical Andes
TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
```

```{r}
st_crs(TApoly)$proj4string
```
```{r}
st_crs(TropicalAndes_IUCNHabitat_Forest)$proj4string
```
```{r}
p1 <- ggplot() +
  geom_sf(data = TApoly)

p1
```
```{r}
p2 <- ggplot() +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest)

p2
```

```{r}
suggested_crs_TA <- suggest_crs(TApoly, limit = 5)
suggested_crs_TA
```
```{r}
plot_map <- function(crs, map) {
  ggplot() +
    geom_sf(data = map) +
    coord_sf(crs = as.numeric(crs)) +
    labs(
      title = paste("CRS:", crs)
    )
}

maps <- lapply(suggested_crs_TA$crs_code, plot_map, map = TApoly)
wrap_plots(maps, ncol = 3, nrow = 2)
```
```{r}
p2 <- ggplot() +
  geom_sf(data = map) +
  coord_sf(crs = 5922)

p1 + p2
```

# One Option 
+proj=aea +lat_0=-32 +lon_0=-60 +lat_1=-5 +lat_2=-42 +x_0=0 +y_0=0 +ellps=aust_SA +units=m +no_defs +type=crs

### 2024-03-07 code/L1/L1_P3_TropicalAndes_plant_traits.Rmd
remove rows without accepted species names, create functions to average trait measurements (get one trait measure per species)

@@
 traits <- rbind(TRY_filtered_subset, BIEN_filtered_subset, GIFT_filtered_subset)
+nrow(traits)
+```
+
+```{r}
+# remove rows without accepted species names
+traits_clean <- traits[!(is.na(traits$Accepted_species) | traits$Accepted_species==""), ]
+nrow(traits_clean)
 ```
@@
 # Standardize trait measurements
+## Create functions
+```{r}
+# Numeric data
+numeric_data_average <- function(data, unit, traitname) {
+# remove rows if TraitValue is non-numeric
+data <- data %>%
+  filter(!is.na(as.numeric(TraitValue))) %>%
+  mutate(TraitValue = as.numeric(TraitValue))
+# remove duplicate rows
+data <- data %>%
+  distinct(.keep_all = TRUE)
+# calculate average value per trait per species
+average_data <- aggregate(TraitValue ~ Accepted_species, data, FUN = mean, na.rm = TRUE)
+# Remove rows with missing or empty species names
+average_data <- average_data[!is.na(average_data$Accepted_species) & average_data$Accepted_species !="", ]
+# Add unit & trait column
+average_data$Unit <- unit
+average_data$TraitName <- traitname
+
+return(average_data)
+}
+```
+
+```{r}
+# categorical data
+factor_data_merge <- function(data, traitname) {
+# remove duplicate rows
+data <- data %>%
+  distinct(.keep_all = TRUE)
+# remove rows with missing or empty species names
+data <- data[!is.na(data$Accepted_species) & data$Accepted_species != "", ]
+# Subset data to just species and TraitValue
+data <- data[,c("Accepted_species", "TraitValue")]
+# Find the most common trait value for each species
+  most_common_values <- aggregate(. ~ Accepted_species, data, function(x) {
+    # Find the mode (most common value) of the trait values
+    mode_value <- names(sort(table(x), decreasing = TRUE))[1]
+    return(mode_value)
+  })
+# Merge most common trait values back to the original data
+merged_data <- merge(data, most_common_values, by = "Accepted_species", all.x = TRUE)
+
+# Rename the merged column
+colnames(merged_data)[ncol(merged_data)] <- "TraitValue"
+
+# Remove column TraitValue.x
+merged_data <- merged_data[, !colnames(merged_data) %in% "TraitValue.x", drop = FALSE]
+
+# remove duplicate rows
+merged_data <- merged_data %>%
+  distinct(.keep_all = TRUE)
+
+# Add unit & trait column
+merged_data$TraitName <- traitname
+
+  return(merged_data)
+}
+```
+
 ## Dispersal syndrome
@@
 traits_dispersal <- traits %>%
-  filter(grepl('Dispersal|dispersal', TraitName))
+  filter(grepl('Dispersal syndrome|whole plant dispersal syndrome|dispersal_syndrome_1|dispersal_syndrome_2', TraitName))
 traits_dispersal$OrigTraitValue <- traits_dispersal$TraitValue
+```
+
+```{r}
+# remove rows containing weight or mass
+traits_dispersal <- traits_dispersal[!grepl("weight|mass|air-dry|oven|mean|SE",traits_dispersal$TraitValue, ignore.case = TRUE),]
+nrow(traits_dispersal)
+length(unique(traits_dispersal$Accepted_species))
 ```
@@
   mutate(TraitValue= case_when(
-    grepl('Eaten|Animal|Bird|Mammal|Monkey', OrigTraitValue, ignore.case = TRUE) ~ "zoochorous",
+    grepl('Eaten|Animal|Bird|Mammal|animals|endozoochor|dysochor|epizoochor|zoochor|Vertebrates|Reptiles+birds|Endo-zoochory|mammals|exozoochory|birds|Zoochory|vertebrate|Dys-zoochory|Ornithochory|Epi-zoochory|endozoochorous|endozoochory|ectozoochorous|zoochorous', OrigTraitValue, ignore.case = TRUE) ~ "zoochorous",
     grepl('Ants', OrigTraitValue, ignore.case = TRUE) ~ "myrmecochorous",
-    grepl('wind|Anemochory', OrigTraitValue, ignore.case = TRUE) ~ "anemochorous",
-    grepl('human', OrigTraitValue, ignore.case = TRUE) ~ "anthropochorous",
-    grepl('No', OrigTraitValue, ignore.case = TRUE) ~ "autochorous"
-  ))
+    grepl('wind|Anemochory|anemochor|meteorochor|chamaechor|Anemochory|anemochorous', OrigTraitValue, ignore.case = TRUE) ~ "anemochorous",
+    grepl('human|agochor|man|hemerochor|anthropochorous', OrigTraitValue, ignore.case = TRUE) ~ "anthropochorous",
+    grepl('No|Unassisted|autochor|blastochor|ombrochor|Autochory|Barochory|ballochor|herpochor|autochorous', OrigTraitValue, ignore.case = TRUE) ~ "autochorous",
+    grepl('water|nautochor|floating|hydrochor|hydrochorous', OrigTraitValue, ignore.case = TRUE) ~ "hydrochorous",
+    grepl('boleochor|combination|non specialized|unspecialized', OrigTraitValue, ignore.case = TRUE) ~ "unspecialized"
+  )) %>%
+  filter(!is.na(TraitValue))
+nrow(traits_dispersal)
+length(unique(traits_dispersal$Accepted_species))
 ```
@@
 ```
+
+```{r}
+traits_dispersal <- traits_dispersal %>%
+  mutate(TraitName = "dispersal_syndrome")
+```
+
+```{r}
+averagetraits_dispersal <- factor_data_merge(traits_dispersal, "Dispersal Syndrome")
+```
+
+## Fruit dry mass
+```{r}
+traits_fruitMass <- traits %>%
+  filter(grepl('Fruit dry mass', TraitName))
+traits_fruitMass$OrigTraitValue <- traits_fruitMass$TraitValue
+```
+
+```{r}
+print(unique(traits_fruitMass$Unit))
+```
+
+```{r}
+averagetraits_fruitMass <- numeric_data_average(traits_fruitMass, "mg", "Fruit Mass")
+```
+
+## Fruit length
+```{r}
+traits_fruitLength <- traits %>%
+  filter(grepl('Fruit length|maximum fruit length|minimum fruit length|fruit_length_min|fruit_length_max|fruit_length_mean', TraitName))
+traits_fruitLength$OrigTraitValue <- traits_fruitLength$TraitValue
+```
+
+```{r}
+print(unique(traits_fruitLength$Unit))
+```
+
+```{r}
+# remove rows if TraitValue is non-numeric
+traits_fruitLength <- traits_fruitLength %>%
+  filter(!is.na(as.numeric(TraitValue))) %>%
+  mutate(TraitValue = as.numeric(TraitValue))
+```
+
+```{r}
+# if units = cm, multiply trait value by 10 and change units to mm
+traits_fruitLength <- traits_fruitLength %>%
+  mutate(TraitValue = if_else(Unit == "cm", TraitValue * 10, TraitValue),
+         Unit = if_else(Unit == "cm", "mm", Unit))
+```
+
+```{r}
+averagetraits_fruitLength <- numeric_data_average(traits_fruitLength, "mm", "Fruit Length")
+```
+
+## Plant height
+```{r}
+traits_plantHeight <- traits %>%
+  filter(grepl('Plant height generative|Plant height vegetative|maximum whole plant height|minimum whole plant height|plant_height|plant_height_max|plant_height_min|whole plant height', TraitName))
+traits_plantHeight$OrigTraitValue <- traits_plantHeight$TraitValue
+```
+
+```{r}
+print(unique(traits_plantHeight$Unit))
+```
+
+```{r}
+averagetraits_plantHeight <- numeric_data_average(traits_plantHeight, "m", "Plant Height")
+```
+
+## Plant lifespan/longevity
+```{r}
+traits_plantLifespan <- traits %>%
+  filter(grepl('Plant lifespan (longevity)|longest whole plant longevity|maximum whole plant longevity|plant_lifespan', TraitName))
+traits_plantLifespan$OrigTraitValue <- traits_plantLifespan$TraitValue
+```
+
+```{r}
+print(unique(traits_plantLifespan$Unit))
+```
+
+```{r}
+averagetraits_plantLifespan <- numeric_data_average(traits_plantLifespan, "years", "Plant Lifespan")
+```
+
+## Fruit type
+```{r}
+traits_fruitType <- traits %>%
+  filter(grepl('Fruit dehiscence type|Fruit surface type|Fruit texture inside (fleshy, leathrey, …)|Fruit type|fruit type|fruit_dryness|fruit_type', TraitName))
+traits_fruitType$OrigTraitValue <- traits_fruitType$TraitValue
+```
+
+```{r}
+print(unique(traits_fruitType$Unit))
+```
+
+```{r}
+averagetraits_fruitType <- factor_data_merge(traits_fruitType, "Fruit Type")
+```
+
+## Seed mass
+```{r}
+traits_seedMass <- traits %>%
+  filter(grepl('Seed dry mass|seed mass|seed_mass_max|seed_mass_mean|seed_mass_min', TraitName))
+traits_seedMass$OrigTraitValue <- traits_seedMass$TraitValue
+```
+
+```{r}
+print(unique(traits_seedMass$Unit))
+```
+
+```{r}
+# remove rows if TraitValue is non-numeric
+traits_seedMass <- traits_seedMass %>%
+  filter(!is.na(as.numeric(TraitValue))) %>%
+  mutate(TraitValue = as.numeric(TraitValue))
+```
+
+```{r}
+# convert rows with unit = g to mg; multiply by 1000
+traits_seedMass <- traits_seedMass %>%
+  mutate(TraitValue = if_else(Unit == "g", TraitValue * 1000, TraitValue),
+         Unit = if_else(Unit == "g", "mg", Unit))
+```
+
+```{r}
+averagetraits_seedMass <- numeric_data_average(traits_seedMass, "mg", "Seed Mass")
+```
+
+## Seed length
+```{r}
+traits_seedLength <- traits %>%
+  filter(grepl('Seed length|seed_length_max|seed_length_mean|seed_length_min', TraitName))
+traits_seedLength$OrigTraitValue <- traits_seedLength$TraitValue
+```
+
+```{r}
+print(unique(traits_seedLength$Unit))
+```
+
+```{r}
+averagetraits_seedLength <- numeric_data_average(traits_seedLength, "mm", "Seed Length")
+```
+
+## Seed width
+```{r}
+traits_seedWidth <- traits %>%
+  filter(grepl('Seed width|seed_width_max|seed_width_mean|seed_width_min'
+, TraitName))
+traits_seedWidth$OrigTraitValue <- traits_seedWidth$TraitValue
+```
+
+```{r}
+print(unique(traits_seedWidth$Unit))
+```
+
+```{r}
+averagetraits_seedWidth <- numeric_data_average(traits_seedWidth, "mm", "Seed Width")
+```
+
+## Growth form
+```{r}
+traits_growthForm <- traits %>%
+  filter(grepl('Plant growth form|growth_form|whole plant growth form|whole plant growth form diversity'
+, TraitName))
+traits_growthForm$OrigTraitValue <- traits_growthForm$TraitValue
+```
+
+```{r}
+averagetraits_growthForm <- factor_data_merge(traits_growthForm, "Growth Form")
+```

### 2024-03-07 code/L1/L1_P1_TropicalAndes_plants_GBIF.Rmd
Add data summary function

@@
 output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+source("~/GitHub/neotropical_plants/functions/data_summary.R")
 ```
@@
 ```{r}
-print("The number of records is")
-nrow(TropicalAndes_GBIF_plant_occ_cleaned)
-print("The number of species is")
-length(unique(TropicalAndes_GBIF_plant_occ_cleaned$species))
-print("The number of genera is" )
-length(unique(TropicalAndes_GBIF_plant_occ_cleaned$genus))
-print("The number of families is")
-length(unique(TropicalAndes_GBIF_plant_occ_cleaned$family))
+data_summary(TropicalAndes_GBIF_plant_occ_cleaned, TropicalAndes_GBIF_plant_occ_cleaned$species, TropicalAndes_GBIF_plant_occ_cleaned$genus, TropicalAndes_GBIF_plant_occ_cleaned$family)
 ```

### 2024-04-18 code/L1/L1_P4_TropicalAndes_plant_imputetraits.Rmd
add code for imputation

@@
 ```{r}
+library(tidyr)
+library(mice)
+library(dplyr)
+```
 
+# Read in data
+```{r}
+plant_traits <- read.csv(file.path(data_path,"TropicalAndes_all_plant_traits_harmonized.csv"))
 ```
+
+# Set the seed for reproducibility
+```{r}
+set.seed(123)
+```
+
+```{r}
+#replace " " with _ in column names
+names(wide_traits) <- sub(" ", "_", names(wide_traits))
+```
+
+# show the missing data pattern
+```{r}
+md.pattern(wide_traits)
+```
+
+# Perform the imputation
+```{r}
+imp_model <- mice(wide_traits, method = "cart", maxit = 20)
+```
+
+```{r}
+imputed_data <- complete(imp_model)
+```

### 2024-04-18 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
add code for multi panel plot

@@
 ```
+
+## 10 km
+ADD CODE HERE!!!
 
@@
 Richness maps
+```{r}
+grid.arrange(plot_1, plot_2, nrow = 1)
+```
+
 # Write data to csv

### 2024-04-18 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
update input file, trait type code

@@
 ```{r}
-
 plant_traits$plant_height <- as.numeric(plant_traits$plant_height)
@@
 
-## Remove the species from PAM that have no occurrences anymore after subsetting to TA
+## Remove the species from PAM that have no occurrences
 ```{r}
@@
 ```{r}
-
 plant_traits_df_subset <- plant_traits %>% filter(species %in% colnames_plant)
@@
 ```{r}
-plant_trait_name <- c("dispersal_syndrome", "plant_height", "plant_lifespan")
-plant_trait_type <- c("N", "Q", "Q") 
-plant_trait_cat <- as.data.frame(cbind(plant_trait_name, plant_trait_type))
+trait_name <- c("dispersal_syndrome", "plant_height", "plant_lifespan")
+trait_type <- c("N", "Q", "Q") 
+plant_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
 ```
@@
 ```{r}
-frug_trait_name <- c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")
-frug_trait_type <- c("N", "Q", "Q", "Q")
-frug_trait_cat <- as.data.frame(cbind(frug_trait_name, frug_trait_type))
+trait_name <- c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")
+trait_type <- c("N", "Q", "Q", "Q")
+frug_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
 ```
@@
 ```{r}
-# View the components of the PCA axes *Remember the firt few components explain the most variation in dissimilarity. Clusters into groups
+# View the components of the PCA axes *Remember the first few components explain the most variation in dissimilarity. Clusters into groups
 
@@
 ## check number of species names
-nrow(sp_faxes_coord_frugivore_sub) #718
-ncol(subset_matrix_frugivore) #712
-
-sp_faxes_coord_frugivore_sub_names <- row.names(sp_faxes_coord_frugivore_sub)
-subset_matrix_frugivore_names <- colnames(subset_matrix_frugivore)
+nrow(sp_faxes_coord_frugivore_sub) #764
+ncol(subset_matrix_frugivore) #782
 
@@
 
-sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore_sub[ which((row.names(sp_faxes_coord_frugivore_sub) %in% subset_matrix_frugivore_names)==TRUE), ]
+sp_faxes_coord_frugivore_sub_names <- row.names(sp_faxes_coord_frugivore_sub)
+subset_matrix_frugivore_names <- colnames(subset_matrix_frugivore)
 
-nrow(sp_faxes_coord_frugivore_sub)
+frugivore_names <- intersect(sp_faxes_coord_frugivore_sub_names, subset_matrix_frugivore_names)
+frugivore_names <- na.omit(frugivore_names)
+
+sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore_sub[ which((row.names(sp_faxes_coord_frugivore_sub) %in% frugivore_names)==TRUE), ]
+
+subset_matrix_frugivore  <- as.data.frame(subset_matrix_frugivore)
+subset_matrix_frugivore <- subset_matrix_frugivore[ ,which((colnames(subset_matrix_frugivore) %in% frugivore_names)==TRUE)]
+
+#remove NAs
+sp_faxes_coord_frugivore_sub <- na.omit(sp_faxes_coord_frugivore_sub)
+subset_matrix_frugivore <- na.omit(subset_matrix_frugivore)
+
+nrow(sp_faxes_coord_frugivore_sub) #757
+ncol(subset_matrix_frugivore) #757
 sp_faxes_coord_frugivore_sub <- as.matrix(sp_faxes_coord_frugivore_sub)
+subset_matrix_frugivore <- as.matrix(subset_matrix_frugivore)
 ```
@@
   sp_faxes_coord   = sp_faxes_coord_frugivore_sub,
-  asb_sp_w         = PAM_plant_site_final,
+  asb_sp_w         = subset_matrix_frugivore,
   ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
@@
 fdis_frugivore <- alpha_fd_indices_frugivore$functional_diversity_indices$fdis
-## subset frugivore to compare for plotting
-fdis_frugivore_sub <- fdis_frugivore[1:25]
 
 plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
-frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)
-## subset frugivore to compare for plotting
 frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)
@@
 # Functional diversity plots
+# Polygons of countries
+```{r}
+worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
+Americas <- ne_countries(continent = c("North America", "South America"), returnclass = "sf")
+#polygon of Tropical Andes
+TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
+```
+
 ```{r}

### 2024-04-25 code/L1/L1_P4_TropicalAndes_plant_imputetraits.Rmd
add code to use BIEN and GIFT to fill in gaps by genus or family

--- 2023-11-01 code/L1/L1_P4_TropicalAndes_plant_imputetraits.Rmd
@@
 ---
-title: "TITLE"
+title: "Plant Trait Imputation"
 author: "Hazel J. Anderson"
@@
 ```{r}
+library(tidyr)
+library(BIEN)
+library(GIFT)
+library(mice)
+library(dplyr)
+```
 
+# Read in data
+```{r}
+plant_traits <- read.csv(file.path(data_path,"TropicalAndes_all_plant_traits_harmonized.csv"))
 ```
+
+# Use BIEN and GIFT to fill in gaps by genus or family.
+Need a list of families and genus for each trait that needs requesting
+2. add genus and family information to dataframe
+3. list by trait of species and genus
+## Create a dataframe of species and TraitNames with TraitValue = NA
+```{r}
+# remove X column
+plant_traits <- plant_traits[, !colnames(plant_traits) %in% "X", drop = FALSE]
+# Convert the wide dataframe to long format
+long_plant_traits <- plant_traits %>%
+  gather(TraitName, TraitValue, -species)
+
+# Filter rows with NA values
+na_traits <- long_plant_traits %>%
+  filter(is.na(TraitValue))
+
+# get a list of species
+na_species_list <- unique(na_traits$species)
+```
+
+## Add genus and family information to dataframe
+```{r}
+taxonomic_info <- BIEN_taxonomy_species(na_species_list)
+```
+
+```{r}
+# keep only columns with family, genus, species
+subset_df <- taxonomic_info[, c("scrubbed_family", "scrubbed_genus", "scrubbed_species_binomial")]
+# remove duplicates
+na_species_taxonomy <- distinct(subset_df)
+# rename column to family, genus, species
+na_species_taxonomy <- na_species_taxonomy %>%
+  rename(family = scrubbed_family,
+         genus = scrubbed_genus,
+         species = scrubbed_species_binomial)
+```
+
+```{r}
+# add to species dataframe
+na_species_df <- as.data.frame(na_species_list)
+names(na_species_df) <- "species"
+na_species_df <- merge(na_species_df, na_species_taxonomy, by = "species", all.x = TRUE)
+
+# retrieve list of species without family & genus information
+species_no_family <- na_species_df %>%
+  filter(is.na(family) | family == "") %>%
+  select(species) %>%
+  pull()
+species_no_genus <- na_species_df %>%
+  filter(is.na(genus) | genus == "") %>%
+  select(species) %>%
+  pull()
+# check if all the species in no_family & no_genus lists are the same
+identical(species_no_family, species_no_genus)
+```
+## Use taxize for species with no family and genus info
+
+```{r}
+library(taxize)
+
+# Function to retrieve taxonomic information for a chunk of species names
+get_taxonomic_info_chunk <- function(chunk_species_names) {
+  # Initialize an empty list to store taxonomic information for each chunk
+  chunk_taxonomic_info <- list()
+
+  # Loop through each species name in the chunk and retrieve taxonomic information
+  for (species_name in chunk_species_names) {
+    # Add a delay of 1 second between consecutive API requests
+    Sys.sleep(1)
+
+    # Make API request to retrieve taxonomic information
+    taxon_info <- tax_name(species_name, get = c("genus", "family"), db = "both")
+
+    # Store taxonomic information for the species in the list
+    chunk_taxonomic_info[[species_name]] <- data.frame(Species = species_name, 
+                                                       Genus = taxon_info$genus, 
+                                                       Family = taxon_info$family)
+  }
+
+  # Return the list of taxonomic information for the chunk
+  return(chunk_taxonomic_info)
+}
+
+# Define the chunk size
+chunk_size <- 100  # Adjust the chunk size as needed
+
+# Split the dataframe into chunks based on the chunk size
+chunks <- split(species_no_family, ceiling(seq_along(species_no_family) / chunk_size))
+
+# Initialize an empty list to store taxonomic information for all chunks
+all_taxonomic_info <- list()
+
+# Iterate over each chunk and retrieve taxonomic information
+for (chunk_index in seq_along(chunks)) {
+  # Get the species names in the current chunk
+  chunk_species_names <- chunks[[chunk_index]]
+
+  # Retrieve taxonomic information for the chunk
+  chunk_taxonomic_info <- get_taxonomic_info_chunk(chunk_species_names)
+
+  # Append taxonomic information for the chunk to the list
+  all_taxonomic_info <- c(all_taxonomic_info, list(chunk_taxonomic_info))
+}
+
+# Combine taxonomic information from all chunks into a single list
+all_taxonomic_info <- do.call(c, all_taxonomic_info)
+
+# Combine the list of dataframes into a single dataframe
+taxonomic_df <- do.call(rbind, all_taxonomic_info)
+```
+
+# Set the seed for reproducibility
+```{r}
+set.seed(123)
+```
+
+```{r}
+#replace " " with _ in column names
+names(plant_traits) <- sub(" ", "_", names(plant_traits))
+```
+
+# show the missing data pattern
+```{r}
+md.pattern(plant_traits)
+```
+# Trying to prevent error
+Error in edit.setup(data, setup, ...) : 
+  `mice` detected constant and/or collinear variables. No predictors were left after their removal.
+```{r}
+# Subset to numeric plant traits
+num_plant_traits <- plant_traits %>%
+  select_if(is.numeric)
+# Check for constant variables
+constant_vars <- sapply(num_plant_traits, function(x) length(unique(x))) == 1
+constant_vars <- names(constant_vars[constant_vars])
+
+# Remove constant variables
+num_plant_traits <- num_plant_traits[, !names(num_plant_traits) %in% constant_vars]
+
+# Check for collinearity
+cor_matrix <- cor(num_plant_traits)
+
+# Find highly correlated variables
+highly_correlated_pairs <- which(abs(cor_matrix) > 0.8 & upper.tri(cor_matrix, diag = TRUE), arr.ind = TRUE)
+
+# Print highly correlated pairs
+print(highly_correlated_pairs)
+```
+
+# Perform the imputation
+```{r}
+imp_model <- mice(plant_traits, method = "cart", maxit = 20)
+```
+
+```{r}
+imputed_data <- complete(imp_model)
+```

### 2024-04-25 code/L1/L1_P4_TropicalAndes_plant_imputetraits.Rmd
add catch and delay to taxize code retriving family and genus information for species from ncbi, fill in NAs of original dataframe and get list of species still missing family info

@@
   for (species_name in chunk_species_names) {
-    # Add a delay of 1 second between consecutive API requests
+    tryCatch({
+      # Add a delay between consecutive API requests
       Sys.sleep(1)
@@
       # Make API request to retrieve taxonomic information
-    taxon_info <- tax_name(species_name, get = c("genus", "family"), db = "both")
+      taxon_info <- tax_name(species_name, get = c("genus", "family"), db = "ncbi")
 
@@
                                                          Family = taxon_info$family)
+    }, error = function(e) {
+      # Print error message
+      cat("Error retrieving taxonomic information for", species_name, ":", conditionMessage(e), "\n")
+    })
   }
@@
 # Define the chunk size
-chunk_size <- 100  # Adjust the chunk size as needed
+chunk_size <- 300 # Adjust the chunk size as needed
 
@@
   all_taxonomic_info <- c(all_taxonomic_info, list(chunk_taxonomic_info))
+
+  # Add a delay 
+  Sys.sleep(5)
 }
@@
 taxonomic_df <- do.call(rbind, all_taxonomic_info)
+```
+
+```{r}
+# remove duplicates
+taxonomic_df <- distinct(taxonomic_df)
+# rename column to family, genus, species
+taxonomic_df <- taxonomic_df %>%
+  rename(family = Family,
+         genus = Genus,
+         species = Species)
+```
+
+```{r}
+# add to species dataframe
+na_species_df_2 <- na_species_df %>%
+  mutate(family = coalesce(family, taxonomic_df$family),
+         genus = coalesce(genus, taxonomic_df$genus))
+
+# retrieve list of species without family & genus information
+species_no_family_2 <- na_species_df_2 %>%
+  filter(is.na(family) | family == "") %>%
+  select(species) %>%
+  pull()
 ```

### 2024-05-01 code/L1/L1_P4_TropicalAndes_plant_imputetraits.Rmd
add genus and family info to na_species dataframes, get species with no family info and retrive that info from powo and add it to the dataframe, combine family and genus info to dataframe with species and traits, create a list of unique genus and families for each trait to pull from BIEN and GIFT

@@
 3. list by trait of species and genus
+
 ## Create a dataframe of species and TraitNames with TraitValue = NA
@@
 ```
-## Use taxize for species with no family and genus info
-
+Use taxize for species with no family and genus info
 ```{r}
@@
 # add to species dataframe
-na_species_df_2 <- na_species_df %>%
-  mutate(family = coalesce(family, taxonomic_df$family),
-         genus = coalesce(genus, taxonomic_df$genus))
+na_species_df$family <- ifelse(is.na(na_species_df$family), taxonomic_df$family[match(na_species_df$species, taxonomic_df$species)], na_species_df$family)
+na_species_df$genus <- ifelse(is.na(na_species_df$genus), taxonomic_df$genus[match(na_species_df$species, taxonomic_df$species)], na_species_df$genus)
 
 # retrieve list of species without family & genus information
-species_no_family_2 <- na_species_df_2 %>%
+species_no_family_2 <- na_species_df %>%
   filter(is.na(family) | family == "") %>%
@@
   pull()
+```
+
+```{r}
+# get powo id
+species_no_family_2_powo <- get_pow(species_no_family_2, messages = TRUE)
+```
+
+```{r}
+# Function to retrieve taxonomic information for a chunk of taxon IDs
+get_taxonomic_info_chunk <- function(chunk_taxon_id) {
+  # Initialize an empty list to store taxonomic information for each chunk
+  chunk_taxonomic_info <- list()
+
+  # Loop through each taxon ID in the chunk and retrieve taxonomic information
+  for (taxon_id in chunk_taxon_id) {
+    tryCatch({
+      # Print taxon ID for debugging
+      print(paste("Processing taxon ID:", taxon_id))
+
+      # Add a delay between consecutive API requests
+      Sys.sleep(1)
+
+      # Query POWO database for taxonomic information
+      taxon_info <- pow_lookup(id = taxon_id)
+
+      # Check if taxon_info is NULL
+      if (is.null(taxon_info)) {
+        # If taxon_info is NULL, return NA values for species, family, and genus
+        chunk_taxonomic_info[[taxon_id]] <- data.frame(taxon_id = taxon_id, species = NA, family = NA, genus = NA)
+      } else {
+        # Store taxonomic information for the taxon ID in the list
+        chunk_taxonomic_info[[taxon_id]] <- data.frame(taxon_id = taxon_id,
+                                                        species = taxon_info$meta$name,
+                                                        genus = taxon_info$meta$genus, 
+                                                        family = taxon_info$meta$family)
+      }
+    }, error = function(e) {
+      # Print error message for debugging
+      print(paste("Error processing taxon ID:", taxon_id))
+      print(e)
+    })
+  }
+
+  # Return the list of taxonomic information for the chunk
+  return(chunk_taxonomic_info)
+}
+
+# Define the chunk size
+chunk_size <- 300 # Adjust the chunk size as needed
+
+# Split the vector of taxon IDs into chunks based on the chunk size
+chunks <- split(species_no_family_2_powo, ceiling(seq_along(species_no_family_2_powo) / chunk_size))
+
+# Initialize an empty list to store taxonomic information for all chunks
+all_taxonomic_info <- list()
+
+# Iterate over each chunk and retrieve taxonomic information
+for (chunk_index in seq_along(chunks)) {
+  # Get the taxon IDs in the current chunk
+  chunk_taxon_id <- chunks[[chunk_index]]
+
+  # Retrieve taxonomic information for the chunk
+  chunk_taxonomic_info <- get_taxonomic_info_chunk(chunk_taxon_id)
+
+  # Append taxonomic information for the chunk to the list
+  all_taxonomic_info <- c(all_taxonomic_info, list(chunk_taxonomic_info))
+
+  # Add a delay 
+  Sys.sleep(5)
+}
+
+# Combine taxonomic information from all chunks into a single list
+all_taxonomic_info <- do.call(c, all_taxonomic_info)
+
+# Combine the list of dataframes into a single dataframe
+taxonomic_df <- do.call(rbind, all_taxonomic_info)
+
+```
+
+```{r}
+# add to species dataframe
+na_species_df$family <- ifelse(is.na(na_species_df$family), taxonomic_df$family[match(na_species_df$species, taxonomic_df$species)], na_species_df$family)
+na_species_df$genus <- ifelse(is.na(na_species_df$genus), taxonomic_df$genus[match(na_species_df$species, taxonomic_df$species)], na_species_df$genus)
+
+# retrieve list of species without family & genus information
+species_no_family_3 <- na_species_df %>%
+  filter(is.na(family) | family == "") %>%
+  select(species) %>%
+  pull()
+```
+
+## Add genus and family info to dataframe with species and traits
+```{r}
+na_traits_family.genus <- merge(na_traits, na_species_df, by = "species", all.x = TRUE)
+```
+
+## Get a list for each trait of species, genus, and family to fill in gaps
+```{r}
+# Get unique traits
+traits <- unique(na_traits_family.genus$TraitName)
+
+# Create an empty list to store dataframes for each trait
+trait_dfs <- list()
+
+# Split the dataframe by traitName and create a dataframe for each trait
+for (trait in traits) {
+  # Subset the dataframe for the current trait
+  trait_df <- subset(na_traits_family.genus, TraitName == trait)
+
+  # Remove the traitName column from the subsetted dataframe
+  trait_df <- trait_df[, !(names(trait_df) %in% "traitName")]
+
+  # Add the dataframe to the list
+  trait_dfs[[trait]] <- trait_df
+}
+
+# Seperate dataframes from list
+Dispersal.Syndrome <- trait_dfs$Dispersal.Syndrome
+Fruit.Length <- trait_dfs$Fruit.Length
+Plant.Height <- trait_dfs$Plant.Height
+Fruit.Type <- trait_dfs$Fruit.Type
+Growth.Form <- trait_dfs$Growth.Form
+```
+
+```{r}
+# Create a list of unique genus and families for each trait
+Dispersal.Syndrome_families <- unique(Dispersal.Syndrome$family)
+Dispersal.Syndrome_genus <- unique(Dispersal.Syndrome$genus)
+
+Fruit.Length_families <- unique(Fruit.Length$family)
+Fruit.Length_genus <- unique(Fruit.Length$genus)
+
+Plant.Height_families <- unique(Plant.Height$family)
+Plant.Height_genus <- unique(Plant.Height$genus)
+
+Fruit.Type_families <- unique(Fruit.Type$family)
+Fruit.Type_genus <- unique(Fruit.Type$genus)
+
+Growth.Form_families <- unique(Growth.Form$family)
+Growth.Form_genus <- unique(Growth.Form$genus)
+```
+
+##Retrive trait records for species using BIEN & GIFT
+```{r}
+
 ```

### 2024-05-03 code/L1/L1_P4_TropicalAndes_plant_imputetraits.Rmd
Update L1_P4_TropicalAndes_plant_imputetraits.Rmd
add code to retrive trait data from BIEN & GIFT using family and genus

@@
 Use taxize for species with no family and genus info
-```{r}
+```{r, results='hide',message=FALSE, warning=FALSE}
 library(taxize)
@@
 ```{r}
+# Dispersal Syndrome
+Dispersal.Syndrome_families_BIEN <- BIEN_trait_traitbyfamily(family = Dispersal.Syndrome_families, trait = "dispersal syndrome")
+Dispersal.Syndrome_genus_BIEN <- BIEN_trait_traitbygenus(genus = Dispersal.Syndrome_genus, trait)
 
+Dispersal.Syndrome_families_GIFT <- GIFT_traits_tax(trait_IDs = "3.3.1")
+Dispersal.Syndrome_families_GIFT <- Dispersal.Syndrome_families_GIFT[Dispersal.Syndrome_families_GIFT$family %in% Dispersal.Syndrome_families, ]
+```
+
+```{r}
+# Fruit Length
+Fruit.Length_families_BIEN <- BIEN_trait_traitbyfamily(family = Fruit.Length_families, trait = c("maximum fruit length", "minimum fruit length"))
+Fruit.Length_genus_BIEN <- BIEN_trait_traitbygenus(genus = Fruit.Length_genus, trait = c("maximum fruit length", "minimum fruit length"))
+
+Fruit.Length_families_GIFT <- GIFT_traits_tax(trait_IDs = "3.13.3")
+Fruit.Length_families_GIFT <- Fruit.Length_families_GIFT[Fruit.Length_families_GIFT$family %in% Fruit.Length_families, ]
+```
+
+```{r}
+# Plant Height
+Plant.Height_families_BIEN <- BIEN_trait_traitbyfamily(family = Plant.Height_families, trait = "whole plant height")
+Plant.Height_genus_BIEN <- BIEN_trait_traitbygenus(genus = Plant.Height_genus, trait = "whole plant height")
+
+Plant.Height_families_GIFT <- GIFT_traits_tax(trait_IDs = "1.6.3")
+Plant.Height_families_GIFT <- Plant.Height_families_GIFT[Plant.Height_families_GIFT$family %in% Plant.Height_families, ]
+```
+
+```{r}
+# Fruit Type
+Fruit.Type_families_BIEN <- BIEN_trait_traitbyfamily(family = Fruit.Type_families, trait = "fruit type")
+Fruit.Type_genus_BIEN <- BIEN_trait_traitbygenus(genus = Fruit.Type_genus, trait = "fruit type")
+
+Fruit.Type_families_GIFT <- GIFT_traits_tax(trait_IDs = "3.16.1")
+Fruit.Type_families_GIFT <- Fruit.Type_families_GIFT[Fruit.Type_families_GIFT$family %in% Fruit.Type_families, ]
+```
+
+```{r}
+# Growth Form
+Growth.Form_families_BIEN <- BIEN_trait_traitbyfamily(family = Growth.Form_families, trait = "whole plant growth form")
+Growth.Form_genus_BIEN <- BIEN_trait_traitbygenus(genus = Growth.Form_genus, trait = "whole plant growth form")
+
+Growth.Form_families_GIFT <- GIFT_traits_tax(trait_IDs = "1.2.1")
+Growth.Form_families_GIFT <- Growth.Form_families_GIFT[Growth.Form_families_GIFT$family %in% Growth.Form_families, ]
 ```

### 2024-05-07 code/L1/L1_P4_TropicalAndes_plant_imputetraits.Rmd
add code to subset dataframes to desired columns for family and genus level traits

@@
 
-##Retrive trait records for species using BIEN & GIFT
+## Retrive trait records for species using BIEN & GIFT
 ```{r}
 # Dispersal Syndrome
-Dispersal.Syndrome_families_BIEN <- BIEN_trait_traitbyfamily(family = Dispersal.Syndrome_families, trait = "dispersal syndrome")
-Dispersal.Syndrome_genus_BIEN <- BIEN_trait_traitbygenus(genus = Dispersal.Syndrome_genus, trait)
+Dispersal.Syndrome_families_BIEN <- BIEN_trait_traitbyfamily(family = Dispersal.Syndrome_families, trait = "whole plant dispersal syndrome")
+Dispersal.Syndrome_genus_BIEN <- BIEN_trait_traitbygenus(genus = Dispersal.Syndrome_genus, trait = "whole plant dispersal syndrome")
 
@@
 
-Fruit.Length_families_GIFT <- GIFT_traits_tax(trait_IDs = "3.13.3")
-Fruit.Length_families_GIFT <- Fruit.Length_families_GIFT[Fruit.Length_families_GIFT$family %in% Fruit.Length_families, ]
+#Error in GIFT_traits_tax(trait_IDs = "3.13.3") : None of the traits asked was available at the taxonomic level.
+#Fruit.Length_families_GIFT <- GIFT_traits_tax(trait_IDs = "3.13.3")
+#Fruit.Length_families_GIFT <- Fruit.Length_families_GIFT[Fruit.Length_families_GIFT$family %in% Fruit.Length_families, ]
 ```
@@
 
-Plant.Height_families_GIFT <- GIFT_traits_tax(trait_IDs = "1.6.3")
-Plant.Height_families_GIFT <- Plant.Height_families_GIFT[Plant.Height_families_GIFT$family %in% Plant.Height_families, ]
+#Error in GIFT_traits_tax(trait_IDs = "1.6.3") :   None of the traits asked was available at the taxonomic level.
+#Plant.Height_families_GIFT <- GIFT_traits_tax(trait_IDs = "1.6.3")
+#Plant.Height_families_GIFT <- Plant.Height_families_GIFT[Plant.Height_families_GIFT$family %in% Plant.Height_families, ]
 ```
@@
 
-Fruit.Type_families_GIFT <- GIFT_traits_tax(trait_IDs = "3.16.1")
-Fruit.Type_families_GIFT <- Fruit.Type_families_GIFT[Fruit.Type_families_GIFT$family %in% Fruit.Type_families, ]
+# Error in GIFT_traits_tax(trait_IDs = "3.16.1") :   None of the traits asked was available at the taxonomic level.
+#Fruit.Type_families_GIFT <- GIFT_traits_tax(trait_IDs = "3.16.1")
+#Fruit.Type_families_GIFT <- Fruit.Type_families_GIFT[Fruit.Type_families_GIFT$family %in% Fruit.Type_families, ]
 ```
@@
 
-Growth.Form_families_GIFT <- GIFT_traits_tax(trait_IDs = "1.2.1")
-Growth.Form_families_GIFT <- Growth.Form_families_GIFT[Growth.Form_families_GIFT$family %in% Growth.Form_families, ]
+# Error in GIFT_traits_tax(trait_IDs = "1.2.1") :   None of the traits asked was available at the taxonomic level.
+#Growth.Form_families_GIFT <- GIFT_traits_tax(trait_IDs = "1.2.1")
+#Growth.Form_families_GIFT <- Growth.Form_families_GIFT[Growth.Form_families_GIFT$family %in% Growth.Form_families, ]
 ```
 
-# Set the seed for reproducibility
+## Combine new info with existing trait table
+Make sure numeric traits match the units of existing trait table. 
+To fill in gaps of missing traits, use genus level data first then family level. Add column to trait dataframe of the level of trait measurement: species, genus, family.
+```{r}
+# for traits to family level from BIEN, keep columns of scrubbed_family, trait_name, trait_value
+Dispersal.Syndrome_families_BIEN_df <- Dispersal.Syndrome_families_BIEN[, c("scrubbed_family", "trait_name", "trait_value")]
+Fruit.Length_families_BIEN_df <- Fruit.Length_families_BIEN[, c("scrubbed_family", "trait_name", "trait_value")]
+Plant.Height_families_BIEN_df <- Plant.Height_families_BIEN[, c("scrubbed_family", "trait_name", "trait_value")]
+Fruit.Type_families_BIEN_df <- Fruit.Type_families_BIEN[, c("scrubbed_family", "trait_name", "trait_value")]
+Growth.Form_families_BIEN_df <- Growth.Form_families_BIEN[, c("scrubbed_family", "trait_name", "trait_value")]
+```
+
+```{r}
+# for traits to genus level from BIEN, keep columns of scrubbed_genus, trait_name, trait_value
+Dispersal.Syndrome_genus_BIEN_df <- Dispersal.Syndrome_genus_BIEN[, c("scrubbed_genus", "trait_name", "trait_value")]
+Fruit.Length_genus_BIEN_df <- Fruit.Length_genus_BIEN[, c("scrubbed_genus", "trait_name", "trait_value")]
+Plant.Height_genus_BIEN_df <- Plant.Height_genus_BIEN[, c("scrubbed_genus", "trait_name", "trait_value")]
+Fruit.Type_genus_BIEN_df <- Fruit.Type_genus_BIEN[, c("scrubbed_genus", "trait_name", "trait_value")]
+Growth.Form_genus_BIEN_df <- Growth.Form_genus_BIEN[, c("scrubbed_genus", "trait_name", "trait_value")]
+```
+
+```{r}
+
+```
+
+# Impute Traits
+## Set the seed for reproducibility
 ```{r}

### 2024-05-09 code/L1/L1_P4_TropicalAndes_plant_imputetraits.Rmd
get a single trait value for each family & genus, combine all family level trait data, combine all genus level trait data, add trait level column to dataframes, check column names for dataframes, combine non-imputed traits and family or genus level traits

@@
 
-# Seperate dataframes from list
+# Separate dataframes from list
 Dispersal.Syndrome <- trait_dfs$Dispersal.Syndrome
@@
 
+Get a single trait value for each family/genus
 ```{r}
+# numeric traits
+numeric_traits_combined <- function(df, level) {
+  df$trait_value <- as.numeric(df$trait_value)
+  if (level == "genus") {
+    df %>%
+      group_by(scrubbed_genus) %>%
+      summarise(trait_value = mean(trait_value, na.rm = TRUE)) %>%
+      ungroup()
+  } else if (level == "family") {
+    df %>%
+      group_by(scrubbed_family) %>%
+      summarise(trait_value = mean(trait_value, na.rm = TRUE)) %>%
+      ungroup()
+  }
+}
+
+# non-numeric traits
+cat_traits_combined <- function(df, level){
+  if (level == "genus") {
+    df %>%
+      group_by(scrubbed_genus) %>%
+      summarise_all(.funs = list(mode_trait = ~names(sort(table(.), decreasing = TRUE))[1])) %>%
+      ungroup()
+  } else if (level == "family") {
+    df %>%
+      group_by(scrubbed_family) %>%
+      summarise_all(.funs = list(mode_trait = ~names(sort(table(.), decreasing = TRUE))[1])) %>%
+      ungroup()
+  }
+}
 
 ```
+
+```{r}
+Dispersal.Syndrome_families_BIEN_df <- cat_traits_combined(Dispersal.Syndrome_families_BIEN_df, "family")
+Dispersal.Syndrome_genus_BIEN_df <- cat_traits_combined(Dispersal.Syndrome_genus_BIEN_df, "genus")
+```
+
+```{r}
+Fruit.Length_families_BIEN_df <- numeric_traits_combined(Fruit.Length_families_BIEN_df, "family")
+Fruit.Length_genus_BIEN_df <- numeric_traits_combined(Fruit.Length_genus_BIEN_df, "genus")
+```
+
+```{r}
+Plant.Height_families_BIEN_df <- numeric_traits_combined(Plant.Height_families_BIEN_df, "family")
+Plant.Height_genus_BIEN_df <- numeric_traits_combined(Plant.Height_genus_BIEN_df, "genus")
+```
+
+```{r}
+Fruit.Type_families_BIEN_df <- cat_traits_combined(Fruit.Type_families_BIEN_df, "family")
+Fruit.Type_genus_BIEN_df <- cat_traits_combined(Fruit.Type_genus_BIEN_df, "genus")
+```
+
+```{r}
+Growth.Form_families_BIEN_df <- cat_traits_combined(Growth.Form_families_BIEN_df, "family")
+Growth.Form_genus_BIEN_df <- cat_traits_combined(Growth.Form_genus_BIEN_df, "genus")
+```
+
+add column of trait level (specices, genus, family)
+```{r}
+# Add new column trait_level to long_plant_trait dataframe
+long_plant_traits <- long_plant_traits %>%
+  mutate(TraitLevel = ifelse(!is.na(TraitValue), "species", NA))
+```
+
+```{r}
+# add family and genus information to long_plant_trait dataframe
+long_plant_traits_tax <- merge(long_plant_traits, na_species_df, by = "species", all.x = TRUE)
+```
+
+```{r}
+# combine all family level trait data
+## add column of trait_name for numeric traits
+Dispersal.Syndrome_families_BIEN_df$trait_name <- "Dispersal.Syndrome"
+Fruit.Length_families_BIEN_df$trait_name <- "Fruit.Length_mm"
+Plant.Height_families_BIEN_df$trait_name <- "Plant.Height_m"
+Fruit.Type_families_BIEN_df$trait_name <- "Fruit.Type"
+Growth.Form_families_BIEN_df$trait_name <- "Growth.Form"
+
+## rename trait_value column so it's consistent with other traits
+Fruit.Type_families_BIEN_df$trait_value <- Fruit.Type_families_BIEN_df$trait_value_mode_trait
+Growth.Form_families_BIEN_df$trait_value <- Growth.Form_families_BIEN_df$trait_value_mode_trait
+
+## convert numeric trait values into character
+Plant.Height_families_BIEN_df$trait_value <- as.character(Plant.Height_families_BIEN_df$trait_value)
+Fruit.Length_families_BIEN_df$trait_value <- as.character(Fruit.Length_families_BIEN_df$trait_value)
+
+all_family_traits <- bind_rows(
+  select(Dispersal.Syndrome_families_BIEN_df, scrubbed_family, trait_name, trait_value),
+  select(Fruit.Length_families_BIEN_df, scrubbed_family, trait_name, trait_value),
+  select(Plant.Height_families_BIEN_df, scrubbed_family, trait_name, trait_value),
+  select(Fruit.Type_families_BIEN_df, scrubbed_family, trait_name, trait_value),
+  select(Growth.Form_families_BIEN_df, scrubbed_family, trait_name, trait_value)
+  )
+
+## add trait level column
+all_family_traits$TraitLevel <- "family"
+```
+
+```{r}
+# combine all genus level trait data
+## add column of trait_name for numeric traits
+Dispersal.Syndrome_genus_BIEN_df$trait_name <- "Dispersal.Syndrome"
+Fruit.Length_genus_BIEN_df$trait_name <- "Fruit.Length_mm"
+Plant.Height_genus_BIEN_df$trait_name <- "Plant.Height_m"
+Fruit.Type_genus_BIEN_df$trait_name <- "Fruit.Type"
+Growth.Form_genus_BIEN_df$trait_name <- "Growth.Form"
+
+## rename trait_value column so it's consistent with other traits
+Fruit.Type_genus_BIEN_df$trait_value <- Fruit.Type_genus_BIEN_df$trait_value_mode_trait
+Growth.Form_genus_BIEN_df$trait_value <- Growth.Form_genus_BIEN_df$trait_value_mode_trait
+
+## convert numeric trait values into character
+Plant.Height_genus_BIEN_df$trait_value <- as.character(Plant.Height_genus_BIEN_df$trait_value)
+Fruit.Length_genus_BIEN_df$trait_value <- as.character(Fruit.Length_genus_BIEN_df$trait_value)
+
+all_genus_traits <- bind_rows(
+  select(Dispersal.Syndrome_genus_BIEN_df, scrubbed_genus, trait_name, trait_value),
+  select(Fruit.Length_genus_BIEN_df, scrubbed_genus, trait_name, trait_value),
+  select(Plant.Height_genus_BIEN_df, scrubbed_genus, trait_name, trait_value),
+  select(Fruit.Type_genus_BIEN_df, scrubbed_genus, trait_name, trait_value),
+  select(Growth.Form_genus_BIEN_df, scrubbed_genus, trait_name, trait_value)
+  )
+
+## add trait level column
+all_genus_traits$TraitLevel <- "genus"
+```
+
+make sure column names match between dataframes before combining
+```{r}
+colnames(all_family_traits)
+colnames(all_genus_traits)
+colnames(long_plant_traits_tax)
+```
+
+```{r}
+colnames(all_family_traits)[which(colnames(all_family_traits) == "scrubbed_family")] <- "family"
+colnames(all_family_traits)[which(colnames(all_family_traits) == "trait_name")] <- "TraitName"
+colnames(all_family_traits)[which(colnames(all_family_traits) == "trait_value")] <- "TraitValue"
+
+colnames(all_genus_traits)[which(colnames(all_genus_traits) == "scrubbed_genus")] <- "genus"
+colnames(all_genus_traits)[which(colnames(all_genus_traits) == "trait_name")] <- "TraitName"
+colnames(all_genus_traits)[which(colnames(all_genus_traits) == "trait_value")] <- "TraitValue"
+```
+
+use family & genus level traits to fill in gaps of original dataframe
+genus level prefered over family
+```{r}
+# Join the dataframes based on genus and TraitName
+filled_long_plant_traits_tax <- long_plant_traits_tax %>%
+  left_join(all_genus_traits, by = c("genus", "TraitName"), relationship = "many-to-many") %>%
+  # Fill NAs in TraitValue column with values from all_genus_traits
+  mutate(TraitValue = ifelse(is.na(TraitValue.x), TraitValue.y, TraitValue.x)) %>%
+  # Remove the columns from all_genus_traits (if needed)
+  select(-TraitValue.y)
+```
+
+Add units to wide trait table numeric trait names
+fruit mass -> Fruit.Mass_mg
+fruit length -> Fruit.Length_mm
+plant height -> Plant.Height_m
+plant lifespan -> Plant.Lifespan_years
+seed mass -> Seed.Mass_mg
+seed length -> Seed.Length_mm
+seed width -> Seed.Width_mm
 

### 2024-05-10 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
update input file for plant occ, create function for richness plots to reduce redundant code

--- 2023-12-13 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
@@
 ```{r}
-TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
+TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned_subset.csv"))
 TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_frugivore_occ_cleaned_subset.csv"))
@@
 
-# Plot TD/Richness
-
-## 5 km resolution
-### Make Grid
 ```{r}
@@
 
+# Plot TD/Richness
+function for richness plots
 ```{r}
-TAGrid_5 <- TApoly %>%
-  st_make_grid(cellsize = c(5000)) %>%
+create_rich_plots <- function(resolution_meters) {
+  # Make Grid
+  TAGrid <- TApoly %>%
+    st_make_grid(cellsize = c(resolution_meters)) %>%
     st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
@@
     mutate(cellid = row_number())
-```
 
-### Plot Empty Grid
-```{r}
-gridPlot <-
-ggplot() +
+  # Plot Empty Grid
+  gridPlot <- ggplot() +
     geom_sf(data = Americas, fill = "white") +
     geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
-  geom_sf(data = TAGrid_5) +
-  labs(title = "Tropical Andes Grid [5 km]", x = "Latitude", y = "Longitude") +
+    geom_sf(data = TAGrid) +
+    labs(title = paste("Tropical Andes Grid [", resolution_meters/1000, " km]", sep=""), x = "Latitude", y = "Longitude") +
     coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
     annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+    annotation_north_arrow(location = "bl", which_north = "true", pad_x = unit(0.3, "in"), 
+                           pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
     theme(panel.background = element_rect(fill = "lightblue"))
 
-gridPlot
-```
+  print(gridPlot)
 
-### Cell Richness
-```{r}
-plant_richness_grid_5 <- TAGrid_5 %>%
+  # Calculate cell richness for plants
+  plant_richness_grid <- TAGrid %>%
     st_join(plants_sf_species) %>%
@@
     summarize(num_species = sum(overlap))
-```
 
-```{r}
-frugivore_richness_grid_5 <- TAGrid_5 %>%
+  # Calculate cell richness for frugivores
+  frugivore_richness_grid <- TAGrid %>%
     st_join(frugivores_sf_species) %>%
@@
     summarize(num_species = sum(overlap))
-```
 
-### Richness plots
-```{r}
-plantgridRichTA_5km <-
-ggplot(plant_richness_grid_5) +
+  # Plot richness for plants
+  plantgridRichTA <- ggplot(plant_richness_grid) +
     geom_sf(data = Americas, fill = "white") +
@@
     scale_fill_scico(palette = "romaO", direction = 1) +
-  labs(title = "Tropical Andes Plant Richness [5 km]", x = "Longitude", y = "Latitude",
+    labs(title = paste("Tropical Andes Plant Richness [", resolution_meters/1000, " km]", sep=""), x = "Longitude", y = "Latitude",
          fill = "Number of species") +
@@
 
-plantgridRichTA_5km
-```
+  print(plantgridRichTA)
 
-```{r}
-frugivoregridRichTA_5km <-
-ggplot(frugivore_richness_grid_5) +
+  # Plot richness for frugivores
+  frugivoregridRichTA <- ggplot(frugivore_richness_grid) +
     geom_sf(data = Americas, fill = "white") +
@@
     scale_fill_scico(palette = "romaO", direction = 1) +
-  labs(title = "Tropical Andes Frugivore Richness [5 km]", x = "Longitude", y = "Latitude",
+    labs(title = paste("Tropical Andes Frugivore Richness [", resolution_meters/1000, " km]", sep=""), x = "Longitude", y = "Latitude",
          fill = "Number of species") +
@@
 
-frugivoregridRichTA_5km
+  print(frugivoregridRichTA)
+
+  # Extract cell values
+  plant_cellRichness <- plantgridRichTA[["data"]]
+  frugivore_cellRichness <- frugivoregridRichTA[["data"]]
+
+  ### Plant vs frugivore richness plot
+  compare_richness_summary <- summary(lm(frugivore_cellRichness$num_species ~ plant_cellRichness$num_species))
+  print(compare_richness_summary)
+
+  # save r-squared value
+  compare_richness_r2 <- compare_richness_summary$r.squared
+
+  plant_frug_richness_compare <- plot(plant_cellRichness$num_species, frugivore_cellRichness$num_species,
+       main = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
+       xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
+  abline(lm(frugivore_cellRichness$num_species ~ plant_cellRichness$num_species))
+
+  print(plant_frug_richness_compare)
+
+  plant_richness_hist <- hist(plant_cellRichness$num_species,
+       main = paste("Histogram of Plant Richness [", resolution_meters/1000, " km] by cell", sep = ""),
+       xlab = "Plant Richness by cell")
+  print(plant_richness_hist)
+  frugivore_richness_hist <- hist(frugivore_cellRichness$num_species,
+       main = paste("Histogram of Frugivore Richness [", resolution_meters/1000, " km] by cell", sep = ""),
+       xlab = "Frugivore Richness by cell")
+  print(frugivore_richness_hist)
+}
+
 ```
 
-## Extract cell values
+## 1 km resolution
 ```{r}
-plant_cellRichness_5km <- plantgridRichTA_5km[["data"]]
-frugivore_cellRichness_5km <- frugivoregridRichTA_5km[["data"]]
+richness_1km <- create_rich_plots(1000)
+# save plots
+plantgridRichTA_1km <- plantgridRichTA
+frugivoregridRichTA_1km <- frugivoregridRichTA
+# save richness data
+plant_cellRichness_1km <- plant_cellRichness
+frugivore_cellRichness_1km <- frugivore_cellRichness
+
+compare_richness_r2_1km <- compare_richness_r2
 ```
 
-### Plant vs frugivore richness plot
+## 5 km resolution
 ```{r}
-compare_richness_summary_5km <- summary(lm(frugivore_cellRichness_5km$num_species ~ plant_cellRichness_5km$num_species))
-compare_richness_summary_5km
+richness_5km <- create_rich_plots(5000)
+# save plots
+plantgridRichTA_5km <- plantgridRichTA
+frugivoregridRichTA_5km <- frugivoregridRichTA
+## Extract cell values
+plant_cellRichness_5km <- plant_cellRichness
+frugivore_cellRichness_5km <- frugivore_cellRichness
 
-# save r-squared value
-compare_richness_r2_5km <- compare_richness_summary_5km$r.squared
+compare_richness_r2_5km <- compare_richness_r2
 ```
 
+## 10 km
 ```{r}
-plot(plant_cellRichness_5km$num_species, frugivore_cellRichness_5km$num_species,
-     main = "Plant Richness vs Frugivore Richness [5 km]",
-     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
-abline(lm(frugivore_cellRichness_5km$num_species ~ plant_cellRichness_5km$num_species))
-```
+richness_10km <- create_rich_plots(10000)
+# save plots
+plantgridRichTA_10km <- plantgridRichTA
+frugivoregridRichTA_10km <- frugivoregridRichTA
+## Extract cell values
+plant_cellRichness_10km <- plant_cellRichness
+frugivore_cellRichness_10km <- frugivore_cellRichness
 
-```{r}
-hist(plant_cellRichness_5km$num_species,
-     main = "Histogram of Plant Richness by cell",
-     xlab = "Plant Richness by cell")
-hist(frugivore_cellRichness_5km$num_species,
-     main = "Histogram of Frugivore Richness by cell",
-     xlab = "Frugivore Richness by cell")
+compare_richness_r2_10km <- compare_richness_r2
 ```
@@
 ## 25 km
-### Make Grid
 ```{r}
-TAGrid_25 <- TApoly %>%
-  st_make_grid(cellsize = c(25000)) %>%
-  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
-  st_cast("MULTIPOLYGON") %>%
-  st_sf() %>%
-  mutate(cellid = row_number())
-```
+richness_25km <- create_rich_plots(25000)
+# save plots
+plantgridRichTA_25km <- plantgridRichTA
+frugivoregridRichTA_25km <- frugivoregridRichTA
+## Extract cell values
+plant_cellRichness_25km <- plant_cellRichness
+frugivore_cellRichness_25km <- frugivore_cellRichness
 
-### Plot Empty Grid
-```{r}
-gridPlot <-
-ggplot() +
-  geom_sf(data = Americas, fill = "white") +
-  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
-  geom_sf(data = TAGrid_25) +
-  labs(title = "Tropical Andes Grid [25 km]", x = "Latitude", y = "Longitude") +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  theme(panel.background = element_rect(fill = "lightblue"))
-
-gridPlot
-```
-
-### Cell Richness
-```{r}
-plant_richness_grid_25 <- TAGrid_25 %>%
-  st_join(plants_sf_species) %>%
-  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
-  group_by(cellid) %>%
-  summarize(num_species = sum(overlap))
-```
-
-```{r}
-frugivore_richness_grid_25 <- TAGrid_25 %>%
-  st_join(frugivores_sf_species) %>%
-  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
-  group_by(cellid) %>%
-  summarize(num_species = sum(overlap))
-```
-
-### Richness plots
-```{r}
-plantgridRichTA_25km <-
-ggplot(plant_richness_grid_25) +
-  geom_sf(data = Americas, fill = "white") +
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_sf(aes(fill = num_species), color = NA) +
-  scale_fill_scico(palette = "romaO", direction = 1) +
-  labs(title = "Tropical Andes Plant Richness [25 km]", x = "Longitude", y = "Latitude",
-       fill = "Number of species") +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  theme(panel.background = element_rect(fill = "lightblue"))
-
-plantgridRichTA_25km
-```
-
-```{r}
-frugivoregridRichTA_25km <-
-ggplot(frugivore_richness_grid_25) +
-  geom_sf(data = Americas, fill = "white") +
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_sf(aes(fill = num_species), color = NA) +
-  scale_fill_scico(palette = "romaO", direction = 1) +
-  labs(title = "Tropical Andes Frugivore Richness [25 km]", x = "Longitude", y = "Latitude",
-       fill = "Number of species") +
-    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  theme(panel.background = element_rect(fill = "lightblue"))
-
-frugivoregridRichTA_25km
-```
-
-## Extract cell values
-```{r}
-plant_cellRichness_25km <- plantgridRichTA_25km[["data"]]
-frugivore_cellRichness_25km <- frugivoregridRichTA_25km[["data"]]
-```
-
-### Plant vs frugivore richness plot
-```{r}
-compare_richness_summary_25km <- summary(lm(frugivore_cellRichness_25km$num_species ~ plant_cellRichness_25km$num_species))
-compare_richness_summary_25km
-
-# save r-squared value
-compare_richness_r2_25km <- compare_richness_summary_25km$r.squared
-```
-
-```{r}
-plot(plant_cellRichness_25km$num_species, frugivore_cellRichness_25km$num_species,
-     main = "Plant Richness vs Frugivore Richness [25 km]",
-     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
-abline(lm(frugivore_cellRichness_25km$num_species ~ plant_cellRichness_25km$num_species))
-```
-
-```{r}
-hist(plant_cellRichness_25km$num_species,
-     main = "Histogram of Plant Richness by cell",
-     xlab = "Plant Richness by cell")
-hist(frugivore_cellRichness_25km$num_species,
-     main = "Histogram of Frugivore Richness by cell",
-     xlab = "Frugivore Richness by cell")
+compare_richness_r2_25km <- compare_richness_r2
 ```
@@
 ## 50 km
-### Make Grid
 ```{r}
-TAGrid_50 <- TApoly %>%
-  st_make_grid(cellsize = c(50000)) %>%
-  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
-  st_cast("MULTIPOLYGON") %>%
-  st_sf() %>%
-  mutate(cellid = row_number())
-```
+richness_50km <- create_rich_plots(50000)
+# save plots
+plantgridRichTA_50km <- plantgridRichTA
+frugivoregridRichTA_50km <- frugivoregridRichTA
+## Extract cell values
+plant_cellRichness_50km <- plant_cellRichness
+frugivore_cellRichness_50km <- frugivore_cellRichness
 
-### Plot Empty Grid
-```{r}
-gridPlot <-
-ggplot() +
-  geom_sf(data = Americas, fill = "white") +
-  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
-  geom_sf(data = TAGrid_50) +
-  labs(title = "Tropical Andes Grid [50 km]", x = "Latitude", y = "Longitude") +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  theme(panel.background = element_rect(fill = "lightblue"))
-
-gridPlot
-```
-
-### Cell Richness
-```{r}
-plant_richness_grid_50 <- TAGrid_50 %>%
-  st_join(plants_sf_species) %>%
-  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
-  group_by(cellid) %>%
-  summarize(num_species = sum(overlap))
-```
-
-```{r}
-frugivore_richness_grid_50 <- TAGrid_50 %>%
-  st_join(frugivores_sf_species) %>%
-  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
-  group_by(cellid) %>%
-  summarize(num_species = sum(overlap))
-```
-
-### Richness plots
-```{r}
-plantgridRichTA_50km <-
-ggplot(plant_richness_grid_50) +
-  geom_sf(data = Americas, fill = "white") +
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_sf(aes(fill = num_species), color = NA) +
-  scale_fill_scico(palette = "romaO", direction = 1) +
-  labs(title = "Tropical Andes Plant Richness [50 km]", x = "Longitude", y = "Latitude",
-       fill = "Number of species") +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  theme(panel.background = element_rect(fill = "lightblue"))
-
-plantgridRichTA_50km
-```
-
-```{r}
-frugivoregridRichTA_50km <-
-ggplot(frugivore_richness_grid_50) +
-  geom_sf(data = Americas, fill = "white") +
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_sf(aes(fill = num_species), color = NA) +
-  scale_fill_scico(palette = "romaO", direction = 1) +
-  labs(title = "Tropical Andes Frugivore Richness [50 km]", x = "Longitude", y = "Latitude",
-       fill = "Number of species") +
-    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  theme(panel.background = element_rect(fill = "lightblue"))
-
-frugivoregridRichTA_50km
-```
-
-## Extract cell values
-```{r}
-plant_cellRichness_50km <- plantgridRichTA_50km[["data"]]
-frugivore_cellRichness_50km <- frugivoregridRichTA_50km[["data"]]
-```
-
-### Plant vs frugivore richness plot
-```{r}
-compare_richness_summary_50km <- summary(lm(frugivore_cellRichness_50km$num_species ~ plant_cellRichness_50km$num_species))
-compare_richness_summary_50km
-
-# save r-squared value
-compare_richness_r2_50km <- compare_richness_summary_50km$r.squared
-```
-
-```{r}
-plot(plant_cellRichness_50km$num_species, frugivore_cellRichness_50km$num_species,
-     main = "Plant Richness vs Frugivore Richness [50 km]",
-     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
-abline(lm(frugivore_cellRichness_50km$num_species ~ plant_cellRichness_50km$num_species))
-```
-
-```{r}
-hist(plant_cellRichness_50km$num_species,
-     main = "Histogram of Plant Richness by cell",
-     xlab = "Plant Richness by cell")
-hist(frugivore_cellRichness_50km$num_species,
-     main = "Histogram of Frugivore Richness by cell",
-     xlab = "Frugivore Richness by cell")
+compare_richness_r2_50km <- compare_richness_r2
 ```
@@
 ## 75 km 
-### Make Grid
 ```{r}
-TAGrid_75 <- TApoly %>%
-  st_make_grid(cellsize = c(75000)) %>%
-  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
-  st_cast("MULTIPOLYGON") %>%
-  st_sf() %>%
-  mutate(cellid = row_number())
+richness_75km <- create_rich_plots(75000)
+# save plots
+plantgridRichTA_75km <- plantgridRichTA
+frugivoregridRichTA_75km <- frugivoregridRichTA
+## Extract cell values
+plant_cellRichness_75km <- plant_cellRichness
+frugivore_cellRichness_75km <- frugivore_cellRichness
+
+compare_richness_r2_75km <- compare_richness_r2
 ```
 
-### Plot Empty Grid
+## 100 km 
 ```{r}
-gridPlot <-
-ggplot() +
-  geom_sf(data = Americas, fill = "white") +
-  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
-  geom_sf(data = TAGrid_75) +
-  labs(title = "Tropical Andes Grid [75 km]", x = "Latitude", y = "Longitude") +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  theme(panel.background = element_rect(fill = "lightblue"))
+richness_100km <- create_rich_plots(100000)
+# save plots
+plantgridRichTA_100km <- plantgridRichTA
+frugivoregridRichTA_100km <- frugivoregridRichTA
+## Extract cell values
+plant_cellRichness_100km <- plant_cellRichness
+frugivore_cellRichness_100km <- frugivore_cellRichness
 
-gridPlot
-```
-
-### Cell Richness
-```{r}
-plant_richness_grid_75 <- TAGrid_75 %>%
-  st_join(plants_sf_species) %>%
-  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
-  group_by(cellid) %>%
-  summarize(num_species = sum(overlap))
-```
-
-```{r}
-frugivore_richness_grid_75 <- TAGrid_75 %>%
-  st_join(frugivores_sf_species) %>%
-  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
-  group_by(cellid) %>%
-  summarize(num_species = sum(overlap))
-```
-
-### Richness plots
-```{r}
-plantgridRichTA_75km <-
-ggplot(plant_richness_grid_75) +
-  geom_sf(data = Americas, fill = "white") +
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_sf(aes(fill = num_species), color = NA) +
-  scale_fill_scico(palette = "romaO", direction = 1) +
-  labs(title = "Tropical Andes Plant Richness [75 km]", x = "Longitude", y = "Latitude",
-       fill = "Number of species") +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  theme(panel.background = element_rect(fill = "lightblue"))
-
-plantgridRichTA_75km
-```
-
-```{r}
-frugivoregridRichTA_75km <-
-ggplot(frugivore_richness_grid_75) +
-  geom_sf(data = Americas, fill = "white") +
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_sf(aes(fill = num_species), color = NA) +
-  scale_fill_scico(palette = "romaO", direction = 1) +
-  labs(title = "Tropical Andes Frugivore Richness [75 km]", x = "Longitude", y = "Latitude",
-       fill = "Number of species") +
-    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  theme(panel.background = element_rect(fill = "lightblue"))
-
-frugivoregridRichTA_75km
-```
-
-## Extract cell values
-```{r}
-plant_cellRichness_75km <- plantgridRichTA_75km[["data"]]
-frugivore_cellRichness_75km <- frugivoregridRichTA_75km[["data"]]
-```
-
-### Plant vs frugivore richness plot
-```{r}
-compare_richness_summary_75km <- summary(lm(frugivore_cellRichness_75km$num_species ~ plant_cellRichness_75km$num_species))
-compare_richness_summary_75km
-
-# save r-squared value
-compare_richness_r2_75km <- compare_richness_summary_75km$r.squared
+compare_richness_r2_100km <- compare_richness_r2
 ```
@@
 Plot
+```{r}
+# create dataframe of resolutions and r-squared values
+resolution <- c("5", "25", "50", "75", "100")
+r_squared <- c(compare_richness_r2_5km, compare_richness_r2_25km, compare_richness_r2_50km, compare_richness_r2_75km, compare_richness_r2_100km)
+r_squared_compare <- data.frame(resolution, r-squared)
+```
+
+```{r}
+ggplot(r_squared_compare, aes(x=resolution, y=r_squared)) +
+  geom_point()+
+  labs(title = "R-squared values of plant-frugivore richness relationships by resolution", x = "Resolution (km)", y = "R-squared",
+       fill = "Number of species") 
+```
+
 # Multi panel plots
 Richness maps
+```{r}
+grid.arrange(plot_1, plot_2, nrow = 1)
+```
+
 # Write data to csv

### 2024-05-10 code/L1/taxharmonization_troubleshooting.R
move files to archive

@@ -1,32 +1,0 @@
-# Trouble shooting adding accepted names from lookup table to occurrence data
-
-library(dplyr)
-
-## Set file paths
-data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
-data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
-output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
-
-## Load files
-GBIF_occ <- read.csv(file.path(data_path_L1, "TropicalAndes_GBIF_plant_occ_cleaned.csv"))
-lookup_table <- read.csv(file.path(output_path,"plant_lookup_table.csv"))
-
-#small subset for testing
-lookup_table_5 <- lookup_table[1:5, 2:3]
-
-GBIF_occ_subset_5 <- GBIF_occ %>%
-  filter(species %in% lookup_table_5$Name_submitted)
-nrow(GBIF_occ_subset_5)
-
-# x is lookuptable  y is occurrences
-GBIF_occ_subset_harmonized_5 <- right_join(lookup_table_5, GBIF_occ_subset_5, join_by(Name_submitted == species), keep = TRUE)
-nrow(GBIF_occ_subset_harmonized_5)
-
-## filter data to species in lookup_table$Names_submitted
-GBIF_occ_subset <- GBIF_occ %>%
-  filter(species %in% lookup_table$Name_submitted)
-nrow(GBIF_occ_subset)
-
-## add accepted species column from lookup table based on species column
-GBIF_occ_subset_harmonized <- right_join(lookup_table, GBIF_occ_subset, join_by(Name_submitted == species), keep = TRUE)
-nrow(GBIF_occ_subset_harmonized)

### 2024-05-10 code/L1/L1_P2_TropicalAndes_taxcleaning.Rmd

@@
                             filter(Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", "", NA)) %>%
-                            select(Name_submitted)
+                            dplyr::select(Name_submitted)
 nrow(all_species_disparities)

### 2024-05-10 code/L1/L1_P3_TropicalAndes_plant_traits.Rmd

@@
 ```
+
+# Combine all traits
+```{r}
+# add unit column to categorical traits
+cat_traits <- rbind(averagetraits_dispersal, averagetraits_fruitType, averagetraits_growthForm)
+cat_traits$Unit <- ""
+clean_traits <- rbind(averagetraits_fruitLength, averagetraits_fruitMass, averagetraits_plantHeight, averagetraits_plantLifespan, averagetraits_seedLength, averagetraits_seedMass, averagetraits_seedWidth, cat_traits)
+nrow(clean_traits)
+```
+
+```{r}
+clean_traits %>%
+  count(TraitName)
+```
+
+# Summary
+```{r}
+print("Number of records: ")
+nrow(clean_traits)
+print("Number of species: ")
+length(unique(clean_traits$Accepted_species))
+```
+
+# Get one trait value per species
+```{r}
+#remove unit column from dataframe
+clean_traits <- clean_traits[, !colnames(clean_traits) %in% "Unit", drop = FALSE]
+```
+
+```{r}
+# separate numeric and non-numeric traits
+numeric_traits <- clean_traits[clean_traits$TraitName %in% c("Fruit Length", "Fruit Mass", "Plant Height", "Plant Lifespan", "Seed Length", "Seed Width", "Seed Mass"), ]
+cat_traits <- clean_traits[clean_traits$TraitName %in% c("Dispersal Syndrome", "Fruit Type", "Growth Form"), ]
+
+numeric_traits$TraitValue <- as.numeric(numeric_traits$TraitValue)
+```
+
+```{r}
+# numeric traits
+## Group by species and TraitName, and calculate the mean TraitValue for each combination
+numeric_traits_combined <- numeric_traits %>%
+  group_by(species, TraitName) %>%
+  summarise(TraitValue = mean(TraitValue, na.rm = TRUE)) %>%
+  ungroup()
+
+# non-numeric traits
+cat_traits_combined <- cat_traits %>%
+  group_by(species) %>%
+  summarise_all(.funs = list(mode_trait = function(x) {
+    mode_value <- names(sort(table(x), decreasing = TRUE))[1]  # Calculate mode
+    return(mode_value)
+  }))
+```
+
+```{r}
+#rename columns to original values
+colnames(cat_traits_combined)[which(colnames(cat_traits_combined) == "TraitValue_mode_trait")] <- "TraitValue"
+colnames(cat_traits_combined)[which(colnames(cat_traits_combined) == "TraitName_mode_trait")] <- "TraitName"
+
+# Remove rows with "?" or numbers in the TraitValue column
+cleaned_cat_traits <- cat_traits_combined %>%
+  filter(!grepl("\\?", TraitValue) & !grepl("\\d", TraitValue))
+```
+
+```{r}
+#combine numeric and non-numeric traits
+combined_clean_traits <- rbind(cat_traits_combined, numeric_traits_combined)
+```
+
+```{r}
+combined_clean_traits %>%
+  count(TraitName)
+```
+
+# Remove traits with less than 10 records
+```{r}
+combined_clean_traits <- combined_clean_traits[!(combined_clean_traits$TraitName %in% c("Fruit Mass", "Seed Mass")), ]
+```
+
+# Transform plant dataframe from long to wide
+```{r}
+wide_traits <- spread(combined_clean_traits, key = TraitName, value = TraitValue)
+
+#remove X column from dataframe
+wide_plant_traits <- wide_traits[, !colnames(wide_traits) %in% "X", drop = FALSE]
+```
+
+```{r}
+# Assign data types
+wide_plant_traits$`Fruit Length` <- as.numeric(wide_plant_traits$`Fruit Length`)
+wide_plant_traits$`Plant Height` <- as.numeric(wide_plant_traits$`Plant Height`)
+wide_plant_traits$`Dispersal Syndrome` <- as.character(wide_plant_traits$`Dispersal Syndrome`)
+wide_plant_traits$`Growth Forme` <- as.character(wide_plant_traits$`Growth Form`)
+wide_plant_traits$`Fruit Type` <- as.character(wide_plant_traits$`Fruit Type`)
+```
+
+# Summary
+```{r}
+print("Number of species: ")
+length(unique(wide_plant_traits$species))
+```
+
+# Write data to csv
+```{r}
+write.csv(wide_plant_traits, file.path(output_path,"TropicalAndes_all_plant_traits_harmonized.csv"))
+write.csv(clean_traits, file.path(output_path,"TropicalAndes_all_plant_traits_harmonized_with_units.csv"))
+```

### 2024-05-10 code/L2/L2_1_TropicalAndes_taxdiv.Rmd

@@
 
-  # Plot Empty Grid
-  gridPlot <- ggplot() +
-    geom_sf(data = Americas, fill = "white") +
-    geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
-    geom_sf(data = TAGrid) +
-    labs(title = paste("Tropical Andes Grid [", resolution_meters/1000, " km]", sep=""), x = "Latitude", y = "Longitude") +
-    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-    annotation_scale(location = "bl", width_hint = 0.5) +
-    annotation_north_arrow(location = "bl", which_north = "true", pad_x = unit(0.3, "in"), 
-                           pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-    theme(panel.background = element_rect(fill = "lightblue"))
-
-  print(gridPlot)
-
   # Calculate cell richness for plants
@@
 
-  # Plot richness for plants
-  plantgridRichTA <- ggplot(plant_richness_grid) +
+  # Helper function to generate plots
+  generate_plot <- function(data, title) {
+    ggplot(data) +
       geom_sf(data = Americas, fill = "white") +
@@
       scale_fill_scico(palette = "romaO", direction = 1) +
-    labs(title = paste("Tropical Andes Plant Richness [", resolution_meters/1000, " km]", sep=""), x = "Longitude", y = "Latitude",
-         fill = "Number of species") +
+      labs(title = title, x = "Longitude", y = "Latitude", fill = "Number of species") +
       coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
@@
       theme(panel.background = element_rect(fill = "lightblue"))
+  }
 
-  print(plantgridRichTA)
+  # Generate plots
+  plant_plot <- generate_plot(plant_richness_grid, paste("Tropical Andes Plant Richness [", resolution_meters/1000, " km]", sep=""))
+  frugivore_plot <- generate_plot(frugivore_richness_grid, paste("Tropical Andes Frugivore Richness [", resolution_meters/1000, " km]", sep=""))
 
-  # Plot richness for frugivores
-  frugivoregridRichTA <- ggplot(frugivore_richness_grid) +
-    geom_sf(data = Americas, fill = "white") +
-    geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-    geom_sf(aes(fill = num_species), color = NA) +
-    scale_fill_scico(palette = "romaO", direction = 1) +
-    labs(title = paste("Tropical Andes Frugivore Richness [", resolution_meters/1000, " km]", sep=""), x = "Longitude", y = "Latitude",
-         fill = "Number of species") +
-    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-    annotation_scale(location = "bl", width_hint = 0.5) +
-    annotation_north_arrow(location = "bl", which_north = "true", 
-                           pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-    theme(panel.background = element_rect(fill = "lightblue"))
+  # Compare plant and frugivore richness
+  compare_richness_summary <- summary(
+    lm(frugivore_richness_grid$num_species ~ plant_richness_grid$num_species))
 
-  print(frugivoregridRichTA)
+  comparison_data <- data.frame(Plant_Richness = plant_richness_grid$num_species,
+                               Frugivore_Richness = frugivore_richness_grid$num_species)
 
-  # Extract cell values
-  plant_cellRichness <- plantgridRichTA[["data"]]
-  frugivore_cellRichness <- frugivoregridRichTA[["data"]]
+  # Create scatter plot
+  plant_frug_richness_compare <- ggplot(comparison_data,
+                                        aes(x = Plant_Richness, y = Frugivore_Richness)) +
+    geom_point() +
+    labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]",
+                       sep = ""), x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
+    geom_smooth(method = "lm", se = FALSE, color = "red")
 
-  ### Plant vs frugivore richness plot
-  compare_richness_summary <- summary(lm(frugivore_cellRichness$num_species ~ plant_cellRichness$num_species))
-  print(compare_richness_summary)
-
-  # save r-squared value
-  compare_richness_r2 <- compare_richness_summary$r.squared
-
-  plant_frug_richness_compare <- plot(plant_cellRichness$num_species, frugivore_cellRichness$num_species,
-       main = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
-       xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
-  abline(lm(frugivore_cellRichness$num_species ~ plant_cellRichness$num_species))
-
-  print(plant_frug_richness_compare)
-
-  plant_richness_hist <- hist(plant_cellRichness$num_species,
+  # Create histograms
+  plant_richness_hist <- hist(plant_richness_grid$num_species, 
     main = paste("Histogram of Plant Richness [", resolution_meters/1000, " km] by cell", sep = ""),
                               xlab = "Plant Richness by cell")
-  print(plant_richness_hist)
-  frugivore_richness_hist <- hist(frugivore_cellRichness$num_species,
+  frugivore_richness_hist <- hist(frugivore_richness_grid$num_species,
     main = paste("Histogram of Frugivore Richness [", resolution_meters/1000, " km] by cell", sep = ""),
                                   xlab = "Frugivore Richness by cell")
-  print(frugivore_richness_hist)
+
+  # Return a list of plots
+  list(plant_plot = plant_plot,
+       frugivoregridRichTA = frugivore_plot,
+       plant_cellRichness = plant_richness_grid,
+       frugivore_cellRichness = frugivore_richness_grid, 
+       compare_richness_summary = compare_richness_summary,
+       plant_frug_richness_compare = plant_frug_richness_compare,
+       plant_richness_hist = plant_richness_hist,
+       frugivore_richness_hist = frugivore_richness_hist)
 }
@@
 # save plots
-plantgridRichTA_1km <- plantgridRichTA
+plantgridRichTA_1km <- plant_plot
 frugivoregridRichTA_1km <- frugivoregridRichTA
@@
 
-compare_richness_r2_1km <- compare_richness_r2
+compare_richness_r2_1km <- compare_richness_summary$r.squared
 ```

### 2024-05-10 code/Manuscript_figures.rmd

Summary plots
# Load required packages
```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tidyr)
# devtools::install_github("pedropark99/ggfunnel")
library(ggfunnel)
library(ggpubr)
```

```{r}
Step <- c("Raw", "After cleaning", "Full trait coverage")
frug_records <- c(8089513, 1721825, 1667892)
frug_species <- c(1222, 1219, 1074)
frug_genera <- c(400, 398, 370)
frug_families <- c(66, 66, 64)

frug_records_df <- data.frame(Step, frug_records)
frug_species_df <- data.frame(Step, frug_species)
frug_genera_df <- data.frame(Step, frug_genera)
frug_families_df <- data.frame(Step, frug_families)
```

```{r}
ggplot(frug_records_df, aes(x = Step, y = frug_records)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_text(aes(label = frug_records), color = "white") + 
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(),axis.ticks.x = element_blank(), axis.title.y = element_blank())
```
```{r}
frug_records_fig <- ggfunnel::funnel(frug_records_df, levels = Step, values = frug_records, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Frugivore records", y = NULL)
frug_records_fig
```

```{r}
frug_species_fig <- ggfunnel::funnel(frug_species_df, levels = Step, values = frug_species, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Frugivore species", y = NULL)
frug_species_fig
```

```{r}
frug_genera_fig <- ggfunnel::funnel(frug_genera_df, levels = Step, values = frug_genera, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Frugivore genera", y = NULL)
frug_genera_fig
```

```{r}
frug_families_fig <- ggfunnel::funnel(frug_families_df, levels = Step, values = frug_families, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Frugivore families", y = NULL)
frug_families_fig
```

```{r}
figure <- ggarrange(frug_records_fig,frug_species_fig, frug_genera_fig, frug_families_fig,
                    labels = c("A", "B", "C", "D"),
                    ncol = 2, nrow = 2)
figure
```

```{r}
devtools::session_info()
```

### 2024-05-10 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
Fix column issues with gbifID and remove geometry column from occurrence output

@@
 output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+source("~/GitHub/neotropical_plants/functions/data_summary.R")
 ```
@@
 occ_download(
-pred("taxonKey", 6),
+pred_in("taxonKey", c(220, 196)),
 pred("hasCoordinate", TRUE),
@@
 ```{r}
-d1 <- occ_download_get("0009835-231002084531237", path = output_path, overwrite = TRUE) %>%
+d1 <- occ_download_get("0061290-231120084113126", path = output_path, overwrite = TRUE) %>%
   occ_download_import()
@@
 
+# Fix column types 
+```{r}
+TropicalAndes_plant_occ_forest$gbifID <- as.character(TropicalAndes_plant_occ_forest$gbifID)
+TropicalAndes_plant_occ_forest$geometry <- NULL
+```
+
 # Summary 
@@
 ```{r}
-print("The number of occurrences is")
-nrow(TropicalAndes_plant_occ_forest)
-print("The number of species is" )
-length(unique(TropicalAndes_plant_occ_forest$species))
-print("The number of genera is" )
-length(unique(TropicalAndes_plant_occ_forest$genus))
-print("The number of families is" )
-length(unique(TropicalAndes_plant_occ_forest$family))
+data_summary(TropicalAndes_plant_occ_forest, TropicalAndes_plant_occ_forest$species, TropicalAndes_plant_occ_forest$genus, TropicalAndes_plant_occ_forest$family)
 ```

### 2024-05-14 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
add figure path, missing packages, and package citation with session info

@@
 output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
 ```
@@
 library(terra)
+library(ggplot2)
+library(ggspatial)
 ```
@@
 labs(title = "Tropical Andes IUCN Montane and Lowland forest", x = "Latitude", y = "Longitude") +
-coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
+coord_sf(xlim = c(-85, -60), ylim = c(-24, 14), expand = FALSE) +
 annotation_scale(location = "bl", width_hint = 0.5) +
@@
 ```{r}
-ggsave(filename = "Tropical Andes Forest Map.png", dpi = 300, device="png")
+ggsave(filename = "Tropical Andes Forest Map.png", dpi = 300, device="png", path = output_path)
 ```
+
+# Cite packages and print session info
+```{r}
+library(report)
+cite_packages()
+
+devtools::session_info()
+```

### 2024-05-14 code/L1/L1_LastStep_DiversityInputObjects.Rmd
inital commit

---
title: ""
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Projecting occurrence data and creating presence-absence matrix for plants and frugivores. "
data input: 
data output: 
date: "2024-05-13"
output: html_document
---

# Load required packages
```{r}
library(letsR)
library(mFD)
library(vegan)
library(rnaturalearth)
library(sf)
library(raster)
library(fasterize)
library(dplyr)
library(funbiogeo)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path_L1 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Read in Data
```{r}
TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned_subset.csv"))
TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_frugivore_occ_cleaned_subset.csv"))
TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
frugivore_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_Frugivoria_traits_Forest.csv"))
plant_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_imputed_plant_traits.csv"))
```

# Convert data to spatial data
```{r}
plants.sf <- st_as_sf(TropicalAndes_plant_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)

frugivores.sf <- st_as_sf(TropicalAndes_frugivore_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
```

# Polygons of countries
```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
Americas <- ne_countries(continent = c("North America", "South America"), returnclass = "sf")
#polygon of Tropical Andes
TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
```

# Transform to projected coordinate reference system (units from degrees to meters)
```{r}
Americas <- st_transform(Americas, 5389)
TApoly <- st_transform(TApoly, 5389)
TropicalAndes_IUCNHabitat_Forest <- st_transform(TropicalAndes_IUCNHabitat_Forest, 5389)
plants.sf <- st_transform(plants.sf, 5389)
frugivores.sf <- st_transform(frugivores.sf, 5389)
```

```{r}
#check units
st_crs(TApoly, parameters = TRUE)$units_gdal
st_crs(TApoly)
st_crs(TropicalAndes_IUCNHabitat_Forest)
```

# Group by species
```{r}
plants_sf_species <- plants.sf %>%
  group_by(species) %>%
  summarise()

frugivores_sf_species <- frugivores.sf %>%
  group_by(species) %>%
  summarise()
```

# Plot Base Map
```{r}
basePlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "forestgreen")+
  labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

basePlot
```

# Plot points
```{r}
plantsPointsPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = plants_sf_species, pch = 21) +
  labs(title = "Tropical Andes Plant Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

plantsPointsPlot
```

```{r}
frugivoresPointsPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = frugivores_sf_species, pch = 21) +
  labs(title = "Tropical Andes Frugivore Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

frugivoresPointsPlot
```

# Create Presence-Absense matrix

Use project data sf object & convert to dataframe
```{r}
plants_occ_projected <- plants.sf %>%
  mutate(lon = sf::st_coordinates(.)[,1],
         lat = sf::st_coordinates(.)[,2]) %>%
  st_drop_geometry()
frugivores_occ_projected <- frugivores.sf %>%
  mutate(lon = sf::st_coordinates(.)[,1],
         lat = sf::st_coordinates(.)[,2]) %>%
  st_drop_geometry()
```

ADD CRS argument to lets.presab.points line (in PROJ4 type description)
```{r}
x <- plants_occ_projected$lon
y <- plants_occ_projected$lat
plant_xy <- cbind(x, y)
plant_species <- plants_occ_projected$species
plants_PAM <- letsR::lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, resol = 10000, crs = "+proj=utm +zone=19 +south +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs +type=crs", count = TRUE)
summary(plants_PAM)
plot(plants_PAM, xlab = "Longitude", ylab = "Latitude", main = "Plant richness map")
```

```{r}
x <- frugivores_occ_projected$lon
y <- frugivores_occ_projected$lat
frugivore_xy <- cbind(x, y)
frugivore_species <- frugivores_occ_projected$species
frugivore_PAM <- letsR::lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, crs = "+proj=utm +zone=19 +south +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs +type=crs", count = TRUE)
summary(frugivore_PAM)
plot(frugivore_PAM, xlab = "Longitude", ylab = "Latitude", main = "Frugivore richness map")
```

Check CRS of occurrence and IUCN layer
```{r}
st_crs(TropicalAndes_frugivore_occ_forest)
st_crs(TropicalAndes_plant_occ_forest)
st_crs(TropicalAndes_IUCNHabitat_Forest)
```

## Perform the spatial intersection
```{r}
intersections_plant_TA <- st_intersection(plants.sf, TropicalAndes_IUCNHabitat_Forest)
intersections_frugivore_TA <- st_intersection(plants.sf, TropicalAndes_IUCNHabitat_Forest)

# Convert the intersected sf object back to a dataframe
coordinates_df_subset_plant <- as.data.frame(intersections_plant_TA)
coordinates_df_subset_frugivore <- as.data.frame(intersections_frugivore_TA)
```

```{r}
#Check CRS
st_crs(intersections_plant_TA)
st_crs(coordinates_df_subset_plant)
```

```{r}
# Print the resulting subset of coordinates
print(coordinates_df_subset_plant)
print(coordinates_df_subset_frugivore)

# Convert the point geometry to a dataframe of longitude and latitude
subset_plant <- as.data.frame(st_coordinates(intersections_plant_TA))
subset_frugivore <- as.data.frame(st_coordinates(intersections_frugivore_TA))
```

## Match the coordinates to the original dataframe
```{r}
#merge the subset and full dataframe together to get final TA dataset
plant_PAM_filter <- merge(plants_PAM_matrix, subset_plant, by.x = c("Longitude(x)", "Latitude(y)"), by.y = c("X", "Y"))
frugivore_PAM_filter <- merge(frugivore_PAM_matrix, subset_frugivore, by.x = c("Longitude(x)", "Latitude(y)"), by.y = c("X", "Y"))
```

## Turn PAM into matrix
```{r}
plant_PAM <- as.matrix(plant_PAM_filter)
frugivore_PAM <- as.matrix(frugivore_PAM_filter)
```

## Add rownames
```{r}
# Get the number of rows in the matrix
num_rows_plant <- nrow(plant_PAM)

# Generate unique row names
row_names_plant <- paste0("cell", 1:num_rows_plant)

# Assign row names to the matrix
rownames(plant_PAM) <- row_names_plant

column_names_plant <- colnames(plant_PAM)
clean_column_names_plant <- gsub("_", " ", column_names_plant)

#Insert clean column names
colnames(plant_PAM) <- clean_column_names_plant
```

```{r}
# Get the number of rows in the matrix
num_rows_frugivore <- nrow(frugivore_PAM)

# Generate unique row names
row_names_frugivore <- paste0("cell", 1:num_rows_frugivore)

# Assign row names to the matrix
rownames(frugivore_PAM) <- row_names_frugivore

column_names_frugivore <- colnames(frugivore_PAM)
clean_column_names_frugivore <- gsub("_", " ", column_names_frugivore)

#Insert clean column names
colnames(frugivore_PAM) <- clean_column_names_frugivore
```

## Turn data into correct data types for inputs into the trait categories dataframe
```{r}
plant_traits$Growth.Form <- as.factor(plant_traits$Growth.Form)
plant_traits$Plant.Height_m <- as.numeric(plant_traits$Plant.Height_m)
plant_traits$Fruit.Length_mm <- as.numeric(plant_traits$Fruit.Length_mm)
plant_traits$Fruit.Type <- as.factor(plant_traits$Fruit.Type)
```

```{r}
frugivore_traits$diet_cat <- as.factor(frugivore_traits$diet_cat)
frugivore_traits$body_mass_e <- as.numeric(frugivore_traits$body_mass_e)
frugivore_traits$body_size_mm <- as.numeric(frugivore_traits$body_size_mm)
frugivore_traits$generation_time <- as.numeric(frugivore_traits$generation_time)
```

## Remove the species from PAM that have no occurrences
```{r}
# Remove columns with sum equal to zero
PAM_plant_site_final <- plant_PAM[, colSums(plant_PAM) != 0]
PAM_frugivore_site_final <- frugivore_PAM[, colSums(frugivore_PAM) != 0]

#Save coordinates for later
site_loc_key_plant <- PAM_plant_site_final[,1:2]
site_loc_key_frugivore <- PAM_frugivore_site_final[,1:2]

columns_to_remove <- c(1,2)
PAM_plant_site_final <- PAM_plant_site_final[,-columns_to_remove]
PAM_frugivore_site_final <- PAM_frugivore_site_final[,-columns_to_remove]

colnames_plant <- colnames(PAM_plant_site_final)
colnames_frugivore <- colnames(PAM_frugivore_site_final)
```

## Remove species names from trait matrix not in the PAM
```{r}
plant_traits_df_subset <- plant_traits %>% filter(species %in% colnames_plant)
frugivore_traits_df_subset <- frugivore_traits %>% filter(IUCN_species_name %in% colnames_frugivore)

# Turn trait dataframe into a matrix
class(plant_traits)
class(frugivore_traits)

plant_traits_matrix <- as.matrix(plant_traits_df_subset)
frugivore_traits_matrix <- as.matrix(frugivore_traits_df_subset)

# Define row names as species names
row_names_plant <- plant_traits_matrix[,1]
row_names_frugivore <- frugivore_traits_matrix[,1]

# Assign row names to the matrix
rownames(plant_traits_matrix) <- row_names_plant
rownames(frugivore_traits_matrix) <- row_names_frugivore

# Turn back into dataframe
plant_traits_df_final <-as.data.frame(plant_traits_matrix)
frugivore_traits_df_final <-as.data.frame(frugivore_traits_matrix)

plant_traits_df_final$X <-NULL
frugivore_traits_df_final$X <-NULL
```

```{r}
#fix types
plant_traits_df_final$Growth.Form <- as.factor(plant_traits_df_final$Growth.Form)
plant_traits_df_final$Plant.Height_m <- as.numeric(plant_traits_df_final$Plant.Height_m)
plant_traits_df_final$Fruit.Length_mm <- as.numeric(plant_traits_df_final$Fruit.Length_mm)
plant_traits_df_final$Fruit.Type <- as.factor(plant_traits_df_final$Fruit.Type)
```

```{r}
frugivore_traits_df_final$diet_cat <- as.factor(frugivore_traits_df_final$diet_cat)
frugivore_traits_df_final$body_mass_e <- as.numeric(frugivore_traits_df_final$body_mass_e)
frugivore_traits_df_final$body_size_mm<- as.numeric(frugivore_traits_df_final$body_size_mm)
frugivore_traits_df_final$generation_time <- as.numeric(frugivore_traits_df_final$generation_time)
```

```{r}
# remove duplicate species name column
plant_traits_df_final <- plant_traits_df_final[, c("Growth.Form", "Plant.Height_m", "Fruit.Length_mm", "Fruit.Type")]
frugivore_traits_df_final <- frugivore_traits_df_final[, c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
```

# Export data
```{r}
# objects to save
## sf objects
save(plants_sf_species, file = file.path(data_path_L1,"plants_sf_species.rdata"))
save(frugivores_sf_species, file = file.path(data_path_L1,"frugivores_sf_species.rdata"))
save(Americas, file = file.path(data_path_L1, "Americas.rdata"))
save(TApoly, file = file.path(data_path_L1,"TApoly.rdata"))
save(TropicalAndes_IUCNHabitat_Forest, file = file.path(data_path_L1,"TropicalAndes_IUCNHabitat_Forest.rdata"))
```

```{r}
# PAM objects
lets.save(plants_PAM, file = file.path(data_path_L1,"plants_PAM.RData"))
lets.save(frugivore_PAM, file = file.path(data_path_L1,"frugivore_PAM.RData"))
```

```{r}
# objects for functional diversity
PAM_plant_site_final
PAM_frugivore_site_final
plant_traits_df_final
frugivore_traits_df_final
```

### 2024-05-14 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
move code to L1_LastStep_Diversity_Input_Objects script

@@
 library(raster)
-library(scico)
 library(ggspatial)
-library(letsR)
-library(mFD)
+library(vegan)
 ```
@@
 frugivore_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_Frugivoria_traits_Forest.csv"))
-plant_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_plant_traits_forest.csv"))
+plant_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_imputed_plant_traits.csv"))
 ```
 
-# Create Presence-Absense matrix
+# Calculating Rao's Q
 ```{r}
-x <- TropicalAndes_plant_occ_forest$decimalLongitude
-y <- TropicalAndes_plant_occ_forest$decimalLatitude
-plant_xy <- cbind(x, y)
-plant_species <- TropicalAndes_plant_occ_forest$species
-plants_PAM <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14)
-summary(plants_PAM)
-plot(plants_PAM, xlab = "Longitude", ylab = "Latitude", main = "Plant richness map")
-plants_PAM_matrix <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE)
-```
+# Calculate Rao's Q using occurrence data and trait dissimilarity
 
-```{r}
-x <- TropicalAndes_frugivore_occ_forest$decimalLongitude
-y <- TropicalAndes_frugivore_occ_forest$decimalLatitude
-frugivore_xy <- cbind(x, y)
-frugivore_species <- TropicalAndes_frugivore_occ_forest$species
-frugivore_PAM <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14)
-summary(frugivore_PAM)
-plot(frugivore_PAM, xlab = "Longitude", ylab = "Latitude", main = "Frugivore richness map")
-frugivore_PAM_matrix <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE)
-```
+## Define different spatial scales (e.g., grid cell sizes)
+spatial_scales <- c(1000, 5000, 10000, 25000, 50000, 75000, 100000)  # In meters
 
-# Prepping data for functional diversity calculations
-```{r}
-plant_traits <- na.omit(plant_traits)
-plant_traits <- plant_traits[, c("species", "plant_height", "dispersal_syndrome", "plant_lifespan")]
-```
+## Create an empty dataframe to store results
+plant_raos_q_results <- data.frame(spatial_scale = numeric(),
+                             raos_q = numeric())
 
-```{r}
-frugivore_traits <- na.omit(frugivore_traits)
-frugivore_traits$body_mass_e <- as.numeric(frugivore_traits$body_mass_e)
-frugivore_traits$body_size_mm <- as.numeric(frugivore_traits$body_size_mm)
-frugivore_traits <- frugivore_traits[, c("IUCN_species_name", "diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
-```
+## Iterate over different spatial scales
+for (scale in spatial_scales) {
+  ### Aggregate occurrence data to the current spatial scale
+  plant_occurrence_aggregated <- aggregate_occurrence(plant_abundance_per_species, scale)
 
-```{r}
-coordinates_sf_plant <- st_as_sf(TropicalAndes_plant_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(TropicalAndes_IUCNHabitat_Forest))
-```
+  ### Calculate dissimilarities using trait data
+  plant_trait_dissimilarities <- vegdist(plant_traits, method = "bray")
 
-```{r}
-coordinates_sf_frugivore <- st_as_sf(TropicalAndes_frugivore_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(TropicalAndes_IUCNHabitat_Forest))
-```
+  ### Calculate dissimilarities using occurrence data
+  plant_occurrence_dissimilarities <- vegdist(plant_occurrence_aggregated, method = "bray")
 
-## Perform the spatial intersection
-```{r}
+  ### Calculate Rao's Q index
+  raos_q <- raosq(plant_trait_dissimilarities, plant_occurrence_dissimilarities)
 
-intersections_plant_TA <- st_intersection(coordinates_sf_plant, TropicalAndes_IUCNHabitat_Forest)
-intersections_frugivore_TA <- st_intersection(coordinates_sf_frugivore, TropicalAndes_IUCNHabitat_Forest)
+  #### Add results to the dataframe
+  plant_raos_q_results <- rbind(raos_q_results, data.frame(spatial_scale = scale, raos_q = raos_q))
+}
 
-# Convert the intersected sf object back to a dataframe
-coordinates_df_subset_plant <- as.data.frame(intersections_plant_TA)
-coordinates_df_subset_frugivore <- as.data.frame(intersections_frugivore_TA)
-```
+## Create an empty dataframe to store results
+frugivore_raos_q_results <- data.frame(spatial_scale = numeric(),
+                             raos_q = numeric())
 
-```{r}
-# Print the resulting subset of coordinates
-print(coordinates_df_subset_plant)
-print(coordinates_df_subset_frugivore)
+## Iterate over different spatial scales
+for (scale in spatial_scales) {
+  ### Aggregate occurrence data to the current spatial scale
+  frugivore_occurrence_aggregated <- aggregate_occurrence(frugivore_abundance_per_species, scale)
 
-# Convert the point geometry to a dataframe of longitude and latitude
-subset_plant <- as.data.frame(st_coordinates(intersections_plant_TA))
-subset_frugivore <- as.data.frame(st_coordinates(intersections_frugivore_TA))
-```
+  ### Calculate dissimilarities using trait data
+  frugivore_trait_dissimilarities <- vegdist(frugivore_traits, method = "bray")
 
-## Match the coordinates to the original dataframe
-```{r}
-#merge the subset and full dataframe together to get final TA dataset
-plant_PAM_filter <- merge(plants_PAM_matrix, subset_plant, by.x = c("Longitude(x)", "Latitude(y)"), by.y = c("X", "Y"))
-frugivore_PAM_filter <- merge(frugivore_PAM_matrix, subset_frugivore, by.x = c("Longitude(x)", "Latitude(y)"), by.y = c("X", "Y"))
-```
+  ### Calculate dissimilarities using occurrence data
+  frugivore_occurrence_dissimilarities <- vegdist(frugivore_occurrence_aggregated, method = "bray")
 
-## Turn PAM into matrix
-```{r}
-plant_PAM <- as.matrix(plant_PAM_filter)
-frugivore_PAM <- as.matrix(frugivore_PAM_filter)
-```
+  ### Calculate Rao's Q index
+  raos_q <- raosq(frugivore_trait_dissimilarities, frugivore_occurrence_dissimilarities)
 
-## Add rownames
-```{r}
-# Get the number of rows in the matrix
-num_rows_plant <- nrow(plant_PAM)
-
-# Generate unique row names
-row_names_plant <- paste0("cell", 1:num_rows_plant)
-
-# Assign row names to the matrix
-rownames(plant_PAM) <- row_names_plant
-
-# Print the matrix with row names
-print(plant_PAM)
-column_names_plant <- colnames(plant_PAM)
-clean_column_names_plant <- gsub("_", " ", column_names_plant)
-
-#Insert clean column names
-colnames(plant_PAM) <- clean_column_names_plant
-```
-
-```{r}
-# Get the number of rows in the matrix
-num_rows_frugivore <- nrow(frugivore_PAM)
-
-# Generate unique row names
-row_names_frugivore <- paste0("cell", 1:num_rows_frugivore)
-
-# Assign row names to the matrix
-rownames(frugivore_PAM) <- row_names_frugivore
-
-# Print the matrix with row names
-print(frugivore_PAM)
-column_names_frugivore <- colnames(frugivore_PAM)
-clean_column_names_frugivore <- gsub("_", " ", column_names_frugivore)
-
-#Insert clean column names
-colnames(frugivore_PAM) <- clean_column_names_frugivore
-```
-
-## Turn data into correct data types for inputs into the trait categories dataframe
-```{r}
-plant_traits$plant_height <- as.numeric(plant_traits$plant_height)
-plant_traits$plant_lifespan<- as.numeric(plant_traits$plant_lifespan)
-plant_traits$dispersal_syndrome <- as.factor(plant_traits$dispersal_syndrome)
-```
-
-```{r}
-frugivore_traits$diet_cat <- as.factor(frugivore_traits$diet_cat)
-frugivore_traits$body_mass_e <- as.numeric(frugivore_traits$body_mass_e)
-frugivore_traits$body_size_mm <- as.numeric(frugivore_traits$body_size_mm)
-frugivore_traits$generation_time <- as.numeric(frugivore_traits$generation_time)
-```
-
-## Remove the species from PAM that have no occurrences
-```{r}
-# Remove columns with sum equal to zero
-PAM_plant_site_final <- plant_PAM[, colSums(plant_PAM) != 0]
-PAM_frugivore_site_final <- frugivore_PAM[, colSums(frugivore_PAM) != 0]
-
-#Save coordinates for later
-site_loc_key_plant <- PAM_plant_site_final[,1:2]
-site_loc_key_frugivore <- PAM_frugivore_site_final[,1:2]
-
-columns_to_remove <- c(1,2)
-PAM_plant_site_final <- PAM_plant_site_final[,-columns_to_remove]
-PAM_frugivore_site_final <- PAM_frugivore_site_final[,-columns_to_remove]
-
-colnames_plant <- colnames(PAM_plant_site_final)
-colnames_frugivore <- colnames(PAM_frugivore_site_final)
-```
-
-## Remove species names from trait matrix not in the PAM
-```{r}
-plant_traits_df_subset <- plant_traits %>% filter(species %in% colnames_plant)
-frugivore_traits_df_subset <- frugivore_traits %>% filter(IUCN_species_name %in% colnames_frugivore)
-
-# Turn trait dataframe into a matrix
-class(plant_traits)
-class(frugivore_traits)
-
-plant_traits_matrix <- as.matrix(plant_traits_df_subset)
-frugivore_traits_matrix <- as.matrix(frugivore_traits_df_subset)
-
-# Define row names as species names
-row_names_plant <- plant_traits_matrix[,1]
-row_names_frugivore <- frugivore_traits_matrix[,1]
-
-# Assign row names to the matrix
-rownames(plant_traits_matrix) <- row_names_plant
-rownames(frugivore_traits_matrix) <- row_names_frugivore
-
-# Turn back into dataframe
-plant_traits_df_final <-as.data.frame(plant_traits_matrix)
-frugivore_traits_df_final <-as.data.frame(frugivore_traits_matrix)
-
-plant_traits_df_final$X <-NULL
-frugivore_traits_df_final$X <-NULL
-```
-
-```{r}
-#fix types
-plant_traits_df_final$plant_height <- as.numeric(plant_traits_df_final$plant_height)
-plant_traits_df_final$plant_lifespan<- as.numeric(plant_traits_df_final$plant_lifespan)
-plant_traits_df_final$dispersal_syndrome <- as.factor(plant_traits_df_final$dispersal_syndrome)
-```
-
-```{r}
-frugivore_traits_df_final$diet_cat <- as.factor(frugivore_traits_df_final$diet_cat)
-frugivore_traits_df_final$body_mass_e <- as.numeric(frugivore_traits_df_final$body_mass_e)
-frugivore_traits_df_final$body_size_mm<- as.numeric(frugivore_traits_df_final$body_size_mm)
-frugivore_traits_df_final$generation_time <- as.numeric(frugivore_traits_df_final$generation_time)
-```
-
-```{r}
-# remove duplicate species name column
-plant_traits_df_final <- plant_traits_df_final[, c("plant_height", "dispersal_syndrome", "plant_lifespan")]
-frugivore_traits_df_final <- frugivore_traits_df_final[, c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
+  #### Add results to the dataframe
+  frugivore_raos_q_results <- rbind(raos_q_results, data.frame(spatial_scale = scale, raos_q = raos_q))
+}
 ```
@@
 ```{r}
-trait_name <- c("dispersal_syndrome", "plant_height", "plant_lifespan")
-trait_type <- c("N", "Q", "Q") 
+trait_name <- c("Growth.Form", "Plant.Height_m", "Fruit.Length_mm", "Fruit.Type")
+trait_type <- c("N", "Q", "Q", "N") 
 plant_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
@@
   stop_if_NA = TRUE)
+```
+
+```{r}
+save(plant_traits_df_final, plant_trait_cat, file = "../../plant_funct.dist_files.rdata")
 ```

### 2024-05-14 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
move code to L1_LastStep_Diversity_Input_Objects script

@@
 library(raster)
-library(scico)
 library(ggspatial)
+library(gridExtra)
 ```
@@
 
-# Polygons of countries
-```{r}
-worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
-Americas <- ne_countries(continent = c("North America", "South America"), returnclass = "sf")
-#polygon of Tropical Andes
-TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
-```
-
-# Convert data to spatial data
-```{r}
-plants.sf <- st_as_sf(TropicalAndes_plant_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
-
-frugivores.sf <- st_as_sf(TropicalAndes_frugivore_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
-```
-
-# Transform to projected coordinate reference system (units from degrees to meters)
-```{r}
-Americas <- st_transform(Americas, 5389)
-TApoly <- st_transform(TApoly, 5389)
-TropicalAndes_IUCNHabitat_Forest <- st_transform(TropicalAndes_IUCNHabitat_Forest, 5389)
-plants.sf <- st_transform(plants.sf, 5389)
-frugivores.sf <- st_transform(frugivores.sf, 5389)
-```
-
-# Group by species
-```{r}
-plants_sf_species <- plants.sf %>%
-  group_by(species) %>%
-  summarise()
-
-frugivores_sf_species <- frugivores.sf %>%
-  group_by(species) %>%
-  summarise()
-```
-
-# Plot Base Map
-```{r}
-basePlot <-
-ggplot() +
-  geom_sf(data = Americas, fill = "white") +
-  geom_sf(data = TApoly) +
-  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "forestgreen")+
-  labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  theme(panel.background = element_rect(fill = "lightblue"))
-
-basePlot
-```
-
-# Plot points
-```{r}
-plantsPointsPlot <-
-ggplot() +
-  geom_sf(data = Americas, fill = "white") +
-  geom_sf(data = TApoly) +
-  geom_sf(data = plants_sf_species, pch = 21) +
-  labs(title = "Tropical Andes Plant Occurrences", x = "Latitude", y = "Longitude") +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  theme(panel.background = element_rect(fill = "lightblue"))
-
-plantsPointsPlot
-```
-
-```{r}
-frugivoresPointsPlot <-
-ggplot() +
-  geom_sf(data = Americas, fill = "white") +
-  geom_sf(data = TApoly) +
-  geom_sf(data = frugivores_sf_species, pch = 21) +
-  labs(title = "Tropical Andes Frugivore Occurrences", x = "Latitude", y = "Longitude") +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  theme(panel.background = element_rect(fill = "lightblue"))
-
-frugivoresPointsPlot
-```
-
-```{r}
-#check units
-st_crs(TApoly, parameters = TRUE)$units_gdal
-st_crs(TApoly)
-st_crs(TropicalAndes_IUCNHabitat_Forest)
-```
-
 # Plot TD/Richness
-function for richness plots
+## Function for richness plots
+Need to set legend scale and color for plants constant (same for frugivores)
+remove cells with richness of 0
 ```{r}
@@
       labs(title = title, x = "Longitude", y = "Latitude", fill = "Number of species") +
-      coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+      coord_sf(xlim = c(-85, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
       annotation_scale(location = "bl", width_hint = 0.5)  +
-      annotation_north_arrow(location = "bl", which_north = "true", 
-                             pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
       theme(panel.background = element_rect(fill = "lightblue"))
@@
   plant_plot <- generate_plot(plant_richness_grid, paste("Tropical Andes Plant Richness [", resolution_meters/1000, " km]", sep=""))
-  frugivore_plot <- generate_plot(frugivore_richness_grid, paste("Tropical Andes Frugivore Richness [", resolution_meters/1000, " km]", sep=""))
+  plant_plot <- generate_plot(frugivore_richness_grid, paste("Tropical Andes Frugivore Richness [", resolution_meters/1000, " km]", sep=""))
+
+  # Extract cell values
+  plant_cellRichness <- plant_plot[["data"]]
+  frugivore_cellRichness <- plant_plot[["data"]]
 
@@
     lm(frugivore_richness_grid$num_species ~ plant_richness_grid$num_species))
+  r_squared <- compare_richness_summary$r.squared
 
@@
 
-  # Create scatter plot
+  ## Create scatter plot
   plant_frug_richness_compare <- ggplot(comparison_data,
@@
   geom_point() +
-    labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]",
-                       sep = ""), x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
-    geom_smooth(method = "lm", se = FALSE, color = "red")
+  labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
+       subtitle =  paste("R-squared = ", round(r_squared, 2), sep = ""),
+       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
+  geom_smooth(method = "lm", se = FALSE, color = "red")  # Add regression line
 
@@
   # Return a list of plots
-  list(plant_plot = plant_plot,
+  list(plantgridRichTA = plant_plot,
        frugivoregridRichTA = frugivore_plot,
@@
 richness_1km <- create_rich_plots(1000)
+print(richness_1km)
+
 # save plots
-plantgridRichTA_1km <- plant_plot
-frugivoregridRichTA_1km <- frugivoregridRichTA
+plantgridRichTA_1km <- richness_1km$plant_plot
+frugivoregridRichTA_1km <- richness_1km$frugivoregridRichTA
 # save richness data
-plant_cellRichness_1km <- plant_cellRichness
-frugivore_cellRichness_1km <- frugivore_cellRichness
+plant_cellRichness_1km <- richness_1km$plant_cellRichness
+frugivore_cellRichness_1km <- richness_1km$frugivore_cellRichness
 
-compare_richness_r2_1km <- compare_richness_summary$r.squared
+compare_richness_r2_1km <- richness_1km$compare_richness_summary$r.squared
 ```
@@
 richness_5km <- create_rich_plots(5000)
+print(richness_5km)
+
 # save plots
-plantgridRichTA_5km <- plantgridRichTA
-frugivoregridRichTA_5km <- frugivoregridRichTA
+plantgridRichTA_5km <- richness_5km$plantgridRichTA
+frugivoregridRichTA_5km <- richness_5km$frugivoregridRichTA
 ## Extract cell values
-plant_cellRichness_5km <- plant_cellRichness
-frugivore_cellRichness_5km <- frugivore_cellRichness
+plant_cellRichness_5km <- richness_5km$plant_cellRichness
+frugivore_cellRichness_5km <- richness_5km$frugivore_cellRichness
 
-compare_richness_r2_5km <- compare_richness_r2
+compare_richness_r2_5km <- richness_5km$compare_richness_summary$r.squared
 ```
@@
 richness_10km <- create_rich_plots(10000)
+print(richness_10km)
+
 # save plots
-plantgridRichTA_10km <- plantgridRichTA
-frugivoregridRichTA_10km <- frugivoregridRichTA
+plantgridRichTA_10km <- richness_10km$plantgridRichTA
+frugivoregridRichTA_10km <- richness_10km$frugivoregridRichTA
 ## Extract cell values
-plant_cellRichness_10km <- plant_cellRichness
-frugivore_cellRichness_10km <- frugivore_cellRichness
+plant_cellRichness_10km <- richness_10km$plant_cellRichness
+frugivore_cellRichness_10km <- richness_10km$frugivore_cellRichness
 
-compare_richness_r2_10km <- compare_richness_r2
+compare_richness_r2_10km <- richness_10km$compare_richness_summary$r.squared
 ```
@@
 richness_25km <- create_rich_plots(25000)
+print(richness_25km)
+
 # save plots
-plantgridRichTA_25km <- plantgridRichTA
-frugivoregridRichTA_25km <- frugivoregridRichTA
+plantgridRichTA_25km <- richness_25km$plantgridRichTA
+frugivoregridRichTA_25km <- richness_25km$frugivoregridRichTA
 ## Extract cell values
-plant_cellRichness_25km <- plant_cellRichness
-frugivore_cellRichness_25km <- frugivore_cellRichness
+plant_cellRichness_25km <- richness_25km$plant_cellRichness
+frugivore_cellRichness_25km <- richness_25km$frugivore_cellRichness
 
-compare_richness_r2_25km <- compare_richness_r2
+compare_richness_r2_25km <- richness_25km$compare_richness_summary$r.squared
 ```
@@
 richness_50km <- create_rich_plots(50000)
+print(richness_50km)
 # save plots
-plantgridRichTA_50km <- plantgridRichTA
-frugivoregridRichTA_50km <- frugivoregridRichTA
+plantgridRichTA_50km <- richness_50km$plantgridRichTA
+frugivoregridRichTA_50km <- richness_50km$frugivoregridRichTA
 ## Extract cell values
-plant_cellRichness_50km <- plant_cellRichness
-frugivore_cellRichness_50km <- frugivore_cellRichness
+plant_cellRichness_50km <- richness_50km$plant_cellRichness
+frugivore_cellRichness_50km <- richness_50km$frugivore_cellRichness
 
-compare_richness_r2_50km <- compare_richness_r2
+compare_richness_r2_50km <- richness_50km$compare_richness_summary$r.squared
 ```
@@
 richness_75km <- create_rich_plots(75000)
+print(richness_75km)
 # save plots
-plantgridRichTA_75km <- plantgridRichTA
-frugivoregridRichTA_75km <- frugivoregridRichTA
+plantgridRichTA_75km <- richness_75km$plantgridRichTA
+frugivoregridRichTA_75km <- richness_75km$frugivoregridRichTA
 ## Extract cell values
-plant_cellRichness_75km <- plant_cellRichness
-frugivore_cellRichness_75km <- frugivore_cellRichness
+plant_cellRichness_75km <- richness_75km$plant_cellRichness
+frugivore_cellRichness_75km <- richness_75km$frugivore_cellRichness
 
-compare_richness_r2_75km <- compare_richness_r2
+compare_richness_r2_75km <- richness_75km$compare_richness_summary$r.squared
 ```
@@
 richness_100km <- create_rich_plots(100000)
+print(richness_100km)
 # save plots
-plantgridRichTA_100km <- plantgridRichTA
-frugivoregridRichTA_100km <- frugivoregridRichTA
+plantgridRichTA_100km <- richness_100km$plantgridRichTA
+frugivoregridRichTA_100km <- richness_100km$frugivoregridRichTA
 ## Extract cell values
-plant_cellRichness_100km <- plant_cellRichness
-frugivore_cellRichness_100km <- frugivore_cellRichness
+plant_cellRichness_100km <- richness_100km$plant_cellRichness
+frugivore_cellRichness_100km <- richness_100km$frugivore_cellRichness
 
-compare_richness_r2_100km <- compare_richness_r2
+compare_richness_r2_100km <- richness_100km$compare_richness_summary$r.squared
+```
+
+# Multi panel plots
+Richness maps
+```{r}
+all_richness_plots <- gridExtra::grid.arrange(#plantgridRichTA_1km, frugivoregridRichTA_1km,
+                                              #plantgridRichTA_5km, frugivoregridRichTA_5km, 
+                                               plantgridRichTA_10km, frugivoregridRichTA_10km,
+                                               plantgridRichTA_25km, frugivoregridRichTA_25km,
+                                               plantgridRichTA_50km, frugivoregridRichTA_50km,
+                                               plantgridRichTA_75km,  frugivoregridRichTA_75km,
+                                               plantgridRichTA_100km, frugivoregridRichTA_100km,
+                                               ncol = 5, nrow = 2)
+all_richness_plots
+ggsave("all_richness_plots.pdf", all_richness_plots, path = output_path, width = 20, height = 8.5, units = "in")
 ```
@@
 # create dataframe of resolutions and r-squared values 
-resolution <- c("5", "25", "50", "75", "100")
-r_squared <- c(compare_richness_r2_5km, compare_richness_r2_25km, compare_richness_r2_50km, compare_richness_r2_75km, compare_richness_r2_100km)
-r_squared_compare <- data.frame(resolution, r-squared)
-```
+#[ADD 1, 5,]
+resolution <- c( 10, 25, 50, 75, 100)
+#ADD compare_richness_r2_1km, compare_richness_r2_5km,
+r_squared <- c( compare_richness_r2_10km, compare_richness_r2_25km, compare_richness_r2_50km, compare_richness_r2_75km, compare_richness_r2_100km)
+r_squared_compare <- data.frame(resolution, r_squared)
 
-```{r}
-ggplot(r_squared_compare, aes(x=resolution, y=r_squared)) +
+compare_r2_by_res <- ggplot(r_squared_compare, aes(x=resolution, y=r_squared)) +
     geom_point()+
-  labs(title = "R-squared values of plant-frugivore richness relationships by resolution", x = "Resolution (km)", y = "R-squared",
-       fill = "Number of species") 
-```
-
-# Multi panel plots
-Richness maps
-```{r}
-grid.arrange(plot_1, plot_2, nrow = 1)
+    labs(title = "R-squared values of plant-frugivore richness relationships by resolution",
+         x = "Resolution (km)", y = "R-squared", fill = "Number of species")
+print(compare_r2_by_res)
 ```
@@
 ```{r}
-write.csv(plant_cellRichness0.75, file.path(output_path,"TropicalAndes_GBIF_plant_richness_sub.csv"))
-write.csv(frugivore_cellRichness0.75, file.path(output_path,"TropicalAndes_frugivore_richness_sub.csv"))
+write.csv(plant_cellRichness_1km, file.path(output_path,"TropicalAndes_plantRichness_1km.csv"))
+write.csv(frugivore_cellRichness_1km, file.path(output_path,"TropicalAndes_frugivoreRichness_1km.csv"))
 ```
+
+```{r}
+write.csv(plant_cellRichness_5km, file.path(output_path,"TropicalAndes_plantRichness_5km.csv"))
+write.csv(frugivore_cellRichness_5km, file.path(output_path,"TropicalAndes_frugivoreRichness_5km.csv"))
+```
+
+```{r}
+write.csv(plant_cellRichness_10km, file.path(output_path,"TropicalAndes_plantRichness_10km.csv"))
+write.csv(frugivore_cellRichness_10km, file.path(output_path,"TropicalAndes_frugivoreRichness_10km.csv"))
+```
+
+```{r}
+write.csv(plant_cellRichness_25km, file.path(output_path,"TropicalAndes_plantRichness_25km.csv"))
+write.csv(frugivore_cellRichness_25km, file.path(output_path,"TropicalAndes_frugivoreRichness_25km.csv"))
+```
+
+```{r}
+write.csv(plant_cellRichness_50km, file.path(output_path,"TropicalAndes_plantRichness_50km.csv"))
+write.csv(frugivore_cellRichness_50km, file.path(output_path,"TropicalAndes_frugivoreRichness_50km.csv"))
+```
+
+```{r}
+write.csv(plant_cellRichness_75km, file.path(output_path,"TropicalAndes_plantRichness_75km.csv"))
+write.csv(frugivore_cellRichness_75km, file.path(output_path,"TropicalAndes_frugivoreRichness_75km.csv"))
+```
+
+```{r}
+write.csv(plant_cellRichness_100km, file.path(output_path,"TropicalAndes_plantRichness_100km.csv"))
+write.csv(frugivore_cellRichness_100km, file.path(output_path,"TropicalAndes_frugivoreRichness_100km.csv"))
+```

### 2024-05-14 code/L1/L1_LastStep_DiversityInputObjects.Rmd
remove projection code for PAM and rename following variables using un-projected objects, export files as RDS files

@@
 
-Use project data sf object & convert to dataframe
+Use unprojected data because projected data is causing issues
 ```{r}
-plants_occ_projected <- plants.sf %>%
-  mutate(lon = sf::st_coordinates(.)[,1],
-         lat = sf::st_coordinates(.)[,2]) %>%
-  st_drop_geometry()
-frugivores_occ_projected <- frugivores.sf %>%
-  mutate(lon = sf::st_coordinates(.)[,1],
-         lat = sf::st_coordinates(.)[,2]) %>%
-  st_drop_geometry()
-```
-
-ADD CRS argument to lets.presab.points line (in PROJ4 type description)
-```{r}
-x <- plants_occ_projected$lon
-y <- plants_occ_projected$lat
+x <- TropicalAndes_plant_occ_forest$decimalLongitude
+y <- TropicalAndes_plant_occ_forest$decimalLatitude
 plant_xy <- cbind(x, y)
-plant_species <- plants_occ_projected$species
-plants_PAM <- letsR::lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, resol = 10000, crs = "+proj=utm +zone=19 +south +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs +type=crs", count = TRUE)
+plant_species <- TropicalAndes_plant_occ_forest$species
+plants_PAM <- letsR::lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, count = TRUE)
 summary(plants_PAM)
@@
 ```{r}
-x <- frugivores_occ_projected$lon
-y <- frugivores_occ_projected$lat
+x <- TropicalAndes_frugivore_occ_forest$decimalLongitude
+y <- TropicalAndes_frugivore_occ_forest$decimalLatitude
 frugivore_xy <- cbind(x, y)
-frugivore_species <- frugivores_occ_projected$species
-frugivore_PAM <- letsR::lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, crs = "+proj=utm +zone=19 +south +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs +type=crs", count = TRUE)
+frugivore_species <- TropicalAndes_frugivore_occ_forest$species
+frugivore_PAM <- letsR::lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, count = TRUE)
 summary(frugivore_PAM)
@@
 
-Check CRS of occurrence and IUCN layer
-```{r}
-st_crs(TropicalAndes_frugivore_occ_forest)
-st_crs(TropicalAndes_plant_occ_forest)
-st_crs(TropicalAndes_IUCNHabitat_Forest)
-```
-
-## Perform the spatial intersection
-```{r}
-intersections_plant_TA <- st_intersection(plants.sf, TropicalAndes_IUCNHabitat_Forest)
-intersections_frugivore_TA <- st_intersection(plants.sf, TropicalAndes_IUCNHabitat_Forest)
-
-# Convert the intersected sf object back to a dataframe
-coordinates_df_subset_plant <- as.data.frame(intersections_plant_TA)
-coordinates_df_subset_frugivore <- as.data.frame(intersections_frugivore_TA)
-```
-
-```{r}
-#Check CRS
-st_crs(intersections_plant_TA)
-st_crs(coordinates_df_subset_plant)
-```
-
-```{r}
-# Print the resulting subset of coordinates
-print(coordinates_df_subset_plant)
-print(coordinates_df_subset_frugivore)
-
-# Convert the point geometry to a dataframe of longitude and latitude
-subset_plant <- as.data.frame(st_coordinates(intersections_plant_TA))
-subset_frugivore <- as.data.frame(st_coordinates(intersections_frugivore_TA))
-```
-
-## Match the coordinates to the original dataframe
-```{r}
-#merge the subset and full dataframe together to get final TA dataset
-plant_PAM_filter <- merge(plants_PAM_matrix, subset_plant, by.x = c("Longitude(x)", "Latitude(y)"), by.y = c("X", "Y"))
-frugivore_PAM_filter <- merge(frugivore_PAM_matrix, subset_frugivore, by.x = c("Longitude(x)", "Latitude(y)"), by.y = c("X", "Y"))
-```
-
 ## Turn PAM into matrix
 ```{r}
-plant_PAM <- as.matrix(plant_PAM_filter)
-frugivore_PAM <- as.matrix(frugivore_PAM_filter)
+plants_PAM_matrix <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE, count = TRUE)
+frugivore_PAM_matrix <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE, count = TRUE)
 ```
@@
 # Get the number of rows in the matrix
-num_rows_plant <- nrow(plant_PAM)
+num_rows_plant <- nrow(plants_PAM_matrix)
 
@@
 # Assign row names to the matrix
-rownames(plant_PAM) <- row_names_plant
+rownames(plants_PAM_matrix) <- row_names_plant
 
-column_names_plant <- colnames(plant_PAM)
-clean_column_names_plant <- gsub("_", " ", column_names_plant)
-
-#Insert clean column names
-colnames(plant_PAM) <- clean_column_names_plant
+column_names_plant <- colnames(plants_PAM_matrix)
 ```
@@
 # Get the number of rows in the matrix
-num_rows_frugivore <- nrow(frugivore_PAM)
+num_rows_frugivore <- nrow(frugivore_PAM_matrix)
 
@@
 # Assign row names to the matrix
-rownames(frugivore_PAM) <- row_names_frugivore
+rownames(frugivore_PAM_matrix) <- row_names_frugivore
 
-column_names_frugivore <- colnames(frugivore_PAM)
-clean_column_names_frugivore <- gsub("_", " ", column_names_frugivore)
-
-#Insert clean column names
-colnames(frugivore_PAM) <- clean_column_names_frugivore
+column_names_frugivore <- colnames(frugivore_PAM_matrix)
 ```
@@
 # Remove columns with sum equal to zero
-PAM_plant_site_final <- plant_PAM[, colSums(plant_PAM) != 0]
-PAM_frugivore_site_final <- frugivore_PAM[, colSums(frugivore_PAM) != 0]
+PAM_plant_site_final <- plants_PAM_matrix[, colSums(plants_PAM_matrix) != 0]
+PAM_frugivore_site_final <- frugivore_PAM_matrix[, colSums(frugivore_PAM_matrix) != 0]
 
@@
 
-columns_to_remove <- c(1,2)
-PAM_plant_site_final <- PAM_plant_site_final[,-columns_to_remove]
-PAM_frugivore_site_final <- PAM_frugivore_site_final[,-columns_to_remove]
+PAM_plant_site_final <- PAM_plant_site_final[,-c(1:2)]
+PAM_frugivore_site_final <- PAM_frugivore_site_final[,-(1:2)]
 
@@
 
+```{r}
+str(PAM_plant_site_final)
+str(PAM_frugivore_site_final)
+```
+
 ## Remove species names from trait matrix not in the PAM
 ```{r}
-plant_traits_df_subset <- plant_traits %>% filter(species %in% colnames_plant)
-frugivore_traits_df_subset <- frugivore_traits %>% filter(IUCN_species_name %in% colnames_frugivore)
+plant_traits_df_subset <- plant_traits %>%
+                            filter(species %in% colnames_plant) %>%
+                            distinct(species, .keep_all = TRUE)
+frugivore_traits_df_subset <- frugivore_traits %>%
+                                filter(IUCN_species_name %in% colnames_frugivore) %>%
+                                distinct(IUCN_species_name, .keep_all = TRUE)
 
-# Turn trait dataframe into a matrix
-class(plant_traits)
-class(frugivore_traits)
+dim(plant_traits)
+dim(plant_traits_df_subset)
 
-plant_traits_matrix <- as.matrix(plant_traits_df_subset)
-frugivore_traits_matrix <- as.matrix(frugivore_traits_df_subset)
-
-# Define row names as species names
-row_names_plant <- plant_traits_matrix[,1]
-row_names_frugivore <- frugivore_traits_matrix[,1]
-
-# Assign row names to the matrix
-rownames(plant_traits_matrix) <- row_names_plant
-rownames(frugivore_traits_matrix) <- row_names_frugivore
-
-# Turn back into dataframe
-plant_traits_df_final <-as.data.frame(plant_traits_matrix)
-frugivore_traits_df_final <-as.data.frame(frugivore_traits_matrix)
-
-plant_traits_df_final$X <-NULL
-frugivore_traits_df_final$X <-NULL
+dim(frugivore_traits)
+dim(frugivore_traits_df_subset)
 ```
@@
 ```{r}
-#fix types
-plant_traits_df_final$Growth.Form <- as.factor(plant_traits_df_final$Growth.Form)
-plant_traits_df_final$Plant.Height_m <- as.numeric(plant_traits_df_final$Plant.Height_m)
-plant_traits_df_final$Fruit.Length_mm <- as.numeric(plant_traits_df_final$Fruit.Length_mm)
-plant_traits_df_final$Fruit.Type <- as.factor(plant_traits_df_final$Fruit.Type)
+# Define row names as species names
+row_names_plant <- plant_traits_df_subset$species
+row_names_frugivore <- frugivore_traits_df_subset$IUCN_species_name
+
+# Assign row names to the matrix
+rownames(plant_traits_df_subset) <- row_names_plant
+rownames(frugivore_traits_df_subset) <- row_names_frugivore
+
+plant_traits_df_subset$X <-NULL
+frugivore_traits_df_subset$X <-NULL
+# remove duplicate species name column
+plant_traits_df_subset <- plant_traits_df_subset[, c("Growth.Form", "Plant.Height_m", "Fruit.Length_mm", "Fruit.Type")]
+frugivore_traits_df_subset <- frugivore_traits_df_subset[, c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
 ```
@@
 ```{r}
-frugivore_traits_df_final$diet_cat <- as.factor(frugivore_traits_df_final$diet_cat)
-frugivore_traits_df_final$body_mass_e <- as.numeric(frugivore_traits_df_final$body_mass_e)
-frugivore_traits_df_final$body_size_mm<- as.numeric(frugivore_traits_df_final$body_size_mm)
-frugivore_traits_df_final$generation_time <- as.numeric(frugivore_traits_df_final$generation_time)
-```
-
-```{r}
-# remove duplicate species name column
-plant_traits_df_final <- plant_traits_df_final[, c("Growth.Form", "Plant.Height_m", "Fruit.Length_mm", "Fruit.Type")]
-frugivore_traits_df_final <- frugivore_traits_df_final[, c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
+str(plant_traits_df_subset)
+str(frugivore_traits_df_subset)
 ```
@@
 ## sf objects
-save(plants_sf_species, file = file.path(data_path_L1,"plants_sf_species.rdata"))
-save(frugivores_sf_species, file = file.path(data_path_L1,"frugivores_sf_species.rdata"))
-save(Americas, file = file.path(data_path_L1, "Americas.rdata"))
-save(TApoly, file = file.path(data_path_L1,"TApoly.rdata"))
-save(TropicalAndes_IUCNHabitat_Forest, file = file.path(data_path_L1,"TropicalAndes_IUCNHabitat_Forest.rdata"))
+saveRDS(plants_sf_species, file = file.path(data_path_L1,"plants_sf_species.rds"))
+saveRDS(frugivores_sf_species, file = file.path(data_path_L1,"frugivores_sf_species.rds"))
+saveRDS(Americas, file = file.path(data_path_L1, "Americas.rds"))
+saveRDS(TApoly, file = file.path(data_path_L1,"TApoly.rds"))
+saveRDS(TropicalAndes_IUCNHabitat_Forest, file = file.path(data_path_L1,"TropicalAndes_IUCNHabitat_Forest.rds"))
 ```
@@
 # PAM objects
-lets.save(plants_PAM, file = file.path(data_path_L1,"plants_PAM.RData"))
-lets.save(frugivore_PAM, file = file.path(data_path_L1,"frugivore_PAM.RData"))
+saveRDS(plants_PAM, file = file.path(data_path_L1,"plants_PAM.rds"))
+saveRDS(frugivore_PAM, file = file.path(data_path_L1,"frugivore_PAM.rds"))
 ```
@@
 # objects for functional diversity
-PAM_plant_site_final
-PAM_frugivore_site_final
-plant_traits_df_final
-frugivore_traits_df_final
+saveRDS(PAM_plant_site_final, file = file.path(data_path_L1,"PAM_plant_site_final.rds"))
+saveRDS(PAM_frugivore_site_final, file = file.path(data_path_L1,"PAM_frugivore_site_final.rds"))
+saveRDS(plant_traits_df_subset, file = file.path(data_path_L1,"plant_traits_df_final.rds"))
+saveRDS(frugivore_traits_df_subset, file = file.path(data_path_L1,"frugivore_traits_df_final.rds"))
 ```

### 2024-05-14 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
re-organize so that all frugivore code is together and all plant code is together

@@
 # Read in Data
+
 ```{r}
-TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned_subset.csv"))
-TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_frugivore_occ_cleaned_subset.csv"))
-TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
-frugivore_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_Frugivoria_traits_Forest.csv"))
-plant_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_imputed_plant_traits.csv"))
+## projected sf objects
+plants_sf_species <- readRDS(file = file.path(data_path_L1,"plants_sf_species.rds"))
+frugivores_sf_species <- readRDS(file = file.path(data_path_L1,"frugivores_sf_species.rds"))
+Americas <- readRDS(file = file.path(data_path_L1, "Americas.rds"))
+TApoly <- readRDS(file = file.path(data_path_L1,"TApoly.rds"))
+TropicalAndes_IUCNHabitat_Forest <- readRDS(file = file.path(data_path_L1,"TropicalAndes_IUCNHabitat_Forest.rds"))
 ```
 
-# Calculating Rao's Q
 ```{r}
-# Calculate Rao's Q using occurrence data and trait dissimilarity
-
-## Define different spatial scales (e.g., grid cell sizes)
-spatial_scales <- c(1000, 5000, 10000, 25000, 50000, 75000, 100000)  # In meters
-
-## Create an empty dataframe to store results
-plant_raos_q_results <- data.frame(spatial_scale = numeric(),
-                             raos_q = numeric())
-
-## Iterate over different spatial scales
-for (scale in spatial_scales) {
-  ### Aggregate occurrence data to the current spatial scale
-  plant_occurrence_aggregated <- aggregate_occurrence(plant_abundance_per_species, scale)
-
-  ### Calculate dissimilarities using trait data
-  plant_trait_dissimilarities <- vegdist(plant_traits, method = "bray")
-
-  ### Calculate dissimilarities using occurrence data
-  plant_occurrence_dissimilarities <- vegdist(plant_occurrence_aggregated, method = "bray")
-
-  ### Calculate Rao's Q index
-  raos_q <- raosq(plant_trait_dissimilarities, plant_occurrence_dissimilarities)
-
-  #### Add results to the dataframe
-  plant_raos_q_results <- rbind(raos_q_results, data.frame(spatial_scale = scale, raos_q = raos_q))
-}
-
-## Create an empty dataframe to store results
-frugivore_raos_q_results <- data.frame(spatial_scale = numeric(),
-                             raos_q = numeric())
-
-## Iterate over different spatial scales
-for (scale in spatial_scales) {
-  ### Aggregate occurrence data to the current spatial scale
-  frugivore_occurrence_aggregated <- aggregate_occurrence(frugivore_abundance_per_species, scale)
-
-  ### Calculate dissimilarities using trait data
-  frugivore_trait_dissimilarities <- vegdist(frugivore_traits, method = "bray")
-
-  ### Calculate dissimilarities using occurrence data
-  frugivore_occurrence_dissimilarities <- vegdist(frugivore_occurrence_aggregated, method = "bray")
-
-  ### Calculate Rao's Q index
-  raos_q <- raosq(frugivore_trait_dissimilarities, frugivore_occurrence_dissimilarities)
-
-  #### Add results to the dataframe
-  frugivore_raos_q_results <- rbind(raos_q_results, data.frame(spatial_scale = scale, raos_q = raos_q))
-}
+# objects for functional diversity
+PAM_plant_site_final <- readRDS(file = file.path(data_path_L1,"PAM_plant_site_final.rds"))
+PAM_frugivore_site_final <- readRDS(file = file.path(data_path_L1,"PAM_frugivore_site_final.rds"))
+plant_traits_df_final <- readRDS(file = file.path(data_path_L1,"plant_traits_df_final.rds"))
+frugivore_traits_df_final <- readRDS(file = file.path(data_path_L1,"frugivore_traits_df_final.rds"))
 ```
 
-# Create trait type table
-```{r}
-trait_name <- c("Growth.Form", "Plant.Height_m", "Fruit.Length_mm", "Fruit.Type")
-trait_type <- c("N", "Q", "Q", "N") 
-plant_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
-```
-
+# Frugivores
+## Create trait type table
 ```{r}
@@
 
-# Summary of the assemblages * species dataframe
+## Summary of the assemblages * species dataframe
 ```{r}
-asb_sp_plant_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_plant_site_final)
-
 asb_sp_frugivore_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_frugivore_site_final)
@@
 
-# Species traits summary
+## Species traits summary
 ```{r}
-plant_traits_summ <- mFD::sp.tr.summary(
-  tr_cat     = plant_trait_cat,   
-  sp_tr      = plant_traits_df_final, 
-  stop_if_NA = TRUE)
-
 frugivore_traits_summ <- mFD::sp.tr.summary(
@@
 
-```{r}
-save(plant_traits_df_final, plant_trait_cat, file = "../../plant_funct.dist_files.rdata")
-```
-
-# Estimate functional trait-based distances between species
-```{r}
-sp_dist_plants <- mFD::funct.dist(
-  sp_tr         = plant_traits_df_final,
-  tr_cat        = plant_trait_cat,
-  metric        = "gower",
-  scale_euclid  = "scale_center",
-  ordinal_var   = "classic",
-  weight_type   = "equal",
-  stop_if_NA    = TRUE)
-```
-
+## Estimate functional trait-based distances between species
 ```{r}
@@
 
-# Generate a multidimensional space
-```{r}
-fspaces_quality_plants <- mFD::quality.fspaces(
-  sp_dist             = sp_dist_plants,
-  maxdim_pcoa         = 10,
-  deviation_weighting = "absolute",
-  fdist_scaling       = FALSE,
-  fdendro             = "average")
-```
-
+## Generate a multidimensional space
 ```{r}
@@
 # Look at the quality spaces only (MAD index looks at mean absolute deviation from the dissimilary matrix; want the deviation to be low, meaning that the true distances have been retained in the PCA)
-round(fspaces_quality_plants$"quality_fspaces", 3)            
 round(fspaces_quality_frugivore$"quality_fspaces", 3)     
 # Plot the quality spaces (chose to look at 3D, 4D, and 5D since they had the lowest MAD). Will go with 4 dimensions.
-mFD::quality.fspaces.plot(
-  fspaces_quality            = fspaces_quality_plants,
-  quality_metric             = "mad",
-  fspaces_plot               = c("pcoa_2d", "pcoa_3d", 
-                                 "pcoa_4d"),
-  name_file                  = NULL,
-  range_dist                 = NULL,
-  range_dev                  = NULL,
-  range_qdev                 = NULL,
-  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
-  gradient_deviation_quality = c(low = "yellow", high = "red"),
-  x_lab                      = "Trait-based distance")
 mFD::quality.fspaces.plot(
@@
 #testing correlation between functional axes and traits
-sp_faxes_coord_plants <- fspaces_quality_plants$"details_fspaces"$"sp_pc_coord"
 sp_faxes_coord_frugivore <- fspaces_quality_frugivore$"details_fspaces"$"sp_pc_coord"
@@
 # Computes linear model for continuous traits and Kruskall-Wallis tests for other types. 
-plant_tr_faxes <- mFD::traits.faxes.cor(
-  sp_tr          = plant_traits_df_final, 
-  sp_faxes_coord = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")], 
-  plot           = TRUE)
-
 frugivore_tr_faxes <- mFD::traits.faxes.cor(
@@
 # Print traits with significant effect:
-plant_tr_faxes$"tr_faxes_stat"[which(plant_tr_faxes$"tr_faxes_stat"$"p.value" < 0.05), ]
 frugivore_tr_faxes$"tr_faxes_stat"[which(frugivore_tr_faxes$"tr_faxes_stat"$"p.value" < 0.05), ]
 # Return plots:
-plant_tr_faxes$"tr_faxes_plot"
 frugivore_tr_faxes$"tr_faxes_plot"
@@
 #plotting functional space
-sp_faxes_coord_plants <- fspaces_quality_plants$"details_fspaces"$"sp_pc_coord"
 sp_faxes_coord_frugivore <- fspaces_quality_frugivore$"details_fspaces"$"sp_pc_coord"
-big_plot_plants <- mFD::funct.space.plot(
-  sp_faxes_coord  = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")],
+
+big_plot_frugivore <- mFD::funct.space.plot(
+  sp_faxes_coord  = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")],
   faxes           = c("PC1", "PC2", "PC3", "PC4"),
@@
   check_input     = TRUE)
+```
 
-big_plot_frugivore <- mFD::funct.space.plot(
-  sp_faxes_coord  = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")],
+```{r}
+big_plot_frugivore
+```
+
+```{r}
+#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
+# Calculate row sums
+row_sums_frugivore <- rowSums(PAM_frugivore_site_final)
+subset_matrix_frugivore <- PAM_frugivore_site_final[row_sums_frugivore >= 4, ]
+```
+
+```{r}
+# match frugivore names
+sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")]
+summary(sp_faxes_coord_frugivore_sub)
+```
+
+```{r}
+## check number of species names
+nrow(sp_faxes_coord_frugivore_sub)
+ncol(subset_matrix_frugivore) 
+```
+
+```{r}
+sp_faxes_coord_frugivore_sub_names <- row.names(sp_faxes_coord_frugivore_sub)
+subset_matrix_frugivore_names <- colnames(subset_matrix_frugivore)
+```
+
+```{r}
+sp_faxes_coord_frugivore_sub <- as.data.frame(sp_faxes_coord_frugivore_sub)
+row.names(sp_faxes_coord_frugivore_sub) <- sp_faxes_coord_frugivore_sub_names
+
+sp_faxes_coord_frugivore_sub_names <- row.names(sp_faxes_coord_frugivore_sub)
+subset_matrix_frugivore_names <- colnames(subset_matrix_frugivore)
+
+frugivore_names <- intersect(sp_faxes_coord_frugivore_sub_names, subset_matrix_frugivore_names)
+frugivore_names <- na.omit(frugivore_names)
+
+sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore_sub[ which((row.names(sp_faxes_coord_frugivore_sub) %in% frugivore_names)==TRUE), ]
+
+subset_matrix_frugivore  <- as.data.frame(subset_matrix_frugivore)
+subset_matrix_frugivore <- subset_matrix_frugivore[ ,which((colnames(subset_matrix_frugivore) %in% frugivore_names)==TRUE)]
+
+#remove NAs
+sp_faxes_coord_frugivore_sub <- na.omit(sp_faxes_coord_frugivore_sub)
+subset_matrix_frugivore <- na.omit(subset_matrix_frugivore)
+```
+
+```{r}
+nrow(sp_faxes_coord_frugivore_sub)
+ncol(subset_matrix_frugivore)
+```
+
+```{r}
+sp_faxes_coord_frugivore_sub <- as.matrix(sp_faxes_coord_frugivore_sub)
+subset_matrix_frugivore <- as.matrix(subset_matrix_frugivore)
+```
+
+```{r}
+#computing FD
+# The number of species per assemblage has to be higher or equal to the number of traits
+
+alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(
+  sp_faxes_coord   = sp_faxes_coord_frugivore_sub,
+  asb_sp_w         = subset_matrix_frugivore,
+  ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
+  scaling          = TRUE,
+  check_input      = TRUE,
+  details_returned = TRUE)
+
+details_list_frugivore <- alpha_fd_indices_frugivore$"details"
+
+plots_alpha_frugivore <- mFD::alpha.multidim.plot(
+  output_alpha_fd_multidim = alpha_fd_indices_frugivore,
+  plot_asb_nm              = c("cell4", "cell20"),
+  ind_nm                   = c("fdis", "fide", "fnnd"
+                        , "fori", "fspe"),
+  faxes                    = NULL,
+  faxes_nm                 = NULL,
+  range_faxes              = c(NA, NA),
+  color_bg                 = "grey95",
+  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
+  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
+  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  color_ch          
+  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
+  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
+  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
+  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
+  shape_centroid_fspe      = 23,
+  color_centroid_fspe      = "black",
+  size_sp_nm               = 3, 
+  color_sp_nm              = "black",
+  plot_sp_nm               = NULL,
+  fontface_sp_nm           = "plain",
+  save_file                = FALSE,
+  check_input              = TRUE)
+
+ #comparing convex hulls of different hulls
+plots_alpha_frugivore$"fric"$"patchwork"
+
+plots_alpha_frugivore$"fdiv"$"patchwork"
+
+plots_alpha_frugivore$"fspe"$"patchwork"
+
+plots_alpha_frugivore$"fdis"$"patchwork"
+
+plots_alpha_frugivore$"fori"$"patchwork"
+```
+
+## Turning the FD values into a raster
+Need to subset the coordinates made at the top to the same set of the assemblages (removed values less than 5 so these coords need to be removed to do this correctly)
+
+```{r}
+# Create an empty raster with desired resolution and extent
+resolution <- c(0.75, 0.75)  # Set the desired resolution
+extent <- c(-85, -54, -24, 14)  # Set the desired extent
+empty_raster <- raster(resolution = resolution, xmn = extent[1], xmx = extent[2], ymn = extent[3], ymx = extent[4])
+```
+
+```{r}
+# Generate coordinates
+subset_coords_frugivore <- site_loc_key_frugivore[rowSums(PAM_frugivore_site_final) >= 5, ]
+subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]
+
+#Get functional dispersion
+fdis_frugivore <- alpha_fd_indices_frugivore$functional_diversity_indices$fdis
+
+frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)
+```
+
+```{r}
+# Convert the dataframe to sf format
+spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
+```
+
+```{r}
+# Rasterize the sf data to create the FD raster
+fd_raster_frugivore <- rasterize(spatial_fdis_frugivore, empty_raster)
+writeRaster(fd_raster_frugivore$fdis_frugivore, filename = file.path(output_path,"FD_frugivore_TA.tif"), format="GTiff",overwrite=T)
+writeRaster(fd_raster_frugivore$ID, filename = file.path(output_path,"FD_frugivore_TA_ID.tif"), format="GTiff",overwrite=T)
+```
+
+# Plants
+## Create trait type table
+```{r}
+trait_name <- c("Growth.Form", "Plant.Height_m", "Fruit.Length_mm", "Fruit.Type")
+trait_type <- c("N", "Q", "Q", "N") 
+plant_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
+```
+
+## Summary of the assemblages * species dataframe
+```{r}
+asb_sp_plant_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_plant_site_final)
+```
+
+## Species traits summary
+```{r}
+plant_traits_summ <- mFD::sp.tr.summary(
+  tr_cat     = plant_trait_cat,   
+  sp_tr      = plant_traits_df_final, 
+  stop_if_NA = TRUE)
+```
+
+```{r}
+save(plant_traits_df_final, plant_trait_cat, file = "../../plant_funct.dist_files.rdata")
+```
+
+## Estimate functional trait-based distances between species
+```{r}
+sp_dist_plants <- mFD::funct.dist(
+  sp_tr         = plant_traits_df_final,
+  tr_cat        = plant_trait_cat,
+  metric        = "gower",
+  scale_euclid  = "scale_center",
+  ordinal_var   = "classic",
+  weight_type   = "equal",
+  stop_if_NA    = TRUE)
+```
+
+## Generate a multidimensional space
+```{r}
+fspaces_quality_plants <- mFD::quality.fspaces(
+  sp_dist             = sp_dist_plants,
+  maxdim_pcoa         = 10,
+  deviation_weighting = "absolute",
+  fdist_scaling       = FALSE,
+  fdendro             = "average")
+```
+
+```{r}
+# Look at the quality spaces only (MAD index looks at mean absolute deviation from the dissimilary matrix; want the deviation to be low, meaning that the true distances have been retained in the PCA)
+round(fspaces_quality_plants$"quality_fspaces", 3)            
+# Plot the quality spaces (chose to look at 3D, 4D, and 5D since they had the lowest MAD). Will go with 4 dimensions.
+mFD::quality.fspaces.plot(
+  fspaces_quality            = fspaces_quality_plants,
+  quality_metric             = "mad",
+  fspaces_plot               = c("pcoa_2d", "pcoa_3d", 
+                                 "pcoa_4d"),
+  name_file                  = NULL,
+  range_dist                 = NULL,
+  range_dev                  = NULL,
+  range_qdev                 = NULL,
+  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
+  gradient_deviation_quality = c(low = "yellow", high = "red"),
+  x_lab                      = "Trait-based distance")
+```
+
+```{r}
+#testing correlation between functional axes and traits
+sp_faxes_coord_plants <- fspaces_quality_plants$"details_fspaces"$"sp_pc_coord"
+```
+
+```{r}
+# View the components of the PCA axes *Remember the first few components explain the most variation in dissimilarity. Clusters into groups
+
+# Computes linear model for continuous traits and Kruskall-Wallis tests for other types. 
+plant_tr_faxes <- mFD::traits.faxes.cor(
+  sp_tr          = plant_traits_df_final, 
+  sp_faxes_coord = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")], 
+  plot           = TRUE)
+
+# Print traits with significant effect:
+plant_tr_faxes$"tr_faxes_stat"[which(plant_tr_faxes$"tr_faxes_stat"$"p.value" < 0.05), ]
+# Return plots:
+plant_tr_faxes$"tr_faxes_plot"
+```
+
+```{r}
+#plotting functional space
+sp_faxes_coord_plants <- fspaces_quality_plants$"details_fspaces"$"sp_pc_coord"
+
+big_plot_plants <- mFD::funct.space.plot(
+  sp_faxes_coord  = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")],
   faxes           = c("PC1", "PC2", "PC3", "PC4"),
@@
 subset_matrix_plant <- PAM_plant_site_final[row_sums_plant >= 4, ]
-row_sums_frugivore <- rowSums(PAM_frugivore_site_final)
-subset_matrix_frugivore <- PAM_frugivore_site_final[row_sums_frugivore >= 4, ]
-
 ```
@@
 big_plot_plants
-big_plot_frugivore
 ```
@@
 ```{r}
-# match frugivore names
-sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")]
-
-## check number of species names
-nrow(sp_faxes_coord_frugivore_sub) #764
-ncol(subset_matrix_frugivore) #782
-
-sp_faxes_coord_frugivore_sub_names <- chartr(".", " ", sp_faxes_coord_frugivore_sub_names)
-sp_faxes_coord_frugivore_sub <- as.data.frame(sp_faxes_coord_frugivore_sub)
-row.names(sp_faxes_coord_frugivore_sub) <- sp_faxes_coord_frugivore_sub_names
-
-sp_faxes_coord_frugivore_sub_names <- row.names(sp_faxes_coord_frugivore_sub)
-subset_matrix_frugivore_names <- colnames(subset_matrix_frugivore)
-
-frugivore_names <- intersect(sp_faxes_coord_frugivore_sub_names, subset_matrix_frugivore_names)
-frugivore_names <- na.omit(frugivore_names)
-
-sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore_sub[ which((row.names(sp_faxes_coord_frugivore_sub) %in% frugivore_names)==TRUE), ]
-
-subset_matrix_frugivore  <- as.data.frame(subset_matrix_frugivore)
-subset_matrix_frugivore <- subset_matrix_frugivore[ ,which((colnames(subset_matrix_frugivore) %in% frugivore_names)==TRUE)]
-
-#remove NAs
-sp_faxes_coord_frugivore_sub <- na.omit(sp_faxes_coord_frugivore_sub)
-subset_matrix_frugivore <- na.omit(subset_matrix_frugivore)
-
-nrow(sp_faxes_coord_frugivore_sub) #757
-ncol(subset_matrix_frugivore) #757
-sp_faxes_coord_frugivore_sub <- as.matrix(sp_faxes_coord_frugivore_sub)
-subset_matrix_frugivore <- as.matrix(subset_matrix_frugivore)
-```
-
-```{r}
 #computing FD
@@
 
-alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(
-  sp_faxes_coord   = sp_faxes_coord_frugivore_sub,
-  asb_sp_w         = subset_matrix_frugivore,
-  ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
-  scaling          = TRUE,
-  check_input      = TRUE,
-  details_returned = TRUE)
-
 details_list_plant <- alpha_fd_indices_plant$"details"
-details_list_frugivore <- alpha_fd_indices_frugivore$"details"
 
@@
 
-plots_alpha_frugivore <- mFD::alpha.multidim.plot(
-  output_alpha_fd_multidim = alpha_fd_indices_frugivore,
-  plot_asb_nm              = c("cell4", "cell20"),
-  ind_nm                   = c("fdis", "fide", "fnnd"
-                        , "fori", "fspe"),
-  faxes                    = NULL,
-  faxes_nm                 = NULL,
-  range_faxes              = c(NA, NA),
-  color_bg                 = "grey95",
-  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
-  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
-  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  color_ch          
-  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
-  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
-  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
-  shape_centroid_fspe      = 23,
-  color_centroid_fspe      = "black",
-  size_sp_nm               = 3, 
-  color_sp_nm              = "black",
-  plot_sp_nm               = NULL,
-  fontface_sp_nm           = "plain",
-  save_file                = FALSE,
-  check_input              = TRUE)
+# comparing convex hulls of different hulls
+plots_alpha_plant$"fric"$"patchwork"
 
-plots_alpha_plant$"fric"$"patchwork"  #comparing convex hulls of different hulls
-plots_alpha_frugivore$"fric"$"patchwork"
 plots_alpha_plant$"fdiv"$"patchwork"
-plots_alpha_frugivore$"fdiv"$"patchwork"
+
 plots_alpha_plant$"fspe"$"patchwork"
-plots_alpha_frugivore$"fspe"$"patchwork"
+
 plots_alpha_plant$"fdis"$"patchwork"
-plots_alpha_frugivore$"fdis"$"patchwork"
+
 plots_alpha_plant$"fori"$"patchwork"
-plots_alpha_frugivore$"fori"$"patchwork"
+
 ```
 
-# Turning the FD values into a raster
+## Turning the FD values into a raster
+Need to subset the coordinates made at the top to the same set of the assemblages (removed values less than 5 so these coords need to be removed to do this correctly)
+
 ```{r}
-
-#Need to subset the coordinates made at the top to the same set of the assemblages (removed values less than 5 so these coords need to be removed to do this correctly)
-
-#Have coordinates and associated values
-
-library(raster)
-
 # Create an empty raster with desired resolution and extent
@@
 empty_raster <- raster(resolution = resolution, xmn = extent[1], xmx = extent[2], ymn = extent[3], ymx = extent[4])
+```
 
+```{r}
 # Generate coordinates
@@
 subset_coords_plant_sp <-subset_coords_plant[,1:2 ]
-subset_coords_frugivore <- site_loc_key_frugivore[rowSums(PAM_frugivore_site_final) >= 5, ]
-subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]
 
@@
 fdis_plant <- alpha_fd_indices_plant$functional_diversity_indices$fdis
-fdis_frugivore <- alpha_fd_indices_frugivore$functional_diversity_indices$fdis
 
 plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
-frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)
 
@@
 spatial_fdis_plant <- st_as_sf(plant_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
-spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
-
 ```
@@
 writeRaster(fd_raster_plant$ID, filename = file.path(output_path,"FD_plant_TA_ID.tif"), format="GTiff")
+```
 
-fd_raster_frugivore <- rasterize(spatial_fdis_frugivore, empty_raster)
-writeRaster(fd_raster_frugivore$fdis_frugivore, filename = file.path(output_path,"FD_frugivore_TA.tif"), format="GTiff",overwrite=T)
-writeRaster(fd_raster_frugivore$ID, filename = file.path(output_path,"FD_frugivore_TA_ID.tif"), format="GTiff")
+# Calculating Rao's Q
+Calculate Rao's Q using occurrence data and trait dissimilarity
+
+```{r}
+## Define different spatial scales (e.g., grid cell sizes)
+spatial_scales <- c(1000, 5000, 10000, 25000, 50000, 75000, 100000)  # In meters
+
+## Create an empty dataframe to store results
+plant_raos_q_results <- data.frame(spatial_scale = numeric(),
+                             raos_q = numeric())
+
+## Iterate over different spatial scales
+for (scale in spatial_scales) {
+  ### Aggregate occurrence data to the current spatial scale
+  plant_occurrence_aggregated <- aggregate_occurrence(plant_abundance_per_species, scale)
+
+  ### Calculate dissimilarities using trait data
+  plant_trait_dissimilarities <- vegdist(plant_traits, method = "bray")
+
+  ### Calculate dissimilarities using occurrence data
+  plant_occurrence_dissimilarities <- vegdist(plant_occurrence_aggregated, method = "bray")
+
+  ### Calculate Rao's Q index
+  raos_q <- raosq(plant_trait_dissimilarities, plant_occurrence_dissimilarities)
+
+  #### Add results to the dataframe
+  plant_raos_q_results <- rbind(raos_q_results, data.frame(spatial_scale = scale, raos_q = raos_q))
+}
+```
+
+```{r}
+## Create an empty dataframe to store results
+frugivore_raos_q_results <- data.frame(spatial_scale = numeric(),
+                             raos_q = numeric())
+
+## Iterate over different spatial scales
+for (scale in spatial_scales) {
+  ### Aggregate occurrence data to the current spatial scale
+  frugivore_occurrence_aggregated <- aggregate_occurrence(frugivore_abundance_per_species, scale)
+
+  ### Calculate dissimilarities using trait data
+  frugivore_trait_dissimilarities <- vegdist(frugivore_traits, method = "bray")
+
+  ### Calculate dissimilarities using occurrence data
+  frugivore_occurrence_dissimilarities <- vegdist(frugivore_occurrence_aggregated, method = "bray")
+
+  ### Calculate Rao's Q index
+  raos_q <- raosq(frugivore_trait_dissimilarities, frugivore_occurrence_dissimilarities)
+
+  #### Add results to the dataframe
+  frugivore_raos_q_results <- rbind(raos_q_results, data.frame(spatial_scale = scale, raos_q = raos_q))
+}
 ```
@@
 # Functional diversity plots
-# Polygons of countries
-```{r}
-worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
-Americas <- ne_countries(continent = c("North America", "South America"), returnclass = "sf")
-#polygon of Tropical Andes
-TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
-```
 
+## Project FD objects
 ```{r}
 # set crs of sf objects
-spatial_fdis_plant <- spatial_fdis_plant %>% st_set_crs(4326)
+#spatial_fdis_plant <- spatial_fdis_plant %>% st_set_crs(4326)
 spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(4326)
@@
 ```{r}
+#spatial_fdis_plant_projected <- st_transform(spatial_fdis_plant, 5389)
+spatial_fdis_frugivore_projected <- st_transform(spatial_fdis_frugivore, 5389)
+```
+
+```{r}
+# determine resolution of spatial_fdis_frugivore_projected 
+st_crs(spatial_fdis_frugivore_projected, parameters = TRUE)$units_gdal
+```
+
+```{r}
 # Raster to raster point
-spec_FD_raster_plant_point <- rasterToPoints(fd_raster_plant, spatial = TRUE)
+#spec_FD_raster_plant_point <- rasterToPoints(fd_raster_plant, spatial = TRUE)
 spec_FD_raster_frugivore_point <- rasterToPoints(fd_raster_frugivore, spatial = TRUE)
 # Convert to a 'conventional' dataframe
-fd_raster_plant_point <- data.frame(spec_FD_raster_plant_point)
+#fd_raster_plant_point <- data.frame(spec_FD_raster_plant_point)
 fd_raster_frugivore_point <- data.frame(spec_FD_raster_frugivore_point)
@@
   scale_fill_scico(palette = "romaO", direction = 1) +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
+  coord_sf(xlim = c(-85, -60), ylim = c(-24, 14), expand = FALSE) +
   theme(

### 2024-05-14 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
update plotting function to set legend scale and color for plants constant (same frugivores); remove cells with richness of 0 from scatterplot, add exponential model to comparison scatterplot with r-squared, re-organize multipanel plot order

@@
 ## Function for richness plots
-Need to set legend scale and color for plants constant (same for frugivores)
-remove cells with richness of 0
 ```{r}
@@
 
-  # Helper function to generate plots
-  generate_plot <- function(data, title) {
+  # Helper functions to generate plots
+  generate_plant_plot <- function(data) {
     ggplot(data) +
@@
       geom_sf(aes(fill = num_species), color = NA) +
-      scale_fill_scico(palette = "romaO", direction = 1) +
+      scale_fill_distiller(pallete = "Greens", limits = c(0,6000)) +
+      labs(title = title, x = "Longitude", y = "Latitude", fill = "Number of species") +
+      coord_sf(xlim = c(-85, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+      annotation_scale(location = "bl", width_hint = 0.5)  +
+      theme(panel.background = element_rect(fill = "lightblue"))
+  }
+  generate_frugivore_plot <- function(data) {
+    ggplot(data) +
+      geom_sf(data = Americas, fill = "white") +
+      geom_sf(data = TApoly, fill = "grey", size = 0.1) +
+      geom_sf(aes(fill = num_species), color = NA) +
+      scale_fill_distiller(pallete = "YlOrBr", limits = c(0,500)) + 
       labs(title = title, x = "Longitude", y = "Latitude", fill = "Number of species") +
@@
   # Generate plots
-  plant_plot <- generate_plot(plant_richness_grid, paste("Tropical Andes Plant Richness [", resolution_meters/1000, " km]", sep=""))
-  plant_plot <- generate_plot(frugivore_richness_grid, paste("Tropical Andes Frugivore Richness [", resolution_meters/1000, " km]", sep=""))
+  plant_plot <- generate_plant_plot(plant_richness_grid)
+  frugivore_plot <- generate_frugivore_plot(frugivore_richness_grid)
 
@@
 
-  # Compare plant and frugivore richness
-  compare_richness_summary <- summary(
-    lm(frugivore_richness_grid$num_species ~ plant_richness_grid$num_species))
-  r_squared <- compare_richness_summary$r.squared
-
     comparison_data <- data.frame(Plant_Richness = plant_richness_grid$num_species,
                                Frugivore_Richness = frugivore_richness_grid$num_species)
+  # Filter out rows with richness = 0 for both plant and frugivore
+  comparison_data_filtered <- comparison_data %>%
+                                filter(Plant_Richness > 0 | Frugivore_Richness > 0)
 
-  ## Create scatter plot
-  plant_frug_richness_compare <- ggplot(comparison_data,
+  # Compare plant and frugivore richness
+  ## Fit a linear model
+  compare_richness_linear_summary <- summary(lm(Frugivore_Richness ~ Plant_Richness),
+                                      data = comparison_data_filtered)
+  linear_r_squared <- compare_richness_linear_summary$r.squared
+
+  ## Fit an exponential model
+  compare_richness_exponential_summary <- summary(lm(log(Frugivore_Richness) ~ log(Plant_Richness),
+                                      data = comparison_data_filtered))
+  exponential_r_squared <- compare_richness_exponential_summary$r.squared
+
+  ## Create scatter plot with linear and exponential lines
+  plant_frug_richness_compare <- ggplot(comparison_data_filtered,
                                         aes(x = Plant_Richness, y = Frugivore_Richness)) +
@@
   labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
-       subtitle =  paste("R-squared = ", round(r_squared, 2), sep = ""),
+       subtitle =  paste("Linear (Red) R-squared = ", round(linear_r_squared, 2), "\n",
+                  "Exponential (Blue) R-squared = ", round(exponential_r_squared, 2)),
        x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
-  geom_smooth(method = "lm", se = FALSE, color = "red")  # Add regression line
+    geom_smooth(method = "lm", se = FALSE, color = "red") +
+    geom_smooth(method = "lm", se = FALSE, color = "blue", formula = y ~ exp(x), linetype = "dashed")
 
   # Create histograms
-  plant_richness_hist <- hist(plant_richness_grid$num_species, 
+  plant_richness_hist <- hist(comparison_data_filtered$PlantRichness, 
     main = paste("Histogram of Plant Richness [", resolution_meters/1000, " km] by cell", sep = ""),
                               xlab = "Plant Richness by cell")
-  frugivore_richness_hist <- hist(frugivore_richness_grid$num_species,
+  frugivore_richness_hist <- hist(comparison_data_filtered$FrugivoreRichness,
     main = paste("Histogram of Frugivore Richness [", resolution_meters/1000, " km] by cell", sep = ""),
@@
        frugivore_cellRichness = frugivore_richness_grid, 
-       compare_richness_summary = compare_richness_summary,
+       compare_richness_linear_summary = compare_richness_linear_summary,
+       compare_richness_exponential_summary = compare_richness_exponential_summary,
        plant_frug_richness_compare = plant_frug_richness_compare,
@@
 ```{r}
-all_richness_plots <- gridExtra::grid.arrange(#plantgridRichTA_1km, frugivoregridRichTA_1km,
-                                              #plantgridRichTA_5km, frugivoregridRichTA_5km, 
-                                               plantgridRichTA_10km, frugivoregridRichTA_10km,
-                                               plantgridRichTA_25km, frugivoregridRichTA_25km,
-                                               plantgridRichTA_50km, frugivoregridRichTA_50km,
-                                               plantgridRichTA_75km,  frugivoregridRichTA_75km,
-                                               plantgridRichTA_100km, frugivoregridRichTA_100km,
+all_richness_plots <- gridExtra::grid.arrange(#plantgridRichTA_1km, plantgridRichTA_5km, 
+                                              plantgridRichTA_10km, plantgridRichTA_25km,
+                                              plantgridRichTA_50km, plantgridRichTA_75km,
+                                              plantgridRichTA_100km,
+                                              #frugivoregridRichTA_1km, frugivoregridRichTA_5km, 
+                                              frugivoregridRichTA_10km, frugivoregridRichTA_25km,
+                                              frugivoregridRichTA_50km, frugivoregridRichTA_75km,
+                                              frugivoregridRichTA_100km, 
                                               ncol = 5, nrow = 2)

### 2024-05-14 code/L1/L1_LastStep_DiversityInputObjects.Rmd
save site location keys for plants and frugivores

@@
 # objects for functional diversity
+saveRDS(site_loc_key_plant, file = file.path(data_path_L1,"site_loc_key_plant.rds"))
+saveRDS(site_loc_key_frugivore, file = file.path(data_path_L1,"site_loc_key_frugivore.rds"))
 saveRDS(PAM_plant_site_final, file = file.path(data_path_L1,"PAM_plant_site_final.rds"))

### 2024-05-14 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
read in site location key files from L1, update extent, reorder more code

@@
 frugivore_traits_df_final <- readRDS(file = file.path(data_path_L1,"frugivore_traits_df_final.rds"))
+site_loc_key_plant <- readRDS(file = file.path(data_path_L1,"site_loc_key_plant.rds"))
+site_loc_key_frugivore <- readRDS(file = file.path(data_path_L1,"site_loc_key_frugivore.rds"))
 ```
@@
 
+```{r}
+# Get functional dispersion
+fdis_frugivore <- alpha_fd_indices_frugivore$functional_diversity_indices$fdis
+```
+
 ## Turning the FD values into a raster
@@
 resolution <- c(0.75, 0.75)  # Set the desired resolution
-extent <- c(-85, -54, -24, 14)  # Set the desired extent
+extent <- c(-85, 60, -24, 14)  # Set the desired extent
 empty_raster <- raster(resolution = resolution, xmn = extent[1], xmx = extent[2], ymn = extent[3], ymx = extent[4])
@@
 subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]
-
-#Get functional dispersion
-fdis_frugivore <- alpha_fd_indices_frugivore$functional_diversity_indices$fdis
 
@@
 writeRaster(fd_raster_frugivore$ID, filename = file.path(output_path,"FD_frugivore_TA_ID.tif"), format="GTiff",overwrite=T)
+```
+
+## Project FD objects
+```{r}
+# set crs of sf objects
+spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(4326)
+```
+
+```{r}
+spatial_fdis_frugivore_projected <- st_transform(spatial_fdis_frugivore, 5389)
+```
+
+```{r}
+# determine resolution of spatial_fdis_frugivore_projected 
+st_crs(spatial_fdis_frugivore_projected, parameters = TRUE)$units_gdal
+```
+
+```{r}
+# Raster to raster point
+spec_FD_raster_frugivore_point <- rasterToPoints(fd_raster_frugivore, spatial = TRUE)
+# Convert to dataframe
+fd_raster_frugivore_point <- data.frame(spec_FD_raster_frugivore_point)
+```
+
+```{r}
+frugivoregridFDTA <-
+  ggplot() +
+  geom_sf(data = Americas, fill = "white")+
+  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
+  geom_tile(data= fd_raster_frugivore_point, aes(x=x, y=y, fill=fdis_frugivore)) +
+  scale_fill_scico(palette = "romaO", direction = 1) +
+  coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE) +
+  theme(
+    plot.background = element_rect(fill = "#f1f2f3"),
+    panel.background = element_rect(fill = "lightblue"),
+    panel.grid = element_blank(),
+    line = element_blank(),
+    rect = element_blank()
+  ) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  labs(title = "Tropical Andes Frugivore Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
+frugivoregridFDTA
 ```
@@
 
+```{r}
+#Get functional dispersion
+fdis_plant <- alpha_fd_indices_plant$functional_diversity_indices$fdis
+```
+
 ## Turning the FD values into a raster
@@
 
-#Get functional dispersion
-fdis_plant <- alpha_fd_indices_plant$functional_diversity_indices$fdis
-
 plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
@@
 writeRaster(fd_raster_plant$ID, filename = file.path(output_path,"FD_plant_TA_ID.tif"), format="GTiff")
+```
+
+## Project FD objects
+```{r}
+# set crs of sf objects
+spatial_fdis_plant <- spatial_fdis_plant %>% st_set_crs(4326)
+```
+
+```{r}
+spatial_fdis_plant_projected <- st_transform(spatial_fdis_plant, 5389)
+```
+
+```{r}
+# Raster to raster point
+spec_FD_raster_plant_point <- rasterToPoints(fd_raster_plant, spatial = TRUE)
+# Convert to a 'conventional' dataframe
+fd_raster_plant_point <- data.frame(spec_FD_raster_plant_point)
+```
+
+```{r}
+plantgridFDTA <-
+  ggplot() +
+  geom_sf(data = Americas, fill = "white")+
+  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
+  geom_tile(data= fd_raster_plant_point, aes(x=x, y=y, fill=fdis_plant)) +
+  scale_fill_scico(palette = "romaO", direction = 1) +
+  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
+  theme(
+    plot.background = element_rect(fill = "#f1f2f3"),
+    panel.background = element_rect(fill = "lightblue"),
+    panel.grid = element_blank(),
+    line = element_blank(),
+    rect = element_blank()
+  ) +
+  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_north_arrow(location = "bl", which_north = "true", 
+  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  labs(title = "Tropical Andes Plant Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
+plantgridFDTA
 ```
@@
 
-# Functional diversity plots
-
-## Project FD objects
-```{r}
-# set crs of sf objects
-#spatial_fdis_plant <- spatial_fdis_plant %>% st_set_crs(4326)
-spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(4326)
-```
-
-```{r}
-#spatial_fdis_plant_projected <- st_transform(spatial_fdis_plant, 5389)
-spatial_fdis_frugivore_projected <- st_transform(spatial_fdis_frugivore, 5389)
-```
-
-```{r}
-# determine resolution of spatial_fdis_frugivore_projected 
-st_crs(spatial_fdis_frugivore_projected, parameters = TRUE)$units_gdal
-```
-
-```{r}
-# Raster to raster point
-#spec_FD_raster_plant_point <- rasterToPoints(fd_raster_plant, spatial = TRUE)
-spec_FD_raster_frugivore_point <- rasterToPoints(fd_raster_frugivore, spatial = TRUE)
-# Convert to a 'conventional' dataframe
-#fd_raster_plant_point <- data.frame(spec_FD_raster_plant_point)
-fd_raster_frugivore_point <- data.frame(spec_FD_raster_frugivore_point)
-```
-
-```{r}
-plantgridFDTA <-
-  ggplot() +
-  geom_sf(data = worldMap, fill = "white")+
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_tile(data= fd_raster_plant_point, aes(x=x, y=y, fill=fdis_plant)) +
-  scale_fill_scico(palette = "romaO", direction = 1) +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  ) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Plant Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
-plantgridFDTA
-```
-
-```{r}
-frugivoregridFDTA <-
-  ggplot() +
-  geom_sf(data = worldMap, fill = "white")+
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_tile(data= fd_raster_frugivore_point, aes(x=x, y=y, fill=fdis_frugivore)) +
-  scale_fill_scico(palette = "romaO", direction = 1) +
-  coord_sf(xlim = c(-85, -60), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  ) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Frugivore Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
-frugivoregridFDTA
-```
-
 # Extract cell values

### 2024-05-14 code/Manuscript_figures.rmd
add plant cleaning visualizations

@@
-Summary plots
+---
+title: "Manuscript Figures"
+author: "Hazel J. Anderson"
+project: "Plant-Frugivore Diversity"
+collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
+overview: "This script generates misc figures for manuscript that corresponds to this project."
+data input: ""
+data output: ""
+date: "2024-05-14"
+output: 
+    html_document:
+      code_link: TRUE
+      code_download: TRUE
+---
+
 # Load required packages
+
 ```{r}
@@
 ```
+
+# Creating cleaning occurrence data figures
+
+These numbers are recorded from the summary function at the end of most L0 and L1 scripts. A table is available in the manuscript supplemential materials.
+
+## Frugivores
 
@@
 ```
+
 ```{r}
@@
 ```{r}
-figure <- ggarrange(frug_records_fig,frug_species_fig, frug_genera_fig, frug_families_fig,
+frug_cleaning <- ggarrange(frug_records_fig,frug_species_fig, frug_genera_fig, frug_families_fig,
                     labels = c("A", "B", "C", "D"),
                     ncol = 2, nrow = 2)
-figure
+frug_cleaning
+```
+
+## Plants
+
+```{r}
+Step <- c("Raw", "After cleaning", "Full trait coverage")
+plant_records <- c(4519792, 1886120, 1700044)
+plant_species <- c(40288, 39972, 33746)
+plant_genera <- c(4083, 3793, 3400)
+plant_families <- c(307, 293, 275)
+
+plant_records_df <- data.frame(Step, plant_records)
+plant_species_df <- data.frame(Step, plant_species)
+plant_genera_df <- data.frame(Step, plant_genera)
+plant_families_df <- data.frame(Step, plant_families)
 ```
@@
 ```{r}
+ggplot(plant_records_df, aes(x = Step, y = plant_records)) +
+  geom_bar(stat = "identity", fill = "darkgreen") +
+  coord_flip() +
+  geom_text(aes(label = plant_records), vjust = -0.3, size = 4) + 
+  labs(title = "Plant Records", y = "Count") +
+  theme_minimal()
+```
+
+```{r}
+plant_records_fig <- ggfunnel::funnel(plant_records_df, levels = Step, values = plant_records, text_specs = list(colour = "white", size = 3)) +
+  labs(title = "Plant records", y = NULL)
+plant_records_fig
+```
+
+```{r}
+plant_species_fig <- ggfunnel::funnel(plant_species_df, levels = Step, values = plant_species, text_specs = list(colour = "white", size = 3)) +
+  labs(title = "Plant species", y = NULL)
+plant_species_fig
+```
+
+```{r}
+plant_genera_fig <- ggfunnel::funnel(plant_genera_df, levels = Step, values = plant_genera, text_specs = list(colour = "white", size = 3)) +
+  labs(title = "Plant genera", y = NULL)
+plant_genera_fig
+```
+
+```{r}
+plant_families_fig <- ggfunnel::funnel(plant_families_df, levels = Step, values = plant_families, text_specs = list(colour = "white", size = 3)) +
+  labs(title = "Plant families", y = NULL)
+plant_families_fig
+```
+
+```{r}
+plant_cleaning <- ggarrange(plant_records_fig,plant_species_fig, plant_genera_fig, plant_families_fig,
+                    labels = c("A", "B", "C", "D"),
+                    ncol = 2, nrow = 2)
+plant_cleaning
+```
+
+# References
+```{r}
+library(report)
+cite_packages()
+```
+
+# Session Info
+```{r, class.source= 'fold-hide'}
 devtools::session_info()

### 2024-05-14 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
update data names and paths to read in, change plot function scale color and limits, update variable names to fix errors

--- 2024-05-14 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
@@
 library(ggspatial)
-library(gridExtra)
+library(ggpubr)
 ```
@@
 ```{r}
-TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned_subset.csv"))
-TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_frugivore_occ_cleaned_subset.csv"))
-TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
+## projected sf objects
+plants_sf_species <- readRDS(file = file.path(data_path_L1,"plants_sf_species.rds"))
+frugivores_sf_species <- readRDS(file = file.path(data_path_L1,"frugivores_sf_species.rds"))
+Americas <- readRDS(file = file.path(data_path_L1, "Americas.rds"))
+TApoly <- readRDS(file = file.path(data_path_L1,"TApoly.rds"))
+TropicalAndes_IUCNHabitat_Forest <- readRDS(file = file.path(data_path_L1,"TropicalAndes_IUCNHabitat_Forest.rds"))
 ```
@@
       geom_sf(aes(fill = num_species), color = NA) +
-      scale_fill_scico(palette = "romaO", direction = 1) +
+      scale_fill_distiller(palette = "Greens", limits = c(0,6000)) +
       labs(title = title, x = "Longitude", y = "Latitude", fill = "Number of species") +
-      coord_sf(xlim = c(-85, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+      coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+      annotation_scale(location = "bl", width_hint = 0.5)  +
+      theme(panel.background = element_rect(fill = "lightblue"))
+  }
+  generate_frugivore_plot <- function(data) {
+    ggplot(data) +
+      geom_sf(data = Americas, fill = "white") +
+      geom_sf(data = TApoly, fill = "grey", size = 0.1) +
+      geom_sf(aes(fill = num_species), color = NA) +
+      scale_fill_distiller(palette = "YlOrBr", limits = c(0,500)) + 
+      labs(title = title, x = "Longitude", y = "Latitude", fill = "Number of species") +
+      coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
       annotation_scale(location = "bl", width_hint = 0.5)  +
@@
   # Generate plots
-  plant_plot <- generate_plot(plant_richness_grid, paste("Tropical Andes Plant Richness [", resolution_meters/1000, " km]", sep=""))
-  plant_plot <- generate_plot(frugivore_richness_grid, paste("Tropical Andes Frugivore Richness [", resolution_meters/1000, " km]", sep=""))
+  plant_plot <- generate_plant_plot(plant_richness_grid)
+  frugivore_plot <- generate_frugivore_plot(frugivore_richness_grid)
 
   # Extract cell values
-  plant_cellRichness <- plant_plot[["data"]]
-  frugivore_cellRichness <- plant_plot[["data"]]
-
-  # Compare plant and frugivore richness
-  compare_richness_summary <- summary(
-    lm(frugivore_richness_grid$num_species ~ plant_richness_grid$num_species))
-  r_squared <- compare_richness_summary$r.squared
-
   comparison_data <- data.frame(Plant_Richness = plant_richness_grid$num_species,
@@
 
-  ## Create scatter plot
+  # Filter out rows where either plant or frugivore richness is zero
+  comparison_data_filtered <- comparison_data %>%
+                                filter(Plant_Richness > 0 & Frugivore_Richness > 0)
+  # Compare plant and frugivore richness
+  ## Fit a linear model
+  compare_richness_linear_summary <- summary(lm(comparison_data_filtered$Frugivore_Richness ~
+                                                  comparison_data_filtered$Plant_Richness))
+  linear_r_squared <- compare_richness_linear_summary$r.squared
+
+  ## Fit an exponential model
+  compare_richness_exponential_summary <- summary(lm(exp(comparison_data_filtered$Frugivore_Richness) ~
+                                                       comparison_data_filtered$Plant_Richness))
+  exponential_r_squared <- compare_richness_exponential_summary$r.squared
+
+  ## Create scatter plot with linear and exponential lines
   plant_frug_richness_compare <- ggplot(comparison_data,
@@
   labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
-       subtitle =  paste("R-squared = ", round(r_squared, 2), sep = ""),
+       subtitle =  paste("Linear (Red) R-squared = ", round(linear_r_squared, 2), "\n",
+                  "Exponential (Blue) R-squared = ", round(exponential_r_squared, 2)),
        x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
-  geom_smooth(method = "lm", se = FALSE, color = "red")  # Add regression line
+    geom_smooth(method = "lm", se = FALSE, color = "red") +
+    geom_smooth(method = "lm", se = FALSE, color = "blue", formula = y ~ exp(x), linetype = "dashed")
 
   # Create histograms
-  plant_richness_hist <- hist(plant_richness_grid$num_species, 
+  plant_richness_hist <- hist(comparison_data$Plant_Richness, 
     main = paste("Histogram of Plant Richness [", resolution_meters/1000, " km] by cell", sep = ""),
                               xlab = "Plant Richness by cell")
-  frugivore_richness_hist <- hist(frugivore_richness_grid$num_species,
+  frugivore_richness_hist <- hist(comparison_data$Frugivore_Richness,
     main = paste("Histogram of Frugivore Richness [", resolution_meters/1000, " km] by cell", sep = ""),
                                   xlab = "Frugivore Richness by cell")
-
-  # Return a list of plots
-  list(plantgridRichTA = plant_plot,
-       frugivoregridRichTA = frugivore_plot,
-       plant_cellRichness = plant_richness_grid,
-       frugivore_cellRichness = frugivore_richness_grid, 
-       compare_richness_summary = compare_richness_summary,
-       plant_frug_richness_compare = plant_frug_richness_compare,
-       plant_richness_hist = plant_richness_hist,
-       frugivore_richness_hist = frugivore_richness_hist)
 }
@@
 ```{r}
-all_richness_plots <- gridExtra::grid.arrange(#plantgridRichTA_1km, frugivoregridRichTA_1km,
-                                              #plantgridRichTA_5km, frugivoregridRichTA_5km, 
-                                               plantgridRichTA_10km, frugivoregridRichTA_10km,
-                                               plantgridRichTA_25km, frugivoregridRichTA_25km,
-                                               plantgridRichTA_50km, frugivoregridRichTA_50km,
-                                               plantgridRichTA_75km,  frugivoregridRichTA_75km,
-                                               plantgridRichTA_100km, frugivoregridRichTA_100km,
-                                               ncol = 5, nrow = 2)
+all_richness_plots <- ggpubr::ggarrange(#plantgridRichTA_1km, plantgridRichTA_5km, 
+                                              plantgridRichTA_10km, plantgridRichTA_25km,
+                                              plantgridRichTA_50km, plantgridRichTA_75km,
+                                              plantgridRichTA_100km,
+                                              #frugivoregridRichTA_1km, frugivoregridRichTA_5km, 
+                                              frugivoregridRichTA_10km, frugivoregridRichTA_25km,
+                                              frugivoregridRichTA_50km, frugivoregridRichTA_75km,
+                                              frugivoregridRichTA_100km, 
+                                              ncol = 5, nrow = 2,
+                                              labels = c("10km", "25km", ))
 all_richness_plots
@@
 #ADD compare_richness_r2_1km, compare_richness_r2_5km,
-r_squared <- c( compare_richness_r2_10km, compare_richness_r2_25km, compare_richness_r2_50km, compare_richness_r2_75km, compare_richness_r2_100km)
+r_squared <- c(compare_richness_r2_10km, compare_richness_r2_25km, compare_richness_r2_50km, compare_richness_r2_75km, compare_richness_r2_100km)
 r_squared_compare <- data.frame(resolution, r_squared)

### 2024-05-15 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
in richness plotting function: add direction for palettes and remove title; update exponential model equation, create seperate scatterplots for exponential and linear models, add return list of plots that accidentially was removed

@@
       geom_sf(aes(fill = num_species), color = NA) +
-      scale_fill_distiller(palette = "Greens", limits = c(0,6000)) +
-      labs(title = title, x = "Longitude", y = "Latitude", fill = "Number of species") +
+      scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,6000)) +
+      labs( x = "Longitude", y = "Latitude", fill = "Number of species") +
       coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
@@
       geom_sf(aes(fill = num_species), color = NA) +
-      scale_fill_distiller(palette = "YlOrBr", limits = c(0,500)) + 
-      labs(title = title, x = "Longitude", y = "Latitude", fill = "Number of species") +
+      scale_fill_distiller(palette = "YlOrBr", direction = 1, limits = c(0,500)) + 
+      labs(x = "Longitude", y = "Latitude", fill = "Number of species") +
       coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
@@
   ## Fit an exponential model
-  compare_richness_exponential_summary <- summary(lm(exp(comparison_data_filtered$Frugivore_Richness) ~
-                                                       comparison_data_filtered$Plant_Richness))
+  compare_richness_exponential_summary <- summary(lm(comparison_data_filtered$Frugivore_Richness ~
+                                                       log(comparison_data_filtered$Plant_Richness)))
   exponential_r_squared <- compare_richness_exponential_summary$r.squared
@@
   ## Create scatter plot with linear and exponential lines
-  plant_frug_richness_compare <- ggplot(comparison_data,
+  compare_richness_linear_plot <- ggplot(comparison_data,
                                         aes(x = Plant_Richness, y = Frugivore_Richness)) +
@@
   labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
-       subtitle =  paste("Linear (Red) R-squared = ", round(linear_r_squared, 2), "\n",
-                  "Exponential (Blue) R-squared = ", round(exponential_r_squared, 2)),
+       subtitle =  paste("Linear (Red) R-squared = ", round(linear_r_squared, 2)),
        x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
-    geom_smooth(method = "lm", se = FALSE, color = "red") +
-    geom_smooth(method = "lm", se = FALSE, color = "blue", formula = y ~ exp(x), linetype = "dashed")
+    geom_smooth(method = "lm", se = FALSE, color = "red")
+
+  compare_richness_exponential_plot <- ggplot(comparison_data,
+                                        aes(x = Plant_Richness, y = Frugivore_Richness)) +
+  geom_point() +
+  labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
+       subtitle =  paste("Exponential (Blue) R-squared = ", round(exponential_r_squared, 2)),
+       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
+    geom_smooth(method = "lm", se = FALSE, color = "blue", formula = y ~ log(x))
 
@@
                                   xlab = "Frugivore Richness by cell")
+  # Return a list of plots
+  list(plantgridRichTA = plant_plot,
+       frugivoregridRichTA = frugivore_plot,
+       plant_cellRichness = plant_richness_grid,
+       frugivore_cellRichness = frugivore_richness_grid, 
+       compare_richness_linear_summary = compare_richness_linear_summary,
+       compare_richness_exponential_summary = compare_richness_exponential_summary,
+       compare_richness_linear_plot = compare_richness_linear_plot,
+       compare_richness_exponential_plot = compare_richness_exponential_plot,
+       plant_richness_hist = plant_richness_hist,
+       frugivore_richness_hist = frugivore_richness_hist)
 }
@@
 frugivore_cellRichness_1km <- richness_1km$frugivore_cellRichness
-
-compare_richness_r2_1km <- richness_1km$compare_richness_summary$r.squared
+# save linear and exponential summaries
+compare_richness_linear_summary_1km <- richness_1km$compare_richness_linear_summary
+compare_richness_exponential_summary_1km <- richness_1km$compare_richness_exponential_summary
+# save r-squared for linear and exponential models
+compare_richness_linear_r2_1km <- richness_1km$compare_richness_linear_summary$r.squared
+compare_richness_exponential_r2_1km <- richness_1km$compare_richness_exponential_summary$r.squared
 ```
@@
 frugivore_cellRichness_5km <- richness_5km$frugivore_cellRichness
-
-compare_richness_r2_5km <- richness_5km$compare_richness_summary$r.squared
+# save linear and exponential summaries
+compare_richness_linear_summary_5km <- richness_5km$compare_richness_linear_summary
+compare_richness_exponential_summary_5km <- richness_5km$compare_richness_exponential_summary
+# save r-squared for linear and exponential models
+compare_richness_linear_r2_5km <- richness_5km$compare_richness_linear_summary$r.squared
+compare_richness_exponential_r2_5km <- richness_5km$compare_richness_exponential_summary$r.squared
 ```
@@
 frugivore_cellRichness_10km <- richness_10km$frugivore_cellRichness
-
-compare_richness_r2_10km <- richness_10km$compare_richness_summary$r.squared
+# save linear and exponential summaries
+compare_richness_linear_summary_10km <- richness_10km$compare_richness_linear_summary
+compare_richness_exponential_summary_10km <- richness_10km$compare_richness_exponential_summary
+# save r-squared for linear and exponential models
+compare_richness_linear_r2_10km <- richness_10km$compare_richness_linear_summary$r.squared
+compare_richness_exponential_r2_10km <- richness_10km$compare_richness_exponential_summary$r.squared
 ```
@@
 frugivore_cellRichness_25km <- richness_25km$frugivore_cellRichness
-
-compare_richness_r2_25km <- richness_25km$compare_richness_summary$r.squared
+# save linear and exponential summaries
+compare_richness_linear_summary_25km <- richness_25km$compare_richness_linear_summary
+compare_richness_exponential_summary_25km <- richness_25km$compare_richness_exponential_summary
+# save r-squared for linear and exponential models
+compare_richness_linear_r2_25km <- richness_25km$compare_richness_linear_summary$r.squared
+compare_richness_exponential_r2_25km <- richness_25km$compare_richness_exponential_summary$r.squared
 ```
@@
 frugivore_cellRichness_50km <- richness_50km$frugivore_cellRichness
-
-compare_richness_r2_50km <- richness_50km$compare_richness_summary$r.squared
+# save linear and exponential summaries
+compare_richness_linear_summary_50km <- richness_50km$compare_richness_linear_summary
+compare_richness_exponential_summary_50km <- richness_50km$compare_richness_exponential_summary
+# save r-squared for linear and exponential models
+compare_richness_linear_r2_50km <- richness_50km$compare_richness_linear_summary$r.squared
+compare_richness_exponential_r2_50km <- richness_50km$compare_richness_exponential_summary$r.squared
 ```
@@
 frugivore_cellRichness_75km <- richness_75km$frugivore_cellRichness
-
-compare_richness_r2_75km <- richness_75km$compare_richness_summary$r.squared
+# save linear and exponential summaries
+compare_richness_linear_summary_75km <- richness_75km$compare_richness_linear_summary
+compare_richness_exponential_summary_75km <- richness_75km$compare_richness_exponential_summary
+# save r-squared for linear and exponential models
+compare_richness_linear_r2_75km <- richness_75km$compare_richness_linear_summary$r.squared
+compare_richness_exponential_r2_75km <- richness_75km$compare_richness_exponential_summary$r.squared
 ```
@@
 frugivore_cellRichness_100km <- richness_100km$frugivore_cellRichness
-
-compare_richness_r2_100km <- richness_100km$compare_richness_summary$r.squared
+# save linear and exponential summaries
+compare_richness_linear_summary_100km <- richness_100km$compare_richness_linear_summary
+compare_richness_exponential_summary_100km <- richness_100km$compare_richness_exponential_summary
+# save r-squared for linear and exponential models
+compare_richness_linear_r2_100km <- richness_100km$compare_richness_linear_summary$r.squared
+compare_richness_exponential_r2_100km <- richness_100km$compare_richness_exponential_summary$r.squared
 ```

### 2024-05-16 code/L0/L0_TropicalAndes_plants_GIFT.Rmd
add date data was downloaded

@@
 overview: "This script retrieves plant trait data from the GIFT database for plant species list."
-data input: ""
-data output: ""
+data input: "none"
+data output: "TropicalAndes_GIFT_plant_traits.csv"
 date: "2023-08-03"
@@
 # Retrive GIFT data
+Data downloaded 2023-10-25
 ```{r}

### 2024-05-16 code/L0/L0_TropicalAndes_plants_TRY.Rmd
add date data was downloaded

@@
 # Load TRY data
-Data was obtained from https://www.try-db.org/TryWeb/Prop0.php 
+Data was obtained from https://www.try-db.org/TryWeb/Prop0.php on 2023-10-24
 ```{r}

### 2024-05-16 code/L0/L0_TropicalAndes_plants_BIEN_traits.Rmd
update file paths, add date data was downloaded, update data summary to use data_summary function

@@
 ```{r}
-data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
-output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+data_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+output_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
 ```
@@
 # Retrived desired trait data from BIEN
-
+Data downloaded on 2023-10-25
 ```{r}
@@
 ```{r}
-print("The number of records is")
-nrow(AllDesired_BIEN_plant_traits)
-print("The number of species is")
-length(unique(AllDesired_BIEN_plant_traits$scrubbed_species_binomial))
-print("The number of genera is" )
-length(unique(AllDesired_BIEN_plant_traits$scrubbed_genus))
-print("The number of families is")
-length(unique(AllDesired_BIEN_plant_traits$scrubbed_family))
+source("~/GitHub/neotropical_plants/code/data_summary.R")
+
+data_summary(AllDesired_BIEN_plant_traits, AllDesired_BIEN_plant_traits$scrubbed_species_binomial, AllDesired_BIEN_plant_traits$scrubbed_genus, AllDesired_BIEN_plant_traits$scrubbed_family)
 ```
@@
 ```{r}
-print("The number of records is")
-nrow(TropicalAndes_BIEN_plant_traits)
-print("The number of species is")
-length(unique(TropicalAndes_BIEN_plant_traits$scrubbed_species_binomial))
-print("The number of genera is" )
-length(unique(TropicalAndes_BIEN_plant_traits$scrubbed_genus))
-print("The number of families is")
-length(unique(TropicalAndes_BIEN_plant_traits$scrubbed_family))
+source("~/GitHub/neotropical_plants/code/data_summary.R")
+
+data_summary(AllDesired_BIEN_plant_traits, TropicalAndes_BIEN_plant_traits$scrubbed_species_binomial, TropicalAndes_BIEN_plant_traits$scrubbed_genus, TropicalAndes_BIEN_plant_traits$scrubbed_family)
 ```

### 2024-05-16 code/L0/L0_TropicalAndes_frugivores_GBIF.Rmd
update file paths, add date data was downloaded, add statement not to re-request data if file exists, update summary with custom function, add package citation and session info at end

@@
 ```{r}
-data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
-output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+data_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+output_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
 ```
@@
 ```{r}
-frugivore_species <- read.csv(file.path(data_path,"TropicalAndes_Frugivoria_frugivore_traits_species.csv"))
+frugivore_species <- read.csv(file.path(data_path_L0,"TropicalAndes_Frugivoria_frugivore_traits_species.csv"))
 ```
 
+# Download data
+Data downloaded on 2023-08-02
+```{r}
+file_name <- "0009771-231002084531237.zip"
+
+if (!file.exists(file.path(output_path_L0, file_name))) {
     # Match species names to GBIF  
-```{r}
     gbif_taxon_keys <- frugivore_species$x %>% 
@@
       pull(usageKey) 
-```
-
-# Download data
-```{r}
+    # Make GBIF request
     occ_download(
@@
     )
-```
-
-# Retrive GBIF download, save file, and load into r
-```{r}
-d <- occ_download_get("0009771-231002084531237", path = output_path, overwrite = TRUE) %>%
+    # Download GBIF data
+  d <- occ_download_get("0009771-231002084531237",
+                        path = output_path_L0, overwrite = TRUE) %>%
     occ_download_import()
+  } else {
+    cat("File already exists:", file_name, "\n")
+    d <- occ_download_get("0009771-231002084531237",
+                        path = output_path_L0, overwrite = TRUE) %>%
+                        curveocc_download_import()
+  }
 ```
@@
 # Subset occurrence to Tropical Andes IUCN Montane and Lowland forest
-
 ## Read in Forest shape
@@
 
+# Fix column type 
+```{r}
+TropicalAndes_frugivore_occ_forest$gbifID <- as.character(TropicalAndes_frugivore_occ_forest$gbifID)
+TropicalAndes_frugivore_occ_forest$geometry <- NULL
+```
+
 # Summary 
@@
 ```{r}
-print("The number of records is")
-nrow(TropicalAndes_frugivore_occ_forest)
-print("The number of species is")
-length(unique(TropicalAndes_frugivore_occ_forest$species))
-print("The number of genera is" )
-length(unique(TropicalAndes_frugivore_occ_forest$genus))
-print("The number of families is")
-length(unique(TropicalAndes_frugivore_occ_forest$family))
+source("~/GitHub/neotropical_plants/code/data_summary.R")
+
+data_summary(TropicalAndes_frugivore_occ_forest, TropicalAndes_frugivore_occ_forest$species, TropicalAndes_frugivore_occ_forest$genus, TropicalAndes_frugivore_occ_forest$family)
 ```
@@
 ```{r}
-write.csv(TropicalAndes_frugivore_occ_forest, file = file.path(output_path,"TropicalAndes_GBIF_frugivore_occ.csv"))
+write.csv(TropicalAndes_frugivore_occ_forest, file = file.path(output_path_L0,"TropicalAndes_GBIF_frugivore_occ.csv"))
 
-write.csv(GBIF_occ_SpeciesList, file = file.path(output_path,"TropicalAndes_GBIF_frugivore_occ_species.csv"))
+write.csv(GBIF_occ_SpeciesList, file = file.path(output_path_L0,"TropicalAndes_GBIF_frugivore_occ_species.csv"))
 ```
+
+# Package Citations and Session Info
+```{r}
+library(report)
+report::cite_packages()
+
+devtools::session_info()
+```

### 2024-05-16 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
remove raster code, try using sf grid instead -- not yet working

@@
 
-## Turning the FD values into a raster
-Need to subset the coordinates made at the top to the same set of the assemblages (removed values less than 5 so these coords need to be removed to do this correctly)
-
-```{r}
-# Create an empty raster with desired resolution and extent
-resolution <- c(0.75, 0.75)  # Set the desired resolution
-extent <- c(-85, 60, -24, 14)  # Set the desired extent
-empty_raster <- raster(resolution = resolution, xmn = extent[1], xmx = extent[2], ymn = extent[3], ymx = extent[4])
-```
-
+## Project FD objects
 ```{r}
@@
 ```{r}
-# Rasterize the sf data to create the FD raster
-fd_raster_frugivore <- rasterize(spatial_fdis_frugivore, empty_raster)
-writeRaster(fd_raster_frugivore$fdis_frugivore, filename = file.path(output_path,"FD_frugivore_TA.tif"), format="GTiff",overwrite=T)
-writeRaster(fd_raster_frugivore$ID, filename = file.path(output_path,"FD_frugivore_TA_ID.tif"), format="GTiff",overwrite=T)
-```
-
-## Project FD objects
-```{r}
 # set crs of sf objects
@@
 ```{r}
-# determine resolution of spatial_fdis_frugivore_projected 
-st_crs(spatial_fdis_frugivore_projected, parameters = TRUE)$units_gdal
+create_fdis_plots <- function(resolution_meters, data, variable_name) {
+# Make Grid
+TAGrid <- TApoly %>%
+  st_make_grid(cellsize = c(resolution_meters)) %>%
+  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+  st_cast("MULTIPOLYGON") %>%
+  st_sf() %>%
+  mutate(cellid = row_number())
+
+  # Calculate functional dispersion for grid
+  fdis_grid <- TAGrid %>%
+    st_join(data) %>%
+    group_by(cellid) %>%
+    summarize(sum_variable = sum(variable_name, na.rm = TRUE))
+
+  # Make plot
+  FDplot <-
+    ggplot() +
+    geom_sf(data = Americas, fill = "white")+
+    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
+    geom_sf(data = fdis_grid, aes(fill = fdis_frugivore)) +
+    scale_fill_gradient(low = "blue", high = "red") +
+    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+    theme(panel.background = element_rect(fill = "lightblue")) +
+    annotation_scale(location = "bl", width_hint = 0.5) +
+    labs(x = "Longitude", y = "Latitude", fill = "FDis")
+
+  # Return a list of data
+  list(FDplot = FDplot)
+}
 ```
@@
 ```{r}
-# Raster to raster point
-spec_FD_raster_frugivore_point <- rasterToPoints(fd_raster_frugivore, spatial = TRUE)
-# Convert to dataframe
-fd_raster_frugivore_point <- data.frame(spec_FD_raster_frugivore_point)
-```
-
-```{r}
-frugivoregridFDTA <-
-  ggplot() +
-  geom_sf(data = Americas, fill = "white")+
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_tile(data= fd_raster_frugivore_point, aes(x=x, y=y, fill=fdis_frugivore)) +
-  scale_fill_scico(palette = "romaO", direction = 1) +
-  coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  ) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  labs(title = "Tropical Andes Frugivore Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
-frugivoregridFDTA
+create_fdis_plots(100000, spatial_fdis_frugivore_projected, fdis_frugivore)
 ```
@@
 # Plants
+Test with small random sample as a subset to make it work before using all data
+```{r}
+all_plant_species <- row.names(plant_traits_df_final)
+plant_random_subset <- sample(all_plant_species, 1000)
+# objects that need to be subset
+PAM_plant_site_final #matrix
+plant_traits_df_final
+```
+
 ## Create trait type table
@@
 ncol(subset_matrix_plant) #82
+```
 
+```{r}
 sp_faxes_coord_plants_sub_names <- row.names(sp_faxes_coord_plants_sub)

### 2024-05-16 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
remove scatterplot from plotting function, work in progress fitting exponential for 100km -- still needs work

--- 2024-05-14 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
@@
                                 filter(Plant_Richness > 0 & Frugivore_Richness > 0)
-  # Compare plant and frugivore richness
-  ## Fit a linear model
-  compare_richness_linear_summary <- summary(lm(comparison_data_filtered$Frugivore_Richness ~
-                                                  comparison_data_filtered$Plant_Richness))
-  linear_r_squared <- compare_richness_linear_summary$r.squared
-
-  ## Fit an exponential model
-  compare_richness_exponential_summary <- summary(lm(exp(comparison_data_filtered$Frugivore_Richness) ~
-                                                       comparison_data_filtered$Plant_Richness))
-  exponential_r_squared <- compare_richness_exponential_summary$r.squared
-
-  ## Create scatter plot with linear and exponential lines
-  plant_frug_richness_compare <- ggplot(comparison_data,
-                                        aes(x = Plant_Richness, y = Frugivore_Richness)) +
-  geom_point() +
-  labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
-       subtitle =  paste("Linear (Red) R-squared = ", round(linear_r_squared, 2), "\n",
-                  "Exponential (Blue) R-squared = ", round(exponential_r_squared, 2)),
-       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
-    geom_smooth(method = "lm", se = FALSE, color = "red") +
-    geom_smooth(method = "lm", se = FALSE, color = "blue", formula = y ~ exp(x), linetype = "dashed")
 
@@
                                   xlab = "Frugivore Richness by cell")
+  # Return a list of data
+  list(plantgridRichTA = plant_plot,
+       frugivoregridRichTA = frugivore_plot,
+       plant_cellRichness = plant_richness_grid,
+       frugivore_cellRichness = frugivore_richness_grid, 
+       comparison_data_filtered = comparison_data_filtered,
+       plant_richness_hist = plant_richness_hist,
+       frugivore_richness_hist = frugivore_richness_hist)
 }
@@
 
-compare_richness_r2_1km <- richness_1km$compare_richness_summary$r.squared
+comparison_data_filtered_1km <- richness_1km$comparison_data_filtered
+```
+
+```{r}
+
 ```
@@
 
-compare_richness_r2_5km <- richness_5km$compare_richness_summary$r.squared
+comparison_data_filtered_5km <- richness_5km$comparison_data_filtered
+```
+
+```{r}
+
 ```
@@
 
-compare_richness_r2_10km <- richness_10km$compare_richness_summary$r.squared
+comparison_data_filtered_10km <- richness_10km$comparison_data_filtered
+```
+
+```{r}
+
 ```
@@
 
-compare_richness_r2_25km <- richness_25km$compare_richness_summary$r.squared
+comparison_data_filtered_25km <- richness_25km$comparison_data_filtered
+```
+
+```{r}
+
 ```
@@
 
-compare_richness_r2_50km <- richness_50km$compare_richness_summary$r.squared
+comparison_data_filtered_50km <- richness_50km$comparison_data_filtered
+```
+
+```{r}
+
 ```
@@
 
-compare_richness_r2_75km <- richness_75km$compare_richness_summary$r.squared
+comparison_data_filtered_75km <- richness_75km$comparison_data_filtered
+```
+
+```{r}
+
 ```
@@
 
-compare_richness_r2_100km <- richness_100km$compare_richness_summary$r.squared
+comparison_data_filtered_100km <- richness_100km$comparison_data_filtered
+```
+
+```{r}
+
+```
+
+```{r}
+# Compare plant and frugivore richness
+  ## Fit a linear model
+  compare_richness_linear_summary_100km <- summary(lm(comparison_data_filtered_100km$Frugivore_Richness ~
+                                                  comparison_data_filtered_100km$Plant_Richness))
+  linear_r_squared_100km <- compare_richness_linear_summary_100km$r.squared
+
+  ## Fit an exponential model
+  compare_richness_exponential_summary_100km <- summary(lm(comparison_data_filtered_100km$Frugivore_Richness ~
+                                                       log(comparison_data_filtered_100km$Plant_Richness)))
+  exponential_r_squared_100km <- compare_richness_exponential_summary_100km$r.squared
+
+  ## Create scatter plot with linear and exponential lines
+  compare_richness_plot_100km <- ggplot(comparison_data_filtered_100km,
+                                        aes(x = Plant_Richness, y = Frugivore_Richness)) +
+  geom_point() +
+  labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
+       subtitle =  paste("Linear (Red) R-squared = ", round(linear_r_squared, 2), "\n",
+                  "Exponential (Blue) R-squared = ", round(exponential_r_squared, 2)),
+       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
+  geom_smooth(method = "lm", se = FALSE, color = "red")+
+  geom_smooth(method = "lm", se = FALSE, color = "blue", formula = y ~ log(x)) +
+  coord_cartesian(xlim = c(0, NA), ylim = c(0, NA))
+```
+
+```{r}
+# Fit the nonlinear model using nls
+nls_model <- nls(Frugivore_Richness ~ b * exp(c * Plant_Richness), 
+                 data = comparison_data_filtered_100km, 
+                 start = list(b = 200, c = -0.00001),
+                 algorithm = "port",
+                 control = nls.control(maxiter = 500, minFactor = 1e-10, warnOnly = TRUE))
+
+# Summary of the model
+summary(nls_model)
 ```

### 2024-05-16 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
move some code to L2_3_TropicalAndes_divRelationships

@@
 
-  # Extract cell values
-  comparison_data <- data.frame(Plant_Richness = plant_richness_grid$num_species,
-                               Frugivore_Richness = frugivore_richness_grid$num_species)
-
-  # Filter out rows where either plant or frugivore richness is zero
-  comparison_data_filtered <- comparison_data %>%
-                                filter(Plant_Richness > 0 & Frugivore_Richness > 0)
-
   # Create histograms
@@
        frugivore_cellRichness = frugivore_richness_grid, 
-       comparison_data_filtered = comparison_data_filtered,
        plant_richness_hist = plant_richness_hist,
@@
 frugivore_cellRichness_1km <- richness_1km$frugivore_cellRichness
-
-comparison_data_filtered_1km <- richness_1km$comparison_data_filtered
-```
-
-```{r}
-
 ```
@@
 frugivore_cellRichness_5km <- richness_5km$frugivore_cellRichness
-
-comparison_data_filtered_5km <- richness_5km$comparison_data_filtered
-```
-
-```{r}
-
 ```
@@
 frugivore_cellRichness_10km <- richness_10km$frugivore_cellRichness
-
-comparison_data_filtered_10km <- richness_10km$comparison_data_filtered
-```
-
-```{r}
-
 ```
@@
 frugivore_cellRichness_25km <- richness_25km$frugivore_cellRichness
-
-comparison_data_filtered_25km <- richness_25km$comparison_data_filtered
 ```
@@
 frugivore_cellRichness_50km <- richness_50km$frugivore_cellRichness
-
-comparison_data_filtered_50km <- richness_50km$comparison_data_filtered
-```
-
-```{r}
-
 ```
@@
 frugivore_cellRichness_75km <- richness_75km$frugivore_cellRichness
-
-comparison_data_filtered_75km <- richness_75km$comparison_data_filtered
-```
-
-```{r}
-
 ```
@@
 frugivore_cellRichness_100km <- richness_100km$frugivore_cellRichness
-
-comparison_data_filtered_100km <- richness_100km$comparison_data_filtered
-```
-
-```{r}
-
-```
-
-```{r}
-# Compare plant and frugivore richness
-  ## Fit a linear model
-  compare_richness_linear_summary_100km <- summary(lm(comparison_data_filtered_100km$Frugivore_Richness ~
-                                                  comparison_data_filtered_100km$Plant_Richness))
-  linear_r_squared_100km <- compare_richness_linear_summary_100km$r.squared
-
-  ## Fit an exponential model
-  compare_richness_exponential_summary_100km <- summary(lm(comparison_data_filtered_100km$Frugivore_Richness ~
-                                                       log(comparison_data_filtered_100km$Plant_Richness)))
-  exponential_r_squared_100km <- compare_richness_exponential_summary_100km$r.squared
-
-  ## Create scatter plot with linear and exponential lines
-  compare_richness_plot_100km <- ggplot(comparison_data_filtered_100km,
-                                        aes(x = Plant_Richness, y = Frugivore_Richness)) +
-  geom_point() +
-  labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
-       subtitle =  paste("Linear (Red) R-squared = ", round(linear_r_squared, 2), "\n",
-                  "Exponential (Blue) R-squared = ", round(exponential_r_squared, 2)),
-       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
-  geom_smooth(method = "lm", se = FALSE, color = "red")+
-  geom_smooth(method = "lm", se = FALSE, color = "blue", formula = y ~ log(x)) +
-  coord_cartesian(xlim = c(0, NA), ylim = c(0, NA))
-```
-
-```{r}
-# Fit the nonlinear model using nls
-nls_model <- nls(Frugivore_Richness ~ b * exp(c * Plant_Richness), 
-                 data = comparison_data_filtered_100km, 
-                 start = list(b = 200, c = -0.00001),
-                 algorithm = "port",
-                 control = nls.control(maxiter = 500, minFactor = 1e-10, warnOnly = TRUE))
-
-# Summary of the model
-summary(nls_model)
 ```
@@
 ggsave("all_richness_plots.pdf", all_richness_plots, path = output_path, width = 20, height = 8.5, units = "in")
-```
-
-# Compare r-squared by resolution
-Plot
-```{r}
-# create dataframe of resolutions and r-squared values 
-#[ADD 1, 5,]
-resolution <- c( 10, 25, 50, 75, 100)
-#ADD compare_richness_r2_1km, compare_richness_r2_5km,
-r_squared <- c(compare_richness_r2_10km, compare_richness_r2_25km, compare_richness_r2_50km, compare_richness_r2_75km, compare_richness_r2_100km)
-r_squared_compare <- data.frame(resolution, r_squared)
-
-compare_r2_by_res <- ggplot(r_squared_compare, aes(x=resolution, y=r_squared)) +
-    geom_point()+
-    labs(title = "R-squared values of plant-frugivore richness relationships by resolution",
-         x = "Resolution (km)", y = "R-squared", fill = "Number of species")
-print(compare_r2_by_res)
 ```

### 2024-05-16 code/L2/L2_3_TropicalAndes_divRelationships.Rmd
add code from L2_1_TropicalAndes_taxdiv

@@
 ```
+
+```{r}
+  # Extract cell values
+  comparison_data <- data.frame(Plant_Richness = plant_richness_grid$num_species,
+                               Frugivore_Richness = frugivore_richness_grid$num_species)
+
+  # Filter out rows where either plant or frugivore richness is zero
+  comparison_data_filtered <- comparison_data %>%
+                                filter(Plant_Richness > 0 & Frugivore_Richness > 0)
+```
+
+```{r}
+# Compare plant and frugivore richness
+  ## Fit a linear model
+  compare_richness_linear_summary_100km <- summary(lm(comparison_data_filtered_100km$Frugivore_Richness ~
+                                                  comparison_data_filtered_100km$Plant_Richness))
+  linear_r_squared_100km <- compare_richness_linear_summary_100km$r.squared
+
+  ## Fit an exponential model
+  compare_richness_exponential_summary_100km <- summary(lm(comparison_data_filtered_100km$Frugivore_Richness ~
+                                                       log(comparison_data_filtered_100km$Plant_Richness)))
+  exponential_r_squared_100km <- compare_richness_exponential_summary_100km$r.squared
+
+  ## Create scatter plot with linear and exponential lines
+  compare_richness_plot_100km <- ggplot(comparison_data_filtered_100km,
+                                        aes(x = Plant_Richness, y = Frugivore_Richness)) +
+  geom_point() +
+  labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
+       subtitle =  paste("Linear (Red) R-squared = ", round(linear_r_squared, 2), "\n",
+                  "Exponential (Blue) R-squared = ", round(exponential_r_squared, 2)),
+       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
+  geom_smooth(method = "lm", se = FALSE, color = "red")+
+  geom_smooth(method = "lm", se = FALSE, color = "blue", formula = y ~ log(x)) +
+  coord_cartesian(xlim = c(0, NA), ylim = c(0, NA))
+```
+
+```{r}
+# Fit the nonlinear model using nls
+nls_model <- nls(Frugivore_Richness ~ b * exp(c * Plant_Richness), 
+                 data = comparison_data_filtered_100km, 
+                 start = list(b = 200, c = -0.00001),
+                 algorithm = "port",
+                 control = nls.control(maxiter = 500, minFactor = 1e-10, warnOnly = TRUE))
+
+# Summary of the model
+summary(nls_model)
+```
+
+# Compare r-squared by resolution
+Plot
+```{r}
+# create dataframe of resolutions and r-squared values 
+#[ADD 1, 5,]
+resolution <- c( 10, 25, 50, 75, 100)
+#ADD compare_richness_r2_1km, compare_richness_r2_5km,
+r_squared <- c(compare_richness_r2_10km, compare_richness_r2_25km, compare_richness_r2_50km, compare_richness_r2_75km, compare_richness_r2_100km)
+r_squared_compare <- data.frame(resolution, r_squared)
+
+compare_r2_by_res <- ggplot(r_squared_compare, aes(x=resolution, y=r_squared)) +
+    geom_point()+
+    labs(title = "R-squared values of plant-frugivore richness relationships by resolution",
+         x = "Resolution (km)", y = "R-squared", fill = "Number of species")
+print(compare_r2_by_res)
+```

### 2024-05-16 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
remove unused packages, update plotting function to match object names

@@
 library(rnaturalearth)
-library(smoothr)
-library(purrr)
-library(raster)
 library(ggspatial)
@@
 output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
+figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
 ```
@@
   # Create histograms
-  plant_richness_hist <- hist(comparison_data$Plant_Richness, 
+  plant_richness_hist <- hist(plant_richness_grid$num_species, 
     main = paste("Histogram of Plant Richness [", resolution_meters/1000, " km] by cell", sep = ""),
                               xlab = "Plant Richness by cell")
-  frugivore_richness_hist <- hist(comparison_data$Frugivore_Richness,
+  frugivore_richness_hist <- hist(frugivore_richness_grid$num_species,
     main = paste("Histogram of Frugivore Richness [", resolution_meters/1000, " km] by cell", sep = ""),
@@
 
-```{r}
-
-```
-
 ## 50 km
@@
 all_richness_plots <- ggpubr::ggarrange(#plantgridRichTA_1km, plantgridRichTA_5km, 
-                                              plantgridRichTA_10km, plantgridRichTA_25km,
+                                              #plantgridRichTA_10km, 
+                                              plantgridRichTA_25km,
                                               plantgridRichTA_50km, plantgridRichTA_75km,
@@
                                               #frugivoregridRichTA_1km, frugivoregridRichTA_5km, 
-                                              frugivoregridRichTA_10km, frugivoregridRichTA_25km,
+                                              #frugivoregridRichTA_10km, 
+                                              frugivoregridRichTA_25km,
                                               frugivoregridRichTA_50km, frugivoregridRichTA_75km,
                                               frugivoregridRichTA_100km, 
-                                              ncol = 5, nrow = 2,
-                                              labels = c("10km", "25km", ))
+                                              ncol = 4, nrow = 2)
 all_richness_plots
-ggsave("all_richness_plots.pdf", all_richness_plots, path = output_path, width = 20, height = 8.5, units = "in")
+ggsave("all_richness_plots.png", all_richness_plots, path = output_path, width = 20, height = 8.5, units = "in")
 ```

### 2024-05-16 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
remove unused packages, add figure path, add code to map without projection -- still doesn't work

@@
 ```{r}
+library(mFD)
 library(sf)
@@
 library(rnaturalearth)
-library(smoothr)
-library(purrr)
-library(raster)
 library(ggspatial)
-library(vegan)
+library(rlang)
 ```
@@
 
+# Mapping without projection 
+Because projecting is being difficult
+```{r}
+TropicalAndes_IUCNHabitat_Forest_latlong <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
+Americas_latlong <- ne_countries(continent = c("North America", "South America"), returnclass = "sf")
+#polygon of Tropical Andes
+TApoly_latlong <- Americas_latlong %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
+```
+
+```{r}
+# Generate coordinates
+subset_coords_frugivore <- site_loc_key_frugivore[rowSums(PAM_frugivore_site_final) >= 5, ]
+subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]
+
+frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)
+```
+
+```{r}
+# Convert the dataframe to sf format
+spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
+```
+
+```{r}
+# set crs of sf objects
+spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(4326)
+```
+
+```{r}
+# Create a grid over the area
+create_grid <- function(data, cellsize) {
+  st_make_grid(st_as_sfc(st_bbox(data)), cellsize = cellsize) %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+}
+
+# Aggregate functional dispersion within each grid cell
+aggregate_to_grid <- function(grid, points, value_col) {
+  points %>%
+    st_join(grid) %>%
+    group_by(cellid) %>%
+    summarize(mean_value = mean(get(value_col), na.rm = TRUE)) %>%
+    st_set_geometry("geometry")
+}
+
+# Example usage with your data
+resolution_meters <- 100000
+variable_name <- "fdis_frugivore"
+
+# Create grid
+grid <- create_grid(spatial_fdis_frugivore, cellsize = resolution_meters)
+
+# Aggregate data to grid
+aggregated_data <- aggregate_to_grid(grid, spatial_fdis_frugivore, variable_name)
+
+```
+
+```{r}
+FDisplot <-
+    ggplot() +
+    geom_sf(data = Americas_latlong, fill = "white")+
+    geom_sf(data = TApoly_latlong, fill = "lightgrey", size = 0.1) +
+    geom_point(data = spatial_fdis_frugivore , aes(fill = fdis_frugivore)) +
+    scale_fill_gradient(low = "blue", high = "red") +
+    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+    theme(panel.background = element_rect(fill = "lightblue")) +
+    annotation_scale(location = "bl", width_hint = 0.5) +
+    labs(x = "Longitude", y = "Latitude", fill = "FDis")
+FDisplot
+```
+
 ## Project FD objects
@@
     group_by(cellid) %>%
-    summarize(sum_variable = sum(variable_name, na.rm = TRUE))
+    summarize(
+      sum_variable = sum(variable_name, na.rm = TRUE),
+      mean_variable = mean(variable_name, na.rm = TRUE)
+    )
 
@@
   # Return a list of data
-  list(FDplot = FDplot)
+  list(fdis_grid = fdis_grid,
+        FDplot = FDplot)
 }
@@
 ```{r}
+set.seed(100)
 all_plant_species <- row.names(plant_traits_df_final)
@@
 # objects that need to be subset
-PAM_plant_site_final #matrix
+PAM_plant_site_final_df <- as.dataframe(PAM_plant_site_final) #matrix
 plant_traits_df_final

### 2024-05-16 code/L2/L2_3_TropicalAndes_divRelationships.Rmd
add libraries, add data inputs from taxdiv outputs

@@
 ```{r}
-
+library(dplyr)
+library(ggplot2)
+library(smoothr)
+library(purrr)
 ```
@@
 ```{r}
+plant_cellRichness_1km <- read.csv(file.path(output_path,"TropicalAndes_plantRichness_1km.csv"))
+frugivore_cellRichness_1km <- read.csv(file.path(output_path,"TropicalAndes_frugivoreRichness_1km.csv"))
 
+plant_cellRichness_5km <- read.csv(file.path(output_path,"TropicalAndes_plantRichness_5km.csv"))
+frugivore_cellRichness_5km <- read.csv(file.path(output_path,"TropicalAndes_frugivoreRichness_5km.csv"))
+
+plant_cellRichness_10km <- read.csv(file.path(output_path,"TropicalAndes_plantRichness_10km.csv"))
+frugivore_cellRichness_10km <- read.csv(file.path(output_path,"TropicalAndes_frugivoreRichness_10km.csv"))
+
+plant_cellRichness_25km <- read.csv(file.path(output_path,"TropicalAndes_plantRichness_25km.csv"))
+frugivore_cellRichness_25km <- read.csv(file.path(output_path,"TropicalAndes_frugivoreRichness_25km.csv"))
+
+plant_cellRichness_50km <- read.csv(file.path(output_path,"TropicalAndes_plantRichness_50km.csv"))
+frugivore_cellRichness_50km <- read.csv(file.path(output_path,"TropicalAndes_frugivoreRichness_50km.csv"))
+
+plant_cellRichness_75km <- read.csv(file.path(output_path,"TropicalAndes_plantRichness_75km.csv"))
+frugivore_cellRichness_75km <- read.csv(file.path(output_path,"TropicalAndes_frugivoreRichness_75km.csv"))
+
+plant_cellRichness_100km <- read.csv(file.path(output_path,"TropicalAndes_plantRichness_100km.csv"))
+frugivore_cellRichness_100km <- read.csv(file.path(output_path,"TropicalAndes_frugivoreRichness_100km.csv"))
 ```
 
+# 1 km
+# 5 km
+# 10 km
+# 25 km
+# 50 km
+# 75 km
+# 100 km
 ```{r}
   # Extract cell values
-  comparison_data <- data.frame(Plant_Richness = plant_richness_grid$num_species,
-                               Frugivore_Richness = frugivore_richness_grid$num_species)
+  comparison_data_100km <- data.frame(Plant_Richness = plant_richness_grid_100km$num_species,
+                               Frugivore_Richness = frugivore_richness_grid_100km$num_species)
 
   # Filter out rows where either plant or frugivore richness is zero
-  comparison_data_filtered <- comparison_data %>%
+  comparison_data_filtered_100km <- comparison_data %>%
                                 filter(Plant_Richness > 0 & Frugivore_Richness > 0)
@@
 # Compare r-squared by resolution
-Plot
+## Taxonomic diversity
 ```{r}
@@
 ```
+## Functional Diversity
+```{r}
+
+```

### 2024-05-16 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
fine tunning multipannel plot for maps

@@
 # Multi panel plots
-Richness maps
+## Richness maps
 ```{r}
-all_richness_plots <- ggpubr::ggarrange(#plantgridRichTA_1km, plantgridRichTA_5km, 
+all_plant_richness_plots <- ggpubr::ggarrange(#plantgridRichTA_1km, plantgridRichTA_5km, 
                                               #plantgridRichTA_10km, 
@@
                                               plantgridRichTA_100km,
-                                              #frugivoregridRichTA_1km, frugivoregridRichTA_5km, 
+                                              ncol = 4, nrow = 1,
+                                              labels = c("25 km", "50 km", "75 km", "100 km"),
+                                              common.legend = TRUE, legend = "right")
+all_plant_richness_plots_labeled <- ggpubr::annotate_figure(all_plant_richness_plots,
+                                                            left = ggpubr::text_grob("Plants", face = "bold", size = 14))
+all_plant_richness_plots_labeled
+```
+
+```{r}
+all_frugivore_richness_plots <- ggpubr::ggarrange(#frugivoregridRichTA_1km, frugivoregridRichTA_5km, 
                                               #frugivoregridRichTA_10km, 
@@
                                               frugivoregridRichTA_100km,
-                                              ncol = 4, nrow = 2)
+                                              ncol = 4, nrow = 1,
+                                              labels = c("25 km", "50 km", "75 km", "100 km"),
+                                              common.legend = TRUE, legend = "right")
+all_frugivore_richness_plots_labeled <- ggpubr::annotate_figure(all_frugivore_richness_plots,
+                                                            left = ggpubr::text_grob("Frugivores", face = "bold", size = 14))
+all_frugivore_richness_plots_labeled
+```
+
+```{r}
+all_richness_plots <- ggpubr::ggarrange(all_plant_richness_plots_labeled,
+                                        all_frugivore_richness_plots_labeled,
+                                        ncol = 1, nrow = 2)
 all_richness_plots
-ggsave("all_richness_plots.png", all_richness_plots, path = output_path, width = 20, height = 8.5, units = "in")
+ggsave("all_richness_plots.png", all_richness_plots, path = figure_path, width = 11, height = 8, units = "in")
 ```

### 2024-05-22 code/L1/L1_P3_TropicalAndes_plant_traits.Rmd
add summary code to get number of traits and species coverage for each, still need to standardize fruit type and plant growth form

@@
 ---
-title: "TITLE"
+title: "Plant trait cleaning and combining"
 author: "Hazel J. Anderson"
@@
 
+```{r}
+TRY_traits %>%
+  count(TraitName)
+```
+
+```{r}
+TRY_traits %>%
+  distinct(Accepted_species, TraitName) %>%
+  count(TraitName)
+```
+
 ## BIEN
@@
 nrow(BIEN_filtered)
+```
+
+```{r}
+BIEN_traits %>%
+  count(trait_name)
+```
+
+```{r}
+BIEN_traits %>%
+  distinct(Accepted_species, trait_name) %>%
+  count(trait_name)
 ```
@@
 
+```{r}
+GIFT_traits %>%
+  count(trait_name)
+```
+
+```{r}
+GIFT_traits %>%
+  distinct(Accepted_species, trait_name) %>%
+  count(trait_name)
+```
+
 # Combine trait information from TRY, BIEN, GIFT
@@
   distinct(DataName)
+```
+
+```{r}
+unique(TRYTraitTypes$TraitName)
 ```
@@
 ```{r}
+# save sources/references
+TRY_references <- unique(TRY_filtered[ , c("Dataset", "Reference")])
+write.csv(TRY_references, file.path(output_path,"TRY_references.csv"))
+```
+
+```{r}
+unique(BIEN_filtered$trait_name)
+```
+
+```{r}
 colnames(BIEN_filtered)
@@
 colnames(BIEN_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue", "Unit")
+```
+
+```{r}
+# save sources/references
+BIEN_references <- unique(BIEN_filtered[ , c("url_source", "source_citation")])
+write.csv(BIEN_references, file.path(output_path,"BIEN_references.csv"))
+```
+
+```{r}
+unique(GIFT_filtered$trait_name)
 ```
@@
 ```
+
+```{r}
+traits %>%
+  distinct(Accepted_species, TraitName) %>%
+  count(TraitName)
+```
+
+```{r}
+# Group by TraitName and summarize the unique TraitValues
+unique_trait_values <- traits %>%
+  group_by(TraitName) %>%
+  summarize(UniqueValues = list(unique(TraitValue)))%>%
+  unnest(UniqueValues)
+```
+
+Final trait list:
+- dispersal syndrome
+- Fruit mass
+- Fruit length
+- Fruit type
+- Plant growth form
+- Plant height
+- Plant Lifespan
+- Seed mass
+- Seed length
+- Seed width
 
@@
 traits_dispersal$OrigTraitValue <- traits_dispersal$TraitValue
+nrow(traits_dispersal)
+length(unique(traits_dispersal$Accepted_species))
 ```
@@
 length(unique(traits_dispersal$Accepted_species))
+```
+
+```{r}
+# remove rows where TraitValue is numeric
+traits_dispersal <- traits_dispersal %>%
+  filter(is.na(as.numeric(TraitValue)))
+nrow(traits_dispersal)
+length(unique(traits_dispersal$Accepted_species))
+```
+
+```{r}
+traits_dispersal %>%
+  distinct(TraitValue)
 ```
@@
 ```{r}
-print(unique(traits_dispersal$TraitValue))
+traits_dispersal %>%
+  distinct(TraitValue)
 ```
@@
 ```{r}
+# standardize trait names
 traits_dispersal <- traits_dispersal %>%
-  mutate(TraitName = "dispersal_syndrome")
+  mutate(TraitName = "DispersalSyndrome")
 ```
@@
 ```{r}
-averagetraits_dispersal <- factor_data_merge(traits_dispersal, "Dispersal Syndrome")
+averagetraits_dispersal <- factor_data_merge(traits_dispersal, "DispersalSyndrome")
+```
+
+```{r}
+nrow(averagetraits_dispersal)
+length(unique(averagetraits_dispersal$Accepted_species))
+```
+
+```{r}
+averagetraits_dispersal %>%
+  count(TraitValue)
+```
+
+```{r}
+summary(averagetraits_dispersal)
 ```
@@
 traits_fruitMass$OrigTraitValue <- traits_fruitMass$TraitValue
+nrow(traits_fruitMass)
+length(unique(traits_fruitMass$Accepted_species))
 ```
@@
 ```{r}
+# remove non-numeric rows
+traits_fruitMass <- traits_fruitMass %>%
+  filter(!is.na(as.numeric(TraitValue))) %>%
+  mutate(TraitValue = as.numeric(TraitValue))
+
+nrow(traits_fruitMass)
+length(unique(traits_fruitMass$Accepted_species))
+```
+
+```{r}
+# check that units match
 print(unique(traits_fruitMass$Unit))
@@
 ```{r}
-averagetraits_fruitMass <- numeric_data_average(traits_fruitMass, "mg", "Fruit Mass")
+# standardize trait names with units
+traits_fruitMass <- traits_fruitMass %>%
+  mutate(TraitName = "FruitMass_mg")
+```
+
+```{r}
+averagetraits_fruitMass <- numeric_data_average(traits_fruitMass, "mg", "FruitMass_mg")
+```
+
+```{r}
+nrow(averagetraits_fruitMass)
+length(unique(averagetraits_fruitMass$Accepted_species))
+```
+
+```{r}
+summary(averagetraits_fruitMass)
 ```
@@
 traits_fruitLength$OrigTraitValue <- traits_fruitLength$TraitValue
-```
 
-```{r}
-print(unique(traits_fruitLength$Unit))
+nrow(traits_fruitLength)
+length(unique(traits_fruitLength$Accepted_species))
 ```
@@
   mutate(TraitValue = as.numeric(TraitValue))
+
+nrow(traits_fruitLength)
+length(unique(traits_fruitLength$Accepted_species))
+```
+
+```{r}
+# check that units match
+print(unique(traits_fruitLength$Unit))
 ```
@@
 ```{r}
-averagetraits_fruitLength <- numeric_data_average(traits_fruitLength, "mm", "Fruit Length")
+# standardize trait names with units
+traits_fruitLength <- traits_fruitLength %>%
+  mutate(TraitName = "FruitLength_mm")
+```
+
+```{r}
+averagetraits_fruitLength <- numeric_data_average(traits_fruitLength, "mm", "FruitLength_mm")
+```
+
+```{r}
+nrow(averagetraits_fruitLength)
+length(unique(averagetraits_fruitLength$Accepted_species))
+```
+
+```{r}
+summary(averagetraits_fruitLength)
 ```
@@
 traits_plantHeight$OrigTraitValue <- traits_plantHeight$TraitValue
+
+nrow(traits_plantHeight)
+length(unique(traits_plantHeight$Accepted_species))
+```
+
+```{r}
+# remove non-numeric rows
+traits_plantHeight <- traits_plantHeight %>%
+  filter(!is.na(as.numeric(TraitValue))) %>%
+  mutate(TraitValue = as.numeric(TraitValue))
+
+nrow(traits_plantHeight)
+length(unique(traits_plantHeight$Accepted_species))
 ```
@@
 ```{r}
-averagetraits_plantHeight <- numeric_data_average(traits_plantHeight, "m", "Plant Height")
+# standardize trait names with units
+traits_plantHeight <- traits_plantHeight %>%
+  mutate(TraitName = "PlantHeight_m")
+```
+
+```{r}
+averagetraits_plantHeight <- numeric_data_average(traits_plantHeight, "m", "PlantHeight_m")
+```
+
+```{r}
+nrow(averagetraits_plantHeight)
+length(unique(averagetraits_plantHeight$Accepted_species))
+```
+
+```{r}
+summary(averagetraits_plantHeight)
 ```
@@
 traits_plantLifespan <- traits %>%
-  filter(grepl('Plant lifespan (longevity)|longest whole plant longevity|maximum whole plant longevity|plant_lifespan', TraitName))
+  filter(grepl('Plant lifespan \\(longevity\\)|longest whole plant longevity|maximum whole plant longevity|plant_lifespan', TraitName))
 traits_plantLifespan$OrigTraitValue <- traits_plantLifespan$TraitValue
+
+nrow(traits_plantLifespan)
+length(unique(traits_plantLifespan$Accepted_species))
 ```
@@
 ```{r}
+# remove rows if TraitValue is non-numeric
+traits_plantLifespan <- traits_plantLifespan %>%
+  filter(!is.na(as.numeric(TraitValue))) %>%
+  mutate(TraitValue = as.numeric(TraitValue))
+
+nrow(traits_plantLifespan)
+length(unique(traits_plantLifespan$Accepted_species))
+```
+
+```{r}
+# check that units match
+print(unique(traits_plantLifespan$Unit))
+traits_plantLifespan$Unit <- "years"
 print(unique(traits_plantLifespan$Unit))
@@
 ```{r}
-averagetraits_plantLifespan <- numeric_data_average(traits_plantLifespan, "years", "Plant Lifespan")
+# standardize trait names with units
+traits_plantLifespan <- traits_plantLifespan %>%
+  mutate(TraitName = "PlantLifespan_years")
+
+```
+
+```{r}
+averagetraits_plantLifespan <- numeric_data_average(traits_plantLifespan, "years", "PlantLifespan_years")
+```
+
+```{r}
+nrow(averagetraits_plantLifespan)
+length(unique(averagetraits_plantLifespan$Accepted_species))
+```
+
+```{r}
+summary(averagetraits_plantLifespan)
 ```
@@
 traits_fruitType <- traits %>%
-  filter(grepl('Fruit dehiscence type|Fruit surface type|Fruit texture inside (fleshy, leathrey, …)|Fruit type|fruit type|fruit_dryness|fruit_type', TraitName))
+  filter(grepl('Fruit type|fruit type|fruit_type', TraitName))
 traits_fruitType$OrigTraitValue <- traits_fruitType$TraitValue
+
+nrow(traits_fruitType)
+length(unique(traits_fruitType$Accepted_species))
 ```
@@
 ```{r}
-print(unique(traits_fruitType$Unit))
+traits_fruitType <- traits_fruitType %>%
+  filter(is.na(as.numeric(TraitValue)))
+
+nrow(traits_fruitType)
+length(unique(traits_fruitType$Accepted_species))
+```
+
+```{r}
+traits_fruitType %>%
+  distinct(TraitValue)
+```
+
+```{r}
+# standardize trait values
+traits_fruitType <- traits_fruitType %>%
+  filter() %>%
+  mutate(TraitValue= case_when(
+    grepl('drupe|ectozoochorous|zoochorous', OrigTraitValue, ignore.case = TRUE) ~ "zoochorous",
+    grepl('capsule|Capsule|Capsule |', OrigTraitValue, ignore.case = TRUE) ~ "myrmecochorous",
+  )) %>%
+  filter(!is.na(TraitValue))
+```
+
+```{r}
+# standardize trait name
+traits_fruitType <- traits_fruitType %>%
+  mutate(TraitName = "FruitType")
 ```
@@
 averagetraits_fruitType <- factor_data_merge(traits_fruitType, "Fruit Type")
+```
+
+```{r}
+nrow(averagetraits_fruitType)
+length(unique(averagetraits_fruitType$Accepted_species))
+```
+
+```{r}
+summary(averagetraits_fruitType)
 ```
@@
 traits_seedMass$OrigTraitValue <- traits_seedMass$TraitValue
+
+nrow(traits_seedMass)
+length(unique(traits_seedMass$Accepted_species))
+```
+
+```{r}
+# remove non-numeric rows
+traits_seedMass <- traits_seedMass %>%
+  filter(!is.na(as.numeric(TraitValue))) %>%
+  mutate(TraitValue = as.numeric(TraitValue))
+
+nrow(traits_seedMass)
+length(unique(traits_seedMass$Accepted_species))
 ```
@@
 print(unique(traits_seedMass$Unit))
-```
-
-```{r}
-# remove rows if TraitValue is non-numeric
-traits_seedMass <- traits_seedMass %>%
-  filter(!is.na(as.numeric(TraitValue))) %>%
-  mutate(TraitValue = as.numeric(TraitValue))
 ```
@@
 ```{r}
-averagetraits_seedMass <- numeric_data_average(traits_seedMass, "mg", "Seed Mass")
+# standardize trait names with units
+traits_seedMass <- traits_seedMass %>%
+  mutate(TraitName = "SeedMass_g")
+```
+
+```{r}
+averagetraits_seedMass <- numeric_data_average(traits_seedMass, "mg", "SeedMass_g")
+```
+
+```{r}
+nrow(averagetraits_seedMass)
+length(unique(averagetraits_seedMass$Accepted_species))
+```
+
+```{r}
+summary(averagetraits_seedMass)
 ```
@@
 traits_seedLength$OrigTraitValue <- traits_seedLength$TraitValue
+
+nrow(traits_seedLength)
+length(unique(traits_seedLength$Accepted_species))
+```
+
+```{r}
+# remove non-numeric rows
+traits_seedLength <- traits_seedLength %>%
+  filter(!is.na(as.numeric(TraitValue))) %>%
+  mutate(TraitValue = as.numeric(TraitValue))
+
+nrow(traits_seedLength)
+length(unique(traits_seedLength$Accepted_species))
 ```
@@
 ```{r}
-averagetraits_seedLength <- numeric_data_average(traits_seedLength, "mm", "Seed Length")
+# standardize trait names with units
+traits_seedLength <- traits_seedLength %>%
+  mutate(TraitName = "SeedLength_mm")
+```
+
+```{r}
+averagetraits_seedLength <- numeric_data_average(traits_seedLength, "mm", "SeedLength_mm")
+```
+
+```{r}
+nrow(averagetraits_seedLength)
+length(unique(averagetraits_seedLength$Accepted_species))
+```
+
+```{r}
+summary(averagetraits_fruitLength)
 ```
@@
 traits_seedWidth$OrigTraitValue <- traits_seedWidth$TraitValue
+
+nrow(traits_seedWidth)
+length(unique(traits_seedWidth$Accepted_species))
+```
+
+```{r}
+# remove non-numeric rows
+traits_seedWidth <- traits_seedWidth %>%
+  filter(!is.na(as.numeric(TraitValue))) %>%
+  mutate(TraitValue = as.numeric(TraitValue))
+
+nrow(traits_seedWidth)
+length(unique(traits_seedWidth$Accepted_species))
 ```
@@
 ```{r}
-averagetraits_seedWidth <- numeric_data_average(traits_seedWidth, "mm", "Seed Width")
+# standardize trait names with units
+traits_seedWidth <- traits_seedWidth %>%
+  mutate(TraitName = "SeedWidth_mm")
+```
+
+```{r}
+averagetraits_seedWidth <- numeric_data_average(traits_seedWidth, "mm", "SeedWidth_mm")
+```
+
+```{r}
+nrow(averagetraits_seedWidth)
+length(unique(averagetraits_seedWidth$Accepted_species))
+```
+
+```{r}
+summary(averagetraits_seedWidth)
 ```
@@
 traits_growthForm <- traits %>%
-  filter(grepl('Plant growth form|growth_form|whole plant growth form|whole plant growth form diversity'
+  filter(grepl('Plant growth form|growth_form|whole plant growth form'
 , TraitName))
 traits_growthForm$OrigTraitValue <- traits_growthForm$TraitValue
+
+nrow(traits_growthForm)
+length(unique(traits_growthForm$Accepted_species))
+```
+
+```{r}
+# remove numeric rows
+traits_growthForm <- traits_growthForm %>%
+  filter(is.na(as.numeric(TraitValue)))
+
+nrow(traits_growthForm)
+length(unique(traits_growthForm$Accepted_species))
+```
+
+```{r}
+# look at unique trait values
+traits_growthForm %>%
+  distinct(TraitValue)
 ```
@@
 wide_plant_traits$`Plant Height` <- as.numeric(wide_plant_traits$`Plant Height`)
-wide_plant_traits$`Dispersal Syndrome` <- as.character(wide_plant_traits$`Dispersal Syndrome`)
-wide_plant_traits$`Growth Forme` <- as.character(wide_plant_traits$`Growth Form`)
-wide_plant_traits$`Fruit Type` <- as.character(wide_plant_traits$`Fruit Type`)
+wide_plant_traits$`Dispersal Syndrome` <- as.factor(wide_plant_traits$`Dispersal Syndrome`)
+wide_plant_traits$`Growth Form` <- as.factor(wide_plant_traits$`Growth Form`)
+wide_plant_traits$`Fruit Type` <- as.factor(wide_plant_traits$`Fruit Type`)
 ```

### 2024-05-22 code/L1/L1_LastStep_DiversityInputObjects.Rmd
project PAM and update object names

@@
 ```{r}
-plants_PAM_matrix <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE, count = TRUE)
+plant_PAM_matrix <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE, count = TRUE)
 frugivore_PAM_matrix <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE, count = TRUE)
+```
+
+## Project PAM
+```{r}
+# check dimensions
+dim(plant_PAM_matrix)
+dim(frugivore_PAM_matrix)
+```
+
+```{r}
+# convert PAM matrix to dataframe
+plant_PAM_df <- as.data.frame(plant_PAM_matrix)
+frugivore_PAM_df <- as.data.frame(frugivore_PAM_matrix)
+```
+
+```{r}
+# check dimensions
+dim(plant_PAM_df)
+dim(frugivore_PAM_df)
+```
+
+```{r}
+# Convert to sf object and assign crs
+plant_PAM.sf <- st_as_sf(plant_PAM_df, coords = c("Longitude(x)", "Latitude(y)"), crs = 4326)
+frugivore_PAM.sf <- st_as_sf(frugivore_PAM_df, coords = c("Longitude(x)", "Latitude(y)"), crs = 4326)
+```
+
+```{r}
+# check dimensions
+dim(plant_PAM.sf)
+dim(frugivore_PAM.sf)
+```
+
+```{r}
+# project to crs to match taxonomic diversity
+plant_PAM.sf <- st_transform(plant_PAM.sf, 5389)
+frugivore_PAM.sf <- st_transform(frugivore_PAM.sf, 5389)
+```
+
+```{r}
+# check units
+st_crs(plant_PAM.sf)
+st_crs(frugivore_PAM.sf)
+```
+
+```{r}
+# convert sf to dataframe and retain lat & long as first two columns
+## extract coordinates
+plant_coords <- st_coordinates(plant_PAM.sf)
+frugivore_coords <- st_coordinates(frugivore_PAM.sf)
+
+## convert sf to dataframe
+plant_PAM.sf_df <- as.data.frame(st_drop_geometry(plant_PAM.sf))
+frugivore_PAM.sf_df <- as.data.frame(st_drop_geometry(frugivore_PAM.sf))
+
+# add coordinates as separate columns
+plant_PAM.sf_df$Longitude <- plant_coords[, 1]
+plant_PAM.sf_df$Latitude <- plant_coords[, 2]
+
+frugivore_PAM.sf_df$Longitude <- frugivore_coords[, 1]
+frugivore_PAM.sf_df$Latitude <- frugivore_coords[, 2]
+
+# rearrange columns to make Longitude and Latitude the first two columns
+plant_PAM.sf_df <- plant_PAM.sf_df[, c("Longitude", "Latitude", setdiff(names(plant_PAM.sf_df), c("Longitude", "Latitude")))]
+frugivore_PAM.sf_df <- frugivore_PAM.sf_df[, c("Longitude", "Latitude", setdiff(names(frugivore_PAM.sf_df), c("Longitude", "Latitude")))]
+```
+
+```{r}
+#check str of dataframes
+plant_PAM.sf_df[1:4, 1:4]
+frugivore_PAM.sf_df[1:4, 1:4]
+```
+
+```{r}
+# convert back to matrix and make sure all columns are numeric
+plant_PAM.sf_matrix <- as.matrix(plant_PAM.sf_df)
+frugivore_PAM.sf_matrix <- as.matrix(frugivore_PAM.sf_df)
+
+plant_PAM.sf_matrix <- apply(plant_PAM.sf_matrix, 2, as.numeric)
+frugivore_PAM.sf_matrix <- apply(frugivore_PAM.sf_matrix, 2, as.numeric)
+```
+
+```{r}
+# check dimensions
+dim(plant_PAM.sf_matrix)
+dim(frugivore_PAM.sf_matrix)
 ```
@@
 # Get the number of rows in the matrix
-num_rows_plant <- nrow(plants_PAM_matrix)
+num_rows_plant <- nrow(plant_PAM.sf_matrix)
 
@@
 # Assign row names to the matrix
-rownames(plants_PAM_matrix) <- row_names_plant
+rownames(plant_PAM.sf_matrix) <- row_names_plant
 
-column_names_plant <- colnames(plants_PAM_matrix)
+column_names_plant <- colnames(plant_PAM.sf_matrix)
 ```
@@
 # Get the number of rows in the matrix
-num_rows_frugivore <- nrow(frugivore_PAM_matrix)
+num_rows_frugivore <- nrow(frugivore_PAM.sf_matrix)
 
@@
 # Assign row names to the matrix
-rownames(frugivore_PAM_matrix) <- row_names_frugivore
+rownames(frugivore_PAM.sf_matrix) <- row_names_frugivore
 
-column_names_frugivore <- colnames(frugivore_PAM_matrix)
+column_names_frugivore <- colnames(frugivore_PAM.sf_matrix)
 ```
@@
 # Remove columns with sum equal to zero
-PAM_plant_site_final <- plants_PAM_matrix[, colSums(plants_PAM_matrix) != 0]
-PAM_frugivore_site_final <- frugivore_PAM_matrix[, colSums(frugivore_PAM_matrix) != 0]
+PAM_plant_site_final <- plant_PAM.sf_matrix[, colSums(plant_PAM.sf_matrix) != 0]
+PAM_frugivore_site_final <- frugivore_PAM.sf_matrix[, colSums(frugivore_PAM.sf_matrix) != 0]
 

### 2024-05-22 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
re-order mapping code

@@
 
+## Mapping FD
+
+```{r}
+# Generate coordinates
+subset_coords_frugivore <- site_loc_key_frugivore[rowSums(PAM_frugivore_site_final) >= 5, ]
+subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]
+
+frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)
+```
+
+```{r}
+# Convert the dataframe to sf format
+spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
+```
+
+```{r}
+# set crs of sf objects
+spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(4326)
+```
+
+```{r}
+# Create a grid over the area
+create_grid <- function(data, cellsize) {
+  st_make_grid(st_as_sfc(st_bbox(data)), cellsize = cellsize) %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+}
+
+# Aggregate functional dispersion within each grid cell
+aggregate_to_grid <- function(grid, points, value_col) {
+  points %>%
+    st_join(grid) %>%
+    group_by(cellid) %>%
+    summarize(mean_value = mean(get(value_col), na.rm = TRUE)) %>%
+    st_set_geometry("geometry")
+}
+
+# Example usage with your data
+resolution_meters <- 100000
+variable_name <- "fdis_frugivore"
+
+# Create grid
+grid <- create_grid(spatial_fdis_frugivore, cellsize = resolution_meters)
+
+# Aggregate data to grid
+aggregated_data <- aggregate_to_grid(grid, spatial_fdis_frugivore, variable_name)
+
+```
+
+```{r}
+FDisplot <-
+    ggplot() +
+    geom_sf(data = Americas_latlong, fill = "white")+
+    geom_sf(data = TApoly_latlong, fill = "lightgrey", size = 0.1) +
+    geom_point(data = spatial_fdis_frugivore , aes(fill = fdis_frugivore)) +
+    scale_fill_gradient(low = "blue", high = "red") +
+    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+    theme(panel.background = element_rect(fill = "lightblue")) +
+    annotation_scale(location = "bl", width_hint = 0.5) +
+    labs(x = "Longitude", y = "Latitude", fill = "FDis")
+FDisplot
+```
+
 # Mapping without projection 
@@
 FDisplot
-```
-
-## Project FD objects
-```{r}
-# Generate coordinates
-subset_coords_frugivore <- site_loc_key_frugivore[rowSums(PAM_frugivore_site_final) >= 5, ]
-subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]
-
-frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)
-```
-
-```{r}
-# Convert the dataframe to sf format
-spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
-```
-
-```{r}
-# set crs of sf objects
-spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(4326)
-```
-
-```{r}
-spatial_fdis_frugivore_projected <- st_transform(spatial_fdis_frugivore, 5389)
-```
-
-```{r}
-create_fdis_plots <- function(resolution_meters, data, variable_name) {
-# Make Grid
-TAGrid <- TApoly %>%
-  st_make_grid(cellsize = c(resolution_meters)) %>%
-  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
-  st_cast("MULTIPOLYGON") %>%
-  st_sf() %>%
-  mutate(cellid = row_number())
-
-  # Calculate functional dispersion for grid
-  fdis_grid <- TAGrid %>%
-    st_join(data) %>%
-    group_by(cellid) %>%
-    summarize(
-      sum_variable = sum(variable_name, na.rm = TRUE),
-      mean_variable = mean(variable_name, na.rm = TRUE)
-    )
-
-  # Make plot
-  FDplot <-
-    ggplot() +
-    geom_sf(data = Americas, fill = "white")+
-    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
-    geom_sf(data = fdis_grid, aes(fill = fdis_frugivore)) +
-    scale_fill_gradient(low = "blue", high = "red") +
-    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-    theme(panel.background = element_rect(fill = "lightblue")) +
-    annotation_scale(location = "bl", width_hint = 0.5) +
-    labs(x = "Longitude", y = "Latitude", fill = "FDis")
-
-  # Return a list of data
-  list(fdis_grid = fdis_grid,
-        FDplot = FDplot)
-}
-```
-
-```{r}
-create_fdis_plots(100000, spatial_fdis_frugivore_projected, fdis_frugivore)
 ```

### 2024-05-22 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
modify richness function to iteratively calculate richness if TAGrid has more than 1000 rows

@@
 
-  # Calculate cell richness for plants
-  plant_richness_grid <- TAGrid %>%
-    st_join(plants_sf_species) %>%
+  # Function to calculate richness for a given dataset
+  calculate_richness <- function(data_sf, species_sf) {
+    result <- data_sf %>%
+      st_join(species_sf) %>%
       mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
       group_by(cellid) %>%
-    summarize(num_species = sum(overlap))
+      summarize(num_species = sum(overlap)) %>%
+      ungroup()
+    return(result)
+  }
 
-  # Calculate cell richness for frugivores
-  frugivore_richness_grid <- TAGrid %>%
-    st_join(frugivores_sf_species) %>%
-    mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
-    group_by(cellid) %>%
-    summarize(num_species = sum(overlap))
+  # Determine the chunk size
+  chunk_size <- 1000
+
+  # Split the TAGrid into chunks if necessary
+  TAGrid_chunks <- split(TAGrid, ceiling(seq_along(TAGrid$cellid) / chunk_size))
+
+  # Initialize empty data frames for results
+  plant_richness_grid <- list()
+  frugivore_richness_grid <- list()
+
+  # Iterate over chunks to calculate richness
+  for (i in seq_along(TAGrid_chunks)) {
+    chunk <- TAGrid_chunks[[i]]
+
+    # Debug: Print chunk info
+    print(paste("Processing chunk", i, "of", length(TAGrid_chunks)))
+
+    plant_richness_chunk <- calculate_richness(chunk, plants_sf_species)
+    frugivore_richness_chunk <- calculate_richness(chunk, frugivores_sf_species)
+
+    # Store the chunks in lists
+    plant_richness_grid[[i]] <- plant_richness_chunk
+    frugivore_richness_grid[[i]] <- frugivore_richness_chunk
+  }
+
+  # Combine all chunks into final data frames
+  plant_richness_grid <- bind_rows(plant_richness_grid)
+  frugivore_richness_grid <- bind_rows(frugivore_richness_grid)
 

### 2024-05-22 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
add 10 km back to multipanel plots

@@
 all_plant_richness_plots <- ggpubr::ggarrange(#plantgridRichTA_1km, plantgridRichTA_5km, 
-                                              #plantgridRichTA_10km, 
-                                              plantgridRichTA_25km,
+                                              plantgridRichTA_10km, plantgridRichTA_25km,
                                               plantgridRichTA_50km, plantgridRichTA_75km,
                                               plantgridRichTA_100km,
-                                              ncol = 4, nrow = 1,
-                                              labels = c("25 km", "50 km", "75 km", "100 km"),
+                                              ncol = 5, nrow = 1,
+                                              labels = c("10 km", "25 km", "50 km", "75 km", "100 km"),
                                               common.legend = TRUE, legend = "right")
+all_plant_richness_plots
 all_plant_richness_plots_labeled <- ggpubr::annotate_figure(all_plant_richness_plots,
@@
 all_frugivore_richness_plots <- ggpubr::ggarrange(#frugivoregridRichTA_1km, frugivoregridRichTA_5km, 
-                                              #frugivoregridRichTA_10km, 
-                                              frugivoregridRichTA_25km,
+                                              frugivoregridRichTA_10km, frugivoregridRichTA_25km,
                                               frugivoregridRichTA_50km, frugivoregridRichTA_75km,
                                               frugivoregridRichTA_100km,
-                                              ncol = 4, nrow = 1,
-                                              labels = c("25 km", "50 km", "75 km", "100 km"),
+                                              ncol = 5, nrow = 1,
+                                              labels = c("10 km", "25 km", "50 km", "75 km", "100 km"),
                                               common.legend = TRUE, legend = "right")

### 2024-05-22 code/Manuscript_figures.rmd
add figure path, add new library for sankey diagram

@@
       code_link: TRUE
-      code_download: TRUE
 ---
 
+# Set file paths
+```{r}
+figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
+```
+
 # Load required packages
-
 ```{r}
@@
 library(ggpubr)
+library(networkD3)
 ```
@@
 
+# Create Sankey diagram for 
+Inspiration figures: https://conbio-onlinelibrary-wiley-com.proxy2.cl.msu.edu/doi/10.1111/csp2.311#csp2311-fig-0003 and https://bsapubs-onlinelibrary-wiley-com.proxy2.cl.msu.edu/doi/full/10.1002/aps3.11575#aps311575-fig-0003
+
+```{r}
+
+```
+
 # References
+
 ```{r}
@@
 
-# Session Info
+<details>
+
+<summary><h1>Session Info</h1></summary>
+
 ```{r, class.source= 'fold-hide'}
@@
 ```
+
+</details>

### 2024-05-22 code/L1/L1_LastStep_DiversityInputObjects.Rmd
add function to creat presence-absence-matrix by resolution -- this has not been tested yet

@@
 
+```{r}
+# Define the function
+create_presence_absence_matrix <- function(resolution_meters, species_sf) {
+  # Make the grid
+  TAGrid <- TApoly %>%
+    st_make_grid(cellsize = c(resolution_meters)) %>%
+    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+    st_cast("MULTIPOLYGON") %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+
+  # Join species data to the grid
+  species_grid <- TAGrid %>%
+    st_join(species_sf) %>%
+    filter(!is.na(species)) %>%
+    select(cellid, species)
+
+  # Create the presence-absence matrix
+  presence_absence_matrix <- species_grid %>%
+    mutate(presence = 1) %>%
+    pivot_wider(names_from = species, values_from = presence, values_fill = list(presence = 0)) %>%
+    as.data.frame()
+
+  # Ensure the cellid column is the first column
+  presence_absence_matrix <- presence_absence_matrix %>%
+    select(cellid, everything())
+
+  return(presence_absence_matrix)
+}
+```
+
+```{r}
+presence_absence_matrix <- create_presence_absence_matrix(resolution_meters, species_sf)
+```
+
 Use unprojected data because projected data is causing issues
@@
 frugivore_PAM_matrix <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE, count = TRUE)
+```
+
+```{r}
+summary(plant_PAM_matrix)
+summary(frugivore_PAM_matrix)
 ```

### 2024-05-22 code/L1/L1_P3_TropicalAndes_plant_traits.Rmd
add note that data is harmonized when reading it in

@@
 
-# Read in data
+# Read in harmonized data
 ```{r}

### 2024-05-22 code/L0/L0_TropicalAndes_plants_FRUBASE.Rmd
inital commit

---
title: "Tropical Andes plant trait data FRUBASE"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script retrives plant trait data from the FRUBASE database for the Tropical Andes."
data input: ""
data output: ""
date: "2024-05-14"
output: 
    html_document:
      code_link: TRUE
---

# Set file paths
```{r}
data_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
#devtools::install_github("pedroj/FRUBASE")
library(frubase)
library(dplyr)
```

# Retrive data from FRUBASE
Data downloaded on 2024-05-22
For more information about FRUBASE, see https://github.com/pedroj/FRUBASE 
```{r}
data(frubase)
```

### 2024-05-26 code/L1/L1_LastStep_DiversityInputObjects.Rmd
add packages, update presence-absence matrix function, run function for each resolution, add outputs from each resolution

--- 2024-05-14 code/L1/L1_LastStep_DiversityInputObjects.Rmd
@@
 library(fasterize)
+library(funbiogeo)
 library(dplyr)
-library(funbiogeo)
+library(tidyr)
+library(ggplot2)
+library(ggpubr)
 ```
@@
 
-Use unprojected data because projected data is causing issues
 ```{r}
-x <- TropicalAndes_plant_occ_forest$decimalLongitude
-y <- TropicalAndes_plant_occ_forest$decimalLatitude
-plant_xy <- cbind(x, y)
-plant_species <- TropicalAndes_plant_occ_forest$species
-plants_PAM <- letsR::lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, count = TRUE)
-summary(plants_PAM)
-plot(plants_PAM, xlab = "Longitude", ylab = "Latitude", main = "Plant richness map")
+# Define the function
+create_presence_absence_matrix <- function(resolution_meters, species_sf) {
+  # Make Grid
+  TAGrid <- TApoly %>%
+    st_make_grid(cellsize = c(resolution_meters)) %>%
+    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+    st_cast("MULTIPOLYGON") %>%
+    st_sf() %>%
+    mutate(cellid = row_number(),
+           centroid = st_centroid(geometry)) %>%
+    mutate(longitude = st_coordinates(centroid)[, 1],
+           latitude = st_coordinates(centroid)[, 2]) %>%
+    dplyr::select(-centroid)
+
+  # Join with species data and filter out NA species
+  species_grid <- TAGrid %>%
+    st_join(species_sf) %>%
+    filter(!is.na(species)) %>%
+    dplyr::select(cellid, longitude, latitude, species)
+
+  # Create Presence-Absence Matrix
+  presence_absence_matrix <- species_grid %>%
+    st_set_geometry(NULL) %>%
+    mutate(presence = 1) %>%
+    pivot_wider(names_from = species, values_from = presence, values_fill = list(presence = 0)) %>%
+    arrange(cellid) %>%
+    dplyr::select(cellid, longitude, latitude, everything())
+
+  # Convert to matrix and set row names
+  presence_absence_matrix <- as.data.frame(presence_absence_matrix)
+  rownames(presence_absence_matrix) <- presence_absence_matrix$cellid
+  presence_absence_matrix <- presence_absence_matrix %>%
+    dplyr::select(-cellid) %>%
+    as.matrix()
+
+  return(presence_absence_matrix)
+}
+```
+
+## 100 km 
+```{r}
+plant_PAM_100km <- create_presence_absence_matrix(100000, plants_sf_species)
+frugivore_PAM_100km <- create_presence_absence_matrix(100000, frugivores_sf_species)
 ```
@@
 ```{r}
-x <- TropicalAndes_frugivore_occ_forest$decimalLongitude
-y <- TropicalAndes_frugivore_occ_forest$decimalLatitude
-frugivore_xy <- cbind(x, y)
-frugivore_species <- TropicalAndes_frugivore_occ_forest$species
-frugivore_PAM <- letsR::lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, count = TRUE)
-summary(frugivore_PAM)
-plot(frugivore_PAM, xlab = "Longitude", ylab = "Latitude", main = "Frugivore richness map")
+#check str 
+plant_PAM_100km[1:4, 1:4]
+frugivore_PAM_100km[1:4, 1:4]
 ```
 
-## Turn PAM into matrix
+### Remove the species from PAM that have no occurrences
 ```{r}
-plants_PAM_matrix <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE, count = TRUE)
-frugivore_PAM_matrix <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE, count = TRUE)
-```
+# Remove columns with sum equal to zero
+PAM_plant_site_final_100km <- plant_PAM_100km[, colSums(plant_PAM_100km) != 0]
+PAM_frugivore_site_final_100km <- frugivore_PAM_100km[, colSums(frugivore_PAM_100km) != 0]
 
-## Add rownames
-```{r}
-# Get the number of rows in the matrix
-num_rows_plant <- nrow(plants_PAM_matrix)
+#Save coordinates for later
+site_loc_key_plant_100km <- PAM_plant_site_final_100km[,1:2]
+site_loc_key_frugivore_100km <- PAM_frugivore_site_final_100km[,1:2]
 
-# Generate unique row names
-row_names_plant <- paste0("cell", 1:num_rows_plant)
+PAM_plant_site_final_100km <- PAM_plant_site_final_100km[,-c(1:2)]
+PAM_frugivore_site_final_100km <- PAM_frugivore_site_final_100km[,-(1:2)]
 
-# Assign row names to the matrix
-rownames(plants_PAM_matrix) <- row_names_plant
-
-column_names_plant <- colnames(plants_PAM_matrix)
+colnames_plant_100km <- colnames(PAM_plant_site_final_100km)
+colnames_frugivore_100km <- colnames(PAM_frugivore_site_final_100km)
 ```
@@
 ```{r}
-# Get the number of rows in the matrix
-num_rows_frugivore <- nrow(frugivore_PAM_matrix)
-
-# Generate unique row names
-row_names_frugivore <- paste0("cell", 1:num_rows_frugivore)
-
-# Assign row names to the matrix
-rownames(frugivore_PAM_matrix) <- row_names_frugivore
-
-column_names_frugivore <- colnames(frugivore_PAM_matrix)
+str(PAM_plant_site_final_100km)
+str(PAM_frugivore_site_final_100km)
 ```
-
-## Turn data into correct data types for inputs into the trait categories dataframe
+## 75 km 
 ```{r}
-plant_traits$Growth.Form <- as.factor(plant_traits$Growth.Form)
-plant_traits$Plant.Height_m <- as.numeric(plant_traits$Plant.Height_m)
-plant_traits$Fruit.Length_mm <- as.numeric(plant_traits$Fruit.Length_mm)
-plant_traits$Fruit.Type <- as.factor(plant_traits$Fruit.Type)
+plant_PAM_75km <- create_presence_absence_matrix(75000, plants_sf_species)
+frugivore_PAM_75km <- create_presence_absence_matrix(75000, frugivores_sf_species)
 ```
@@
 ```{r}
-frugivore_traits$diet_cat <- as.factor(frugivore_traits$diet_cat)
-frugivore_traits$body_mass_e <- as.numeric(frugivore_traits$body_mass_e)
-frugivore_traits$body_size_mm <- as.numeric(frugivore_traits$body_size_mm)
-frugivore_traits$generation_time <- as.numeric(frugivore_traits$generation_time)
+#check str 
+plant_PAM_75km[1:4, 1:4]
+frugivore_PAM_75km[1:4, 1:4]
 ```
 
-## Remove the species from PAM that have no occurrences
+### Remove the species from PAM that have no occurrences
 ```{r}
 # Remove columns with sum equal to zero
-PAM_plant_site_final <- plants_PAM_matrix[, colSums(plants_PAM_matrix) != 0]
-PAM_frugivore_site_final <- frugivore_PAM_matrix[, colSums(frugivore_PAM_matrix) != 0]
+PAM_plant_site_final_75km <- plant_PAM_75km[, colSums(plant_PAM_75km) != 0]
+PAM_frugivore_site_final_75km <- frugivore_PAM_75km[, colSums(frugivore_PAM_75km) != 0]
 
 #Save coordinates for later
-site_loc_key_plant <- PAM_plant_site_final[,1:2]
-site_loc_key_frugivore <- PAM_frugivore_site_final[,1:2]
+site_loc_key_plant_75km <- PAM_plant_site_final_75km[,1:2]
+site_loc_key_frugivore_75km <- PAM_frugivore_site_final_75km[,1:2]
 
-PAM_plant_site_final <- PAM_plant_site_final[,-c(1:2)]
-PAM_frugivore_site_final <- PAM_frugivore_site_final[,-(1:2)]
+PAM_plant_site_final_75km <- PAM_plant_site_final_75km[,-c(1:2)]
+PAM_frugivore_site_final_75km <- PAM_frugivore_site_final_75km[,-(1:2)]
 
-colnames_plant <- colnames(PAM_plant_site_final)
-colnames_frugivore <- colnames(PAM_frugivore_site_final)
+colnames_plant_75km <- colnames(PAM_plant_site_final_75km)
+colnames_frugivore_75km <- colnames(PAM_frugivore_site_final_75km)
 ```
@@
 ```{r}
-str(PAM_plant_site_final)
-str(PAM_frugivore_site_final)
+str(PAM_plant_site_final_75km)
+str(PAM_frugivore_site_final_75km)
+```
+
+## 50 km 
+```{r}
+plant_PAM_50km <- create_presence_absence_matrix(50000, plants_sf_species)
+frugivore_PAM_50km <- create_presence_absence_matrix(50000, frugivores_sf_species)
+```
+
+```{r}
+#check str 
+plant_PAM_50km[1:4, 1:4]
+frugivore_PAM_50km[1:4, 1:4]
+```
+
+### Remove the species from PAM that have no occurrences
+```{r}
+# Remove columns with sum equal to zero
+PAM_plant_site_final_50km <- plant_PAM_50km[, colSums(plant_PAM_50km) != 0]
+PAM_frugivore_site_final_50km <- frugivore_PAM_50km[, colSums(frugivore_PAM_50km) != 0]
+
+#Save coordinates for later
+site_loc_key_plant_50km <- PAM_plant_site_final_50km[,1:2]
+site_loc_key_frugivore_50km <- PAM_frugivore_site_final_50km[,1:2]
+
+PAM_plant_site_final_50km <- PAM_plant_site_final_50km[,-c(1:2)]
+PAM_frugivore_site_final_50km <- PAM_frugivore_site_final_50km[,-(1:2)]
+
+colnames_plant_50km <- colnames(PAM_plant_site_final_50km)
+colnames_frugivore_50km <- colnames(PAM_frugivore_site_final_50km)
+```
+
+```{r}
+str(PAM_plant_site_final_50km)
+str(PAM_frugivore_site_final_50km)
+```
+
+## 25 km 
+```{r}
+plant_PAM_25km <- create_presence_absence_matrix(25000, plants_sf_species)
+frugivore_PAM_25km <- create_presence_absence_matrix(25000, frugivores_sf_species)
+```
+
+```{r}
+#check str 
+plant_PAM_25km[1:4, 1:4]
+frugivore_PAM_25km[1:4, 1:4]
+```
+
+### Remove the species from PAM that have no occurrences
+```{r}
+# Remove columns with sum equal to zero
+PAM_plant_site_final_25km <- plant_PAM_25km[, colSums(plant_PAM_25km) != 0]
+PAM_frugivore_site_final_25km <- frugivore_PAM_25km[, colSums(frugivore_PAM_25km) != 0]
+
+#Save coordinates for later
+site_loc_key_plant_25km <- PAM_plant_site_final_25km[,1:2]
+site_loc_key_frugivore_25km <- PAM_frugivore_site_final_25km[,1:2]
+
+PAM_plant_site_final_25km <- PAM_plant_site_final_25km[,-c(1:2)]
+PAM_frugivore_site_final_25km <- PAM_frugivore_site_final_25km[,-(1:2)]
+
+colnames_plant_25km <- colnames(PAM_plant_site_final_25km)
+colnames_frugivore_25km <- colnames(PAM_frugivore_site_final_25km)
+```
+
+```{r}
+str(PAM_plant_site_final_25km)
+str(PAM_frugivore_site_final_25km)
+```
+
+## 10 km 
+```{r}
+plant_PAM_10km <- create_presence_absence_matrix(10000, plants_sf_species)
+frugivore_PAM_10km <- create_presence_absence_matrix(10000, frugivores_sf_species)
+```
+
+```{r}
+#check str 
+plant_PAM_10km[1:4, 1:4]
+frugivore_PAM_10km[1:4, 1:4]
+```
+
+### Remove the species from PAM that have no occurrences
+```{r}
+# Remove columns with sum equal to zero
+PAM_plant_site_final_10km <- plant_PAM_10km[, colSums(plant_PAM_10km) != 0]
+PAM_frugivore_site_final_10km <- frugivore_PAM_10km[, colSums(frugivore_PAM_10km) != 0]
+
+#Save coordinates for later
+site_loc_key_plant_10km <- PAM_plant_site_final_10km[,1:2]
+site_loc_key_frugivore_10km <- PAM_frugivore_site_final_10km[,1:2]
+
+PAM_plant_site_final_10km <- PAM_plant_site_final_10km[,-c(1:2)]
+PAM_frugivore_site_final_10km <- PAM_frugivore_site_final_10km[,-(1:2)]
+
+colnames_plant_10km <- colnames(PAM_plant_site_final_10km)
+colnames_frugivore_10km <- colnames(PAM_frugivore_site_final_10km)
+```
+
+```{r}
+str(PAM_plant_site_final_10km)
+str(PAM_frugivore_site_final_10km)
+```
+
+## 5 km 
+```{r}
+plant_PAM_5km <- create_presence_absence_matrix(5000, plants_sf_species)
+frugivore_PAM_5km <- create_presence_absence_matrix(5000, frugivores_sf_species)
+```
+
+```{r}
+#check str 
+plant_PAM_5km[1:4, 1:4]
+frugivore_PAM_5km[1:4, 1:4]
+```
+
+### Remove the species from PAM that have no occurrences
+```{r}
+# Remove columns with sum equal to zero
+PAM_plant_site_final_5km <- plant_PAM_5km[, colSums(plant_PAM_5km) != 0]
+PAM_frugivore_site_final_5km <- frugivore_PAM_5km[, colSums(frugivore_PAM_5km) != 0]
+
+#Save coordinates for later
+site_loc_key_plant_5km <- PAM_plant_site_final_5km[,1:2]
+site_loc_key_frugivore_5km <- PAM_frugivore_site_final_5km[,1:2]
+
+PAM_plant_site_final_5km <- PAM_plant_site_final_5km[,-c(1:2)]
+PAM_frugivore_site_final_5km <- PAM_frugivore_site_final_5km[,-(1:2)]
+
+colnames_plant_5km <- colnames(PAM_plant_site_final_5km)
+colnames_frugivore_5km <- colnames(PAM_frugivore_site_final_5km)
+```
+
+```{r}
+str(PAM_plant_site_final_5km)
+str(PAM_frugivore_site_final_5km)
+```
+
+## 1 km 
+```{r}
+plant_PAM_1km <- create_presence_absence_matrix(1000, plants_sf_species)
+frugivore_PAM_1km <- create_presence_absence_matrix(1000, frugivores_sf_species)
+```
+
+```{r}
+#check str 
+plant_PAM_1km[1:4, 1:4]
+frugivore_PAM_1km[1:4, 1:4]
+```
+
+### Remove the species from PAM that have no occurrences
+```{r}
+# Remove columns with sum equal to zero
+PAM_plant_site_final_1km <- plant_PAM_1km[, colSums(plant_PAM_1km) != 0]
+PAM_frugivore_site_final_1km <- frugivore_PAM_1km[, colSums(frugivore_PAM_1km) != 0]
+
+#Save coordinates for later
+site_loc_key_plant_1km <- PAM_plant_site_final_1km[,1:2]
+site_loc_key_frugivore_1km <- PAM_frugivore_site_final_1km[,1:2]
+
+PAM_plant_site_final_1km <- PAM_plant_site_final_1km[,-c(1:2)]
+PAM_frugivore_site_final_1km <- PAM_frugivore_site_final_1km[,-(1:2)]
+
+colnames_plant_1km <- colnames(PAM_plant_site_final_1km)
+colnames_frugivore_1km <- colnames(PAM_frugivore_site_final_1km)
+```
+
+```{r}
+str(PAM_plant_site_final_1km)
+str(PAM_frugivore_site_final_1km)
 ```
@@
 plant_traits_df_subset <- plant_traits %>%
-                            filter(species %in% colnames_plant) %>%
+                            filter(species %in% colnames_plant_100km) %>%
                             distinct(species, .keep_all = TRUE)
 frugivore_traits_df_subset <- frugivore_traits %>%
-                                filter(IUCN_species_name %in% colnames_frugivore) %>%
+                                filter(IUCN_species_name %in% colnames_frugivore_100km) %>%
                                 distinct(IUCN_species_name, .keep_all = TRUE)
@@
 # Export data
+## sf objects
 ```{r}
-# objects to save
-## sf objects
 saveRDS(plants_sf_species, file = file.path(data_path_L1,"plants_sf_species.rds"))
@@
 
+# PAM objects
 ```{r}
-# PAM objects
-saveRDS(plants_PAM, file = file.path(data_path_L1,"plants_PAM.rds"))
-saveRDS(frugivore_PAM, file = file.path(data_path_L1,"frugivore_PAM.rds"))
+saveRDS(plant_PAM_100km, file = file.path(data_path_L1,"plant_PAM_100km.rds"))
+saveRDS(frugivore_PAM_100km, file = file.path(data_path_L1,"frugivore_PAM_100km.rds"))
 ```
@@
 ```{r}
-# objects for functional diversity
-saveRDS(site_loc_key_plant, file = file.path(data_path_L1,"site_loc_key_plant.rds"))
-saveRDS(site_loc_key_frugivore, file = file.path(data_path_L1,"site_loc_key_frugivore.rds"))
-saveRDS(PAM_plant_site_final, file = file.path(data_path_L1,"PAM_plant_site_final.rds"))
-saveRDS(PAM_frugivore_site_final, file = file.path(data_path_L1,"PAM_frugivore_site_final.rds"))
+saveRDS(plant_PAM_75km, file = file.path(data_path_L1,"plant_PAM_75km.rds"))
+saveRDS(frugivore_PAM_75km, file = file.path(data_path_L1,"frugivore_PAM_75km.rds"))
+```
+
+```{r}
+saveRDS(plant_PAM_50km, file = file.path(data_path_L1,"plant_PAM_50km.rds"))
+saveRDS(frugivore_PAM_50km, file = file.path(data_path_L1,"frugivore_PAM_50km.rds"))
+```
+
+```{r}
+saveRDS(plant_PAM_25km, file = file.path(data_path_L1,"plant_PAM_25km.rds"))
+saveRDS(frugivore_PAM_25km, file = file.path(data_path_L1,"frugivore_PAM_25km.rds"))
+```
+
+```{r}
+saveRDS(plant_PAM_10km, file = file.path(data_path_L1,"plant_PAM_10km.rds"))
+saveRDS(frugivore_PAM_10km, file = file.path(data_path_L1,"frugivore_PAM_10km.rds"))
+```
+
+```{r}
+saveRDS(plant_PAM_5km, file = file.path(data_path_L1,"plant_PAM_5km.rds"))
+saveRDS(frugivore_PAM_5km, file = file.path(data_path_L1,"frugivore_PAM_5km.rds"))
+```
+
+```{r}
+saveRDS(plant_PAM_1km, file = file.path(data_path_L1,"plant_PAM_1km.rds"))
+saveRDS(frugivore_PAM_1km, file = file.path(data_path_L1,"frugivore_PAM_1km.rds"))
+```
+
+## objects for functional diversity
+```{r}
 saveRDS(plant_traits_df_subset, file = file.path(data_path_L1,"plant_traits_df_final.rds"))
@@
 ```
+
+```{r}
+saveRDS(site_loc_key_plant_100km, file = file.path(data_path_L1,"site_loc_key_plant_100km.rds"))
+saveRDS(site_loc_key_frugivore_100km, file = file.path(data_path_L1,"site_loc_key_frugivore_100km.rds"))
+saveRDS(PAM_plant_site_final_100km, file = file.path(data_path_L1,"PAM_plant_site_final_100km.rds"))
+saveRDS(PAM_frugivore_site_final_100km, file = file.path(data_path_L1,"PAM_frugivore_site_final_100km.rds"))
+```
+
+```{r}
+saveRDS(site_loc_key_plant_75km, file = file.path(data_path_L1,"site_loc_key_plant_75km.rds"))
+saveRDS(site_loc_key_frugivore_75km, file = file.path(data_path_L1,"site_loc_key_frugivore_75km.rds"))
+saveRDS(PAM_plant_site_final_75km, file = file.path(data_path_L1,"PAM_plant_site_final_75km.rds"))
+saveRDS(PAM_frugivore_site_final_75km, file = file.path(data_path_L1,"PAM_frugivore_site_final_75km.rds"))
+```
+
+```{r}
+saveRDS(site_loc_key_plant_50km, file = file.path(data_path_L1,"site_loc_key_plant_50km.rds"))
+saveRDS(site_loc_key_frugivore_50km, file = file.path(data_path_L1,"site_loc_key_frugivore_50km.rds"))
+saveRDS(PAM_plant_site_final_50km, file = file.path(data_path_L1,"PAM_plant_site_final_50km.rds"))
+saveRDS(PAM_frugivore_site_final_50km, file = file.path(data_path_L1,"PAM_frugivore_site_final_50km.rds"))
+```
+
+```{r}
+saveRDS(site_loc_key_plant_25km, file = file.path(data_path_L1,"site_loc_key_plant_25km.rds"))
+saveRDS(site_loc_key_frugivore_25km, file = file.path(data_path_L1,"site_loc_key_frugivore_25km.rds"))
+saveRDS(PAM_plant_site_final_25km, file = file.path(data_path_L1,"PAM_plant_site_final_25km.rds"))
+saveRDS(PAM_frugivore_site_final_25km, file = file.path(data_path_L1,"PAM_frugivore_site_final_25km.rds"))
+```
+
+```{r}
+saveRDS(site_loc_key_plant_10km, file = file.path(data_path_L1,"site_loc_key_plant_10km.rds"))
+saveRDS(site_loc_key_frugivore_10km, file = file.path(data_path_L1,"site_loc_key_frugivore_10km.rds"))
+saveRDS(PAM_plant_site_final_10km, file = file.path(data_path_L1,"PAM_plant_site_final_10km.rds"))
+saveRDS(PAM_frugivore_site_final_10km, file = file.path(data_path_L1,"PAM_frugivore_site_final_10km.rds"))
+```
+
+```{r}
+saveRDS(site_loc_key_plant_5km, file = file.path(data_path_L1,"site_loc_key_plant_5km.rds"))
+saveRDS(site_loc_key_frugivore_5km, file = file.path(data_path_L1,"site_loc_key_frugivore_5km.rds"))
+saveRDS(PAM_plant_site_final_5km, file = file.path(data_path_L1,"PAM_plant_site_final_5km.rds"))
+saveRDS(PAM_frugivore_site_final_5km, file = file.path(data_path_L1,"PAM_frugivore_site_final_5km.rds"))
+```
+
+```{r}
+saveRDS(site_loc_key_plant_1km, file = file.path(data_path_L1,"site_loc_key_plant_1km.rds"))
+saveRDS(site_loc_key_frugivore_1km, file = file.path(data_path_L1,"site_loc_key_frugivore_1km.rds"))
+saveRDS(PAM_plant_site_final_1km, file = file.path(data_path_L1,"PAM_plant_site_final_1km.rds"))
+saveRDS(PAM_frugivore_site_final_1km, file = file.path(data_path_L1,"PAM_frugivore_site_final_1km.rds"))
+```

### 2024-05-29 code/L1/L1_P3_TropicalAndes_plant_fruiting_subset.Rmd
inital commit

--- 2024-05-10 code/L1/L1_P3_TropicalAndes_plant_traits.Rmd
@@
 ---
-title: "TITLE"
+title: "Subsetting plant species list by fruiting species"
 author: "Hazel J. Anderson"
@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: "This script subsets plant trait data by synonyms species list and combine."
+overview: "This script subsets plant occurrence and trait data by fruiting plant species."
 data input: "ADD"
 data output: "ADD"
-date: "2023-11-04"
+date: "2024-05-29"
 output: html_document
@@
 
+```{r}
+TRY_traits %>%
+  distinct(Accepted_species, TraitName) %>%
+  count(TraitName)
+```
+
 ## BIEN
@@
 
-# Standardize trait measurements
-## Create functions
 ```{r}
-# Numeric data
-numeric_data_average <- function(data, unit, traitname) {
-# remove rows if TraitValue is non-numeric
-data <- data %>%
-  filter(!is.na(as.numeric(TraitValue))) %>%
-  mutate(TraitValue = as.numeric(TraitValue))
-# remove duplicate rows
-data <- data %>%
-  distinct(.keep_all = TRUE)
-# calculate average value per trait per species
-average_data <- aggregate(TraitValue ~ Accepted_species, data, FUN = mean, na.rm = TRUE)
-# Remove rows with missing or empty species names
-average_data <- average_data[!is.na(average_data$Accepted_species) & average_data$Accepted_species !="", ]
-# Add unit & trait column
-average_data$Unit <- unit
-average_data$TraitName <- traitname
-
-return(average_data)
-}
+traits %>%
+  distinct(Accepted_species, TraitName) %>%
+  count(TraitName)
 ```
 
+```{r}
+# Group by TraitName and summarize the unique TraitValues
+unique_trait_values <- traits %>%
+  group_by(TraitName) %>%
+  summarize(UniqueValues = list(unique(TraitValue)))%>%
+  unnest(UniqueValues)
+```
+
+# Function to get one value per species
 ```{r}
@@
 
-## Dispersal syndrome
+# Fruit type
 ```{r}
-traits_dispersal <- traits %>%
-  filter(grepl('Dispersal syndrome|whole plant dispersal syndrome|dispersal_syndrome_1|dispersal_syndrome_2', TraitName))
-traits_dispersal$OrigTraitValue <- traits_dispersal$TraitValue
+traits_fruitType <- traits %>%
+  filter(grepl('Fruit type|fruit type|fruit_type', TraitName))
+traits_fruitType$OrigTraitValue <- traits_fruitType$TraitValue
+
+nrow(traits_fruitType)
+length(unique(traits_fruitType$Accepted_species))
 ```
@@
 ```{r}
-# remove rows containing weight or mass
-traits_dispersal <- traits_dispersal[!grepl("weight|mass|air-dry|oven|mean|SE",traits_dispersal$TraitValue, ignore.case = TRUE),]
-nrow(traits_dispersal)
-length(unique(traits_dispersal$Accepted_species))
-```
+traits_fruitType <- traits_fruitType %>%
+  filter(is.na(as.numeric(TraitValue)))
 
-Reclassify to the following:
-anemochorous - wind dispersal
-anthropochorous - human dispersal
-autochorous - non-aided dispersal
-hydrochorous - water dispersal
-myrmecochorous - ant dispersal
-zoochorous - animal dispersal
-unspecialized - multiple
-```{r}
-traits_dispersal <- traits_dispersal %>%
-  filter() %>%
-  mutate(TraitValue= case_when(
-    grepl('Eaten|Animal|Bird|Mammal|animals|endozoochor|dysochor|epizoochor|zoochor|Vertebrates|Reptiles+birds|Endo-zoochory|mammals|exozoochory|birds|Zoochory|vertebrate|Dys-zoochory|Ornithochory|Epi-zoochory|endozoochorous|endozoochory|ectozoochorous|zoochorous', OrigTraitValue, ignore.case = TRUE) ~ "zoochorous",
-    grepl('Ants', OrigTraitValue, ignore.case = TRUE) ~ "myrmecochorous",
-    grepl('wind|Anemochory|anemochor|meteorochor|chamaechor|Anemochory|anemochorous', OrigTraitValue, ignore.case = TRUE) ~ "anemochorous",
-    grepl('human|agochor|man|hemerochor|anthropochorous', OrigTraitValue, ignore.case = TRUE) ~ "anthropochorous",
-    grepl('No|Unassisted|autochor|blastochor|ombrochor|Autochory|Barochory|ballochor|herpochor|autochorous', OrigTraitValue, ignore.case = TRUE) ~ "autochorous",
-    grepl('water|nautochor|floating|hydrochor|hydrochorous', OrigTraitValue, ignore.case = TRUE) ~ "hydrochorous",
-    grepl('boleochor|combination|non specialized|unspecialized', OrigTraitValue, ignore.case = TRUE) ~ "unspecialized"
-  )) %>%
-  filter(!is.na(TraitValue))
-nrow(traits_dispersal)
-length(unique(traits_dispersal$Accepted_species))
+nrow(traits_fruitType)
+length(unique(traits_fruitType$Accepted_species))
 ```
@@
 ```{r}
-print(unique(traits_dispersal$TraitValue))
+traits_fruitType %>%
+  count(TraitValue)
 ```
@@
 ```{r}
-traits_dispersal <- traits_dispersal %>%
-  mutate(TraitName = "dispersal_syndrome")
+# standardize trait values
+traits_fruitType <- traits_fruitType %>%
+  filter() %>%
+  mutate(TraitValue= case_when(
+    grepl('Berry|berry|Berry ', OrigTraitValue, ignore.case = TRUE) ~ "berry",
+    grepl('capsule|Capsule|Capsule ', OrigTraitValue, ignore.case = TRUE) ~ "capsule",
+    grepl('follicle|Follicle', OrigTraitValue, ignore.case = TRUE) ~ "follicle",
+  )) %>%
+  filter(!is.na(TraitValue))
 ```
@@
 ```{r}
-averagetraits_dispersal <- factor_data_merge(traits_dispersal, "Dispersal Syndrome")
-```
-
-## Fruit dry mass
-```{r}
-traits_fruitMass <- traits %>%
-  filter(grepl('Fruit dry mass', TraitName))
-traits_fruitMass$OrigTraitValue <- traits_fruitMass$TraitValue
-```
-
-```{r}
-print(unique(traits_fruitMass$Unit))
-```
-
-```{r}
-averagetraits_fruitMass <- numeric_data_average(traits_fruitMass, "mg", "Fruit Mass")
-```
-
-## Fruit length
-```{r}
-traits_fruitLength <- traits %>%
-  filter(grepl('Fruit length|maximum fruit length|minimum fruit length|fruit_length_min|fruit_length_max|fruit_length_mean', TraitName))
-traits_fruitLength$OrigTraitValue <- traits_fruitLength$TraitValue
-```
-
-```{r}
-print(unique(traits_fruitLength$Unit))
-```
-
-```{r}
-# remove rows if TraitValue is non-numeric
-traits_fruitLength <- traits_fruitLength %>%
-  filter(!is.na(as.numeric(TraitValue))) %>%
-  mutate(TraitValue = as.numeric(TraitValue))
-```
-
-```{r}
-# if units = cm, multiply trait value by 10 and change units to mm
-traits_fruitLength <- traits_fruitLength %>%
-  mutate(TraitValue = if_else(Unit == "cm", TraitValue * 10, TraitValue),
-         Unit = if_else(Unit == "cm", "mm", Unit))
-```
-
-```{r}
-averagetraits_fruitLength <- numeric_data_average(traits_fruitLength, "mm", "Fruit Length")
-```
-
-## Plant height
-```{r}
-traits_plantHeight <- traits %>%
-  filter(grepl('Plant height generative|Plant height vegetative|maximum whole plant height|minimum whole plant height|plant_height|plant_height_max|plant_height_min|whole plant height', TraitName))
-traits_plantHeight$OrigTraitValue <- traits_plantHeight$TraitValue
-```
-
-```{r}
-print(unique(traits_plantHeight$Unit))
-```
-
-```{r}
-averagetraits_plantHeight <- numeric_data_average(traits_plantHeight, "m", "Plant Height")
-```
-
-## Plant lifespan/longevity
-```{r}
-traits_plantLifespan <- traits %>%
-  filter(grepl('Plant lifespan (longevity)|longest whole plant longevity|maximum whole plant longevity|plant_lifespan', TraitName))
-traits_plantLifespan$OrigTraitValue <- traits_plantLifespan$TraitValue
-```
-
-```{r}
-print(unique(traits_plantLifespan$Unit))
-```
-
-```{r}
-averagetraits_plantLifespan <- numeric_data_average(traits_plantLifespan, "years", "Plant Lifespan")
-```
-
-## Fruit type
-```{r}
-traits_fruitType <- traits %>%
-  filter(grepl('Fruit dehiscence type|Fruit surface type|Fruit texture inside (fleshy, leathrey, …)|Fruit type|fruit type|fruit_dryness|fruit_type', TraitName))
-traits_fruitType$OrigTraitValue <- traits_fruitType$TraitValue
-```
-
-```{r}
-print(unique(traits_fruitType$Unit))
+# standardize trait name
+traits_fruitType <- traits_fruitType %>%
+  mutate(TraitName = "FruitType")
 ```
@@
 
-## Seed mass
 ```{r}
-traits_seedMass <- traits %>%
-  filter(grepl('Seed dry mass|seed mass|seed_mass_max|seed_mass_mean|seed_mass_min', TraitName))
-traits_seedMass$OrigTraitValue <- traits_seedMass$TraitValue
+nrow(averagetraits_fruitType)
+length(unique(averagetraits_fruitType$Accepted_species))
 ```
@@
 ```{r}
-print(unique(traits_seedMass$Unit))
+summary(averagetraits_fruitType)
 ```
-
-```{r}
-# remove rows if TraitValue is non-numeric
-traits_seedMass <- traits_seedMass %>%
-  filter(!is.na(as.numeric(TraitValue))) %>%
-  mutate(TraitValue = as.numeric(TraitValue))
-```
-
-```{r}
-# convert rows with unit = g to mg; multiply by 1000
-traits_seedMass <- traits_seedMass %>%
-  mutate(TraitValue = if_else(Unit == "g", TraitValue * 1000, TraitValue),
-         Unit = if_else(Unit == "g", "mg", Unit))
-```
-
-```{r}
-averagetraits_seedMass <- numeric_data_average(traits_seedMass, "mg", "Seed Mass")
-```
-
-## Seed length
-```{r}
-traits_seedLength <- traits %>%
-  filter(grepl('Seed length|seed_length_max|seed_length_mean|seed_length_min', TraitName))
-traits_seedLength$OrigTraitValue <- traits_seedLength$TraitValue
-```
-
-```{r}
-print(unique(traits_seedLength$Unit))
-```
-
-```{r}
-averagetraits_seedLength <- numeric_data_average(traits_seedLength, "mm", "Seed Length")
-```
-
-## Seed width
-```{r}
-traits_seedWidth <- traits %>%
-  filter(grepl('Seed width|seed_width_max|seed_width_mean|seed_width_min'
-, TraitName))
-traits_seedWidth$OrigTraitValue <- traits_seedWidth$TraitValue
-```
-
-```{r}
-print(unique(traits_seedWidth$Unit))
-```
-
-```{r}
-averagetraits_seedWidth <- numeric_data_average(traits_seedWidth, "mm", "Seed Width")
-```
-
-## Growth form
-```{r}
-traits_growthForm <- traits %>%
-  filter(grepl('Plant growth form|growth_form|whole plant growth form|whole plant growth form diversity'
-, TraitName))
-traits_growthForm$OrigTraitValue <- traits_growthForm$TraitValue
-```
-
-```{r}
-averagetraits_growthForm <- factor_data_merge(traits_growthForm, "Growth Form")
-```
-
-# Combine all traits
-```{r}
-# add unit column to categorical traits
-cat_traits <- rbind(averagetraits_dispersal, averagetraits_fruitType, averagetraits_growthForm)
-cat_traits$Unit <- ""
-clean_traits <- rbind(averagetraits_fruitLength, averagetraits_fruitMass, averagetraits_plantHeight, averagetraits_plantLifespan, averagetraits_seedLength, averagetraits_seedMass, averagetraits_seedWidth, cat_traits)
-nrow(clean_traits)
-```
-
-```{r}
-clean_traits %>%
-  count(TraitName)
-```
-
-# Summary
-```{r}
-print("Number of records: ")
-nrow(clean_traits)
-print("Number of species: ")
-length(unique(clean_traits$Accepted_species))
-```
-
-# Get one trait value per species
-```{r}
-#remove unit column from dataframe
-clean_traits <- clean_traits[, !colnames(clean_traits) %in% "Unit", drop = FALSE]
-```
-
-```{r}
-# separate numeric and non-numeric traits
-numeric_traits <- clean_traits[clean_traits$TraitName %in% c("Fruit Length", "Fruit Mass", "Plant Height", "Plant Lifespan", "Seed Length", "Seed Width", "Seed Mass"), ]
-cat_traits <- clean_traits[clean_traits$TraitName %in% c("Dispersal Syndrome", "Fruit Type", "Growth Form"), ]
-
-numeric_traits$TraitValue <- as.numeric(numeric_traits$TraitValue)
-```
-
-```{r}
-# numeric traits
-## Group by species and TraitName, and calculate the mean TraitValue for each combination
-numeric_traits_combined <- numeric_traits %>%
-  group_by(species, TraitName) %>%
-  summarise(TraitValue = mean(TraitValue, na.rm = TRUE)) %>%
-  ungroup()
-
-# non-numeric traits
-cat_traits_combined <- cat_traits %>%
-  group_by(species) %>%
-  summarise_all(.funs = list(mode_trait = function(x) {
-    mode_value <- names(sort(table(x), decreasing = TRUE))[1]  # Calculate mode
-    return(mode_value)
-  }))
-```
-
-```{r}
-#rename columns to original values
-colnames(cat_traits_combined)[which(colnames(cat_traits_combined) == "TraitValue_mode_trait")] <- "TraitValue"
-colnames(cat_traits_combined)[which(colnames(cat_traits_combined) == "TraitName_mode_trait")] <- "TraitName"
-
-# Remove rows with "?" or numbers in the TraitValue column
-cleaned_cat_traits <- cat_traits_combined %>%
-  filter(!grepl("\\?", TraitValue) & !grepl("\\d", TraitValue))
-```
-
-```{r}
-#combine numeric and non-numeric traits
-combined_clean_traits <- rbind(cat_traits_combined, numeric_traits_combined)
-```
-
-```{r}
-combined_clean_traits %>%
-  count(TraitName)
-```
-
-# Remove traits with less than 10 records
-```{r}
-combined_clean_traits <- combined_clean_traits[!(combined_clean_traits$TraitName %in% c("Fruit Mass", "Seed Mass")), ]
-```
-
-# Transform plant dataframe from long to wide
-```{r}
-wide_traits <- spread(combined_clean_traits, key = TraitName, value = TraitValue)
-
-#remove X column from dataframe
-wide_plant_traits <- wide_traits[, !colnames(wide_traits) %in% "X", drop = FALSE]
-```
-
-```{r}
-# Assign data types
-wide_plant_traits$`Fruit Length` <- as.numeric(wide_plant_traits$`Fruit Length`)
-wide_plant_traits$`Plant Height` <- as.numeric(wide_plant_traits$`Plant Height`)
-wide_plant_traits$`Dispersal Syndrome` <- as.character(wide_plant_traits$`Dispersal Syndrome`)
-wide_plant_traits$`Growth Forme` <- as.character(wide_plant_traits$`Growth Form`)
-wide_plant_traits$`Fruit Type` <- as.character(wide_plant_traits$`Fruit Type`)
-```
-
-# Summary
-```{r}
-print("Number of species: ")
-length(unique(wide_plant_traits$species))
-```
-
-# Write data to csv
-```{r}
-write.csv(wide_plant_traits, file.path(output_path,"TropicalAndes_all_plant_traits_harmonized.csv"))
-write.csv(clean_traits, file.path(output_path,"TropicalAndes_all_plant_traits_harmonized_with_units.csv"))
-```

### 2024-05-29 code/L1/L1_P3_TropicalAndes_plant_traits.Rmd
move code to new script

--- 2024-05-22 code/L1/L1_P3_TropicalAndes_plant_traits.Rmd
@@
 
-# Read in data
+# Read in subset fruiting species data
 ```{r}
@@
 
-# Get species list from plant_occ data
-```{r}
-plant_species <- unique(plant_occ$Accepted_species)
-length(plant_species)
-```
-
-# Subset TRY, BIEN, and GIFT by species list
-## TRY
-```{r}
-nrow(TRY_traits)
-TRY_filtered <- TRY_traits %>%
-  filter(Accepted_species %in% plant_species)
-nrow(TRY_filtered)
-```
-
-```{r}
-TRY_traits %>%
-  count(TraitName)
-```
-
-```{r}
-TRY_traits %>%
-  distinct(Accepted_species, TraitName) %>%
-  count(TraitName)
-```
-
-## BIEN
-```{r}
-nrow(BIEN_traits)
-BIEN_filtered <- BIEN_traits %>%
-  filter(Accepted_species %in% plant_species)
-nrow(BIEN_filtered)
-```
-
-```{r}
-BIEN_traits %>%
-  count(trait_name)
-```
-
-```{r}
-BIEN_traits %>%
-  distinct(Accepted_species, trait_name) %>%
-  count(trait_name)
-```
-
-## GIFT
-```{r}
-nrow(GIFT_traits)
-GIFT_filtered <- GIFT_traits %>%
-  filter(Accepted_species %in% plant_species)
-nrow(GIFT_filtered)
-```
-
-```{r}
-GIFT_traits %>%
-  count(trait_name)
-```
-
-```{r}
-GIFT_traits %>%
-  distinct(Accepted_species, trait_name) %>%
-  count(trait_name)
-```
-
-# Combine trait information from TRY, BIEN, GIFT
-
-```{r}
-TRYTraitTypes <- TRY_filtered %>%
-  group_by(TraitName) %>%
-  distinct(DataName)
-```
-
-```{r}
-unique(TRYTraitTypes$TraitName)
-```
-
-```{r}
-colnames(TRY_filtered)
-```
-columns to keep : Accepted_species, TraitName, OrigValueStr
-
-```{r}
-TRY_filtered_subset <- TRY_filtered[ , c("Accepted_species", "TraitName", "OrigValueStr", "UnitName")]
-colnames(TRY_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue", "Unit")
-```
-
-```{r}
-# save sources/references
-TRY_references <- unique(TRY_filtered[ , c("Dataset", "Reference")])
-write.csv(TRY_references, file.path(output_path,"TRY_references.csv"))
-```
-
-```{r}
-unique(BIEN_filtered$trait_name)
-```
-
-```{r}
-colnames(BIEN_filtered)
-```
-columns to keep : Accepted_species, trait_name, trait_value, unit
-
-```{r}
-BIEN_filtered_subset <- BIEN_filtered[ , c("Accepted_species", "trait_name", "trait_value", "unit")]
-colnames(BIEN_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue", "Unit")
-```
-
-```{r}
-# save sources/references
-BIEN_references <- unique(BIEN_filtered[ , c("url_source", "source_citation")])
-write.csv(BIEN_references, file.path(output_path,"BIEN_references.csv"))
-```
-
-```{r}
-unique(GIFT_filtered$trait_name)
-```
-
-```{r}
-colnames(GIFT_filtered)
-```
-columns to keep : Accepted_species, trait_value, trait_name
-
-```{r}
-GIFT_filtered_subset <- GIFT_filtered[ , c("Accepted_species", "trait_name", "trait_value")]
-colnames(GIFT_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue")
-```
-
-Add Unit column to GIFT dataframe
-```{r}
-GIFT_filtered_subset <- GIFT_filtered_subset %>%
-  mutate(Unit = case_when(
-  TraitName == "plant_height" ~ "m",
-  TraitName == "plant_height_max" ~ "m",
-  TraitName == "plant_height_min" ~ "m",
-  TraitName == "seed_mass_mean" ~ "g",
-  TraitName == "seed_mass_min" ~ "g",
-  TraitName == "seed_mass_max" ~ "g",
-  TraitName == "fruit_length_min" ~ "cm",
-  TraitName == "fruit_length_max" ~ "cm",
-  TraitName == "fruit_length_mean" ~ "cm",
-  TraitName == "seed_length_max" ~ "mm",
-  TraitName == "seed_length_min" ~ "mm",
-  TraitName == "seed_length_mean" ~ "mm",
-  TraitName == "seed_width_max" ~ "mm",
-  TraitName == "seed_width_min" ~ "mm",
-  TraitName == "seed_width_mean" ~ "mm",
-  TraitName == "plant_lifespan" ~ "years",
-  .default = ""
-  ))
-```
-
-```{r}
-traits <- rbind(TRY_filtered_subset, BIEN_filtered_subset, GIFT_filtered_subset)
-nrow(traits)
-```
-
-```{r}
-# remove rows without accepted species names
-traits_clean <- traits[!(is.na(traits$Accepted_species) | traits$Accepted_species==""), ]
-nrow(traits_clean)
-```
-
+# Look at trait numbers
 ```{r}
@@
   count(TraitName)
-```
-
-```{r}
-# Group by TraitName and summarize the unique TraitValues
-unique_trait_values <- traits %>%
-  group_by(TraitName) %>%
-  summarize(UniqueValues = list(unique(TraitValue)))%>%
-  unnest(UniqueValues)
 ```
@@
 
-## Fruit type
-```{r}
-traits_fruitType <- traits %>%
-  filter(grepl('Fruit type|fruit type|fruit_type', TraitName))
-traits_fruitType$OrigTraitValue <- traits_fruitType$TraitValue
-
-nrow(traits_fruitType)
-length(unique(traits_fruitType$Accepted_species))
-```
-
-```{r}
-traits_fruitType <- traits_fruitType %>%
-  filter(is.na(as.numeric(TraitValue)))
-
-nrow(traits_fruitType)
-length(unique(traits_fruitType$Accepted_species))
-```
-
-```{r}
-traits_fruitType %>%
-  distinct(TraitValue)
-```
-
-```{r}
-# standardize trait values
-traits_fruitType <- traits_fruitType %>%
-  filter() %>%
-  mutate(TraitValue= case_when(
-    grepl('drupe|ectozoochorous|zoochorous', OrigTraitValue, ignore.case = TRUE) ~ "zoochorous",
-    grepl('capsule|Capsule|Capsule |', OrigTraitValue, ignore.case = TRUE) ~ "myrmecochorous",
-  )) %>%
-  filter(!is.na(TraitValue))
-```
-
-```{r}
-# standardize trait name
-traits_fruitType <- traits_fruitType %>%
-  mutate(TraitName = "FruitType")
-```
-
-```{r}
-averagetraits_fruitType <- factor_data_merge(traits_fruitType, "Fruit Type")
-```
-
-```{r}
-nrow(averagetraits_fruitType)
-length(unique(averagetraits_fruitType$Accepted_species))
-```
-
-```{r}
-summary(averagetraits_fruitType)
-```
-
 ## Seed mass

### 2024-05-29 code/L1/L1_P5_TropicalAndes_plant_imputetraits.Rmd
rename L1 plant scripts to update number order

--- 2024-05-09 code/L1/L1_P4_TropicalAndes_plant_imputetraits.Rmd
@@
 plant_traits <- plant_traits[, !colnames(plant_traits) %in% "X", drop = FALSE]
+# remove Growth Forme trait
+plant_traits <- plant_traits %>%
+  select(-Growth.Forme)
 # Convert the wide dataframe to long format
@@
 ## rename trait_value column so it's consistent with other traits
-Fruit.Type_families_BIEN_df$trait_value <- Fruit.Type_families_BIEN_df$trait_value_mode_trait
-Growth.Form_families_BIEN_df$trait_value <- Growth.Form_families_BIEN_df$trait_value_mode_trait
+Fruit.Type_families_BIEN_df$TraitValue <- Fruit.Type_families_BIEN_df$trait_value_mode_trait
+Growth.Form_families_BIEN_df$TraitValue <- Growth.Form_families_BIEN_df$trait_value_mode_trait
 
 ## convert numeric trait values into character
-Plant.Height_families_BIEN_df$trait_value <- as.character(Plant.Height_families_BIEN_df$trait_value)
-Fruit.Length_families_BIEN_df$trait_value <- as.character(Fruit.Length_families_BIEN_df$trait_value)
+Plant.Height_families_BIEN_df$TraitValue <- as.character(Plant.Height_families_BIEN_df$TraitValue)
+Fruit.Length_families_BIEN_df$TraitValue <- as.character(Fruit.Length_families_BIEN_df$TraitValue)
 
 all_family_traits <- bind_rows(
-  select(Dispersal.Syndrome_families_BIEN_df, scrubbed_family, trait_name, trait_value),
-  select(Fruit.Length_families_BIEN_df, scrubbed_family, trait_name, trait_value),
-  select(Plant.Height_families_BIEN_df, scrubbed_family, trait_name, trait_value),
-  select(Fruit.Type_families_BIEN_df, scrubbed_family, trait_name, trait_value),
-  select(Growth.Form_families_BIEN_df, scrubbed_family, trait_name, trait_value)
+  select(Dispersal.Syndrome_families_BIEN_df, scrubbed_family, trait_name, TraitValue),
+  select(Fruit.Length_families_BIEN_df, scrubbed_family, trait_name, TraitValue),
+  select(Plant.Height_families_BIEN_df, scrubbed_family, trait_name, TraitValue),
+  select(Fruit.Type_families_BIEN_df, scrubbed_family, trait_name, TraitValue),
+  select(Growth.Form_families_BIEN_df, scrubbed_family, trait_name, TraitValue)
   )
@@
 
-## rename trait_value column so it's consistent with other traits
-Fruit.Type_genus_BIEN_df$trait_value <- Fruit.Type_genus_BIEN_df$trait_value_mode_trait
-Growth.Form_genus_BIEN_df$trait_value <- Growth.Form_genus_BIEN_df$trait_value_mode_trait
+## rename TraitValue column so it's consistent with other traits
+Fruit.Type_genus_BIEN_df$TraitValue <- Fruit.Type_genus_BIEN_df$trait_value_mode_trait
+Growth.Form_genus_BIEN_df$TraitValue <- Growth.Form_genus_BIEN_df$trait_value_mode_trait
 
 ## convert numeric trait values into character
-Plant.Height_genus_BIEN_df$trait_value <- as.character(Plant.Height_genus_BIEN_df$trait_value)
-Fruit.Length_genus_BIEN_df$trait_value <- as.character(Fruit.Length_genus_BIEN_df$trait_value)
+Plant.Height_genus_BIEN_df$TraitValue <- as.character(Plant.Height_genus_BIEN_df$TraitValue)
+Fruit.Length_genus_BIEN_df$TraitValue <- as.character(Fruit.Length_genus_BIEN_df$TraitValue)
 
 all_genus_traits <- bind_rows(
-  select(Dispersal.Syndrome_genus_BIEN_df, scrubbed_genus, trait_name, trait_value),
-  select(Fruit.Length_genus_BIEN_df, scrubbed_genus, trait_name, trait_value),
-  select(Plant.Height_genus_BIEN_df, scrubbed_genus, trait_name, trait_value),
-  select(Fruit.Type_genus_BIEN_df, scrubbed_genus, trait_name, trait_value),
-  select(Growth.Form_genus_BIEN_df, scrubbed_genus, trait_name, trait_value)
+  select(Dispersal.Syndrome_genus_BIEN_df, scrubbed_genus, trait_name, TraitValue),
+  select(Fruit.Length_genus_BIEN_df, scrubbed_genus, trait_name, TraitValue),
+  select(Plant.Height_genus_BIEN_df, scrubbed_genus, trait_name, TraitValue),
+  select(Fruit.Type_genus_BIEN_df, scrubbed_genus, trait_name, TraitValue),
+  select(Growth.Form_genus_BIEN_df, scrubbed_genus, trait_name, TraitValue)
   )
@@
 all_genus_traits$TraitLevel <- "genus"
+```
+
+rename traits to match
+```{r}
+# Replace trait names to match
+long_plant_traits_tax <- long_plant_traits_tax %>%
+  mutate(TraitName = ifelse(TraitName == "Plant.Height", "Plant.Height_m", TraitName)) %>%
+  mutate(TraitName = ifelse(TraitName == "Fruit.Length", "Fruit.Length_mm", TraitName))
+unique(long_plant_traits_tax$TraitName)
+
+na_traits_family.genus <- na_traits_family.genus %>%
+  mutate(TraitName = ifelse(TraitName == "Plant.Height", "Plant.Height_m", TraitName)) %>%
+  mutate(TraitName = ifelse(TraitName == "Fruit.Length", "Fruit.Length_mm", TraitName))
+unique(na_traits_family.genus$TraitName)
 ```
@@
 colnames(long_plant_traits_tax)
+colnames(na_traits_family.genus)
 ```
@@
 
-use family & genus level traits to fill in gaps of original dataframe
+use family & genus level traits to fill in gaps of na_traits_family.genus
 genus level prefered over family
@@
 # Join the dataframes based on genus and TraitName
-filled_long_plant_traits_tax <- long_plant_traits_tax %>%
+filled_na_traits_family.genus <- na_traits_family.genus %>%
   left_join(all_genus_traits, by = c("genus", "TraitName"), relationship = "many-to-many") %>%
-  # Fill NAs in TraitValue column with values from all_genus_traits
-  mutate(TraitValue = ifelse(is.na(TraitValue.x), TraitValue.y, TraitValue.x)) %>%
-  # Remove the columns from all_genus_traits (if needed)
-  select(-TraitValue.y)
+  # Fill in TraitValue column with values from all_genus_traits
+  mutate(TraitValue = coalesce(TraitValue.x, TraitValue.y)) %>%
+  # Remove the columns
+  select(-TraitValue.y, -TraitValue.x)
 ```
 
-Add units to wide trait table numeric trait names
-fruit mass -> Fruit.Mass_mg
-fruit length -> Fruit.Length_mm
-plant height -> Plant.Height_m
-plant lifespan -> Plant.Lifespan_years
-seed mass -> Seed.Mass_mg
-seed length -> Seed.Length_mm
-seed width -> Seed.Width_mm
+```{r}
+na_traits_after_genus <- filled_na_traits_family.genus %>%
+  filter(is.na(TraitValue))
+filled_traits_with_genus <- filled_na_traits_family.genus %>%
+  filter(!is.na(TraitValue))
+```
+
+```{r}
+# Join the dataframes based on family and TraitName
+## remove trait level column from na_traits_after_genus
+na_traits_after_genus <- na_traits_after_genus %>%
+  select(-TraitLevel)
+filled_na_traits_family <- na_traits_after_genus %>%
+  left_join(all_family_traits, by = c("family", "TraitName"), relationship = "many-to-many") %>%
+  # Fill in TraitValue column with values from all_genus_traits
+  mutate(TraitValue = coalesce(TraitValue.x, TraitValue.y)) %>%
+  # Remove the columns
+  select(-TraitValue.y, -TraitValue.x)
+```
+
+```{r}
+filled_traits_with_family <- filled_na_traits_family %>%
+  filter(!is.na(TraitValue))
+na_traits_after_family <- filled_na_traits_family %>%
+  filter(is.na(TraitValue))
+```
+
+```{r}
+nonimputed_traits <- long_plant_traits_tax %>%
+  filter(!is.na(TraitValue))
+```
+
+combine filled_traits_with_genus, filled_traits_with_family, and nonimputed_traits
+```{r}
+colnames(filled_traits_with_genus)
+colnames(filled_traits_with_family)
+colnames(nonimputed_traits)
+```
+
+```{r}
+all_traits_with_tax <- rbind(filled_traits_with_genus, filled_traits_with_family, nonimputed_traits)
+```
+
+```{r}
+all_traits_with_tax %>%
+  group_by(TraitName) %>%
+  summarise(num_species = n_distinct(species))
+```
+
+```{r}
+na_traits_after_family %>%
+  group_by(TraitName) %>%
+  summarise(num_species = n_distinct(species))
+```
+
+add NAs back to all traits
+```{r}
+colnames(all_traits_with_tax)
+colnames(na_traits_after_family)
+```
+
+```{r}
+all_traits_with_NAs <- rbind(all_traits_with_tax, na_traits_after_family)
+```
+
+### Save long dataframe
+```{r}
+write.csv(all_traits_with_NAs, file.path(output_path,"TropicalAndes_all_plant_traits_filled_with_family_genus_long.csv"))
+```
+
+## Convert long dataframe to wide
+```{r}
+#remove unused columns
+all_traits_with_NAs_long <- all_traits_with_NAs %>%
+  select(-TraitLevel, -family, -genus)
+wide_traits <- pivot_wider(all_traits_with_NAs_long, names_from = TraitName, values_from = TraitValue)
+```
+make NA if:
+- Growth.Form is numeric
+- Dispersal.Syndrome is numeric or contains multiple values
+- Plant.Height = 0
+- Fruit.Type is numeric
+```{r}
+library(stringr)
+wide_traits$Growth.Form <- ifelse(is.numeric(wide_traits$Growth.Form), NA, wide_traits$Growth.Form)
+wide_traits$Dispersal.Syndrome[stringr::str_detect(wide_traits$Dispersal.Syndrome, ",")] <- NA
+
+# Convert Plant.Height_m to numeric 
+wide_traits$Plant.Height_m[wide_traits$Plant.Height_m == "0"] <- NA
+
+wide_traits$Fruit.Type <- ifelse(is.numeric(wide_traits$Fruit.Type), NA, wide_traits$Fruit.Type)
+```
+
+```{r}
+wide_traits <- wide_traits %>%
+  mutate_all(~sapply(., paste, collapse = ", "))
+
+# Replace values with multiple values with NA
+wide_traits <- wide_traits %>%
+  mutate(Dispersal.Syndrome = if_else(str_count(Dispersal.Syndrome, ", ") > 0, NA_character_, Dispersal.Syndrome), 
+         Growth.Form = if_else(str_count(Growth.Form, ", ") > 0, NA_character_, Growth.Form),
+         Plant.Height_m = if_else(str_count(Plant.Height_m, ", ") > 0, NA_character_, Plant.Height_m), 
+         Fruit.Length_mm = if_else(str_count(Fruit.Length_mm, ", ") > 0, NA_character_, Fruit.Length_mm), 
+         Fruit.Type = if_else(str_count(Fruit.Type, ", ") > 0, NA_character_, Fruit.Type))
+```
+
+```{r}
+wide_traits$Plant.Height_m <- unlist(wide_traits$Plant.Height_m)
+wide_traits$Growth.Form <- unlist(wide_traits$Growth.Form)
+wide_traits$Dispersal.Syndrome <- unlist(wide_traits$Dispersal.Syndrome)
+wide_traits$Fruit.Length_mm <- unlist(wide_traits$Fruit.Length_mm)
+wide_traits$Fruit.Type <- unlist(wide_traits$Fruit.Type)
+
+#convert column types from list to character & numeric
+wide_traits$Plant.Height_m <- as.numeric(wide_traits$Plant.Height_m)
+wide_traits$Growth.Form <- as.character(wide_traits$Growth.Form)
+wide_traits$Dispersal.Syndrome <- as.character(wide_traits$Dispersal.Syndrome)
+wide_traits$Fruit.Length_mm <- as.numeric(wide_traits$Fruit.Length_mm)
+wide_traits$Fruit.Type <- as.character(wide_traits$Fruit.Type)
+```
 
@@
 
-```{r}
-#replace " " with _ in column names
-names(plant_traits) <- sub(" ", "_", names(plant_traits))
-```
-
 # show the missing data pattern
 ```{r}
-md.pattern(plant_traits)
+md.pattern(wide_traits)
 ```
-# Trying to prevent error
-Error in edit.setup(data, setup, ...) : 
-  `mice` detected constant and/or collinear variables. No predictors were left after their removal.
-```{r}
-# Subset to numeric plant traits
-num_plant_traits <- plant_traits %>%
-  select_if(is.numeric)
-# Check for constant variables
-constant_vars <- sapply(num_plant_traits, function(x) length(unique(x))) == 1
-constant_vars <- names(constant_vars[constant_vars])
-
-# Remove constant variables
-num_plant_traits <- num_plant_traits[, !names(num_plant_traits) %in% constant_vars]
-
-# Check for collinearity
-cor_matrix <- cor(num_plant_traits)
-
-# Find highly correlated variables
-highly_correlated_pairs <- which(abs(cor_matrix) > 0.8 & upper.tri(cor_matrix, diag = TRUE), arr.ind = TRUE)
-
-# Print highly correlated pairs
-print(highly_correlated_pairs)
-```
-
 # Perform the imputation
 ```{r}
-imp_model <- mice(plant_traits, method = "cart", maxit = 20)
+imp_model <- mice(wide_traits, method = "cart", maxit = 20)
 ```
@@
 ```
+
+remove dispersal syndrome because I need to fix it 
+```{r}
+imputed_data <- imputed_data %>%
+  select(-Dispersal.Syndrome)
+```
+
+```{r}
+imputed_data %>%
+  summarise_at(vars(Growth.Form:Fruit.Type), ~ sum(is.na(.)))
+```
+
+# Save imputed data
+```{r}
+write.csv(imputed_data, file.path(output_path,"TropicalAndes_imputed_plant_traits.csv"))
+```

### 2024-05-29 code/L1/L1_P6_TropicalAndes_plant_occ.Rmd
rename L1 plant scripts to update number order

---
title: "Plant occurrence subset by species with complete trait records"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script subsets plant occurrence data to species with complete trait coverage."
data input: "ADD"
data output: "ADD"
date: "2023-10-04"
output: html_document
---

# Set file paths
```{r}
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path_L1 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(dplyr)
#run if not installed remotes::install_github("FRBCesab/funbiogeo")
library(funbiogeo)
```

# Read in data
```{r}
plant_traits <- read.csv(file.path(data_path_L1, file = "TropicalAndes_imputed_plant_traits.csv"))
plant_occ <- read.csv(file.path(data_path_L1, file = "TropicalAndes_GBIF_plant_occ_cleaned.csv"))
```

# Check trait coverage
```{r}
summary(plant_traits)
```

```{r}
fb_plot_number_species_by_trait(plant_traits)
```

```{r}
Plant_complete <- fb_filter_species_by_trait_coverage(plant_traits, threshold_traits_proportion =  1)
```

# Create list of species with complete trait records
```{r}
plant_species <- unique(Plant_complete$species)
```

# Subset frugivore occurrence data by species list
```{r}
plant_occ_subset <- plant_occ %>%
  filter(species %in% plant_species)
```

# Summary
```{r}
glimpse(plant_occ_subset)
```

```{r}
source("~/GitHub/neotropical_plants/functions/data_summary.R")
```

```{r}
data_summary(plant_occ_subset, plant_occ_subset$species, plant_occ_subset$genus, plant_occ_subset$family)
```

# Write data to csv
```{r}
write.csv(plant_occ_subset, file.path(output_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned_subset.csv"))
```

### 2024-05-30 code/L1/L1_P3_TropicalAndes_plant_fruiting_subset.Rmd
add subset of trait and occurrence data by fruitng species list and write resulting data to csv

@@
 overview: "This script subsets plant occurrence and trait data by fruiting plant species."
-data input: "ADD"
+data input: "TropicalAndes_GBIF_plant_occ_harmonized.csv, TropicalAndes_TRY_plant_traits_harmonized.csv,
+             TropicalAndes_BIEN_plant_traits_harmonized.csv, TropicalAndes_GIFT_plant_traits_harmonized.csv"
 data output: "ADD"
@@
 ```{r}
-traits %>%
+traits_clean %>%
   count(TraitName)
@@
 ```{r}
-traits %>%
+traits_clean %>%
   distinct(Accepted_species, TraitName) %>%
   count(TraitName)
-```
-
-```{r}
-# Group by TraitName and summarize the unique TraitValues
-unique_trait_values <- traits %>%
-  group_by(TraitName) %>%
-  summarize(UniqueValues = list(unique(TraitValue)))%>%
-  unnest(UniqueValues)
 ```
@@
 ```{r}
-traits_fruitType <- traits %>%
+traits_fruitType <- traits_clean %>%
   filter(grepl('Fruit type|fruit type|fruit_type', TraitName))
@@
     grepl('capsule|Capsule|Capsule ', OrigTraitValue, ignore.case = TRUE) ~ "capsule",
-    grepl('follicle|Follicle', OrigTraitValue, ignore.case = TRUE) ~ "follicle",
+    grepl('drupe|Drupe|Drupe |Drupaceous', OrigTraitValue, ignore.case = TRUE) ~ "drupe",
+    grepl('aggregate berries', OrigTraitValue, ignore.case = TRUE) ~ "aggregate berries",
+    grepl('aggregate drupelets', OrigTraitValue, ignore.case = TRUE) ~ "aggregate drupelets",
+    grepl('fleshy', OrigTraitValue, ignore.case = TRUE) ~ "fleshy",
+    grepl('pome', OrigTraitValue, ignore.case = TRUE) ~ "pome",
+    grepl('Squash', OrigTraitValue, ignore.case = TRUE) ~ "squash",
   )) %>%
   filter(!is.na(TraitValue))
+```
+
+```{r}
+traits_fruitType %>%
+  count(TraitValue)
 ```
@@
 ```{r}
+averagetraits_fruitType$TraitValue <- as.factor(averagetraits_fruitType$TraitValue)
 summary(averagetraits_fruitType)
 ```
+
+```{r}
+fruiting_species <- averagetraits_fruitType$Accepted_species
+```
+
+# Subset trait data and occurrence data by fruiting species list
+```{r}
+nrow(plant_occ)
+plant_occ_subset <- plant_occ %>%
+  filter(Accepted_species %in% fruiting_species)
+nrow(plant_occ_subset)
+```
+
+```{r}
+nrow(traits_clean)
+traits_subset <- traits_clean%>%
+  filter(Accepted_species %in% fruiting_species)
+nrow(traits_subset)
+```
+
+```{r}
+traits_subset %>%
+  distinct(Accepted_species, TraitName) %>%
+  count(TraitName)
+```
+
+# Write data to csv
+```{r}
+write.csv(plant_occ_subset, file.path(output_path, "TropicalAndes_GBIF_plant_occ_harmonized_subset.csv"))
+write.csv(traits_subset, file.path(output_path, "TropicalAndes_all_plant_traits_harmonized_subset.csv"))
+```

### 2024-05-30 code/L1/L1_P4_TropicalAndes_plant_traits.Rmd
add new packages, update input files, modify functions to include record count and variance, include summary figures to see trait coverage before imputation

--- 2024-05-29 code/L1/L1_P3_TropicalAndes_plant_traits.Rmd
@@
 library(tidyr)
+library(purrr)
+library(funbiogeo)
+library(visdat)
 ```
@@
 output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
 ```
@@
 ```{r}
-plant_occ <- read.csv(file.path(data_path_L1, "TropicalAndes_GBIF_plant_occ_harmonized.csv"))
-TRY_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_TRY_plant_traits_harmonized.csv"))
-BIEN_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_BIEN_plant_traits_harmonized.csv"))
-GIFT_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_GIFT_plant_traits_harmonized.csv"))
+plant_traits <- read.csv(file.path(data_path_L1, "TropicalAndes_all_plant_traits_harmonized_subset.csv"))
 ```
@@
 ```{r}
-traits %>%
+plant_traits %>%
   count(TraitName)
@@
 ```{r}
-traits %>%
+plant_traits %>%
   distinct(Accepted_species, TraitName) %>%
@@
 - Fruit length
-- Fruit type
 - Plant growth form
@@
 ```{r}
-# Numeric data
 numeric_data_average <- function(data, unit, traitname) {
-# remove rows if TraitValue is non-numeric
+  # Remove rows if TraitValue is non-numeric and convert to numeric
   data <- data %>%
@@
     mutate(TraitValue = as.numeric(TraitValue))
-# remove duplicate rows
+
+  # Remove duplicate rows
   data <- data %>%
     distinct(.keep_all = TRUE)
-# calculate average value per trait per species
-average_data <- aggregate(TraitValue ~ Accepted_species, data, FUN = mean, na.rm = TRUE)
+
   # Remove rows with missing or empty species names
-average_data <- average_data[!is.na(average_data$Accepted_species) & average_data$Accepted_species !="", ]
-# Add unit & trait column
-average_data$Unit <- unit
-average_data$TraitName <- traitname
+  data <- data[!is.na(data$Accepted_species) & data$Accepted_species != "", ]
 
-return(average_data)
+  # Calculate average value, count, and variance per trait per species
+  summary_data <- data %>%
+    group_by(Accepted_species) %>%
+    summarise(
+      TraitValue_mean = mean(TraitValue, na.rm = TRUE),
+      records_used = n(),
+      variance = var(TraitValue, na.rm = TRUE)
+    )
+
+  # Add trait name column
+  summary_data$TraitName <- traitname
+
+  # Rename the mean column to TraitValue
+  summary_data <- summary_data %>%
+    rename(TraitValue = TraitValue_mean)
+
+  return(summary_data)
 }
+
 ```
@@
 ```{r}
-# categorical data
 factor_data_merge <- function(data, traitname) {
-# remove duplicate rows
-data <- data %>%
-  distinct(.keep_all = TRUE)
-# remove rows with missing or empty species names
+
+  # Remove rows with missing or empty species names
   data <- data[!is.na(data$Accepted_species) & data$Accepted_species != "", ]
+
   # Subset data to just species and TraitValue
-data <- data[,c("Accepted_species", "TraitValue")]
-# Find the most common trait value for each species
-  most_common_values <- aggregate(. ~ Accepted_species, data, function(x) {
-    # Find the mode (most common value) of the trait values
-    mode_value <- names(sort(table(x), decreasing = TRUE))[1]
-    return(mode_value)
-  })
-# Merge most common trait values back to the original data
-merged_data <- merge(data, most_common_values, by = "Accepted_species", all.x = TRUE)
+  data <- data[, c("Accepted_species", "TraitValue")]
 
-# Rename the merged column
-colnames(merged_data)[ncol(merged_data)] <- "TraitValue"
+  # Calculate mode, count, and variance-like measure (using mode frequency)
+  summary_data <- data %>%
+    group_by(Accepted_species) %>%
+    summarise(
+      ModeValue = names(sort(table(TraitValue), decreasing = TRUE))[1],
+      records_used = n(),
+      mode_freq = max(table(TraitValue)),
+      variance = 1 - (max(table(TraitValue)) / n())
+    )
 
-# Remove column TraitValue.x
-merged_data <- merged_data[, !colnames(merged_data) %in% "TraitValue.x", drop = FALSE]
+  # Add trait name column
+  summary_data$TraitName <- traitname
 
-# remove duplicate rows
-merged_data <- merged_data %>%
-  distinct(.keep_all = TRUE)
+  # Rename the mean column to TraitValue
+  summary_data <- summary_data %>%
+    rename(TraitValue = ModeValue)
 
-# Add unit & trait column
-merged_data$TraitName <- traitname
-
-  return(merged_data)
+  return(summary_data)
 }
@@
 ```{r}
-traits_dispersal <- traits %>%
+traits_dispersal <- plant_traits %>%
   filter(grepl('Dispersal syndrome|whole plant dispersal syndrome|dispersal_syndrome_1|dispersal_syndrome_2', TraitName))
@@
 ```{r}
-traits_fruitMass <- traits %>%
+traits_fruitMass <- plant_traits %>%
   filter(grepl('Fruit dry mass', TraitName))
@@
 ```{r}
-traits_fruitLength <- traits %>%
+traits_fruitLength <- plant_traits %>%
   filter(grepl('Fruit length|maximum fruit length|minimum fruit length|fruit_length_min|fruit_length_max|fruit_length_mean', TraitName))
@@
 ```{r}
-traits_plantHeight <- traits %>%
+traits_plantHeight <- plant_traits %>%
   filter(grepl('Plant height generative|Plant height vegetative|maximum whole plant height|minimum whole plant height|plant_height|plant_height_max|plant_height_min|whole plant height', TraitName))
@@
 ## Growth form
+Just using data from GIFT 
 ```{r}
-traits_growthForm <- traits %>%
-  filter(grepl('Plant growth form|growth_form|whole plant growth form'
-, TraitName))
+traits_growthForm <- plant_traits %>%
+  filter(TraitName == 'growth_form')
 traits_growthForm$OrigTraitValue <- traits_growthForm$TraitValue
@@
 ```{r}
-averagetraits_growthForm <- factor_data_merge(traits_growthForm, "Growth Form")
+# standardize trait names with units
+traits_growthForm <- traits_growthForm %>%
+  mutate(TraitName = "GrowthForm")
+```
+
+```{r}
+averagetraits_growthForm <- factor_data_merge(traits_growthForm, "GrowthForm")
+```
+
+```{r}
+nrow(averagetraits_growthForm)
+length(unique(averagetraits_growthForm$Accepted_species))
+```
+
+```{r}
+summary(averagetraits_growthForm)
 ```
@@
 ```{r}
-# add unit column to categorical traits
-cat_traits <- rbind(averagetraits_dispersal, averagetraits_fruitType, averagetraits_growthForm)
-cat_traits$Unit <- ""
-clean_traits <- rbind(averagetraits_fruitLength, averagetraits_fruitMass, averagetraits_plantHeight, averagetraits_plantLifespan, averagetraits_seedLength, averagetraits_seedMass, averagetraits_seedWidth, cat_traits)
+trait_dfs <- list(averagetraits_fruitLength, averagetraits_fruitMass, averagetraits_plantHeight, averagetraits_plantLifespan, averagetraits_seedLength, averagetraits_seedMass, averagetraits_seedWidth, averagetraits_dispersal, averagetraits_growthForm)
+
+# Function to combine trait dataframes only matching columns
+combine_matching_columns <- function(df_list) {
+  # Find the common columns across all data frames
+  common_cols <- Reduce(intersect, lapply(df_list, colnames))
+
+  # Ensure TraitValue column is of the same type across all data frames
+  df_list <- lapply(df_list, function(df) {
+    if ("TraitValue" %in% colnames(df)) {
+      df <- df %>% mutate(TraitValue = as.character(TraitValue))
+    }
+    return(df)
+  })
+
+  # Filter each data frame to keep only the common columns
+  df_list_filtered <- lapply(df_list, function(df) {
+    df %>% select(all_of(common_cols))
+  })
+
+  # Combine the data frames using reduce and full_join
+  combined_df <- reduce(df_list_filtered, full_join, by = common_cols)
+
+  return(combined_df)
+}
+clean_traits <- combine_matching_columns(trait_dfs)
 nrow(clean_traits)
@@
 
-# Summary
-```{r}
-print("Number of records: ")
-nrow(clean_traits)
-print("Number of species: ")
-length(unique(clean_traits$Accepted_species))
-```
-
-# Get one trait value per species
-```{r}
-#remove unit column from dataframe
-clean_traits <- clean_traits[, !colnames(clean_traits) %in% "Unit", drop = FALSE]
-```
-
-```{r}
-# separate numeric and non-numeric traits
-numeric_traits <- clean_traits[clean_traits$TraitName %in% c("Fruit Length", "Fruit Mass", "Plant Height", "Plant Lifespan", "Seed Length", "Seed Width", "Seed Mass"), ]
-cat_traits <- clean_traits[clean_traits$TraitName %in% c("Dispersal Syndrome", "Fruit Type", "Growth Form"), ]
-
-numeric_traits$TraitValue <- as.numeric(numeric_traits$TraitValue)
-```
-
-```{r}
-# numeric traits
-## Group by species and TraitName, and calculate the mean TraitValue for each combination
-numeric_traits_combined <- numeric_traits %>%
-  group_by(species, TraitName) %>%
-  summarise(TraitValue = mean(TraitValue, na.rm = TRUE)) %>%
-  ungroup()
-
-# non-numeric traits
-cat_traits_combined <- cat_traits %>%
-  group_by(species) %>%
-  summarise_all(.funs = list(mode_trait = function(x) {
-    mode_value <- names(sort(table(x), decreasing = TRUE))[1]  # Calculate mode
-    return(mode_value)
-  }))
-```
-
-```{r}
-#rename columns to original values
-colnames(cat_traits_combined)[which(colnames(cat_traits_combined) == "TraitValue_mode_trait")] <- "TraitValue"
-colnames(cat_traits_combined)[which(colnames(cat_traits_combined) == "TraitName_mode_trait")] <- "TraitName"
-
-# Remove rows with "?" or numbers in the TraitValue column
-cleaned_cat_traits <- cat_traits_combined %>%
-  filter(!grepl("\\?", TraitValue) & !grepl("\\d", TraitValue))
-```
-
-```{r}
-#combine numeric and non-numeric traits
-combined_clean_traits <- rbind(cat_traits_combined, numeric_traits_combined)
-```
-
-```{r}
-combined_clean_traits %>%
-  count(TraitName)
-```
-
-# Remove traits with less than 10 records
-```{r}
-combined_clean_traits <- combined_clean_traits[!(combined_clean_traits$TraitName %in% c("Fruit Mass", "Seed Mass")), ]
-```
-
 # Transform plant dataframe from long to wide
 ```{r}
-wide_traits <- spread(combined_clean_traits, key = TraitName, value = TraitValue)
+wide_traits <- spread(clean_traits, key = TraitName, value = TraitValue)
 
 #remove X column from dataframe
-wide_plant_traits <- wide_traits[, !colnames(wide_traits) %in% "X", drop = FALSE]
+wide_plant_traits <- wide_traits[, !colnames(wide_traits) %in% c("X", "records_used", "variance"), drop = FALSE]
 ```
@@
 # Assign data types
-wide_plant_traits$`Fruit Length` <- as.numeric(wide_plant_traits$`Fruit Length`)
-wide_plant_traits$`Plant Height` <- as.numeric(wide_plant_traits$`Plant Height`)
-wide_plant_traits$`Dispersal Syndrome` <- as.factor(wide_plant_traits$`Dispersal Syndrome`)
-wide_plant_traits$`Growth Form` <- as.factor(wide_plant_traits$`Growth Form`)
-wide_plant_traits$`Fruit Type` <- as.factor(wide_plant_traits$`Fruit Type`)
+wide_plant_traits$`FruitLength_mm` <- as.numeric(wide_plant_traits$`FruitLength_mm`)
+wide_plant_traits$`FruitMass_mg` <- as.numeric(wide_plant_traits$`FruitMass_mg`)
+wide_plant_traits$`PlantHeight_m` <- as.numeric(wide_plant_traits$`PlantHeight_m`)
+wide_plant_traits$`PlantLifespan_years` <- as.numeric(wide_plant_traits$`PlantLifespan_years`)
+wide_plant_traits$`SeedLength_mm` <- as.numeric(wide_plant_traits$`SeedLength_mm`)
+wide_plant_traits$`SeedMass_g` <- as.numeric(wide_plant_traits$`SeedMass_g`)
+wide_plant_traits$`SeedWidth_mm` <- as.numeric(wide_plant_traits$`SeedWidth_mm`)
+wide_plant_traits$`DispersalSyndrome` <- as.factor(wide_plant_traits$`DispersalSyndrome`)
+wide_plant_traits$`GrowthForm` <- as.factor(wide_plant_traits$`GrowthForm`)
+```
+
+```{r}
+# rename accepted_species column to species
+wide_plant_traits <- wide_plant_traits %>%
+  rename(species = Accepted_species)
 ```
@@
 ```{r}
-print("Number of species: ")
-length(unique(wide_plant_traits$species))
+cat("Number of records:", nrow(clean_traits), "\n")
+cat("Number of species:", length(unique(clean_traits$Accepted_species)), "\n")
+```
+
+```{r}
+fb_plot_species_traits_completeness(wide_plant_traits)
+fb_plot_number_species_by_trait(wide_plant_traits)
+fb_table_trait_summary(wide_plant_traits)
+```
+
+```{r}
+vis_dat(wide_plant_traits)
+vis_miss(wide_plant_traits)
 ```
@@
 ```{r}
-write.csv(wide_plant_traits, file.path(output_path,"TropicalAndes_all_plant_traits_harmonized.csv"))
-write.csv(clean_traits, file.path(output_path,"TropicalAndes_all_plant_traits_harmonized_with_units.csv"))
+write.csv(wide_plant_traits, file.path(output_path,"TropicalAndes_all_plant_traits_standardized.csv"))
+write.csv(clean_traits, file.path(output_path,"TropicalAndes_all_plant_traits_standardized_with_recordCount.csv"))
 ```

### 2024-05-30 code/L1/L1_P5_TropicalAndes_plant_imputetraits.Rmd
update input paths

@@
 ```{r}
-data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
-output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
 ```
@@
 ```{r}
-plant_traits <- read.csv(file.path(data_path,"TropicalAndes_all_plant_traits_harmonized.csv"))
+plant_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_all_plant_traits_standardized.csv"))
 ```
@@
 plant_traits <- plant_traits[, !colnames(plant_traits) %in% "X", drop = FALSE]
-# remove Growth Forme trait
-plant_traits <- plant_traits %>%
-  select(-Growth.Forme)
+
 # Convert the wide dataframe to long format
@@
 ```
+
 Use taxize for species with no family and genus info

### 2024-06-03 code/L1/L1_LastStep_DiversityInputObjects.Rmd
update function to assign cellid as row names

@@
 ```{r}
-# Define the function
 create_presence_absence_matrix <- function(resolution_meters, species_sf) {
@@
     st_sf() %>%
-    mutate(cellid = row_number(),
-           centroid = st_centroid(geometry)) %>%
-    mutate(longitude = st_coordinates(centroid)[, 1],
-           latitude = st_coordinates(centroid)[, 2]) %>%
-    dplyr::select(-centroid)
+    mutate(cellid = row_number())
 
@@
     filter(!is.na(species)) %>%
-    dplyr::select(cellid, longitude, latitude, species)
+    dplyr::select(cellid, species, geometry)
+
+  # Extract the coordinates for each grid cell
+  species_grid_coords <- species_grid %>%
+    st_centroid() %>%
+    st_coordinates() %>%
+    as.data.frame() %>%
+    rename(Longitude = X, Latitude = Y)
+
+  # Combine coordinates with the species grid
+  species_grid <- bind_cols(species_grid, species_grid_coords)
 
@@
     mutate(presence = 1) %>%
-    pivot_wider(names_from = species, values_from = presence, values_fill = list(presence = 0)) %>%
-    arrange(cellid) %>%
-    dplyr::select(cellid, longitude, latitude, everything())
+    pivot_wider(names_from = species, values_from = presence, values_fill = list(presence = 0))
 
-  # Convert to matrix and set row names
-  presence_absence_matrix <- as.data.frame(presence_absence_matrix)
-  rownames(presence_absence_matrix) <- presence_absence_matrix$cellid
-  presence_absence_matrix <- presence_absence_matrix %>%
-    dplyr::select(-cellid) %>%
-    as.matrix()
+  # Convert to matrix
+  presence_absence_matrix_matrix <- as.matrix(presence_absence_matrix)
 
-  return(presence_absence_matrix)
+  # Set row names as "cell_rowNumber"
+  rownames(presence_absence_matrix_matrix) <- paste0("cell_", seq_len(nrow(presence_absence_matrix_matrix)))
+
+  # Ensure latitude and longitude are the first columns
+  presence_absence_matrix_matrix <- presence_absence_matrix_matrix[, c("Latitude", "Longitude", setdiff(colnames(presence_absence_matrix_matrix), c("Latitude", "Longitude", "cellid")))]
+
+  return(presence_absence_matrix_matrix)
 }
@@
 ```{r}
+dim(plant_PAM_100km)
+dim(frugivore_PAM_100km)
 # Remove columns with sum equal to zero
@@
 
-#Save coordinates for later
+dim(PAM_plant_site_final_100km)
+dim(PAM_frugivore_site_final_100km)
+
+# Save coordinates for later
 site_loc_key_plant_100km <- PAM_plant_site_final_100km[,1:2]
@@
 colnames_frugivore_100km <- colnames(PAM_frugivore_site_final_100km)
+
 ```

### 2024-06-03 code/L1/L1_P5_TropicalAndes_plant_imputetraits.Rmd
update to add additional traits

@@
 
-# Convert the wide dataframe to long format
-long_plant_traits <- plant_traits %>%
-  gather(TraitName, TraitValue, -species)
+# get a list of species
+na_species_list <- unique(plant_traits$species)
+```
 
-# Filter rows with NA values
-na_traits <- long_plant_traits %>%
-  filter(is.na(TraitValue))
-
-# get a list of species
-na_species_list <- unique(na_traits$species)
+```{r}
+length(na_species_list)
 ```
@@
 # Separate dataframes from list
-Dispersal.Syndrome <- trait_dfs$Dispersal.Syndrome
-Fruit.Length <- trait_dfs$Fruit.Length
-Plant.Height <- trait_dfs$Plant.Height
-Fruit.Type <- trait_dfs$Fruit.Type
-Growth.Form <- trait_dfs$Growth.Form
+DispersalSyndrome <- trait_dfs$DispersalSyndrome
+PlantLifespan_years <- trait_dfs$PlantLifespan_years
+FruitMass_mg <- trait_dfs$FruitMass_mg
+PlantHeight_m <- trait_dfs$PlantHeight_m
+GrowthForm <- trait_dfs$GrowthForm
+FruitLength_mm <- trait_dfs$FruitLength_mm
+SeedWidth_mm <- trait_dfs$SeedWidth_mm
+SeedLength_mm <- trait_dfs$SeedLength_mm
+SeedMass_g <- trait_dfs$SeedMass_g
+```
+
+## Create a list of unique genus and families for each trait
+```{r}
+DispersalSyndrome_families <- unique(DispersalSyndrome$family)
+DispersalSyndrome_genus <- unique(DispersalSyndrome$genus)
+length(DispersalSyndrome_families)
+length(DispersalSyndrome_genus)
 ```
@@
 ```{r}
-# Create a list of unique genus and families for each trait
-Dispersal.Syndrome_families <- unique(Dispersal.Syndrome$family)
-Dispersal.Syndrome_genus <- unique(Dispersal.Syndrome$genus)
+PlantLifespan_years_families <- unique(PlantLifespan_years$family)
+PlantLifespan_years_genus <- unique(PlantLifespan_years$genus)
+length(PlantLifespan_years_families)
+length(PlantLifespan_years_genus)
+```
 
-Fruit.Length_families <- unique(Fruit.Length$family)
-Fruit.Length_genus <- unique(Fruit.Length$genus)
+```{r}
 
-Plant.Height_families <- unique(Plant.Height$family)
-Plant.Height_genus <- unique(Plant.Height$genus)
+FruitMass_mg_families <- unique(FruitMass_mg$family)
+FruitMass_mg_genus <- unique(FruitMass_mg$genus)
 
-Fruit.Type_families <- unique(Fruit.Type$family)
-Fruit.Type_genus <- unique(Fruit.Type$genus)
+PlantHeight_m_families <- unique(PlantHeight_m$family)
+PlantHeight_m_genus <- unique(PlantHeight_m$genus)
 
-Growth.Form_families <- unique(Growth.Form$family)
-Growth.Form_genus <- unique(Growth.Form$genus)
+GrowthForm_families <- unique(GrowthForm$family)
+GrowthForm_genus <- unique(GrowthForm$genus)
+
+FruitLength_mm_families <- unique(FruitLength_mm$family)
+FruitLength_mm_genus <- unique(FruitLength_mm$genus)
+
+SeedWidth_mm_families <- unique(SeedWidth_mm$family)
+SeedWidth_mm_genus <- unique(SeedWidth_mm$genus)
+
+SeedLength_mm_families <- unique(SeedLength_mm$family)
+SeedLength_mm_genus <- unique(SeedLength_mm$genus)
+
+SeedMass_g_families <- unique(SeedMass_g$family)
+SeedMass_g_genus <- unique(SeedMass_g$genus)
 ```
@@
 ## Retrive trait records for species using BIEN & GIFT
+### Dispersal Syndrome
 ```{r}
-# Dispersal Syndrome
-Dispersal.Syndrome_families_BIEN <- BIEN_trait_traitbyfamily(family = Dispersal.Syndrome_families, trait = "whole plant dispersal syndrome")
-Dispersal.Syndrome_genus_BIEN <- BIEN_trait_traitbygenus(genus = Dispersal.Syndrome_genus, trait = "whole plant dispersal syndrome")
+DispersalSyndrome_families_BIEN <- BIEN_trait_traitbyfamily(family = DispersalSyndrome_families, trait = "whole plant dispersal syndrome")
+DispersalSyndrome_genus_BIEN <- BIEN_trait_traitbygenus(genus = DispersalSyndrome_genus, trait = "whole plant dispersal syndrome")
 
-Dispersal.Syndrome_families_GIFT <- GIFT_traits_tax(trait_IDs = "3.3.1")
-Dispersal.Syndrome_families_GIFT <- Dispersal.Syndrome_families_GIFT[Dispersal.Syndrome_families_GIFT$family %in% Dispersal.Syndrome_families, ]
+DispersalSyndrome_families_GIFT <- GIFT_traits_tax(trait_IDs = "3.3.1")
+DispersalSyndrome_families_GIFT <- DispersalSyndrome_families_GIFT[DispersalSyndrome_families_GIFT$family %in% DispersalSyndrome_families, ]
 ```
@@
 ```{r}
-# Fruit Length
+nrow(DispersalSyndrome_families_BIEN)
+length(unique(DispersalSyndrome_families_BIEN$scrubbed_family))
+unique(DispersalSyndrome_families_BIEN$trait_value)
+```
+```{r}
+nrow(DispersalSyndrome_genus_BIEN)
+length(unique(DispersalSyndrome_genus_BIEN$scrubbed_genus))
+unique(DispersalSyndrome_genus_BIEN$trait_value)
+```
+
+```{r}
+nrow(DispersalSyndrome_families_GIFT)
+```
+
+### Fruit Length
+```{r}
 Fruit.Length_families_BIEN <- BIEN_trait_traitbyfamily(family = Fruit.Length_families, trait = c("maximum fruit length", "minimum fruit length"))
@@
 #Fruit.Length_families_GIFT <- Fruit.Length_families_GIFT[Fruit.Length_families_GIFT$family %in% Fruit.Length_families, ]
+```
+
+```{r}
+nrow(DispersalSyndrome_families_BIEN)
+length(unique(DispersalSyndrome_families_BIEN$scrubbed_family))
+unique(DispersalSyndrome_families_BIEN$trait_value)
 ```
@@
 
-add column of trait level (specices, genus, family)
+add column of trait level (species, genus, family)
 ```{r}

### 2024-06-03 code/L2/L2_3_TropicalAndes_divRelationships.Rmd
update input files

@@
 data output: "ADD"
-date: "2023-10-04"
+date: "2024-05-16"
 output: html_document
@@
 ```{r}
-data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
-output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
+data_path_L2 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
+output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
+figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
 ```
@@
 ```{r}
-plant_cellRichness_1km <- read.csv(file.path(output_path,"TropicalAndes_plantRichness_1km.csv"))
-frugivore_cellRichness_1km <- read.csv(file.path(output_path,"TropicalAndes_frugivoreRichness_1km.csv"))
+plant_cellRichness_1km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_1km.csv"))
+frugivore_cellRichness_1km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_1km.csv"))
 
-plant_cellRichness_5km <- read.csv(file.path(output_path,"TropicalAndes_plantRichness_5km.csv"))
-frugivore_cellRichness_5km <- read.csv(file.path(output_path,"TropicalAndes_frugivoreRichness_5km.csv"))
+plant_cellRichness_5km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_5km.csv"))
+frugivore_cellRichness_5km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_5km.csv"))
 
-plant_cellRichness_10km <- read.csv(file.path(output_path,"TropicalAndes_plantRichness_10km.csv"))
-frugivore_cellRichness_10km <- read.csv(file.path(output_path,"TropicalAndes_frugivoreRichness_10km.csv"))
+plant_cellRichness_10km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_10km.csv"))
+frugivore_cellRichness_10km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_10km.csv"))
 
-plant_cellRichness_25km <- read.csv(file.path(output_path,"TropicalAndes_plantRichness_25km.csv"))
-frugivore_cellRichness_25km <- read.csv(file.path(output_path,"TropicalAndes_frugivoreRichness_25km.csv"))
+plant_cellRichness_25km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_25km.csv"))
+frugivore_cellRichness_25km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_25km.csv"))
 
-plant_cellRichness_50km <- read.csv(file.path(output_path,"TropicalAndes_plantRichness_50km.csv"))
-frugivore_cellRichness_50km <- read.csv(file.path(output_path,"TropicalAndes_frugivoreRichness_50km.csv"))
+plant_cellRichness_50km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_50km.csv"))
+frugivore_cellRichness_50km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_50km.csv"))
 
-plant_cellRichness_75km <- read.csv(file.path(output_path,"TropicalAndes_plantRichness_75km.csv"))
-frugivore_cellRichness_75km <- read.csv(file.path(output_path,"TropicalAndes_frugivoreRichness_75km.csv"))
+plant_cellRichness_75km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_75km.csv"))
+frugivore_cellRichness_75km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_75km.csv"))
 
-plant_cellRichness_100km <- read.csv(file.path(output_path,"TropicalAndes_plantRichness_100km.csv"))
-frugivore_cellRichness_100km <- read.csv(file.path(output_path,"TropicalAndes_frugivoreRichness_100km.csv"))
+plant_cellRichness_100km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_100km.csv"))
+frugivore_cellRichness_100km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_100km.csv"))
 ```

### 2024-06-03 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
working on plotting - not working still, update input files and object names

@@
 ```{r}
-# objects for functional diversity
-PAM_plant_site_final <- readRDS(file = file.path(data_path_L1,"PAM_plant_site_final.rds"))
-PAM_frugivore_site_final <- readRDS(file = file.path(data_path_L1,"PAM_frugivore_site_final.rds"))
 plant_traits_df_final <- readRDS(file = file.path(data_path_L1,"plant_traits_df_final.rds"))
 frugivore_traits_df_final <- readRDS(file = file.path(data_path_L1,"frugivore_traits_df_final.rds"))
-site_loc_key_plant <- readRDS(file = file.path(data_path_L1,"site_loc_key_plant.rds"))
-site_loc_key_frugivore <- readRDS(file = file.path(data_path_L1,"site_loc_key_frugivore.rds"))
+```
+
+```{r}
+site_loc_key_plant_100km <- readRDS(file = file.path(data_path_L1,"site_loc_key_plant_100km.rds"))
+site_loc_key_frugivore_100km <- readRDS(file = file.path(data_path_L1,"site_loc_key_frugivore_100km.rds"))
+PAM_plant_site_final_100km <- readRDS(file = file.path(data_path_L1,"PAM_plant_site_final_100km.rds"))
+PAM_frugivore_site_final_100km <- readRDS(file = file.path(data_path_L1,"PAM_frugivore_site_final_100km.rds"))
 ```
@@
 ```{r}
-asb_sp_frugivore_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_frugivore_site_final)
+asb_sp_frugivore_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_frugivore_site_final_100km)
 ```
@@
 # Calculate row sums
-row_sums_frugivore <- rowSums(PAM_frugivore_site_final)
-subset_matrix_frugivore <- PAM_frugivore_site_final[row_sums_frugivore >= 4, ]
+row_sums_frugivore <- rowSums(PAM_frugivore_site_final_100km)
+subset_matrix_frugivore <- PAM_frugivore_site_final_100km[row_sums_frugivore >= 4, ]
 ```
@@
   asb_sp_w         = subset_matrix_frugivore,
-  ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
+  ind_vect         = "fdis",
+  #did not run fdiv or fric (slows it down); c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe")
   scaling          = TRUE,
@@
 details_list_frugivore <- alpha_fd_indices_frugivore$"details"
-
-plots_alpha_frugivore <- mFD::alpha.multidim.plot(
-  output_alpha_fd_multidim = alpha_fd_indices_frugivore,
-  plot_asb_nm              = c("cell4", "cell20"),
-  ind_nm                   = c("fdis", "fide", "fnnd"
-                        , "fori", "fspe"),
-  faxes                    = NULL,
-  faxes_nm                 = NULL,
-  range_faxes              = c(NA, NA),
-  color_bg                 = "grey95",
-  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
-  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
-  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  color_ch          
-  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
-  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
-  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
-  shape_centroid_fspe      = 23,
-  color_centroid_fspe      = "black",
-  size_sp_nm               = 3, 
-  color_sp_nm              = "black",
-  plot_sp_nm               = NULL,
-  fontface_sp_nm           = "plain",
-  save_file                = FALSE,
-  check_input              = TRUE)
-
- #comparing convex hulls of different hulls
-plots_alpha_frugivore$"fric"$"patchwork"
-
-plots_alpha_frugivore$"fdiv"$"patchwork"
-
-plots_alpha_frugivore$"fspe"$"patchwork"
-
-plots_alpha_frugivore$"fdis"$"patchwork"
-
-plots_alpha_frugivore$"fori"$"patchwork"
 ```
@@
 # Generate coordinates
-subset_coords_frugivore <- site_loc_key_frugivore[rowSums(PAM_frugivore_site_final) >= 5, ]
+subset_coords_frugivore <-site_loc_key_frugivore_100km[rowSums(PAM_frugivore_site_final_100km) >= 4, ]
 subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]
@@
 # Convert the dataframe to sf format
-spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
+spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude", "Latitude"))
 ```
@@
 # set crs of sf objects
-spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(4326)
+spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(5389)
 ```
@@
     summarize(mean_value = mean(get(value_col), na.rm = TRUE)) %>%
+    ungroup() %>%
     st_set_geometry("geometry")
@@
 
-# Example usage with your data
+# use funtion
 resolution_meters <- 100000
@@
 
+# Convert geometries to centroids
+aggregated_data_centroids <- st_centroid(aggregated_data)
+
+# Extract coordinates from centroids
+aggregated_data_coords <- st_coordinates(aggregated_data_centroids)
+
+# Add coordinates as separate columns
+aggregated_data <- aggregated_data %>%
+  mutate(x = aggregated_data_coords[, 1], y = aggregated_data_coords[, 2])
+
+# Remove rows with NA values in mean_value
+aggregated_data <- aggregated_data %>%
+  filter(!is.na(mean_value))
 ```
@@
     ggplot() +
-    geom_sf(data = Americas_latlong, fill = "white")+
-    geom_sf(data = TApoly_latlong, fill = "lightgrey", size = 0.1) +
-    geom_point(data = spatial_fdis_frugivore , aes(fill = fdis_frugivore)) +
+    geom_sf(data = Americas, fill = "white")+
+    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
+    geom_sf(data = grid, fill = "grey")+
+    geom_sf(data = aggregated_data, aes(fill = mean_value), color = NA) +
+    labs(x = "Longitude", y = "Latitude", fill = "FDis") +
     scale_fill_gradient(low = "blue", high = "red") +
@@
     theme(panel.background = element_rect(fill = "lightblue")) +
-    annotation_scale(location = "bl", width_hint = 0.5) +
-    labs(x = "Longitude", y = "Latitude", fill = "FDis")
+    theme(text = element_text(size = 12)) # Explicitly setting text size
 FDisplot

### 2024-06-03 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
trying to fix functional diversity mapping by gridding point fdis data - produced map but doesn't look like 100km

@@
 ```{r}
-# Create a grid over the area
-create_grid <- function(data, cellsize) {
-  st_make_grid(st_as_sfc(st_bbox(data)), cellsize = cellsize) %>%
+# Extract coordinates 
+spatial_fdis_frugivore_coords <- st_coordinates(spatial_fdis_frugivore)
+
+# Add coordinates as separate columns
+spatial_fdis_frugivore <- spatial_fdis_frugivore %>%
+  mutate(x = spatial_fdis_frugivore_coords[, 1], y = spatial_fdis_frugivore_coords[, 2])
+```
+
+```{r}
+summary(spatial_fdis_frugivore)
+```
+
+```{r}
+TAGrid_100km <- TApoly %>%
+  st_make_grid(cellsize = c(100000)) %>%
+  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+  st_cast("MULTIPOLYGON") %>%
   st_sf() %>%
   mutate(cellid = row_number())
-}
 
-# Aggregate functional dispersion within each grid cell
-aggregate_to_grid <- function(grid, points, value_col) {
-  points %>%
-    st_join(grid) %>%
+spatial_fdis_frugivore$cellid <- sapply(1:nrow(TAGrid_100km), function(i) {
+  contains <- st_contains(TAGrid_100km[i, ], spatial_fdis_frugivore)
+  cellid <- ifelse(any(contains), i, NA)
+  return(cellid)
+})
+
+spatial_fdis_frugivore_grid <- TAGrid_100km %>%
+   st_join(spatial_fdis_frugivore) %>%
    group_by(cellid) %>%
-    summarize(mean_value = mean(get(value_col), na.rm = TRUE)) %>%
-    ungroup() %>%
-    st_set_geometry("geometry")
-}
+   summarize(fdis_value = sum(fdis_frugivore, na.rm = TRUE)) %>%
+   ungroup()
+```
 
-# use funtion
-resolution_meters <- 100000
-variable_name <- "fdis_frugivore"
-
-# Create grid
-grid <- create_grid(spatial_fdis_frugivore, cellsize = resolution_meters)
-
-# Aggregate data to grid
-aggregated_data <- aggregate_to_grid(grid, spatial_fdis_frugivore, variable_name)
-
-# Convert geometries to centroids
-aggregated_data_centroids <- st_centroid(aggregated_data)
-
-# Extract coordinates from centroids
-aggregated_data_coords <- st_coordinates(aggregated_data_centroids)
-
-# Add coordinates as separate columns
-aggregated_data <- aggregated_data %>%
-  mutate(x = aggregated_data_coords[, 1], y = aggregated_data_coords[, 2])
-
-# Remove rows with NA values in mean_value
-aggregated_data <- aggregated_data %>%
-  filter(!is.na(mean_value))
+```{r}
+summary(spatial_fdis_frugivore_grid)
 ```
@@
     geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
-    geom_sf(data = grid, fill = "grey")+
-    geom_sf(data = aggregated_data, aes(fill = mean_value), color = NA) +
+    geom_sf(data = spatial_fdis_frugivore_grid, aes(fill = fdis_value), color = NA) +
     labs(x = "Longitude", y = "Latitude", fill = "FDis") +
-    scale_fill_gradient(low = "blue", high = "red") +
+    scale_fill_distiller(palette = "YlOrBr", direction = 1) +
     coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-    theme(panel.background = element_rect(fill = "lightblue")) +
-    theme(text = element_text(size = 12)) # Explicitly setting text size
-FDisplot
-```
-
-# Mapping without projection 
-Because projecting is being difficult
-```{r}
-TropicalAndes_IUCNHabitat_Forest_latlong <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
-Americas_latlong <- ne_countries(continent = c("North America", "South America"), returnclass = "sf")
-#polygon of Tropical Andes
-TApoly_latlong <- Americas_latlong %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
-```
-
-```{r}
-# Generate coordinates
-subset_coords_frugivore <- site_loc_key_frugivore[rowSums(PAM_frugivore_site_final) >= 5, ]
-subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]
-
-frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)
-```
-
-```{r}
-# Convert the dataframe to sf format
-spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
-```
-
-```{r}
-# set crs of sf objects
-spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(4326)
-```
-
-```{r}
-# Create a grid over the area
-create_grid <- function(data, cellsize) {
-  st_make_grid(st_as_sfc(st_bbox(data)), cellsize = cellsize) %>%
-    st_sf() %>%
-    mutate(cellid = row_number())
-}
-
-# Aggregate functional dispersion within each grid cell
-aggregate_to_grid <- function(grid, points, value_col) {
-  points %>%
-    st_join(grid) %>%
-    group_by(cellid) %>%
-    summarize(mean_value = mean(get(value_col), na.rm = TRUE)) %>%
-    st_set_geometry("geometry")
-}
-
-# Example usage with your data
-resolution_meters <- 100000
-variable_name <- "fdis_frugivore"
-
-# Create grid
-grid <- create_grid(spatial_fdis_frugivore, cellsize = resolution_meters)
-
-# Aggregate data to grid
-aggregated_data <- aggregate_to_grid(grid, spatial_fdis_frugivore, variable_name)
-
-```
-
-```{r}
-FDisplot <-
-    ggplot() +
-    geom_sf(data = Americas_latlong, fill = "white")+
-    geom_sf(data = TApoly_latlong, fill = "lightgrey", size = 0.1) +
-    geom_point(data = spatial_fdis_frugivore , aes(fill = fdis_frugivore)) +
-    scale_fill_gradient(low = "blue", high = "red") +
-    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-    theme(panel.background = element_rect(fill = "lightblue")) +
-    annotation_scale(location = "bl", width_hint = 0.5) +
-    labs(x = "Longitude", y = "Latitude", fill = "FDis")
+    theme(panel.background = element_rect(fill = "lightblue"),
+        text = element_text(size = 12))
 FDisplot
@@
 # Plants
-Test with small random sample as a subset to make it work before using all data
-```{r}
-set.seed(100)
-all_plant_species <- row.names(plant_traits_df_final)
-plant_random_subset <- sample(all_plant_species, 1000)
-# objects that need to be subset
-PAM_plant_site_final_df <- as.dataframe(PAM_plant_site_final) #matrix
-plant_traits_df_final
-```
-
 ## Create trait type table
@@
 
+```
+
+```{r}
 ## Create an empty dataframe to store results

### 2024-06-03 code/L1/L1_LastStep_DiversityInputObjects.Rmd
remove 1 km

@@
 
-## 1 km 
-```{r}
-plant_PAM_1km <- create_presence_absence_matrix(1000, plants_sf_species)
-frugivore_PAM_1km <- create_presence_absence_matrix(1000, frugivores_sf_species)
-```
-
-```{r}
-#check str 
-plant_PAM_1km[1:4, 1:4]
-frugivore_PAM_1km[1:4, 1:4]
-```
-
-### Remove the species from PAM that have no occurrences
-```{r}
-# Remove columns with sum equal to zero
-PAM_plant_site_final_1km <- plant_PAM_1km[, colSums(plant_PAM_1km) != 0]
-PAM_frugivore_site_final_1km <- frugivore_PAM_1km[, colSums(frugivore_PAM_1km) != 0]
-
-#Save coordinates for later
-site_loc_key_plant_1km <- PAM_plant_site_final_1km[,1:2]
-site_loc_key_frugivore_1km <- PAM_frugivore_site_final_1km[,1:2]
-
-PAM_plant_site_final_1km <- PAM_plant_site_final_1km[,-c(1:2)]
-PAM_frugivore_site_final_1km <- PAM_frugivore_site_final_1km[,-(1:2)]
-
-colnames_plant_1km <- colnames(PAM_plant_site_final_1km)
-colnames_frugivore_1km <- colnames(PAM_frugivore_site_final_1km)
-```
-
-```{r}
-str(PAM_plant_site_final_1km)
-str(PAM_frugivore_site_final_1km)
-```
-
 ## Remove species names from trait matrix not in the PAM
@@
 
-```{r}
-saveRDS(plant_PAM_1km, file = file.path(data_path_L1,"plant_PAM_1km.rds"))
-saveRDS(frugivore_PAM_1km, file = file.path(data_path_L1,"frugivore_PAM_1km.rds"))
-```
-
 ## objects for functional diversity
@@
 ```
-
-```{r}
-saveRDS(site_loc_key_plant_1km, file = file.path(data_path_L1,"site_loc_key_plant_1km.rds"))
-saveRDS(site_loc_key_frugivore_1km, file = file.path(data_path_L1,"site_loc_key_frugivore_1km.rds"))
-saveRDS(PAM_plant_site_final_1km, file = file.path(data_path_L1,"PAM_plant_site_final_1km.rds"))
-saveRDS(PAM_frugivore_site_final_1km, file = file.path(data_path_L1,"PAM_frugivore_site_final_1km.rds"))
-```

### 2024-06-03 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
remove 1 km

--- 2024-05-22 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
@@
 
-## 1 km resolution
-```{r}
-richness_1km <- create_rich_plots(1000)
-print(richness_1km)
-
-# save plots
-plantgridRichTA_1km <- richness_1km$plant_plot
-frugivoregridRichTA_1km <- richness_1km$frugivoregridRichTA
-# save richness data
-plant_cellRichness_1km <- richness_1km$plant_cellRichness
-frugivore_cellRichness_1km <- richness_1km$frugivore_cellRichness
-```
-
 ## 5 km resolution
@@
 ```{r}
-all_plant_richness_plots <- ggpubr::ggarrange(#plantgridRichTA_1km, plantgridRichTA_5km, 
-                                              #plantgridRichTA_10km, 
-                                              plantgridRichTA_25km,
+all_plant_richness_plots <- ggpubr::ggarrange(# plantgridRichTA_5km, 
+                                              plantgridRichTA_10km, plantgridRichTA_25km,
                                               plantgridRichTA_50km, plantgridRichTA_75km,
                                               plantgridRichTA_100km,
-                                              ncol = 4, nrow = 1,
-                                              labels = c("25 km", "50 km", "75 km", "100 km"),
+                                              ncol = 5, nrow = 1,
+                                              labels = c("10 km", "25 km", "50 km", "75 km", "100 km"),
                                               common.legend = TRUE, legend = "right")
+all_plant_richness_plots
 all_plant_richness_plots_labeled <- ggpubr::annotate_figure(all_plant_richness_plots,
@@
 ```{r}
-all_frugivore_richness_plots <- ggpubr::ggarrange(#frugivoregridRichTA_1km, frugivoregridRichTA_5km, 
-                                              #frugivoregridRichTA_10km, 
-                                              frugivoregridRichTA_25km,
+all_frugivore_richness_plots <- ggpubr::ggarrange(#frugivoregridRichTA_5km, 
+                                              frugivoregridRichTA_10km, frugivoregridRichTA_25km,
                                               frugivoregridRichTA_50km, frugivoregridRichTA_75km,
                                               frugivoregridRichTA_100km,
-                                              ncol = 4, nrow = 1,
-                                              labels = c("25 km", "50 km", "75 km", "100 km"),
+                                              ncol = 5, nrow = 1,
+                                              labels = c("10 km", "25 km", "50 km", "75 km", "100 km"),
                                               common.legend = TRUE, legend = "right")
@@
 # Write data to csv
-```{r}
-write.csv(plant_cellRichness_1km, file.path(output_path,"TropicalAndes_plantRichness_1km.csv"))
-write.csv(frugivore_cellRichness_1km, file.path(output_path,"TropicalAndes_frugivoreRichness_1km.csv"))
-```
 

### 2024-06-03 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
remove unnecessary code and add histogram

@@
 
-spatial_fdis_frugivore$cellid <- sapply(1:nrow(TAGrid_100km), function(i) {
-  contains <- st_contains(TAGrid_100km[i, ], spatial_fdis_frugivore)
-  cellid <- ifelse(any(contains), i, NA)
-  return(cellid)
-})
+spatial_fdis_frugivore$cellid <- st_intersects(spatial_fdis_frugivore, TAGrid_100km)$nn
 
@@
 ```{r}
-FDisplot <-
+frugivoregridFDisTA_100km <-
     ggplot() +
@@
         text = element_text(size = 12))
-FDisplot
+frugivoregridFDisTA_100km
+```
+
+```{r}
+# Extract cell values
+frugivore_cellFDis_100km <- spatial_fdis_frugivore_grid
+hist(frugivore_cellFDis_100km$fdis_value)
 ```
@@
 
-## Turning the FD values into a raster
-Need to subset the coordinates made at the top to the same set of the assemblages (removed values less than 5 so these coords need to be removed to do this correctly)
-
-```{r}
-# Create an empty raster with desired resolution and extent
-resolution <- c(0.75, 0.75)  # Set the desired resolution
-extent <- c(-85, -54, -24, 14)  # Set the desired extent
-empty_raster <- raster(resolution = resolution, xmn = extent[1], xmx = extent[2], ymn = extent[3], ymx = extent[4])
-```
-
-```{r}
-# Generate coordinates
-subset_coords_plant <- site_loc_key_plant[rowSums(PAM_plant_site_final) >= 5, ]
-subset_coords_plant_sp <-subset_coords_plant[,1:2 ]
-
-plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
-
-# Convert the dataframe to sf format
-spatial_fdis_plant <- st_as_sf(plant_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
-```
-
-```{r}
-# Rasterize the sf data to create the FD raster
-fd_raster_plant <- rasterize(spatial_fdis_plant, empty_raster)
-writeRaster(fd_raster_plant$fdis_plant, filename = file.path(output_path,"FD_plant_TA.tif"), format="GTiff",overwrite=T)
-writeRaster(fd_raster_plant$ID, filename = file.path(output_path,"FD_plant_TA_ID.tif"), format="GTiff")
-```
-
-## Project FD objects
-```{r}
-# set crs of sf objects
-spatial_fdis_plant <- spatial_fdis_plant %>% st_set_crs(4326)
-```
-
-```{r}
-spatial_fdis_plant_projected <- st_transform(spatial_fdis_plant, 5389)
-```
-
-```{r}
-# Raster to raster point
-spec_FD_raster_plant_point <- rasterToPoints(fd_raster_plant, spatial = TRUE)
-# Convert to a 'conventional' dataframe
-fd_raster_plant_point <- data.frame(spec_FD_raster_plant_point)
-```
-
-```{r}
-plantgridFDTA <-
-  ggplot() +
-  geom_sf(data = Americas, fill = "white")+
-  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
-  geom_tile(data= fd_raster_plant_point, aes(x=x, y=y, fill=fdis_plant)) +
-  scale_fill_scico(palette = "romaO", direction = 1) +
-  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
-  theme(
-    plot.background = element_rect(fill = "#f1f2f3"),
-    panel.background = element_rect(fill = "lightblue"),
-    panel.grid = element_blank(),
-    line = element_blank(),
-    rect = element_blank()
-  ) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
-  labs(title = "Tropical Andes Plant Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
-plantgridFDTA
-```
-
-# Calculating Rao's Q
-Calculate Rao's Q using occurrence data and trait dissimilarity
-
-```{r}
-## Define different spatial scales (e.g., grid cell sizes)
-spatial_scales <- c(1000, 5000, 10000, 25000, 50000, 75000, 100000)  # In meters
-
-```
-
-```{r}
-## Create an empty dataframe to store results
-plant_raos_q_results <- data.frame(spatial_scale = numeric(),
-                             raos_q = numeric())
-
-## Iterate over different spatial scales
-for (scale in spatial_scales) {
-  ### Aggregate occurrence data to the current spatial scale
-  plant_occurrence_aggregated <- aggregate_occurrence(plant_abundance_per_species, scale)
-
-  ### Calculate dissimilarities using trait data
-  plant_trait_dissimilarities <- vegdist(plant_traits, method = "bray")
-
-  ### Calculate dissimilarities using occurrence data
-  plant_occurrence_dissimilarities <- vegdist(plant_occurrence_aggregated, method = "bray")
-
-  ### Calculate Rao's Q index
-  raos_q <- raosq(plant_trait_dissimilarities, plant_occurrence_dissimilarities)
-
-  #### Add results to the dataframe
-  plant_raos_q_results <- rbind(raos_q_results, data.frame(spatial_scale = scale, raos_q = raos_q))
-}
-```
-
-```{r}
-## Create an empty dataframe to store results
-frugivore_raos_q_results <- data.frame(spatial_scale = numeric(),
-                             raos_q = numeric())
-
-## Iterate over different spatial scales
-for (scale in spatial_scales) {
-  ### Aggregate occurrence data to the current spatial scale
-  frugivore_occurrence_aggregated <- aggregate_occurrence(frugivore_abundance_per_species, scale)
-
-  ### Calculate dissimilarities using trait data
-  frugivore_trait_dissimilarities <- vegdist(frugivore_traits, method = "bray")
-
-  ### Calculate dissimilarities using occurrence data
-  frugivore_occurrence_dissimilarities <- vegdist(frugivore_occurrence_aggregated, method = "bray")
-
-  ### Calculate Rao's Q index
-  raos_q <- raosq(frugivore_trait_dissimilarities, frugivore_occurrence_dissimilarities)
-
-  #### Add results to the dataframe
-  frugivore_raos_q_results <- rbind(raos_q_results, data.frame(spatial_scale = scale, raos_q = raos_q))
-}
-```
-
 # Extract cell values
@@
 ```
-
-```{r}
-plot(spatial_fdis_plant$fdis_plant, spatial_fdis_frugivore$fdis_frugivore,
-     main = "Plant FD vs Frugivore FD",
-     xlab = "Plant Functional Dispersion by cell", ylab = "Frugivore Functional Dispersion by cell")
-abline(lm(spatial_fdis_frugivore$fdis_frugivore ~ spatial_fdis_plant$fdis_plant))
-summary(lm(spatial_fdis_frugivore$fdis_frugivore ~ spatial_fdis_plant$fdis_plant))
-```

### 2024-06-10 code/L1/L1_P3_TropicalAndes_plant_fruiting_subset.Rmd
update to create species list with any species that has information related to a fruiting trait

--- 2024-05-29 code/L1/L1_P3_TropicalAndes_plant_fruiting_subset.Rmd
@@
 overview: "This script subsets plant occurrence and trait data by fruiting plant species."
-data input: "ADD"
-data output: "ADD"
+data input: "TropicalAndes_GBIF_plant_occ_harmonized.csv, TropicalAndes_TRY_plant_traits_harmonized.csv,
+             TropicalAndes_BIEN_plant_traits_harmonized.csv, TropicalAndes_GIFT_plant_traits_harmonized.csv"
+data output: "TropicalAndes_GBIF_plant_occ_harmonized_subset.csv, TropicalAndes_all_plant_traits_harmonized_subset.csv"
 date: "2024-05-29"
@@
 ```{r}
+# add trait database source column to each subset
+TRY_filtered_subset$DatabaseSource <- "TRY"
+BIEN_filtered_subset$DatabaseSource <- "BIEN"
+GIFT_filtered_subset$DatabaseSource <- "GIFT"
+
+# combine all subsets into one dataframe
 traits <- rbind(TRY_filtered_subset, BIEN_filtered_subset, GIFT_filtered_subset)
-nrow(traits)
+dim(traits)
 ```
@@
 ```{r}
-traits %>%
+traits_clean %>%
   distinct(Accepted_species, TraitName) %>%
@@
 
+# Create a species list of species with fruiting traits
 ```{r}
-# Group by TraitName and summarize the unique TraitValues
-unique_trait_values <- traits %>%
-  group_by(TraitName) %>%
-  summarize(UniqueValues = list(unique(TraitValue)))%>%
-  unnest(UniqueValues)
+fruit_traits <- c("Fruit dehiscence type", "Fruit dry mass", "Fruit length", "Fruit surface type",
+                  "Fruit texture inside (fleshy, leathrey, …)", "Fruit type", "fruit type", "fruit_dryness", "fruit_length_max",
+                  "fruit_length_mean", "fruit_length_min", "fruit_type", "maximum fruit length", "minimum fruit length")
+
+# Filter the dataframe for fruiting traits
+fruiting_df <- traits_clean %>%
+  filter(TraitName %in% fruit_traits)
+dim(fruiting_df)
+
+# Extract unique species names
+fruiting_species <- unique(fruiting_df$Accepted_species)
+length(fruiting_species)
 ```
 
-# Function to get one value per species
+# Subset trait data and occurrence data by fruiting species list
+
 ```{r}
-# categorical data
-factor_data_merge <- function(data, traitname) {
-# remove duplicate rows
-data <- data %>%
-  distinct(.keep_all = TRUE)
-# remove rows with missing or empty species names
-data <- data[!is.na(data$Accepted_species) & data$Accepted_species != "", ]
-# Subset data to just species and TraitValue
-data <- data[,c("Accepted_species", "TraitValue")]
-# Find the most common trait value for each species
-  most_common_values <- aggregate(. ~ Accepted_species, data, function(x) {
-    # Find the mode (most common value) of the trait values
-    mode_value <- names(sort(table(x), decreasing = TRUE))[1]
-    return(mode_value)
-  })
-# Merge most common trait values back to the original data
-merged_data <- merge(data, most_common_values, by = "Accepted_species", all.x = TRUE)
-
-# Rename the merged column
-colnames(merged_data)[ncol(merged_data)] <- "TraitValue"
-
-# Remove column TraitValue.x
-merged_data <- merged_data[, !colnames(merged_data) %in% "TraitValue.x", drop = FALSE]
-
-# remove duplicate rows
-merged_data <- merged_data %>%
-  distinct(.keep_all = TRUE)
-
-# Add unit & trait column
-merged_data$TraitName <- traitname
-
-  return(merged_data)
-}
-```
-
-# Fruit type
-```{r}
-traits_fruitType <- traits %>%
-  filter(grepl('Fruit type|fruit type|fruit_type', TraitName))
-traits_fruitType$OrigTraitValue <- traits_fruitType$TraitValue
-
-nrow(traits_fruitType)
-length(unique(traits_fruitType$Accepted_species))
+dim(plant_occ)
+plant_occ_subset <- plant_occ %>%
+  filter(Accepted_species %in% fruiting_species)
+dim(plant_occ_subset)
 ```
@@
 ```{r}
-traits_fruitType <- traits_fruitType %>%
-  filter(is.na(as.numeric(TraitValue)))
-
-nrow(traits_fruitType)
-length(unique(traits_fruitType$Accepted_species))
+dim(traits_clean)
+traits_subset <- traits_clean%>%
+  filter(Accepted_species %in% fruiting_species)
+dim(traits_subset)
 ```
@@
 ```{r}
-traits_fruitType %>%
-  count(TraitValue)
+traits_subset %>%
+  distinct(Accepted_species, TraitName) %>%
+  count(TraitName)
 ```
 
+# Write data to csv
 ```{r}
-# standardize trait values
-traits_fruitType <- traits_fruitType %>%
-  filter() %>%
-  mutate(TraitValue= case_when(
-    grepl('Berry|berry|Berry ', OrigTraitValue, ignore.case = TRUE) ~ "berry",
-    grepl('capsule|Capsule|Capsule ', OrigTraitValue, ignore.case = TRUE) ~ "capsule",
-    grepl('follicle|Follicle', OrigTraitValue, ignore.case = TRUE) ~ "follicle",
-  )) %>%
-  filter(!is.na(TraitValue))
+write.csv(plant_occ_subset, file.path(output_path, "TropicalAndes_GBIF_plant_occ_harmonized_subset.csv"))
+write.csv(traits_subset, file.path(output_path, "TropicalAndes_all_plant_traits_harmonized_subset.csv"))
 ```
-
-```{r}
-# standardize trait name
-traits_fruitType <- traits_fruitType %>%
-  mutate(TraitName = "FruitType")
-```
-
-```{r}
-averagetraits_fruitType <- factor_data_merge(traits_fruitType, "Fruit Type")
-```
-
-```{r}
-nrow(averagetraits_fruitType)
-length(unique(averagetraits_fruitType$Accepted_species))
-```
-
-```{r}
-summary(averagetraits_fruitType)
-```

### 2024-06-10 code/L1/L1_F1_TropicalAndes_frugivores_GBIF.Rmd
update to use data_summary function

@@
 ```{r}
-print("The number of records is")
-nrow(TropicalAndes_GBIF_frugivore_occ_cleaned)
-print("The number of species is")
-length(unique(TropicalAndes_GBIF_frugivore_occ_cleaned$species))
-print("The number of genera is" )
-length(unique(TropicalAndes_GBIF_frugivore_occ_cleaned$genus))
-print("The number of families is")
-length(unique(TropicalAndes_GBIF_frugivore_occ_cleaned$family))
+source("~/GitHub/neotropical_plants/code/data_summary.R")
+
+data_summary(TropicalAndes_GBIF_frugivore_occ_cleaned, TropicalAndes_GBIF_frugivore_occ_cleaned$species, TropicalAndes_GBIF_frugivore_occ_cleaned$genus, TropicalAndes_GBIF_frugivore_occ_cleaned$family)
 ```

### 2024-06-10 code/L1/L1_F2_TropicalAndes_frugivore_traits.Rmd
add code to visualize trait record completeness

@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: "Subsetting Frugivoria to the species and traits of interest"
-data input: "TropicalAndes_Frugivoria_frugivore_traits.csv", "TropicalAndes_GBIF_frugivore_occ_cleaned.csv"
+overview: "Subsetting Frugivoria to the species and traits of interest."
+data input: "TropicalAndes_Frugivoria_frugivore_traits.csv, TropicalAndes_GBIF_frugivore_occ_cleaned.csv"
 data output: "TropicalAndes_Frugivoria_traits_subset.csv"
 date: "2023-10-31"
-output: html_document
+output: 
+  html_document:
+    code-link: true
 ---
@@
 library(dplyr)
+library(funbiogeo)
+library(visdat)
 ```
@@
 ```{r}
-source("~/GitHub/neotropical_plants/code/L0/data_summary.R")
+source("~/GitHub/neotropical_plants/code/data_summary.R")
 ```
@@
 
+```{r}
+fb_plot_species_traits_completeness(Frugivoria_subset)
+fb_plot_number_species_by_trait(Frugivoria_subset)
+fb_table_trait_summary(Frugivoria_subset)
+```
+
+```{r}
+vis_dat(Frugivoria_subset)
+vis_miss(Frugivoria_subset)
+```
+
 # Write data to csv
@@
 ```
+
+# Package citations and session info
+```{r}
+library(report)
+report::cite_packages()
+
+devtools::session_info()
+```

### 2024-06-10 code/L1/L1_P4_TropicalAndes_plant_traits.Rmd
add code to detect and report outliers, as well as summarize species records

@@
 overview: "This script subsets plant trait data by synonyms species list and combine."
-data input: "ADD"
+data input: "TropicalAndes_all_plant_traits_harmonized_subset.csv"
 data output: "ADD"
@@
 library(purrr)
+library(rstatix)
 library(funbiogeo)
@@
 ```{r}
-data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
 data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
@@
 - Seed width
+- Fruit type
 
@@
 
-  # Calculate average value, count, and variance per trait per species
+  # Calculate geometic mean value, count, and variance per trait per species
   summary_data <- data %>%
@@
     summarise(
-      TraitValue_mean = mean(TraitValue, na.rm = TRUE),
+      TraitValue_mean = exp(mean(log(TraitValue))),
       records_used = n(),
@@
   return(summary_data)
+}
+```
+
+```{r}
+# Detect outliers and report species with multiple entries per trait
+detect_outliers <- function(data) {
+  outlier_results <- data %>%
+    group_by(Accepted_species) %>%
+    filter(n() >= 3) %>%
+    nest() %>%
+    mutate(outliers = map(data, ~identify_outliers(.x, variable = "TraitValue"))) %>%
+    unnest(outliers) %>%
+    select(Accepted_species, TraitName, TraitValue, is.outlier, is.extreme) %>%
+    ungroup()
+
+  return(outlier_results)
+}
+
+# Get the species with outliers
+species_with_outliers <- function(data) {
+  outliers_by_species <- data %>%
+    group_by(Accepted_species) %>%
+    summarize(outlier_count = sum(is.outlier, na.rm = TRUE)) %>%
+    filter(outlier_count > 0) %>%
+    ungroup()
+
+  return(outliers_by_species)
+}
+```
+
+```{r}
+summarize_species_records <- function(df) {
+  # Count the total number of species
+  total_species <- df %>% 
+    distinct(Accepted_species) %>% 
+    nrow()
+
+  # Count the number of species with only one trait record
+  species_with_one_record <- df %>% 
+    group_by(Accepted_species) %>% 
+    summarize(record_count = n()) %>% 
+    filter(record_count == 1) %>% 
+    nrow()
+
+  # Count the number of species with more than three records
+  species_with_more_than_three_records <- df %>% 
+    group_by(Accepted_species) %>% 
+    summarize(record_count = n()) %>% 
+    filter(record_count > 3) %>% 
+    nrow()
+
+  # Print the results
+  cat("Total number of species:", total_species, "\n")
+  cat("Number of species with only one trait record:", species_with_one_record, "\n")
+  cat("Number of species with more than three records:", species_with_more_than_three_records, "\n")
 }
@@
 ```{r}
+summarize_species_records(traits_dispersal)
+```
+
+```{r}
 averagetraits_dispersal <- factor_data_merge(traits_dispersal, "DispersalSyndrome")
@@
   mutate(TraitName = "FruitMass_mg")
+```
+
+```{r}
+dotchart(traits_fruitMass$TraitValue)
+```
+
+```{r}
+# detect outliers
+fruitMass_outliers <- detect_outliers(traits_fruitMass)
+print(fruitMass_outliers)
+fruitMass_outliers_species <- species_with_outliers(fruitMass_outliers)
+print(fruitMass_outliers_species)
+```
+
+```{r}
+summarize_species_records(traits_fruitMass)
 ```
@@
 ```{r}
+dotchart(traits_fruitLength$TraitValue)
+```
+
+```{r}
+# detect outliers
+fruitLength_outliers <- detect_outliers(traits_fruitLength)
+print(fruitLength_outliers)
+fruitLength_outliers_species <- species_with_outliers(fruitLength_outliers)
+print(fruitLength_outliers_species)
+```
+
+```{r}
+summarize_species_records(traits_fruitLength)
+```
+
+```{r}
 averagetraits_fruitLength <- numeric_data_average(traits_fruitLength, "mm", "FruitLength_mm")
@@
   mutate(TraitName = "PlantHeight_m")
+```
+
+```{r}
+dotchart(traits_plantHeight$TraitValue)
+```
+
+```{r}
+# detect outliers
+plantHeight_outliers <- detect_outliers(traits_plantHeight)
+print(plantHeight_outliers)
+plantHeight_outliers_species <- species_with_outliers(plantHeight_outliers)
+print(plantHeight_outliers_species)
+```
+
+```{r}
+summarize_species_records(traits_plantHeight)
 ```
@@
 
+```
+
+```{r}
+dotchart(traits_plantLifespan$TraitValue)
+```
+
+```{r}
+# detect outliers
+plantLifespan_outliers <- detect_outliers(traits_plantLifespan)
+print(plantLifespan_outliers)
+plantLifespan_outliers_species <- species_with_outliers(plantLifespan_outliers)
+print(plantLifespan_outliers_species)
+```
+
+```{r}
+summarize_species_records(traits_plantLifespan)
 ```
@@
 ```{r}
+dotchart(traits_seedMass$TraitValue)
+```
+
+```{r}
+# detect outliers
+seedMass_outliers <- detect_outliers(traits_seedMass)
+print(seedMass_outliers)
+seedMass_outliers_species <- species_with_outliers(seedMass_outliers)
+print(seedMass_outliers_species)
+```
+
+```{r}
+summarize_species_records(traits_seedMass)
+```
+
+```{r}
 averagetraits_seedMass <- numeric_data_average(traits_seedMass, "mg", "SeedMass_g")
@@
 ```{r}
+dotchart(traits_seedLength$TraitValue)
+```
+
+```{r}
+# detect outliers
+seedLength_outliers <- detect_outliers(traits_seedLength)
+print(seedLength_outliers)
+seedLength_outliers_species <- species_with_outliers(seedLength_outliers)
+print(seedLength_outliers_species)
+```
+
+```{r}
+summarize_species_records(traits_seedLength)
+```
+
+```{r}
 averagetraits_seedLength <- numeric_data_average(traits_seedLength, "mm", "SeedLength_mm")
@@
 ```{r}
-summary(averagetraits_fruitLength)
+summary(averagetraits_seedLength)
 ```
@@
   mutate(TraitName = "SeedWidth_mm")
+```
+
+```{r}
+dotchart(traits_seedWidth$TraitValue)
+```
+
+```{r}
+# detect outliers
+seedWidth_outliers <- detect_outliers(traits_seedWidth)
+print(seedWidth_outliers)
+seedWidth_outliers_species <- species_with_outliers(seedWidth_outliers)
+print(seedWidth_outliers_species)
+```
+
+```{r}
+summarize_species_records(traits_seedWidth)
 ```
@@
 ```{r}
+summarize_species_records(traits_growthForm)
+```
+
+```{r}
 averagetraits_growthForm <- factor_data_merge(traits_growthForm, "GrowthForm")
@@
 
+# Fruit type
+```{r}
+traits_fruitType <- plant_traits %>%
+  filter(grepl('Fruit type|fruit type|fruit_type', TraitName))
+traits_fruitType$OrigTraitValue <- traits_fruitType$TraitValue
+
+nrow(traits_fruitType)
+length(unique(traits_fruitType$Accepted_species))
+```
+
+```{r}
+# remove numeric values
+traits_fruitType <- traits_fruitType %>%
+  filter(is.na(as.numeric(TraitValue)))
+
+nrow(traits_fruitType)
+length(unique(traits_fruitType$Accepted_species))
+```
+
+```{r}
+traits_fruitType %>%
+  count(TraitValue)
+```
+
+```{r}
+# standardize trait values
+traits_fruitType <- traits_fruitType %>%
+  filter() %>%
+  mutate(TraitValue= case_when(
+    grepl('Berry|berry|Berry ', OrigTraitValue, ignore.case = TRUE) ~ "berry",
+    grepl('capsule|Capsule|Capsule |Pixidium|Dehiscent capsule|Indehiscent capsule', OrigTraitValue, ignore.case = TRUE) ~ "capsule",
+    grepl('drupe|Drupe|Drupe |Drupaceous', OrigTraitValue, ignore.case = TRUE) ~ "drupe",
+    grepl('aggregate berries', OrigTraitValue, ignore.case = TRUE) ~ "aggregate berries",
+    grepl('aggregate drupelets', OrigTraitValue, ignore.case = TRUE) ~ "aggregate drupelets",
+    grepl('fleshy', OrigTraitValue, ignore.case = TRUE) ~ "fleshy",
+    grepl('pome', OrigTraitValue, ignore.case = TRUE) ~ "pome",
+    grepl('Squash', OrigTraitValue, ignore.case = TRUE) ~ "squash",
+    grepl('aggregate follicles', OrigTraitValue, ignore.case = TRUE) ~ "aggregate follicles",
+    grepl('aggregate nutlets', OrigTraitValue, ignore.case = TRUE) ~ "aggregate nutlets",
+    grepl('achene', OrigTraitValue, ignore.case = TRUE) ~ "achene",
+    grepl('follicle|Follicle ', OrigTraitValue, ignore.case = TRUE) ~ "follicle",
+    grepl('legume|Legume ', OrigTraitValue, ignore.case = TRUE) ~ "legume",
+    grepl('lomentum', OrigTraitValue, ignore.case = TRUE) ~ "lomentum",
+    grepl('nut', OrigTraitValue, ignore.case = TRUE) ~ "nut",
+    grepl('pod|Pod|Pod ', OrigTraitValue, ignore.case = TRUE) ~ "pod",
+    grepl('Samara|Samaroid', OrigTraitValue, ignore.case = TRUE) ~ "samara",
+    grepl('schizocarp', OrigTraitValue, ignore.case = TRUE) ~ "schizocarp",
+    grepl('siliqua', OrigTraitValue, ignore.case = TRUE) ~ "siliqua",
+    grepl('syncarpous', OrigTraitValue, ignore.case = TRUE) ~ "syncarpous",
+    grepl('utricle', OrigTraitValue, ignore.case = TRUE) ~ "utricle",
+    grepl('other|apocarpous|pericarp, aril, arillode|pseudosyncarpous|vegetative|dry|fleshy', OrigTraitValue, ignore.case = TRUE) ~ "unspecified",
+  )) %>%
+  filter(!is.na(TraitValue))
+```
+
+```{r}
+traits_fruitType %>%
+  count(TraitValue)
+```
+
+```{r}
+# standardize trait name
+traits_fruitType <- traits_fruitType %>%
+  mutate(TraitName = "FruitType")
+```
+
+```{r}
+summarize_species_records(traits_fruitType)
+```
+
+```{r}
+averagetraits_fruitType <- factor_data_merge(traits_fruitType, "FruitType")
+```
+
+```{r}
+nrow(averagetraits_fruitType)
+length(unique(averagetraits_fruitType$Accepted_species))
+```
+
+```{r}
+summary(averagetraits_fruitType)
+```
+
 # Combine all traits
 ```{r}
-trait_dfs <- list(averagetraits_fruitLength, averagetraits_fruitMass, averagetraits_plantHeight, averagetraits_plantLifespan, averagetraits_seedLength, averagetraits_seedMass, averagetraits_seedWidth, averagetraits_dispersal, averagetraits_growthForm)
+trait_dfs <- list(averagetraits_fruitLength, averagetraits_fruitMass, averagetraits_plantHeight, averagetraits_plantLifespan, averagetraits_seedLength, averagetraits_seedMass, averagetraits_seedWidth, averagetraits_dispersal, averagetraits_growthForm, averagetraits_fruitType)
 
@@
 # Assign data types
+wide_plant_traits$`DispersalSyndrome` <- as.factor(wide_plant_traits$`DispersalSyndrome`)
 wide_plant_traits$`FruitLength_mm` <- as.numeric(wide_plant_traits$`FruitLength_mm`)
 wide_plant_traits$`FruitMass_mg` <- as.numeric(wide_plant_traits$`FruitMass_mg`)
+wide_plant_traits$`FruitType` <- as.factor(wide_plant_traits$`FruitType`)
+wide_plant_traits$`GrowthForm` <- as.factor(wide_plant_traits$`GrowthForm`)
 wide_plant_traits$`PlantHeight_m` <- as.numeric(wide_plant_traits$`PlantHeight_m`)
@@
 wide_plant_traits$`SeedWidth_mm` <- as.numeric(wide_plant_traits$`SeedWidth_mm`)
-wide_plant_traits$`DispersalSyndrome` <- as.factor(wide_plant_traits$`DispersalSyndrome`)
-wide_plant_traits$`GrowthForm` <- as.factor(wide_plant_traits$`GrowthForm`)
 ```

### 2024-06-11 code/L1/L1_P4_TropicalAndes_plant_traits.Rmd
add output, summary graphs, save figures into figure folder

@@
 data input: "TropicalAndes_all_plant_traits_harmonized_subset.csv"
-data output: "ADD"
+data output: "TropicalAndes_all_plant_traits_standardized.csv, TropicalAndes_all_plant_traits_standardized_with_recordCount.csv"
 date: "2023-11-04"
-output: html_document
+output: 
+  html_document:
+    code-link: true
+    df_print: paged
 ---
@@
 library(visdat)
+library(ggplot2)
 ```
@@
 plant_traits <- read.csv(file.path(data_path_L1, "TropicalAndes_all_plant_traits_harmonized_subset.csv"))
+```
+
+```{r}
+# remove X column
+plant_traits <- plant_traits[, !colnames(plant_traits) %in% "X", drop = FALSE]
+
+# Convert all character columns to UTF-8 encoding to handle encoding issues
+plant_traits <- plant_traits %>%
+    mutate(across(where(is.character), ~ iconv(., from = "", to = "UTF-8", sub = "byte")))
+
+# Replace the specific value in TraitName column
+plant_traits <- plant_traits %>%
+  mutate(TraitName = gsub("Fruit texture inside \\(fleshy, leathrey, .+\\)", "Fruit texture inside", TraitName))
 ```
@@
 ```{r}
-traits_plantLifespan <- traits %>%
+traits_plantLifespan <- plant_traits %>%
   filter(grepl('Plant lifespan \\(longevity\\)|longest whole plant longevity|maximum whole plant longevity|plant_lifespan', TraitName))
@@
 ```{r}
-traits_seedMass <- traits %>%
+traits_seedMass <- plant_traits %>%
   filter(grepl('Seed dry mass|seed mass|seed_mass_max|seed_mass_mean|seed_mass_min', TraitName))
@@
 ```{r}
-traits_seedLength <- traits %>%
+traits_seedLength <- plant_traits %>%
   filter(grepl('Seed length|seed_length_max|seed_length_mean|seed_length_min', TraitName))
@@
 ```{r}
-traits_seedWidth <- traits %>%
+traits_seedWidth <- plant_traits %>%
   filter(grepl('Seed width|seed_width_max|seed_width_mean|seed_width_min'
@@
 ```{r}
+# for summary
+unaveraged_traits <- rbind(traits_dispersal, traits_fruitLength, traits_fruitMass, traits_fruitType, traits_growthForm, traits_plantHeight, traits_plantLifespan, traits_seedLength, traits_seedMass, traits_seedWidth)
+```
+
+```{r}
 trait_dfs <- list(averagetraits_fruitLength, averagetraits_fruitMass, averagetraits_plantHeight, averagetraits_plantLifespan, averagetraits_seedLength, averagetraits_seedMass, averagetraits_seedWidth, averagetraits_dispersal, averagetraits_growthForm, averagetraits_fruitType)
@@
 combine_matching_columns <- function(df_list) {
-  # Find the common columns across all data frames
+  # Only use specified columns
   common_cols <- Reduce(intersect, lapply(df_list, colnames))
@@
 ```{r}
-wide_traits <- spread(clean_traits, key = TraitName, value = TraitValue)
-
-#remove X column from dataframe
-wide_plant_traits <- wide_traits[, !colnames(wide_traits) %in% c("X", "records_used", "variance"), drop = FALSE]
+wide_plant_traits <- clean_traits %>%
+  select(Accepted_species, TraitName, TraitValue) %>%
+  pivot_wider(names_from = TraitName, values_from = TraitValue)
+nrow(wide_plant_traits)
 ```
@@
 fb_plot_species_traits_completeness(wide_plant_traits)
+ggsave("plant_trait_completeness.png", plot = last_plot(), path = figure_path)
 fb_plot_number_species_by_trait(wide_plant_traits)
+ggsave("plant_number_species_trait.png", plot = last_plot(), path = figure_path)
 fb_table_trait_summary(wide_plant_traits)
@@
 vis_dat(wide_plant_traits)
+ggsave("plant_trait_visual_type.png", plot = last_plot(), path = figure_path)
 vis_miss(wide_plant_traits)
+ggsave("plant_trait_visual_missing.png", plot = last_plot(), path = figure_path)
+```
+
+```{r}
+# Count the number of traits per DatabaseSource and TraitName
+trait_counts <- unaveraged_traits %>%
+  count(DatabaseSource, TraitName)
+
+# Create the faceted bar plot with free y-axis scales
+ggplot(trait_counts, aes(x = DatabaseSource, y = n, fill = DatabaseSource)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Trait Counts per Database",
+       x = "Trait Name",
+       y = "Trait Count",
+       fill = "Database Source") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  facet_wrap(~ TraitName, scales = "free_y")
+ggsave("plant_trait_counts_per_database.png", plot = last_plot(), path = figure_path)
 ```
@@
 write.csv(clean_traits, file.path(output_path,"TropicalAndes_all_plant_traits_standardized_with_recordCount.csv"))
+write.csv(unaveraged_traits, file.path(output_path,"TropicalAndes_all_plant_traits_cleaned_unaveraged.csv"))
 ```
+
+# Package citations and session info
+```{r}
+library(report)
+report::cite_packages()
+
+devtools::session_info()
+```

### 2024-06-11 code/L1/L1_P5_TropicalAndes_plant_imputetraits.Rmd
add row counts and length outputs, update trait list, harmonize and clean new traits

@@
 overview: "This script fill plant trait gaps with imputation."
-data input: "ADD"
-data output: "ADD"
+data input: "TropicalAndes_all_plant_traits_standardized.csv"
+data output: "TropicalAndes_imputed_plant_traits.csv"
 date: "2023-10-04"
@@
 ```{r}
-data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+data_path_L1 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
 output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
@@
 library(GIFT)
+library(purrr)
 library(mice)
@@
 
+# Convert the wide dataframe to long format
+long_plant_traits <- plant_traits %>%
+  gather(TraitName, TraitValue, -species)
+# Filter rows with NA values
+na_traits <- long_plant_traits %>%
+  filter(is.na(TraitValue))
+dim(na_traits)
+
 # get a list of species
 na_species_list <- unique(plant_traits$species)
+length(na_species_list)
 ```
@@
   pull()
+
 # check if all the species in no_family & no_genus lists are the same
 identical(species_no_family, species_no_genus)
+
+length(species_no_family)
 ```
@@
   pull()
+length(species_no_family_2)
 ```
@@
   pull()
+length(species_no_family_3)
 ```
@@
 # Separate dataframes from list
+PlantLifespan_years <- trait_dfs$PlantLifespan_years
+FruitType <- trait_dfs$FruitType
 DispersalSyndrome <- trait_dfs$DispersalSyndrome
-PlantLifespan_years <- trait_dfs$PlantLifespan_years
 FruitMass_mg <- trait_dfs$FruitMass_mg
-PlantHeight_m <- trait_dfs$PlantHeight_m
+SeedLength_mm <- trait_dfs$SeedLength_mm
+SeedWidth_mm <- trait_dfs$SeedWidth_mm
 GrowthForm <- trait_dfs$GrowthForm
 FruitLength_mm <- trait_dfs$FruitLength_mm
-SeedWidth_mm <- trait_dfs$SeedWidth_mm
-SeedLength_mm <- trait_dfs$SeedLength_mm
 SeedMass_g <- trait_dfs$SeedMass_g
+PlantHeight_m <- trait_dfs$PlantHeight_m
 ```
@@
 ## Create a list of unique genus and families for each trait
+```{r}
+PlantLifespan_years_families <- unique(PlantLifespan_years$family)
+PlantLifespan_years_genus <- unique(PlantLifespan_years$genus)
+length(PlantLifespan_years_families)
+length(PlantLifespan_years_genus)
+```
+
+```{r}
+FruitType_families <- unique(FruitType$family)
+FruitType_genus <- unique(FruitType$genus)
+length(FruitType_families)
+length(FruitType_genus)
+```
+
 ```{r}
@@
 ```{r}
-PlantLifespan_years_families <- unique(PlantLifespan_years$family)
-PlantLifespan_years_genus <- unique(PlantLifespan_years$genus)
-length(PlantLifespan_years_families)
-length(PlantLifespan_years_genus)
+FruitMass_mg_families <- unique(FruitMass_mg$family)
+FruitMass_mg_genus <- unique(FruitMass_mg$genus)
+length(FruitMass_mg_families)
+length(FruitMass_mg_genus)
 ```
@@
 ```{r}
+SeedLength_mm_families <- unique(SeedLength_mm$family)
+SeedLength_mm_genus <- unique(SeedLength_mm$genus)
+length(SeedLength_mm_families)
+length(SeedLength_mm_genus)
+```
 
-FruitMass_mg_families <- unique(FruitMass_mg$family)
-FruitMass_mg_genus <- unique(FruitMass_mg$genus)
+```{r}
+SeedWidth_mm_families <- unique(SeedWidth_mm$family)
+SeedWidth_mm_genus <- unique(SeedWidth_mm$genus)
+length(SeedWidth_mm_families)
+length(SeedWidth_mm_genus)
+```
 
+```{r}
+GrowthForm_families <- unique(GrowthForm$family)
+GrowthForm_genus <- unique(GrowthForm$genus)
+length(GrowthForm_families)
+length(GrowthForm_genus)
+```
+
+```{r}
+FruitLength_mm_families <- unique(FruitLength_mm$family)
+FruitLength_mm_genus <- unique(FruitLength_mm$genus)
+length(FruitLength_mm_families)
+length(FruitLength_mm_genus)
+```
+
+```{r}
+SeedMass_g_families <- unique(SeedMass_g$family)
+SeedMass_g_genus <- unique(SeedMass_g$genus)
+length(SeedMass_g_families)
+length(SeedMass_g_genus)
+```
+
+```{r}
 PlantHeight_m_families <- unique(PlantHeight_m$family)
 PlantHeight_m_genus <- unique(PlantHeight_m$genus)
-
-GrowthForm_families <- unique(GrowthForm$family)
-GrowthForm_genus <- unique(GrowthForm$genus)
-
-FruitLength_mm_families <- unique(FruitLength_mm$family)
-FruitLength_mm_genus <- unique(FruitLength_mm$genus)
-
-SeedWidth_mm_families <- unique(SeedWidth_mm$family)
-SeedWidth_mm_genus <- unique(SeedWidth_mm$genus)
-
-SeedLength_mm_families <- unique(SeedLength_mm$family)
-SeedLength_mm_genus <- unique(SeedLength_mm$genus)
-
-SeedMass_g_families <- unique(SeedMass_g$family)
-SeedMass_g_genus <- unique(SeedMass_g$genus)
+length(PlantHeight_m_families)
+length(PlantHeight_m_genus)
 ```
@@
 ## Retrive trait records for species using BIEN & GIFT
+### Plant Lifespan
+```{r}
+PlantLifespan_years_families_BIEN <- BIEN_trait_traitbyfamily(family = PlantLifespan_years_families, trait = "maximum whole plant longevity")
+PlantLifespan_years_genus_BIEN <- BIEN_trait_traitbygenus(genus = PlantLifespan_years_genus, trait = "maximum whole plant longevity")
+
+# Error in GIFT_traits_tax(trait_IDs = "2.2.1") :   None of the traits asked was available at the taxonomic level.
+#PlantLifespan_years_families_GIFT <- GIFT_traits_tax(trait_IDs = "2.2.1")
+#PlantLifespan_years_families_GIFT <- PlantLifespan_years_families_GIFT[PlantLifespan_years_families_GIFT$family %in% PlantLifespan_years_families, ]
+```
+
+```{r}
+nrow(PlantLifespan_years_families_BIEN)
+length(unique(PlantLifespan_years_families_BIEN$scrubbed_family))
+unique(PlantLifespan_years_families_BIEN$trait_value)
+```
+
+```{r}
+nrow(PlantLifespan_years_genus_BIEN)
+length(unique(PlantLifespan_years_genus_BIEN$scrubbed_genus))
+unique(PlantLifespan_years_genus_BIEN$trait_value)
+```
+
+### Fruit Type
+```{r}
+FruitType_families_BIEN <- BIEN_trait_traitbyfamily(family = FruitType_families, trait = "fruit type")
+FruitType_genus_BIEN <- BIEN_trait_traitbygenus(genus = FruitType_genus, trait = "fruit type")
+
+# Error in GIFT_traits_tax(trait_IDs = "3.16.1") :   None of the traits asked was available at the taxonomic level.
+#Fruit.Type_families_GIFT <- GIFT_traits_tax(trait_IDs = "3.16.1")
+#Fruit.Type_families_GIFT <- Fruit.Type_families_GIFT[Fruit.Type_families_GIFT$family %in% Fruit.Type_families, ]
+```
+
+```{r}
+nrow(FruitType_families_BIEN)
+length(unique(FruitType_families_BIEN$scrubbed_family))
+unique(FruitType_families_BIEN$trait_value)
+```
+
+```{r}
+nrow(FruitType_genus_BIEN)
+length(unique(FruitType_genus_BIEN$scrubbed_genus))
+unique(FruitType_genus_BIEN$trait_value)
+```
+
 ### Dispersal Syndrome
@@
 ```
+This information is not useful to fill in trait gaps.
+
+### Fruit Mass
+No information in BIEN or GIFT
+
+### Seed Length
+Only in GIFT: Seed_length_min (3.10.1), Seed_length_max (3.10.2), Seed_length_mean (3.10.3)
+```{r}
+SeedLength_mm_families_GIFT <- GIFT_traits_tax(trait_IDs = c("3.10.1", "3.10.2", "3.10.3"))
+SeedLength_mm_families_GIFT <- SeedLength_mm_families_GIFT[SeedLength_mm_families_GIFT$family %in% SeedLength_mm_families, ]
+```
+
+```{r}
+nrow(SeedLength_mm_families_GIFT)
+```
+
+### Seed Width
+Only in GIFT: Seed_width_min (3.11.1), Seed_width_max (3.11.2), Seed_width_mean (3.11.3)
+```{r}
+#Error in GIFT_traits_tax(trait_IDs = c("3.11.1", "3.11.2", "3.11.3")) :   None of the traits asked was available at the taxonomic level.
+#SeedWidth_mm_families_GIFT <- GIFT_traits_tax(trait_IDs = c("3.11.1", "3.11.2", "3.11.3"))
+#SeedWidth_mm_families_GIFT <- SeedWidth_mm_families_GIFT[SeedWidth_mm_families_GIFT$family %in% SeedWidth_mm_families, ]
+```
+
+### Growth Form
+```{r}
+# Growth Form
+GrowthForm_families_BIEN <- BIEN_trait_traitbyfamily(family = GrowthForm_families, trait = "whole plant growth form")
+GrowthForm_genus_BIEN <- BIEN_trait_traitbygenus(genus = GrowthForm_genus, trait = "whole plant growth form")
+
+# Error in GIFT_traits_tax(trait_IDs = "1.2.1") :   None of the traits asked was available at the taxonomic level.
+#Growth.Form_families_GIFT <- GIFT_traits_tax(trait_IDs = "1.2.1")
+#Growth.Form_families_GIFT <- Growth.Form_families_GIFT[Growth.Form_families_GIFT$family %in% Growth.Form_families, ]
+```
+
+```{r}
+nrow(GrowthForm_families_BIEN)
+length(unique(GrowthForm_families_BIEN$scrubbed_family))
+unique(GrowthForm_families_BIEN$trait_value)
+```
+
+```{r}
+nrow(GrowthForm_genus_BIEN)
+length(unique(GrowthForm_genus_BIEN$scrubbed_genus))
+unique(GrowthForm_genus_BIEN$trait_value)
+```
 
@@
 ```{r}
-Fruit.Length_families_BIEN <- BIEN_trait_traitbyfamily(family = Fruit.Length_families, trait = c("maximum fruit length", "minimum fruit length"))
-Fruit.Length_genus_BIEN <- BIEN_trait_traitbygenus(genus = Fruit.Length_genus, trait = c("maximum fruit length", "minimum fruit length"))
+FruitLength_mm_families_BIEN <- BIEN_trait_traitbyfamily(family = FruitLength_mm_families, trait = c("maximum fruit length", "minimum fruit length"))
+FruitLength_mm_genus_BIEN <- BIEN_trait_traitbygenus(genus = FruitLength_mm_genus, trait = c("maximum fruit length", "minimum fruit length"))
 
@@
 ```{r}
-nrow(DispersalSyndrome_families_BIEN)
-length(unique(DispersalSyndrome_families_BIEN$scrubbed_family))
-unique(DispersalSyndrome_families_BIEN$trait_value)
+nrow(FruitLength_mm_families_BIEN)
+length(unique(FruitLength_mm_families_BIEN$scrubbed_family))
+unique(FruitLength_mm_families_BIEN$trait_value)
 ```
@@
 ```{r}
-# Plant Height
-Plant.Height_families_BIEN <- BIEN_trait_traitbyfamily(family = Plant.Height_families, trait = "whole plant height")
-Plant.Height_genus_BIEN <- BIEN_trait_traitbygenus(genus = Plant.Height_genus, trait = "whole plant height")
+nrow(FruitLength_mm_genus_BIEN)
+length(unique(FruitLength_mm_genus_BIEN$scrubbed_genus))
+unique(FruitLength_mm_genus_BIEN$trait_value)
+```
+
+### Seed Mass
+```{r}
+SeedMass_g_families_BIEN <- BIEN_trait_traitbyfamily(family = SeedMass_g_families, trait = "seed mass")
+SeedMass_g_genus_BIEN <- BIEN_trait_traitbygenus(genus = SeedMass_g_genus, trait = "seed mass")
+
+SeedMass_g_families_GIFT <- GIFT_traits_tax(trait_IDs = c("3.2.1", "3.2.2", "3.2.3"))
+SeedMass_g_families_GIFT <- SeedMass_g_families_GIFT[SeedMass_g_families_GIFT$family %in% SeedMass_g_families, ]
+```
+
+```{r}
+nrow(SeedMass_g_families_BIEN)
+length(unique(SeedMass_g_families_BIEN$scrubbed_family))
+unique(SeedMass_g_families_BIEN$trait_value)
+```
+
+```{r}
+nrow(SeedMass_g_genus_BIEN)
+length(unique(SeedMass_g_genus_BIEN$scrubbed_genus))
+unique(SeedMass_g_genus_BIEN$trait_value)
+```
+
+```{r}
+nrow(SeedMass_g_families_GIFT)
+```
+
+### Plant Height
+```{r}
+PlantHeight_m_families_BIEN <- BIEN_trait_traitbyfamily(family = PlantHeight_m_families, trait = "whole plant height")
+PlantHeight_m_genus_BIEN <- BIEN_trait_traitbygenus(genus = PlantHeight_m_genus, trait = "whole plant height")
 
@@
 ```{r}
-# Fruit Type
-Fruit.Type_families_BIEN <- BIEN_trait_traitbyfamily(family = Fruit.Type_families, trait = "fruit type")
-Fruit.Type_genus_BIEN <- BIEN_trait_traitbygenus(genus = Fruit.Type_genus, trait = "fruit type")
-
-# Error in GIFT_traits_tax(trait_IDs = "3.16.1") :   None of the traits asked was available at the taxonomic level.
-#Fruit.Type_families_GIFT <- GIFT_traits_tax(trait_IDs = "3.16.1")
-#Fruit.Type_families_GIFT <- Fruit.Type_families_GIFT[Fruit.Type_families_GIFT$family %in% Fruit.Type_families, ]
+nrow(PlantHeight_m_families_BIEN)
+length(unique(PlantHeight_m_families_BIEN$scrubbed_family))
+unique(PlantHeight_m_families_BIEN$trait_value)
 ```
@@
 ```{r}
-# Growth Form
-Growth.Form_families_BIEN <- BIEN_trait_traitbyfamily(family = Growth.Form_families, trait = "whole plant growth form")
-Growth.Form_genus_BIEN <- BIEN_trait_traitbygenus(genus = Growth.Form_genus, trait = "whole plant growth form")
+nrow(PlantHeight_m_genus_BIEN)
+length(unique(PlantHeight_m_genus_BIEN$scrubbed_genus))
+unique(PlantHeight_m_genus_BIEN$trait_value)
+```
 
-# Error in GIFT_traits_tax(trait_IDs = "1.2.1") :   None of the traits asked was available at the taxonomic level.
-#Growth.Form_families_GIFT <- GIFT_traits_tax(trait_IDs = "1.2.1")
-#Growth.Form_families_GIFT <- Growth.Form_families_GIFT[Growth.Form_families_GIFT$family %in% Growth.Form_families, ]
+## Clean new trait data
+Make sure numeric traits match the units of existing trait table. Numeric traits should only contain numeric values and non-numeric traits should only have non-numeric trait values. 
+### # Lifespan
+remove non-numeric trait values
+```{r}
+PlantLifespan_years_families_BIEN <- PlantLifespan_years_families_BIEN %>%
+  filter(!is.na(as.numeric(trait_value))) %>%
+  mutate(trait_value = as.numeric(trait_value))
+
+nrow(PlantLifespan_years_families_BIEN)
+unique(PlantLifespan_years_families_BIEN$unit)
+
+PlantLifespan_years_genus_BIEN <- PlantLifespan_years_genus_BIEN %>%
+  filter(!is.na(as.numeric(trait_value))) %>%
+  mutate(trait_value = as.numeric(trait_value))
+
+nrow(PlantLifespan_years_genus_BIEN)
+unique(PlantLifespan_years_genus_BIEN$unit)
+```
+
+# Fruit type
+haromonize trait values
+```{r}
+unique(FruitType_families_BIEN$trait_value)
+unique(FruitType_genus_BIEN$trait_value)
+FruitType_families_BIEN <- FruitType_families_BIEN %>%
+  filter() %>%
+  mutate(trait_value = case_when(
+    grepl('Berry|berry|Berry ', trait_value, ignore.case = TRUE) ~ "berry",
+    grepl('capsule|Capsule|Capsule ', trait_value, ignore.case = TRUE) ~ "capsule",
+    grepl('drupe|Drupe|Drupe |Drupaceous', trait_value, ignore.case = TRUE) ~ "drupe",
+    grepl('Squash', trait_value, ignore.case = TRUE) ~ "squash",
+    grepl('pod|Pod|Pod ', trait_value, ignore.case = TRUE) ~ "pod",
+    grepl('Samara|Samaroid', trait_value, ignore.case = TRUE) ~ "samara"
+  )) %>%
+  filter(!is.na(trait_value))
+FruitType_genus_BIEN <- FruitType_genus_BIEN %>%
+  filter() %>%
+  mutate(trait_value = case_when(
+    grepl('Berry|berry|Berry ', trait_value, ignore.case = TRUE) ~ "berry",
+    grepl('capsule|Capsule|Capsule ', trait_value, ignore.case = TRUE) ~ "capsule",
+    grepl('drupe|Drupe|Drupe |Drupaceous', trait_value, ignore.case = TRUE) ~ "drupe",
+    grepl('Squash', trait_value, ignore.case = TRUE) ~ "squash",
+    grepl('pod|Pod|Pod ', trait_value, ignore.case = TRUE) ~ "pod",
+    grepl('Samara|Samaroid', trait_value, ignore.case = TRUE) ~ "samara"
+  )) %>%
+  filter(!is.na(trait_value))
+unique(FruitType_families_BIEN$trait_value)
+unique(FruitType_genus_BIEN$trait_value)
+```
+
+# Growth form
+harmonize trait values and remove numeric values
+```{r}
+# remove numeric rows
+GrowthForm_families_BIEN <- GrowthForm_families_BIEN %>%
+  filter(is.na(as.numeric(trait_value)))
+nrow(GrowthForm_families_BIEN)
+GrowthForm_genus_BIEN <- GrowthForm_genus_BIEN %>%
+  filter(is.na(as.numeric(trait_value)))
+nrow(GrowthForm_genus_BIEN)
+```
+
+```{r}
+# classify values as herb, shrub, tree, or other
+GrowthForm_families_BIEN <- GrowthForm_families_BIEN %>%
+  mutate(Category = case_when(
+    grepl("herb", trait_value, ignore.case = TRUE) ~ "herb",
+    grepl("shrub", trait_value, ignore.case = TRUE) ~ "shrub",
+    grepl("tree", trait_value, ignore.case = TRUE) ~ "tree",
+    TRUE ~ "other"
+  ))
+GrowthForm_genus_BIEN <- GrowthForm_genus_BIEN %>%
+  mutate(Category = case_when(
+    grepl("herb", trait_value, ignore.case = TRUE) ~ "herb",
+    grepl("shrub", trait_value, ignore.case = TRUE) ~ "shrub",
+    grepl("tree", trait_value, ignore.case = TRUE) ~ "tree",
+    TRUE ~ "other"
+  ))
+```
+
+### Fruit length
+remove non-numeric trait values and check units
+```{r}
+FruitLength_mm_families_BIEN <- FruitLength_mm_families_BIEN %>%
+  filter(!is.na(as.numeric(trait_value))) %>%
+  mutate(trait_value = as.numeric(trait_value))
+
+nrow(FruitLength_mm_families_BIEN)
+unique(FruitLength_mm_families_BIEN$unit)
+
+FruitLength_mm_genus_BIEN <- FruitLength_mm_genus_BIEN %>%
+  filter(!is.na(as.numeric(trait_value))) %>%
+  mutate(trait_value = as.numeric(trait_value))
+
+nrow(FruitLength_mm_genus_BIEN)
+unique(FruitLength_mm_genus_BIEN$unit)
+```
+
+### Seed mass
+remove non-numeric trait values and check units
+```{r}
+SeedMass_g_families_BIEN <- SeedMass_g_families_BIEN %>%
+  filter(!is.na(as.numeric(trait_value))) %>%
+  mutate(trait_value = as.numeric(trait_value))
+
+nrow(SeedMass_g_families_BIEN)
+unique(SeedMass_g_families_BIEN$unit)
+
+SeedMass_g_genus_BIEN <- SeedMass_g_genus_BIEN %>%
+  filter(!is.na(as.numeric(trait_value))) %>%
+  mutate(trait_value = as.numeric(trait_value))
+
+nrow(SeedMass_g_genus_BIEN)
+unique(SeedMass_g_genus_BIEN$unit)
+
+```
+
+### Plant Height
+remove non-numeric trait values and check units
+```{r}
+PlantHeight_m_families_BIEN <- PlantHeight_m_families_BIEN %>%
+  filter(!is.na(as.numeric(trait_value))) %>%
+  mutate(trait_value = as.numeric(trait_value))
+
+nrow(PlantHeight_m_families_BIEN)
+unique(PlantHeight_m_families_BIEN$unit)
+
+PlantHeight_m_genus_BIEN <- PlantHeight_m_genus_BIEN %>%
+  filter(!is.na(as.numeric(trait_value))) %>%
+  mutate(trait_value = as.numeric(trait_value))
+
+nrow(PlantHeight_m_genus_BIEN)
+unique(PlantHeight_m_genus_BIEN$unit)
 ```
@@
 ## Combine new info with existing trait table
-Make sure numeric traits match the units of existing trait table. 
+
 To fill in gaps of missing traits, use genus level data first then family level. Add column to trait dataframe of the level of trait measurement: species, genus, family.
+
+### Subset dataframe to only keep required columns
 ```{r}
 # for traits to family level from BIEN, keep columns of scrubbed_family, trait_name, trait_value
-Dispersal.Syndrome_families_BIEN_df <- Dispersal.Syndrome_families_BIEN[, c("scrubbed_family", "trait_name", "trait_value")]
-Fruit.Length_families_BIEN_df <- Fruit.Length_families_BIEN[, c("scrubbed_family", "trait_name", "trait_value")]
-Plant.Height_families_BIEN_df <- Plant.Height_families_BIEN[, c("scrubbed_family", "trait_name", "trait_value")]
-Fruit.Type_families_BIEN_df <- Fruit.Type_families_BIEN[, c("scrubbed_family", "trait_name", "trait_value")]
-Growth.Form_families_BIEN_df <- Growth.Form_families_BIEN[, c("scrubbed_family", "trait_name", "trait_value")]
+PlantLifespan_years_families_BIEN_df <- PlantLifespan_years_families_BIEN[, c("scrubbed_family", "trait_name", "trait_value")]
+FruitType_families_BIEN_df <- FruitType_families_BIEN[, c("scrubbed_family", "trait_name", "trait_value")]
+GrowthForm_families_BIEN_df <- GrowthForm_families_BIEN[, c("scrubbed_family", "trait_name", "trait_value")]
+FruitLength_mm_families_BIEN_df <- FruitLength_mm_families_BIEN[, c("scrubbed_family", "trait_name", "trait_value")]
+SeedMass_g_families_BIEN_df <- SeedMass_g_families_BIEN[, c("scrubbed_family", "trait_name", "trait_value")]
+PlantHeight_m_families_BIEN_df <- PlantHeight_m_families_BIEN[, c("scrubbed_family", "trait_name", "trait_value")]
 ```
@@
 # for traits to genus level from BIEN, keep columns of scrubbed_genus, trait_name, trait_value
-Dispersal.Syndrome_genus_BIEN_df <- Dispersal.Syndrome_genus_BIEN[, c("scrubbed_genus", "trait_name", "trait_value")]
-Fruit.Length_genus_BIEN_df <- Fruit.Length_genus_BIEN[, c("scrubbed_genus", "trait_name", "trait_value")]
-Plant.Height_genus_BIEN_df <- Plant.Height_genus_BIEN[, c("scrubbed_genus", "trait_name", "trait_value")]
-Fruit.Type_genus_BIEN_df <- Fruit.Type_genus_BIEN[, c("scrubbed_genus", "trait_name", "trait_value")]
-Growth.Form_genus_BIEN_df <- Growth.Form_genus_BIEN[, c("scrubbed_genus", "trait_name", "trait_value")]
+PlantLifespan_years_genus_BIEN_df <- PlantLifespan_years_genus_BIEN[, c("scrubbed_genus", "trait_name", "trait_value")]
+FruitType_genus_BIEN_df <- FruitType_genus_BIEN[, c("scrubbed_genus", "trait_name", "trait_value")]
+GrowthForm_genus_BIEN_df <- GrowthForm_genus_BIEN[, c("scrubbed_genus", "trait_name", "trait_value")]
+FruitLength_mm_genus_BIEN_df <- FruitLength_mm_genus_BIEN[, c("scrubbed_genus", "trait_name", "trait_value")]
+SeedMass_g_genus_BIEN_df <- SeedMass_g_genus_BIEN[, c("scrubbed_genus", "trait_name", "trait_value")]
+PlantHeight_m_genus_BIEN_df <- PlantHeight_m_genus_BIEN[, c("scrubbed_genus", "trait_name", "trait_value")]
 ```
 
-Get a single trait value for each family/genus
+### Get a single trait value for each family/genus
 ```{r}
 # numeric traits
-numeric_traits_combined <- function(df, level) {
+numeric_traits_combined <- function(df, level, traitname) {
   df$trait_value <- as.numeric(df$trait_value)
   if (level == "genus") {
-    df %>%
+    summary_data <- df %>%
       group_by(scrubbed_genus) %>%
-      summarise(trait_value = mean(trait_value, na.rm = TRUE)) %>%
-      ungroup()
+          summarise(
+            trait_value_mean = exp(mean(log(trait_value))),
+            records_used = n(),
+            variance = var(trait_value, na.rm = TRUE)
+          )
+      # Add trait name column
+      summary_data$TraitName <- traitname
+
+      # Rename the mean column to TraitValue
+      summary_data <- summary_data %>%
+        rename(trait_value = trait_value_mean)
+
+    return(summary_data)
   } else if (level == "family") {
-    df %>%
+    summary_data <- df %>%
       group_by(scrubbed_family) %>%
-      summarise(trait_value = mean(trait_value, na.rm = TRUE)) %>%
-      ungroup()
+      summarise(
+            trait_value_mean = exp(mean(log(trait_value))),
+            records_used = n(),
+            variance = var(trait_value, na.rm = TRUE)
+          )
+      # Add trait name column
+      summary_data$TraitName <- traitname
+
+      # Rename the mean column to TraitValue
+      summary_data <- summary_data %>%
+        rename(TraitValue = trait_value_mean)
+
+    return(summary_data)
   }
@@
 # non-numeric traits
-cat_traits_combined <- function(df, level){
+cat_traits_combined <- function(df, level, traitname){
   if (level == "genus") {
-    df %>%
+    summary_data <- df %>%
       group_by(scrubbed_genus) %>%
-      summarise_all(.funs = list(mode_trait = ~names(sort(table(.), decreasing = TRUE))[1])) %>%
-      ungroup()
+      summarise(
+        ModeValue = names(sort(table(trait_value), decreasing = TRUE))[1],
+        records_used = n(),
+        mode_freq = max(table(trait_value)),
+        variance = 1 - (max(table(trait_value)) / n())
+    )
+
+    # Add trait name column
+    summary_data$TraitName <- traitname
+
+    # Rename the mean column to TraitValue
+    summary_data <- summary_data %>%
+      rename(TraitValue = ModeValue)
+
+  return(summary_data)
   } else if (level == "family") {
-    df %>%
+    summary_data <- df %>%
       group_by(scrubbed_family) %>%
-      summarise_all(.funs = list(mode_trait = ~names(sort(table(.), decreasing = TRUE))[1])) %>%
-      ungroup()
+      summarise(
+        ModeValue = names(sort(table(trait_value), decreasing = TRUE))[1],
+        records_used = n(),
+        mode_freq = max(table(trait_value)),
+        variance = 1 - (max(table(trait_value)) / n())
+    )
+
+    # Add trait name column
+    summary_data$TraitName <- traitname
+
+    # Rename the mean column to TraitValue
+    summary_data <- summary_data %>%
+      rename(TraitValue = ModeValue)
+
+  return(summary_data)
   }
@@
 ```{r}
-Dispersal.Syndrome_families_BIEN_df <- cat_traits_combined(Dispersal.Syndrome_families_BIEN_df, "family")
-Dispersal.Syndrome_genus_BIEN_df <- cat_traits_combined(Dispersal.Syndrome_genus_BIEN_df, "genus")
+dim(PlantLifespan_years_families_BIEN_df)
+PlantLifespan_years_families_BIEN_df_average <- numeric_traits_combined(PlantLifespan_years_families_BIEN_df, "family", "PlantLifespan_years")
+dim(PlantLifespan_years_families_BIEN_df_average)
+
+dim(PlantLifespan_years_genus_BIEN_df)
+PlantLifespan_years_genus_BIEN_df_average <- numeric_traits_combined(PlantLifespan_years_genus_BIEN_df, "genus", "PlantLifespan_years")
+dim(PlantLifespan_years_genus_BIEN_df_average)
 ```
@@
 ```{r}
-Fruit.Length_families_BIEN_df <- numeric_traits_combined(Fruit.Length_families_BIEN_df, "family")
-Fruit.Length_genus_BIEN_df <- numeric_traits_combined(Fruit.Length_genus_BIEN_df, "genus")
+dim(FruitType_families_BIEN_df)
+FruitType_families_BIEN_df_average <- cat_traits_combined(FruitType_families_BIEN_df, "family", "FruitType")
+dim(FruitType_families_BIEN_df_average)
+
+dim(FruitType_genus_BIEN_df)
+FruitType_genus_BIEN_df_average <- cat_traits_combined(FruitType_genus_BIEN_df, "genus", "FruitType")
+dim(FruitType_genus_BIEN_df_average)
 ```
@@
 ```{r}
-Plant.Height_families_BIEN_df <- numeric_traits_combined(Plant.Height_families_BIEN_df, "family")
-Plant.Height_genus_BIEN_df <- numeric_traits_combined(Plant.Height_genus_BIEN_df, "genus")
+dim(GrowthForm_families_BIEN_df)
+GrowthForm_families_BIEN_df_average <- cat_traits_combined(GrowthForm_families_BIEN_df, "family", "GrowthForm")
+dim(GrowthForm_families_BIEN_df_average)
+
+dim(GrowthForm_genus_BIEN_df)
+GrowthForm_genus_BIEN_df_average <- cat_traits_combined(GrowthForm_genus_BIEN_df, "genus", "GrowthForm")
+dim(GrowthForm_genus_BIEN_df_average)
 ```
@@
 ```{r}
-Fruit.Type_families_BIEN_df <- cat_traits_combined(Fruit.Type_families_BIEN_df, "family")
-Fruit.Type_genus_BIEN_df <- cat_traits_combined(Fruit.Type_genus_BIEN_df, "genus")
+dim(FruitLength_mm_families_BIEN_df)
+FruitLength_mm_families_BIEN_df_average <- numeric_traits_combined(FruitLength_mm_families_BIEN_df, "family", "FruitLength_mm")
+dim(FruitLength_mm_families_BIEN_df_average)
+
+dim(FruitLength_mm_genus_BIEN_df)
+FruitLength_mm_genus_BIEN_df_average <- numeric_traits_combined(FruitLength_mm_genus_BIEN_df, "genus", "FruitLength_mm")
+dim(FruitLength_mm_genus_BIEN_df_average)
 ```
@@
 ```{r}
-Growth.Form_families_BIEN_df <- cat_traits_combined(Growth.Form_families_BIEN_df, "family")
-Growth.Form_genus_BIEN_df <- cat_traits_combined(Growth.Form_genus_BIEN_df, "genus")
+dim(SeedMass_g_families_BIEN_df)
+SeedMass_g_families_BIEN_df_average <- numeric_traits_combined(SeedMass_g_families_BIEN_df, "family", "SeedMass_g")
+dim(PlantLifespan_years_families_BIEN_df_average)
+
+dim(SeedMass_g_genus_BIEN_df)
+SeedMass_g_genus_BIEN_df_average <- numeric_traits_combined(SeedMass_g_genus_BIEN_df, "genus", "SeedMass_g")
+dim(SeedMass_g_genus_BIEN_df_average)
 ```
 
-add column of trait level (species, genus, family)
+```{r}
+dim(PlantHeight_m_families_BIEN_df)
+PlantHeight_m_families_BIEN_df_average <- numeric_traits_combined(PlantHeight_m_families_BIEN_df, "family", "PlantHeight_m")
+dim(PlantHeight_m_families_BIEN_df_average)
+
+dim(PlantHeight_m_genus_BIEN_df)
+PlantHeight_m_genus_BIEN_df_average <- numeric_traits_combined(PlantHeight_m_genus_BIEN_df, "genus", "PlantHeight_m")
+dim(PlantHeight_m_genus_BIEN_df_average)
+```
+
+### Add column of trait level (species, genus, family)
 ```{r}
@@
 
+#### Combine all family level trait data
 ```{r}
-# combine all family level trait data
-## add column of trait_name for numeric traits
-Dispersal.Syndrome_families_BIEN_df$trait_name <- "Dispersal.Syndrome"
-Fruit.Length_families_BIEN_df$trait_name <- "Fruit.Length_mm"
-Plant.Height_families_BIEN_df$trait_name <- "Plant.Height_m"
-Fruit.Type_families_BIEN_df$trait_name <- "Fruit.Type"
-Growth.Form_families_BIEN_df$trait_name <- "Growth.Form"
+colnames(PlantLifespan_years_families_BIEN_df_average)
+colnames(FruitType_families_BIEN_df_average)
+colnames(GrowthForm_families_BIEN_df_average)
+colnames(FruitLength_mm_families_BIEN_df_average)
+colnames(SeedMass_g_families_BIEN_df_average)
+colnames(PlantHeight_m_families_BIEN_df_average)
+```
 
-## rename trait_value column so it's consistent with other traits
-Fruit.Type_families_BIEN_df$TraitValue <- Fruit.Type_families_BIEN_df$trait_value_mode_trait
-Growth.Form_families_BIEN_df$TraitValue <- Growth.Form_families_BIEN_df$trait_value_mode_trait
+```{r}
+family_level_trait_data <- list(PlantLifespan_years_families_BIEN_df_average, FruitType_families_BIEN_df_average, GrowthForm_families_BIEN_df_average, FruitLength_mm_families_BIEN_df_average, SeedMass_g_families_BIEN_df_average, PlantHeight_m_genus_BIEN_df_average)
 
-## convert numeric trait values into character
-Plant.Height_families_BIEN_df$TraitValue <- as.character(Plant.Height_families_BIEN_df$TraitValue)
-Fruit.Length_families_BIEN_df$TraitValue <- as.character(Fruit.Length_families_BIEN_df$TraitValue)
+# Function to select and convert columns to character
+select_and_convert <- function(df) {
+  df %>%
+    select(scrubbed_family, TraitName, TraitValue) %>%
+    mutate(across(everything(), as.character))
+}
 
-all_family_traits <- bind_rows(
-  select(Dispersal.Syndrome_families_BIEN_df, scrubbed_family, trait_name, TraitValue),
-  select(Fruit.Length_families_BIEN_df, scrubbed_family, trait_name, TraitValue),
-  select(Plant.Height_families_BIEN_df, scrubbed_family, trait_name, TraitValue),
-  select(Fruit.Type_families_BIEN_df, scrubbed_family, trait_name, TraitValue),
-  select(Growth.Form_families_BIEN_df, scrubbed_family, trait_name, TraitValue)
-  )
+# Apply the function to each dataframe in the list and combine them
+all_family_traits <- family_level_trait_data %>%
+  purrr::map_df(select_and_convert)
+```
 
+```{r}
 ## add trait level column
@@
 
+#### Combine all genus level trait data
 ```{r}
-# combine all genus level trait data
 ## add column of trait_name for numeric traits

### 2024-06-11 code/L1/L1_P5_TropicalAndes_plant_imputetraits.Rmd
add visualization packages and code

@@
 library(dplyr)
+library(funbiogeo)
+library(visdat)
+library(ggplot2)
+library(forcats)
 ```
@@
       summary_data <- summary_data %>%
-        rename(trait_value = trait_value_mean)
+        rename(TraitValue = trait_value_mean)
 
@@
 ```{r}
-family_level_trait_data <- list(PlantLifespan_years_families_BIEN_df_average, FruitType_families_BIEN_df_average, GrowthForm_families_BIEN_df_average, FruitLength_mm_families_BIEN_df_average, SeedMass_g_families_BIEN_df_average, PlantHeight_m_genus_BIEN_df_average)
+family_level_trait_data <- list(PlantLifespan_years_families_BIEN_df_average, 
+                                FruitType_families_BIEN_df_average, 
+                                GrowthForm_families_BIEN_df_average, 
+                                FruitLength_mm_families_BIEN_df_average, 
+                                SeedMass_g_families_BIEN_df_average, 
+                                PlantHeight_m_families_BIEN_df_average)
 
@@
   df %>%
-    select(scrubbed_family, TraitName, TraitValue) %>%
+    dplyr::select(scrubbed_family, TraitName, TraitValue) %>%
     mutate(across(everything(), as.character))
@@
 all_family_traits <- family_level_trait_data %>%
-  purrr::map_df(select_and_convert)
+  map_df(select_and_convert) %>%
+  bind_rows()
 ```
@@
 ```{r}
-## add column of trait_name for numeric traits
-Dispersal.Syndrome_genus_BIEN_df$trait_name <- "Dispersal.Syndrome"
-Fruit.Length_genus_BIEN_df$trait_name <- "Fruit.Length_mm"
-Plant.Height_genus_BIEN_df$trait_name <- "Plant.Height_m"
-Fruit.Type_genus_BIEN_df$trait_name <- "Fruit.Type"
-Growth.Form_genus_BIEN_df$trait_name <- "Growth.Form"
-
-## rename TraitValue column so it's consistent with other traits
-Fruit.Type_genus_BIEN_df$TraitValue <- Fruit.Type_genus_BIEN_df$trait_value_mode_trait
-Growth.Form_genus_BIEN_df$TraitValue <- Growth.Form_genus_BIEN_df$trait_value_mode_trait
-
-## convert numeric trait values into character
-Plant.Height_genus_BIEN_df$TraitValue <- as.character(Plant.Height_genus_BIEN_df$TraitValue)
-Fruit.Length_genus_BIEN_df$TraitValue <- as.character(Fruit.Length_genus_BIEN_df$TraitValue)
-
-all_genus_traits <- bind_rows(
-  select(Dispersal.Syndrome_genus_BIEN_df, scrubbed_genus, trait_name, TraitValue),
-  select(Fruit.Length_genus_BIEN_df, scrubbed_genus, trait_name, TraitValue),
-  select(Plant.Height_genus_BIEN_df, scrubbed_genus, trait_name, TraitValue),
-  select(Fruit.Type_genus_BIEN_df, scrubbed_genus, trait_name, TraitValue),
-  select(Growth.Form_genus_BIEN_df, scrubbed_genus, trait_name, TraitValue)
-  )
+colnames(PlantLifespan_years_genus_BIEN_df_average)
+colnames(FruitType_genus_BIEN_df_average)
+colnames(GrowthForm_genus_BIEN_df_average)
+colnames(FruitLength_mm_genus_BIEN_df_average)
+colnames(SeedMass_g_genus_BIEN_df_average)
+colnames(PlantHeight_m_genus_BIEN_df_average)
+```
+
+```{r}
+genus_level_trait_data <- list(PlantLifespan_years_genus_BIEN_df_average, 
+                                FruitType_genus_BIEN_df_average, 
+                                GrowthForm_genus_BIEN_df_average, 
+                                FruitLength_mm_genus_BIEN_df_average, 
+                                SeedMass_g_genus_BIEN_df_average, 
+                                PlantHeight_m_genus_BIEN_df_average)
+
+# Function to select and convert columns to character
+select_and_convert <- function(df) {
+  df %>%
+    dplyr::select(scrubbed_genus, TraitName, TraitValue) %>%
+    mutate(across(everything(), as.character))
+}
+
+# Apply the function to each dataframe in the list and combine them
+all_genus_traits <- genus_level_trait_data %>%
+  map_df(select_and_convert) %>%
+  bind_rows()
 
@@
 all_genus_traits$TraitLevel <- "genus"
-```
-
-rename traits to match
-```{r}
-# Replace trait names to match
-long_plant_traits_tax <- long_plant_traits_tax %>%
-  mutate(TraitName = ifelse(TraitName == "Plant.Height", "Plant.Height_m", TraitName)) %>%
-  mutate(TraitName = ifelse(TraitName == "Fruit.Length", "Fruit.Length_mm", TraitName))
-unique(long_plant_traits_tax$TraitName)
-
-na_traits_family.genus <- na_traits_family.genus %>%
-  mutate(TraitName = ifelse(TraitName == "Plant.Height", "Plant.Height_m", TraitName)) %>%
-  mutate(TraitName = ifelse(TraitName == "Fruit.Length", "Fruit.Length_mm", TraitName))
-unique(na_traits_family.genus$TraitName)
 ```
@@
 colnames(all_family_traits)[which(colnames(all_family_traits) == "scrubbed_family")] <- "family"
-colnames(all_family_traits)[which(colnames(all_family_traits) == "trait_name")] <- "TraitName"
-colnames(all_family_traits)[which(colnames(all_family_traits) == "trait_value")] <- "TraitValue"
 
 colnames(all_genus_traits)[which(colnames(all_genus_traits) == "scrubbed_genus")] <- "genus"
-colnames(all_genus_traits)[which(colnames(all_genus_traits) == "trait_name")] <- "TraitName"
-colnames(all_genus_traits)[which(colnames(all_genus_traits) == "trait_value")] <- "TraitValue"
 ```
@@
   filter(is.na(TraitValue))
+nrow(na_traits_after_genus)
 filled_traits_with_genus <- filled_na_traits_family.genus %>%
   filter(!is.na(TraitValue))
+nrow(filled_traits_with_genus)
 ```
@@
   filter(!is.na(TraitValue))
+nrow(filled_traits_with_family)
 na_traits_after_family <- filled_na_traits_family %>%
   filter(is.na(TraitValue))
+nrow(na_traits_after_family)
 ```
@@
 
+```{r}
+# Count number of NA records
+na_records_count <- sum(is.na(all_traits_with_NAs$TraitValue))
+
+# Count number of unique species with NA records
+species_with_na_count <- all_traits_with_NAs %>%
+  filter(is.na(TraitValue)) %>%
+  distinct(species) %>%
+  nrow()
+
+# Count number of unique traits with NA records
+traits_with_na_count <- all_traits_with_NAs %>%
+  filter(is.na(TraitValue)) %>%
+  distinct(TraitName) %>%
+  nrow()
+
+# Print results
+cat("Number of NA records:", na_records_count, "\n")
+cat("Number of species with NA records:", species_with_na_count, "\n")
+cat("Number of traits with NA records:", traits_with_na_count, "\n")
+```
+
 ### Save long dataframe
@@
 all_traits_with_NAs_long <- all_traits_with_NAs %>%
-  select(-TraitLevel, -family, -genus)
+  select(-TraitLevel, -family, -genus) %>%
+  distinct()
 wide_traits <- pivot_wider(all_traits_with_NAs_long, names_from = TraitName, values_from = TraitValue)
 ```
-make NA if:
-- Growth.Form is numeric
-- Dispersal.Syndrome is numeric or contains multiple values
-- Plant.Height = 0
-- Fruit.Type is numeric
-```{r}
-library(stringr)
-wide_traits$Growth.Form <- ifelse(is.numeric(wide_traits$Growth.Form), NA, wide_traits$Growth.Form)
-wide_traits$Dispersal.Syndrome[stringr::str_detect(wide_traits$Dispersal.Syndrome, ",")] <- NA
-
-# Convert Plant.Height_m to numeric 
-wide_traits$Plant.Height_m[wide_traits$Plant.Height_m == "0"] <- NA
-
-wide_traits$Fruit.Type <- ifelse(is.numeric(wide_traits$Fruit.Type), NA, wide_traits$Fruit.Type)
-```
-
-```{r}
-wide_traits <- wide_traits %>%
-  mutate_all(~sapply(., paste, collapse = ", "))
-
-# Replace values with multiple values with NA
-wide_traits <- wide_traits %>%
-  mutate(Dispersal.Syndrome = if_else(str_count(Dispersal.Syndrome, ", ") > 0, NA_character_, Dispersal.Syndrome), 
-         Growth.Form = if_else(str_count(Growth.Form, ", ") > 0, NA_character_, Growth.Form),
-         Plant.Height_m = if_else(str_count(Plant.Height_m, ", ") > 0, NA_character_, Plant.Height_m), 
-         Fruit.Length_mm = if_else(str_count(Fruit.Length_mm, ", ") > 0, NA_character_, Fruit.Length_mm), 
-         Fruit.Type = if_else(str_count(Fruit.Type, ", ") > 0, NA_character_, Fruit.Type))
-```
-
-```{r}
-wide_traits$Plant.Height_m <- unlist(wide_traits$Plant.Height_m)
-wide_traits$Growth.Form <- unlist(wide_traits$Growth.Form)
-wide_traits$Dispersal.Syndrome <- unlist(wide_traits$Dispersal.Syndrome)
-wide_traits$Fruit.Length_mm <- unlist(wide_traits$Fruit.Length_mm)
-wide_traits$Fruit.Type <- unlist(wide_traits$Fruit.Type)
-
-#convert column types from list to character & numeric
-wide_traits$Plant.Height_m <- as.numeric(wide_traits$Plant.Height_m)
-wide_traits$Growth.Form <- as.character(wide_traits$Growth.Form)
-wide_traits$Dispersal.Syndrome <- as.character(wide_traits$Dispersal.Syndrome)
-wide_traits$Fruit.Length_mm <- as.numeric(wide_traits$Fruit.Length_mm)
-wide_traits$Fruit.Type <- as.character(wide_traits$Fruit.Type)
+
+```{r}
+glimpse(wide_traits)
+```
+
+```{r}
+#convert column types from character to factor & numeric
+wide_traits$PlantLifespan_years <- as.numeric(wide_traits$PlantLifespan_years)
+wide_traits$GrowthForm <- as.factor(wide_traits$GrowthForm)
+wide_traits$SeedMass_g <- as.numeric(wide_traits$SeedMass_g)
+wide_traits$PlantHeight_m <- as.numeric(wide_traits$PlantHeight_m)
+wide_traits$FruitType <- as.factor(wide_traits$FruitType)
+wide_traits$FruitLength_mm <- as.numeric(wide_traits$FruitLength_mm)
+wide_traits$SeedWidth_mm <- as.numeric(wide_traits$SeedWidth_mm)
+wide_traits$FruitMass_mg <- as.numeric(wide_traits$FruitMass_mg)
+wide_traits$SeedLength_mm <- as.numeric(wide_traits$SeedLength_mm)
+wide_traits$DispersalSyndrome <- as.factor(wide_traits$DispersalSyndrome)
 ```
@@
 
-# show the missing data pattern
-```{r}
-md.pattern(wide_traits)
-```
 # Perform the imputation
@@
 
-remove dispersal syndrome because I need to fix it 
-```{r}
-imputed_data <- imputed_data %>%
-  select(-Dispersal.Syndrome)
-```
-
-```{r}
-imputed_data %>%
-  summarise_at(vars(Growth.Form:Fruit.Type), ~ sum(is.na(.)))
+```{r}
+glimpse(imputed_data)
+```
+
+# Summary
+```{r}
+fb_plot_species_traits_completeness(wide_traits)
+ggsave("plant_trait_completeness_familygenus.png", plot = last_plot(), path = figure_path)
+fb_plot_number_species_by_trait(wide_traits)
+ggsave("plant_number_species_trait_familygenus.png", plot = last_plot(), path = figure_path)
+fb_table_trait_summary(wide_traits)
+```
+
+```{r}
+fb_plot_species_traits_completeness(imputed_data)
+ggsave("plant_trait_completeness_imputed.png", plot = last_plot(), path = figure_path)
+fb_plot_number_species_by_trait(imputed_data)
+ggsave("plant_number_species_trait_imputed.png", plot = last_plot(), path = figure_path)
+fb_table_trait_summary(imputed_data)
+```
+
+```{r}
+# Reassign NA values in TraitLevel as "Imputed"
+all_traits_with_NAs$TraitLevel[is.na(all_traits_with_NAs$TraitLevel)] <- "Imputed"
+
+# Count the number of traits per TraitLevel and TraitName
+trait_counts <- all_traits_with_NAs %>%
+  count(TraitLevel, TraitName) %>%
+  mutate(TraitLevel = fct_relevel(TraitLevel, "species", "genus", "family", "Imputed"))
+
+# Create the faceted bar plot with free y-axis scales
+ggplot(trait_counts, aes(x = TraitLevel, y = n, fill = TraitLevel)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Trait Counts per Level",
+       x = "Trait Level",
+       y = "Trait Count",
+       fill = "Level") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.position = "none") +  # Remove legend
+  facet_wrap(~ TraitName, scales = "free_y") +
+  guides(fill = "none")  # Remove legend
+ggsave("plant_trait_counts_per_level.png", plot = last_plot(), path = figure_path)
+```
+
+```{r}
+# Count the number of traits per TraitLevel
+trait_counts_overall <- all_traits_with_NAs %>%
+  count(TraitLevel) %>%
+  mutate(TraitLevel = fct_relevel(TraitLevel, "species", "genus", "family", "Imputed"))
+
+# Create the faceted bar plot with free y-axis scales
+ggplot(trait_counts_overall, aes(x = TraitLevel, y = n, fill = TraitLevel)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Trait Counts per Level",
+       x = "Trait Level",
+       y = "Trait Count",
+       fill = "Level") +
+  theme_minimal() +
+  guides(fill = "none")  # Remove legend
+ggsave("plant_trait_counts_per_level_overall.png", plot = last_plot(), path = figure_path)
 ```

### 2024-06-11 code/L1/L1_P3_TropicalAndes_plant_fruiting_subset.Rmd
add summary code

@@
 
+# Summary
+```{r}
+cat("Number of occurrence records:", nrow(plant_occ_subset), "\n")
+cat("Number of species with occurrence records:", length(unique(plant_occ_subset$Accepted_species)), "\n")
+```
+
+```{r}
+# the number of records and species per trait database
+database_summary <- traits_subset %>%
+  group_by(DatabaseSource) %>%
+  summarise(
+    num_records = n(),
+    num_species = n_distinct(Accepted_species)
+  )
+
+# Print the result
+print(database_summary)
+```
+
 # Write data to csv

### 2024-06-12 code/L1/L1_P6_TropicalAndes_plant_occ.Rmd
add summary code and visualize taxonomic coverage

@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: "This script subsets plant occurrence data to species with complete trait coverage."
-data input: "ADD"
-data output: "ADD"
+overview: "This script subsets plant occurrence data to species with complete trait coverage and summarizes the taxonomic coverage."
+data input: "TropicalAndes_imputed_plant_traits.csv, TropicalAndes_GBIF_plant_occ_harmonized_subset.csv, TropicalAndes_plant_checklist.csv"
+data output: "TropicalAndes_GBIF_plant_occ_harmonized_subset_final.csv"
 date: "2023-10-04"
@@
 ```{r}
+data_path_L0 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
 data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
 output_path_L1 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
+
 ```
@@
 ```{r}
+library(GIFT)
 library(dplyr)
-#run if not installed remotes::install_github("FRBCesab/funbiogeo")
-library(funbiogeo)
+library(sf)
+library(tidyr)
+library(ggplot2)
 ```
@@
 plant_traits <- read.csv(file.path(data_path_L1, file = "TropicalAndes_imputed_plant_traits.csv"))
-plant_occ <- read.csv(file.path(data_path_L1, file = "TropicalAndes_GBIF_plant_occ_cleaned.csv"))
-```
+plant_occ <- read.csv(file.path(data_path_L1, file = "TropicalAndes_GBIF_plant_occ_harmonized_subset.csv"))
 
-# Check trait coverage
-```{r}
-summary(plant_traits)
-```
-
-```{r}
-fb_plot_number_species_by_trait(plant_traits)
-```
-
-```{r}
-Plant_complete <- fb_filter_species_by_trait_coverage(plant_traits, threshold_traits_proportion =  1)
+# Load Tropical Andes Forest shape
+TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
 ```
@@
 ```{r}
-plant_species <- unique(Plant_complete$species)
+plant_species <- unique(plant_traits$species)
+length(plant_species)
 ```
@@
   filter(species %in% plant_species)
+nrow(plant_occ_subset)
+```
+
+# Get a checklist of Angiosperms (flowering plants) in the Tropical Andes Forest shapefile
+```{r}
+checklist <- GIFT_checklists(taxon_name = "Angiospermae", taxonomic_group = TRUE, floristic_group = "all", shp = TropicalAndes_IUCNHabitat_Forest, overlap = "extent_intersect")
+```
+## Add genus based on work_ID
+```{r}
+full <- GIFT_taxonomy()
+```
+
+```{r}
+genus_ID <- checklist$checklists$genus_ID
+genus_list <- full %>% filter(taxon_ID %in% genus_ID) %>% select(taxon_ID, taxon_name)
+names(genus_list) <- c("genus_ID", "genus")
+```
+
+```{r}
+checklist_tax <- dplyr::inner_join(checklist$checklists, genus_list)
+```
+
+```{r}
+reference_ID_list <- unique(checklist_tax$ref_ID)
+ref <- GIFT_references()
+checklist_ref <- ref[which(ref$ref_ID %in% reference_ID_list),
+           c("ref_ID", "ref_long", "geo_entity_ref")]
+checklist_ref %>%
+  mutate(ref_long = iconv(ref_long, from = "UTF-8", to = "ASCII//TRANSLIT"))
 ```
@@
 ```{r}
-source("~/GitHub/neotropical_plants/functions/data_summary.R")
+source("~/GitHub/neotropical_plants/code/data_summary.R")
 ```
@@
 ```{r}
-data_summary(plant_occ_subset, plant_occ_subset$species, plant_occ_subset$genus, plant_occ_subset$family)
+data_summary(plant_occ_subset, plant_occ_subset$Accepted_species, plant_occ_subset$genus, plant_occ_subset$family)
+```
+
+```{r}
+data_summary(checklist_tax, checklist_tax$work_species, checklist_tax$genus, checklist_tax$family)
+```
+
+# Visualize taxonomic coverage
+```{r}
+GBIF_occ_summary <- plant_occ_subset %>%
+  summarise(
+    num_species = n_distinct(Accepted_species),
+    num_families = n_distinct(family),
+    num_genera = n_distinct(genus)
+  )
+GIFT_checklist_summary <- checklist_tax %>%
+  summarise(
+    num_species = n_distinct(work_species),
+    num_families = n_distinct(family),
+    num_genera = n_distinct(genus)
+  )
+```
+
+```{r}
+# Combine the summaries into one dataframe
+combined_summary <- bind_rows(
+  GBIF_occ_summary %>% mutate(source = "Data"),
+  GIFT_checklist_summary %>% mutate(source = "Checklist")
+)
+
+# Reshape to long format for easier plotting
+combined_summary_long <- combined_summary %>%
+  pivot_longer(cols = c(num_species, num_families, num_genera), 
+               names_to = "category", 
+               values_to = "count")
+```
+
+```{r}
+ggplot(combined_summary_long, aes(x = source, y = count, fill = source)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Comparison of Species, Genera, and Families",
+       x = "Source",
+       y = "Count",
+       fill = "Source") +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),  # Angled x-axis labels
+    plot.title = element_text(hjust = 0.5),             # Center the plot title
+    plot.margin = margin(t = 10, r = 10, b = 10, l = 10)
+  ) +
+  facet_wrap(~ category, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 12, face = "bold"))
+ggsave("plant_taxonomic_coverage.png", plot = last_plot(), path = figure_path)
 ```
@@
 ```{r}
-write.csv(plant_occ_subset, file.path(output_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned_subset.csv"))
+write.csv(plant_occ_subset, file.path(output_path_L1,"TropicalAndes_GBIF_plant_occ_harmonized_subset_final.csv"))
 ```

### 2024-06-14 code/L0/L0_TropicalAndes_frugivores_frugivoria_traits.Rmd
update file paths, add dplyr package, update summary with summary function, update output paths, add citations and package info

--- 2023-07-30 code/L0/L0_TropicalAndes_frugivores_frugivoria_traits.Rmd
@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: ""
-data input: ""
-data output: ""
+overview: "This script retrived the Frugivoria database from EDI."
+#data input: "None"
+#data output: "TropicalAndes_Frugivoria_frugivore_traits.csv, TropicalAndes_Frugivoria_frugivore_traits_species.csv, TropicalAndes_frugivore_LookupTable.csv"
 date: "2023-07-18"
-output: html_document
+output:
+  html_document:
+    highlight_downlit: TRUE
+    df_print: tibble
 ---
@@
 ```{r}
-data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
-output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
-output_path2<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+data_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+output_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+output_path_L1 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+```
+
+# Load required packages
+```{r}
+library(dplyr)
 ```
@@
 ```{r}
-print("The number of records is")
-nrow(Frugivoria)
-print("The number of species is" )
-length(unique(Frugivoria$species))
-print("The number of families is" )
-length(unique(Frugivoria$family))
+source("~/GitHub/neotropical_plants/code/data_summary.R")
+```
+
+```{r}
+data_summary(Frugivoria, Frugivoria$IUCN_species_name, Frugivoria$genus, Frugivoria$family)
 ```
@@
 ```{r}
-Frugivoria_SpeciesList <- unique(Frugivoria$species)
+Frugivoria_SpeciesList <- unique(Frugivoria$IUCN_species_name)
+length(Frugivoria_SpeciesList)
 ```
@@
 ```{r}
-TropicalAndes_Frugivoria_path <- file.path(output_path,"TropicalAndes_Frugivoria_frugivore_traits.csv") 
-write.csv(Frugivoria, file = TropicalAndes_Frugivoria_path)
-Frugivoria_species_path <- file.path(output_path,"TropicalAndes_Frugivoria_frugivore_traits_species.csv") 
-write.csv(Frugivoria_SpeciesList, file = Frugivoria_species_path)
-Frugivore_LookupTable_path <- file.path(output_path2,"TropicalAndes_frugivore_LookupTable.csv") 
-write.csv(Frugivore_LookupTable, file = Frugivore_LookupTable_path)
+write.csv(Frugivoria, file = file.path(output_path_L0,"TropicalAndes_Frugivoria_frugivore_traits.csv"))
+write.csv(Frugivoria_SpeciesList, file = file.path(output_path_L0,"TropicalAndes_Frugivoria_frugivore_traits_species.csv"))
+write.csv(Frugivore_LookupTable, file = file.path(output_path_L1,"TropicalAndes_frugivore_LookupTable.csv"))
 ```
+
+# Package citations and session info
+```{r}
+library(report)
+report::cite_packages()
+
+devtools::session_info()
+```

### 2024-06-17 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
add code for parallel processing, save individual plots to file

@@
 Note: some code is adapted from https://luisdva.github.io/rstats/richness/."
-data input: "TropicalAndes_plant_occ.csv, TropicalAndes_frugivore_occ.csv"
-data output: "TropicalAndes_frugivore_richness.csv, TropicalAndes_GBIF_plant_richness.csv"
+#data input: "TropicalAndes_plant_occ.csv, TropicalAndes_frugivore_occ.csv"
+#data output: "TropicalAndes_frugivore_richness.csv, TropicalAndes_GBIF_plant_richness.csv"
 date: "2023-08-01"
-output: html_document
+output: 
+  html_document:
+    highlight_downlit = TRUE
+    df_print: tibble
 ---
@@
 library(ggplot2)
-library(rnaturalearth)
+library(parallel)
+library(foreach)
+library(doParallel)
 library(ggspatial)
@@
 ```{r}
-data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
-data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+data_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+data_path_L1 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
 output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
@@
 # Plot TD/Richness
-## Function for richness plots
+## Function for richness plots using parallel processing
 ```{r}
-create_rich_plots <- function(resolution_meters) {
- # Make Grid
-  TAGrid <- TApoly %>%
-    st_make_grid(cellsize = c(resolution_meters)) %>%
-    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
-    st_cast("MULTIPOLYGON") %>%
-    st_sf() %>%
-    mutate(cellid = row_number())
-
-  # Function to calculate richness for a given dataset
+# Define the calculate_richness function outside
 calculate_richness <- function(data_sf, species_sf) {
@@
 
+create_rich_plots <- function(resolution_meters) {
+  # Make Grid
+  TAGrid <- TApoly %>%
+    st_make_grid(cellsize = c(resolution_meters)) %>%
+    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+    st_cast("MULTIPOLYGON") %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+
   # Determine the chunk size
@@
 
-  # Split the TAGrid into chunks if necessary
+  # Split the TAGrid into chunks
   TAGrid_chunks <- split(TAGrid, ceiling(seq_along(TAGrid$cellid) / chunk_size))
 
-  # Initialize empty data frames for results
-  plant_richness_grid <- list()
-  frugivore_richness_grid <- list()
+  # Set up parallel processing
+  no_cores <- detectCores() - 1
+  cl <- makeCluster(no_cores)
+  registerDoParallel(cl)
 
-  # Iterate over chunks to calculate richness
-  for (i in seq_along(TAGrid_chunks)) {
-    chunk <- TAGrid_chunks[[i]]
+  # Export necessary objects to the cluster
+  clusterExport(cl, varlist = c("plants_sf_species", "frugivores_sf_species", "calculate_richness"))
 
-    # Debug: Print chunk info
-    print(paste("Processing chunk", i, "of", length(TAGrid_chunks)))
-
-    plant_richness_chunk <- calculate_richness(chunk, plants_sf_species)
-    frugivore_richness_chunk <- calculate_richness(chunk, frugivores_sf_species)
-
-    # Store the chunks in lists
-    plant_richness_grid[[i]] <- plant_richness_chunk
-    frugivore_richness_grid[[i]] <- frugivore_richness_chunk
+  # Process chunks in parallel
+  plant_richness_grid <- foreach(chunk = TAGrid_chunks, .combine = bind_rows, .packages = c("dplyr", "sf")) %dopar% {
+    calculate_richness(chunk, plants_sf_species)
   }
 
-  # Combine all chunks into final data frames
-  plant_richness_grid <- bind_rows(plant_richness_grid)
-  frugivore_richness_grid <- bind_rows(frugivore_richness_grid)
+  frugivore_richness_grid <- foreach(chunk = TAGrid_chunks, .combine = bind_rows, .packages = c("dplyr", "sf")) %dopar% {
+    calculate_richness(chunk, frugivores_sf_species)
+  }
+
+  # Stop the parallel cluster
+  stopCluster(cl)
 
@@
       geom_sf(aes(fill = num_species), color = NA) +
-      scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,6000)) +
-      labs( x = "Longitude", y = "Latitude", fill = "Number of species") +
+      scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,1200)) +
+      labs(x = "Longitude", y = "Latitude", fill = "S") +
       coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
@@
   }
+
   generate_frugivore_plot <- function(data) {
@@
       scale_fill_distiller(palette = "YlOrBr", direction = 1, limits = c(0,500)) +
-      labs(x = "Longitude", y = "Latitude", fill = "Number of species") +
+      labs(x = "Longitude", y = "Latitude", fill = "S") +
       coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
@@
   plant_richness_hist <- hist(plant_richness_grid$num_species,
-    main = paste("Histogram of Plant Richness [", resolution_meters/1000, " km] by cell", sep = ""),
+                              main = paste("Histogram of Plant Richness [", resolution_meters / 1000, " km] by cell", sep = ""),
                               xlab = "Plant Richness by cell")
   frugivore_richness_hist <- hist(frugivore_richness_grid$num_species,
-    main = paste("Histogram of Frugivore Richness [", resolution_meters/1000, " km] by cell", sep = ""),
+                                  main = paste("Histogram of Frugivore Richness [", resolution_meters / 1000, " km] by cell", sep = ""),
                                   xlab = "Frugivore Richness by cell")
+
   # Return a list of data
@@
 }
-
 ```
@@
 richness_5km <- create_rich_plots(5000)
+```
+
+```{r}
 print(richness_5km)
@@
 richness_10km <- create_rich_plots(10000)
+```
+
+```{r}
 print(richness_10km)
@@
 richness_25km <- create_rich_plots(25000)
+```
+
+```{r}
 print(richness_25km)
@@
 richness_50km <- create_rich_plots(50000)
+```
+
+```{r}
 print(richness_50km)
@@
 richness_75km <- create_rich_plots(75000)
+```
+
+```{r}
 print(richness_75km)
@@
 richness_100km <- create_rich_plots(100000)
+```
+
+```{r}
 print(richness_100km)
@@
 
-# Multi panel plots
-## Richness maps
+# Save richness objects to file
 ```{r}
-all_plant_richness_plots <- ggpubr::ggarrange(# plantgridRichTA_5km, 
-                                              plantgridRichTA_10km, plantgridRichTA_25km,
-                                              plantgridRichTA_50km, plantgridRichTA_75km,
-                                              plantgridRichTA_100km,
-                                              ncol = 5, nrow = 1,
-                                              labels = c("10 km", "25 km", "50 km", "75 km", "100 km"),
-                                              common.legend = TRUE, legend = "right")
-all_plant_richness_plots
-all_plant_richness_plots_labeled <- ggpubr::annotate_figure(all_plant_richness_plots,
-                                                            left = ggpubr::text_grob("Plants", face = "bold", size = 14))
-all_plant_richness_plots_labeled
+saveRDS(richness_5km, file = file.path(output_path,"richness_5km.rds"))
+saveRDS(richness_10km, file = file.path(output_path,"richness_10km.rds"))
+saveRDS(richness_25km, file = file.path(output_path,"richness_25km.rds"))
+saveRDS(richness_50km, file = file.path(output_path,"richness_50km.rds"))
+saveRDS(richness_75km, file = file.path(output_path,"richness_75km.rds"))
+saveRDS(richness_100km, file = file.path(output_path,"richness_100km.rds"))
 ```
 
+# Save plots to file
+```{r}
+plot_list <- list(plantgridRichTA_5km = plantgridRichTA_5km, plantgridRichTA_10km = plantgridRichTA_10km,
+                  plantgridRichTA_25km = plantgridRichTA_25km, plantgridRichTA_50km = plantgridRichTA_50km,
+                  plantgridRichTA_75km = plantgridRichTA_75km, plantgridRichTA_100km = plantgridRichTA_100km,
+                  frugivoregridRichTA_5km = frugivoregridRichTA_5km, frugivoregridRichTA_10km = frugivoregridRichTA_10km,
+                  frugivoregridRichTA_25km = frugivoregridRichTA_25km, frugivoregridRichTA_50km = frugivoregridRichTA_50km,
+                  frugivoregridRichTA_75km = frugivoregridRichTA_75km, frugivoregridRichTA_100km = frugivoregridRichTA_100km)
+#histogram_list <- list(plant_richness_hist_5km = richness_5km$plant_richness_hist, 
+#                       frugivore_richness_hist_5km = richness_5km$frugivore_richness_hist,
+#                       plant_richness_hist_10km = richness_10km$plant_richness_hist,
+#                       frugivore_richness_hist_10km = richness_10km$frugivore_richness_hist,
+#                       plant_richness_hist_25km = richness_25km$plant_richness_hist,
+#                       frugivore_richness_hist_55km = richness_25km$frugivore_richness_hist,
+#                       plant_richness_hist_50km = richness_50km$plant_richness_hist,
+#                       frugivore_richness_hist_50km = richness_50km$frugivore_richness_hist,
+#                       plant_richness_hist_75km = richness_75km$plant_richness_hist,
+#                       frugivore_richness_hist_75km = richness_75km$frugivore_richness_hist,
+#                       plant_richness_hist_100km = richness_100km$plant_richness_hist,
+#                       frugivore_richness_hist_100km = richness_100km$frugivore_richness_hist)
+#histograms would have to be converted to ggplot to save
+```
+
+```{r}
+save_plots <- function(plot_list, figure_path) {
+  # Iterate over the list and save each plot
+  for (plot_name in names(plot_list)) {
+    plot_path <- file.path(figure_path, paste0(plot_name, ".png"))
+
+    # Try to save the plot and handle any potential errors
+    tryCatch({
+      ggsave(filename = plot_path, 
+             plot = plot_list[[plot_name]])
+      cat("Saved plot:", plot_path, "\n")
+    }, error = function(e) {
+      cat("Failed to save plot:", plot_name, "\nError message:", e$message, "\n")
+    })
+  }
+}
+```
+
+```{r}
+# Save the plots
+save_plots(plot_list, figure_path)
+```
+
+# Multi panel plots of richness maps
+## Plants
+```{r}
+# Add labels and remove axis labels from all plots
+plantgridRichTA_5km_labeled <- plantgridRichTA_5km + ggtitle("5 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
+plantgridRichTA_10km_labeled <- plantgridRichTA_10km + ggtitle("10 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
+plantgridRichTA_25km_labeled <- plantgridRichTA_25km + ggtitle("25 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
+plantgridRichTA_50km_labeled <- plantgridRichTA_50km + ggtitle("50 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
+plantgridRichTA_75km_labeled <- plantgridRichTA_75km + ggtitle("75 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
+plantgridRichTA_100km_labeled <- plantgridRichTA_100km + ggtitle("100 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
+plantgridRichTA_100km_labeled
+```
+
+```{r}
+# Arrange the plots with common legend
+all_plant_richness_plots <- ggarrange(plantgridRichTA_5km_labeled, plantgridRichTA_10km_labeled,
+                                      plantgridRichTA_25km_labeled, plantgridRichTA_50km_labeled,
+                                      plantgridRichTA_75km_labeled, plantgridRichTA_100km_labeled,
+                                      ncol = 6, nrow = 1,
+                                      common.legend = TRUE, legend = "left")
+
+all_plant_richness_plots
+```
+
+```{r}
+# Add "Plants" label above the legend
+all_plant_richness_plots_labeled <- all_plant_richness_plots + theme(legend.title = element_text("Plants\n S"))
+all_plant_richness_plots_labeled
+ggsave("plant_richness_plots.png", all_plant_richness_plots_labeled, path = figure_path, width = 18, height = 8, units = "in")
+```
+
+## Frugivores
 ```{r}
@@
                                               common.legend = TRUE, legend = "right")
+```
+
+```{r}
 all_frugivore_richness_plots_labeled <- ggpubr::annotate_figure(all_frugivore_richness_plots,
@@
 all_frugivore_richness_plots_labeled
+ggsave("frugivore_richness_plots.png", all_frugivore_richness_plots_labeled, path = figure_path, width = 11, height = 8, units = "in")
 ```
 
+## Combine plants & frugivores
 ```{r}
@@
 # Write data to csv
-
 ```{r}

### 2024-06-18 code/L1/L1_LastStep_DiversityInputObjects.Rmd
add figure save code, update input and outputs

@@
 ---
-title: ""
+title: "Creating Diversity Input Objects"
 author: "Hazel J. Anderson"
@@
 overview: "Projecting occurrence data and creating presence-absence matrix for plants and frugivores. "
-data input: 
-data output: 
+#data input: ""
+#data output: ""
 date: "2024-05-13"
-output: html_document
+output: 
+  html_document:
+    highlight_downlit = TRUE
+    df_print: tibble
 ---
@@
 library(tidyr)
+library(ggspatial)
 library(ggplot2)
@@
 output_path_L1 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
 ```
@@
 ```{r}
-TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned_subset.csv"))
+TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_plant_occ_harmonized_subset_final.csv"))
 TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_frugivore_occ_cleaned_subset.csv"))
@@
 basePlot
+ggsave("tropical_andes_forest_map.png", plot = last_plot(), path = figure_path)
 ```
@@
 plantsPointsPlot
+ggsave("plant_occurrence_points_map.png", plot = last_plot(), path = figure_path)
 ```
@@
   theme(panel.background = element_rect(fill = "lightblue"))
+frugivoresPointsPlot
+ggsave("frugivore_occurrence_points_map.png", plot = last_plot(), path = figure_path)
+```
 
-frugivoresPointsPlot
+```{r}
+both_points_maps <- ggpubr::ggarrange(plantsPointsPlot,
+                                        frugivoresPointsPlot,
+                                        ncol = 2, nrow = 1)
+both_points_maps
+ggsave("both_points_maps.png", both_points_maps, path = figure_path)
 ```
@@
 # Create Presence-Absense matrix
-
 ```{r}
@@
 # remove duplicate species name column
-plant_traits_df_subset <- plant_traits_df_subset[, c("Growth.Form", "Plant.Height_m", "Fruit.Length_mm", "Fruit.Type")]
+plant_traits_df_subset$species <- NULL
 frugivore_traits_df_subset <- frugivore_traits_df_subset[, c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")]

### 2024-06-18 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
add labels and remove axis labels from all plots

@@
 ```{r}
-# Add "Plants" label above the legend
-all_plant_richness_plots_labeled <- all_plant_richness_plots + theme(legend.title = element_text("Plants\n S"))
+all_plant_richness_plots_labeled <- ggpubr::annotate_figure(all_plant_richness_plots,
+                                                            left = ggpubr::text_grob("Plants", face = "bold", size = 14, 
+                                                                                     rot = 90))
 all_plant_richness_plots_labeled
-ggsave("plant_richness_plots.png", all_plant_richness_plots_labeled, path = figure_path, width = 18, height = 8, units = "in")
+ggsave("plant_richness_plots.png", all_plant_richness_plots_labeled, path = figure_path, width = 20, height = 5, units = "in")
 ```
@@
 ```{r}
-all_frugivore_richness_plots <- ggpubr::ggarrange(#frugivoregridRichTA_5km, 
-                                              frugivoregridRichTA_10km, frugivoregridRichTA_25km,
-                                              frugivoregridRichTA_50km, frugivoregridRichTA_75km,
-                                              frugivoregridRichTA_100km,
-                                              ncol = 5, nrow = 1,
-                                              labels = c("10 km", "25 km", "50 km", "75 km", "100 km"),
-                                              common.legend = TRUE, legend = "right")
+# Add labels and remove axis labels from all plots
+frugivoregridRichTA_5km_labeled <- frugivoregridRichTA_5km + ggtitle("5 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
+frugivoregridRichTA_10km_labeled <- frugivoregridRichTA_10km + ggtitle("10 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
+frugivoregridRichTA_25km_labeled <- frugivoregridRichTA_25km + ggtitle("25 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
+frugivoregridRichTA_50km_labeled <- frugivoregridRichTA_50km + ggtitle("50 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
+frugivoregridRichTA_75km_labeled <- frugivoregridRichTA_75km + ggtitle("75 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
+frugivoregridRichTA_100km_labeled <- frugivoregridRichTA_100km + ggtitle("100 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
+frugivoregridRichTA_100km_labeled
+```
+
+```{r}
+all_frugivore_richness_plots <- ggpubr::ggarrange(frugivoregridRichTA_5km_labeled,frugivoregridRichTA_10km_labeled,
+                                                  frugivoregridRichTA_25km_labeled, frugivoregridRichTA_50km_labeled,
+                                                  frugivoregridRichTA_75km_labeled, frugivoregridRichTA_100km_labeled,
+                                                  ncol = 6, nrow = 1,
+                                                  common.legend = TRUE, legend = "left")
+all_frugivore_richness_plots
 ```
@@
 all_frugivore_richness_plots_labeled <- ggpubr::annotate_figure(all_frugivore_richness_plots,
-                                                            left = ggpubr::text_grob("Frugivores", face = "bold", size = 14))
+                                                            left = ggpubr::text_grob("Frugivores", face = "bold", size = 14,
+                                                                                     rot = 90))
 all_frugivore_richness_plots_labeled
-ggsave("frugivore_richness_plots.png", all_frugivore_richness_plots_labeled, path = figure_path, width = 11, height = 8, units = "in")
+ggsave("frugivore_richness_plots.png", all_frugivore_richness_plots_labeled, path = figure_path, width = 20, height = 4, units = "in")
 ```
@@
                                         ncol = 1, nrow = 2)
+# Add a white background to the plot
+all_richness_plots <- all_richness_plots + theme(plot.background = element_rect(fill = "white", color = NA))
 all_richness_plots
-ggsave("all_richness_plots.png", all_richness_plots, path = figure_path, width = 11, height = 8, units = "in")
+
+ggsave("all_richness_plots.png", all_richness_plots, path = figure_path, width = 20, height = 8, units = "in")
 ```

### 2024-06-19 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
add new input files for all spatial grains, clean up code to reduce number of lines for 100 km fdis calculations for plants and frugivores

@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: ""
-data input: ""
-data output: ""
+overview: "This script calculates functional diversity as functional dispersion for plants and frugivores."
 date: "2023-08-03"
@@
 library(rlang)
+library(doParallel)
+library(foreach)
 ```
@@
 
-# Frugivores
-## Create trait type table
+```{r}
+site_loc_key_plant_75km <- readRDS(file = file.path(data_path_L1,"site_loc_key_plant_75km.rds"))
+site_loc_key_frugivore_75km <- readRDS(file = file.path(data_path_L1,"site_loc_key_frugivore_75km.rds"))
+PAM_plant_site_final_75km <- readRDS(file = file.path(data_path_L1,"PAM_plant_site_final_75km.rds"))
+PAM_frugivore_site_final_75km <- readRDS(file = file.path(data_path_L1,"PAM_frugivore_site_final_75km.rds"))
+```
+
+```{r}
+site_loc_key_plant_50km <- readRDS(file = file.path(data_path_L1,"site_loc_key_plant_50km.rds"))
+site_loc_key_frugivore_50km <- readRDS(file = file.path(data_path_L1,"site_loc_key_frugivore_50km.rds"))
+PAM_plant_site_final_50km <- readRDS(file = file.path(data_path_L1,"PAM_plant_site_final_50km.rds"))
+PAM_frugivore_site_final_50km <- readRDS(file = file.path(data_path_L1,"PAM_frugivore_site_final_50km.rds"))
+```
+
+```{r}
+site_loc_key_plant_25km <- readRDS(file = file.path(data_path_L1,"site_loc_key_plant_25km.rds"))
+site_loc_key_frugivore_25km <- readRDS(file = file.path(data_path_L1,"site_loc_key_frugivore_25km.rds"))
+PAM_plant_site_final_25km <- readRDS(file = file.path(data_path_L1,"PAM_plant_site_final_25km.rds"))
+PAM_frugivore_site_final_25km <- readRDS(file = file.path(data_path_L1,"PAM_frugivore_site_final_25km.rds"))
+```
+
+```{r}
+site_loc_key_plant_10km <- readRDS(file = file.path(data_path_L1,"site_loc_key_plant_10km.rds"))
+site_loc_key_frugivore_10km <- readRDS(file = file.path(data_path_L1,"site_loc_key_frugivore_10km.rds"))
+PAM_plant_site_final_10km <- readRDS(file = file.path(data_path_L1,"PAM_plant_site_final_10km.rds"))
+PAM_frugivore_site_final_10km <- readRDS(file = file.path(data_path_L1,"PAM_frugivore_site_final_10km.rds"))
+```
+
+```{r}
+site_loc_key_plant_5km <- readRDS(file = file.path(data_path_L1,"site_loc_key_plant_5km.rds"))
+site_loc_key_frugivore_5km <- readRDS(file = file.path(data_path_L1,"site_loc_key_frugivore_5km.rds"))
+PAM_plant_site_final_5km <- readRDS(file = file.path(data_path_L1,"PAM_plant_site_final_5km.rds"))
+PAM_frugivore_site_final_5km <- readRDS(file = file.path(data_path_L1,"PAM_frugivore_site_final_5km.rds"))
+```
+
+# Create trait type tables
 ```{r}
@@
 
-## Summary of the assemblages * species dataframe
 ```{r}
+trait_name <- c("PlantLifespan_years","GrowthForm", "SeedMass_g", "PlantHeight_m", "FruitType", "FruitLength_mm",
+                "SeedWidth_mm", "FruitMass_mg", "SeedLength_mm", "DispersalSyndrome")
+trait_type <- c("Q", "N", "Q", "Q", "N", "Q", "Q", "Q", "Q", "N") 
+plant_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
+```
+
+```{r}
+# fix Nominal traits as factor
+frugivore_traits_df_final$diet_cat <- as.factor(frugivore_traits_df_final$diet_cat)
+
+plant_traits_df_final$GrowthForm <- as.factor(plant_traits_df_final$GrowthForm)
+plant_traits_df_final$FruitType <- as.factor(plant_traits_df_final$FruitType)
+plant_traits_df_final$DispersalSyndrome <- as.factor(plant_traits_df_final$DispersalSyndrome)
+```
+
+# Create function
+```{r}
+
+```
+
+# Frugivores
+## 100 km
+```{r}
+# Summary of the assemblages * species dataframe
 asb_sp_frugivore_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_frugivore_site_final_100km)
@@
 
-## Species traits summary
 ```{r}
-frugivore_traits_summ <- mFD::sp.tr.summary(
-  tr_cat     = frug_trait_cat,   
-  sp_tr      = frugivore_traits_df_final, 
-  stop_if_NA = TRUE)
-```
-
-## Estimate functional trait-based distances between species
-```{r}
-sp_dist_frugivore <- mFD::funct.dist(
-  sp_tr         = frugivore_traits_df_final,
-  tr_cat        = frug_trait_cat,
-  metric        = "gower",
-  scale_euclid  = "scale_center",
-  ordinal_var   = "classic",
-  weight_type   = "equal",
-  stop_if_NA    = TRUE)
-```
-
-## Generate a multidimensional space
-```{r}
-fspaces_quality_frugivore <- mFD::quality.fspaces(
-  sp_dist             = sp_dist_frugivore,
-  maxdim_pcoa         = 10,
-  deviation_weighting = "absolute",
-  fdist_scaling       = FALSE,
-  fdendro             = "average")
+# Species traits summary
+frugivore_traits_summ <- mFD::sp.tr.summary(tr_cat = frug_trait_cat, sp_tr = frugivore_traits_df_final)
 ```
@@
 ```{r}
-# Look at the quality spaces only (MAD index looks at mean absolute deviation from the dissimilary matrix; want the deviation to be low, meaning that the true distances have been retained in the PCA)
+# Estimate functional trait-based distances between species
+sp_dist_frugivore <- mFD::funct.dist( sp_tr = frugivore_traits_df_final, tr_cat = frug_trait_cat, metric = "gower",
+  scale_euclid = "scale_center", ordinal_var = "classic", weight_type = "equal", stop_if_NA = TRUE)
+```
+
+```{r}
+# Generate a multidimensional space
+fspaces_quality_frugivore <- mFD::quality.fspaces(sp_dist = sp_dist_frugivore, maxdim_pcoa = 10, deviation_weighting = "absolute",
+                                                  fdist_scaling = FALSE, fdendro = "average")
+```
+
+```{r}
+# Look at the quality spaces only (MAD index looks at the mean absolute deviation from the dissimilarity matrix; want the deviation to be low meaning that the true distances have been retained in the PCA)
 round(fspaces_quality_frugivore$"quality_fspaces", 3)
-# Plot the quality spaces (chose to look at 3D, 4D, and 5D since they had the lowest MAD). Will go with 4 dimensions.
-mFD::quality.fspaces.plot(
-  fspaces_quality            = fspaces_quality_frugivore,
-  quality_metric             = "mad",
-  fspaces_plot               = c("pcoa_3d", "pcoa_4d", 
-                                 "pcoa_5d"),
-  name_file                  = NULL,
-  range_dist                 = NULL,
-  range_dev                  = NULL,
-  range_qdev                 = NULL,
-  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
-  gradient_deviation_quality = c(low = "yellow", high = "red"),
-  x_lab                      = "Trait-based distance")
+```
+
+```{r}
+# Plot the quality spaces (chose to look at 3D, 4D, and 5D since they had the lowest MAD). Will go with 3 dimensions
+mFD::quality.fspaces.plot(fspaces_quality = fspaces_quality_frugivore, quality_metric = "mad",
+                          fspaces_plot = c("pcoa_3d", "pcoa_4d", "pcoa_5d"))
 ```
@@
 ```{r}
-# View the components of the PCA axes *Remember the first few components explain the most variation in dissimilarity. Clusters into groups
-
 # Computes linear model for continuous traits and Kruskall-Wallis tests for other types. 
-frugivore_tr_faxes <- mFD::traits.faxes.cor(
-  sp_tr          = frugivore_traits_df_final, 
+frugivore_tr_faxes <- mFD::traits.faxes.cor(sp_tr = frugivore_traits_df_final, 
                                             sp_faxes_coord = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")],
@@
 
-big_plot_frugivore <- mFD::funct.space.plot(
-  sp_faxes_coord  = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")],
-  faxes           = c("PC1", "PC2", "PC3", "PC4"),
-  name_file       = NULL,
-  faxes_nm        = NULL,
-  range_faxes     = c(NA, NA),
-  color_bg        = "grey95",
-  color_pool      = "darkgreen",
-  fill_pool       = "white",
-  shape_pool      = 21,
-  size_pool       = 1,
-  plot_ch         = TRUE,
-  color_ch        = "black",
-  fill_ch         = "white",
-  alpha_ch        = 0.5,
-  plot_vertices   = TRUE,
-  color_vert      = "blueviolet",
-  fill_vert       = "blueviolet",
-  shape_vert      = 6,
-  size_vert       = 1,
-  plot_sp_nm      = NULL,
-  nm_size         = 3,
-  nm_color        = "black",
-  nm_fontface     = "plain",
-  check_input     = TRUE)
-```
-
-```{r}
+big_plot_frugivore <- mFD::funct.space.plot(sp_faxes_coord = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")],
+                                            faxes = c("PC1", "PC2", "PC3", "PC4"), alpha_ch = 0.5, shape_vert = 6)
 big_plot_frugivore
@@
 ```{r}
-#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
+#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (3)
 # Calculate row sums
 row_sums_frugivore <- rowSums(PAM_frugivore_site_final_100km)
-subset_matrix_frugivore <- PAM_frugivore_site_final_100km[row_sums_frugivore >= 4, ]
+subset_matrix_frugivore <- PAM_frugivore_site_final_100km[row_sums_frugivore >= 3, ]
 ```
@@
 ```{r}
-#computing FD
+# Computing FD
 # The number of species per assemblage has to be higher or equal to the number of traits
 
-alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(
-  sp_faxes_coord   = sp_faxes_coord_frugivore_sub,
-  asb_sp_w         = subset_matrix_frugivore,
-  ind_vect         = "fdis",
-  #did not run fdiv or fric (slows it down); c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe")
-  scaling          = TRUE,
-  check_input      = TRUE,
-  details_returned = TRUE)
+alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(sp_faxes_coord = sp_faxes_coord_frugivore_sub, asb_sp_w = subset_matrix_frugivore,
+                                                     ind_vect = "fdis", details_returned = TRUE)
 
@@
 
-## Mapping FD
+### Mapping FD
 
@@
 # Generate coordinates
-subset_coords_frugivore <-site_loc_key_frugivore_100km[rowSums(PAM_frugivore_site_final_100km) >= 4, ]
+subset_coords_frugivore <-site_loc_key_frugivore_100km[rowSums(PAM_frugivore_site_final_100km) >= 3, ]
 subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]
@@
     geom_sf(data = spatial_fdis_frugivore_grid, aes(fill = fdis_value), color = NA) +
-    labs(x = "Longitude", y = "Latitude", fill = "FDis") +
+    labs(fill = "FDis") +
+    ggtitle("100 km") +
     scale_fill_distiller(palette = "YlOrBr", direction = 1) +
@@
     theme(panel.background = element_rect(fill = "lightblue"),
-        text = element_text(size = 12))
+        text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
 frugivoregridFDisTA_100km
@@
 # Plants
-## Create trait type table
+
+## 100 km
 ```{r}
-trait_name <- c("Growth.Form", "Plant.Height_m", "Fruit.Length_mm", "Fruit.Type")
-trait_type <- c("N", "Q", "Q", "N") 
-plant_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
-```
-
-## Summary of the assemblages * species dataframe
-```{r}
-asb_sp_plant_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_plant_site_final)
-```
-
-## Species traits summary
-```{r}
-plant_traits_summ <- mFD::sp.tr.summary(
-  tr_cat     = plant_trait_cat,   
-  sp_tr      = plant_traits_df_final, 
-  stop_if_NA = TRUE)
+# Summary of the assemblages * species dataframe
+asb_sp_plant_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_plant_site_final_100km)
 ```
@@
 ```{r}
-save(plant_traits_df_final, plant_trait_cat, file = "../../plant_funct.dist_files.rdata")
-```
-
-## Estimate functional trait-based distances between species
-```{r}
-sp_dist_plants <- mFD::funct.dist(
-  sp_tr         = plant_traits_df_final,
-  tr_cat        = plant_trait_cat,
-  metric        = "gower",
-  scale_euclid  = "scale_center",
-  ordinal_var   = "classic",
-  weight_type   = "equal",
-  stop_if_NA    = TRUE)
-```
-
-## Generate a multidimensional space
-```{r}
-fspaces_quality_plants <- mFD::quality.fspaces(
-  sp_dist             = sp_dist_plants,
-  maxdim_pcoa         = 10,
-  deviation_weighting = "absolute",
-  fdist_scaling       = FALSE,
-  fdendro             = "average")
+# Species traits summary
+plant_traits_summ <- mFD::sp.tr.summary(tr_cat = plant_trait_cat, sp_tr = plant_traits_df_final)
 ```
@@
 ```{r}
-# Look at the quality spaces only (MAD index looks at mean absolute deviation from the dissimilary matrix; want the deviation to be low, meaning that the true distances have been retained in the PCA)
-round(fspaces_quality_plants$"quality_fspaces", 3)            
-# Plot the quality spaces (chose to look at 3D, 4D, and 5D since they had the lowest MAD). Will go with 4 dimensions.
-mFD::quality.fspaces.plot(
-  fspaces_quality            = fspaces_quality_plants,
-  quality_metric             = "mad",
-  fspaces_plot               = c("pcoa_2d", "pcoa_3d", 
-                                 "pcoa_4d"),
-  name_file                  = NULL,
-  range_dist                 = NULL,
-  range_dev                  = NULL,
-  range_qdev                 = NULL,
-  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
-  gradient_deviation_quality = c(low = "yellow", high = "red"),
-  x_lab                      = "Trait-based distance")
+# Estimate functional trait-based distances between species
+sp_dist_plant <- mFD::funct.dist( sp_tr = plant_traits_df_final, tr_cat = plant_trait_cat, metric = "gower",
+  scale_euclid = "scale_center", ordinal_var = "classic", weight_type = "equal", stop_if_NA = TRUE)
+```
+
+```{r}
+# Generate a multidimensional space
+fspaces_quality_plant <- mFD::quality.fspaces(sp_dist = sp_dist_plant, maxdim_pcoa = 10, deviation_weighting = "absolute",
+                                                  fdist_scaling = FALSE, fdendro = "average")
+```
+
+```{r}
+# Look at the quality spaces only (MAD index looks at the mean absolute deviation from the dissimilarity matrix; want the deviation to be low meaning that the true distances have been retained in the PCA)
+round(fspaces_quality_plant$"quality_fspaces", 3)
+```
+
+```{r}
+# Plot the quality spaces (chose to look at )
+mFD::quality.fspaces.plot(fspaces_quality = fspaces_quality_plant, quality_metric = "mad",
+                          fspaces_plot = c("pcoa_6d", "pcoa_7d", "pcoa_8d"))
 ```
@@
 #testing correlation between functional axes and traits
-sp_faxes_coord_plants <- fspaces_quality_plants$"details_fspaces"$"sp_pc_coord"
+sp_faxes_coord_plant <- fspaces_quality_plant$"details_fspaces"$"sp_pc_coord"
 ```
@@
 ```{r}
-# View the components of the PCA axes *Remember the first few components explain the most variation in dissimilarity. Clusters into groups
-
 # Computes linear model for continuous traits and Kruskall-Wallis tests for other types. 
-plant_tr_faxes <- mFD::traits.faxes.cor(
-  sp_tr          = plant_traits_df_final, 
-  sp_faxes_coord = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")], 
+plant_tr_faxes <- mFD::traits.faxes.cor(sp_tr = plant_traits_df_final, 
+                                            sp_faxes_coord = sp_faxes_coord_plant[ , c("PC1", "PC2", "PC3", "PC4")],
                                             plot = TRUE)
@@
 #plotting functional space
-sp_faxes_coord_plants <- fspaces_quality_plants$"details_fspaces"$"sp_pc_coord"
+sp_faxes_coord_plant <- fspaces_quality_plant$"details_fspaces"$"sp_pc_coord"
 
-big_plot_plants <- mFD::funct.space.plot(
-  sp_faxes_coord  = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")],
-  faxes           = c("PC1", "PC2", "PC3", "PC4"),
-  name_file       = NULL,
-  faxes_nm        = NULL,
-  range_faxes     = c(NA, NA),
-  color_bg        = "grey95",
-  color_pool      = "darkgreen",
-  fill_pool       = "white",
-  shape_pool      = 21,
-  size_pool       = 1,
-  plot_ch         = TRUE,
-  color_ch        = "black",
-  fill_ch         = "white",
-  alpha_ch        = 0.5,
-  plot_vertices   = TRUE,
-  color_vert      = "blueviolet",
-  fill_vert       = "blueviolet",
-  shape_vert      = 6,
-  size_vert       = 1,
-  plot_sp_nm      = NULL,
-  nm_size         = 3,
-  nm_color        = "black",
-  nm_fontface     = "plain",
-  check_input     = TRUE)
-
-#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
-# Calculate row sums
-row_sums_plant <- rowSums(PAM_plant_site_final)
-subset_matrix_plant <- PAM_plant_site_final[row_sums_plant >= 4, ]
+big_plot_plant <- mFD::funct.space.plot(sp_faxes_coord = sp_faxes_coord_plant[ , c("PC1", "PC2", "PC3", "PC4")],
+                                            faxes = c("PC1", "PC2", "PC3", "PC4"), alpha_ch = 0.5, shape_vert = 6)
+big_plot_plant
 ```
@@
 ```{r}
-big_plot_plants
+#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
+# Calculate row sums
+row_sums_plant <- rowSums(PAM_plant_site_final_100km)
+subset_matrix_plant <- PAM_plant_site_final_100km[row_sums_plant >= 4, ]
 ```
@@
 ```{r}
-# getting sp_faxes_coord_plants and subset_matrix_plants names to match
-sp_faxes_coord_plants_sub <- sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")]
-## check number of species names
-nrow(sp_faxes_coord_plants_sub) # 54
-ncol(subset_matrix_plant) #82
+# match plant names
+sp_faxes_coord_plant_sub <- sp_faxes_coord_plant[ , c("PC1", "PC2", "PC3", "PC4")]
+summary(sp_faxes_coord_plant_sub)
 ```
@@
 ```{r}
-sp_faxes_coord_plants_sub_names <- row.names(sp_faxes_coord_plants_sub)
+## check number of species names
+nrow(sp_faxes_coord_plant_sub)
+ncol(subset_matrix_plant) 
+```
+
+```{r}
+sp_faxes_coord_plant_sub_names <- row.names(sp_faxes_coord_plant_sub)
+subset_matrix_plant_names <- colnames(subset_matrix_plant)
+```
+
+```{r}
+sp_faxes_coord_plant_sub <- as.data.frame(sp_faxes_coord_plant_sub)
+row.names(sp_faxes_coord_plant_sub) <- sp_faxes_coord_plant_sub_names
+
+sp_faxes_coord_plant_sub_names <- row.names(sp_faxes_coord_plant_sub)
 subset_matrix_plant_names <- colnames(subset_matrix_plant)
 
-sp_faxes_coord_plants_sub_df <- as.data.frame(sp_faxes_coord_plants_sub)
-subset_matrix_plant_df <- as.data.frame(subset_matrix_plant)
+plant_names <- intersect(sp_faxes_coord_plant_sub_names, subset_matrix_plant_names)
+plant_names <- na.omit(plant_names)
 
-subset_matrix_plant <- subset_matrix_plant_df[, which((names(subset_matrix_plant_df) %in% sp_faxes_coord_plants_sub_names)==TRUE)]
+sp_faxes_coord_plant_sub <- sp_faxes_coord_plant_sub[ which((row.names(sp_faxes_coord_plant_sub) %in% plant_names)==TRUE), ]
 
+subset_matrix_plant  <- as.data.frame(subset_matrix_plant)
+subset_matrix_plant <- subset_matrix_plant[ ,which((colnames(subset_matrix_plant) %in% plant_names)==TRUE)]
+
+#remove NAs
+sp_faxes_coord_plant_sub <- na.omit(sp_faxes_coord_plant_sub)
+subset_matrix_plant <- na.omit(subset_matrix_plant)
+```
+
+```{r}
+nrow(sp_faxes_coord_plant_sub)
 ncol(subset_matrix_plant)
+```
+
+```{r}
+sp_faxes_coord_plant_sub <- as.matrix(sp_faxes_coord_plant_sub)
 subset_matrix_plant <- as.matrix(subset_matrix_plant)
@@
 ```{r}
-#computing FD
+# Computing FD
 # The number of species per assemblage has to be higher or equal to the number of traits
-alpha_fd_indices_plant <- mFD::alpha.fd.multidim(
-  sp_faxes_coord   = sp_faxes_coord_plants_sub,
-  asb_sp_w         = PAM_plant_site_final,
-  ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
-  scaling          = TRUE,
-  check_input      = TRUE,
-  details_returned = TRUE)
+
+alpha_fd_indices_plant <- mFD::alpha.fd.multidim(sp_faxes_coord = sp_faxes_coord_plant_sub, asb_sp_w = subset_matrix_plant,
+                                                     ind_vect = "fdis", details_returned = TRUE)
 
 details_list_plant <- alpha_fd_indices_plant$"details"
-
-#plot
-plots_alpha_plant <- mFD::alpha.multidim.plot(
-  output_alpha_fd_multidim = alpha_fd_indices_plant,
-  plot_asb_nm              = c("cell4", "cell20"),
-  ind_nm                   = c("fdis", "fide", "fnnd"
-                        , "fori", "fspe"),
-  faxes                    = NULL,
-  faxes_nm                 = NULL,
-  range_faxes              = c(NA, NA),
-  color_bg                 = "grey95",
-  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
-  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
-  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  color_ch          
-  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
-  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
-  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
-  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
-  shape_centroid_fspe      = 23,
-  color_centroid_fspe      = "black",
-  size_sp_nm               = 3, 
-  color_sp_nm              = "black",
-  plot_sp_nm               = NULL,
-  fontface_sp_nm           = "plain",
-  save_file                = FALSE,
-  check_input              = TRUE)
-
-# comparing convex hulls of different hulls
-plots_alpha_plant$"fric"$"patchwork"
-
-plots_alpha_plant$"fdiv"$"patchwork"
-
-plots_alpha_plant$"fspe"$"patchwork"
-
-plots_alpha_plant$"fdis"$"patchwork"
-
-plots_alpha_plant$"fori"$"patchwork"
-
 ```
@@
 ```{r}
-#Get functional dispersion
+# Get functional dispersion
 fdis_plant <- alpha_fd_indices_plant$functional_diversity_indices$fdis
@@
 
+## Mapping FD
+
+```{r}
+# Generate coordinates
+subset_coords_plant <-site_loc_key_plant_100km[rowSums(PAM_plant_site_final_100km) >= 4, ]
+subset_coords_plant_sp <-subset_coords_plant[,1:2 ]
+
+plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
+```
+
+```{r}
+# Convert the dataframe to sf format
+spatial_fdis_plant <- st_as_sf(plant_fd_sp, coords = c("Longitude", "Latitude"))
+```
+
+```{r}
+# set crs of sf objects
+spatial_fdis_plant <- spatial_fdis_plant %>% st_set_crs(5389)
+```
+
+```{r}
+# Extract coordinates 
+spatial_fdis_plant_coords <- st_coordinates(spatial_fdis_plant)
+
+# Add coordinates as separate columns
+spatial_fdis_plant <- spatial_fdis_plant %>%
+  mutate(x = spatial_fdis_plant_coords[, 1], y = spatial_fdis_plant_coords[, 2])
+```
+
+```{r}
+summary(spatial_fdis_plant)
+```
+
+```{r}
+TAGrid_100km <- TApoly %>%
+  st_make_grid(cellsize = c(100000)) %>%
+  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+  st_cast("MULTIPOLYGON") %>%
+  st_sf() %>%
+  mutate(cellid = row_number())
+
+spatial_fdis_plant$cellid <- st_intersects(spatial_fdis_plant, TAGrid_100km)$nn
+
+spatial_fdis_plant_grid <- TAGrid_100km %>%
+   st_join(spatial_fdis_plant) %>%
+   group_by(cellid) %>%
+   summarize(fdis_value = sum(fdis_plant, na.rm = TRUE)) %>%
+   ungroup()
+```
+
+```{r}
+summary(spatial_fdis_plant_grid)
+```
+
+```{r}
+plantgridFDisTA_100km <-
+    ggplot() +
+    geom_sf(data = Americas, fill = "white")+
+    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
+    geom_sf(data = spatial_fdis_plant_grid, aes(fill = fdis_value), color = NA) +
+    labs(fill = "FDis") +
+    ggtitle("100 km") +
+    scale_fill_distiller(palette = "Greens", direction = 1) +
+    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+    theme(panel.background = element_rect(fill = "lightblue"),
+        text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
+plantgridFDisTA_100km
+```
+
+```{r}
 # Extract cell values
+plant_cellFDis_100km <- spatial_fdis_plant_grid
+hist(plant_cellFDis_100km$fdis_value)
+```
+
+## 75 km
+
 ```{r}
-plantgridFDTA <- plantgridFDTA[["data"]]
-frugivoregridFDTA <- frugivoregridFDTA[["data"]]
+#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
+# Calculate row sums
+row_sums_plant <- rowSums(PAM_plant_site_final_75km)
+subset_matrix_plant <- PAM_plant_site_final_75km[row_sums_plant >= 4, ]
 ```
+
+```{r}
+# match plant names
+sp_faxes_coord_plant_sub <- sp_faxes_coord_plant[ , c("PC1", "PC2", "PC3", "PC4")]
+summary(sp_faxes_coord_plant_sub)
+```
+
+```{r}
+## check number of species names
+nrow(sp_faxes_coord_plant_sub)
+ncol(subset_matrix_plant) 
+```
+
+```{r}
+sp_faxes_coord_plant_sub_names <- row.names(sp_faxes_coord_plant_sub)
+subset_matrix_plant_names <- colnames(subset_matrix_plant)
+```
+
+```{r}
+sp_faxes_coord_plant_sub <- as.data.frame(sp_faxes_coord_plant_sub)
+row.names(sp_faxes_coord_plant_sub) <- sp_faxes_coord_plant_sub_names
+
+sp_faxes_coord_plant_sub_names <- row.names(sp_faxes_coord_plant_sub)
+subset_matrix_plant_names <- colnames(subset_matrix_plant)
+
+plant_names <- intersect(sp_faxes_coord_plant_sub_names, subset_matrix_plant_names)
+plant_names <- na.omit(plant_names)
+
+sp_faxes_coord_plant_sub <- sp_faxes_coord_plant_sub[ which((row.names(sp_faxes_coord_plant_sub) %in% plant_names)==TRUE), ]
+
+subset_matrix_plant  <- as.data.frame(subset_matrix_plant)
+subset_matrix_plant <- subset_matrix_plant[ ,which((colnames(subset_matrix_plant) %in% plant_names)==TRUE)]
+
+#remove NAs
+sp_faxes_coord_plant_sub <- na.omit(sp_faxes_coord_plant_sub)
+subset_matrix_plant <- na.omit(subset_matrix_plant)
+```
+
+```{r}
+nrow(sp_faxes_coord_plant_sub)
+ncol(subset_matrix_plant)
+```
+
+```{r}
+sp_faxes_coord_plant_sub <- as.matrix(sp_faxes_coord_plant_sub)
+subset_matrix_plant <- as.matrix(subset_matrix_plant)
+```
+
+```{r}
+# Computing FD
+# The number of species per assemblage has to be higher or equal to the number of traits
+
+alpha_fd_indices_plant <- mFD::alpha.fd.multidim(sp_faxes_coord = sp_faxes_coord_plant_sub, asb_sp_w = subset_matrix_plant,
+                                                     ind_vect = "fdis", details_returned = TRUE)
+
+details_list_plant <- alpha_fd_indices_plant$"details"
+```
+
+```{r}
+# Get functional dispersion
+fdis_plant <- alpha_fd_indices_plant$functional_diversity_indices$fdis
+```
+
+## Mapping FD
+
+```{r}
+# Generate coordinates
+subset_coords_plant <-site_loc_key_plant_75km[rowSums(PAM_plant_site_final_75km) >= 4, ]
+subset_coords_plant_sp <-subset_coords_plant[,1:2 ]
+
+plant_fd_sp_75km <- data.frame(subset_coords_plant_sp, fdis_plant)
+```
+
+```{r}
+# Convert the dataframe to sf format
+spatial_fdis_plant_75km <- st_as_sf(plant_fd_sp_75km, coords = c("Longitude", "Latitude"))
+```
+
+```{r}
+# set crs of sf objects
+spatial_fdis_plant_75km <- spatial_fdis_plant_75km %>% st_set_crs(5389)
+```
+
+```{r}
+# Extract coordinates 
+spatial_fdis_plant_coords_75km <- st_coordinates(spatial_fdis_plant_75km)
+
+# Add coordinates as separate columns
+spatial_fdis_plant_75km <- spatial_fdis_plant_75km %>%
+  mutate(x = spatial_fdis_plant_coords_75km[, 1], y = spatial_fdis_plant_coords_75km[, 2])
+```
+
+```{r}
+summary(spatial_fdis_plant_75km)
+```
+
+```{r}
+TAGrid__75km <- TApoly %>%
+  st_make_grid(cellsize = c(75000)) %>%
+  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+  st_cast("MULTIPOLYGON") %>%
+  st_sf() %>%
+  mutate(cellid = row_number())
+
+spatial_fdis_plant_75km$cellid <- st_intersects(spatial_fdis_plant_75km, TAGrid_75km)$nn
+
+spatial_fdis_plant_grid_75km <- TAGrid_75km %>%
+   st_join(spatial_fdis_plant_75km) %>%
+   group_by(cellid) %>%
+   summarize(fdis_value = sum(fdis_plant, na.rm = TRUE)) %>%
+   ungroup()
+```
+
+```{r}
+summary(spatial_fdis_plant_grid_75km)
+```
+
+```{r}
+plantgridFDisTA_75km <-
+    ggplot() +
+    geom_sf(data = Americas, fill = "white")+
+    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
+    geom_sf(data = spatial_fdis_plant_grid_75km, aes(fill = fdis_value), color = NA) +
+    labs(fill = "FDis") +
+    ggtitle("100 km") +
+    scale_fill_distiller(palette = "Greens", direction = 1) +
+    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+    theme(panel.background = element_rect(fill = "lightblue"),
+        text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
+plantgridFDisTA_75km
+```

### 2024-06-24 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
create functions to calculate and map FDis, make multi-panel plots for maps

@@
 library(foreach)
+library(purrr)
+library(ggpubr)
 ```
@@
 ```{r}
-data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
+data_path_L1 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
 output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
@@
 plant_traits_df_final$DispersalSyndrome <- as.factor(plant_traits_df_final$DispersalSyndrome)
-```
-
-# Create function
-```{r}
-
 ```
@@
 ```{r}
+ggsave("frugivoregridFDisTA_100km.png", frugivoregridFDisTA_100km, path = figure_path)
 # Extract cell values
@@
 hist(frugivore_cellFDis_100km$fdis_value)
+```
+
+## Create function to calculate FD as fdis for frugivores for other spatial grains
+Uses inputs from trait objects created above to avoid redundancies 
+```{r}
+calculate_fd_as_fdis_frugivores <- function(PAM_site_final) {
+  # Calculate row sums and subset matrix early to avoid unnecessary calculations
+  row_sums_frugivore <- rowSums(PAM_site_final)
+  subset_matrix_frugivore <- PAM_site_final[row_sums_frugivore >= 4, ]
+
+  # Select the necessary columns from sp_faxes_coord_frugivore
+  sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore[, c("PC1", "PC2", "PC3", "PC4")]
+
+  ## check number of species names
+  print(nrow(sp_faxes_coord_frugivore_sub))
+  print(ncol(subset_matrix_frugivore))
+
+  sp_faxes_coord_frugivore_sub_names <- row.names(sp_faxes_coord_frugivore_sub)
+  subset_matrix_frugivore_names <- colnames(subset_matrix_frugivore)
+
+  sp_faxes_coord_frugivore_sub <- as.data.frame(sp_faxes_coord_frugivore_sub)
+  row.names(sp_faxes_coord_frugivore_sub) <- sp_faxes_coord_frugivore_sub_names
+
+  sp_faxes_coord_frugivore_sub_names <- row.names(sp_faxes_coord_frugivore_sub)
+  subset_matrix_frugivore_names <- colnames(subset_matrix_frugivore)
+
+  frugivore_names <- intersect(sp_faxes_coord_frugivore_sub_names, subset_matrix_frugivore_names)
+  frugivore_names <- na.omit(frugivore_names)
+
+  sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore_sub[ which((row.names(sp_faxes_coord_frugivore_sub) %in%
+                                                                         frugivore_names)==TRUE), ]
+
+  subset_matrix_frugivore  <- as.data.frame(subset_matrix_frugivore)
+  subset_matrix_frugivore <- subset_matrix_frugivore[ ,which((colnames(subset_matrix_frugivore) %in% frugivore_names)==TRUE)]
+
+  #remove NAs
+  sp_faxes_coord_frugivore_sub <- na.omit(sp_faxes_coord_frugivore_sub)
+  subset_matrix_frugivore <- na.omit(subset_matrix_frugivore)
+
+  print(nrow(sp_faxes_coord_frugivore_sub))
+  print(ncol(subset_matrix_frugivore))
+
+  if (nrow(sp_faxes_coord_frugivore_sub) != ncol(subset_matrix_frugivore)) {
+    stop("Number of rows in sp_faxes_coord_frugivore_sub does not match number of columns in subset_matrix_frugivore")
+  }
+
+  sp_faxes_coord_frugivore_sub <- as.matrix(sp_faxes_coord_frugivore_sub)
+  subset_matrix_frugivore <- as.matrix(subset_matrix_frugivore)
+
+  # Use three fewer cores than available
+  num_cores <- parallel::detectCores() - 3
+
+  # Set up parallel backend
+  cl <- makeCluster(num_cores)
+  registerDoParallel(cl)
+
+  tryCatch({
+    # Define chunk size
+    chunk_size <- ceiling(nrow(subset_matrix_frugivore) / num_cores)
+
+    # Split data into chunks for parallel processing
+    chunks <- split(seq_len(nrow(subset_matrix_frugivore)), ceiling(seq_len(nrow(subset_matrix_frugivore)) / chunk_size))
+
+    # Process chunks in parallel
+    results_list <- foreach(chunk_indices = chunks, .combine = rbind, .packages = "mFD") %dopar% {
+      subset_matrix_frugivore_chunk <- subset_matrix_frugivore[chunk_indices, , drop = FALSE]
+
+      alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(
+        sp_faxes_coord = sp_faxes_coord_frugivore_sub, 
+        asb_sp_w = subset_matrix_frugivore_chunk, 
+        ind_vect = "fdis", 
+        details_returned = TRUE
+      )
+     # Extract fdis values and match with corresponding cell numbers
+      fdis_values <- alpha_fd_indices_frugivore$functional_diversity_indices$fdis
+      cell_numbers <- rownames(subset_matrix_frugivore)[chunk_indices]
+
+      # Create a data frame for this chunk
+      chunk_results <- data.frame(Cell_Number = cell_numbers, fdis = fdis_values)
+      chunk_results
+    }
+
+    # Combine all chunk results into a single data frame
+    all_results <- do.call(rbind, results_list)
+
+    # Stop the parallel backend
+    stopCluster(cl)
+
+    return(results_list)
+  }, error = function(e) {
+    stop("Error in parallel processing: ", conditionMessage(e))
+  })
+}
+
+```
+
+```{r}
+fdis_frugivore_75km <- calculate_fd_as_fdis_frugivores(PAM_frugivore_site_final_75km)
+```
+
+```{r}
+fdis_frugivore_50km <- calculate_fd_as_fdis_frugivores(PAM_frugivore_site_final_50km)
+```
+
+```{r}
+fdis_frugivore_25km <- calculate_fd_as_fdis_frugivores(PAM_frugivore_site_final_25km)
+```
+
+```{r}
+fdis_frugivore_10km <- calculate_fd_as_fdis_frugivores(PAM_frugivore_site_final_10km)
+```
+
+```{r}
+fdis_frugivore_5km <- calculate_fd_as_fdis_frugivores(PAM_frugivore_site_final_5km)
+```
+
+## Map FD for spatial grains other than 100 km
+```{r}
+mapping_fd_as_fdis_frugivores <- function(PAM_frugivore_site, site_loc_key_frugivore, fdis_frugivore, resolution_meters) {
+  # Generate coordinates
+  subset_coords_frugivore <- site_loc_key_frugivore[rowSums(PAM_frugivore_site) >= 4, ]
+  subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]
+
+  frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)
+
+  # Convert the dataframe to sf format
+  spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude", "Latitude"))
+
+  # set crs of sf objects
+  spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(5389)
+
+  # Extract coordinates 
+  spatial_fdis_frugivore_coords <- st_coordinates(spatial_fdis_frugivore)
+
+  # Add coordinates as separate columns
+  spatial_fdis_frugivore <- spatial_fdis_frugivore %>%
+    mutate(x = spatial_fdis_frugivore_coords[, 1], y = spatial_fdis_frugivore_coords[, 2])
+
+  TAGrid <- TApoly %>%
+    st_make_grid(cellsize = c(resolution_meters)) %>%
+    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+    st_cast("MULTIPOLYGON") %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+
+  spatial_fdis_frugivore$cellid <- st_intersects(spatial_fdis_frugivore, TAGrid)$nn
+
+  spatial_fdis_frugivore_grid <- TAGrid %>%
+   st_join(spatial_fdis_frugivore) %>%
+   group_by(cellid) %>%
+   summarize(fdis_value = sum(fdis_frugivore, na.rm = TRUE)) %>%
+   ungroup()
+
+  frugivoregridFDisTA <-
+    ggplot() +
+    geom_sf(data = Americas, fill = "white")+
+    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
+    geom_sf(data = spatial_fdis_frugivore_grid, aes(fill = fdis_value), color = NA) +
+    labs(fill = "FDis") +
+    scale_fill_distiller(palette = "YlOrBr", direction = 1) +
+    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+    theme(panel.background = element_rect(fill = "lightblue"),
+        text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
+  frugivoregridFDisTA
+
+  list(frugivoregridFDisTA = frugivoregridFDisTA,
+         spatial_fdis_frugivore_grid = spatial_fdis_frugivore_grid)
+}
+```
+
+```{r}
+mapping_75km <- mapping_fd_as_fdis_frugivores(PAM_frugivore_site_final_75km, site_loc_key_frugivore_75km, fdis_frugivore_75km$fdis, 75000)
+frugivoregridFDisTA_75km <- mapping_75km$frugivoregridFDisTA
+frugivore_cellFDis_75km <- mapping_75km$spatial_fdis_frugivore_grid
+```
+
+```{r}
+# add title to maps
+(frugivoregridFDisTA_75km <- frugivoregridFDisTA_75km +
+    ggtitle("75 km")
+)
+ggsave("frugivoregridFDisTA_75km.png", frugivoregridFDisTA_75km, path = figure_path)
+```
+
+```{r}
+mapping_50km <- mapping_fd_as_fdis_frugivores(PAM_frugivore_site_final_50km, site_loc_key_frugivore_50km, fdis_frugivore_50km$fdis, 50000)
+frugivoregridFDisTA_50km <- mapping_50km$frugivoregridFDisTA
+frugivore_cellFDis_50km <- mapping_50km$spatial_fdis_frugivore_grid
+```
+
+```{r}
+# add title to maps
+(frugivoregridFDisTA_50km <- frugivoregridFDisTA_50km +
+    ggtitle("50 km")
+)
+ggsave("frugivoregridFDisTA_50km.png", frugivoregridFDisTA_50km, path = figure_path)
+```
+
+```{r}
+mapping_25km <- mapping_fd_as_fdis_frugivores(PAM_frugivore_site_final_25km, site_loc_key_frugivore_25km, fdis_frugivore_25km$fdis, 25000)
+frugivoregridFDisTA_25km <- mapping_25km$frugivoregridFDisTA
+frugivore_cellFDis_25km <- mapping_25km$spatial_fdis_frugivore_grid
+```
+
+```{r}
+# add title to maps
+(frugivoregridFDisTA_25km <- frugivoregridFDisTA_25km +
+    ggtitle("25 km")
+)
+ggsave("frugivoregridFDisTA_25km.png", frugivoregridFDisTA_25km, path = figure_path)
+```
+
+```{r}
+mapping_10km <- mapping_fd_as_fdis_frugivores(PAM_frugivore_site_final_10km, site_loc_key_frugivore_10km, fdis_frugivore_10km$fdis, 10000)
+frugivoregridFDisTA_10km <- mapping_10km$frugivoregridFDisTA
+frugivore_cellFDis_10km <- mapping_10km$spatial_fdis_frugivore_grid
+```
+
+```{r}
+# add title to maps
+(frugivoregridFDisTA_10km <- frugivoregridFDisTA_10km +
+    ggtitle("10 km")
+)
+ggsave("frugivoregridFDisTA_10km.png", frugivoregridFDisTA_10km, path = figure_path)
+```
+
+```{r}
+# Generate coordinates
+subset_coords_frugivore_5km <-site_loc_key_frugivore_5km[rowSums(PAM_frugivore_site_final_5km) >= 4, ]
+subset_coords_frugivore_sp_5km <-subset_coords_frugivore_5km[,1:2 ]
+
+frugivore_fd_sp_5km <- data.frame(subset_coords_frugivore_sp_5km, fdis_frugivore_5km)
+```
+
+```{r}
+# Convert the dataframe to sf format
+spatial_fdis_frugivore_5km <- st_as_sf(frugivore_fd_sp_5km, coords = c("Longitude", "Latitude"))
+```
+
+```{r}
+# set crs of sf objects
+spatial_fdis_frugivore_5km <- spatial_fdis_frugivore_5km %>% st_set_crs(5389)
+```
+
+```{r}
+# Extract coordinates 
+spatial_fdis_frugivore_coords_5km <- st_coordinates(spatial_fdis_frugivore_5km)
+
+# Add coordinates as separate columns
+spatial_fdis_frugivore_5km <- spatial_fdis_frugivore_5km %>%
+  mutate(x = spatial_fdis_frugivore_coords_5km[, 1], y = spatial_fdis_frugivore_coords_5km[, 2])
+```
+
+```{r}
+TAGrid_5km <- TApoly %>%
+  st_make_grid(cellsize = c(5000)) %>%
+  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+  st_cast("MULTIPOLYGON") %>%
+  st_sf() %>%
+  mutate(cellid = row_number())
+
+spatial_fdis_frugivore_5km$cellid <- st_intersects(spatial_fdis_frugivore_5km, TAGrid_5km)$nn
+
+```
+
+```{r}
+# Determine number of cores to use
+num_cores <- detectCores() - 2
+
+# Register parallel backend
+cl <- makeCluster(num_cores)
+registerDoParallel(cl)
+
+# Ensure the cluster stops when done
+on.exit(stopCluster(cl))
+
+# Split TAGrid_5km indices into chunks for parallel processing
+chunk_indices <- split(seq_len(nrow(TAGrid_5km)), rep(1:num_cores, each = ceiling(nrow(TAGrid_5km) / num_cores), length.out = nrow(TAGrid_5km)))
+
+# Perform spatial join and aggregation in parallel
+results_list <- foreach(indices = chunk_indices, .combine = 'c', .packages = c("sf", "dplyr")) %dopar% {
+  grid_chunk <- TAGrid_5km[indices, ]
+
+  # Perform spatial join
+  joined <- st_join(grid_chunk, spatial_fdis_frugivore_5km)
+
+  # Group by cellid and summarize fdis values
+  summarized <- joined %>%
+    group_by(cellid) %>%
+    summarize(fdis_value = sum(fdis, na.rm = TRUE)) %>%
+    ungroup()
+
+  # Ensure that summarized is a data frame and add debugging information
+  summarized_df <- as.data.frame(summarized)
+
+  # Debugging: Check the structure of the summarized data frame
+  message("Processed chunk: ", indices[1], " to ", indices[length(indices)])
+  str(summarized_df)
+
+  list(summarized_df)  # Return as list to avoid unintended simplification
+}
+
+# Flatten the results_list and ensure each element is a data frame
+flattened_results <- map_dfr(results_list, ~ as.data.frame(.x))
+
+# Check the structure of the combined result
+str(flattened_results)
+
+# Convert to sf object if applicable
+if ("geometry" %in% colnames(flattened_results)) {
+  spatial_fdis_frugivore_grid <- st_as_sf(flattened_results)
+} else {
+  spatial_fdis_frugivore_grid <- flattened_results
+}
+
+# Check the structure of the final output
+str(spatial_fdis_frugivore_grid)
+
+```
+
+```{r}
+frugivoregridFDisTA_5km <-
+    ggplot() +
+    geom_sf(data = Americas, fill = "white")+
+    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
+    geom_sf(data = spatial_fdis_frugivore_grid, aes(fill = fdis_value), color = NA) +
+    labs(fill = "FDis") +
+    scale_fill_distiller(palette = "YlOrBr", direction = 1) +
+    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+    theme(panel.background = element_rect(fill = "lightblue"),
+        text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
+  frugivoregridFDisTA_5km
+frugivore_cellFDis_5km <- spatial_fdis_frugivore_grid
+```
+
+```{r}
+# add title to maps
+(frugivoregridFDisTA_5km <- frugivoregridFDisTA_5km +
+    ggtitle("5 km")
+)
+ggsave("frugivoregridFDisTA_5km.png", frugivoregridFDisTA_5km, path = figure_path)
 ```
@@
     ggtitle("100 km") +
-    scale_fill_distiller(palette = "Greens", direction = 1) +
+    scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,7.5)) +
     coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
@@
 ```{r}
+ggsave("plantgridFDisTA_100km.png", plantgridFDisTA_100km, path = figure_path)
 # Extract cell values
@@
 
-## 75 km
+## Create function to calculate FD as fdis for plants for other spatial grains
+Uses inputs from trait objects created above to avoid redundancies 
+```{r}
+calculate_fd_as_fdis_plants <- function(PAM_site_final) {
+  # Calculate row sums and subset matrix early to avoid unnecessary calculations
+  row_sums_plant <- rowSums(PAM_site_final)
+  subset_matrix_plant <- PAM_site_final[row_sums_plant >= 4, ]
 
-```{r}
-#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
-# Calculate row sums
-row_sums_plant <- rowSums(PAM_plant_site_final_75km)
-subset_matrix_plant <- PAM_plant_site_final_75km[row_sums_plant >= 4, ]
+  # Select the necessary columns from sp_faxes_coord_plant
+  sp_faxes_coord_plant_sub <- sp_faxes_coord_plant[, c("PC1", "PC2", "PC3", "PC4")]
+
+  if (nrow(sp_faxes_coord_plant_sub) != ncol(subset_matrix_plant)) {
+    stop("Number of rows in sp_faxes_coord_plant_sub does not match number of columns in subset_matrix_plant")
+  }
+
+  sp_faxes_coord_plant_sub <- as.matrix(sp_faxes_coord_plant_sub)
+  subset_matrix_plant <- as.matrix(subset_matrix_plant)
+
+  # Use three fewer cores than available
+  num_cores <- parallel::detectCores() - 3
+
+  # Set up parallel backend
+  cl <- makeCluster(num_cores)
+  registerDoParallel(cl)
+
+  tryCatch({
+    # Define chunk size
+    chunk_size <- ceiling(nrow(subset_matrix_plant) / num_cores)
+
+    # Split data into chunks for parallel processing
+    chunks <- split(seq_len(nrow(subset_matrix_plant)), ceiling(seq_len(nrow(subset_matrix_plant)) / chunk_size))
+
+    # Process chunks in parallel
+    results_list <- foreach(chunk_indices = chunks, .combine = rbind, .packages = "mFD") %dopar% {
+      subset_matrix_plant_chunk <- subset_matrix_plant[chunk_indices, , drop = FALSE]
+
+      alpha_fd_indices_plant <- mFD::alpha.fd.multidim(
+        sp_faxes_coord = sp_faxes_coord_plant_sub, 
+        asb_sp_w = subset_matrix_plant_chunk, 
+        ind_vect = "fdis", 
+        details_returned = TRUE
+      )
+     # Extract fdis values and match with corresponding cell numbers
+      fdis_values <- alpha_fd_indices_plant$functional_diversity_indices$fdis
+      cell_numbers <- rownames(subset_matrix_plant)[chunk_indices]
+
+      # Create a data frame for this chunk
+      chunk_results <- data.frame(Cell_Number = cell_numbers, fdis = fdis_values)
+      chunk_results
+    }
+
+    # Combine all chunk results into a single data frame
+    all_results <- do.call(rbind, results_list)
+
+    # Stop the parallel backend
+    stopCluster(cl)
+
+    return(results_list)
+  }, error = function(e) {
+    stop("Error in parallel processing: ", conditionMessage(e))
+  })
+}
+
 ```
@@
 ```{r}
-# match plant names
-sp_faxes_coord_plant_sub <- sp_faxes_coord_plant[ , c("PC1", "PC2", "PC3", "PC4")]
-summary(sp_faxes_coord_plant_sub)
+fdis_plant_75km <- calculate_fd_as_fdis_plants(PAM_site_final = PAM_plant_site_final_75km)
 ```
@@
 ```{r}
-## check number of species names
-nrow(sp_faxes_coord_plant_sub)
-ncol(subset_matrix_plant) 
+fdis_plant_50km <- calculate_fd_as_fdis_plants(PAM_site_final = PAM_plant_site_final_50km)
 ```
@@
 ```{r}
-sp_faxes_coord_plant_sub_names <- row.names(sp_faxes_coord_plant_sub)
-subset_matrix_plant_names <- colnames(subset_matrix_plant)
+fdis_plant_25km <- calculate_fd_as_fdis_plants(PAM_site_final = PAM_plant_site_final_25km)
 ```
@@
 ```{r}
-sp_faxes_coord_plant_sub <- as.data.frame(sp_faxes_coord_plant_sub)
-row.names(sp_faxes_coord_plant_sub) <- sp_faxes_coord_plant_sub_names
-
-sp_faxes_coord_plant_sub_names <- row.names(sp_faxes_coord_plant_sub)
-subset_matrix_plant_names <- colnames(subset_matrix_plant)
-
-plant_names <- intersect(sp_faxes_coord_plant_sub_names, subset_matrix_plant_names)
-plant_names <- na.omit(plant_names)
-
-sp_faxes_coord_plant_sub <- sp_faxes_coord_plant_sub[ which((row.names(sp_faxes_coord_plant_sub) %in% plant_names)==TRUE), ]
-
-subset_matrix_plant  <- as.data.frame(subset_matrix_plant)
-subset_matrix_plant <- subset_matrix_plant[ ,which((colnames(subset_matrix_plant) %in% plant_names)==TRUE)]
-
-#remove NAs
-sp_faxes_coord_plant_sub <- na.omit(sp_faxes_coord_plant_sub)
-subset_matrix_plant <- na.omit(subset_matrix_plant)
+fdis_plant_10km <- calculate_fd_as_fdis_plants(PAM_site_final = PAM_plant_site_final_10km)
 ```
@@
 ```{r}
-nrow(sp_faxes_coord_plant_sub)
-ncol(subset_matrix_plant)
+fdis_plant_5km <- calculate_fd_as_fdis_plants(PAM_site_final = PAM_plant_site_final_5km)
 ```
 
+## Map FD for spatial grains other than 100 km
 ```{r}
-sp_faxes_coord_plant_sub <- as.matrix(sp_faxes_coord_plant_sub)
-subset_matrix_plant <- as.matrix(subset_matrix_plant)
-```
+mapping_fd_as_fdis_plants <- function(PAM_plant_site, site_loc_key_plant, fdis_plant, resolution_meters) {
+  # Generate coordinates
+  subset_coords_plant <- site_loc_key_plant[rowSums(PAM_plant_site) >= 4, ]
+  subset_coords_plant_sp <- subset_coords_plant[, 1:2]
 
-```{r}
-# Computing FD
-# The number of species per assemblage has to be higher or equal to the number of traits
+  plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
 
-alpha_fd_indices_plant <- mFD::alpha.fd.multidim(sp_faxes_coord = sp_faxes_coord_plant_sub, asb_sp_w = subset_matrix_plant,
-                                                     ind_vect = "fdis", details_returned = TRUE)
+  # Convert the dataframe to sf format
+  spatial_fdis_plant <- st_as_sf(plant_fd_sp, coords = c("Longitude", "Latitude"))
 
-details_list_plant <- alpha_fd_indices_plant$"details"
-```
+  # set crs of sf objects
+  spatial_fdis_plant <- spatial_fdis_plant %>% st_set_crs(5389)
 
-```{r}
-# Get functional dispersion
-fdis_plant <- alpha_fd_indices_plant$functional_diversity_indices$fdis
-```
-
-## Mapping FD
-
-```{r}
-# Generate coordinates
-subset_coords_plant <-site_loc_key_plant_75km[rowSums(PAM_plant_site_final_75km) >= 4, ]
-subset_coords_plant_sp <-subset_coords_plant[,1:2 ]
-
-plant_fd_sp_75km <- data.frame(subset_coords_plant_sp, fdis_plant)
-```
-
-```{r}
-# Convert the dataframe to sf format
-spatial_fdis_plant_75km <- st_as_sf(plant_fd_sp_75km, coords = c("Longitude", "Latitude"))
-```
-
-```{r}
-# set crs of sf objects
-spatial_fdis_plant_75km <- spatial_fdis_plant_75km %>% st_set_crs(5389)
-```
-
-```{r}
   # Extract coordinates 
-spatial_fdis_plant_coords_75km <- st_coordinates(spatial_fdis_plant_75km)
+  spatial_fdis_plant_coords <- st_coordinates(spatial_fdis_plant)
 
   # Add coordinates as separate columns
-spatial_fdis_plant_75km <- spatial_fdis_plant_75km %>%
-  mutate(x = spatial_fdis_plant_coords_75km[, 1], y = spatial_fdis_plant_coords_75km[, 2])
-```
+  spatial_fdis_plant <- spatial_fdis_plant %>%
+    mutate(x = spatial_fdis_plant_coords[, 1], y = spatial_fdis_plant_coords[, 2])
 
-```{r}
-summary(spatial_fdis_plant_75km)
-```
+  # Use three fewer cores than available
+  num_cores <- detectCores() - 3
+  cl <- makeCluster(num_cores)
+  registerDoParallel(cl)
 
-```{r}
-TAGrid__75km <- TApoly %>%
-  st_make_grid(cellsize = c(75000)) %>%
+  on.exit(stopCluster(cl)) # Ensure the cluster is stopped when the function exits
+
+  tryCatch({
+    # Generate the grid
+    TAGrid <- TApoly %>%
+      st_make_grid(cellsize = resolution_meters) %>%
       st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
@@
 
-spatial_fdis_plant_75km$cellid <- st_intersects(spatial_fdis_plant_75km, TAGrid_75km)$nn
+    # Check if TAGrid is empty
+    if (nrow(TAGrid) == 0) {
+      stop("Error: No valid grid cells after intersection.")
+    }
 
-spatial_fdis_plant_grid_75km <- TAGrid_75km %>%
-   st_join(spatial_fdis_plant_75km) %>%
+    # Divide the indices of the grid into chunks for parallel processing
+    chunk_indices <- split(seq_len(nrow(TAGrid)), rep(1:num_cores, each = ceiling(nrow(TAGrid) / num_cores), length.out = nrow(TAGrid)))
+
+    # Process chunks in parallel
+    results_list <- foreach(indices = chunk_indices, .combine = rbind, .packages = c("sf", "dplyr")) %dopar% {
+      chunk <- TAGrid[indices, ]
+
+      # Perform the intersection
+      intersected <- st_intersection(spatial_fdis_plant, chunk)
+
+      # Calculate cellid for intersected
+      intersected$cellid <- st_intersects(intersected, chunk)$nn
+
+      # Ensure that intersected is a data frame
+      as.data.frame(intersected)
+    }
+
+    # Combine all chunk results into a single data frame
+    spatial_fdis_plant <- do.call(rbind, results_list)
+
+    # Ensure the combined result is a data frame
+    spatial_fdis_plant <- as.data.frame(spatial_fdis_plant)
+
+    # Debugging: Check if 'cellid' column exists
+    if (!"cellid" %in% colnames(spatial_fdis_plant)) {
+      stop("Error: 'cellid' column not found in spatial_fdis_plant.")
+    }
+
+    # Summarize fdis values per grid cell
+    spatial_fdis_plant_grid <- spatial_fdis_plant %>%
+      group_by(cellid) %>%
+      summarize(fdis_value = sum(fdis_plant, na.rm = TRUE)) %>%
+      right_join(as.data.frame(TAGrid), by = "cellid")
+
+    # Convert the grid to sf
+    spatial_fdis_plant_grid <- st_as_sf(spatial_fdis_plant_grid)
+
+    # Create the plot
+    plantgridFDisTA <- ggplot() +
+      geom_sf(data = Americas, fill = "white") +
+      geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
+      geom_sf(data = spatial_fdis_plant_grid, aes(fill = fdis_value), color = NA) +
+      labs(fill = "FDis") +
+      scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0, 7.5)) +
+      coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+      theme(panel.background = element_rect(fill = "lightblue"),
+            text = element_text(size = 12),
+            plot.title = element_text(hjust = 0.5),
+            axis.text.x = element_text(angle = 45, hjust = 1))
+
+    list(plantgridFDisTA = plantgridFDisTA, spatial_fdis_plant_grid = spatial_fdis_plant_grid)
+  }, error = function(e) {
+    message("Error in parallel processing: ", conditionMessage(e))
+    NULL
+  })
+}
+```
+
+```{r}
+mapping_75km <- mapping_fd_as_fdis_plants(PAM_plant_site_final_75km, site_loc_key_plant_75km, fdis_plant_75km$fdis, 75000)
+plantgridFDisTA_75km <- mapping_75km$plantgridFDisTA
+plant_cellFDis_75km <- mapping_75km$spatial_fdis_plant_grid
+```
+
+```{r}
+# add title to maps
+(plantgridFDisTA_75km <- plantgridFDisTA_75km +
+    ggtitle("75 km")
+)
+ggsave("plantgridFDisTA_75km.png", plantgridFDisTA_75km, path = figure_path)
+```
+
+```{r}
+mapping_50km <- mapping_fd_as_fdis_plants(PAM_plant_site_final_50km, site_loc_key_plant_50km, fdis_plant_50km$fdis, 50000)
+plantgridFDisTA_50km <- mapping_50km$plantgridFDisTA
+plant_cellFDis_50km <- mapping_50km$spatial_fdis_plant_grid
+```
+
+```{r}
+# add title to maps
+(plantgridFDisTA_50km <- plantgridFDisTA_50km +
+    ggtitle("50 km")
+)
+ggsave("plantgridFDisTA_50km.png", plantgridFDisTA_50km, path = figure_path)
+```
+
+```{r}
+mapping_25km <- mapping_fd_as_fdis_plants(PAM_plant_site_final_25km, site_loc_key_plant_25km, fdis_plant_25km$fdis, 25000)
+plantgridFDisTA_25km <- mapping_25km$plantgridFDisTA
+plant_cellFDis_25km <- mapping_25km$spatial_fdis_plant_grid
+```
+
+```{r}
+# add title to maps
+(plantgridFDisTA_25km <- plantgridFDisTA_25km +
+    ggtitle("25 km")
+)
+ggsave("plantgridFDisTA_25km.png", plantgridFDisTA_25km, path = figure_path)
+```
+
+```{r}
+mapping_10km <- mapping_fd_as_fdis_plants(PAM_plant_site_final_10km, site_loc_key_plant_10km, fdis_plant_10km$fdis, 10000)
+plantgridFDisTA_10km <- mapping_10km$plantgridFDisTA
+plant_cellFDis_10km <- mapping_10km$spatial_fdis_plant_grid
+```
+
+```{r}
+# add title to maps
+(plantgridFDisTA_10km <- plantgridFDisTA_10km +
+    ggtitle("10 km")
+)
+ggsave("plantgridFDisTA_10km.png", plantgridFDisTA_10km, path = figure_path)
+```
+
+```{r}
+# Generate coordinates
+subset_coords_plant_5km <-site_loc_key_plant_5km[rowSums(PAM_plant_site_final_5km) >= 4, ]
+subset_coords_plant_sp_5km <-subset_coords_plant_5km[,1:2 ]
+
+plant_fd_sp_5km <- data.frame(subset_coords_plant_sp_5km, fdis_plant_5km)
+```
+
+```{r}
+# Convert the dataframe to sf format
+spatial_fdis_plant_5km <- st_as_sf(plant_fd_sp_5km, coords = c("Longitude", "Latitude"))
+```
+
+```{r}
+# set crs of sf objects
+spatial_fdis_plant_5km <- spatial_fdis_plant_5km %>% st_set_crs(5389)
+```
+
+```{r}
+# Extract coordinates 
+spatial_fdis_plant_coords_5km <- st_coordinates(spatial_fdis_plant_5km)
+
+# Add coordinates as separate columns
+spatial_fdis_plant_5km <- spatial_fdis_plant_5km %>%
+  mutate(x = spatial_fdis_plant_coords_5km[, 1], y = spatial_fdis_plant_coords_5km[, 2])
+```
+
+```{r}
+summary(spatial_fdis_plant_5km)
+```
+
+```{r}
+TAGrid_5km <- TApoly %>%
+  st_make_grid(cellsize = c(5000)) %>%
+  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+  st_cast("MULTIPOLYGON") %>%
+  st_sf() %>%
+  mutate(cellid = row_number())
+
+spatial_fdis_plant_5km$cellid <- st_intersects(spatial_fdis_plant_5km, TAGrid_5km)$nn
+
+```
+
+```{r}
+# Determine number of cores to use
+num_cores <- detectCores() - 2
+
+# Register parallel backend
+cl <- makeCluster(num_cores)
+registerDoParallel(cl)
+
+# Ensure the cluster stops when done
+on.exit(stopCluster(cl))
+
+# Split TAGrid_5km indices into chunks for parallel processing
+chunk_indices <- split(seq_len(nrow(TAGrid_5km)), rep(1:num_cores, each = ceiling(nrow(TAGrid_5km) / num_cores), length.out = nrow(TAGrid_5km)))
+
+# Perform spatial join and aggregation in parallel
+results_list <- foreach(indices = chunk_indices, .combine = 'c', .packages = c("sf", "dplyr")) %dopar% {
+  grid_chunk <- TAGrid_5km[indices, ]
+
+  # Perform spatial join
+  joined <- st_join(grid_chunk, spatial_fdis_plant)
+
+  # Group by cellid and summarize fdis values
+  summarized <- joined %>%
     group_by(cellid) %>%
@@
     ungroup()
+
+  # Ensure that summarized is a data frame and add debugging information
+  summarized_df <- as.data.frame(summarized)
+
+  # Debugging: Check the structure of the summarized data frame
+  message("Processed chunk: ", indices[1], " to ", indices[length(indices)])
+  str(summarized_df)
+
+  list(summarized_df)  # Return as list to avoid unintended simplification
+}
+
+# Flatten the results_list and ensure each element is a data frame
+flattened_results <- map_dfr(results_list, ~ as.data.frame(.x))
+
+# Check the structure of the combined result
+str(flattened_results)
+
+# Convert to sf object if applicable
+if ("geometry" %in% colnames(flattened_results)) {
+  spatial_fdis_plant_grid <- st_as_sf(flattened_results)
+} else {
+  spatial_fdis_plant_grid <- flattened_results
+}
+
+# Check the structure of the final output
+str(spatial_fdis_plant_grid)
+
 ```
@@
 ```{r}
-summary(spatial_fdis_plant_grid_75km)
+plantgridFDisTA_5km <-
+    ggplot() +
+    geom_sf(data = Americas, fill = "white")+
+    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
+    geom_sf(data = spatial_fdis_plant_grid, aes(fill = fdis_value), color = NA) +
+    labs(fill = "FDis") +
+    ggtitle("5 km") +
+    scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,6.5)) +
+    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+    theme(panel.background = element_rect(fill = "lightblue"),
+        text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
+plantgridFDisTA_5km
 ```
@@
 ```{r}
-plantgridFDisTA_75km <-
-    ggplot() +
-    geom_sf(data = Americas, fill = "white")+
-    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
-    geom_sf(data = spatial_fdis_plant_grid_75km, aes(fill = fdis_value), color = NA) +
-    labs(fill = "FDis") +
-    ggtitle("100 km") +
-    scale_fill_distiller(palette = "Greens", direction = 1) +
-    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-    theme(panel.background = element_rect(fill = "lightblue"),
-        text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
-plantgridFDisTA_75km
+plant_cellFDis_5km <- spatial_fdis_plant_grid
 ```
+
+```{r}
+# add title to maps
+(plantgridFDisTA_5km <- plantgridFDisTA_5km +
+    ggtitle("5 km")
+)
+ggsave("plantgridFDisTA_5km.png", plantgridFDisTA_5km, path = figure_path)
+```
+
+# Multi panel plots of FD maps
+## Make sure all plots for the same taxa have the same legend scale
+```{r}
+plant_plot_legend <- function(plot){
+  plot + scale_fill_distiller(palette = "Greens", direction = 1,limits = c(0, 6.5),
+                              oob = scales::squish, breaks = seq(0, 6.5, by = 1)) +
+    theme(legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10)) +
+      annotation_scale(location = "bl", width_hint = 0.5)
+}
+frugivore_plot_legend <- function(plot){
+  plot + scale_fill_distiller(palette = "YlOrBr", direction = 1,limits = c(0, 8.5),
+                              oob = scales::squish, breaks = seq(0, 8.5, by = 1)) +
+    theme(legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10)) +
+      annotation_scale(location = "bl", width_hint = 0.5)
+}
+```
+
+```{r}
+(plantgridFDisTA_100km <- plant_plot_legend(plantgridFDisTA_100km))
+(plantgridFDisTA_75km <- plant_plot_legend(plantgridFDisTA_75km))
+(plantgridFDisTA_50km <- plant_plot_legend(plantgridFDisTA_50km))
+(plantgridFDisTA_25km <- plant_plot_legend(plantgridFDisTA_25km))
+(plantgridFDisTA_10km <- plant_plot_legend(plantgridFDisTA_10km))
+(plantgridFDisTA_5km <- plant_plot_legend(plantgridFDisTA_5km))
+```
+
+```{r}
+(frugivoregridFDisTA_100km <- frugivore_plot_legend(frugivoregridFDisTA_100km))
+(frugivoregridFDisTA_75km <- frugivore_plot_legend(frugivoregridFDisTA_75km))
+(frugivoregridFDisTA_50km <- frugivore_plot_legend(frugivoregridFDisTA_50km))
+(frugivoregridFDisTA_25km <- frugivore_plot_legend(frugivoregridFDisTA_25km))
+(frugivoregridFDisTA_10km <- frugivore_plot_legend(frugivoregridFDisTA_10km))
+(frugivoregridFDisTA_5km <- frugivore_plot_legend(frugivoregridFDisTA_5km))
+```
+
+## Plants
+```{r}
+# Arrange the plots with common legend
+all_plant_fdis_plots <- ggarrange(plantgridFDisTA_5km, plantgridFDisTA_10km,
+                                      plantgridFDisTA_25km, plantgridFDisTA_50km,
+                                      plantgridFDisTA_75km, plantgridFDisTA_100km,
+                                      ncol = 6, nrow = 1,
+                                      common.legend = TRUE, legend = "left")
+
+all_plant_fdis_plots
+```
+
+```{r}
+all_plant_fdis_plots_labeled <- ggpubr::annotate_figure(all_plant_fdis_plots,
+                                                            left = ggpubr::text_grob("Plants", face = "bold", size = 20, 
+                                                                                     rot = 90))
+all_plant_fdis_plots_labeled
+ggsave("plant_fdis_plots.png", all_plant_fdis_plots_labeled, path = figure_path, width = 16, height = 5, units = "in")
+```
+
+## Frugivores
+```{r}
+all_frugivore_fdis_plots <- ggpubr::ggarrange(frugivoregridFDisTA_5km,frugivoregridFDisTA_10km,
+                                                  frugivoregridFDisTA_25km, frugivoregridFDisTA_50km,
+                                                  frugivoregridFDisTA_75km, frugivoregridFDisTA_100km,
+                                                  ncol = 6, nrow = 1,
+                                                  common.legend = TRUE, legend = "left")
+all_frugivore_fdis_plots
+```
+
+```{r}
+all_frugivore_fdis_plots_labeled <- ggpubr::annotate_figure(all_frugivore_fdis_plots,
+                                                            left = ggpubr::text_grob("Frugivores", face = "bold", size = 20,
+                                                                                     rot = 90))
+all_frugivore_fdis_plots_labeled
+ggsave("frugivore_fdis_plots.png", all_frugivore_fdis_plots_labeled, path = figure_path, width = 16, height = 5, units = "in")
+```
+
+## Combine plants & frugivores
+```{r}
+all_fdis_plots <- ggpubr::ggarrange(all_plant_fdis_plots_labeled,
+                                        all_frugivore_fdis_plots_labeled,
+                                        ncol = 1, nrow = 2)
+# Add a white background to the plot
+all_fdis_plots <- all_fdis_plots + theme(plot.background = element_rect(fill = "white", color = NA))
+all_fdis_plots
+
+ggsave("all_fdis_plots.png", all_fdis_plots, path = figure_path, width = 16, height = 10, units = "in")
+```
+
+# Write data to csv
+Have to remove geometry before saving to csv
+```{r}
+plant_cellFDis_5km_df <- data.frame(cellid = plant_cellFDis_5km$fdis_value,
+                                        fdis_value = plant_cellFDis_5km$fdis_value)
+frugivore_cellFDis_5km_df <- data.frame(cellid = frugivore_cellFDis_5km$fdis_value,
+                                        fdis_value = frugivore_cellFDis_5km$fdis_value)
+
+plant_cellFDis_10km_df <- data.frame(cellid = plant_cellFDis_10km$fdis_value,
+                                        fdis_value = plant_cellFDis_10km$fdis_value)
+frugivore_cellFDis_10km_df <- data.frame(cellid = frugivore_cellFDis_10km$fdis_value,
+                                        fdis_value = frugivore_cellFDis_10km$fdis_value)
+
+plant_cellFDis_25km_df <- data.frame(cellid = plant_cellFDis_25km$fdis_value,
+                                        fdis_value = plant_cellFDis_25km$fdis_value)
+frugivore_cellFDis_25km_df <- data.frame(cellid = frugivore_cellFDis_25km$fdis_value,
+                                        fdis_value = frugivore_cellFDis_25km$fdis_value)
+
+plant_cellFDis_50km_df <- data.frame(cellid = plant_cellFDis_50km$fdis_value,
+                                        fdis_value = plant_cellFDis_50km$fdis_value)
+frugivore_cellFDis_50km_df <- data.frame(cellid = frugivore_cellFDis_50km$fdis_value,
+                                        fdis_value = frugivore_cellFDis_50km$fdis_value)
+
+plant_cellFDis_75km_df <- data.frame(cellid = plant_cellFDis_75km$fdis_value,
+                                        fdis_value = plant_cellFDis_75km$fdis_value)
+frugivore_cellFDis_75km_df <- data.frame(cellid = frugivore_cellFDis_75km$fdis_value,
+                                        fdis_value = frugivore_cellFDis_75km$fdis_value)
+
+plant_cellFDis_100km_df <- data.frame(cellid = plant_cellFDis_100km$fdis_value,
+                                        fdis_value = plant_cellFDis_100km$fdis_value)
+frugivore_cellFDis_100km_df <- data.frame(cellid = frugivore_cellFDis_100km$fdis_value,
+                                        fdis_value = frugivore_cellFDis_100km$fdis_value)
+```
+
+```{r}
+write.csv(plant_cellFDis_5km_df, file.path(output_path,"TropicalAndes_plantFDis_5km.csv"), row.names = FALSE)
+write.csv(frugivore_cellFDis_5km_df, file.path(output_path,"TropicalAndes_frugivoreFDis_5km.csv"), row.names = FALSE)
+```
+
+```{r}
+write.csv(plant_cellFDis_10km_df, file.path(output_path,"TropicalAndes_plantFDis_10km.csv"), row.names = FALSE)
+write.csv(frugivore_cellFDis_10km_df, file.path(output_path,"TropicalAndes_frugivoreFDis_10km.csv"), row.names = FALSE)
+```
+
+```{r}
+write.csv(plant_cellFDis_25km_df, file.path(output_path,"TropicalAndes_plantFDis_25km.csv"), row.names = FALSE)
+write.csv(frugivore_cellFDis_25km_df, file.path(output_path,"TropicalAndes_frugivoreFDis_25km.csv"), row.names = FALSE)
+```
+
+```{r}
+write.csv(plant_cellFDis_50km_df, file.path(output_path,"TropicalAndes_plantFDis_50km.csv"), row.names = FALSE)
+write.csv(frugivore_cellFDis_50km_df, file.path(output_path,"TropicalAndes_frugivoreFDis_50km.csv"), row.names = FALSE)
+```
+
+```{r}
+write.csv(plant_cellFDis_75km_df, file.path(output_path,"TropicalAndes_plantFDis_75km.csv"), row.names = FALSE)
+write.csv(frugivore_cellFDis_75km_df, file.path(output_path,"TropicalAndes_frugivoreFDis_75km.csv"), row.names = FALSE)
+```
+
+```{r}
+write.csv(plant_cellFDis_100km_df, file.path(output_path,"TropicalAndes_plantFDis_100km.csv"), row.names = FALSE)
+write.csv(frugivore_cellFDis_100km_df, file.path(output_path,"TropicalAndes_frugivoreFDis_100km.csv"), row.names = FALSE)
+```

### 2024-06-24 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
adjust plots before combining into multi-panel plots, remove geometry from richness dataframes before saving as csv

@@
 Note: some code is adapted from https://luisdva.github.io/rstats/richness/."
-#data input: "TropicalAndes_plant_occ.csv, TropicalAndes_frugivore_occ.csv"
-#data output: "TropicalAndes_frugivore_richness.csv, TropicalAndes_GBIF_plant_richness.csv"
 date: "2023-08-01"
@@
 # Add labels and remove axis labels from all plots
-plantgridRichTA_5km_labeled <- plantgridRichTA_5km + ggtitle("5 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
-plantgridRichTA_10km_labeled <- plantgridRichTA_10km + ggtitle("10 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
-plantgridRichTA_25km_labeled <- plantgridRichTA_25km + ggtitle("25 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
-plantgridRichTA_50km_labeled <- plantgridRichTA_50km + ggtitle("50 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
-plantgridRichTA_75km_labeled <- plantgridRichTA_75km + ggtitle("75 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
-plantgridRichTA_100km_labeled <- plantgridRichTA_100km + ggtitle("100 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
+plantgridRichTA_5km_labeled <- plantgridRichTA_5km + ggtitle("5 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
+plantgridRichTA_10km_labeled <- plantgridRichTA_10km + ggtitle("10 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
+plantgridRichTA_25km_labeled <- plantgridRichTA_25km + ggtitle("25 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
+plantgridRichTA_50km_labeled <- plantgridRichTA_50km + ggtitle("50 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
+plantgridRichTA_75km_labeled <- plantgridRichTA_75km + ggtitle("75 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
+plantgridRichTA_100km_labeled <- plantgridRichTA_100km + ggtitle("100 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
 plantgridRichTA_100km_labeled
@@
 all_plant_richness_plots_labeled <- ggpubr::annotate_figure(all_plant_richness_plots,
-                                                            left = ggpubr::text_grob("Plants", face = "bold", size = 14, 
+                                                            left = ggpubr::text_grob("Plants", face = "bold", size = 20, 
                                                                                      rot = 90))
 all_plant_richness_plots_labeled
-ggsave("plant_richness_plots.png", all_plant_richness_plots_labeled, path = figure_path, width = 20, height = 5, units = "in")
+ggsave("plant_richness_plots.png", all_plant_richness_plots_labeled, path = figure_path, width = 16, height = 5, units = "in")
 ```
@@
 # Add labels and remove axis labels from all plots
-frugivoregridRichTA_5km_labeled <- frugivoregridRichTA_5km + ggtitle("5 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
-frugivoregridRichTA_10km_labeled <- frugivoregridRichTA_10km + ggtitle("10 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
-frugivoregridRichTA_25km_labeled <- frugivoregridRichTA_25km + ggtitle("25 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
-frugivoregridRichTA_50km_labeled <- frugivoregridRichTA_50km + ggtitle("50 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
-frugivoregridRichTA_75km_labeled <- frugivoregridRichTA_75km + ggtitle("75 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
-frugivoregridRichTA_100km_labeled <- frugivoregridRichTA_100km + ggtitle("100 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank())
+frugivoregridRichTA_5km_labeled <- frugivoregridRichTA_5km + ggtitle("5 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
+frugivoregridRichTA_10km_labeled <- frugivoregridRichTA_10km + ggtitle("10 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
+frugivoregridRichTA_25km_labeled <- frugivoregridRichTA_25km + ggtitle("25 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
+frugivoregridRichTA_50km_labeled <- frugivoregridRichTA_50km + ggtitle("50 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
+frugivoregridRichTA_75km_labeled <- frugivoregridRichTA_75km + ggtitle("75 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
+frugivoregridRichTA_100km_labeled <- frugivoregridRichTA_100km + ggtitle("100 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
 frugivoregridRichTA_100km_labeled
@@
 all_frugivore_richness_plots_labeled <- ggpubr::annotate_figure(all_frugivore_richness_plots,
-                                                            left = ggpubr::text_grob("Frugivores", face = "bold", size = 14,
+                                                            left = ggpubr::text_grob("Frugivores", face = "bold", size = 20,
                                                                                      rot = 90))
 all_frugivore_richness_plots_labeled
-ggsave("frugivore_richness_plots.png", all_frugivore_richness_plots_labeled, path = figure_path, width = 20, height = 4, units = "in")
+ggsave("frugivore_richness_plots.png", all_frugivore_richness_plots_labeled, path = figure_path, width = 16, height = 5, units = "in")
 ```
@@
 
-ggsave("all_richness_plots.png", all_richness_plots, path = figure_path, width = 20, height = 8, units = "in")
+ggsave("all_richness_plots.png", all_richness_plots, path = figure_path, width = 16, height = 10, units = "in")
 ```
@@
 # Write data to csv
+Have to remove geometry before saving to csv
 ```{r}
-write.csv(plant_cellRichness_5km, file.path(output_path,"TropicalAndes_plantRichness_5km.csv"))
-write.csv(frugivore_cellRichness_5km, file.path(output_path,"TropicalAndes_frugivoreRichness_5km.csv"))
+plant_cellRichness_5km_df <- data.frame(cellid = plant_cellRichness_5km$cellid,
+                                        num_species = plant_cellRichness_5km$num_species)
+frugivore_cellRichness_5km_df <- data.frame(cellid = frugivore_cellRichness_5km$cellid,
+                                            num_species = frugivore_cellRichness_5km$num_species)
+
+plant_cellRichness_10km_df <- data.frame(cellid = plant_cellRichness_10km$cellid,
+                                        num_species = plant_cellRichness_10km$num_species)
+frugivore_cellRichness_10km_df <- data.frame(cellid = frugivore_cellRichness_10km$cellid,
+                                            num_species = frugivore_cellRichness_10km$num_species)
+
+plant_cellRichness_25km_df <- data.frame(cellid = plant_cellRichness_25km$cellid,
+                                        num_species = plant_cellRichness_25km$num_species)
+frugivore_cellRichness_25km_df <- data.frame(cellid = frugivore_cellRichness_25km$cellid,
+                                            num_species = frugivore_cellRichness_25km$num_species)
+
+plant_cellRichness_50km_df <- data.frame(cellid = plant_cellRichness_50km$cellid,
+                                        num_species = plant_cellRichness_50km$num_species)
+frugivore_cellRichness_50km_df <- data.frame(cellid = frugivore_cellRichness_50km$cellid,
+                                            num_species = frugivore_cellRichness_50km$num_species)
+
+plant_cellRichness_75km_df <- data.frame(cellid = plant_cellRichness_75km$cellid,
+                                        num_species = plant_cellRichness_75km$num_species)
+frugivore_cellRichness_75km_df <- data.frame(cellid = frugivore_cellRichness_75km$cellid,
+                                            num_species = frugivore_cellRichness_75km$num_species)
+
+plant_cellRichness_100km_df <- data.frame(cellid = plant_cellRichness_100km$cellid,
+                                        num_species = plant_cellRichness_100km$num_species)
+frugivore_cellRichness_100km_df <- data.frame(cellid = frugivore_cellRichness_100km$cellid,
+                                            num_species = frugivore_cellRichness_100km$num_species)
 ```
@@
 ```{r}
-write.csv(plant_cellRichness_10km, file.path(output_path,"TropicalAndes_plantRichness_10km.csv"))
-write.csv(frugivore_cellRichness_10km, file.path(output_path,"TropicalAndes_frugivoreRichness_10km.csv"))
+write.csv(plant_cellRichness_5km_df, file.path(output_path,"TropicalAndes_plantRichness_5km.csv"), row.names = FALSE)
+write.csv(frugivore_cellRichness_5km_df, file.path(output_path,"TropicalAndes_frugivoreRichness_5km.csv"), row.names = FALSE)
 ```
@@
 ```{r}
-write.csv(plant_cellRichness_25km, file.path(output_path,"TropicalAndes_plantRichness_25km.csv"))
-write.csv(frugivore_cellRichness_25km, file.path(output_path,"TropicalAndes_frugivoreRichness_25km.csv"))
+write.csv(plant_cellRichness_10km_df, file.path(output_path,"TropicalAndes_plantRichness_10km.csv"), row.names = FALSE)
+write.csv(frugivore_cellRichness_10km_df, file.path(output_path,"TropicalAndes_frugivoreRichness_10km.csv"), row.names = FALSE)
 ```
@@
 ```{r}
-write.csv(plant_cellRichness_50km, file.path(output_path,"TropicalAndes_plantRichness_50km.csv"))
-write.csv(frugivore_cellRichness_50km, file.path(output_path,"TropicalAndes_frugivoreRichness_50km.csv"))
+write.csv(plant_cellRichness_25km_df, file.path(output_path,"TropicalAndes_plantRichness_25km.csv"), row.names = FALSE)
+write.csv(frugivore_cellRichness_25km_df, file.path(output_path,"TropicalAndes_frugivoreRichness_25km.csv"), row.names = FALSE)
 ```
@@
 ```{r}
-write.csv(plant_cellRichness_75km, file.path(output_path,"TropicalAndes_plantRichness_75km.csv"))
-write.csv(frugivore_cellRichness_75km, file.path(output_path,"TropicalAndes_frugivoreRichness_75km.csv"))
+write.csv(plant_cellRichness_50km_df, file.path(output_path,"TropicalAndes_plantRichness_50km.csv"), row.names = FALSE)
+write.csv(frugivore_cellRichness_50km_df, file.path(output_path,"TropicalAndes_frugivoreRichness_50km.csv"), row.names = FALSE)
 ```
@@
 ```{r}
-write.csv(plant_cellRichness_100km, file.path(output_path,"TropicalAndes_plantRichness_100km.csv"))
-write.csv(frugivore_cellRichness_100km, file.path(output_path,"TropicalAndes_frugivoreRichness_100km.csv"))
+write.csv(plant_cellRichness_75km_df, file.path(output_path,"TropicalAndes_plantRichness_75km.csv"), row.names = FALSE)
+write.csv(frugivore_cellRichness_75km_df, file.path(output_path,"TropicalAndes_frugivoreRichness_75km.csv"), row.names = FALSE)
 ```
+
+```{r}
+write.csv(plant_cellRichness_100km_df, file.path(output_path,"TropicalAndes_plantRichness_100km.csv"), row.names = FALSE)
+write.csv(frugivore_cellRichness_100km_df, file.path(output_path,"TropicalAndes_frugivoreRichness_100km.csv"), row.names = FALSE)
+```

### 2024-06-24 code/L2/L2_3_TropicalAndes_divRelationships.Rmd
create plots to compare tax and fun div between plants and frugivores using multiple models

@@
 collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
-overview: "This script compares taxonomic and functional diversity relationships for plants and frugivores for various spatial grains and combined diversity for each grain."
-data input: "ADD"
-data output: "ADD"
+overview: "This script compares taxonomic and functional diversity relationships for plants and frugivores for various spatial grains."
 date: "2024-05-16"
@@
 library(purrr)
+library(ggtrendline)
+library(ggpubr)
+library(tidyr)
 ```
@@
 ```{r}
-plant_cellRichness_1km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_1km.csv"))
-frugivore_cellRichness_1km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_1km.csv"))
-
 plant_cellRichness_5km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_5km.csv"))
@@
 
-# 1 km
-# 5 km
-# 10 km
-# 25 km
-# 50 km
-# 75 km
-# 100 km
 ```{r}
-  # Extract cell values
-  comparison_data_100km <- data.frame(Plant_Richness = plant_richness_grid_100km$num_species,
-                               Frugivore_Richness = frugivore_richness_grid_100km$num_species)
+plant_cellFDis_5km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantFDis_5km.csv"))
+frugivore_cellFDis_5km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreFDis_5km.csv"))
 
-  # Filter out rows where either plant or frugivore richness is zero
-  comparison_data_filtered_100km <- comparison_data %>%
-                                filter(Plant_Richness > 0 & Frugivore_Richness > 0)
+plant_cellFDis_10km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantFDis_10km.csv"))
+frugivore_cellFDis_10km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreFDis_10km.csv"))
+
+plant_cellFDis_25km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantFDis_25km.csv"))
+frugivore_cellFDis_25km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreFDis_25km.csv"))
+
+plant_cellFDis_50km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantFDis_50km.csv"))
+frugivore_cellFDis_50km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreFDis_50km.csv"))
+
+plant_cellFDis_75km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantFDis_75km.csv"))
+frugivore_cellFDis_75km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreFDis_75km.csv"))
+
+plant_cellFDis_100km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantFDis_100km.csv"))
+frugivore_cellFDis_100km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreFDis_100km.csv"))
+```
+
+# Preliminary plots
+Check the data and inspect possible relationships
+
+```{r}
+plot(plant_cellRichness_100km$num_species, frugivore_cellRichness_100km$num_species)
+plot(plant_cellRichness_75km$num_species, frugivore_cellRichness_75km$num_species)
+plot(plant_cellRichness_50km$num_species, frugivore_cellRichness_50km$num_species)
+plot(plant_cellRichness_25km$num_species, frugivore_cellRichness_25km$num_species)
+plot(plant_cellRichness_10km$num_species, frugivore_cellRichness_10km$num_species)
+plot(plant_cellRichness_5km$num_species, frugivore_cellRichness_5km$num_species)
 ```
@@
 ```{r}
-# Compare plant and frugivore richness
-  ## Fit a linear model
-  compare_richness_linear_summary_100km <- summary(lm(comparison_data_filtered_100km$Frugivore_Richness ~
-                                                  comparison_data_filtered_100km$Plant_Richness))
-  linear_r_squared_100km <- compare_richness_linear_summary_100km$r.squared
-
-  ## Fit an exponential model
-  compare_richness_exponential_summary_100km <- summary(lm(comparison_data_filtered_100km$Frugivore_Richness ~
-                                                       log(comparison_data_filtered_100km$Plant_Richness)))
-  exponential_r_squared_100km <- compare_richness_exponential_summary_100km$r.squared
-
-  ## Create scatter plot with linear and exponential lines
-  compare_richness_plot_100km <- ggplot(comparison_data_filtered_100km,
-                                        aes(x = Plant_Richness, y = Frugivore_Richness)) +
-  geom_point() +
-  labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
-       subtitle =  paste("Linear (Red) R-squared = ", round(linear_r_squared, 2), "\n",
-                  "Exponential (Blue) R-squared = ", round(exponential_r_squared, 2)),
-       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
-  geom_smooth(method = "lm", se = FALSE, color = "red")+
-  geom_smooth(method = "lm", se = FALSE, color = "blue", formula = y ~ log(x)) +
-  coord_cartesian(xlim = c(0, NA), ylim = c(0, NA))
+plot(plant_cellFDis_100km$fdis_value, frugivore_cellFDis_100km$fdis_value)
+plot(plant_cellFDis_75km$fdis_value, frugivore_cellFDis_75km$fdis_value)
+plot(plant_cellFDis_50km$fdis_value, frugivore_cellFDis_50km$fdis_value)
+plot(plant_cellFDis_25km$fdis_value, frugivore_cellFDis_25km$fdis_value)
+plot(plant_cellFDis_10km$fdis_value, frugivore_cellFDis_10km$fdis_value)
+plot(plant_cellFDis_5km$fdis_value, frugivore_cellFDis_5km$fdis_value)
 ```
@@
 ```{r}
-# Fit the nonlinear model using nls
-nls_model <- nls(Frugivore_Richness ~ b * exp(c * Plant_Richness), 
-                 data = comparison_data_filtered_100km, 
-                 start = list(b = 200, c = -0.00001),
-                 algorithm = "port",
-                 control = nls.control(maxiter = 500, minFactor = 1e-10, warnOnly = TRUE))
+plot(plant_cellRichness_5km$num_species, plant_cellFDis_5km$fdis_value)
+plot(plant_cellRichness_10km$num_species, plant_cellFDis_10km$fdis_value)
+plot(plant_cellRichness_25km$num_species, plant_cellFDis_25km$fdis_value)
+plot(plant_cellRichness_50km$num_species, plant_cellFDis_50km$fdis_value)
+plot(plant_cellRichness_75km$num_species, plant_cellFDis_75km$fdis_value)
+plot(plant_cellRichness_100km$num_species, plant_cellFDis_100km$fdis_value)
+```
 
-# Summary of the model
-summary(nls_model)
+```{r}
+plot(frugivore_cellRichness_5km$num_species, frugivore_cellFDis_5km$fdis_value)
+plot(frugivore_cellRichness_10km$num_species, frugivore_cellFDis_10km$fdis_value)
+plot(frugivore_cellRichness_25km$num_species, frugivore_cellFDis_25km$fdis_value)
+plot(frugivore_cellRichness_50km$num_species, frugivore_cellFDis_50km$fdis_value)
+plot(frugivore_cellRichness_75km$num_species, frugivore_cellFDis_75km$fdis_value)
+plot(frugivore_cellRichness_100km$num_species, frugivore_cellFDis_100km$fdis_value)
+```
+
+# Compare diversity relationships between plants and frugivores
+## Function for comparing diversity relationships
+For comparing richness, use exp3P and line2P for line types
+For comparing FDis, use line3P and line2P for line types
+
+line colors: line2P is "red", line3P is "blue", exp3P is "purple"
+```{r}
+compare_Richness <- function(plant_Richness_grid, frugivore_Richness_grid, resolution_meters){
+  # Extract cell values
+  comparison_Richness <- data.frame(Plant_Richness = plant_Richness_grid$num_species,
+                               Frugivore_Richness = frugivore_Richness_grid$num_species)
+
+  # Filter out rows where either plant or frugivore richness is zero
+  comparison_Richness_filtered <- comparison_Richness %>%
+                                filter(Plant_Richness > 0 & Frugivore_Richness > 0)
+
+  # Calculate Annotation position
+  mean_x <- max(comparison_Richness_filtered$Plant_Richness) / 2
+  annot_y <- max(comparison_Richness_filtered$Frugivore_Richness) + 25
+
+  # Create scatter plots
+  compare_Richness_plot_line2P <- ggtrendline(comparison_Richness_filtered$Plant_Richness,
+               comparison_Richness_filtered$Frugivore_Richness, model = "line2P", linecolor = "red", linewidth = 1,
+               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, rrp.x = mean_x, rrp.y = annot_y, 
+               text.col = "red", eSize = 4) +
+      geom_point(aes(x = comparison_Richness_filtered$Plant_Richness,
+                                           y = comparison_Richness_filtered$Frugivore_Richness), alpha = 0.25, size = 1) +
+      labs(title = paste("Comparing Richness [", resolution_meters/1000, " km]", sep = ""),
+           subtitle = "Fitted with a linear model",
+       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
+      scale_x_continuous(expand = c(0, 0)) + 
+      scale_y_continuous(expand = c(0, 0)) + 
+      theme_minimal() +
+      theme(panel.grid.major = element_blank(), 
+            panel.grid.minor = element_blank(),
+            axis.line = element_line(color = "black"),
+            axis.ticks = element_line(color = "black"),
+            axis.text = element_text(color = "black"))
+
+  compare_Richness_plot_exp3P <- ggtrendline(comparison_Richness_filtered$Plant_Richness,
+               comparison_Richness_filtered$Frugivore_Richness, model = "exp3P", linecolor = "purple", linewidth = 1,
+               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, rrp.x = mean_x, rrp.y = annot_y, 
+               text.col = "purple", eSize = 4) +
+      geom_point(aes(x = comparison_Richness_filtered$Plant_Richness,
+                                           y = comparison_Richness_filtered$Frugivore_Richness), alpha = 0.25, size = 1) +
+      labs(title = paste("Comparing Richness [", resolution_meters/1000, " km]", sep = ""),
+           subtitle = "Fitted with an exponential model",
+       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
+      scale_x_continuous(expand = c(0, 0)) + 
+      scale_y_continuous(expand = c(0, 0)) + 
+      theme_minimal() +
+      theme(panel.grid.major = element_blank(), 
+            panel.grid.minor = element_blank(),
+            axis.line = element_line(color = "black"),
+            axis.ticks = element_line(color = "black"),
+            axis.text = element_text(color = "black"))
+
+    # Summaries of models used for trendlines
+  line2P_summary <- trendline_sum(comparison_Richness_filtered$Plant_Richness,
+               comparison_Richness_filtered$Frugivore_Richness, model = "line2P")
+  exp3P_summary <- trendline_sum(comparison_Richness_filtered$Plant_Richness,
+               comparison_Richness_filtered$Frugivore_Richness, model = "exp3P")
+
+  # Combine trendlines into one plot
+  compare_Richness_plot_both <- ggtrendline(comparison_Richness_filtered$Plant_Richness,
+               comparison_Richness_filtered$Frugivore_Richness, model = "exp3P", linecolor = "purple", linewidth = 1,
+               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, show.Rsquare = FALSE) +
+      geom_smooth(aes(x = comparison_Richness_filtered$Plant_Richness, y = comparison_Richness_filtered$Frugivore_Richness), 
+              method = "lm", se = FALSE, color = "red", linetype = "solid") +
+      geom_point(aes(x = comparison_Richness_filtered$Plant_Richness,
+                                           y = comparison_Richness_filtered$Frugivore_Richness), alpha = 0.25, size = 1) +
+      labs(title = paste("Comparing Richness [", resolution_meters/1000, " km]", sep = ""),
+       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
+      scale_x_continuous(expand = c(0, 0)) + 
+      scale_y_continuous(expand = c(0, 0)) + 
+      theme_minimal() +
+      theme(panel.grid.major = element_blank(), 
+            panel.grid.minor = element_blank(),
+            axis.line = element_line(color = "black"),
+            axis.ticks = element_line(color = "black"),
+            axis.text = element_text(color = "black")) +
+      annotate("text", x = mean_x, y = annot_y-10, label = paste("Linear R² =", round(line2P_summary$R.squared, 3)),
+               color = "red", size = 4) +
+      annotate("text", x = mean_x, y = annot_y-30, label = paste("Exponential R² =", round(exp3P_summary$R.squared, 3)),
+               color = "purple", size = 4)
+
+  list(compare_Richness_plot_line2P = compare_Richness_plot_line2P,
+       compare_Richness_plot_exp3P = compare_Richness_plot_exp3P,
+       line2P_summary = line2P_summary,
+       exp3P_summary = exp3P_summary,
+       compare_Richness_plot_both = compare_Richness_plot_both)
+}
+```
+
+```{r}
+compare_FDis <- function(plant_FDis_grid, frugivore_FDis_grid, resolution_meters){
+  # Extract cell values
+  comparison_FDis <- data.frame(Plant_FDis = plant_FDis_grid$fdis_value,
+                               Frugivore_FDis = frugivore_FDis_grid$fdis_value)
+
+  # Filter out rows where either plant or frugivore FDis is zero
+  comparison_FDis_filtered <- comparison_FDis %>%
+                                filter(Plant_FDis > 0 & Frugivore_FDis > 0)
+  # Calculate Annotation position
+  mean_x <- max(comparison_FDis_filtered$Plant_FDis) / 2
+  annot_y <- max(comparison_FDis_filtered$Frugivore_FDis) + 0.25
+
+  ## Create scatter plots
+  compare_FDis_plot_line2P <- ggtrendline(comparison_FDis_filtered$Plant_FDis,
+               comparison_FDis_filtered$Frugivore_FDis, model = "line2P", linecolor = "red", linewidth = 1,
+               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, rrp.x = mean_x, rrp.y = annot_y, 
+               text.col = "red", eSize = 4) +
+      geom_point(aes(x = comparison_FDis_filtered$Plant_FDis,
+                                           y = comparison_FDis_filtered$Frugivore_FDis), alpha = 0.25, size = 1) +
+      labs(title = paste("Compare FDis [", resolution_meters/1000, " km]", sep = ""),
+           subtitle = "Fitted with a linear model",
+       x = "Plant FDis by cell", y = "Frugivore FDis by cell") +
+      scale_x_continuous(expand = c(0, 0)) + 
+      scale_y_continuous(expand = c(0, 0)) + 
+      theme_minimal() +
+      theme(panel.grid.major = element_blank(), 
+            panel.grid.minor = element_blank(),
+            axis.line = element_line(color = "black"),
+            axis.ticks = element_line(color = "black"),
+            axis.text = element_text(color = "black"))
+
+  compare_FDis_plot_line3P <- ggtrendline(comparison_FDis_filtered$Plant_FDis,
+               comparison_FDis_filtered$Frugivore_FDis, model = "line3P", linecolor = "blue", linewidth = 1,
+               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, rrp.x = mean_x, rrp.y = annot_y, 
+               text.col = "blue", eSize = 4) +
+      geom_point(aes(x = comparison_FDis_filtered$Plant_FDis,
+                                           y = comparison_FDis_filtered$Frugivore_FDis), alpha = 0.25, size = 1) +
+      labs(title = paste("Compare FDis [", resolution_meters/1000, " km]", sep = ""),
+           subtitle = "Fitted with a quadratic model",
+       x = "Plant FDis by cell", y = "Frugivore FDis by cell") +
+      scale_x_continuous(expand = c(0, 0)) + 
+      scale_y_continuous(expand = c(0, 0)) + 
+      theme_minimal() +
+      theme(panel.grid.major = element_blank(), 
+            panel.grid.minor = element_blank(),
+            axis.line = element_line(color = "black"),
+            axis.ticks = element_line(color = "black"),
+            axis.text = element_text(color = "black"))
+
+  # Summaries of models used for trendlines
+  line2P_summary <- trendline_sum(comparison_FDis_filtered$Plant_FDis,
+               comparison_FDis_filtered$Frugivore_FDis, model = "line2P")
+  line3P_summary <- trendline_sum(comparison_FDis_filtered$Plant_FDis,
+               comparison_FDis_filtered$Frugivore_FDis, model = "line3P")
+
+  # Combine trendlines into one plot
+  compare_FDis_plot_both <- ggtrendline(comparison_FDis_filtered$Plant_FDis,
+               comparison_FDis_filtered$Frugivore_FDis, model = "line3P", linecolor = "blue", linewidth = 1,
+               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, show.Rsquare = FALSE) +
+      geom_smooth(aes(x = comparison_FDis_filtered$Plant_FDis, y = comparison_FDis_filtered$Frugivore_FDis), 
+              method = "lm", se = FALSE, color = "red", linetype = "solid") +
+      geom_point(aes(x = comparison_FDis_filtered$Plant_FDis,
+                                           y = comparison_FDis_filtered$Frugivore_FDis), alpha = 0.25, size = 1) +
+      labs(title = paste("Comparing FDis [", resolution_meters/1000, " km]", sep = ""),
+       x = "Plant FDis by cell", y = "Frugivore FDis by cell") +
+      scale_x_continuous(expand = c(0, 0)) + 
+      scale_y_continuous(expand = c(0, 0)) + 
+      theme_minimal() +
+      theme(panel.grid.major = element_blank(), 
+            panel.grid.minor = element_blank(),
+            axis.line = element_line(color = "black"),
+            axis.ticks = element_line(color = "black"),
+            axis.text = element_text(color = "black")) +
+      annotate("text", x = mean_x, y = annot_y, label = paste("Linear R² =", round(line2P_summary$R.squared, 3)),
+               color = "red", size = 4) +
+      annotate("text", x = mean_x, y = annot_y-0.5, label = paste("Quadratic R² =", round(line3P_summary$R.squared, 3)),
+               color = "blue", size = 4)
+
+  list(compare_FDis_plot_line2P = compare_FDis_plot_line2P,
+       compare_FDis_plot_line3P = compare_FDis_plot_line3P,
+       line2P_summary = line2P_summary,
+       line3P_summary = line3P_summary,
+       compare_FDis_plot_both = compare_FDis_plot_both)
+}
+```
+
+## 5 km
+```{r}
+compare_Richness_5km <- compare_Richness(plant_cellRichness_5km, frugivore_cellRichness_5km, 5000)
+(compare_Richness_5km_linear <- compare_Richness_5km$compare_Richness_plot_line2P)
+(compare_Richness_5km_exp <- compare_Richness_5km$compare_Richness_plot_exp3P)
+(compare_Richness_5km_plots <- compare_Richness_5km$compare_Richness_plot_both)
+compare_Richness_r2_5km_linear <- compare_Richness_5km$line2P_summary$R.squared
+compare_Richness_r2_5km_exp <- compare_Richness_5km$exp3P_summary$R.squared
+compare_Richness_slopecoefficent_5km_linear <- compare_Richness_5km$line2P_summary$parameter["a"]
+compare_Richness_slopecoefficent_5km_exp <- compare_Richness_5km$exp3P_summary$parameter["b"]
+```
+
+```{r}
+compare_FDis_5km <- compare_FDis(plant_cellFDis_5km, frugivore_cellFDis_5km, 5000)
+(compare_FDis_5km_linear <- compare_FDis_5km$compare_FDis_plot_line2P)
+(compare_FDis_5km_quad <- compare_FDis_5km$compare_FDis_plot_line3P)
+(compare_FDis_5km_plots <- compare_FDis_5km$compare_FDis_plot_both)
+compare_FDis_r2_5km_linear <- compare_FDis_5km$line2P_summary$R.squared
+compare_FDis_r2_5km_quad <- compare_FDis_5km$line3P_summary$R.squared
+compare_FDis_slopecoefficent_5km_linear <- compare_FDis_5km$line2P_summary$parameter["a"]
+compare_FDis_slopecoefficent_5km_quad <- compare_FDis_5km$line3P_summary$parameter["b"]
+```
+
+## 10 km
+```{r}
+compare_Richness_10km <- compare_Richness(plant_cellRichness_10km, frugivore_cellRichness_10km, 10000)
+(compare_Richness_10km_linear <- compare_Richness_10km$compare_Richness_plot_line2P)
+(compare_Richness_10km_exp <- compare_Richness_10km$compare_Richness_plot_exp3P)
+(compare_Richness_10km_plots <- compare_Richness_10km$compare_Richness_plot_both)
+compare_Richness_r2_10km_linear <- compare_Richness_10km$line2P_summary$R.squared
+compare_Richness_r2_10km_exp <- compare_Richness_10km$exp3P_summary$R.squared
+compare_Richness_slopecoefficent_10km_linear <- compare_Richness_10km$line2P_summary$parameter["a"]
+compare_Richness_slopecoefficent_10km_exp <- compare_Richness_10km$exp3P_summary$parameter["b"]
+```
+
+```{r}
+compare_FDis_10km <- compare_FDis(plant_cellFDis_10km, frugivore_cellFDis_10km, 10000)
+(compare_FDis_10km_linear <- compare_FDis_10km$compare_FDis_plot_line2P)
+(compare_FDis_10km_quad <- compare_FDis_10km$compare_FDis_plot_line3P)
+(compare_FDis_10km_plots <- compare_FDis_10km$compare_FDis_plot_both)
+compare_FDis_r2_10km_linear <- compare_FDis_10km$line2P_summary$R.squared
+compare_FDis_r2_10km_quad <- compare_FDis_10km$line3P_summary$R.squared
+compare_FDis_slopecoefficent_10km_linear <- compare_FDis_10km$line2P_summary$parameter["a"]
+compare_FDis_slopecoefficent_10km_quad <- compare_FDis_10km$line3P_summary$parameter["b"]
+```
+
+## 25 km
+```{r}
+compare_Richness_25km <- compare_Richness(plant_cellRichness_25km, frugivore_cellRichness_25km, 25000)
+(compare_Richness_25km_linear <- compare_Richness_25km$compare_Richness_plot_line2P)
+(compare_Richness_25km_exp <- compare_Richness_25km$compare_Richness_plot_exp3P)
+(compare_Richness_25km_plots <- compare_Richness_25km$compare_Richness_plot_both)
+compare_Richness_r2_25km_linear <- compare_Richness_25km$line2P_summary$R.squared
+compare_Richness_r2_25km_exp <- compare_Richness_25km$exp3P_summary$R.squared
+compare_Richness_slopecoefficent_25km_linear <- compare_Richness_25km$line2P_summary$parameter["a"]
+compare_Richness_slopecoefficent_25km_exp <- compare_Richness_25km$exp3P_summary$parameter["b"]
+```
+
+```{r}
+compare_FDis_25km <- compare_FDis(plant_cellFDis_25km, frugivore_cellFDis_25km, 25000)
+(compare_FDis_25km_linear <- compare_FDis_25km$compare_FDis_plot_line2P)
+(compare_FDis_25km_quad <- compare_FDis_25km$compare_FDis_plot_line3P)
+(compare_FDis_25km_plots <- compare_FDis_25km$compare_FDis_plot_both)
+compare_FDis_r2_25km_linear <- compare_FDis_25km$line2P_summary$R.squared
+compare_FDis_r2_25km_quad <- compare_FDis_25km$line3P_summary$R.squared
+compare_FDis_slopecoefficent_25km_linear <- compare_FDis_25km$line2P_summary$parameter["a"]
+compare_FDis_slopecoefficent_25km_quad <- compare_FDis_25km$line3P_summary$parameter["b"]
+```
+
+## 50 km
+```{r}
+compare_Richness_50km <- compare_Richness(plant_cellRichness_50km, frugivore_cellRichness_50km, 50000)
+(compare_Richness_50km_linear <- compare_Richness_50km$compare_Richness_plot_line2P)
+(compare_Richness_50km_exp <- compare_Richness_50km$compare_Richness_plot_exp3P)
+(compare_Richness_50km_plots <- compare_Richness_50km$compare_Richness_plot_both)
+compare_Richness_r2_50km_linear <- compare_Richness_50km$line2P_summary$R.squared
+compare_Richness_r2_50km_exp <- compare_Richness_50km$exp3P_summary$R.squared
+compare_Richness_slopecoefficent_50km_linear <- compare_Richness_50km$line2P_summary$parameter["a"]
+compare_Richness_slopecoefficent_50km_exp <- compare_Richness_50km$exp3P_summary$parameter["b"]
+```
+
+```{r}
+compare_FDis_50km <- compare_FDis(plant_cellFDis_50km, frugivore_cellFDis_50km, 50000)
+(compare_FDis_50km_linear <- compare_FDis_50km$compare_FDis_plot_line2P)
+(compare_FDis_50km_quad <- compare_FDis_50km$compare_FDis_plot_line3P)
+(compare_FDis_50km_plots <- compare_FDis_50km$compare_FDis_plot_both)
+compare_FDis_r2_50km_linear <- compare_FDis_50km$line2P_summary$R.squared
+compare_FDis_r2_50km_quad <- compare_FDis_50km$line3P_summary$R.squared
+compare_FDis_slopecoefficent_50km_linear <- compare_FDis_50km$line2P_summary$parameter["a"]
+compare_FDis_slopecoefficent_50km_quad <- compare_FDis_50km$line3P_summary$parameter["b"]
+```
+
+## 75 km
+```{r}
+compare_Richness_75km <- compare_Richness(plant_cellRichness_75km, frugivore_cellRichness_75km, 75000)
+(compare_Richness_75km_linear <- compare_Richness_75km$compare_Richness_plot_line2P)
+(compare_Richness_75km_exp <- compare_Richness_75km$compare_Richness_plot_exp3P)
+(compare_Richness_75km_plots <- compare_Richness_75km$compare_Richness_plot_both)
+compare_Richness_r2_75km_linear <- compare_Richness_75km$line2P_summary$R.squared
+compare_Richness_r2_75km_exp <- compare_Richness_75km$exp3P_summary$R.squared
+compare_Richness_slopecoefficent_75km_linear <- compare_Richness_75km$line2P_summary$parameter["a"]
+compare_Richness_slopecoefficent_75km_exp <- compare_Richness_75km$exp3P_summary$parameter["b"]
+```
+
+```{r}
+compare_FDis_75km <- compare_FDis(plant_cellFDis_75km, frugivore_cellFDis_75km, 75000)
+(compare_FDis_75km_linear <- compare_FDis_75km$compare_FDis_plot_line2P)
+(compare_FDis_75km_quad <- compare_FDis_75km$compare_FDis_plot_line3P)
+(compare_FDis_75km_plots <- compare_FDis_75km$compare_FDis_plot_both)
+compare_FDis_r2_75km_linear <- compare_FDis_75km$line2P_summary$R.squared
+compare_FDis_r2_75km_quad <- compare_FDis_75km$line3P_summary$R.squared
+compare_FDis_slopecoefficent_75km_linear <- compare_FDis_75km$line2P_summary$parameter["a"]
+compare_FDis_slopecoefficent_75km_quad <- compare_FDis_75km$line3P_summary$parameter["b"]
+```
+
+## 100 km
+```{r}
+compare_Richness_100km <- compare_Richness(plant_cellRichness_100km, frugivore_cellRichness_100km, 100000)
+(compare_Richness_100km_linear <- compare_Richness_100km$compare_Richness_plot_line2P)
+(compare_Richness_100km_exp <- compare_Richness_100km$compare_Richness_plot_exp3P)
+(compare_Richness_100km_plots <- compare_Richness_100km$compare_Richness_plot_both)
+compare_Richness_r2_100km_linear <- compare_Richness_100km$line2P_summary$R.squared
+compare_Richness_r2_100km_exp <- compare_Richness_100km$exp3P_summary$R.squared
+compare_Richness_slopecoefficent_100km_linear <- compare_Richness_100km$line2P_summary$parameter["a"]
+compare_Richness_slopecoefficent_100km_exp <- compare_Richness_100km$exp3P_summary$parameter["b"]
+```
+
+```{r}
+compare_FDis_100km <- compare_FDis(plant_cellFDis_100km, frugivore_cellFDis_100km, 100000)
+(compare_FDis_100km_linear <- compare_FDis_100km$compare_FDis_plot_line2P)
+(compare_FDis_100km_quad <- compare_FDis_100km$compare_FDis_plot_line3P)
+(compare_FDis_100km_plots <- compare_FDis_100km$compare_FDis_plot_both)
+compare_FDis_r2_100km_linear <- compare_FDis_100km$line2P_summary$R.squared
+compare_FDis_r2_100km_quad <- compare_FDis_100km$line3P_summary$R.squared
+compare_FDis_slopecoefficent_100km_linear <- compare_FDis_100km$line2P_summary$parameter["a"]
+compare_FDis_slopecoefficent_100km_quad <- compare_FDis_100km$line3P_summary$parameter["b"]
+```
+
+## Combine plots using for richness and fdis into one figure
+
+```{r}
+(compare_Richness_all_plots <- ggarrange(compare_Richness_5km_plots, compare_Richness_10km_plots,
+                                     compare_Richness_25km_plots, compare_Richness_50km_plots,
+                                     compare_Richness_75km_plots, compare_Richness_100km_plots,
+                                        ncol = 3, nrow = 2))
+```
+
+```{r}
+(compare_FDis_all_plots <- ggarrange(compare_FDis_5km_plots, compare_FDis_10km_plots,
+                                     compare_FDis_25km_plots, compare_FDis_50km_plots,
+                                     compare_FDis_75km_plots, compare_FDis_100km_plots,
+                                        ncol = 3, nrow = 2))
 ```
@@
 # create dataframe of resolutions and r-squared values 
-#[ADD 1, 5,]
-resolution <- c( 10, 25, 50, 75, 100)
-#ADD compare_richness_r2_1km, compare_richness_r2_5km,
-r_squared <- c(compare_richness_r2_10km, compare_richness_r2_25km, compare_richness_r2_50km, compare_richness_r2_75km, compare_richness_r2_100km)
-r_squared_compare <- data.frame(resolution, r_squared)
+resolution <- c(5, 10, 25, 50, 75, 100)
+taxdiv_r_squared_linear <- c(compare_Richness_r2_5km_linear, compare_Richness_r2_10km_linear, compare_Richness_r2_25km_linear, compare_Richness_r2_50km_linear, compare_Richness_r2_75km_linear, compare_Richness_r2_100km_linear)
+taxdiv_r_squared_exp <- c(compare_Richness_r2_5km_exp, compare_Richness_r2_10km_exp, compare_Richness_r2_25km_exp, compare_Richness_r2_50km_exp, compare_Richness_r2_75km_exp, compare_Richness_r2_100km_exp)
+taxdiv_r_squared_compare <- data.frame(resolution, taxdiv_r_squared_linear, taxdiv_r_squared_exp)
 
-compare_r2_by_res <- ggplot(r_squared_compare, aes(x=resolution, y=r_squared)) +
-    geom_point()+
-    labs(title = "R-squared values of plant-frugivore richness relationships by resolution",
-         x = "Resolution (km)", y = "R-squared", fill = "Number of species")
-print(compare_r2_by_res)
+# reshape to long format for ease of plotting
+taxdiv_r_squared_long <- taxdiv_r_squared_compare %>%
+  pivot_longer(cols = c(taxdiv_r_squared_linear, taxdiv_r_squared_exp), 
+               names_to = "model", 
+               values_to = "taxdiv_r_squared")
+
 ```
+
+```{r}
+# Define custom colors for the points
+custom_colors <- c(taxdiv_r_squared_linear = "red", taxdiv_r_squared_exp = "purple")
+
+# Define custom legend labels
+custom_labels <- c(taxdiv_r_squared_linear = "Linear Model", taxdiv_r_squared_exp = "Exponential Model")
+
+(taxdiv_r_squared_plot <- ggplot(taxdiv_r_squared_long,aes(x = resolution, y = taxdiv_r_squared, color = model)) +
+  geom_line() +
+  geom_point(size = 3) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = taxdiv_r_squared_compare$resolution) + 
+  labs(title = "Comparison of Richness R-squared Values by Resolution",
+       x = "Resolution (km)",
+       y = "R-squared Value",
+       color = "Model Type") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = "inside",
+        legend.position.inside = c(.85, .3),
+        legend.box.margin = margin(10, 10, 10, 10)))
+
+# Calculate R-squared values for linear model
+taxdiv_lm_linear <- lm(taxdiv_r_squared_linear ~ resolution, data = taxdiv_r_squared_compare)
+taxdiv_r_squared_linear <- summary(taxdiv_lm_linear)$r.squared
+taxdiv_slopecoefficent_linear <- coef(taxdiv_lm_linear)["resolution"]
+
+# Calculate R-squared values for exponential model
+taxdiv_lm_exp <- lm(taxdiv_r_squared_exp ~ resolution, data = taxdiv_r_squared_compare)
+taxdiv_r_squared_exp <- summary(taxdiv_lm_exp)$r.squared
+taxdiv_slopecoefficent_exp <- coef(taxdiv_lm_exp)["resolution"]
+
+(taxdiv_r_squared_plot_trendlines <- ggplot() +
+  geom_smooth(aes(x = taxdiv_r_squared_compare$resolution, y = taxdiv_r_squared_compare$taxdiv_r_squared_linear), 
+              method = "lm", se = FALSE, color = "red", linetype = "solid") +
+  geom_smooth(aes(x = taxdiv_r_squared_compare$resolution, y = taxdiv_r_squared_compare$taxdiv_r_squared_exp), 
+              method = "lm", se = FALSE, color = "purple", linetype = "solid") +
+  geom_point(aes(x = taxdiv_r_squared_long$resolution, y = taxdiv_r_squared_long$taxdiv_r_squared, color = taxdiv_r_squared_long$model),
+             size = 3) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = taxdiv_r_squared_compare$resolution) + 
+  labs(title = "Comparison of Richness R-squared Values by Resolution",
+       x = "Resolution (km)",
+       y = "R-squared Value",
+       color = "Model Type") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = "inside",
+        legend.position.inside = c(.85, .3),
+        legend.box.margin = margin(10, 10, 10, 10)) +
+  annotate("text", x = 55, y = 0.4, label = paste("R² =", round(taxdiv_r_squared_linear, 2)), color = "red", size = 4) +
+  annotate("text", x = 60, y = 0.7, label = paste("R² =", round(taxdiv_r_squared_exp, 2)), color = "purple", size = 4))
+```
+
+```{r}
+ggtrendline(taxdiv_r_squared_compare$resolution, taxdiv_r_squared_compare$taxdiv_r_squared_linear, model = "line2P", summary = FALSE, linecolor = "red", linewidth = 1, CI.level = NA, show.eq = FALSE, show.pvalue = FALSE, text.col = "red", eSize = 4)
+ggtrendline(taxdiv_r_squared_compare$resolution, taxdiv_r_squared_compare$taxdiv_r_squared_exp, model = "line2P", summary = FALSE, linecolor = "purple", linewidth = 1, CI.level = NA, show.eq = FALSE, show.pvalue = FALSE, text.col = "purple", eSize = 4)
+
+ggtrendline(taxdiv_r_squared_compare$resolution, taxdiv_r_squared_compare$taxdiv_r_squared_linear, model = "line3P", summary = FALSE, linecolor = "red", linewidth = 1, CI.level = NA, show.eq = FALSE, show.pvalue = FALSE, text.col = "red", eSize = 4)
+ggtrendline(taxdiv_r_squared_compare$resolution, taxdiv_r_squared_compare$taxdiv_r_squared_exp, model = "line3P", summary = FALSE, linecolor = "purple", linewidth = 1, CI.level = NA, show.eq = FALSE, show.pvalue = FALSE, text.col = "purple", eSize = 4)
+```
+
 ## Functional Diversity
 ```{r}
+# create dataframe of resolutions and r-squared values 
+resolution <- c(5, 10, 25, 50, 75, 100)
+fundiv_r_squared_linear <- c(compare_FDis_r2_5km_linear, compare_FDis_r2_10km_linear, compare_FDis_r2_25km_linear, compare_FDis_r2_50km_linear, compare_FDis_r2_75km_linear, compare_FDis_r2_100km_linear)
+fundiv_r_squared_quad <- c(compare_FDis_r2_5km_quad, compare_FDis_r2_10km_quad, compare_FDis_r2_25km_quad, compare_FDis_r2_50km_quad, compare_FDis_r2_75km_quad, compare_FDis_r2_100km_quad)
+fundiv_r_squared_compare <- data.frame(resolution, fundiv_r_squared_linear, fundiv_r_squared_quad)
+
+# reshape to long format for ease of plotting
+fundiv_r_squared_long <- fundiv_r_squared_compare %>%
+  pivot_longer(cols = c(fundiv_r_squared_linear, fundiv_r_squared_quad), 
+               names_to = "model", 
+               values_to = "fundiv_r_squared")
 
 ```
+
+```{r}
+# Define custom colors for the points
+custom_colors <- c(fundiv_r_squared_linear = "red", fundiv_r_squared_quad = "blue")
+
+# Define custom legend labels
+custom_labels <- c(fundiv_r_squared_quad = "Quadratic Model", fundiv_r_squared_linear = "Linear Model")
+
+(fundiv_r_squared_plot <- ggplot(fundiv_r_squared_long,aes(x = resolution, y = fundiv_r_squared, color = model)) +
+  geom_line() +
+  geom_point(size = 3) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = fundiv_r_squared_compare$resolution) + 
+  labs(title = "Comparison of FDis R-squared Values by Resolution",
+       x = "Resolution (km)",
+       y = "R-squared Value",
+       color = "Model Type") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = "inside",
+        legend.position.inside = c(.9, .2),
+        legend.box.margin = margin(10, 10, 10, 10)))
+
+# Calculate R-squared values for linear model
+fundiv_lm_linear <- lm(fundiv_r_squared_linear ~ resolution, data = fundiv_r_squared_compare)
+fundiv_r_squared_linear <- summary(fundiv_lm_linear)$r.squared
+
+# Calculate R-squared values for quadratic model
+fundiv_lm_quad <- lm(fundiv_r_squared_quad ~ resolution, data = fundiv_r_squared_compare)
+fundiv_r_squared_quad <- summary(fundiv_lm_quad)$r.squared
+
+(fundiv_r_squared_plot_trendlines <- ggplot() +
+  geom_smooth(aes(x = fundiv_r_squared_compare$resolution, y = fundiv_r_squared_compare$fundiv_r_squared_linear), 
+              method = "lm", se = FALSE, color = "red", linetype = "solid") +
+  geom_smooth(aes(x = fundiv_r_squared_compare$resolution, y = fundiv_r_squared_compare$fundiv_r_squared_quad), 
+              method = "lm", se = FALSE, color = "blue", linetype = "solid") +
+  geom_point(aes(x = fundiv_r_squared_long$resolution, y = fundiv_r_squared_long$fundiv_r_squared, color = fundiv_r_squared_long$model),
+             size = 3) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = fundiv_r_squared_compare$resolution) + 
+  labs(title = "Comparison of FDis R-squared Values by Resolution",
+       x = "Resolution (km)",
+       y = "R-squared Value",
+       color = "Model Type") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = "inside",
+        legend.position.inside = c(.9, .2),
+        legend.box.margin = margin(10, 10, 10, 10)) +
+  annotate("text", x = 70, y = 0.2, label = paste("R² =", round(fundiv_r_squared_linear, 2)), color = "red", size = 4) +
+  annotate("text", x = 70, y = 0.6, label = paste("R² =", round(fundiv_r_squared_quad, 2)), color = "blue", size = 4))
+```
+
+# Compare slope coefficents by resolution
+## Taxonomic diversity
+```{r}
+# Create a tibble
+taxdiv_slopecoefficent_compare <- tibble(
+  resolution = c(5, 10, 25, 50, 75, 100),
+  taxdiv_slopecoefficent_linear = c(
+    as.numeric(compare_Richness_slopecoefficent_5km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_10km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_25km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_50km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_75km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_100km_linear$a)
+  ),
+  taxdiv_slopecoefficent_exp = c(
+    as.numeric(compare_Richness_slopecoefficent_5km_exp$b), 
+    as.numeric(compare_Richness_slopecoefficent_10km_exp$b), 
+    as.numeric(compare_Richness_slopecoefficent_25km_exp$b), 
+    as.numeric(compare_Richness_slopecoefficent_50km_exp$b), 
+    as.numeric(compare_Richness_slopecoefficent_75km_exp$b), 
+    as.numeric(compare_Richness_slopecoefficent_100km_exp$b)
+  )
+)
+
+# reshape to long format for ease of plotting
+taxdiv_slopecoefficent_long <- taxdiv_slopecoefficent_compare %>%
+  pivot_longer(cols = c(taxdiv_slopecoefficent_linear, taxdiv_slopecoefficent_exp), 
+               names_to = "model", 
+               values_to = "slope_coefficent")
+```
+
+```{r}
+# Define custom colors for the points
+custom_colors <- c(taxdiv_slopecoefficent_linear = "red", taxdiv_slopecoefficent_exp = "purple")
+
+# Define custom legend labels
+custom_labels <- c(taxdiv_slopecoefficent_linear = "Linear Model", taxdiv_slopecoefficent_exp = "Exponential Model")
+
+(taxdiv_slopecoefficent_plot <- ggplot(taxdiv_slopecoefficent_long,aes(x = resolution, y = slope_coefficent, color = model)) +
+  geom_line() +
+  geom_point(size = 3) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = taxdiv_slopecoefficent_compare$resolution) + 
+  labs(title = "Comparison of Richness Slope Coefficents by Resolution",
+       x = "Resolution (km)",
+       y = "Slope Coefficent",
+       color = "Model Type") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = "inside",
+        legend.position.inside = c(.85, .3),
+        legend.box.margin = margin(10, 10, 10, 10)))
+```
+
+## Functional diversity
+```{r}
+# Create a tibble
+fundiv_slopecoefficent_compare <- tibble(
+  resolution = c(5, 10, 25, 50, 75, 100),
+  fundiv_slopecoefficent_linear = c(
+    as.numeric(compare_FDis_slopecoefficent_5km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_10km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_25km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_50km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_75km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_100km_linear$a)
+  ),
+  fundiv_slopecoefficent_quad = c(
+    as.numeric(compare_FDis_slopecoefficent_5km_quad$b), 
+    as.numeric(compare_FDis_slopecoefficent_10km_quad$b), 
+    as.numeric(compare_FDis_slopecoefficent_25km_quad$b), 
+    as.numeric(compare_FDis_slopecoefficent_50km_quad$b), 
+    as.numeric(compare_FDis_slopecoefficent_75km_quad$b), 
+    as.numeric(compare_FDis_slopecoefficent_100km_quad$b)
+  )
+)
+
+# reshape to long format for ease of plotting
+fundiv_slopecoefficent_long <- fundiv_slopecoefficent_compare %>%
+  pivot_longer(cols = c(fundiv_slopecoefficent_linear, fundiv_slopecoefficent_quad), 
+               names_to = "model", 
+               values_to = "slope_coefficent")
+```
+
+```{r}
+# Define custom colors for the points
+custom_colors <- c(fundiv_slopecoefficent_linear = "red", fundiv_slopecoefficent_quad = "blue")
+
+# Define custom legend labels
+custom_labels <- c(fundiv_slopecoefficent_linear = "Linear Model", fundiv_slopecoefficent_quad = "Quadratic Model")
+
+(fundiv_slopecoefficent_plot <- ggplot(fundiv_slopecoefficent_long,aes(x = resolution, y = slope_coefficent, color = model)) +
+  geom_line() +
+  geom_point(size = 3) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = fundiv_slopecoefficent_compare$resolution) + 
+  labs(title = "Comparison of FDis Slope Coefficents by Resolution",
+       x = "Resolution (km)",
+       y = "Slope Coefficent",
+       color = "Model Type") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = "inside",
+        legend.position.inside = c(.85, .2),
+        legend.box.margin = margin(10, 10, 10, 10)))
+```
+
+## Compare the linear model slope coefficents for taxonomic and functional diversity
+```{r}
+# Create a tibble
+bothdiv_slopecoefficent_compare <- tibble(
+  resolution = c(5, 10, 25, 50, 75, 100),
+  taxdiv_slopecoefficent_linear = c(
+    as.numeric(compare_Richness_slopecoefficent_5km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_10km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_25km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_50km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_75km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_100km_linear$a)
+  ),
+  fundiv_slopecoefficent_linear = c(
+    as.numeric(compare_FDis_slopecoefficent_5km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_10km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_25km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_50km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_75km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_100km_linear$a)
+  )
+)
+
+# reshape to long format for ease of plotting
+bothdiv_slopecoefficent_long <- bothdiv_slopecoefficent_compare %>%
+  pivot_longer(cols = c(taxdiv_slopecoefficent_linear, fundiv_slopecoefficent_linear), 
+               names_to = "type", 
+               values_to = "slope_coefficent")
+```
+
+```{r}
+# Define custom colors for the points
+custom_colors <- c(taxdiv_slopecoefficent_linear = "goldenrod", fundiv_slopecoefficent_linear = "forestgreen")
+
+# Define custom legend labels
+custom_labels <- c(taxdiv_slopecoefficent_linear = "Taxonomic Diversity", fundiv_slopecoefficent_linear = "Functional Diversity")
+
+(bothdiv_slopecoefficent_plot <- ggplot(bothdiv_slopecoefficent_long,aes(x = resolution, y = slope_coefficent, color = type)) +
+  geom_line() +
+  geom_point(size = 3) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = bothdiv_slopecoefficent_compare$resolution) + 
+  labs(x = "Resolution (km)",
+       y = "Slope Coefficent",
+       color = "Diversity Measure") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = "inside",
+        legend.position.inside = c(.85, .2),
+        legend.box.margin = margin(10, 10, 10, 10)))
+```
+
+# Save plots to file
+```{r}
+ggsave("compare_Richness_all_plots.png", compare_Richness_all_plots, path = figure_path, width = 10, height = 8, units = "in")
+ggsave("compare_FDis_all_plots.png", compare_FDis_all_plots, path = figure_path, width = 10, height = 8, units = "in")
+```
+
+```{r}
+ggsave("taxdiv_r_squared_plot.png", taxdiv_r_squared_plot, path = figure_path)
+ggsave("taxdiv_r_squared_plot_trendlines.png", taxdiv_r_squared_plot_trendlines, path = figure_path)
+ggsave("fundiv_r_squared_plot.png", fundiv_r_squared_plot, path = figure_path)
+ggsave("fundiv_r_squared_plot_trendlines.png", fundiv_r_squared_plot_trendlines, path = figure_path)
+```
+
+```{r}
+ggsave("taxdiv_slopecoefficent_plot.png", taxdiv_slopecoefficent_plot, path = figure_path)
+ggsave("fundiv_slopecoefficent_plot.png", fundiv_slopecoefficent_plot, path = figure_path)
+ggsave("bothdiv_slopecoefficent_plot.png", bothdiv_slopecoefficent_plot, path = figure_path)
+```
+
+# Compare same tax taxanomic and functional diversity
+```{r}
+same_taxa_diversity <- function(richness_grid, fdis_grid, resolution_meters, z){
+  # Extract cell values
+  comparison_diversity <- data.frame(Richness = richness_grid$num_species,
+                                    FDis = fdis_grid$fdis_value)
+
+  # Filter out rows where either richness or FDis is zero
+  comparison_diversity_filtered <- comparison_diversity %>%
+                                filter(Richness > 0 & FDis > 0)
+
+  ## Create scatter plots
+  (compare_diversity_plot_line2P <- ggtrendline(comparison_diversity_filtered$Richness,
+               comparison_diversity_filtered$FDis, model = z, linewidth = 1, CI.level = NA, eSize = 4) +
+      geom_point(aes(x = comparison_diversity_filtered$Richness,
+                                           y = comparison_diversity_filtered$FDis)) +
+      labs(title = paste("Richness vs FDis [", resolution_meters/1000, " km]", sep = ""),
+       x = "Richness by cell", y = "FDis by cell"))
+
+}
+```
+
+```{r}
+same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "line2P")
+same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "line3P")
+same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "log2P")
+same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "exp2P")
+same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "exp3P")
+same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "power2P")
+same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "power3P")
+```
+
+```{r}
+same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "line2P")
+same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "line3P")
+same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "log2P")
+same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "exp2P")
+same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "exp3P")
+same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "power2P")
+same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "power3P")
+```

### 2024-06-24 code/Manuscript_figures.rmd
modify summary figures, try to generate sankey diagram but unsucessful

--- 2024-05-10 code/Manuscript_figures.rmd
@@
-Summary plots
+---
+title: "Manuscript Figures"
+author: "Hazel J. Anderson"
+project: "Plant-Frugivore Diversity"
+collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
+overview: "This script generates misc figures for manuscript that corresponds to this project."
+date: "2024-05-14"
+output: 
+    html_document:
+      code_link: TRUE
+---
+
+# Set file paths
+```{r}
+figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
+```
+
 # Load required packages
@@
 library(tidyr)
-# devtools::install_github("pedropark99/ggfunnel")
-library(ggfunnel)
 library(ggpubr)
+library(networkD3)
+#devtools::install_github("davidsjoberg/ggsankey")
+library(ggsankey)
+```
+
+# Creating cleaning occurrence data figures
+
+These numbers are recorded from the summary function at the end of most L0 and L1 scripts. A table is available in the manuscript supplemental materials.
+
+## Frugivores
+```{r}
+Step <- c("Raw", "After cleaning", "Full trait coverage")
+frug_records <- c(8885330, 1721825, 1667892)
+frug_species <- c(1360, 1219, 1074)
+frug_genera <- c(400, 398, 370)
+frug_families <- c(66, 66, 64)
+
+frug_summary_df <- data.frame(Step, frug_records, frug_species, frug_genera, frug_families)
+
+# Reshape data to long format suitable for plotting
+frug_summary_long <- frug_summary_df %>%
+  pivot_longer(cols = starts_with("frug"), names_to = "Category", values_to = "Value")
 ```
@@
 ```{r}
-Step <- c("Raw", "After cleaning", "Full trait coverage")
-frug_records <- c(8089513, 1721825, 1667892)
-frug_species <- c(1222, 1219, 1074)
-frug_genera <- c(400, 398, 370)
-frug_families <- c(66, 66, 64)
+# Define the order of steps with "Raw" at the top
+step_order <- c("Full trait coverage", "After cleaning", "Raw")
 
-frug_records_df <- data.frame(Step, frug_records)
-frug_species_df <- data.frame(Step, frug_species)
-frug_genera_df <- data.frame(Step, frug_genera)
-frug_families_df <- data.frame(Step, frug_families)
+# Convert Step to factor with defined order
+frug_summary_long$Step <- factor(frug_summary_long$Step, levels = step_order)
+
+# Replace Category labels with more descriptive names
+frug_summary_long <- frug_summary_long %>%
+  mutate(Category = case_when(
+    Category == "frug_records"   ~ "Records",
+    Category == "frug_species"   ~ "Species",
+    Category == "frug_genera"    ~ "Genera",
+    Category == "frug_families"  ~ "Families",
+    TRUE                         ~ as.character(Category)  # Handle any other cases
+  ))
+
+# Define the order of categories
+category_order <- c("Records", "Species", "Genera", "Families")
+
+# Convert Category to factor with defined order
+frug_summary_long$Category <- factor(frug_summary_long$Category, levels = category_order)
+
 ```
@@
 ```{r}
-ggplot(frug_records_df, aes(x = Step, y = frug_records)) +
-  geom_bar(stat = "identity") +
-  coord_flip() +
-  geom_text(aes(label = frug_records), color = "white") + 
-  theme(axis.title.x = element_blank(), axis.text.x = element_blank(),axis.ticks.x = element_blank(), axis.title.y = element_blank())
+(frugivore_summary_plot <- ggplot(frug_summary_long, aes(y = Step, x = Value, fill = Category)) +
+  geom_bar(stat = "identity", position = "stack") +
+  facet_wrap(. ~ Category, scales = "free_x") +
+  labs(title = "Frugivore Data Summary",
+       x = "Value",
+       y = NULL) +
+  theme_minimal() +
+  theme(strip.background = element_blank(),  # Remove background from facet labels
+        strip.text = element_text(size = 12, face = "bold"),
+        legend.position = "none"))
+ggsave("frugivore_summary_plot.png", frugivore_summary_plot, path = figure_path)
 ```
+
+## Plants
+
 ```{r}
-frug_records_fig <- ggfunnel::funnel(frug_records_df, levels = Step, values = frug_records, text_specs = list(colour = "white", size = 3)) +
-  labs(title = "Frugivore records", y = NULL)
-frug_records_fig
+Step <- c("Raw", "After cleaning", "Full trait coverage")
+plant_records <- c(4519792, 1886120, 1700044)
+plant_species <- c(40288, 39972, 33746)
+plant_genera <- c(4083, 3793, 3400)
+plant_families <- c(307, 293, 275)
+
+plant_summary_df <- data.frame(Step, plant_records, plant_species, plant_genera, plant_families)
+
+# Reshape data to long format suitable for plotting
+plant_summary_long <- plant_summary_df %>%
+  pivot_longer(cols = starts_with("plant"), names_to = "Category", values_to = "Value")
 ```
@@
 ```{r}
-frug_species_fig <- ggfunnel::funnel(frug_species_df, levels = Step, values = frug_species, text_specs = list(colour = "white", size = 3)) +
-  labs(title = "Frugivore species", y = NULL)
-frug_species_fig
+# Define the order of steps with "Raw" at the top
+step_order <- c("Full trait coverage", "After cleaning", "Raw")
+
+# Convert Step to factor with defined order
+plant_summary_long$Step <- factor(plant_summary_long$Step, levels = step_order)
+
+# Replace Category labels with more descriptive names
+plant_summary_long <- frug_summary_long %>%
+  mutate(Category = case_when(
+    Category == "plant_records"   ~ "Records",
+    Category == "plant_species"   ~ "Species",
+    Category == "plant_genera"    ~ "Genera",
+    Category == "plant_families"  ~ "Families",
+    TRUE                         ~ as.character(Category)  # Handle any other cases
+  ))
+
+# Define the order of categories
+category_order <- c("Records", "Species", "Genera", "Families")
+
+# Convert Category to factor with defined order
+plant_summary_long$Category <- factor(plant_summary_long$Category, levels = category_order)
+
 ```
@@
 ```{r}
-frug_genera_fig <- ggfunnel::funnel(frug_genera_df, levels = Step, values = frug_genera, text_specs = list(colour = "white", size = 3)) +
-  labs(title = "Frugivore genera", y = NULL)
-frug_genera_fig
+(plant_summary_plot <- ggplot(plant_summary_long, aes(y = Step, x = Value, fill = Category)) +
+  geom_bar(stat = "identity", position = "stack") +
+  facet_wrap(. ~ Category, scales = "free_x") +
+  labs(title = "Plant Data Summary",
+       x = "Value",
+       y = NULL) +
+  theme_minimal() +
+  theme(strip.background = element_blank(),  # Remove background from facet labels
+        strip.text = element_text(size = 12, face = "bold"),
+        legend.position = "none"))
+ggsave("plant_summary_plot.png", plant_summary_plot, path = figure_path)
+```
+
+# Create Sankey diagram for taxonomic harmonization
+Inspiration figures: https://conbio-onlinelibrary-wiley-com.proxy2.cl.msu.edu/doi/10.1111/csp2.311#csp2311-fig-0003 and https://bsapubs-onlinelibrary-wiley-com.proxy2.cl.msu.edu/doi/full/10.1002/aps3.11575#aps311575-fig-0003
+
+Package used (databases)    Plant species inputted    Matches   Disparities
+TNRS                        487,201                   472,680   14,548
+bdc (COL)                   14,548                    5,177     9,371
+
+*The following code doesn't generate the figure like I want.*
+```{r}
+# Create connection dataframe
+links <- data.frame(
+  source = c("Plant species inputted", "TNRS", "TNRS", "Disparities", "bdc (COL)", "bdc (COL)", "Matches"),
+  target = c("TNRS", "Matches", "Disparities", "bdc (COL)", "Matches", "Species dropped", "Species harmonized"),
+  value = c(487201, 472680, 14548, 14548, 5177, 9371, 476010)
+)
+
+# Create nodes
+nodes <- data.frame(
+  name = c(as.character(links$source),
+           as.character(links$target) %>% unique())
+)
+
+# Reformat connections
+links$IDsource <- match(links$source, nodes$name)-1
+links$IDtarget <- match(links$target, nodes$name)-1
+
+# Create the Sankey diagram
+sankeyNetwork(Links = links, Nodes = nodes, Source = "IDsource", Target = "IDtarget",
+              Value = "value", NodeID = "name")
 ```
@@
 ```{r}
-frug_families_fig <- ggfunnel::funnel(frug_families_df, levels = Step, values = frug_families, text_specs = list(colour = "white", size = 3)) +
-  labs(title = "Frugivore families", y = NULL)
-frug_families_fig
+# Create the data frames for the nodes and links
+sankey_data <- data.frame(
+  x = c("source", "target", "target", "source", "target", "target", "source", "target", "source"),
+  node = c("Plant species inputted", "TNRS", "TNRS", "Disparities", "bdc (COL)", "bdc (COL)", "Species dropped", "Matches", "Species harmonized"),
+  next_x = c("target", "source", "source", "target", "source", "source", NA, "source", NA),
+  next_node = c("TNRS", "Matches", "Disparities", "bdc (COL)", "Matches", "Species dropped", NA, "Species harmonized", NA),
+  value = c(487201, 472680, 14548, 14548, 5177, 9371, 9371, 476010, 476010)
+)
+
 ```
@@
 ```{r}
-figure <- ggarrange(frug_records_fig,frug_species_fig, frug_genera_fig, frug_families_fig,
-                    labels = c("A", "B", "C", "D"),
-                    ncol = 2, nrow = 2)
-figure
+# Create the Sankey diagram
+ggplot(sankey_data, aes(x = x, next_x = next_x, node = node, next_node = next_node, value = value)) +
+  geom_sankey(flow.fill = "darkgrey", node.fill = "darkblue", node.color = "darkblue", alpha = 0.5) +
+  geom_sankey_label(aes(label = node), size = 3, fill = "white") +
+  theme_sankey(base_size = 12) +
+  theme_void() +
+  labs(title = "Plant Species Matching Workflow",
+       subtitle = "Visualization of species input, TNRS and BDC matching process",
+       x = NULL,
+       y = NULL) +
+  theme(legend.position = "none")
 ```
 
+# References
+
 ```{r}
+library(report)
+cite_packages()
+```
+
+```{r, class.source= 'fold-hide'}
 devtools::session_info()

### 2024-06-24 code/L1/L1_LastStep_DiversityInputObjects.Rmd
update point plots to make points smaller and add forest fill behind

@@
   geom_sf(data = TApoly) +
-  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "forestgreen")+
+  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "forestgreen") +
   labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
@@
   geom_sf(data = TApoly) +
-  geom_sf(data = plants_sf_species, pch = 21) +
+  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "forestgreen") + 
+  geom_sf(data = plants_sf_species, pch = 16, alpha = 0.1, size = 0.01) +
   labs(title = "Tropical Andes Plant Occurrences", x = "Latitude", y = "Longitude") +
@@
   annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  annotation_north_arrow(location = "tr", which_north = "true",
+                         height = unit(0.5, "in"), width = unit(0.5, "in"),
+                         pad_x = unit(0.1, "in"), pad_y = unit(0.1, "in"),
+                         style = north_arrow_fancy_orienteering) +
   theme(panel.background = element_rect(fill = "lightblue"))
@@
 plantsPointsPlot
-ggsave("plant_occurrence_points_map.png", plot = last_plot(), path = figure_path)
+#ggsave("plant_occurrence_points_map.png", plot = last_plot(), path = figure_path)
 ```
@@
   geom_sf(data = TApoly) +
-  geom_sf(data = frugivores_sf_species, pch = 21) +
+  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "forestgreen") +
+  geom_sf(data = frugivores_sf_species, pch = 16, alpha = 0.1, size = 0.01) +
   labs(title = "Tropical Andes Frugivore Occurrences", x = "Latitude", y = "Longitude") +
@@
   annotation_scale(location = "bl", width_hint = 0.5) +
-  annotation_north_arrow(location = "bl", which_north = "true", 
-    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
+  annotation_north_arrow(location = "tr", which_north = "true",
+                         height = unit(0.5, "in"), width = unit(0.5, "in"),
+                         pad_x = unit(0.1, "in"), pad_y = unit(0.1, "in"),
+                         style = north_arrow_fancy_orienteering) +
   theme(panel.background = element_rect(fill = "lightblue"))
 frugivoresPointsPlot
-ggsave("frugivore_occurrence_points_map.png", plot = last_plot(), path = figure_path)
+#ggsave("frugivore_occurrence_points_map.png", plot = last_plot(), path = figure_path)
 ```
@@
 both_points_maps
-ggsave("both_points_maps.png", both_points_maps, path = figure_path)
+ggsave("both_points_maps.png", both_points_maps, path = figure_path, height =  7, width = 8, units = "in")
 ```

### 2024-09-03 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
comment out saving individual maps, add histograms, changing summarizing fdis value to mean instead of sum (fixing issue), and other minor fixes

@@
 spatial_fdis_frugivore_grid <- TAGrid_100km %>%
-   st_join(spatial_fdis_frugivore) %>%
+   st_join(spatial_fdis_frugivore, join = st_contains) %>%
    group_by(cellid) %>%
-   summarize(fdis_value = sum(fdis_frugivore, na.rm = TRUE)) %>%
+   summarize(fdis_value = mean(fdis_frugivore, na.rm = TRUE)) %>%
    ungroup()
@@
     ggtitle("100 km") +
-    scale_fill_distiller(palette = "YlOrBr", direction = 1) +
+    scale_fill_distiller(palette = "YlOrBr", direction = 1, limits = c(0,0.8)) +
     coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
@@
 ```{r}
-ggsave("frugivoregridFDisTA_100km.png", frugivoregridFDisTA_100km, path = figure_path)
+#ggsave("frugivoregridFDisTA_100km.png", frugivoregridFDisTA_100km, path = figure_path)
 # Extract cell values
@@
    group_by(cellid) %>%
-   summarize(fdis_value = sum(fdis_frugivore, na.rm = TRUE)) %>%
+   summarize(fdis_value = mean(fdis_frugivore, na.rm = TRUE)) %>%
    ungroup()
@@
     geom_sf(data = Americas, fill = "white")+
-    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
+    geom_sf(data = TApoly, fill = "lightgrey") +
     geom_sf(data = spatial_fdis_frugivore_grid, aes(fill = fdis_value), color = NA) +
     labs(fill = "FDis") +
-    scale_fill_distiller(palette = "YlOrBr", direction = 1) +
+    scale_fill_distiller(palette = "YlOrBr", direction = 1, limits = c(0,0.8)) +
     coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
@@
 )
-ggsave("frugivoregridFDisTA_75km.png", frugivoregridFDisTA_75km, path = figure_path)
+#ggsave("frugivoregridFDisTA_75km.png", frugivoregridFDisTA_75km, path = figure_path)
 ```
@@
 )
-ggsave("frugivoregridFDisTA_50km.png", frugivoregridFDisTA_50km, path = figure_path)
+#ggsave("frugivoregridFDisTA_50km.png", frugivoregridFDisTA_50km, path = figure_path)
 ```
@@
 )
-ggsave("frugivoregridFDisTA_25km.png", frugivoregridFDisTA_25km, path = figure_path)
+#ggsave("frugivoregridFDisTA_25km.png", frugivoregridFDisTA_25km, path = figure_path)
 ```
@@
 )
-ggsave("frugivoregridFDisTA_10km.png", frugivoregridFDisTA_10km, path = figure_path)
-```
-
+#ggsave("frugivoregridFDisTA_10km.png", frugivoregridFDisTA_10km, path = figure_path)
+```
+
+```{r}
+hist(frugivore_cellFDis_75km$fdis_value)
+hist(frugivore_cellFDis_50km$fdis_value)
+hist(frugivore_cellFDis_25km$fdis_value)
+hist(frugivore_cellFDis_10km$fdis_value)
+```
+
+## 5 km
 ```{r}
@@
     group_by(cellid) %>%
-    summarize(fdis_value = sum(fdis, na.rm = TRUE)) %>%
+    summarize(fdis_value = mean(fdis, na.rm = TRUE)) %>%
     ungroup()
@@
     geom_sf(data = Americas, fill = "white")+
-    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
+    geom_sf(data = TApoly, fill = "lightgrey") +
     geom_sf(data = spatial_fdis_frugivore_grid, aes(fill = fdis_value), color = NA) +
     labs(fill = "FDis") +
-    scale_fill_distiller(palette = "YlOrBr", direction = 1) +
+    scale_fill_distiller(palette = "YlOrBr", direction = 1, limits = c(0,0.8)) +
     coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
     theme(panel.background = element_rect(fill = "lightblue"),
-        text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
-  frugivoregridFDisTA_5km
+        text = element_text(size = 12), plot.title = element_text(hjust = 0.5),
+        axis.text.x = element_text(angle = 45, hjust = 1))
 frugivore_cellFDis_5km <- spatial_fdis_frugivore_grid
@@
 )
-ggsave("frugivoregridFDisTA_5km.png", frugivoregridFDisTA_5km, path = figure_path)
+#ggsave("frugivoregridFDisTA_5km.png", frugivoregridFDisTA_5km, path = figure_path)
+hist(frugivore_cellFDis_5km$fdis_value)
 ```
@@
    group_by(cellid) %>%
-   summarize(fdis_value = sum(fdis_plant, na.rm = TRUE)) %>%
+   summarize(fdis_value = mean(fdis_plant, na.rm = TRUE)) %>%
    ungroup()
@@
     ggtitle("100 km") +
-    scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,7.5)) +
+    scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,0.8)) +
     coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
@@
 ```{r}
-ggsave("plantgridFDisTA_100km.png", plantgridFDisTA_100km, path = figure_path)
+#ggsave("plantgridFDisTA_100km.png", plantgridFDisTA_100km, path = figure_path)
 # Extract cell values
@@
 
-  # Use three fewer cores than available
-  num_cores <- detectCores() - 3
-  cl <- makeCluster(num_cores)
-  registerDoParallel(cl)
-
-  on.exit(stopCluster(cl)) # Ensure the cluster is stopped when the function exits
-
-  tryCatch({
   # Generate the grid
@@
 
-    # Check if TAGrid is empty
-    if (nrow(TAGrid) == 0) {
-      stop("Error: No valid grid cells after intersection.")
-    }
-
-    # Divide the indices of the grid into chunks for parallel processing
-    chunk_indices <- split(seq_len(nrow(TAGrid)), rep(1:num_cores, each = ceiling(nrow(TAGrid) / num_cores), length.out = nrow(TAGrid)))
-
-    # Process chunks in parallel
-    results_list <- foreach(indices = chunk_indices, .combine = rbind, .packages = c("sf", "dplyr")) %dopar% {
-      chunk <- TAGrid[indices, ]
-
-      # Perform the intersection
-      intersected <- st_intersection(spatial_fdis_plant, chunk)
-
-      # Calculate cellid for intersected
-      intersected$cellid <- st_intersects(intersected, chunk)$nn
-
-      # Ensure that intersected is a data frame
-      as.data.frame(intersected)
-    }
-
-    # Combine all chunk results into a single data frame
-    spatial_fdis_plant <- do.call(rbind, results_list)
-
-    # Ensure the combined result is a data frame
-    spatial_fdis_plant <- as.data.frame(spatial_fdis_plant)
-
-    # Debugging: Check if 'cellid' column exists
-    if (!"cellid" %in% colnames(spatial_fdis_plant)) {
-      stop("Error: 'cellid' column not found in spatial_fdis_plant.")
-    }
-
-    # Summarize fdis values per grid cell
-    spatial_fdis_plant_grid <- spatial_fdis_plant %>%
+  spatial_fdis_plant$cellid <- st_intersects(spatial_fdis_plant, TAGrid)$nn
+
+  spatial_fdis_plant_grid <- TAGrid %>%
+   st_join(spatial_fdis_plant) %>%
    group_by(cellid) %>%
-      summarize(fdis_value = sum(fdis_plant, na.rm = TRUE)) %>%
-      right_join(as.data.frame(TAGrid), by = "cellid")
-
-    # Convert the grid to sf
-    spatial_fdis_plant_grid <- st_as_sf(spatial_fdis_plant_grid)
+   summarize(fdis_value = mean(fdis_plant, na.rm = TRUE)) %>%
+   ungroup()
 
@@
       geom_sf(data = Americas, fill = "white") +
-      geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
+      geom_sf(data = TApoly, fill = "lightgrey") +
       geom_sf(data = spatial_fdis_plant_grid, aes(fill = fdis_value), color = NA) +
       labs(fill = "FDis") +
-      scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0, 7.5)) +
+      scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,0.8)) +
       coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
@@
     list(plantgridFDisTA = plantgridFDisTA, spatial_fdis_plant_grid = spatial_fdis_plant_grid)
-  }, error = function(e) {
-    message("Error in parallel processing: ", conditionMessage(e))
-    NULL
-  })
   }
@@
 )
-ggsave("plantgridFDisTA_75km.png", plantgridFDisTA_75km, path = figure_path)
+#ggsave("plantgridFDisTA_75km.png", plantgridFDisTA_75km, path = figure_path)
 ```
@@
 )
-ggsave("plantgridFDisTA_50km.png", plantgridFDisTA_50km, path = figure_path)
+#ggsave("plantgridFDisTA_50km.png", plantgridFDisTA_50km, path = figure_path)
 ```
@@
 )
-ggsave("plantgridFDisTA_25km.png", plantgridFDisTA_25km, path = figure_path)
+#ggsave("plantgridFDisTA_25km.png", plantgridFDisTA_25km, path = figure_path)
 ```
@@
 )
-ggsave("plantgridFDisTA_10km.png", plantgridFDisTA_10km, path = figure_path)
-```
-
+#ggsave("plantgridFDisTA_10km.png", plantgridFDisTA_10km, path = figure_path)
+```
+
+```{r}
+hist(plant_cellFDis_75km$fdis_value)
+hist(plant_cellFDis_50km$fdis_value)
+hist(plant_cellFDis_25km$fdis_value)
+hist(plant_cellFDis_10km$fdis_value)
+```
+
+## 5 km 
 ```{r}
@@
 # Determine number of cores to use
-num_cores <- detectCores() - 2
+num_cores <- detectCores() - 3
 
@@
   # Perform spatial join
-  joined <- st_join(grid_chunk, spatial_fdis_plant)
+  joined <- st_join(grid_chunk, spatial_fdis_plant_5km)
 
@@
     group_by(cellid) %>%
-    summarize(fdis_value = sum(fdis_plant, na.rm = TRUE)) %>%
+    summarize(fdis_value = mean(fdis, na.rm = TRUE)) %>%
     ungroup()
@@
 
-  # Debugging: Check the structure of the summarized data frame
-  message("Processed chunk: ", indices[1], " to ", indices[length(indices)])
-  str(summarized_df)
-
-  list(summarized_df)  # Return as list to avoid unintended simplification
+  list(as.data.frame(summarized_df))
 }
@@
 if ("geometry" %in% colnames(flattened_results)) {
-  spatial_fdis_plant_grid <- st_as_sf(flattened_results)
+  spatial_fdis_plant_grid_5km <- st_as_sf(flattened_results)
 } else {
-  spatial_fdis_plant_grid <- flattened_results
+  spatial_fdis_plant_grid_5km <- flattened_results
 }
@@
 # Check the structure of the final output
-str(spatial_fdis_plant_grid)
+str(spatial_fdis_plant_grid_5km)
 
@@
     geom_sf(data = Americas, fill = "white")+
-    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
-    geom_sf(data = spatial_fdis_plant_grid, aes(fill = fdis_value), color = NA) +
+    geom_sf(data = TApoly, fill = "lightgrey") +
+    geom_sf(data = spatial_fdis_plant_grid_5km, aes(fill = fdis_value), color = NA) +
     labs(fill = "FDis") +
-    ggtitle("5 km") +
-    scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,6.5)) +
+    scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,0.8)) +
     coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
@@
         text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
-plantgridFDisTA_5km
-```
-
-```{r}
-plant_cellFDis_5km <- spatial_fdis_plant_grid
+
+plant_cellFDis_5km <- spatial_fdis_plant_grid_5km
 ```
@@
 )
-ggsave("plantgridFDisTA_5km.png", plantgridFDisTA_5km, path = figure_path)
+#ggsave("plantgridFDisTA_5km.png", plantgridFDisTA_5km, path = figure_path)
+```
+
+```{r}
+hist(plant_cellFDis_5km$fdis_value)
 ```
@@
 plant_plot_legend <- function(plot){
-  plot + scale_fill_distiller(palette = "Greens", direction = 1,limits = c(0, 6.5),
-                              oob = scales::squish, breaks = seq(0, 6.5, by = 1)) +
+  plot + scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,0.8),
+                              oob = scales::squish, breaks = seq(0, 0.8, by = 0.2)) +
     theme(legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10)) +
@@
 frugivore_plot_legend <- function(plot){
-  plot + scale_fill_distiller(palette = "YlOrBr", direction = 1,limits = c(0, 8.5),
-                              oob = scales::squish, breaks = seq(0, 8.5, by = 1)) +
+  plot + scale_fill_distiller(palette = "YlOrBr", direction = 1, limits = c(0,0.8),
+                              oob = scales::squish, breaks = seq(0, 0.8, by = 0.2)) +
     theme(legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10)) +

### 2024-09-03 code/Manuscript_figures.rmd
change figure colors, combine figures into images for manuscripts

@@
 
+# Define custom colors
+frug_colors <- c("Records" = "darkgoldenrod", "Species" = "darkgoldenrod3", "Genera" = "darkgoldenrod2", "Families" = "darkgoldenrod1")
+
 ```
@@
   geom_bar(stat = "identity", position = "stack") +
-  facet_wrap(. ~ Category, scales = "free_x") +
-  labs(title = "Frugivore Data Summary",
+  facet_wrap(. ~ Category, scales = "free_x", nrow = 1) +
+  scale_fill_manual(values = frug_colors) +
+  labs(title = "Frugivore",
        x = "Value",
@@
         strip.text = element_text(size = 12, face = "bold"),
-        legend.position = "none"))
+        legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1)))
 ggsave("frugivore_summary_plot.png", frugivore_summary_plot, path = figure_path)
@@
 Step <- c("Raw", "After cleaning", "Full trait coverage")
-plant_records <- c(4519792, 1886120, 1700044)
-plant_species <- c(40288, 39972, 33746)
-plant_genera <- c(4083, 3793, 3400)
-plant_families <- c(307, 293, 275)
+plant_records <- c(4519792, 1886120, 658609)
+plant_species <- c(40288, 39972, 3128)
+plant_genera <- c(4083, 3793, 1359)
+plant_families <- c(307, 293, 213)
 
@@
 # Replace Category labels with more descriptive names
-plant_summary_long <- frug_summary_long %>%
+plant_summary_long <- plant_summary_long %>%
   mutate(Category = case_when(
@@
 
+# Define custom colors
+plant_colors <- c("Records" = "palegreen4", "Species" = "palegreen3", "Genera" = "palegreen2", "Families" = "palegreen1")
+
 ```
@@
   geom_bar(stat = "identity", position = "stack") +
-  facet_wrap(. ~ Category, scales = "free_x") +
-  labs(title = "Plant Data Summary",
+  facet_wrap(. ~ Category, scales = "free_x", nrow = 1) +
+  scale_fill_manual(values = plant_colors) +
+  labs(title = "Plant",
        x = "Value",
@@
         strip.text = element_text(size = 12, face = "bold"),
-        legend.position = "none"))
+        legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1)))
 ggsave("plant_summary_plot.png", plant_summary_plot, path = figure_path)
+```
+
+## Combine data summary figures
+```{r}
+(both_summary_plots <- ggarrange(plant_summary_plot, frugivore_summary_plot,
+                                        ncol = 1, nrow = 2))
+ggsave("both_summary_plots.png", plot = both_summary_plots, path = figure_path, height = 4, width = 10, units = "in")
+```
+
+# Create spatial grain maps
+```{r}
+library(rnaturalearth)
+library(sf)
+```
+
+```{r}
+data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
+TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
+```
+
+```{r}
+worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
+Americas <- ne_countries(continent = c("North America", "South America"), returnclass = "sf")
+#polygon of Tropical Andes
+TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
+Ecuador <- worldMap %>% filter(sovereignt == "Ecuador")
+```
+
+```{r}
+# Check and set CRS
+# Assuming your data is initially in WGS 84 (EPSG:4326), reproject to a suitable projected CRS
+TApoly <- st_transform(TApoly, crs = 32718) # UTM Zone 18S for the Andes region
+TropicalAndes_IUCNHabitat_Forest <- st_transform(TropicalAndes_IUCNHabitat_Forest, crs = 32718)
+Ecuador <- st_transform(Ecuador, crs = 32718)
+Americas <- st_transform(Americas, crs = 32718)
+```
+
+Make & map grid for TA
+```{r}
+TAGrid_5km <- TApoly %>%
+    st_make_grid(cellsize = c(5000, 5000)) %>%
+    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+    st_cast("MULTIPOLYGON") %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+```
+
+```{r}
+(TAGrid_5km_plot <- ggplot(data = TAGrid_5km) +
+    geom_sf(data = Americas) +
+    geom_sf(data = TAGrid_5km) +
+    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+    theme_minimal() +
+    labs(title = "5 km") +
+    theme(plot.title = element_text(hjust = 0.5)))
+```
+
+```{r}
+TAGrid_10km <- TApoly %>%
+    st_make_grid(cellsize = c(10000, 10000)) %>%
+    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+    st_cast("MULTIPOLYGON") %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+```
+
+```{r}
+(TAGrid_10km_plot <- ggplot(data = TAGrid_10km) +
+    geom_sf(data = Americas) +
+    geom_sf(data = TAGrid_10km) +
+    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+    theme_minimal() +
+    labs(title = "10 km") +
+    theme(plot.title = element_text(hjust = 0.5)))
+```
+
+```{r}
+TAGrid_25km <- TApoly %>%
+    st_make_grid(cellsize = c(25000, 25000)) %>%
+    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+    st_cast("MULTIPOLYGON") %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+```
+
+```{r}
+(TAGrid_25km_plot <- ggplot(data = TAGrid_25km) +
+    geom_sf(data = Americas) +
+    geom_sf(data = TAGrid_25km) +
+    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+    theme_minimal() +
+    labs(title = "25 km") +
+    theme(plot.title = element_text(hjust = 0.5)))
+```
+
+```{r}
+TAGrid_50km <- TApoly %>%
+    st_make_grid(cellsize = c(50000, 50000)) %>%
+    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+    st_cast("MULTIPOLYGON") %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+```
+
+```{r}
+(TAGrid_50km_plot <- ggplot(data = TAGrid_50km) +
+    geom_sf(data = Americas) +
+    geom_sf(data = TAGrid_50km) +
+    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+    theme_minimal() +
+    labs(title = "50 km") +
+    theme(plot.title = element_text(hjust = 0.5)))
+```
+
+```{r}
+TAGrid_75km <- TApoly %>%
+    st_make_grid(cellsize = c(75000, 75000)) %>%
+    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+    st_cast("MULTIPOLYGON") %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+```
+
+```{r}
+(TAGrid_75km_plot <- ggplot(data = TAGrid_75km) +
+    geom_sf(data = Americas) +
+    geom_sf(data = TAGrid_75km) +
+    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+    theme_minimal() +
+    labs(title = "75 km") +
+    theme(plot.title = element_text(hjust = 0.5)))
+```
+
+```{r}
+TAGrid_100km <- TApoly %>%
+    st_make_grid(cellsize = c(100000, 100000)) %>%
+    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
+    st_cast("MULTIPOLYGON") %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+```
+
+```{r}
+(TAGrid_100km_plot <- ggplot(data = TAGrid_100km) +
+    geom_sf(data = Americas) +
+    geom_sf(data = TAGrid_100km) +
+    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
+    theme_minimal() +
+    labs(title = "100 km") +
+    theme(plot.title = element_text(hjust = 0.5)))
+```
+
+Combine grid plots into one figure
+```{r}
+library(ggpubr)
+(TAGrid_all_plots <- ggpubr::ggarrange(TAGrid_5km_plot, TAGrid_10km_plot, TAGrid_25km_plot,
+                  TAGrid_50km_plot, TAGrid_75km_plot, TAGrid_100km_plot,
+                  ncol = 6, nrow = 1))
+ggsave("TAGrid_all_plots.png", TAGrid_all_plots, path = figure_path, width = 16, height = 10, units = "in")
+```
+
+```{r}
+resolution <- c(5, 10, 25, 50, 75, 100)
+num_cells <- c(length(TAGrid_5km$cellid), length(TAGrid_10km$cellid),length(TAGrid_25km$cellid),
+               length(TAGrid_50km$cellid), length(TAGrid_75km$cellid),length(TAGrid_100km$cellid))
+plot(resolution, num_cells)
+
+```
+
+Make & map grid for just Ecuador
+```{r}
+EcGrid_5km <- TApoly %>%
+    st_make_grid(cellsize = c(5000, 5000)) %>%
+    st_intersection(Ecuador) %>%
+    st_cast("MULTIPOLYGON") %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+```
+
+```{r}
+(EcGrid_5km_plot <- ggplot(data = EcGrid_5km) +
+    geom_sf() +
+    theme_minimal() +
+    labs(title = "5 km") +
+    coord_sf(xlim = c(-81.5, -75), ylim = c(-5.5, 2), expand = FALSE, crs = 4326))
+(Grid_5km_plot <- ggplot(data = EcGrid_5km) +
+    geom_sf() +
+    theme_minimal() +
+    labs(title = "5 km") +
+    theme(axis.text = element_blank())+
+    coord_sf(xlim = c(-79.5, -77.3), ylim = c(-2.5, -0.5), expand = FALSE, crs = 4326))
+```
+
+```{r}
+EcGrid_10km <- TApoly %>%
+    st_make_grid(cellsize = c(10000, 10000)) %>%
+    st_intersection(Ecuador) %>%
+    st_cast("MULTIPOLYGON") %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+```
+
+```{r}
+(EcGrid_10km_plot <- ggplot(data = EcGrid_10km) +
+    geom_sf() +
+    theme_minimal() +
+    labs(title = "10 km") +
+    coord_sf(xlim = c(-81.5, -75), ylim = c(-5.5, 2), expand = FALSE, crs = 4326))
+(Grid_10km_plot <- ggplot(data = EcGrid_10km) +
+    geom_sf() +
+    theme_minimal() +
+    labs(title = "10 km") +
+    theme(axis.text = element_blank())+
+    coord_sf(xlim = c(-79.5, -77.3), ylim = c(-2.5, -0.5), expand = FALSE, crs = 4326))
+```
+
+```{r}
+EcGrid_25km <- TApoly %>%
+    st_make_grid(cellsize = c(25000, 25000)) %>%
+    st_intersection(Ecuador) %>%
+    st_cast("MULTIPOLYGON") %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+```
+
+```{r}
+(EcGrid_25km_plot <- ggplot(data = EcGrid_25km) +
+    geom_sf() +
+    theme_minimal() +
+    labs(title = "25 km") +
+    coord_sf(xlim = c(-81.5, -75), ylim = c(-5.5, 2), expand = FALSE, crs = 4326))
+(Grid_25km_plot <- ggplot(data = EcGrid_25km) +
+    geom_sf() +
+    theme_minimal() +
+    labs(title = "25 km") +
+    theme(axis.text = element_blank())+
+    coord_sf(xlim = c(-79.5, -77.3), ylim = c(-2.5, -0.5), expand = FALSE, crs = 4326))
+```
+
+```{r}
+EcGrid_50km <- TApoly %>%
+    st_make_grid(cellsize = c(50000, 50000)) %>%
+    st_intersection(Ecuador) %>%
+    st_cast("MULTIPOLYGON") %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+```
+
+```{r}
+(EcGrid_50km_plot <- ggplot(data = EcGrid_50km) +
+    geom_sf() +
+    theme_minimal() +
+    labs(title = "50 km") +
+    coord_sf(xlim = c(-81.5, -75), ylim = c(-5.5, 2), expand = FALSE, crs = 4326))
+(Grid_50km_plot <- ggplot(data = EcGrid_50km) +
+    geom_sf() +
+    theme_minimal() +
+    labs(title = "50 km") +
+    theme(axis.text = element_blank())+ 
+    coord_sf(xlim = c(-79.5, -77.3), ylim = c(-2.5, -0.5), expand = FALSE, crs = 4326))
+```
+
+```{r}
+EcGrid_75km <- TApoly %>%
+    st_make_grid(cellsize = c(75000, 75000)) %>%
+    st_intersection(Ecuador) %>%
+    st_cast("MULTIPOLYGON") %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+```
+
+```{r}
+(EcGrid_75km_plot <- ggplot(data = EcGrid_75km) +
+    geom_sf() +
+    theme_minimal() +
+    labs(title = "75 km") +
+    coord_sf(xlim = c(-81.5, -75), ylim = c(-5.5, 2), expand = FALSE, crs = 4326))
+(Grid_75km_plot <- ggplot(data = EcGrid_75km) +
+    geom_sf() +
+    theme_minimal() +
+    labs(title = "75 km") +
+    theme(axis.text = element_blank())+
+    coord_sf(xlim = c(-79.5, -77.3), ylim = c(-2.5, -0.5), expand = FALSE, crs = 4326))
+```
+
+```{r}
+EcGrid_100km <- TApoly %>%
+    st_make_grid(cellsize = c(100000, 100000)) %>%
+    st_intersection(Ecuador) %>%
+    st_cast("MULTIPOLYGON") %>%
+    st_sf() %>%
+    mutate(cellid = row_number())
+```
+
+```{r}
+(EcGrid_100km_plot <- ggplot(data = EcGrid_100km) +
+    geom_sf() +
+    theme_minimal() +
+    labs(title = "100 km") +
+    coord_sf(xlim = c(-81.5, -75), ylim = c(-5.5, 2), expand = FALSE, crs = 4326))
+(Grid_100km_plot <- ggplot(data = EcGrid_100km) +
+    geom_sf() +
+    theme_minimal() +
+    labs(title = "100 km") +
+    theme(axis.text = element_blank())+
+    coord_sf(xlim = c(-79.5, -77.3), ylim = c(-2.5, -0.5), expand = FALSE, crs = 4326))
+```
+
+Combine Ecuador grid plots into one figure
+```{r}
+library(ggpubr)
+(Grid_all_plots <- ggpubr::ggarrange(Grid_5km_plot, Grid_10km_plot, Grid_25km_plot,
+                  Grid_50km_plot, Grid_75km_plot, Grid_100km_plot,
+                  ncol = 6, nrow = 1))
+ggsave("Grid_all_plots.png", Grid_all_plots, path = figure_path, width = 16, height = 10, units = "in")
 ```

### 2024-09-03 code/L1/L1_LastStep_DiversityInputObjects.Rmd
update colors and change plot title, and other minor changes

@@
   geom_sf(data = TApoly) +
-  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "forestgreen") +
-  labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
+  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "#D9F7D9") +
+  labs(title = "Tropical Andes Forest", x = "Latitude", y = "Longitude") +
   coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_scale(location = "bl", width_hint = 0.3, style = "ticks") +
   annotation_north_arrow(location = "bl", which_north = "true", 
@@
 basePlot
-ggsave("tropical_andes_forest_map.png", plot = last_plot(), path = figure_path)
+#ggsave("tropical_andes_forest_map.png", plot = last_plot(), path = figure_path)
 ```
@@
   geom_sf(data = TApoly) +
-  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "forestgreen") + 
-  geom_sf(data = plants_sf_species, pch = 16, alpha = 0.1, size = 0.01) +
-  labs(title = "Tropical Andes Plant Occurrences", x = "Latitude", y = "Longitude") +
+  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "#D9F7D9") + 
+  geom_sf(data = plants_sf_species, pch = 16, size = 0.05) +
+  labs(title = "Fruiting Plant Occurrences", x = "Latitude", y = "Longitude") +
   coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_scale(location = "bl",width_hint = 0.3, style = "ticks") +
   annotation_north_arrow(location = "tr", which_north = "true",
@@
   geom_sf(data = TApoly) +
-  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "forestgreen") +
-  geom_sf(data = frugivores_sf_species, pch = 16, alpha = 0.1, size = 0.01) +
-  labs(title = "Tropical Andes Frugivore Occurrences", x = "Latitude", y = "Longitude") +
+  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "#D9F7D9") +
+  geom_sf(data = frugivores_sf_species, pch = 16, size = 0.01) +
+  labs(title = "Frugivore Occurrences", x = "Latitude", y = "Longitude") +
   coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
-  annotation_scale(location = "bl", width_hint = 0.5) +
+  annotation_scale(location = "bl",width_hint = 0.3, style = "ticks") +
   annotation_north_arrow(location = "tr", which_north = "true",

### 2024-09-03 code/L1/L1_P4_TropicalAndes_plant_traits.Rmd
change plot colors and other plot customizations

@@
 
-# Create the faceted bar plot with free y-axis scales
+# Customize fill color of database
+custom_colors <- c("BIEN" = "#a6cf3c" , "GIFT" = "#009900", "TRY" = "#40a353")
+
+# Create the faceted bar plot
 ggplot(trait_counts, aes(x = DatabaseSource, y = n, fill = DatabaseSource)) +
@@
        fill = "Database Source") +
+  scale_fill_manual(values = custom_colors) +
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  facet_wrap(~ TraitName, scales = "free_y")
-ggsave("plant_trait_counts_per_database.png", plot = last_plot(), path = figure_path)
+  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
+        legend.position = "none") +
+  facet_wrap(~ TraitName, scales = "free_y", nrow = 2)
+ggsave("plant_trait_counts_per_database.png", plot = last_plot(), path = figure_path, width = 12, height = 5)
 ```

### 2024-09-05 code/L2/L2_3_TropicalAndes_divRelationships.Rmd
add code of determining model types, update plant-frugivore TD or FD plotting code to get labels centered and visible, long multi-plot code, add code for TD-FD relationships

--- 2024-06-03 code/L2/L2_3_TropicalAndes_divRelationships.Rmd
@@
 ```{r}
-# Fit the nonlinear model using nls
-nls_model <- nls(Frugivore_Richness ~ b * exp(c * Plant_Richness), 
-                 data = comparison_data_filtered_100km, 
-                 start = list(b = 200, c = -0.00001),
-                 algorithm = "port",
-                 control = nls.control(maxiter = 500, minFactor = 1e-10, warnOnly = TRUE))
-
-# Summary of the model
-summary(nls_model)
+plot(plant_cellRichness_5km$num_species, plant_cellFDis_5km$fdis_value)
+plot(plant_cellRichness_10km$num_species, plant_cellFDis_10km$fdis_value)
+plot(plant_cellRichness_25km$num_species, plant_cellFDis_25km$fdis_value)
+plot(plant_cellRichness_50km$num_species, plant_cellFDis_50km$fdis_value)
+plot(plant_cellRichness_75km$num_species, plant_cellFDis_75km$fdis_value)
+plot(plant_cellRichness_100km$num_species, plant_cellFDis_100km$fdis_value)
 ```
 
-# Compare r-squared by resolution
+```{r}
+plot(frugivore_cellRichness_5km$num_species, frugivore_cellFDis_5km$fdis_value)
+plot(frugivore_cellRichness_10km$num_species, frugivore_cellFDis_10km$fdis_value)
+plot(frugivore_cellRichness_25km$num_species, frugivore_cellFDis_25km$fdis_value)
+plot(frugivore_cellRichness_50km$num_species, frugivore_cellFDis_50km$fdis_value)
+plot(frugivore_cellRichness_75km$num_species, frugivore_cellFDis_75km$fdis_value)
+plot(frugivore_cellRichness_100km$num_species, frugivore_cellFDis_100km$fdis_value)
+```
+
+```{r}
+plot(plant_cellRichness_5km$num_species, frugivore_cellFDis_5km$fdis_value)
+plot(plant_cellRichness_10km$num_species, frugivore_cellFDis_10km$fdis_value)
+plot(plant_cellRichness_25km$num_species, frugivore_cellFDis_25km$fdis_value)
+plot(plant_cellRichness_50km$num_species, frugivore_cellFDis_50km$fdis_value)
+plot(plant_cellRichness_75km$num_species, frugivore_cellFDis_75km$fdis_value)
+plot(plant_cellRichness_100km$num_species,frugivore_cellFDis_100km$fdis_value)
+```
+
+```{r}
+plot(frugivore_cellRichness_5km$num_species, plant_cellFDis_5km$fdis_value)
+plot(frugivore_cellRichness_10km$num_species, plant_cellFDis_10km$fdis_value)
+plot(frugivore_cellRichness_25km$num_species, plant_cellFDis_25km$fdis_value)
+plot(frugivore_cellRichness_50km$num_species, plant_cellFDis_50km$fdis_value)
+plot(frugivore_cellRichness_75km$num_species, plant_cellFDis_75km$fdis_value)
+plot(frugivore_cellRichness_100km$num_species, plant_cellFDis_100km$fdis_value)
+```
+
+# Compare diversity relationships between plants and frugivores
+## Determine model types
+### Richness
+```{r}
+diff_taxa_richness <- function(plant_Richness_grid, frugivore_Richness_grid, resolution_meters, z){
+  # Extract cell values
+  comparison_Richness <- data.frame(Plant_Richness = plant_Richness_grid$num_species,
+                               Frugivore_Richness = frugivore_Richness_grid$num_species)
+
+  # Filter out rows where either plant or frugivore richness is zero
+  comparison_Richness_filtered <- comparison_Richness %>%
+                                filter(Plant_Richness > 0 & Frugivore_Richness > 0)
+
+  ## Create scatter plots
+  (compare_Richness_plot <- ggtrendline(comparison_Richness_filtered$Plant_Richness,
+               comparison_Richness_filtered$Frugivore_Richness, model = z, linewidth = 1, CI.level = NA, eSize = 4) +
+      geom_point(aes(x = comparison_Richness_filtered$Plant_Richness,
+                                           y = comparison_Richness_filtered$Frugivore_Richness)) +
+      labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
+       x = "Plant Richness by cell", y = "Frugivore Richness by cell"))
+
+}
+
+```
+
+```{r}
+diff_taxa_richness(plant_cellRichness_100km, frugivore_cellRichness_100km, 100000, "line2P")
+diff_taxa_richness(plant_cellRichness_100km, frugivore_cellRichness_100km, 100000, "line3P")
+diff_taxa_richness(plant_cellRichness_100km, frugivore_cellRichness_100km, 100000, "log2P")
+diff_taxa_richness(plant_cellRichness_100km, frugivore_cellRichness_100km, 100000, "exp2P")
+diff_taxa_richness(plant_cellRichness_100km, frugivore_cellRichness_100km, 100000, "exp3P")
+diff_taxa_richness(plant_cellRichness_100km, frugivore_cellRichness_100km, 100000, "power2P")
+diff_taxa_richness(plant_cellRichness_100km, frugivore_cellRichness_100km, 100000, "power3P")
+```
+
+```{r}
+diff_taxa_richness(plant_cellRichness_75km, frugivore_cellRichness_75km, 75000, "line2P")
+diff_taxa_richness(plant_cellRichness_75km, frugivore_cellRichness_75km, 75000, "line3P")
+diff_taxa_richness(plant_cellRichness_75km, frugivore_cellRichness_75km, 75000, "log2P")
+diff_taxa_richness(plant_cellRichness_75km, frugivore_cellRichness_75km, 75000, "exp2P")
+diff_taxa_richness(plant_cellRichness_75km, frugivore_cellRichness_75km, 75000, "exp3P")
+diff_taxa_richness(plant_cellRichness_75km, frugivore_cellRichness_75km, 75000, "power2P")
+diff_taxa_richness(plant_cellRichness_75km, frugivore_cellRichness_75km, 75000, "power3P")
+```
+
+```{r}
+diff_taxa_richness(plant_cellRichness_50km, frugivore_cellRichness_50km, 50000, "line2P")
+diff_taxa_richness(plant_cellRichness_50km, frugivore_cellRichness_50km, 50000, "line3P")
+diff_taxa_richness(plant_cellRichness_50km, frugivore_cellRichness_50km, 50000, "log2P")
+diff_taxa_richness(plant_cellRichness_50km, frugivore_cellRichness_50km, 50000, "exp2P")
+diff_taxa_richness(plant_cellRichness_50km, frugivore_cellRichness_50km, 50000, "exp3P")
+diff_taxa_richness(plant_cellRichness_50km, frugivore_cellRichness_50km, 50000, "power2P")
+diff_taxa_richness(plant_cellRichness_50km, frugivore_cellRichness_50km, 50000, "power3P")
+```
+
+```{r}
+diff_taxa_richness(plant_cellRichness_25km, frugivore_cellRichness_25km, 25000, "line2P")
+diff_taxa_richness(plant_cellRichness_25km, frugivore_cellRichness_25km, 25000, "line3P")
+diff_taxa_richness(plant_cellRichness_25km, frugivore_cellRichness_25km, 25000, "log2P")
+diff_taxa_richness(plant_cellRichness_25km, frugivore_cellRichness_25km, 25000, "exp2P")
+diff_taxa_richness(plant_cellRichness_25km, frugivore_cellRichness_25km, 25000, "exp3P")
+diff_taxa_richness(plant_cellRichness_25km, frugivore_cellRichness_25km, 25000, "power2P")
+diff_taxa_richness(plant_cellRichness_25km, frugivore_cellRichness_25km, 25000, "power3P")
+```
+
+```{r}
+diff_taxa_richness(plant_cellRichness_10km, frugivore_cellRichness_10km, 10000, "line2P")
+diff_taxa_richness(plant_cellRichness_10km, frugivore_cellRichness_10km, 10000, "line3P")
+diff_taxa_richness(plant_cellRichness_10km, frugivore_cellRichness_10km, 10000, "log2P")
+diff_taxa_richness(plant_cellRichness_10km, frugivore_cellRichness_10km, 10000, "exp2P")
+diff_taxa_richness(plant_cellRichness_10km, frugivore_cellRichness_10km, 10000, "exp3P")
+diff_taxa_richness(plant_cellRichness_10km, frugivore_cellRichness_10km, 10000, "power2P")
+diff_taxa_richness(plant_cellRichness_10km, frugivore_cellRichness_10km, 10000, "power3P")
+```
+
+```{r}
+diff_taxa_richness(plant_cellRichness_5km, frugivore_cellRichness_5km, 5000, "line2P")
+diff_taxa_richness(plant_cellRichness_5km, frugivore_cellRichness_5km, 5000, "line3P")
+diff_taxa_richness(plant_cellRichness_5km, frugivore_cellRichness_5km, 5000, "log2P")
+diff_taxa_richness(plant_cellRichness_5km, frugivore_cellRichness_5km, 5000, "exp2P")
+diff_taxa_richness(plant_cellRichness_5km, frugivore_cellRichness_5km, 5000, "exp3P")
+diff_taxa_richness(plant_cellRichness_5km, frugivore_cellRichness_5km, 5000, "power2P")
+diff_taxa_richness(plant_cellRichness_5km, frugivore_cellRichness_5km, 5000, "power3P")
+```
+
+### FDis
+```{r}
+diff_taxa_FDis <- function(plant_FDis_grid, frugivore_FDis_grid, resolution_meters, z){
+  # Extract cell values
+  comparison_FDis <- data.frame(Plant_FDis = plant_FDis_grid$fdis_value,
+                               Frugivore_FDis = frugivore_FDis_grid$fdis_value)
+
+  # Filter out rows where either plant or frugivore FDis is zero
+  comparison_FDis_filtered <- comparison_FDis %>%
+                                filter(Plant_FDis > 0 & Frugivore_FDis > 0)
+
+  ## Create scatter plots
+  (compare_diversity_plot <- ggtrendline(comparison_FDis_filtered$Plant_FDis,
+               comparison_FDis_filtered$Frugivore_FDis, model = z, linewidth = 1, CI.level = NA, eSize = 4) +
+      geom_point(aes(x = comparison_FDis_filtered$Plant_FDis,
+                                           y = comparison_FDis_filtered$Frugivore_FDis)) +
+      labs(title = paste("Plant_FDis vs Frugivore_FDis [", resolution_meters/1000, " km]", sep = ""),
+       x = "Plant_FDis by cell", y = "Frugivore_FDis by cell"))
+
+}
+```
+
+```{r}
+diff_taxa_FDis(plant_cellFDis_100km, frugivore_cellFDis_100km, 100000, "line2P")
+diff_taxa_FDis(plant_cellFDis_100km, frugivore_cellFDis_100km, 100000, "line3P")
+diff_taxa_FDis(plant_cellFDis_100km, frugivore_cellFDis_100km, 100000, "log2P")
+diff_taxa_FDis(plant_cellFDis_100km, frugivore_cellFDis_100km, 100000, "exp2P")
+#diff_taxa_FDis(plant_cellFDis_100km, frugivore_cellFDis_100km, 100000, "exp3P")
+diff_taxa_FDis(plant_cellFDis_100km, frugivore_cellFDis_100km, 100000, "power2P")
+#diff_taxa_FDis(plant_cellFDis_100km, frugivore_cellFDis_100km, 100000, "power3P")
+```
+
+```{r}
+diff_taxa_FDis(plant_cellFDis_75km, frugivore_cellFDis_75km, 75000, "line2P")
+diff_taxa_FDis(plant_cellFDis_75km, frugivore_cellFDis_75km, 75000, "line3P")
+diff_taxa_FDis(plant_cellFDis_75km, frugivore_cellFDis_75km, 75000, "log2P")
+diff_taxa_FDis(plant_cellFDis_75km, frugivore_cellFDis_75km, 75000, "exp2P")
+#diff_taxa_FDis(plant_cellFDis_75km, frugivore_cellFDis_75km, 75000, "exp3P")
+diff_taxa_FDis(plant_cellFDis_75km, frugivore_cellFDis_75km, 75000, "power2P")
+diff_taxa_FDis(plant_cellFDis_75km, frugivore_cellFDis_75km, 75000, "power3P")
+```
+
+```{r}
+diff_taxa_FDis(plant_cellFDis_50km, frugivore_cellFDis_50km, 50000, "line2P")
+diff_taxa_FDis(plant_cellFDis_50km, frugivore_cellFDis_50km, 50000, "line3P")
+diff_taxa_FDis(plant_cellFDis_50km, frugivore_cellFDis_50km, 50000, "log2P")
+diff_taxa_FDis(plant_cellFDis_50km, frugivore_cellFDis_50km, 50000, "exp2P")
+diff_taxa_FDis(plant_cellFDis_50km, frugivore_cellFDis_50km, 50000, "exp3P")
+diff_taxa_FDis(plant_cellFDis_50km, frugivore_cellFDis_50km, 50000, "power2P")
+diff_taxa_FDis(plant_cellFDis_50km, frugivore_cellFDis_50km, 50000, "power3P")
+```
+
+```{r}
+diff_taxa_FDis(plant_cellFDis_25km, frugivore_cellFDis_25km, 25000, "line2P")
+diff_taxa_FDis(plant_cellFDis_25km, frugivore_cellFDis_25km, 25000, "line3P")
+diff_taxa_FDis(plant_cellFDis_25km, frugivore_cellFDis_25km, 25000, "log2P")
+diff_taxa_FDis(plant_cellFDis_25km, frugivore_cellFDis_25km, 25000, "exp2P")
+diff_taxa_FDis(plant_cellFDis_25km, frugivore_cellFDis_25km, 25000, "exp3P")
+diff_taxa_FDis(plant_cellFDis_25km, frugivore_cellFDis_25km, 25000, "power2P")
+diff_taxa_FDis(plant_cellFDis_25km, frugivore_cellFDis_25km, 25000, "power3P")
+```
+
+```{r}
+diff_taxa_FDis(plant_cellFDis_10km, frugivore_cellFDis_10km, 10000, "line2P")
+diff_taxa_FDis(plant_cellFDis_10km, frugivore_cellFDis_10km, 10000, "line3P")
+diff_taxa_FDis(plant_cellFDis_10km, frugivore_cellFDis_10km, 10000, "log2P")
+diff_taxa_FDis(plant_cellFDis_10km, frugivore_cellFDis_10km, 10000, "exp2P")
+#diff_taxa_FDis(plant_cellFDis_10km, frugivore_cellFDis_10km, 10000, "exp3P")
+diff_taxa_FDis(plant_cellFDis_10km, frugivore_cellFDis_10km, 10000, "power2P")
+diff_taxa_FDis(plant_cellFDis_10km, frugivore_cellFDis_10km, 10000, "power3P")
+```
+
+```{r}
+diff_taxa_FDis(plant_cellFDis_5km, frugivore_cellFDis_5km, 5000, "line2P")
+diff_taxa_FDis(plant_cellFDis_5km, frugivore_cellFDis_5km, 5000, "line3P")
+diff_taxa_FDis(plant_cellFDis_5km, frugivore_cellFDis_5km, 5000, "log2P")
+diff_taxa_FDis(plant_cellFDis_5km, frugivore_cellFDis_5km, 5000, "exp2P")
+diff_taxa_FDis(plant_cellFDis_5km, frugivore_cellFDis_5km, 5000, "exp3P")
+diff_taxa_FDis(plant_cellFDis_5km, frugivore_cellFDis_5km, 5000, "power2P")
+diff_taxa_FDis(plant_cellFDis_5km, frugivore_cellFDis_5km, 5000, "power3P")
+```
+
+## Function for comparing diversity relationships plots
+For comparing richness, use exp3P and line2P for line types
+For comparing FDis, use line3P and line2P for line types
+
+line colors: line2P is "red", line3P is "blue", exp3P is "purple"
+```{r}
+compare_Richness <- function(plant_Richness_grid, frugivore_Richness_grid, resolution_meters, custom_y1, custom_y2){
+  # Extract cell values
+  comparison_Richness <- data.frame(Plant_Richness = plant_Richness_grid$num_species,
+                               Frugivore_Richness = frugivore_Richness_grid$num_species)
+
+  # Filter out rows where either plant or frugivore richness is zero
+  comparison_Richness_filtered <- comparison_Richness %>%
+                                filter(Plant_Richness > 0 & Frugivore_Richness > 0)
+
+  # Calculate Annotation position
+  mean_x <- max(comparison_Richness_filtered$Plant_Richness) / 2
+  annot_y <- max(comparison_Richness_filtered$Frugivore_Richness) + 25
+
+  # Create scatter plots
+  compare_Richness_plot_line2P <- ggtrendline(comparison_Richness_filtered$Plant_Richness,
+               comparison_Richness_filtered$Frugivore_Richness, model = "line2P", linecolor = "red", linewidth = 1,
+               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, rrp.x = mean_x, rrp.y = annot_y, 
+               text.col = "red", eSize = 4) +
+      geom_point(aes(x = comparison_Richness_filtered$Plant_Richness,
+                                           y = comparison_Richness_filtered$Frugivore_Richness), alpha = 0.25, size = 1) +
+      labs(title = paste("[", resolution_meters/1000, " km]", sep = ""),
+           subtitle = "Fitted with a linear model",
+       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
+      scale_x_continuous(expand = c(0, 0)) + 
+      scale_y_continuous(expand = c(0, 0)) + 
+      theme_minimal() +
+      theme(panel.grid.major = element_blank(), 
+            panel.grid.minor = element_blank(),
+            axis.line = element_line(color = "black"),
+            axis.ticks = element_line(color = "black"),
+            axis.text = element_text(color = "black"))
+
+  compare_Richness_plot_exp3P <- ggtrendline(comparison_Richness_filtered$Plant_Richness,
+               comparison_Richness_filtered$Frugivore_Richness, model = "exp3P", linecolor = "purple", linewidth = 1,
+               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, rrp.x = mean_x, rrp.y = annot_y, 
+               text.col = "purple", eSize = 4) +
+      geom_point(aes(x = comparison_Richness_filtered$Plant_Richness,
+                                           y = comparison_Richness_filtered$Frugivore_Richness), alpha = 0.25, size = 1) +
+      labs(title = paste("[", resolution_meters/1000, " km]", sep = ""),
+           subtitle = "Fitted with an exponential model",
+       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
+      scale_x_continuous(expand = c(0, 0)) + 
+      scale_y_continuous(expand = c(0, 0)) + 
+      theme_minimal() +
+      theme(panel.grid.major = element_blank(), 
+            panel.grid.minor = element_blank(),
+            axis.line = element_line(color = "black"),
+            axis.ticks = element_line(color = "black"),
+            axis.text = element_text(color = "black"))
+
+    # Summaries of models used for trendlines
+  line2P_summary <- trendline_sum(comparison_Richness_filtered$Plant_Richness,
+               comparison_Richness_filtered$Frugivore_Richness, model = "line2P")
+  exp3P_summary <- trendline_sum(comparison_Richness_filtered$Plant_Richness,
+               comparison_Richness_filtered$Frugivore_Richness, model = "exp3P")
+
+  # Combine trendlines into one plot
+  compare_Richness_plot_both <- ggtrendline(comparison_Richness_filtered$Plant_Richness,
+               comparison_Richness_filtered$Frugivore_Richness, model = "exp3P", linecolor = "purple", linewidth = 1,
+               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, show.Rsquare = FALSE) +
+      geom_smooth(aes(x = comparison_Richness_filtered$Plant_Richness, y = comparison_Richness_filtered$Frugivore_Richness), 
+              method = "lm", se = FALSE, color = "red", linetype = "solid") +
+      geom_point(aes(x = comparison_Richness_filtered$Plant_Richness,
+                                           y = comparison_Richness_filtered$Frugivore_Richness), alpha = 0.25, size = 1) +
+      labs(title = paste("[", resolution_meters/1000, " km]", sep = ""),
+       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
+      scale_x_continuous(expand = c(0, 0)) + 
+      scale_y_continuous(expand = c(0, 0)) + 
+      theme_minimal() +
+      theme(panel.grid.major = element_blank(), 
+            panel.grid.minor = element_blank(),
+            axis.line = element_line(color = "black"),
+            axis.ticks = element_line(color = "black"),
+            axis.text = element_text(color = "black"),
+            text = element_text(size = 18)) +
+      annotate("text", x = mean_x, y = custom_y1, label = paste("Linear R² =", round(line2P_summary$R.squared, 5)),
+               color = "red", size = 6) +
+      annotate("text", x = mean_x, y = custom_y2, label = paste("Exponential R² =", round(exp3P_summary$R.squared, 5)),
+               color = "purple", size = 6)
+
+  list(compare_Richness_plot_line2P = compare_Richness_plot_line2P,
+       compare_Richness_plot_exp3P = compare_Richness_plot_exp3P,
+       line2P_summary = line2P_summary,
+       exp3P_summary = exp3P_summary,
+       compare_Richness_plot_both = compare_Richness_plot_both)
+}
+```
+
+```{r}
+compare_FDis <- function(plant_FDis_grid, frugivore_FDis_grid, resolution_meters, custom_y1, custom_y2, custom_x){
+  # Extract cell values
+  comparison_FDis <- data.frame(Plant_FDis = plant_FDis_grid$fdis_value,
+                               Frugivore_FDis = frugivore_FDis_grid$fdis_value)
+
+  # Filter out rows where either plant or frugivore FDis is zero
+  comparison_FDis_filtered <- comparison_FDis %>%
+                                filter(Plant_FDis > 0 & Frugivore_FDis > 0)
+  ## Create scatter plots
+  compare_FDis_plot_line2P <- ggtrendline(comparison_FDis_filtered$Plant_FDis,
+               comparison_FDis_filtered$Frugivore_FDis, model = "line2P", linecolor = "red", linewidth = 1,
+               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, 
+               text.col = "red", eSize = 4) +
+      geom_point(aes(x = comparison_FDis_filtered$Plant_FDis,
+                                           y = comparison_FDis_filtered$Frugivore_FDis), alpha = 0.25, size = 1) +
+      labs(title = paste("Compare FDis [", resolution_meters/1000, " km]", sep = ""),
+           subtitle = "Fitted with a linear model",
+       x = "Plant FDis by cell", y = "Frugivore FDis by cell") +
+      scale_x_continuous(expand = c(0, 0)) + 
+      scale_y_continuous(expand = c(0, 0)) + 
+      theme_minimal() +
+      theme(panel.grid.major = element_blank(), 
+            panel.grid.minor = element_blank(),
+            axis.line = element_line(color = "black"),
+            axis.ticks = element_line(color = "black"),
+            axis.text = element_text(color = "black"))
+
+  compare_FDis_plot_line3P <- ggtrendline(comparison_FDis_filtered$Plant_FDis,
+               comparison_FDis_filtered$Frugivore_FDis, model = "line3P", linecolor = "blue", linewidth = 1,
+               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, 
+               text.col = "blue", eSize = 4) +
+      geom_point(aes(x = comparison_FDis_filtered$Plant_FDis,
+                                           y = comparison_FDis_filtered$Frugivore_FDis), alpha = 0.25, size = 1) +
+      labs(title = paste("Compare FDis [", resolution_meters/1000, " km]", sep = ""),
+           subtitle = "Fitted with a quadratic model",
+       x = "Plant FDis by cell", y = "Frugivore FDis by cell") +
+      scale_x_continuous(expand = c(0, 0)) + 
+      scale_y_continuous(expand = c(0, 0)) + 
+      theme_minimal() +
+      theme(panel.grid.major = element_blank(), 
+            panel.grid.minor = element_blank(),
+            axis.line = element_line(color = "black"),
+            axis.ticks = element_line(color = "black"),
+            axis.text = element_text(color = "black"))
+
+  # Summaries of models used for trendlines
+  line2P_summary <- trendline_sum(comparison_FDis_filtered$Plant_FDis,
+               comparison_FDis_filtered$Frugivore_FDis, model = "line2P")
+  line3P_summary <- trendline_sum(comparison_FDis_filtered$Plant_FDis,
+               comparison_FDis_filtered$Frugivore_FDis, model = "line3P")
+
+  # Combine trendlines into one plot
+  compare_FDis_plot_both <- ggtrendline(comparison_FDis_filtered$Plant_FDis,
+               comparison_FDis_filtered$Frugivore_FDis, model = "line3P", linecolor = "blue", linewidth = 1,
+               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, show.Rsquare = FALSE) +
+      geom_smooth(aes(x = comparison_FDis_filtered$Plant_FDis, y = comparison_FDis_filtered$Frugivore_FDis), 
+              method = "lm", se = FALSE, color = "red", linetype = "solid") +
+      geom_point(aes(x = comparison_FDis_filtered$Plant_FDis,
+                                           y = comparison_FDis_filtered$Frugivore_FDis), alpha = 0.25, size = 1) +
+      labs(title = paste("[", resolution_meters/1000, " km]", sep = ""),
+       x = "Plant FDis by cell", y = "Frugivore FDis by cell") +
+      scale_x_continuous(expand = c(0, 0)) + 
+      scale_y_continuous(expand = c(0, 0)) + 
+      theme_minimal() +
+      theme(panel.grid.major = element_blank(), 
+            panel.grid.minor = element_blank(),
+            axis.line = element_line(color = "black"),
+            axis.ticks = element_line(color = "black"),
+            axis.text = element_text(color = "black"),
+            text = element_text(size = 18)) +
+      annotate("text", x = custom_x, y = custom_y1, label = paste("Linear R² =", round(line2P_summary$R.squared, 5)),
+               color = "red", size = 6) +
+      annotate("text", x = custom_x, y = custom_y2, label = paste("Quadratic R² =", round(line3P_summary$R.squared, 5)),
+               color = "blue", size = 6)
+
+  list(compare_FDis_plot_line2P = compare_FDis_plot_line2P,
+       compare_FDis_plot_line3P = compare_FDis_plot_line3P,
+       line2P_summary = line2P_summary,
+       line3P_summary = line3P_summary,
+       compare_FDis_plot_both = compare_FDis_plot_both)
+}
+```
+
+## 5 km
+```{r}
+compare_Richness_5km <- compare_Richness(plant_cellRichness_5km, frugivore_cellRichness_5km, 5000, 275, 255)
+(compare_Richness_5km_linear <- compare_Richness_5km$compare_Richness_plot_line2P)
+(compare_Richness_5km_exp <- compare_Richness_5km$compare_Richness_plot_exp3P)
+(compare_Richness_5km_plots <- compare_Richness_5km$compare_Richness_plot_both)
+compare_Richness_r2_5km_linear <- compare_Richness_5km$line2P_summary$R.squared
+compare_Richness_r2_5km_exp <- compare_Richness_5km$exp3P_summary$R.squared
+compare_Richness_slopecoefficent_5km_linear <- compare_Richness_5km$line2P_summary$parameter["a"]
+compare_Richness_slopecoefficent_5km_exp <- compare_Richness_5km$exp3P_summary$parameter["b"]
+```
+
+```{r}
+compare_FDis_5km <- compare_FDis(plant_cellFDis_5km, frugivore_cellFDis_5km, 5000, 0.8, 0.75, 0.15)
+(compare_FDis_5km_linear <- compare_FDis_5km$compare_FDis_plot_line2P)
+(compare_FDis_5km_quad <- compare_FDis_5km$compare_FDis_plot_line3P)
+(compare_FDis_5km_plots <- compare_FDis_5km$compare_FDis_plot_both)
+compare_FDis_r2_5km_linear <- compare_FDis_5km$line2P_summary$R.squared
+compare_FDis_r2_5km_quad <- compare_FDis_5km$line3P_summary$R.squared
+compare_FDis_slopecoefficent_5km_linear <- compare_FDis_5km$line2P_summary$parameter["a"]
+compare_FDis_slopecoefficent_5km_quad <- compare_FDis_5km$line3P_summary$parameter["b"]
+```
+
+## 10 km
+```{r}
+compare_Richness_10km <- compare_Richness(plant_cellRichness_10km, frugivore_cellRichness_10km, 10000, 400, 370)
+(compare_Richness_10km_linear <- compare_Richness_10km$compare_Richness_plot_line2P)
+(compare_Richness_10km_exp <- compare_Richness_10km$compare_Richness_plot_exp3P)
+(compare_Richness_10km_plots <- compare_Richness_10km$compare_Richness_plot_both)
+compare_Richness_r2_10km_linear <- compare_Richness_10km$line2P_summary$R.squared
+compare_Richness_r2_10km_exp <- compare_Richness_10km$exp3P_summary$R.squared
+compare_Richness_slopecoefficent_10km_linear <- compare_Richness_10km$line2P_summary$parameter["a"]
+compare_Richness_slopecoefficent_10km_exp <- compare_Richness_10km$exp3P_summary$parameter["b"]
+```
+
+```{r}
+compare_FDis_10km <- compare_FDis(plant_cellFDis_10km, frugivore_cellFDis_10km, 10000, 0.8, 0.75, 0.15)
+(compare_FDis_10km_linear <- compare_FDis_10km$compare_FDis_plot_line2P)
+(compare_FDis_10km_quad <- compare_FDis_10km$compare_FDis_plot_line3P)
+(compare_FDis_10km_plots <- compare_FDis_10km$compare_FDis_plot_both)
+compare_FDis_r2_10km_linear <- compare_FDis_10km$line2P_summary$R.squared
+compare_FDis_r2_10km_quad <- compare_FDis_10km$line3P_summary$R.squared
+compare_FDis_slopecoefficent_10km_linear <- compare_FDis_10km$line2P_summary$parameter["a"]
+compare_FDis_slopecoefficent_10km_quad <- compare_FDis_10km$line3P_summary$parameter["b"]
+```
+
+## 25 km
+```{r}
+compare_Richness_25km <- compare_Richness(plant_cellRichness_25km, frugivore_cellRichness_25km, 25000, 475, 440)
+(compare_Richness_25km_linear <- compare_Richness_25km$compare_Richness_plot_line2P)
+(compare_Richness_25km_exp <- compare_Richness_25km$compare_Richness_plot_exp3P)
+(compare_Richness_25km_plots <- compare_Richness_25km$compare_Richness_plot_both)
+compare_Richness_r2_25km_linear <- compare_Richness_25km$line2P_summary$R.squared
+compare_Richness_r2_25km_exp <- compare_Richness_25km$exp3P_summary$R.squared
+compare_Richness_slopecoefficent_25km_linear <- compare_Richness_25km$line2P_summary$parameter["a"]
+compare_Richness_slopecoefficent_25km_exp <- compare_Richness_25km$exp3P_summary$parameter["b"]
+```
+
+```{r}
+compare_FDis_25km <- compare_FDis(plant_cellFDis_25km, frugivore_cellFDis_25km, 25000, 0.7, 0.65, 0.18)
+(compare_FDis_25km_linear <- compare_FDis_25km$compare_FDis_plot_line2P)
+(compare_FDis_25km_quad <- compare_FDis_25km$compare_FDis_plot_line3P)
+(compare_FDis_25km_plots <- compare_FDis_25km$compare_FDis_plot_both)
+compare_FDis_r2_25km_linear <- compare_FDis_25km$line2P_summary$R.squared
+compare_FDis_r2_25km_quad <- compare_FDis_25km$line3P_summary$R.squared
+compare_FDis_slopecoefficent_25km_linear <- compare_FDis_25km$line2P_summary$parameter["a"]
+compare_FDis_slopecoefficent_25km_quad <- compare_FDis_25km$line3P_summary$parameter["b"]
+```
+
+## 50 km
+```{r}
+compare_Richness_50km <- compare_Richness(plant_cellRichness_50km, frugivore_cellRichness_50km, 50000, 500, 465)
+(compare_Richness_50km_linear <- compare_Richness_50km$compare_Richness_plot_line2P)
+(compare_Richness_50km_exp <- compare_Richness_50km$compare_Richness_plot_exp3P)
+(compare_Richness_50km_plots <- compare_Richness_50km$compare_Richness_plot_both)
+compare_Richness_r2_50km_linear <- compare_Richness_50km$line2P_summary$R.squared
+compare_Richness_r2_50km_exp <- compare_Richness_50km$exp3P_summary$R.squared
+compare_Richness_slopecoefficent_50km_linear <- compare_Richness_50km$line2P_summary$parameter["a"]
+compare_Richness_slopecoefficent_50km_exp <- compare_Richness_50km$exp3P_summary$parameter["b"]
+```
+
+```{r}
+compare_FDis_50km <- compare_FDis(plant_cellFDis_50km, frugivore_cellFDis_50km, 50000, 0.7, 0.65, 0.34)
+(compare_FDis_50km_linear <- compare_FDis_50km$compare_FDis_plot_line2P)
+(compare_FDis_50km_quad <- compare_FDis_50km$compare_FDis_plot_line3P)
+(compare_FDis_50km_plots <- compare_FDis_50km$compare_FDis_plot_both)
+compare_FDis_r2_50km_linear <- compare_FDis_50km$line2P_summary$R.squared
+compare_FDis_r2_50km_quad <- compare_FDis_50km$line3P_summary$R.squared
+compare_FDis_slopecoefficent_50km_linear <- compare_FDis_50km$line2P_summary$parameter["a"]
+compare_FDis_slopecoefficent_50km_quad <- compare_FDis_50km$line3P_summary$parameter["b"]
+```
+
+## 75 km
+```{r}
+compare_Richness_75km <- compare_Richness(plant_cellRichness_75km, frugivore_cellRichness_75km, 75000, 500, 465)
+(compare_Richness_75km_linear <- compare_Richness_75km$compare_Richness_plot_line2P)
+(compare_Richness_75km_exp <- compare_Richness_75km$compare_Richness_plot_exp3P)
+(compare_Richness_75km_plots <- compare_Richness_75km$compare_Richness_plot_both)
+compare_Richness_r2_75km_linear <- compare_Richness_75km$line2P_summary$R.squared
+compare_Richness_r2_75km_exp <- compare_Richness_75km$exp3P_summary$R.squared
+compare_Richness_slopecoefficent_75km_linear <- compare_Richness_75km$line2P_summary$parameter["a"]
+compare_Richness_slopecoefficent_75km_exp <- compare_Richness_75km$exp3P_summary$parameter["b"]
+```
+
+```{r}
+compare_FDis_75km <- compare_FDis(plant_cellFDis_75km, frugivore_cellFDis_75km, 75000, 0.59, 0.56, 0.25)
+(compare_FDis_75km_linear <- compare_FDis_75km$compare_FDis_plot_line2P)
+(compare_FDis_75km_quad <- compare_FDis_75km$compare_FDis_plot_line3P)
+(compare_FDis_75km_plots <- compare_FDis_75km$compare_FDis_plot_both)
+compare_FDis_r2_75km_linear <- compare_FDis_75km$line2P_summary$R.squared
+compare_FDis_r2_75km_quad <- compare_FDis_75km$line3P_summary$R.squared
+compare_FDis_slopecoefficent_75km_linear <- compare_FDis_75km$line2P_summary$parameter["a"]
+compare_FDis_slopecoefficent_75km_quad <- compare_FDis_75km$line3P_summary$parameter["b"]
+```
+
+## 100 km
+```{r}
+compare_Richness_100km <- compare_Richness(plant_cellRichness_100km, frugivore_cellRichness_100km, 100000, 525, 480)
+(compare_Richness_100km_linear <- compare_Richness_100km$compare_Richness_plot_line2P)
+(compare_Richness_100km_exp <- compare_Richness_100km$compare_Richness_plot_exp3P)
+(compare_Richness_100km_plots <- compare_Richness_100km$compare_Richness_plot_both)
+compare_Richness_r2_100km_linear <- compare_Richness_100km$line2P_summary$R.squared
+compare_Richness_r2_100km_exp <- compare_Richness_100km$exp3P_summary$R.squared
+compare_Richness_slopecoefficent_100km_linear <- compare_Richness_100km$line2P_summary$parameter["a"]
+compare_Richness_slopecoefficent_100km_exp <- compare_Richness_100km$exp3P_summary$parameter["b"]
+```
+
+```{r}
+compare_FDis_100km <- compare_FDis(plant_cellFDis_100km, frugivore_cellFDis_100km, 100000, 0.48, 0.45, 0.37)
+(compare_FDis_100km_linear <- compare_FDis_100km$compare_FDis_plot_line2P)
+(compare_FDis_100km_quad <- compare_FDis_100km$compare_FDis_plot_line3P)
+(compare_FDis_100km_plots <- compare_FDis_100km$compare_FDis_plot_both)
+compare_FDis_r2_100km_linear <- compare_FDis_100km$line2P_summary$R.squared
+compare_FDis_r2_100km_quad <- compare_FDis_100km$line3P_summary$R.squared
+compare_FDis_slopecoefficent_100km_linear <- compare_FDis_100km$line2P_summary$parameter["a"]
+compare_FDis_slopecoefficent_100km_quad <- compare_FDis_100km$line3P_summary$parameter["b"]
+```
+
+## Combine plots using for richness and fdis into one figure
+```{r}
+(compare_Richness_all_plots <- ggarrange(compare_Richness_5km_plots, compare_Richness_10km_plots,
+                                     compare_Richness_25km_plots, compare_Richness_50km_plots,
+                                     compare_Richness_75km_plots, compare_Richness_100km_plots,
+                                        ncol = 3, nrow = 2))
+```
+
+```{r}
+(compare_Richness_all_plots_long <- ggarrange(compare_Richness_5km_plots, compare_Richness_10km_plots,
+                                     compare_Richness_25km_plots, compare_Richness_50km_plots,
+                                     compare_Richness_75km_plots, compare_Richness_100km_plots,
+                                        ncol = 6, nrow = 1))
+```
+
+```{r}
+(compare_FDis_all_plots <- ggarrange(compare_FDis_5km_plots, compare_FDis_10km_plots,
+                                     compare_FDis_25km_plots, compare_FDis_50km_plots,
+                                     compare_FDis_75km_plots, compare_FDis_100km_plots,
+                                        ncol = 3, nrow = 2))
+```
+
+```{r}
+(compare_FDis_all_plots_long <- ggarrange(compare_FDis_5km_plots, compare_FDis_10km_plots,
+                                     compare_FDis_25km_plots, compare_FDis_50km_plots,
+                                     compare_FDis_75km_plots, compare_FDis_100km_plots,
+                                        ncol = 6, nrow = 1))
+```
+
+# Compare R² by resolution
 ## Taxonomic diversity
 ```{r}
-# create dataframe of resolutions and r-squared values 
-#[ADD 1, 5,]
-resolution <- c( 10, 25, 50, 75, 100)
-#ADD compare_richness_r2_1km, compare_richness_r2_5km,
-r_squared <- c(compare_richness_r2_10km, compare_richness_r2_25km, compare_richness_r2_50km, compare_richness_r2_75km, compare_richness_r2_100km)
-r_squared_compare <- data.frame(resolution, r_squared)
+# create dataframe of resolutions and R² values 
+resolution <- c(5, 10, 25, 50, 75, 100)
+taxdiv_r_squared_linear <- c(compare_Richness_r2_5km_linear, compare_Richness_r2_10km_linear, compare_Richness_r2_25km_linear, compare_Richness_r2_50km_linear, compare_Richness_r2_75km_linear, compare_Richness_r2_100km_linear)
+taxdiv_r_squared_exp <- c(compare_Richness_r2_5km_exp, compare_Richness_r2_10km_exp, compare_Richness_r2_25km_exp, compare_Richness_r2_50km_exp, compare_Richness_r2_75km_exp, compare_Richness_r2_100km_exp)
+taxdiv_r_squared_compare <- data.frame(resolution, taxdiv_r_squared_linear, taxdiv_r_squared_exp)
 
-compare_r2_by_res <- ggplot(r_squared_compare, aes(x=resolution, y=r_squared)) +
-    geom_point()+
-    labs(title = "R-squared values of plant-frugivore richness relationships by resolution",
-         x = "Resolution (km)", y = "R-squared", fill = "Number of species")
-print(compare_r2_by_res)
+# reshape to long format for ease of plotting
+taxdiv_r_squared_long <- taxdiv_r_squared_compare %>%
+  pivot_longer(cols = c(taxdiv_r_squared_linear, taxdiv_r_squared_exp), 
+               names_to = "model", 
+               values_to = "taxdiv_r_squared")
+
 ```
+
+```{r}
+# Define custom colors for the points
+custom_colors <- c(taxdiv_r_squared_linear = "red", taxdiv_r_squared_exp = "purple")
+
+# Define custom legend labels
+custom_labels <- c(taxdiv_r_squared_linear = "Linear Model", taxdiv_r_squared_exp = "Exponential Model")
+
+(taxdiv_r_squared_plot <- ggplot(taxdiv_r_squared_long,aes(x = resolution, y = taxdiv_r_squared, color = model)) +
+  geom_line() +
+  geom_point(size = 3) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = taxdiv_r_squared_compare$resolution) + 
+  labs(title = "Comparison of Richness R² Values by Resolution",
+       x = "Grain (km)",
+       y = "R² Value",
+       color = "Model Type") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = "inside",
+        legend.position.inside = c(.85, .3),
+        legend.box.margin = margin(10, 10, 10, 10),
+        text = element_text(size = 18)))
+
+# Calculate R² values for linear model
+taxdiv_lm_linear <- lm(taxdiv_r_squared_linear ~ resolution, data = taxdiv_r_squared_compare)
+taxdiv_r_squared_linear <- summary(taxdiv_lm_linear)$r.squared
+taxdiv_slopecoefficent_linear <- coef(taxdiv_lm_linear)["resolution"]
+
+# Calculate R² values for exponential model
+taxdiv_lm_exp <- lm(taxdiv_r_squared_exp ~ resolution, data = taxdiv_r_squared_compare)
+taxdiv_r_squared_exp <- summary(taxdiv_lm_exp)$r.squared
+taxdiv_slopecoefficent_exp <- coef(taxdiv_lm_exp)["resolution"]
+
+(taxdiv_r_squared_plot_trendlines <- ggplot() +
+  geom_smooth(aes(x = taxdiv_r_squared_compare$resolution, y = taxdiv_r_squared_compare$taxdiv_r_squared_linear), 
+              method = "lm", se = FALSE, color = "red", linetype = "solid") +
+  geom_smooth(aes(x = taxdiv_r_squared_compare$resolution, y = taxdiv_r_squared_compare$taxdiv_r_squared_exp), 
+              method = "lm", se = FALSE, color = "purple", linetype = "solid") +
+  geom_point(aes(x = taxdiv_r_squared_long$resolution, y = taxdiv_r_squared_long$taxdiv_r_squared, color = taxdiv_r_squared_long$model),
+             size = 3) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = taxdiv_r_squared_compare$resolution) + 
+  labs(title = "Comparison of Richness R² Values by Resolution",
+       x = "Grain (km)",
+       y = "R² Value",
+       color = "Model Type") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = "inside",
+        legend.position.inside = c(0.2, 0.85),
+        legend.box.margin = margin(10, 10, 10, 10),
+        text = element_text(size = 18)) +
+  annotate("text", x = 55, y = 0.4, label = paste("R² =", round(taxdiv_r_squared_linear, 2)), color = "red", size = 6) +
+  annotate("text", x = 60, y = 0.7, label = paste("R² =", round(taxdiv_r_squared_exp, 2)), color = "purple", size = 6))
+```
+
+```{r}
+(taxdiv_r_squared_plot_both <- ggplot(taxdiv_r_squared_long, aes(x = resolution, y = taxdiv_r_squared, color = model, group = model)) +
+  geom_line(alpha = 0.25) +
+  geom_point(size = 3) +
+  geom_smooth(data = taxdiv_r_squared_compare, aes(x = resolution, y = taxdiv_r_squared_linear), 
+              method = "lm", se = FALSE, color = "red", linetype = "dashed", inherit.aes = FALSE) +
+  geom_smooth(data = taxdiv_r_squared_compare, aes(x = resolution, y = taxdiv_r_squared_exp), 
+              method = "lm", se = FALSE, color = "purple", linetype = "dashed", inherit.aes = FALSE) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = taxdiv_r_squared_compare$resolution) + 
+  labs(title = "Comparison of Richness R² Values by Resolution",
+       x = "Grain (km)",
+       y = "R² Value",
+       color = "Model Type") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = "inside",
+        legend.position.inside = c(0.2, 0.85),
+        legend.box.margin = margin(10, 10, 10, 10),
+        text = element_text(size = 18)) +
+  annotate("text", x = 55, y = 0.4, label = paste("R² =", round(taxdiv_r_squared_linear, 2)), color = "red", size = 6) +
+  annotate("text", x = 60, y = 0.7, label = paste("R² =", round(taxdiv_r_squared_exp, 2)), color = "purple", size = 6))
+
+```
+
+```{r}
+ggtrendline(taxdiv_r_squared_compare$resolution, taxdiv_r_squared_compare$taxdiv_r_squared_linear, model = "line2P", summary = FALSE, linecolor = "red", linewidth = 1, CI.level = NA, show.eq = FALSE, show.pvalue = FALSE, text.col = "red", eSize = 4)
+ggtrendline(taxdiv_r_squared_compare$resolution, taxdiv_r_squared_compare$taxdiv_r_squared_exp, model = "line2P", summary = FALSE, linecolor = "purple", linewidth = 1, CI.level = NA, show.eq = FALSE, show.pvalue = FALSE, text.col = "purple", eSize = 4)
+
+ggtrendline(taxdiv_r_squared_compare$resolution, taxdiv_r_squared_compare$taxdiv_r_squared_linear, model = "line3P", summary = FALSE, linecolor = "red", linewidth = 1, CI.level = NA, show.eq = FALSE, show.pvalue = FALSE, text.col = "red", eSize = 4)
+ggtrendline(taxdiv_r_squared_compare$resolution, taxdiv_r_squared_compare$taxdiv_r_squared_exp, model = "line3P", summary = FALSE, linecolor = "purple", linewidth = 1, CI.level = NA, show.eq = FALSE, show.pvalue = FALSE, text.col = "purple", eSize = 4)
+```
+
 ## Functional Diversity
 ```{r}
+# create dataframe of resolutions and R² values 
+resolution <- c(5, 10, 25, 50, 75, 100)
+fundiv_r_squared_linear <- c(compare_FDis_r2_5km_linear, compare_FDis_r2_10km_linear, compare_FDis_r2_25km_linear, compare_FDis_r2_50km_linear, compare_FDis_r2_75km_linear, compare_FDis_r2_100km_linear)
+fundiv_r_squared_quad <- c(compare_FDis_r2_5km_quad, compare_FDis_r2_10km_quad, compare_FDis_r2_25km_quad, compare_FDis_r2_50km_quad, compare_FDis_r2_75km_quad, compare_FDis_r2_100km_quad)
+fundiv_r_squared_compare <- data.frame(resolution, fundiv_r_squared_linear, fundiv_r_squared_quad)
+
+# reshape to long format for ease of plotting
+fundiv_r_squared_long <- fundiv_r_squared_compare %>%
+  pivot_longer(cols = c(fundiv_r_squared_linear, fundiv_r_squared_quad), 
+               names_to = "model", 
+               values_to = "fundiv_r_squared")
 
 ```
+
+```{r}
+# Define custom colors for the points
+custom_colors <- c(fundiv_r_squared_linear = "red", fundiv_r_squared_quad = "blue")
+
+# Define custom legend labels
+custom_labels <- c(fundiv_r_squared_quad = "Quadratic Model", fundiv_r_squared_linear = "Linear Model")
+
+(fundiv_r_squared_plot <- ggplot(fundiv_r_squared_long,aes(x = resolution, y = fundiv_r_squared, color = model)) +
+  geom_line() +
+  geom_point(size = 3) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = fundiv_r_squared_compare$resolution) + 
+  labs(title = "Comparison of FDis R² Values by Resolution",
+       x = "Grain (km)",
+       y = "R² Value",
+       color = "Model Type") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = c(0.2, 0.85),
+        legend.box.margin = margin(10, 10, 10, 10),
+        text = element_text(size = 18)))
+
+# Calculate R² values for linear model
+fundiv_lm_linear <- lm(fundiv_r_squared_linear ~ resolution, data = fundiv_r_squared_compare)
+fundiv_r_squared_linear <- summary(fundiv_lm_linear)$r.squared
+
+# Calculate R² values for quadratic model
+fundiv_lm_quad <- lm(fundiv_r_squared_quad ~ resolution, data = fundiv_r_squared_compare)
+fundiv_r_squared_quad <- summary(fundiv_lm_quad)$r.squared
+
+(fundiv_r_squared_plot_trendlines <- ggplot() +
+  geom_smooth(aes(x = fundiv_r_squared_compare$resolution, y = fundiv_r_squared_compare$fundiv_r_squared_linear), 
+              method = "lm", se = FALSE, color = "red", linetype = "solid") +
+  geom_smooth(aes(x = fundiv_r_squared_compare$resolution, y = fundiv_r_squared_compare$fundiv_r_squared_quad), 
+              method = "lm", se = FALSE, color = "blue", linetype = "solid") +
+  geom_point(aes(x = fundiv_r_squared_long$resolution, y = fundiv_r_squared_long$fundiv_r_squared, color = fundiv_r_squared_long$model),
+             size = 3) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = fundiv_r_squared_compare$resolution) + 
+  labs(title = "Comparison of FDis R² Values by Resolution",
+       x = "Grain (km)",
+       y = "R² Value",
+       color = "Model Type") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = c(0.2, 0.85),
+        legend.box.margin = margin(10, 10, 10, 10),
+        text = element_text(size = 18)) +
+  annotate("text", x = 60, y = 0.02, label = paste("R² =", round(fundiv_r_squared_linear, 2)), color = "red", size = 6) +
+  annotate("text", x = 25, y = 0.05, label = paste("R² =", round(fundiv_r_squared_quad, 2)), color = "blue", size = 6))
+```
+
+```{r}
+(fundiv_r_squared_plot_both <- ggplot(fundiv_r_squared_long, aes(x = resolution, y = fundiv_r_squared, color = model, group = model)) +
+  geom_line(alpha = 0.25) +
+  geom_point(size = 3) +
+  geom_smooth(data = fundiv_r_squared_compare, aes(x = resolution, y = fundiv_r_squared_linear), 
+              method = "lm", se = FALSE, color = "red", linetype = "dashed", inherit.aes = FALSE) +
+  geom_smooth(data = fundiv_r_squared_compare, aes(x = resolution, y = fundiv_r_squared_quad), 
+              method = "lm", se = FALSE, color = "blue", linetype = "dashed", inherit.aes = FALSE) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = fundiv_r_squared_compare$resolution) + 
+  labs(title = "Comparison of FDis R² Values by Resolution",
+       x = "Grain (km)",
+       y = "R² Value",
+       color = "Model Type") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = c(0.2, 0.85),
+        legend.box.margin = margin(10, 10, 10, 10),
+        text = element_text(size = 18)) +
+  annotate("text", x = 40, y = 0.02, label = paste("R² =", round(fundiv_r_squared_linear, 2)), color = "red", size = 6) +
+  annotate("text", x = 25, y = 0.05, label = paste("R² =", round(fundiv_r_squared_quad, 2)), color = "blue", size = 6))
+
+```
+
+## Compare the linear model slope coefficents for taxonomic and functional diversity
+```{r}
+# Create a tibble
+bothdiv_r2_compare <- tibble(
+  resolution = c(5, 10, 25, 50, 75, 100),
+  taxdiv_r2_linear = c(
+    as.numeric(compare_Richness_r2_5km_linear), 
+    as.numeric(compare_Richness_r2_10km_linear), 
+    as.numeric(compare_Richness_r2_25km_linear), 
+    as.numeric(compare_Richness_r2_50km_linear), 
+    as.numeric(compare_Richness_r2_75km_linear), 
+    as.numeric(compare_Richness_r2_100km_linear)
+  ),
+  fundiv_r2_linear = c(
+    as.numeric(compare_FDis_r2_5km_linear), 
+    as.numeric(compare_FDis_r2_10km_linear), 
+    as.numeric(compare_FDis_r2_25km_linear), 
+    as.numeric(compare_FDis_r2_50km_linear), 
+    as.numeric(compare_FDis_r2_75km_linear), 
+    as.numeric(compare_FDis_r2_100km_linear)
+  )
+)
+
+# reshape to long format for ease of plotting
+bothdiv_r2_long <- bothdiv_r2_compare %>%
+  pivot_longer(cols = c(taxdiv_r2_linear, fundiv_r2_linear), 
+               names_to = "type", 
+               values_to = "r_squared")
+```
+
+```{r}
+# Define custom colors for the points
+custom_colors <- c(taxdiv_r2_linear = "goldenrod", fundiv_r2_linear = "forestgreen")
+
+# Define custom legend labels
+custom_labels <- c(taxdiv_r2_linear = "Taxonomic Diversity", fundiv_r2_linear = "Functional Diversity")
+
+(bothdiv_r2_plot <- ggplot(bothdiv_r2_long,aes(x = resolution, y = r_squared, color = type)) +
+  geom_line() +
+  geom_point(size = 3) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = bothdiv_r2_compare$resolution) + 
+  labs(x = "Grain (km)",
+       y = "R² value",
+       color = "Diversity Measure") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = "inside",
+        legend.position.inside = c(.2, .8),
+        legend.box.margin = margin(10, 10, 10, 10),
+        text = element_text(size = 18)))
+```
+
+# Compare slope coefficents by resolution
+## Taxonomic diversity
+```{r}
+# Create a tibble
+taxdiv_slopecoefficent_compare <- tibble(
+  resolution = c(5, 10, 25, 50, 75, 100),
+  taxdiv_slopecoefficent_linear = c(
+    as.numeric(compare_Richness_slopecoefficent_5km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_10km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_25km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_50km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_75km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_100km_linear$a)
+  ),
+  taxdiv_slopecoefficent_exp = c(
+    as.numeric(compare_Richness_slopecoefficent_5km_exp$b), 
+    as.numeric(compare_Richness_slopecoefficent_10km_exp$b), 
+    as.numeric(compare_Richness_slopecoefficent_25km_exp$b), 
+    as.numeric(compare_Richness_slopecoefficent_50km_exp$b), 
+    as.numeric(compare_Richness_slopecoefficent_75km_exp$b), 
+    as.numeric(compare_Richness_slopecoefficent_100km_exp$b)
+  )
+)
+
+# reshape to long format for ease of plotting
+taxdiv_slopecoefficent_long <- taxdiv_slopecoefficent_compare %>%
+  pivot_longer(cols = c(taxdiv_slopecoefficent_linear, taxdiv_slopecoefficent_exp), 
+               names_to = "model", 
+               values_to = "slope_coefficent")
+```
+
+```{r}
+# Define custom colors for the points
+custom_colors <- c(taxdiv_slopecoefficent_linear = "red", taxdiv_slopecoefficent_exp = "purple")
+
+# Define custom legend labels
+custom_labels <- c(taxdiv_slopecoefficent_linear = "Linear Model", taxdiv_slopecoefficent_exp = "Exponential Model")
+
+(taxdiv_slopecoefficent_plot <- ggplot(taxdiv_slopecoefficent_long,aes(x = resolution, y = slope_coefficent, color = model)) +
+  geom_line() +
+  geom_point(size = 3) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = taxdiv_slopecoefficent_compare$resolution) + 
+  labs(title = "Comparison of Richness Slope Coefficents by Resolution",
+       x = "Grain (km)",
+       y = "Slope Coefficent",
+       color = "Model Type") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = "inside",
+        legend.position.inside = c(.85, .9),
+        legend.box.margin = margin(10, 10, 10, 10),
+        text = element_text(size = 18)))
+```
+
+## Functional diversity
+```{r}
+# Create a tibble
+fundiv_slopecoefficent_compare <- tibble(
+  resolution = c(5, 10, 25, 50, 75, 100),
+  fundiv_slopecoefficent_linear = c(
+    as.numeric(compare_FDis_slopecoefficent_5km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_10km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_25km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_50km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_75km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_100km_linear$a)
+  ),
+  fundiv_slopecoefficent_quad = c(
+    as.numeric(compare_FDis_slopecoefficent_5km_quad$b), 
+    as.numeric(compare_FDis_slopecoefficent_10km_quad$b), 
+    as.numeric(compare_FDis_slopecoefficent_25km_quad$b), 
+    as.numeric(compare_FDis_slopecoefficent_50km_quad$b), 
+    as.numeric(compare_FDis_slopecoefficent_75km_quad$b), 
+    as.numeric(compare_FDis_slopecoefficent_100km_quad$b)
+  )
+)
+
+# reshape to long format for ease of plotting
+fundiv_slopecoefficent_long <- fundiv_slopecoefficent_compare %>%
+  pivot_longer(cols = c(fundiv_slopecoefficent_linear, fundiv_slopecoefficent_quad), 
+               names_to = "model", 
+               values_to = "slope_coefficent")
+```
+
+```{r}
+# Define custom colors for the points
+custom_colors <- c(fundiv_slopecoefficent_linear = "red", fundiv_slopecoefficent_quad = "blue")
+
+# Define custom legend labels
+custom_labels <- c(fundiv_slopecoefficent_linear = "Linear Model", fundiv_slopecoefficent_quad = "Quadratic Model")
+
+(fundiv_slopecoefficent_plot <- ggplot(fundiv_slopecoefficent_long,aes(x = resolution, y = slope_coefficent, color = model)) +
+  geom_line() +
+  geom_point(size = 3) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = fundiv_slopecoefficent_compare$resolution) + 
+  labs(title = "Comparison of FDis Slope Coefficents by Resolution",
+       x = "Grain (km)",
+       y = "Slope Coefficent",
+       color = "Model Type") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = "inside",
+        legend.position.inside = c(.85, .2),
+        legend.box.margin = margin(10, 10, 10, 10),
+        text = element_text(size = 18)))
+```
+
+## Compare the linear model slope coefficents for taxonomic and functional diversity
+```{r}
+# Create a tibble
+bothdiv_slopecoefficent_compare <- tibble(
+  resolution = c(5, 10, 25, 50, 75, 100),
+  taxdiv_slopecoefficent_linear = c(
+    as.numeric(compare_Richness_slopecoefficent_5km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_10km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_25km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_50km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_75km_linear$a), 
+    as.numeric(compare_Richness_slopecoefficent_100km_linear$a)
+  ),
+  fundiv_slopecoefficent_linear = c(
+    as.numeric(compare_FDis_slopecoefficent_5km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_10km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_25km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_50km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_75km_linear$a), 
+    as.numeric(compare_FDis_slopecoefficent_100km_linear$a)
+  )
+)
+
+# reshape to long format for ease of plotting
+bothdiv_slopecoefficent_long <- bothdiv_slopecoefficent_compare %>%
+  pivot_longer(cols = c(taxdiv_slopecoefficent_linear, fundiv_slopecoefficent_linear), 
+               names_to = "type", 
+               values_to = "slope_coefficent")
+```
+
+```{r}
+# Define custom colors for the points
+custom_colors <- c(taxdiv_slopecoefficent_linear = "goldenrod", fundiv_slopecoefficent_linear = "forestgreen")
+
+# Define custom legend labels
+custom_labels <- c(taxdiv_slopecoefficent_linear = "Taxonomic Diversity", fundiv_slopecoefficent_linear = "Functional Diversity")
+
+(bothdiv_slopecoefficent_plot <- ggplot(bothdiv_slopecoefficent_long,aes(x = resolution, y = slope_coefficent, color = type)) +
+  geom_line() +
+  geom_point(size = 3) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = bothdiv_slopecoefficent_compare$resolution) + 
+  labs(x = "Grain (km)",
+       y = "Slope Coefficent",
+       color = "Diversity Measure") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = "inside",
+        legend.position.inside = c(.85, .9),
+        legend.box.margin = margin(10, 10, 10, 10),
+        text = element_text(size = 18)))
+```
+
+# Save plots to file
+```{r}
+ggsave("compare_Richness_all_plots.png", compare_Richness_all_plots, path = figure_path, width = 21, height = 10, units = "in")
+ggsave("compare_FDis_all_plots.png", compare_FDis_all_plots, path = figure_path, width = 21, height = 10, units = "in")
+```
+
+```{r}
+ggsave("compare_Richness_all_plots_long.png", compare_Richness_all_plots_long, path = figure_path, width = 20, height = 5, units = "in")
+ggsave("compare_FDis_all_plots_long.png", compare_FDis_all_plots_long, path = figure_path, width = 20, height = 5, units = "in")
+```
+
+```{r}
+ggsave("taxdiv_r_squared_plot_large.png", taxdiv_r_squared_plot, path = figure_path)
+ggsave("taxdiv_r_squared_plot_trendlines_large.png", taxdiv_r_squared_plot_trendlines, path = figure_path)
+ggsave("taxdiv_r_squared_plot_both_large.png", taxdiv_r_squared_plot_both, path = figure_path)
+ggsave("fundiv_r_squared_plot_large.png", fundiv_r_squared_plot, path = figure_path)
+ggsave("fundiv_r_squared_plot_trendlines_large.png", fundiv_r_squared_plot_trendlines, path = figure_path)
+ggsave("fundiv_r_squared_plot_both_large.png", fundiv_r_squared_plot_both, path = figure_path)
+ggsave("bothdiv_r2_plot_large.png", bothdiv_r2_plot, path = figure_path)
+
+```
+
+```{r}
+ggsave("taxdiv_slopecoefficent_plot_large.png", taxdiv_slopecoefficent_plot, path = figure_path)
+ggsave("fundiv_slopecoefficent_plot_large.png", fundiv_slopecoefficent_plot, path = figure_path)
+ggsave("bothdiv_slopecoefficent_plot_large.png", bothdiv_slopecoefficent_plot, path = figure_path)
+```
+
+# Compare same tax taxanomic and functional diversity
+## Function for same taxa diversity comparisons
+```{r}
+same_taxa_diversity <- function(richness_grid, fdis_grid, resolution_meters, z){
+  # Extract cell values
+  comparison_diversity <- data.frame(Richness = richness_grid$num_species,
+                                    FDis = fdis_grid$fdis_value)
+
+  # Filter out rows where either richness or FDis is zero
+  comparison_diversity_filtered <- comparison_diversity %>%
+                                filter(Richness > 0 & FDis > 0)
+
+  ## Create scatter plots
+  (compare_diversity_plot_line2P <- ggtrendline(comparison_diversity_filtered$Richness,
+               comparison_diversity_filtered$FDis, model = z, linewidth = 1, CI.level = NA, eSize = 4) +
+      geom_point(aes(x = comparison_diversity_filtered$Richness,
+                                           y = comparison_diversity_filtered$FDis)) +
+      labs(title = paste("Richness vs FDis [", resolution_meters/1000, " km]", sep = ""),
+       x = "Richness by cell", y = "FDis by cell"))
+
+}
+```
+
+## Plants
+```{r}
+same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "line2P")
+same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "line3P")
+same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "log2P")
+same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "exp2P")
+same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "exp3P")
+same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "power2P")
+same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "power3P")
+```
+
+```{r}
+same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "line2P")
+same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "line3P")
+same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "log2P")
+same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "exp2P")
+same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "exp3P")
+same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "power2P")
+#same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "power3P")
+```
+
+```{r}
+same_taxa_diversity(plant_cellRichness_50km, plant_cellFDis_50km, 50000, "line2P")
+same_taxa_diversity(plant_cellRichness_50km, plant_cellFDis_50km, 50000, "line3P")
+same_taxa_diversity(plant_cellRichness_50km, plant_cellFDis_50km, 50000, "log2P")
+same_taxa_diversity(plant_cellRichness_50km, plant_cellFDis_50km, 50000, "exp2P")
+same_taxa_diversity(plant_cellRichness_50km, plant_cellFDis_50km, 50000, "exp3P")
+same_taxa_diversity(plant_cellRichness_50km, plant_cellFDis_50km, 50000, "power2P")
+same_taxa_diversity(plant_cellRichness_50km, plant_cellFDis_50km, 50000, "power3P")
+```
+
+```{r}
+same_taxa_diversity(plant_cellRichness_25km, plant_cellFDis_25km, 25000, "line2P")
+same_taxa_diversity(plant_cellRichness_25km, plant_cellFDis_25km, 25000, "line3P")
+same_taxa_diversity(plant_cellRichness_25km, plant_cellFDis_25km, 25000, "log2P")
+same_taxa_diversity(plant_cellRichness_25km, plant_cellFDis_25km, 25000, "exp2P")
+same_taxa_diversity(plant_cellRichness_25km, plant_cellFDis_25km, 25000, "exp3P")
+same_taxa_diversity(plant_cellRichness_25km, plant_cellFDis_25km, 25000, "power2P")
+#same_taxa_diversity(plant_cellRichness_25km, plant_cellFDis_25km, 25000, "power3P")
+```
+
+```{r}
+same_taxa_diversity(plant_cellRichness_10km, plant_cellFDis_10km, 10000, "line2P")
+same_taxa_diversity(plant_cellRichness_10km, plant_cellFDis_10km, 10000, "line3P")
+same_taxa_diversity(plant_cellRichness_10km, plant_cellFDis_10km, 10000, "log2P")
+same_taxa_diversity(plant_cellRichness_10km, plant_cellFDis_10km, 10000, "exp2P")
+same_taxa_diversity(plant_cellRichness_10km, plant_cellFDis_10km, 10000, "exp3P")
+same_taxa_diversity(plant_cellRichness_10km, plant_cellFDis_10km, 10000, "power2P")
+#same_taxa_diversity(plant_cellRichness_10km, plant_cellFDis_10km, 10000, "power3P")
+```
+
+```{r}
+same_taxa_diversity(plant_cellRichness_5km, plant_cellFDis_5km, 5000, "line2P")
+same_taxa_diversity(plant_cellRichness_5km, plant_cellFDis_5km, 5000, "line3P")
+same_taxa_diversity(plant_cellRichness_5km, plant_cellFDis_5km, 5000, "log2P")
+same_taxa_diversity(plant_cellRichness_5km, plant_cellFDis_5km, 5000, "exp2P")
+same_taxa_diversity(plant_cellRichness_5km, plant_cellFDis_5km, 5000, "exp3P")
+same_taxa_diversity(plant_cellRichness_5km, plant_cellFDis_5km, 5000, "power2P")
+#same_taxa_diversity(plant_cellRichness_5km, plant_cellFDis_5km, 5000, "power3P")
+```
+
+## Frugivores
+```{r}
+same_taxa_diversity(frugivore_cellRichness_100km, frugivore_cellFDis_100km, 100000, "line2P")
+same_taxa_diversity(frugivore_cellRichness_100km, frugivore_cellFDis_100km, 100000, "line3P")
+same_taxa_diversity(frugivore_cellRichness_100km, frugivore_cellFDis_100km, 100000, "log2P")
+same_taxa_diversity(frugivore_cellRichness_100km, frugivore_cellFDis_100km, 100000, "exp2P")
+same_taxa_diversity(frugivore_cellRichness_100km, frugivore_cellFDis_100km, 100000, "exp3P")
+same_taxa_diversity(frugivore_cellRichness_100km, frugivore_cellFDis_100km, 100000, "power2P")
+same_taxa_diversity(frugivore_cellRichness_100km, frugivore_cellFDis_100km, 100000, "power3P")
+```
+
+```{r}
+same_taxa_diversity(frugivore_cellRichness_75km, frugivore_cellFDis_75km, 75000, "line2P")
+same_taxa_diversity(frugivore_cellRichness_75km, frugivore_cellFDis_75km, 75000, "line3P")
+same_taxa_diversity(frugivore_cellRichness_75km, frugivore_cellFDis_75km, 75000, "log2P")
+same_taxa_diversity(frugivore_cellRichness_75km, frugivore_cellFDis_75km, 75000, "exp2P")
+same_taxa_diversity(frugivore_cellRichness_75km, frugivore_cellFDis_75km, 75000, "exp3P")
+same_taxa_diversity(frugivore_cellRichness_75km, frugivore_cellFDis_75km, 75000, "power2P")
+#same_taxa_diversity(frugivore_cellRichness_75km, frugivore_cellFDis_75km, 75000, "power3P")
+```
+
+```{r}
+same_taxa_diversity(frugivore_cellRichness_50km, frugivore_cellFDis_50km, 50000, "line2P")
+same_taxa_diversity(frugivore_cellRichness_50km, frugivore_cellFDis_50km, 50000, "line3P")
+same_taxa_diversity(frugivore_cellRichness_50km, frugivore_cellFDis_50km, 50000, "log2P")
+same_taxa_diversity(frugivore_cellRichness_50km, frugivore_cellFDis_50km, 50000, "exp2P")
+same_taxa_diversity(frugivore_cellRichness_50km, frugivore_cellFDis_50km, 50000, "exp3P")
+same_taxa_diversity(frugivore_cellRichness_50km, frugivore_cellFDis_50km, 50000, "power2P")
+same_taxa_diversity(frugivore_cellRichness_50km, frugivore_cellFDis_50km, 50000, "power3P")
+```
+
+```{r}
+same_taxa_diversity(frugivore_cellRichness_25km, frugivore_cellFDis_25km, 25000, "line2P")
+same_taxa_diversity(frugivore_cellRichness_25km, frugivore_cellFDis_25km, 25000, "line3P")
+same_taxa_diversity(frugivore_cellRichness_25km, frugivore_cellFDis_25km, 25000, "log2P")
+same_taxa_diversity(frugivore_cellRichness_25km, frugivore_cellFDis_25km, 25000, "exp2P")
+#same_taxa_diversity(frugivore_cellRichness_25km, frugivore_cellFDis_25km, 25000, "exp3P")
+same_taxa_diversity(frugivore_cellRichness_25km, frugivore_cellFDis_25km, 25000, "power2P")
+#same_taxa_diversity(frugivore_cellRichness_25km, frugivore_cellFDis_25km, 25000, "power3P")
+```
+
+```{r}
+same_taxa_diversity(frugivore_cellRichness_10km, frugivore_cellFDis_10km, 10000, "line2P")
+same_taxa_diversity(frugivore_cellRichness_10km, frugivore_cellFDis_10km, 10000, "line3P")
+same_taxa_diversity(frugivore_cellRichness_10km, frugivore_cellFDis_10km, 10000, "log2P")
+same_taxa_diversity(frugivore_cellRichness_10km, frugivore_cellFDis_10km, 10000, "exp2P")
+same_taxa_diversity(frugivore_cellRichness_10km, frugivore_cellFDis_10km, 10000, "exp3P")
+same_taxa_diversity(frugivore_cellRichness_10km, frugivore_cellFDis_10km, 10000, "power2P")
+#same_taxa_diversity(frugivore_cellRichness_10km, frugivore_cellFDis_10km, 10000, "power3P")
+```
+
+```{r}
+same_taxa_diversity(frugivore_cellRichness_5km, frugivore_cellFDis_5km, 5000, "line2P")
+same_taxa_diversity(frugivore_cellRichness_5km, frugivore_cellFDis_5km, 5000, "line3P")
+same_taxa_diversity(frugivore_cellRichness_5km, frugivore_cellFDis_5km, 5000, "log2P")
+same_taxa_diversity(frugivore_cellRichness_5km, frugivore_cellFDis_5km, 5000, "exp2P")
+same_taxa_diversity(frugivore_cellRichness_5km, frugivore_cellFDis_5km, 5000, "exp3P")
+same_taxa_diversity(frugivore_cellRichness_5km, frugivore_cellFDis_5km, 5000, "power2P")
+#same_taxa_diversity(frugivore_cellRichness_5km, frugivore_cellFDis_5km, 5000, "power3P")
+```
+
+## Create plots for manuscript
+```{r}
+compare_TD.FD <- function(Richness_grid, FDis_grid, resolution_meters, x.label, y.label){
+
+  # Scatter plot
+  compare_TD.FD_plot <- ggtrendline(Richness_grid$num_species, FDis_grid$fdis_value, model = "line2P",
+                                    linecolor = "black", linewidth = 1, CI.level = NA, summary = FALSE,
+                                    show.eq = FALSE, show.pvalue = FALSE, text.col = "black", eSize = 4) +
+      geom_point(aes(x = Richness_grid$num_species, y = FDis_grid$fdis_value), alpha = 0.25, size = 1) +
+      labs(title = paste("[", resolution_meters/1000, " km]", sep = ""),
+       x = paste(" ",x.label," Richness by cell", sep = ""), y = paste(" ",y.label," FDis by cell", sep = "")) +
+      scale_x_continuous(expand = c(0, 0)) + 
+      scale_y_continuous(expand = c(0, 0)) + 
+      theme_minimal() +
+      theme(panel.grid.major = element_blank(), 
+            panel.grid.minor = element_blank(),
+            axis.line = element_line(color = "black"),
+            axis.ticks = element_line(color = "black"),
+            axis.text = element_text(color = "black"))
+
+  # Model summary
+  line2P_summary <- trendline_sum(Richness_grid$num_species, FDis_grid$fdis_value, model = "line2P")
+
+  list(compare_TD.FD_plot = compare_TD.FD_plot,
+       line2P_summary = line2P_summary)
+}
+```
+
+## Plant TD - Plant FD
+```{r}
+compare_plantTD.FD_5km <- compare_TD.FD(plant_cellRichness_5km, plant_cellFDis_5km, 5000, "Plant", "Plant")
+(compare_plantTD.FD_5km_plot <- compare_plantTD.FD_5km$compare_TD.FD_plot)
+
+compare_plantTD.FD_r2_5km_linear <- compare_plantTD.FD_5km$line2P_summary$R.squared
+compare_plantTD.FD_slopecoefficent_5km_linear <- compare_plantTD.FD_5km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_plantTD.FD_10km <- compare_TD.FD(plant_cellRichness_10km, plant_cellFDis_10km, 10000, "Plant", "Plant")
+(compare_plantTD.FD_10km_plot <- compare_plantTD.FD_10km$compare_TD.FD_plot)
+
+compare_plantTD.FD_r2_10km_linear <- compare_plantTD.FD_10km$line2P_summary$R.squared
+compare_plantTD.FD_slopecoefficent_10km_linear <- compare_plantTD.FD_10km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_plantTD.FD_25km <- compare_TD.FD(plant_cellRichness_25km, plant_cellFDis_25km, 25000, "Plant", "Plant")
+(compare_plantTD.FD_25km_plot <- compare_plantTD.FD_25km$compare_TD.FD_plot)
+
+compare_plantTD.FD_r2_25km_linear <- compare_plantTD.FD_25km$line2P_summary$R.squared
+compare_plantTD.FD_slopecoefficent_25km_linear <- compare_plantTD.FD_25km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_plantTD.FD_50km <- compare_TD.FD(plant_cellRichness_50km, plant_cellFDis_50km, 50000, "Plant", "Plant")
+(compare_plantTD.FD_50km_plot <- compare_plantTD.FD_50km$compare_TD.FD_plot)
+
+compare_plantTD.FD_r2_50km_linear <- compare_plantTD.FD_50km$line2P_summary$R.squared
+compare_plantTD.FD_slopecoefficent_50km_linear <- compare_plantTD.FD_50km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_plantTD.FD_75km <- compare_TD.FD(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "Plant", "Plant")
+(compare_plantTD.FD_75km_plot <- compare_plantTD.FD_75km$compare_TD.FD_plot)
+
+compare_plantTD.FD_r2_75km_linear <- compare_plantTD.FD_75km$line2P_summary$R.squared
+compare_plantTD.FD_slopecoefficent_75km_linear <- compare_plantTD.FD_75km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_plantTD.FD_100km <- compare_TD.FD(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "Plant", "Plant")
+(compare_plantTD.FD_100km_plot <- compare_plantTD.FD_100km$compare_TD.FD_plot)
+
+compare_plantTD.FD_r2_100km_linear <- compare_plantTD.FD_100km$line2P_summary$R.squared
+compare_plantTD.FD_slopecoefficent_100km_linear <- compare_plantTD.FD_100km$line2P_summary$parameter["a"]
+```
+
+## Frugivore TD - Frugivore FD
+```{r}
+compare_frugivoreTD.FD_5km <- compare_TD.FD(frugivore_cellRichness_5km, frugivore_cellFDis_5km, 5000, "Frugivore", "Frugivore")
+(compare_frugivoreTD.FD_5km_plot <- compare_frugivoreTD.FD_5km$compare_TD.FD_plot)
+
+compare_frugivoreTD.FD_r2_5km_linear <- compare_frugivoreTD.FD_5km$line2P_summary$R.squared
+compare_frugivoreTD.FD_slopecoefficent_5km_linear <- compare_frugivoreTD.FD_5km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_frugivoreTD.FD_10km <- compare_TD.FD(frugivore_cellRichness_10km, frugivore_cellFDis_10km, 10000, "Frugivore", "Frugivore")
+(compare_frugivoreTD.FD_10km_plot <- compare_frugivoreTD.FD_10km$compare_TD.FD_plot)
+
+compare_frugivoreTD.FD_r2_10km_linear <- compare_frugivoreTD.FD_10km$line2P_summary$R.squared
+compare_frugivoreTD.FD_slopecoefficent_10km_linear <- compare_frugivoreTD.FD_10km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_frugivoreTD.FD_25km <- compare_TD.FD(frugivore_cellRichness_25km, frugivore_cellFDis_25km, 25000, "Frugivore", "Frugivore")
+(compare_frugivoreTD.FD_25km_plot <- compare_frugivoreTD.FD_25km$compare_TD.FD_plot)
+
+compare_frugivoreTD.FD_r2_25km_linear <- compare_frugivoreTD.FD_25km$line2P_summary$R.squared
+compare_frugivoreTD.FD_slopecoefficent_25km_linear <- compare_frugivoreTD.FD_25km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_frugivoreTD.FD_50km <- compare_TD.FD(frugivore_cellRichness_50km, frugivore_cellFDis_50km, 50000, "Frugivore", "Frugivore")
+(compare_frugivoreTD.FD_50km_plot <- compare_frugivoreTD.FD_50km$compare_TD.FD_plot)
+
+compare_frugivoreTD.FD_r2_50km_linear <- compare_frugivoreTD.FD_50km$line2P_summary$R.squared
+compare_frugivoreTD.FD_slopecoefficent_50km_linear <- compare_frugivoreTD.FD_50km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_frugivoreTD.FD_75km <- compare_TD.FD(frugivore_cellRichness_75km, frugivore_cellFDis_75km, 75000, "Frugivore", "Frugivore")
+(compare_frugivoreTD.FD_75km_plot <- compare_frugivoreTD.FD_75km$compare_TD.FD_plot)
+
+compare_frugivoreTD.FD_r2_75km_linear <- compare_frugivoreTD.FD_75km$line2P_summary$R.squared
+compare_frugivoreTD.FD_slopecoefficent_75km_linear <- compare_frugivoreTD.FD_75km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_frugivoreTD.FD_100km <- compare_TD.FD(frugivore_cellRichness_100km, frugivore_cellFDis_100km, 100000, "Frugivore", "Frugivore")
+(compare_frugivoreTD.FD_100km_plot <- compare_frugivoreTD.FD_100km$compare_TD.FD_plot)
+
+compare_frugivoreTD.FD_r2_100km_linear <- compare_frugivoreTD.FD_100km$line2P_summary$R.squared
+compare_frugivoreTD.FD_slopecoefficent_100km_linear <- compare_frugivoreTD.FD_100km$line2P_summary$parameter["a"]
+```
+
+## Plant TD - Frugivore FD
+```{r}
+compare_plantTD.frugivoreFD_5km <- compare_TD.FD(plant_cellRichness_5km, frugivore_cellFDis_5km, 5000, "Plant", "Frugivore")
+(compare_plantTD.frugivoreFD_5km_plot <- compare_plantTD.frugivoreFD_5km$compare_TD.FD_plot)
+
+compare_plantTD.frugivoreFD_r2_5km_linear <- compare_plantTD.frugivoreFD_5km$line2P_summary$R.squared
+compare_plantTD.frugivoreFD_slopecoefficent_5km_linear <- compare_plantTD.frugivoreFD_5km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_plantTD.frugivoreFD_10km <- compare_TD.FD(plant_cellRichness_10km, frugivore_cellFDis_10km, 10000, "Plant", "Frugivore")
+(compare_plantTD.frugivoreFD_10km_plot <- compare_plantTD.frugivoreFD_10km$compare_TD.FD_plot)
+
+compare_plantTD.frugivoreFD_r2_10km_linear <- compare_plantTD.frugivoreFD_10km$line2P_summary$R.squared
+compare_plantTD.frugivoreFD_slopecoefficent_10km_linear <- compare_plantTD.frugivoreFD_10km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_plantTD.frugivoreFD_25km <- compare_TD.FD(plant_cellRichness_25km, frugivore_cellFDis_25km, 25000, "Plant", "Frugivore")
+(compare_plantTD.frugivoreFD_25km_plot <- compare_plantTD.frugivoreFD_25km$compare_TD.FD_plot)
+
+compare_plantTD.frugivoreFD_r2_25km_linear <- compare_plantTD.frugivoreFD_25km$line2P_summary$R.squared
+compare_plantTD.frugivoreFD_slopecoefficent_25km_linear <- compare_plantTD.frugivoreFD_25km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_plantTD.frugivoreFD_50km <- compare_TD.FD(plant_cellRichness_50km, frugivore_cellFDis_50km, 50000, "Plant", "Frugivore")
+(compare_plantTD.frugivoreFD_50km_plot <- compare_plantTD.frugivoreFD_50km$compare_TD.FD_plot)
+
+compare_plantTD.frugivoreFD_r2_50km_linear <- compare_plantTD.frugivoreFD_50km$line2P_summary$R.squared
+compare_plantTD.frugivoreFD_slopecoefficent_50km_linear <- compare_plantTD.frugivoreFD_50km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_plantTD.frugivoreFD_75km <- compare_TD.FD(plant_cellRichness_75km, frugivore_cellFDis_75km, 75000, "Plant", "Frugivore")
+(compare_plantTD.frugivoreFD_75km_plot <- compare_plantTD.frugivoreFD_75km$compare_TD.FD_plot)
+
+compare_plantTD.frugivoreFD_r2_75km_linear <- compare_plantTD.frugivoreFD_75km$line2P_summary$R.squared
+compare_plantTD.frugivoreFD_slopecoefficent_75km_linear <- compare_plantTD.frugivoreFD_75km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_plantTD.frugivoreFD_100km <- compare_TD.FD(plant_cellRichness_100km, frugivore_cellFDis_100km, 100000, "Plant", "Frugivore")
+(compare_plantTD.frugivoreFD_100km_plot <- compare_plantTD.frugivoreFD_100km$compare_TD.FD_plot)
+
+compare_plantTD.frugivoreFD_r2_100km_linear <- compare_plantTD.frugivoreFD_100km$line2P_summary$R.squared
+compare_plantTD.frugivoreFD_slopecoefficent_100km_linear <- compare_plantTD.frugivoreFD_100km$line2P_summary$parameter["a"]
+```
+
+## Frugivore TD - Plant FD
+```{r}
+compare_frugivoreTD.plantFD_5km <- compare_TD.FD(frugivore_cellRichness_5km, plant_cellFDis_5km, 5000, "Frugivore", "Plant")
+(compare_frugivoreTD.plantFD_5km_plot <- compare_frugivoreTD.plantFD_5km$compare_TD.FD_plot)
+
+compare_frugivoreTD.plantFD_r2_5km_linear <- compare_frugivoreTD.plantFD_5km$line2P_summary$R.squared
+compare_frugivoreTD.plantFD_slopecoefficent_5km_linear <- compare_frugivoreTD.plantFD_5km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_frugivoreTD.plantFD_10km <- compare_TD.FD(frugivore_cellRichness_10km, plant_cellFDis_10km, 10000, "Frugivore", "Plant")
+(compare_frugivoreTD.plantFD_10km_plot <- compare_frugivoreTD.plantFD_10km$compare_TD.FD_plot)
+
+compare_frugivoreTD.plantFD_r2_10km_linear <- compare_frugivoreTD.plantFD_10km$line2P_summary$R.squared
+compare_frugivoreTD.plantFD_slopecoefficent_10km_linear <- compare_frugivoreTD.plantFD_10km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_frugivoreTD.plantFD_25km <- compare_TD.FD(frugivore_cellRichness_25km, plant_cellFDis_25km, 25000, "Frugivore", "Plant")
+(compare_frugivoreTD.plantFD_25km_plot <- compare_frugivoreTD.plantFD_25km$compare_TD.FD_plot)
+
+compare_frugivoreTD.plantFD_r2_25km_linear <- compare_frugivoreTD.plantFD_25km$line2P_summary$R.squared
+compare_frugivoreTD.plantFD_slopecoefficent_25km_linear <- compare_frugivoreTD.plantFD_25km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_frugivoreTD.plantFD_50km <- compare_TD.FD(frugivore_cellRichness_50km, plant_cellFDis_50km, 50000, "Frugivore", "Plant")
+(compare_frugivoreTD.plantFD_50km_plot <- compare_frugivoreTD.plantFD_50km$compare_TD.FD_plot)
+
+compare_frugivoreTD.plantFD_r2_50km_linear <- compare_frugivoreTD.plantFD_50km$line2P_summary$R.squared
+compare_frugivoreTD.plantFD_slopecoefficent_50km_linear <- compare_frugivoreTD.plantFD_50km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_frugivoreTD.plantFD_75km <- compare_TD.FD(frugivore_cellRichness_75km, plant_cellFDis_75km, 75000, "Frugivore", "Plant")
+(compare_frugivoreTD.plantFD_75km_plot <- compare_frugivoreTD.plantFD_75km$compare_TD.FD_plot)
+
+compare_frugivoreTD.plantFD_r2_75km_linear <- compare_frugivoreTD.plantFD_75km$line2P_summary$R.squared
+compare_frugivoreTD.plantFD_slopecoefficent_75km_linear <- compare_frugivoreTD.plantFD_75km$line2P_summary$parameter["a"]
+```
+
+```{r}
+compare_frugivoreTD.plantFD_100km <- compare_TD.FD(frugivore_cellRichness_100km, plant_cellFDis_100km, 100000, "Frugivore", "Plant")
+(compare_frugivoreTD.plantFD_100km_plot <- compare_frugivoreTD.plantFD_100km$compare_TD.FD_plot)
+
+compare_frugivoreTD.plantFD_r2_100km_linear <- compare_frugivoreTD.plantFD_100km$line2P_summary$R.squared
+compare_frugivoreTD.plantFD_slopecoefficent_100km_linear <- compare_frugivoreTD.plantFD_100km$line2P_summary$parameter["a"]
+```
+
+## Multi-panel plots
+```{r}
+(compare_plantTD.FD_all_plots <- ggarrange(compare_plantTD.FD_5km_plot, compare_plantTD.FD_10km_plot,
+                                     compare_plantTD.FD_25km_plot, compare_plantTD.FD_50km_plot,
+                                     compare_plantTD.FD_75km_plot, compare_plantTD.FD_100km_plot,
+                                        ncol = 2, nrow = 3))
+```
+
+```{r}
+(compare_frugivoreTD.FD_all_plots <- ggarrange(compare_frugivoreTD.FD_5km_plot, compare_frugivoreTD.FD_10km_plot,
+                                     compare_frugivoreTD.FD_25km_plot, compare_frugivoreTD.FD_50km_plot,
+                                     compare_frugivoreTD.FD_75km_plot, compare_frugivoreTD.FD_100km_plot,
+                                        ncol = 2, nrow = 3))
+```
+
+```{r}
+(compare_plantTD.frugivoreFD_all_plots <- ggarrange(compare_plantTD.frugivoreFD_5km_plot,
+                                                    compare_plantTD.frugivoreFD_10km_plot,
+                                                    compare_plantTD.frugivoreFD_25km_plot,
+                                                    compare_plantTD.frugivoreFD_50km_plot,
+                                                    compare_plantTD.frugivoreFD_75km_plot,
+                                                    compare_plantTD.frugivoreFD_100km_plot,
+                                                      ncol = 2, nrow = 3))
+```
+
+```{r}
+(compare_frugivoreTD.plantFD_all_plots <- ggarrange(compare_frugivoreTD.plantFD_5km_plot,
+                                                    compare_frugivoreTD.plantFD_10km_plot,
+                                                    compare_frugivoreTD.plantFD_25km_plot,
+                                                    compare_frugivoreTD.plantFD_50km_plot,
+                                                    compare_frugivoreTD.plantFD_75km_plot,
+                                                    compare_frugivoreTD.plantFD_100km_plot,
+                                                      ncol = 2, nrow = 3))
+```
+
+```{r}
+# save plots
+ggsave("compare_plantTD.FD_all_plots.png", compare_plantTD.FD_all_plots, path = figure_path,
+       width = 10, height = 12, units = "in")
+ggsave("compare_frugivoreTD.FD_all_plots.png", compare_frugivoreTD.FD_all_plots, path = figure_path,
+       width = 10, height = 12, units = "in")
+ggsave("compare_plantTD.frugivoreFD_all_plots.png", compare_plantTD.frugivoreFD_all_plots, path = figure_path,
+       width = 10, height = 12, units = "in")
+ggsave("compare_frugivoreTD.plantFD_all_plots.png", compare_frugivoreTD.plantFD_all_plots, path = figure_path,
+       width = 10, height = 12, units = "in")
+```
+
+## Compare TD-FD r-squared and correlation coefficent
+```{r}
+
+# Create a tibble
+TD.FD_r2_compare <- tibble(
+  resolution = c(5, 10, 25, 50, 75, 100),
+  plantTD.FD_r2_linear = c(
+    as.numeric(compare_plantTD.FD_r2_5km_linear), 
+    as.numeric(compare_plantTD.FD_r2_10km_linear), 
+    as.numeric(compare_plantTD.FD_r2_25km_linear), 
+    as.numeric(compare_plantTD.FD_r2_50km_linear), 
+    as.numeric(compare_plantTD.FD_r2_75km_linear), 
+    as.numeric(compare_plantTD.FD_r2_100km_linear)
+  ),
+  frugivoreTD.FD_r2_linear = c(
+    as.numeric(compare_frugivoreTD.FD_r2_5km_linear), 
+    as.numeric(compare_frugivoreTD.FD_r2_10km_linear), 
+    as.numeric(compare_frugivoreTD.FD_r2_25km_linear), 
+    as.numeric(compare_frugivoreTD.FD_r2_50km_linear), 
+    as.numeric(compare_frugivoreTD.FD_r2_75km_linear), 
+    as.numeric(compare_frugivoreTD.FD_r2_100km_linear)
+  ),
+  plantTD.frugivoreFD_r2_linear = c(
+    as.numeric(compare_plantTD.frugivoreFD_r2_5km_linear), 
+    as.numeric(compare_plantTD.frugivoreFD_r2_10km_linear), 
+    as.numeric(compare_plantTD.frugivoreFD_r2_25km_linear), 
+    as.numeric(compare_plantTD.frugivoreFD_r2_50km_linear), 
+    as.numeric(compare_plantTD.frugivoreFD_r2_75km_linear), 
+    as.numeric(compare_plantTD.frugivoreFD_r2_100km_linear)
+  ),
+  frugivoreTD.plantFD_r2_linear = c(
+    as.numeric(compare_frugivoreTD.plantFD_r2_5km_linear), 
+    as.numeric(compare_frugivoreTD.plantFD_r2_10km_linear), 
+    as.numeric(compare_frugivoreTD.plantFD_r2_25km_linear), 
+    as.numeric(compare_frugivoreTD.plantFD_r2_50km_linear), 
+    as.numeric(compare_frugivoreTD.plantFD_r2_75km_linear), 
+    as.numeric(compare_frugivoreTD.plantFD_r2_100km_linear)
+  )
+)
+
+# reshape to long format for ease of plotting
+TD.FD_r2_long <- TD.FD_r2_compare %>%
+  pivot_longer(cols = c(plantTD.FD_r2_linear, frugivoreTD.FD_r2_linear,
+                        plantTD.frugivoreFD_r2_linear, frugivoreTD.plantFD_r2_linear), 
+               names_to = "Relationship", 
+               values_to = "r_squared")
+```
+
+```{r}
+# Define custom colors for the points
+custom_colors <- c(plantTD.FD_r2_linear = "brown", frugivoreTD.FD_r2_linear = "orange",
+                   plantTD.frugivoreFD_r2_linear = "tan", frugivoreTD.plantFD_r2_linear = "darkorange")
+
+# Define custom legend labels
+custom_labels <- c(plantTD.FD_r2_linear = "Plant TD - Plant FD",
+                   frugivoreTD.FD_r2_linear = "Frugivore TD - Frugivore FD",
+                   plantTD.frugivoreFD_r2_linear = "Plant TD - Frugivore FD",
+                   frugivoreTD.plantFD_r2_linear = "Frugivore TD - Plant FD")
+
+(TD.FD_r2_plot <- ggplot(TD.FD_r2_long,aes(x = resolution, y = r_squared, color = Relationship)) +
+  geom_line() +
+  geom_point(size = 3) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = TD.FD_r2_compare$resolution) + 
+  labs(x = "Grain (km)",
+       y = "R² value",
+       color = "Relationship") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = "inside",
+        legend.position.inside = c(.25, .75),
+        legend.box.margin = margin(10, 10, 10, 10),
+        text = element_text(size = 18)))
+```
+
+```{r}
+TD.FD_slopecoefficent_compare <- tibble(
+  resolution = c(5, 10, 25, 50, 75, 100),
+  plantTD.FD_slopecoefficent_linear = c(
+    as.numeric(compare_plantTD.FD_slopecoefficent_5km_linear), 
+    as.numeric(compare_plantTD.FD_slopecoefficent_10km_linear), 
+    as.numeric(compare_plantTD.FD_slopecoefficent_25km_linear), 
+    as.numeric(compare_plantTD.FD_slopecoefficent_50km_linear), 
+    as.numeric(compare_plantTD.FD_slopecoefficent_75km_linear), 
+    as.numeric(compare_plantTD.FD_slopecoefficent_100km_linear)
+  ),
+  frugivoreTD.FD_slopecoefficent_linear = c(
+    as.numeric(compare_frugivoreTD.FD_slopecoefficent_5km_linear), 
+    as.numeric(compare_frugivoreTD.FD_slopecoefficent_10km_linear), 
+    as.numeric(compare_frugivoreTD.FD_slopecoefficent_25km_linear), 
+    as.numeric(compare_frugivoreTD.FD_slopecoefficent_50km_linear), 
+    as.numeric(compare_frugivoreTD.FD_slopecoefficent_75km_linear), 
+    as.numeric(compare_frugivoreTD.FD_slopecoefficent_100km_linear)
+  ),
+  plantTD.frugivoreFD_slopecoefficent_linear = c(
+    as.numeric(compare_plantTD.frugivoreFD_slopecoefficent_5km_linear), 
+    as.numeric(compare_plantTD.frugivoreFD_slopecoefficent_10km_linear), 
+    as.numeric(compare_plantTD.frugivoreFD_slopecoefficent_25km_linear), 
+    as.numeric(compare_plantTD.frugivoreFD_slopecoefficent_50km_linear), 
+    as.numeric(compare_plantTD.frugivoreFD_slopecoefficent_75km_linear), 
+    as.numeric(compare_plantTD.frugivoreFD_slopecoefficent_100km_linear)
+  ),
+  frugivoreTD.plantFD_slopecoefficent_linear = c(
+    as.numeric(compare_frugivoreTD.plantFD_slopecoefficent_5km_linear), 
+    as.numeric(compare_frugivoreTD.plantFD_slopecoefficent_10km_linear), 
+    as.numeric(compare_frugivoreTD.plantFD_slopecoefficent_25km_linear), 
+    as.numeric(compare_frugivoreTD.plantFD_slopecoefficent_50km_linear), 
+    as.numeric(compare_frugivoreTD.plantFD_slopecoefficent_75km_linear), 
+    as.numeric(compare_frugivoreTD.plantFD_slopecoefficent_100km_linear)
+  )
+)
+
+# reshape to long format for ease of plotting
+TD.FD_slopecoefficent_long <- TD.FD_slopecoefficent_compare %>%
+  pivot_longer(cols = c(plantTD.FD_slopecoefficent_linear, frugivoreTD.FD_slopecoefficent_linear,
+                        plantTD.frugivoreFD_slopecoefficent_linear, frugivoreTD.plantFD_slopecoefficent_linear), 
+               names_to = "Relationship", 
+               values_to = "slope_coefficent")
+```
+
+```{r}
+# Define custom colors for the points
+custom_colors <- c(plantTD.FD_slopecoefficent_linear = "brown", frugivoreTD.FD_slopecoefficent_linear = "orange",
+                   plantTD.frugivoreFD_slopecoefficent_linear = "tan", frugivoreTD.plantFD_slopecoefficent_linear = "darkorange")
+
+# Define custom legend labels
+custom_labels <- c(plantTD.FD_slopecoefficent_linear = "Plant TD - Plant FD",
+                   frugivoreTD.FD_slopecoefficent_linear = "Frugivore TD - Frugivore FD",
+                   plantTD.frugivoreFD_slopecoefficent_linear = "Plant TD - Frugivore FD",
+                   frugivoreTD.plantFD_slopecoefficent_linear = "Frugivore TD - Plant FD")
+
+(TD.FD_slopecoefficent_plot <- ggplot(TD.FD_slopecoefficent_long,aes(x = resolution, y = slope_coefficent, color = Relationship)) +
+  geom_line() +
+  geom_point(size = 3, shape = 15) +
+  scale_color_manual(values = custom_colors, labels = custom_labels) +
+  scale_x_continuous(breaks = TD.FD_slopecoefficent_compare$resolution) + 
+  labs(x = "Grain (km)",
+       y = "Slope Coefficent",
+       color = "Relationship") +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.line = element_line(color = "black"),
+        axis.ticks = element_line(color = "black"),
+        axis.text = element_text(color = "black"),
+        legend.position = "inside",
+        legend.position.inside = c(.75, .75),
+        legend.box.margin = margin(10, 10, 10, 10),
+        text = element_text(size = 18)))
+```
+
+```{r}
+(compare_TD.FD_all_plots <- ggarrange(TD.FD_r2_plot, TD.FD_slopecoefficent_plot, ncol = 2, nrow = 1))
+```
+
+```{r}
+ggsave("compare_TD.FD_all_plots.png", compare_TD.FD_all_plots, path = figure_path,
+       width = 14, height = 8, units = "in")
+```

### 2024-09-05 code/L1/L1_P5_TropicalAndes_plant_imputetraits.Rmd
minor edits to plotting and figure saving code

@@
 library(forcats)
+library(viridis)
 ```
@@
 ```{r}
-# Reassign NA values in TraitLevel as "Imputed"
-all_traits_with_NAs$TraitLevel[is.na(all_traits_with_NAs$TraitLevel)] <- "Imputed"
+# Reassign NA values in TraitLevel as "imputed"
+all_traits_with_NAs$TraitLevel[is.na(all_traits_with_NAs$TraitLevel)] <- "imputed"
 
@@
   count(TraitLevel, TraitName) %>%
-  mutate(TraitLevel = fct_relevel(TraitLevel, "species", "genus", "family", "Imputed"))
+  mutate(TraitLevel = fct_relevel(TraitLevel, "species", "genus", "family", "imputed"))
 
 # Create the faceted bar plot with free y-axis scales
-ggplot(trait_counts, aes(x = TraitLevel, y = n, fill = TraitLevel)) +
+(trait_count_level_plot <- ggplot(trait_counts, aes(x = TraitLevel, y = n, fill = TraitLevel)) +
   geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Trait Counts per Level",
-       x = "Trait Level",
+  scale_fill_viridis(discrete = TRUE) +
+  labs(x = "Trait Level",
        y = "Trait Count",
@@
         legend.position = "none") +  # Remove legend
-  facet_wrap(~ TraitName, scales = "free_y") +
-  guides(fill = "none")  # Remove legend
-ggsave("plant_trait_counts_per_level.png", plot = last_plot(), path = figure_path)
+  facet_wrap(~ TraitName, scales = "free_y", nrow = 2) +
+  guides(fill = "none"))  # Remove legend
+ggsave("plant_trait_counts_per_level.png", plot = trait_count_level_plot, path = figure_path, height = 6, width = 8, units = "in")
 ```
@@
   count(TraitLevel) %>%
-  mutate(TraitLevel = fct_relevel(TraitLevel, "species", "genus", "family", "Imputed"))
+  mutate(TraitLevel = fct_relevel(TraitLevel, "species", "genus", "family", "imputed"))
 
 # Create the faceted bar plot with free y-axis scales
-ggplot(trait_counts_overall, aes(x = TraitLevel, y = n, fill = TraitLevel)) +
+(all_trait_count_plot <- ggplot(trait_counts_overall, aes(x = TraitLevel, y = n, fill = TraitLevel)) +
   geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Trait Counts per Level",
+  scale_fill_viridis(discrete = TRUE) +
+  labs(title = "All traits",
        x = "Trait Level",
@@
   theme_minimal() +
-  guides(fill = "none")  # Remove legend
-ggsave("plant_trait_counts_per_level_overall.png", plot = last_plot(), path = figure_path)
+  theme(plot.title = element_text(hjust = 0.5)) +
+  guides(fill = "none"))  # Remove legend
+ggsave("plant_trait_counts_per_level_overall.png", plot = all_trait_count_plot, path = figure_path)
+```
+
+```{r}
+(both_trait_counts <- ggarrange(trait_count_level_plot, all_trait_count_plot,
+                                        ncol = 2, nrow = 1, widths = c(2,1)))
+ggsave("plant_trait_counts_both.png", plot = both_trait_counts, path = figure_path, height = 6, width = 12, units = "in")
+```
+
+```{r}
+# Calculate the total sum of the values
+total_sum <- sum(trait_counts_overall$n)
+
+# Calculate the percentage of each value
+trait_counts_overall$percentage <- (trait_counts_overall$n / total_sum) * 100
+
+# Print the dataframe with the percentages
+print(trait_counts_overall)
 ```

### 2024-09-05 code/L1/L1_P6_TropicalAndes_plant_occ.Rmd
update summary code

@@
                values_to = "count")
+
+# Reorder and rename the categories
+combined_summary_long <- combined_summary_long %>%
+  mutate(
+    source = factor(source, levels = c("Data", "Checklist")),
+    category = fct_recode(category,
+                          "Species" = "num_species",
+                          "Genera" = "num_genera",
+                          "Families" = "num_families")
+  )
+
+# Reorder the category variable
+combined_summary_long <- combined_summary_long %>%
+  mutate(
+    category = factor(category, levels = c("Species", "Genera", "Families"))
+  )
+
+# Custom colors for source
+custom_colors <- c("Data" = "seagreen", "Checklist" = "darkgrey")
 ```
@@
   geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Comparison of Species, Genera, and Families",
-       x = "Source",
+  labs(x = "Source",
        y = "Count",
        fill = "Source") +
+  scale_fill_manual(values = custom_colors) + 
   theme_minimal() +
   theme(
-    axis.text.x = element_text(angle = 45, hjust = 1),  # Angled x-axis labels
-    plot.title = element_text(hjust = 0.5),             # Center the plot title
-    plot.margin = margin(t = 10, r = 10, b = 10, l = 10)
+    plot.margin = margin(t = 10, r = 10, b = 10, l = 10),
+    legend.position = "none"
   ) +

### 2024-09-05 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
minor plotting changes

@@
       geom_sf(data = Americas, fill = "white") +
-      geom_sf(data = TApoly, fill = "grey", size = 0.1) +
+      geom_sf(data = TApoly, fill = "grey") +
       geom_sf(aes(fill = num_species), color = NA) +

### 2025-08-12 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
Added name, no problems with code

@@
 project: "Plant-Frugivore Diversity"
-collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
+collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske, Jenna B. Baljunas"
 overview: "Create raster of IUCN habitat classification Forest-Subtropical moist montane and Forest - Subtropical moist lowland for Tropical Andes."

### 2025-08-14 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
Added name and few minor changes;
computer did not like running line 78- return to this

@@
 project: "Plant-Frugivore Diversity"
-collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
+collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske, and Jenna B. Baljunas"
 overview: "This script retrives plant occurrence data from GBIF using a subset of species with IUCN habitat designations ‘Forest-Subtropical/Tropical Moist Montane’ and/or ‘Forest-Subtropical/Tropical Moist Lowland’.

### 2025-08-14 code/L0/L0_TropicalAndes_frugivores_GBIF.Rmd
Added name and few minor changes;
computer did not like running line 78- return to this

@@
 title: "Tropical Andes montane and lowland forest bird and mammal species GBIF"
-author: "Hazel J. Anderson"
+author: "Hazel J. Anderson, Jenna B. Baljunas"
 project: "Plant-Frugivore Diversity"
@@
                         path = output_path_L0, overwrite = TRUE) %>%
-                        curveocc_download_import()
+                        occ_download_import()
   }
@@
 ```{r}
-TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path, "Forest_sf.shp"), layer = "Forest_sf")
+TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
 ```

### 2025-08-17 code/L0/L0_TropicalAndes_plants_GBIF.Rmd

--- 2024-05-10 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
@@
 title: "Tropical Andes plant occurrence data GBIF"
-author: "Hazel J. Anderson"
+author: "Hazel J. Anderson, Jenna B. Baljunas"
 project: "Plant-Frugivore Diversity"
@@
 TropicalAndes_plant_occ_forest <- TropicalAndes_plant_occ_forest %>%
-    mutate(decimalLongitude = sf::st_coordinates(.)[,1],
-           decimalLatitude = sf::st_coordinates(.)[,2])
+    mutate(decimalLongitude = st_coordinates(.)[,1],
+           decimalLatitude = st_coordinates(.)[,2])
 ```

### 2025-08-17 code/L0/L0_TropicalAndes_frugivores_frugivoria_traits.Rmd
Added name and a function for Frug_mammal (return for a closer look to see if it does the same as original code); if good, similar function can be added for Frug_bird

@@
 title: "Tropical Andes frugivore trait data Frugivoria"
-author: "Hazel J. Anderson"
+author: "Hazel J. Anderson, Jenna B. Baljunas"
 project: "Plant-Frugivore Diversity"
@@
 
+col_class <- function(x){
+
+ # extract cols and categorize them into what they should be
+ cols <- colnames(x)
+ numeric_cols <- c("diet_breadth", "body_mass_e", "body_size_mm", "longevity", "home_range_size", "generation_time", "habitat_breadth", "mean_CHELSA_bio1_1981.2010_V.2.1", "mean_CHELSA_bio12_1981.2010_V.2.1", "mean_human_fp_range_2010", "percent_change_hf_2010_2020","inferred_range_sqkm")
+ factor_cols <- setdiff(cols, numeric_cols)
+
+ # convert to numeric
+  for (col in numeric_cols) {
+    if (col %in% names(x)) {
+      if (is.factor(x[[col]])) {
+        x[[col]] <- as.numeric(levels(x[[col]]))[as.integer(x[[col]])]
+      } else if (is.character(x[[col]])) {
+        x[[col]] <- as.numeric(x[[col]])
+      }
+    }
+  }
+
+ # convert to factor
+ for (col in factor_cols) {
+    if (col %in% names(x) && !is.factor(x[[col]])) {
+      x[[col]] <- as.factor(x[[col]])
+    }
+  }
+
+ return(x)
+}
+
+d <- col_class(Frug_mammal)
+
+# also, replacing NA strings might not be necessary since there are none:
+d == "NA"
+which(d == "NA", arr.ind = TRUE)
+
+# need to double-check but function could replace what's below (and could adjust accordingly)
+
 if (class(Frug_mammal$IUCN_species_name)!="factor") Frug_mammal$IUCN_species_name<- as.factor(Frug_mammal$IUCN_species_name)
@@
 if (class(Frug_mammal$diet_source_e)!="factor") Frug_mammal$diet_source_e<- as.factor(Frug_mammal$diet_source_e)
+
 if (class(Frug_mammal$diet_breadth)=="factor") Frug_mammal$diet_breadth <-as.numeric(levels(Frug_mammal$diet_breadth))[as.integer(Frug_mammal$diet_breadth) ]       
 if (class(Frug_mammal$diet_breadth)=="character") Frug_mammal$diet_breadth <-as.numeric(Frug_mammal$diet_breadth)
+
 if (class(Frug_mammal$diet_level)!="factor") Frug_mammal$diet_level<- as.factor(Frug_mammal$diet_level)
@@
 if (class(Frug_mammal$for_strat_certainty_e)!="factor") Frug_mammal$for_strat_certainty_e<- as.factor(Frug_mammal$for_strat_certainty_e)
+
 if (class(Frug_mammal$body_mass_e)=="factor") Frug_mammal$body_mass_e <-as.numeric(levels(Frug_mammal$body_mass_e))[as.integer(Frug_mammal$body_mass_e) ]         
 if (class(Frug_mammal$body_mass_e)=="character") Frug_mammal$body_mass_e <-as.numeric(Frug_mammal$body_mass_e)
+
 if (class(Frug_mammal$body_mass_level_e)!="factor") Frug_mammal$body_mass_level_e<- as.factor(Frug_mammal$body_mass_level_e)
 if (class(Frug_mammal$body_mass_source_e)!="factor") Frug_mammal$body_mass_source_e<- as.factor(Frug_mammal$body_mass_source_e)
+
 if (class(Frug_mammal$body_size_mm)=="factor") Frug_mammal$body_size_mm <-as.numeric(levels(Frug_mammal$body_size_mm))[as.integer(Frug_mammal$body_size_mm) ]       
 if (class(Frug_mammal$body_size_mm)=="character") Frug_mammal$body_size_mm <-as.numeric(Frug_mammal$body_size_mm)
+
 if (class(Frug_mammal$body_size_notes)!="factor") Frug_mammal$body_size_notes<- as.factor(Frug_mammal$body_size_notes)
@@
 if (class(Frug_mammal$sexual_dimorphism_source)!="factor") Frug_mammal$sexual_dimorphism_source<- as.factor(Frug_mammal$sexual_dimorphism_source)
+
 if (class(Frug_mammal$longevity)=="factor") Frug_mammal$longevity <-as.numeric(levels(Frug_mammal$longevity))[as.integer(Frug_mammal$longevity) ]             
 if (class(Frug_mammal$longevity)=="character") Frug_mammal$longevity <-as.numeric(Frug_mammal$longevity)
+
 if (class(Frug_mammal$longevity_notes)!="factor") Frug_mammal$longevity_notes<- as.factor(Frug_mammal$longevity_notes)
@@
 if (class(Frug_mammal$longevity_source)!="factor") Frug_mammal$longevity_source<- as.factor(Frug_mammal$longevity_source)
+
 if (class(Frug_mammal$home_range_size)=="factor") Frug_mammal$home_range_size <-as.numeric(levels(Frug_mammal$home_range_size))[as.integer(Frug_mammal$home_range_size) ]
 if (class(Frug_mammal$home_range_size)=="character") Frug_mammal$home_range_size <-as.numeric(Frug_mammal$home_range_size)
+
 if (class(Frug_mammal$home_range_notes)!="factor") Frug_mammal$home_range_notes<- as.factor(Frug_mammal$home_range_notes)
@@
 if (class(Frug_mammal$habitat_specialization_source)!="factor") Frug_mammal$habitat_specialization_source<- as.factor(Frug_mammal$habitat_specialization_source)
+
 if (class(Frug_mammal$generation_time)=="factor") Frug_mammal$generation_time <-as.numeric(levels(Frug_mammal$generation_time))[as.integer(Frug_mammal$generation_time) ] 
 if (class(Frug_mammal$generation_time)=="character") Frug_mammal$generation_time <-as.numeric(Frug_mammal$generation_time)
+
 if (class(Frug_mammal$generation_time_notes)!="factor") Frug_mammal$generation_time_notes<- as.factor(Frug_mammal$generation_time_notes)
@@
 if (class(Frug_mammal$habitat)!="factor") Frug_mammal$habitat<- as.factor(Frug_mammal$habitat)
+
 if (class(Frug_mammal$habitat_breadth)=="factor") Frug_mammal$habitat_breadth <-as.numeric(levels(Frug_mammal$habitat_breadth))[as.integer(Frug_mammal$habitat_breadth) ] 
 if (class(Frug_mammal$habitat_breadth)=="character") Frug_mammal$habitat_breadth <-as.numeric(Frug_mammal$habitat_breadth)
+
 if (class(Frug_mammal$habitat_breadth_source)!="factor") Frug_mammal$habitat_breadth_source<- as.factor(Frug_mammal$habitat_breadth_source)
 if (class(Frug_mammal$habitat_level)!="factor") Frug_mammal$habitat_level<- as.factor(Frug_mammal$habitat_level)
+
 if (class(Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1)=="factor") Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1 <-as.numeric(levels(Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1))[as.integer(Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1) ]
 if (class(Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1)=="character") Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1 <-as.numeric(Frug_mammal$mean_CHELSA_bio1_1981.2010_V.2.1)
+
 if (class(Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1)=="factor") Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1 <-as.numeric(levels(Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1))[as.integer(Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1) ]
 if (class(Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1)=="character") Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1 <-as.numeric(Frug_mammal$mean_CHELSA_bio12_1981.2010_V.2.1)
+
 if (class(Frug_mammal$mean_human_fp_range_2010)=="factor") Frug_mammal$mean_human_fp_range_2010 <-as.numeric(levels(Frug_mammal$mean_human_fp_range_2010))[as.integer(Frug_mammal$mean_human_fp_range_2010) ]
 if (class(Frug_mammal$mean_human_fp_range_2010)=="character") Frug_mammal$mean_human_fp_range_2010 <-as.numeric(Frug_mammal$mean_human_fp_range_2010)
+
 if (class(Frug_mammal$mean_human_fp_range_2020)=="factor") Frug_mammal$mean_human_fp_range_2020 <-as.numeric(levels(Frug_mammal$mean_human_fp_range_2020))[as.integer(Frug_mammal$mean_human_fp_range_2020) ]
 if (class(Frug_mammal$mean_human_fp_range_2020)=="character") Frug_mammal$mean_human_fp_range_2020 <-as.numeric(Frug_mammal$mean_human_fp_range_2020)
+
 if (class(Frug_mammal$percent_change_hf_2010_2020)=="factor") Frug_mammal$percent_change_hf_2010_2020 <-as.numeric(levels(Frug_mammal$percent_change_hf_2010_2020))[as.integer(Frug_mammal$percent_change_hf_2010_2020) ]
 if (class(Frug_mammal$percent_change_hf_2010_2020)=="character") Frug_mammal$percent_change_hf_2010_2020 <-as.numeric(Frug_mammal$percent_change_hf_2010_2020)
+
 if (class(Frug_mammal$inferred_range_sqkm)=="factor") Frug_mammal$inferred_range_sqkm <-as.numeric(levels(Frug_mammal$inferred_range_sqkm))[as.integer(Frug_mammal$inferred_range_sqkm) ]       
 if (class(Frug_mammal$inferred_range_sqkm)=="character") Frug_mammal$inferred_range_sqkm <-as.numeric(Frug_mammal$inferred_range_sqkm)
+
 if (class(Frug_mammal$date_data_obtained)!="factor") Frug_mammal$date_data_obtained<- as.factor(Frug_mammal$date_data_obtained)

### 2025-08-18 code/L0/L0_TropicalAndes_plants_BIEN_traits.Rmd
Added name; cannot run line 45 (error regarding vector size), but the tropical Andes subset of plant trait data (line 50) works

@@
 project: "Plant-Frugivore Diversity"
-collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
+collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske, and Jenna B. Baljunas"
 overview: "This script retrieves plant trait data from the BIEN database for Tropical Andean countries."

### 2025-08-18 code/L0/L0_TropicalAndes_plants_TRY.Rmd
Added name, couldn't import data (vector size too big)

@@
 project: "Plant-Frugivore Diversity"
-collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
+collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske, and Jenna B. Baljunas"
 overview: "This script retrieves plant trait data from the TRY database for plant species list."

### 2025-08-19 code/L1/L1_F1_TropicalAndes_frugivores_GBIF.Rmd
Added name; TropicalAndes_GBIF_frugivore_occ.csv file does not exist- was unable to produce it on my computer in L0_TropicalAndes_frugivores_GBIF, but it should be in shared drive anyway and it's not

@@
 project: "Plant-Frugivore Diversity"
-collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
+collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske, and Jenna B. Baljunas"
 overview: "Uses CoordinateCleaner to flag and remove problematic records from GBIF frugivore occurrence records."

### 2025-08-21 code/L0/L0_TropicalAndes_frugivores_frugivoria_traits.Rmd
Changed csv names (previously overrode originals but will see if I can recover them)

@@
 ```{r}
-source("~/GitHub/neotropical_plants/code/data_summary.R")
+source("C:/GitHub_projects/plant-frugivore diversity/neotropical_plants/code/data_summary.R")
 ```
@@
 ```{r}
-write.csv(Frugivoria, file = file.path(output_path_L0,"TropicalAndes_Frugivoria_frugivore_traits.csv"))
-write.csv(Frugivoria_SpeciesList, file = file.path(output_path_L0,"TropicalAndes_Frugivoria_frugivore_traits_species.csv"))
-write.csv(Frugivore_LookupTable, file = file.path(output_path_L1,"TropicalAndes_frugivore_LookupTable.csv"))
+write.csv(Frugivoria, file = file.path(output_path_L0,"TropicalAndes_Frugivoria_frugivore_traits_21Aug2025.csv"))
+write.csv(Frugivoria_SpeciesList, file = file.path(output_path_L0,"TropicalAndes_Frugivoria_frugivore_traits_species_21Aug2025.csv"))
+write.csv(Frugivore_LookupTable, file = file.path(output_path_L1,"TropicalAndes_frugivore_LookupTable_21Aug2025.csv"))
 ```

### 2025-08-27 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
Added name, no problems with code
Small rearrangement

@@
 project: "Plant-Frugivore Diversity"
-collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske, Jenna B. Baljunas"
+collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske, and Jenna B. Baljunas"
 overview: "Create raster of IUCN habitat classification Forest-Subtropical moist montane and Forest - Subtropical moist lowland for Tropical Andes."
@@
 iucn_moist_montane_forest <- raster(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004/lvl2_frac_1km_ver004/iucn_habitatclassification_fraction_lvl2__109_Forest – Subtropical-tropical moist montane__ver004.tif"))
-iucn_moist_lowland_forest <- raster(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004/lvl2_frac_1km_ver004/iucn_habitatclassification_fraction_lvl2__106_Forest – Subtropical-tropical moist lowland__ver004.tif"))
+plot(iucn_moist_montane_forest)
 ```
@@
 ```{r}
-plot(iucn_moist_montane_forest)
+iucn_moist_lowland_forest <- raster(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004/lvl2_frac_1km_ver004/iucn_habitatclassification_fraction_lvl2__106_Forest – Subtropical-tropical moist lowland__ver004.tif"))
 plot(iucn_moist_lowland_forest)

### 2025-08-27 code/L0/L0_TropicalAndes_plants_BIEN_traits.Rmd
Small error in output_path name; could not download AllDesired_BIEN_plant_traits, only TropicalAndes

@@
 ```{r}
-source("~/GitHub/neotropical_plants/code/data_summary.R")
+source("C:/GitHub_projects/plant-frugivore diversity/neotropical_plants/code/data_summary.R")
 
@@
 ```{r}
-write.csv(AllDesired_BIEN_plant_traits, file.path(output_path,"AllDesired_BIEN_plant_traits.csv"))
-write.csv(TropicalAndes_BIEN_plant_traits, file.path(output_path,"TropicalAndes_BIEN_plant_traits.csv"))
+write.csv(AllDesired_BIEN_plant_traits, file.path(output_path_L0,"AllDesired_BIEN_plant_traits.csv"))
+write.csv(TropicalAndes_BIEN_plant_traits, file.path(output_path_L0,"TropicalAndes_BIEN_plant_traits.csv"))
 ```

### 2025-08-27 code/L0/L0_TropicalAndes_plants_GIFT.Rmd
Added name; code works fine but could probably add a function later

@@
 title: "Tropical Andes plant trait data GIFT"
-author: "Hazel J. Anderson"
+author: "Hazel J. Anderson, Jenna B. Baljunas"
 project: "Plant-Frugivore Diversity"
@@
 ```{r}
-# 
+# maybe make into function later?
 height_mean <- GIFT_traits(trait_IDs = c("1.6.3"))

### 2025-08-27 code/L0/L0_TropicalAndes_plants_TRY.Rmd
Still cannot import TRY data

@@
 ```{r}
-try_data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/29548_24102023041242/29548.txt")
+try_data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY_2023/29548_24102023041242/29548.txt")
 ```

### 2025-08-27 code/L0/L0_TropicalAndes_plants_GBIF.Rmd
Source function fix

@@
 output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
-source("~/GitHub/neotropical_plants/functions/data_summary.R")
+source("C:/GitHub_projects/plant-frugivore diversity/neotropical_plants/code/data_summary.R")
 ```
